### 2011-12-24 paper/paper.tex
Add start of paper

\documentclass{sig-alternate}

\begin{document}

\conferenceinfo{HPDC}{'12 Delft, The Netherlands}
%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.

\title{Something About Workflow Scheduling in Clouds}
%\subtitle{[Extended Abstract]}

\numberofauthors{2}
\author{
    \alignauthor Maciej Malawski and Jarek Nabrzyski\\
       \affaddr{University of Notre Dame}\\
       \affaddr{South Bend, IN, USA}\\
       \email{\{mmalawsk,naber\}@nd.edu}
    \alignauthor Gideon Juve and Ewa Deelman\\
       \affaddr{USC Information Sciences Institute}\\
       \affaddr{Marina del Rey, CA, USA}\\
       \email{\{gideon,deelman\}@isi.edu}
}

\maketitle

\begin{abstract}
Abstract goes here.
\end{abstract}

%\category{H.4}{Information Systems Applications}{Miscellaneous}
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Theory}

\keywords{Scientific workflows, DAG scheduling, simulation}

\section{Introduction}
Here is a citation~\cite{Yu2005} just to make bibtex work.

\section{Related Work}

\section{Problem Description}

\section{Algorithms}

\section{Performance Evaluation}

\section{Conclusions}
Conclusions go here.

%\section{Acknowledgments}
%Acknowledgements go here.

\bibliographystyle{abbrv}
\bibliography{paper}

%\appendix
%\section{Headings in Appendices}

%\balancecolumns

\end{document}

### 2011-12-26 paper/paper.tex
Add some notes for related work

@@
        \affaddr{University of Notre Dame}\\
-       \affaddr{South Bend, IN, USA}\\
+       \affaddr{111 Information Technology Center}\\
+       \affaddr{Notre Dame, IN, USA}\\
        \email{\{mmalawsk,naber\}@nd.edu}
@@
        \affaddr{USC Information Sciences Institute}\\
+       \affaddr{4676 Admiralty Way}\\
        \affaddr{Marina del Rey, CA, USA}\\
@@
 \section{Introduction}
-Here is a citation~\cite{Yu2005} just to make bibtex work.
 
 \section{Related Work}
+General policy and rule-based approaches to dynamic provisioning (e.g. Amazon Auto Scaling\footnote{http://aws.amazon.com/autoscaling} and RightScale\footnote{http://www.rightscale.com}).
+
+Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010})
+
+Workflow scheduling in utility grids (e.g. \cite{Yu2005})
+
+Workflow ensemble scheduling in clouds (e.g. \cite{Mao2011})
 
 \section{Problem Description}
+\subsection{Application Model}
+\subsection{System Model}
 
 \section{Algorithms}
+\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
+\subsection{Workflow-Aware DPDS (WA-DPDS)}
+\subsection{Static Provisioning Static Scheduling (SPSS)}
 
@@
 
-\section{Conclusions}
-Conclusions go here.
+\section{Conclusion and Future Work}
 

### 2012-01-05 paper/paper.tex
Add some related work, algorithm example

@@
 \documentclass{sig-alternate}
+\usepackage{algorithm,algpseudocode}
 
@@
 
-\title{Something About Workflow Scheduling in Clouds}
+\title{Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds}
 %\subtitle{[Extended Abstract]}
@@
 
-Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010})
+Policy-based approaches for scientific workloads (e.g.
+\cite{Marshall2010, Kim2011}).
+Our approach is different in that we consider workflows, while policy based approaches typically consider bags of independent tasks or unpredictable batch workloads.
+This enables us to take advantage of scheduling heuristics that cannot be applied to independent tasks.
 
-Workflow scheduling in utility grids (e.g. \cite{Yu2005})
+Deadline-constrained cost-minimization workflow scheduling.
+Our work is different from \cite{Yu2005, Abrishami2010} in that we consider ensembles of workflows in IaaS clouds, which allow one to provision resources on a per-hour billing model, rather than utility grids, which allow one to choose from a pool of existing resources with a per-job billing model.
+Our work is different from \cite{Mao2011} in that we consider ensembles of workflows rather than unpredictable workloads containing workflows.
+We also have budget constraints rather than cost minimization as a goal.
+In other words, we assume that there is more work to be done than the available budget, so some work must be rejected.
+Therefore cost is not something we optimize, but rather a constraint.
 
-Workflow ensemble scheduling in clouds (e.g. \cite{Mao2011})
+Budget-constrained workflow scheduling \cite{Sakellariou2007}.
+
+Bi-criteria scheduling and multi-criteria scheduling.
+These approaches are similar to ours in that we have two scheduling criteria: cost and makespan.
+The challenge in multi-criteria scheduling is to derive an objective function that takes into account all of the criteria.
+These approaches typically use metaheuristics that run for a long time before producing good results.
 
@@
 \subsection{Static Provisioning Static Scheduling (SPSS)}
+
+% This is an example of how we can do algorithms
+\begin{algorithm}
+\caption{Euclid's algorithm}
+\label{alg:euclid}
+\begin{algorithmic}[1]
+\Procedure{Euclid}{$a,b$}       \Comment{The g.c.d. of a and b}
+    \State $r\gets a\bmod b$
+    \While{$r\not=0$}           \Comment{We have the answer if r is 0}
+        \State $a\gets b$
+        \State $b\gets r$
+        \State $r\gets a\bmod b$
+    \EndWhile
+    \State \textbf{return} $b$  \Comment{The gcd is b}
+\EndProcedure
+\end{algorithmic} 
+\end{algorithm}
 

### 2012-01-10 paper/paper.tex
Add outline of algorithm
added DPDS pseudocode

@@
        \affaddr{University of Notre Dame}\\
+       \affaddr{Center for Research Computing}\\
        \affaddr{111 Information Technology Center}\\
@@
 \section{Introduction}
+
+Workflows are important but workflow ensembles are more important, see
+CyberShake or other examples.
+
+Clouds are interesting and different from grids.
+
+Define problems: scheduling and provisioning.
 
@@
 \subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
+
+% This is an example of how we can do algorithms
+\begin{algorithm}
+\caption{Priority-based scheduling}
+\label{alg:ds}
+\begin{algorithmic}[1]
+\Procedure{schedule}{}
+    \State $Q\gets set\ of\ released\ tasks$
+    \State $P\gets empty\ priority\ queue$
+    \While{$Q \neq \emptyset$}
+    	\For{$task\ t\in Q$}
+    		\State \Call{Insert}{$t,P$} 
+    	\EndFor
+    	\While{$FreeVMs \neq \emptyset \wedge P \neq \emptyset $}
+    		\State $VM\gets$ \Call{SelectRandom}{$FreeVMs$}
+    		\State $t\gets$ \Call{Pop}{$P$}
+    		\State \Call{Submit}{$t,VM$}
+    	\EndWhile
+    	\State \Call{Wait}{$task\ finished$}
+    	\State \Call{Update}{$Q$}
+    \EndWhile
+\EndProcedure
+\end{algorithmic} 
+\end{algorithm}
+
 \subsection{Workflow-Aware DPDS (WA-DPDS)}
@@
 \begin{algorithm}
-\caption{Euclid's algorithm}
-\label{alg:euclid}
+\caption{Admit DAGs}
+\label{alg:admit}
 \begin{algorithmic}[1]
-\Procedure{Euclid}{$a,b$}       \Comment{The g.c.d. of a and b}
-    \State $r\gets a\bmod b$
-    \While{$r\not=0$}           \Comment{We have the answer if r is 0}
-        \State $a\gets b$
-        \State $b\gets r$
-        \State $r\gets a\bmod b$
+\Procedure{AdmitDAGs}{$W,b,d$}
+    \State $P\gets empty\ plan$
+    \State $A\gets \emptyset$ \Comment{Set of admitted DAGs}
+    \For{$w\ in\ W$}
+        \State $P^\prime \gets$\ \Call{PlanDAG}{$w,P,d$}
+        \If{$P^\prime\ is\ a\ valid\ plan$}
+            \If{$Cost(P^\prime) \le b$}
+                \State $P\gets\ P^\prime$ \Comment{Accept new plan}
+                \State $A \gets A\ +\ w$ \Comment{Admit w}
+            \EndIf
+        \EndIf
+    \EndFor
+    \State \textbf{return} $P,A$
+\EndProcedure
+\end{algorithmic} 
+\end{algorithm}
+
+\begin{algorithm}
+\caption{PlanDAG}
+\label{alg:plandag}
+\begin{algorithmic}[1]
+\Procedure{PlanDAG}{$w,P,d$}
+    \State $P^\prime\gets \emptyset$
+    \State assign each task in w to its cheapest resource type
+    \State $cp \gets\ CriticalPath(w)$
+    \While{$cp > d$}
+        \State upgrade resource type of task with greatest improvement in runtime for least cost
+        \State $cp \gets\ CriticalPath(w)$
     \EndWhile
-    \State \textbf{return} $b$  \Comment{The gcd is b}
+    \State \Call{DeadlineDistribution}{w,d}
+    \For{$t\ in\ w\ sorted\ by\ Deadline(t)$}
+        \State choose resource r in $P^\prime$ that minimizes Cost(t)
+        \If{$AFT(t,r) < Deadline(t)$}
+            \State Schedule(t,r)
+        \Else
+            \State provision a new resource r
+            \State Schedule(t,r)
+        \EndIf
+    \EndFor
+    \State \textbf{return} $P^\prime$
+\EndProcedure
+\end{algorithmic} 
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Deadline Distribution}
+\label{alg:deadlinedistribution}
+\begin{algorithmic}[1]
+\Procedure{DeadlineDistribution}{$w,d$}
+    \State assign $Level(t)$ to each task t in w
+    \For{$level\ l\ in\ w$}
+        \State $Tasks(l) \gets$ number of tasks in l
+        \State $Runtime(l) \gets$ runtime of tasks in l
+    \EndFor
+    \State $T \gets$ total number of tasks in w
+    \State $R \gets$ total runtime of tasks in w
+    \State $CP \gets$ critical path of w
+    \For{$level\ l\ in\ w$}
+        \State $r \gets \alpha * (Runtime(l)/R)$
+        \State $t \gets (1-\alpha) * (Tasks(l)/T)$
+        \State $FT(l) \gets (r + t) * (d - CP)$
+    \EndFor
+    \For{$task\ t\ in\ w$}
+        \State $est \gets 0$
+        \For{p in Pred(t)}
+            \State $est \gets Deadline(p)$
+        \EndFor
+        \State $Deadline(t) \gets est + RT(t) + FT(Level(t))$
+    \EndFor
 \EndProcedure
@@
 \section{Performance Evaluation}
+
+\subsection{Simulator}
+
+\subsection{Scenarios}
+
+\subsection{Results}
 

### 2012-01-10 src/cws/core/Cloud.java
Add SPSS static algorithm

@@ -104,7 +104,7 @@ public class Cloud extends SimEntity implements WorkflowEvent {
     }
     
     private void vmTerminated(VM vm) {
-     // Sanity check
+        // Sanity check
         if (!vms.contains(vm)) {
             throw new RuntimeException("Unknown VM");
         }

### 2012-01-10 src/cws/core/EnsembleManager.java

@@ -61,7 +61,6 @@ public class EnsembleManager extends SimEntity implements WorkflowEvent {
     
     @Override
     public void startEntity() {
-        System.out.println("Starting em");
         // Submit all DAGs
         while(!dags.isEmpty()) {
             submitDAG(dags.pop());

### 2012-01-10 src/cws/core/SPSS.java

@@ -1,0 +1,826 @@
+package cws.core;
+
+import java.io.File;
+import java.util.ArrayList;
+import java.util.Collections;
+import java.util.Comparator;
+import java.util.HashSet;
+import java.util.LinkedHashSet;
+import java.util.LinkedList;
+import java.util.List;
+import java.util.HashMap;
+import java.util.Queue;
+import java.util.SortedSet;
+import java.util.TreeMap;
+
+import org.cloudbus.cloudsim.core.CloudSim;
+
+import cws.core.Job.Result;
+import cws.core.dag.DAG;
+import cws.core.dag.Task;
+import cws.core.dag.DAGParser;
+import cws.core.dag.algorithms.CriticalPath;
+import cws.core.dag.algorithms.TopologicalOrder;
+
+/**
+ * @author Gideon Juve <juve@usc.edu>
+ */
+public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, JobListener, DAGJobListener {
+    
+    private WorkflowEngine engine;
+    private EnsembleManager manager;
+    private Cloud cloud;
+    private double budget;
+    private double ensembleDeadline;
+    
+    private Plan plan = new Plan();
+    
+    /** List of all dags in ensemble */
+    private List<DAG> allDAGs;
+    
+    /** List of DAGs that were admitted to run */
+    private List<DAG> admittedDAGs = new LinkedList<DAG>();
+    
+    /** Set of jobs that are ready to run arranged by Task */
+    private HashMap<Task, Job> readyJobs = new HashMap<Task, Job>();
+    
+    /** Mapping of Task to the VM that will run the task */
+    private HashMap<Task, VM> taskMap = new HashMap<Task, VM>();
+    
+    /** Schedule of tasks for each VM */
+    private HashMap<VM, Queue<Task>> vmQueues = new HashMap<VM, Queue<Task>>();
+    
+    /** Set of busy VMs */
+    private HashSet<VM> busy = new HashSet<VM>();
+    
+    public SPSS(double budget, double deadline, List<DAG> dags) {
+        this.budget = budget;
+        this.ensembleDeadline = deadline;
+        this.allDAGs = dags;
+    }
+    
+    public void setCloud(Cloud c) {
+        if (cloud != null)
+            cloud.removeVMListener(this);
+        cloud = c;
+        cloud.addVMListener(this);
+    }
+    
+    public void setWorkflowEngine(WorkflowEngine e) {
+        if (engine != null)
[…]

### 2012-01-10 src/cws/core/VM.java

@@ -279,23 +279,21 @@ public class VM extends SimEntity implements WorkflowEvent {
     private void terminateVM() {
         // Can no longer accept jobs 
         running = false;
-
+        
         // cancel future events
         Predicate p = new PredicateType(JOB_FINISHED);
-		CloudSim.cancelAll(getId(), p);
-		        
-		// Move running jobs back to the queue...
-		
-		jobs.addAll(runningJobs);
-		runningJobs.clear();
+        CloudSim.cancelAll(getId(), p);
+        
+        // Move running jobs back to the queue...
+        jobs.addAll(runningJobs);
+        runningJobs.clear();
         
         // ... and fail all queued jobs
         for (Job job : jobs) {
             job.setResult(Job.Result.FAILURE);
             send(job.getOwner(), 0.0, JOB_FINISHED, job);
-    		Log.printLine(CloudSim.clock() + " Terminating job " + job.getID() + " on VM " + job.getVM().getId());
+            Log.printLine(CloudSim.clock() + " Terminating job " + job.getID() + " on VM " + job.getVM().getId());
         }
-        
         
         // Reset dynamic state
         jobs.clear();
@@ -342,12 +340,12 @@ public class VM extends SimEntity implements WorkflowEvent {
     }
     
     private void jobFinish(Job job) {
-    	
+        
         // Sanity check
         if (!running) {
             throw new RuntimeException("Cannot finish job: VM not running");
         }
-    	
+        
         // remove from the running set
         runningJobs.remove(job);
         

### 2012-01-10 src/cws/core/WorkflowEngine.java

@@ -101,8 +101,7 @@ public class WorkflowEngine extends SimEntity implements WorkflowEvent {
     public double getBudget() {
         return budget;
     }
-
-
+    
     public void setBudget(double budget) {
         this.budget = budget;
     }
@@ -123,14 +122,22 @@ public class WorkflowEngine extends SimEntity implements WorkflowEvent {
         return busyVMs;
     }
     
-    public void addJobListener(JobListener jobListener) {
-        jobListeners.add(jobListener);
+    public LinkedList<DAGJob> getAllDags() {
+        return allDAGJobs;
+    }
+    
+    public void addJobListener(JobListener l) {
+        jobListeners.add(l);
+    }
+    
+    public void removeJobListener(JobListener l) {
+        jobListeners.remove(l);
     }
     
     @Override
     public void startEntity() {
         // send the first provisioning request
-        send(this.getId(), 10.0, PROVISIONING_REQUEST);
+        send(this.getId(), 0.0, PROVISIONING_REQUEST);
     }
     
     @Override
@@ -212,7 +219,16 @@ public class WorkflowEngine extends SimEntity implements WorkflowEvent {
             j.setDAGJob(dj);
             j.setTask(t);
             j.setOwner(getId());
-            queue.add(j);
+            jobReleased(j);
+        }
+    }
+    
+    private void jobReleased(Job j) {
+        queue.add(j);
+        
+        // Notify listeners that job was released
+        for (JobListener jl : jobListeners) {
+            jl.jobReleased(j);
         }
     }
     
@@ -225,6 +241,7 @@ public class WorkflowEngine extends SimEntity implements WorkflowEvent {
     
     private void jobFinished(Job j) {
         // Notify the listeners
+        // IT IS IMPORTANT THAT THIS HAPPENS FIRST
         for (JobListener jl : jobListeners) {
             jl.jobFinished(j);
         }
@@ -258,13 +275,9 @@ public class WorkflowEngine extends SimEntity implements WorkflowEvent {
             // Retry the job
             Log.printLine(CloudSim.clock() + " Job " + j.getID() + " failed on VM " + j.getVM().getId() + " resubmitting...");
             Job retry = jobFactory.createJob(dj, t, getId());
-            queue.add(retry);
+            jobReleased(retry);
         }
         
         scheduler.scheduleJobs(this);
     }
-
-	public LinkedList<DAGJob> getAllDags() {
-		return allDAGJobs;
-	}
 }

### 2012-01-10 src/cws/core/dag/Task.java

@@ -22,6 +22,14 @@ public class Task {
         this.size = size;
     }
     
+    /* IT IS IMPORTANT THAT THESE ARE NOT IMPLEMENTED
+    
+    Using the default implementation allows us to put Tasks from
+    different DAGs that have the same task ID into a single HashMap
+    or HashSet. That way we don't have to maintain a reference from
+    the Task to the DAG that owns it--we can mix tasks from different
+    DAGs in the same data structure.
+    
     @Override
     public int hashCode() {
         return id.hashCode();
@@ -35,6 +43,7 @@ public class Task {
         Task t = (Task)o;
         return this.id.equals(t.id);
     }
+    */
     
     @Override
     public String toString() {

### 2012-01-10 src/cws/core/dag/algorithms/CriticalPath.java

@@ -3,7 +3,6 @@ package cws.core.dag.algorithms;
 import java.util.HashMap;
 import java.util.Map;
 
-import cws.core.dag.DAG;
 import cws.core.dag.Task;
 
 
@@ -16,18 +15,35 @@ public class CriticalPath {
     private Map<Task, Double> eft;
     private Double length = null;
     
-    public CriticalPath(DAG dag, TopologicalOrder order) {
-        eft = new HashMap<Task, Double>();
-
+    public CriticalPath(TopologicalOrder order) {
+        this(order, null);
+    }
+    
+    public CriticalPath(TopologicalOrder order, Map<Task, Double> runtimes) {
+        this.eft = new HashMap<Task, Double>();
+        
+        /* XXX By default use the task size as its runtime. This is not strictly
+         * correct because the size is in MI and the runtime depends on the VM
+         * type that the task runs on.
+         */
+        if (runtimes == null) {
+            runtimes = new HashMap<Task, Double>();
+            for (Task task : order) {
+                runtimes.put(task, task.size);
+            }
+        }
+        
         // Initially the finish time is whatever the runtime is
         for (Task task : order) {
-            eft.put(task, task.size);
+            eft.put(task, runtimes.get(task));
         }
         
         // Now we adjust the values in the topological order
         for (Task task : order) {
             for (Task child : task.children) {
-                eft.put(child, Math.max(eft.get(child), eft.get(task)+child.size));
+                eft.put(child, 
+                        Math.max(eft.get(child), 
+                                 eft.get(task)+runtimes.get(child)));
             }
         }
     }
@@ -37,13 +53,6 @@ public class CriticalPath {
      */
     public double eft(Task task) {
         return eft.get(task);
-    }
-    
-    /** 
-     * @return Earliest start time of task
-     */
-    public double est(Task task) {
-        return eft.get(task) - task.size;
     }
     
     /**

### 2012-01-10 test/cws/core/dag/algorithms/CriticalPathTest.java

@@ -16,7 +16,7 @@ public class CriticalPathTest {
     public void cptest() {
         DAG dag = DAGParser.parseDAG(new File("dags/cptest.dag"));
         TopologicalOrder order = new TopologicalOrder(dag);
-        CriticalPath cp = new CriticalPath(dag, order);
+        CriticalPath cp = new CriticalPath(order);
         
         Task A = dag.getTask("A");
         Task B = dag.getTask("B");
@@ -24,19 +24,14 @@ public class CriticalPathTest {
         Task D = dag.getTask("D");
         Task E = dag.getTask("E");
         
-        assertEquals(0.0, cp.est(A), 0.00001);
         assertEquals(1.0, cp.eft(A), 0.00001);
         
-        assertEquals(1.0, cp.est(B), 0.00001);
         assertEquals(2.0, cp.eft(B), 0.00001);
         
-        assertEquals(2.0, cp.est(C), 0.00001);
         assertEquals(3.0, cp.eft(C), 0.00001);
         
-        assertEquals(3.0, cp.est(D), 0.00001);
         assertEquals(4.0, cp.eft(D), 0.00001);
         
-        assertEquals(4.0, cp.est(E), 0.00001);
         assertEquals(5.0, cp.eft(E), 0.00001);
         
         assertEquals(5.0, cp.getCriticalPathLength(), 0.00001);
@@ -46,7 +41,7 @@ public class CriticalPathTest {
     public void test() {
         DAG dag = DAGParser.parseDAG(new File("dags/test.dag"));
         TopologicalOrder order = new TopologicalOrder(dag);
-        CriticalPath cp = new CriticalPath(dag, order);
+        CriticalPath cp = new CriticalPath(order);
         assertEquals(21, cp.getCriticalPathLength(), 0.00001);
     }
     
@@ -54,7 +49,7 @@ public class CriticalPathTest {
     public void cybershake30() {
         DAG dag = DAGParser.parseDAG(new File("dags/CyberShake_30.dag"));
         TopologicalOrder order = new TopologicalOrder(dag);
-        CriticalPath cp = new CriticalPath(dag, order);
+        CriticalPath cp = new CriticalPath(order);
         assertEquals(221.84, cp.getCriticalPathLength(), 0.00001);
     }
 }

### 2012-01-10 paper/paper.tex
Add some assumptions/problem definition

@@
 \section{Problem Description}
+\subsection{Resource Model}
+Describe the Amazon EC2 resource model.
+Instances are requested on-demand.
+Instance pricing is per-hour.
+Multiple VM types are available, but for this paper we focus on a single VM type because we assume that there will typically be one VM type with the best price/performance ratio for the application.
+
 \subsection{Application Model}
-\subsection{System Model}
+The target applications for this paper are scientific workflows that can be modeled as Directed Acyclic Graphs (DAGs).
+
+We assume that we have runtime estimates for each task in the workflow.
+
+Currently we do not consider data sizes, but rather assume that data is stored in a shared cloud storage system such as Amazon S3 and that data transfer times are included in task runtime estimates.
+Further, data transfer times are equal across the VMs.
+
+An ensemble consists of several related workflows. Each workflow is given a priority.
+
+The goal is to complete as many workflows as possible given a budget and a deadline.
 
@@
 \subsection{Workflow-Aware DPDS (WA-DPDS)}
+
 \subsection{Static Provisioning Static Scheduling (SPSS)}
 
-% This is an example of how we can do algorithms
+We assume that the workflows are very wide such that the critical path of any given workflow is much less than the deadline.
+
+The approach is to schedule each workflow in the ensemble in priority order, rejecting any workflows that cannot be completed by the deadline or that cause the cost of the plan to exceed the budget.
+Once the plan is complete, then the workflows are executed according to the plan.
+
 \begin{algorithm}
-\caption{Admit DAGs}
+\caption{Plan Ensemble}
 \label{alg:admit}
 \begin{algorithmic}[1]
-\Procedure{AdmitDAGs}{$W,b,d$}
+\Procedure{PlanEnsemble}{$W,b,d$}
     \State $P\gets empty\ plan$

### 2012-01-10 paper/paper.tex
Paper: Update SPSS algorithm

@@
 \begin{algorithm}
-\caption{Plan Ensemble}
+\caption{Ensemble planning algorithm}
 \label{alg:admit}
 \begin{algorithmic}[1]
+\Require W: workflow ensemble, b: budget, d: deadline
+\Ensure Schedule as much of W as possible given b and d
 \Procedure{PlanEnsemble}{$W,b,d$}
@@
 \begin{algorithm}
-\caption{PlanDAG}
+\caption{DAG planning algorithm}
 \label{alg:plandag}
 \begin{algorithmic}[1]
+\Require $w$: DAG, $P$: current plan, $d$: deadline
+\Ensure Create plan for $w$ that minimizes cost and meets deadline $d$
 \Procedure{PlanDAG}{$w,P,d$}
-    \State $P^\prime\gets \emptyset$
-    \State assign each task in w to its cheapest resource type
-    \State $cp \gets\ CriticalPath(w)$
-    \While{$cp > d$}
-        \State upgrade resource type of task with greatest improvement in runtime for least cost
-        \State $cp \gets\ CriticalPath(w)$
-    \EndWhile
+    \State $P^\prime\gets$ copy of $P$
     \State \Call{DeadlineDistribution}{w,d}
     \For{$t\ in\ w\ sorted\ by\ Deadline(t)$}
-        \State choose resource r in $P^\prime$ that minimizes Cost(t)
+        \State Choose resource r in $P^\prime$ that minimizes Cost(t)
         \If{$AFT(t,r) < Deadline(t)$}
@@
         \Else
-            \State provision a new resource r
+            \State Provision a new resource r
             \State Schedule(t,r)
@@
 \begin{algorithm}
-\caption{Deadline Distribution}
+\caption{Deadline distribution algorithm}
 \label{alg:deadlinedistribution}
 \begin{algorithmic}[1]
+\Require w: DAG, d: global deadline
+\Ensure Assign sub-deadlines to each task in w
 \Procedure{DeadlineDistribution}{$w,d$}

### 2012-01-10 src/cws/core/dag/DAG.java
Add some sanity checking to DAG parser

@@ -22,29 +22,33 @@ public class DAG {
     }
     
     public void addFile(String name, double size) {
+        if (size < 0) {
+            throw new RuntimeException(
+                    "Invalid size for file '"+name+"': "+size);
+        }
         files.put(name, size);
     }
     
     public void addEdge(String parent, String child) {
         Task p = tasks.get(parent);
         if (p == null) {
-            throw new RuntimeException("Invalid parent: "+parent);
+            throw new RuntimeException("Invalid edge: Parent not found: "+parent);
         }
         Task c = tasks.get(child);
         if (c == null) {
-            throw new RuntimeException("Invalid child: "+child);
+            throw new RuntimeException("Invalid edge: Child not found: "+child);
         }
         p.children.add(c);
         c.parents.add(p);
     }
     
     public void setInputs(String task, List<String> inputs) {
-        Task t = tasks.get(task);
+        Task t = getTask(task);
         t.inputs = inputs;
     }
     
     public void setOutputs(String task, List<String> outputs) {
-        Task t = tasks.get(task);
+        Task t = getTask(task);
         t.outputs = outputs;
     }
     
@@ -57,10 +61,16 @@ public class DAG {
     }
     
     public Task getTask(String id) {
+        if (!tasks.containsKey(id)) {
+            throw new RuntimeException("Task not found: "+id);
+        }
         return tasks.get(id);
     }
     
     public double getFileSize(String name) {
+        if (!files.containsKey(name)) {
+            throw new RuntimeException("File not found: "+name);
+        }
         return files.get(name);
     }
     

### 2012-01-10 src/cws/core/dag/Task.java

@@ -20,6 +20,10 @@ public class Task {
         this.id = id;
         this.transformation = transformation;
         this.size = size;
+        if (size <= 0) {
+            throw new RuntimeException(
+                    "Invalid size for task "+id+" ("+transformation+"): "+size);
+        }
     }
     
     /* IT IS IMPORTANT THAT THESE ARE NOT IMPLEMENTED

### 2012-01-10 src/cws/core/SPSS.java
SPSS bug fixes

@@ -50,8 +50,8 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     /** Schedule of tasks for each VM */
     private HashMap<VM, Queue<Task>> vmQueues = new HashMap<VM, Queue<Task>>();
     
-    /** Set of busy VMs */
-    private HashSet<VM> busy = new HashSet<VM>();
+    /** Set of idle VMs */
+    private HashSet<VM> idle = new HashSet<VM>();
     
     public SPSS(double budget, double deadline, List<DAG> dags) {
         this.budget = budget;
@@ -121,19 +121,18 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     public void plan() {
         // We assume the dags are in priority order
         for (DAG dag : allDAGs) {
-            Plan newPlan = planDAG(dag, plan);
-            if (newPlan == null) {
-                // Plan was not feasible
-                System.out.println("Rejecting DAG: plan not feasible");
-            } else {
+            try {
+                Plan newPlan = planDAG(dag, plan);
                 // Plan was feasible
                 if (newPlan.getCost() <= budget) {
                     admittedDAGs.add(dag);
                     plan = newPlan;
                     System.out.println("Admitting DAG. Cost of new plan "+plan.getCost());
                 } else {
-                    System.out.println("Rejecting DAG: budget exceeded");
+                    System.out.println("Rejecting DAG: New plan exceeds budget");
                 }
+            } catch (NoFeasiblePlan m) {
+                System.out.println("Rejecting DAG: "+m.getMessage());
             }
         }
         
@@ -172,15 +171,16 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     /**
      * Develop a plan for a single DAG
      */
-    Plan planDAG(DAG dag, Plan currentPlan) {
+    Plan planDAG(DAG dag, Plan currentPlan) throws NoFeasiblePlan {
         TopologicalOrder order = new TopologicalOrder(dag);
         
         // Initial task assignment
         HashMap<Task, VMType> vmTypes = new HashMap<Task, VMType>();
         HashMap<Task, Double> runtimes = new HashMap<Task, Double>();
+        double minCost = 0.0;
         for (Task t : order) {
             
-            // FIXME For now assign each task to the SMALL VM type
+            // Initially we assign each VM to the SMALL type
[…]

### 2012-01-10 src/cws/core/SPSS.java
Make alpha parameter explicit

@@ -53,10 +53,14 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     /** Set of idle VMs */
     private HashSet<VM> idle = new HashSet<VM>();
     
-    public SPSS(double budget, double deadline, List<DAG> dags) {
+    /** Tuning parameter for deadline distribution (low alpha = runtime, high alpha = tasks) */
+    private double alpha;
+    
+    public SPSS(double budget, double deadline, List<DAG> dags, double alpha) {
         this.budget = budget;
         this.ensembleDeadline = deadline;
         this.allDAGs = dags;
+        this.alpha = alpha;
     }
     
     public void setCloud(Cloud c) {
@@ -208,7 +212,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
         double criticalPath = path.getCriticalPathLength();
         if (criticalPath > ensembleDeadline) {
             throw new NoFeasiblePlan(
-                    "Cannot plan DAG: best critical path ("+criticalPath+") " +
+                    "Best critical path ("+criticalPath+") " +
                     "> deadline ("+ensembleDeadline+")");
         }
         
@@ -223,7 +227,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
         
         // Get deadlines for each task (deadline distribution)
         final HashMap<Task, Double> deadlines = 
-                deadlineDistribution(order, runtimes, 0.7);
+                deadlineDistribution(order, runtimes, this.alpha);
         
         // Sort tasks by deadline
         LinkedList<Task> sortedTasks = new LinkedList<Task>();
@@ -659,11 +663,11 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
         
         List<DAG> dags = new ArrayList<DAG>();
         for (int i = 0; i < 10; i++) {
-            DAG dag = DAGParser.parseDAG(new File("dags/Inspiral_1000.dag"));
+            DAG dag = DAGParser.parseDAG(new File("dags/Montage_25.dag"));
             dags.add(dag);
         }
         
-        SPSS spss = new SPSS(23.50, 40000, dags);
+        SPSS spss = new SPSS(0.45, 300, dags, 0.7);
         
         Cloud cloud = new Cloud();
         WorkflowEngine engine = new WorkflowEngine(spss, spss);
[…]

### 2012-01-10 src/cws/core/experiment/DAGListGenerator.java
Ensemble generator based on Pareto distribution

@@ -1,5 +1,13 @@
 package cws.core.experiment;
 
+import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.Collections;
+import java.util.HashMap;
+import java.util.Random;
+
+import org.cloudbus.cloudsim.distributions.ParetoDistr;
+
  /**
   * Class for generating list of DAG files for experiments.
   * @author malawski
@@ -32,4 +40,39 @@ public class DAGListGenerator {
 		}
 		return dags;
 	}
+	
+	public static String[] generateDAGListPareto(Random seed, String name, int length) {
+		ArrayList<String> dags = new ArrayList<String>(length);
+
+		ParetoDistr pareto = new ParetoDistr(seed, 1, 50);
+		HashMap<Integer, Integer> distr = new HashMap<Integer, Integer>();
+		for (int i = 0; i < length; i++) {
+			double d = pareto.sample();
+			int n;
+			if (d < 100)
+				n = 50;
+			else if (d > 1000)
+				n = 1000;
+			else
+				n = (int) Math.floor(d / 100) * 100;
+			if (!distr.containsKey(n))
+				distr.put(n, 1);
+			else
+				distr.put(n, distr.get(n) + 1);
+			System.out.println(d + "\t" + n + "\t");
+		}
+		Integer[] sizes = distr.keySet().toArray(new Integer[0]);
+		Arrays.sort(sizes);
+		for (int size : sizes) {
+			int count = distr.get(size);
+			for (int i = 0; i < count; i++) {
+				String dag = name + ".n." + size + "." + i % 20 + ".dag";
+				System.out.println(dag);
+				dags.add(dag);
+			}
+		}
+		Collections.reverse(dags);
+		return dags.toArray(new String[0]);
+	}
+	
 }

### 2012-01-10 src/cws/core/experiment/DAGListGeneratorTest.java

@@ -2,6 +2,10 @@ package cws.core.experiment;
 
 import static org.junit.Assert.*;
 
+import java.util.HashMap;
+import java.util.Random;
+
+import org.cloudbus.cloudsim.distributions.ParetoDistr;
 import org.junit.Test;
 
 public class DAGListGeneratorTest {
@@ -9,8 +13,24 @@ public class DAGListGeneratorTest {
 	@Test
 	public void testGenerateDAGList() {
 		String[] dags = DAGListGenerator.generateDAGList("SIPHT", new int [] {900,1000}, 20);
-		assertEquals("SIPHT.n.900.0.dax", dags[0]);
-		assertEquals("SIPHT.n.1000.0.dax", dags[20]);
-		assertEquals("SIPHT.n.1000.19.dax", dags[39]);	
+		assertEquals("SIPHT.n.900.0.dag", dags[0]);
+		assertEquals("SIPHT.n.1000.0.dag", dags[20]);
+		assertEquals("SIPHT.n.1000.19.dag", dags[39]);	
+	}
+	
+	@Test
+	public void testPareto() {
+		String[] dags = DAGListGenerator.generateDAGListPareto(new Random(0), "SIPHT", 20);
+		assertEquals("SIPHT.n.1000.0.dag", dags[0]);
+		assertEquals("SIPHT.n.50.0.dag", dags[19]);
+		assertEquals("SIPHT.n.300.1.dag", dags[2]);	
+	}
+	
+	@Test
+	public void testPareto40() {
+		String[] dags = DAGListGenerator.generateDAGListPareto(new Random(0), "SIPHT", 40);
+		assertEquals("SIPHT.n.1000.1.dag", dags[0]);
+		assertEquals("SIPHT.n.50.0.dag", dags[39]);
+		assertEquals("SIPHT.n.300.1.dag", dags[7]);	
 	}
 }

### 2012-01-11 paper/paper.tex
Fix SPSS DD algorithm
added SPDS algorithm and example

@@
 \section{Algorithms}
-\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
+
+\subsection{Static Provisioning Dynamic Scheduling (SPDS)}
+
+This is the simplest strategy based on static provisioning of resources and
+on-line scheduling of workflow tasks. Given the budget $B$ and deadline $d$
+together with the hourly price $p$ of a VM it is possible to compute the number
+of VMs $N_{VM}$ to provision so that they consume the whole budget between the
+deadline.
+
+\begin{equation}
+\label{eq:static-plan}
+N_{VM} = \lceil B / d / p \rceil
+\end{equation}
+
+Although simple, such provisioning plan has the advantage that it
+minimizes the number of provisioning and deprovisioning requests.
+Once the resources are provisioned, the tasks are scheduled to VMs using dynamic
+priority-based algorithm shown in Algorithm~\ref{alg:ds}.
 
@@
 
+Initially all the
+released tasks from all the workflows are submitted to the queue $Q$. Then, they
+are inserted into priority queue $P$ based on the priority of workflows. If
+there are free VMs available, the tasks from the priority queue are submitted to
+the randomly choosen VMs. The process is repeated each time a task finishes.
+Using the priority queue guarantees that tasks from lower priority workflows are
+always submitted later than the higher priority ones and that the lower priority
+tasks can occupy idle VMs. However, as there is no preemption, long-running
+low-priority tasks may delay execution of high-priority tasks when such ones
+become ready. Fig.~\ref{fig:spss-example} shows the example plan generated using
+SPSS algorithm.
+
+\begin{figure}[htb] 
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
+ \caption{Example of schedule generated using SPSS algorithm: tasks are labelled
+ by colors depending on the workflow. }
+\label{fig:spss-example}
+\end{figure}
+
+\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
+
 \subsection{Workflow-Aware DPDS (WA-DPDS)}
@@
 \begin{algorithmic}[1]
-\Require W: workflow ensemble, b: budget, d: deadline
-\Ensure Schedule as much of W as possible given b and d
+\Require $W$: workflow ensemble; $b$: budget; $d$: deadline
+\Ensure Schedule as much of $W$ as possible given $b$ and $d$
 \Procedure{PlanEnsemble}{$W,b,d$}
@@
 \begin{algorithmic}[1]
-\Require $w$: DAG, $P$: current plan, $d$: deadline
+\Require $w$: DAG; $P$: current plan; $d$: deadline
 \Ensure Create plan for $w$ that minimizes cost and meets deadline $d$
@@
 \begin{algorithmic}[1]
-\Require w: DAG, d: global deadline
+\Require $w$: DAG; $d$: global deadline
 \Ensure Assign sub-deadlines to each task in w
 \Procedure{DeadlineDistribution}{$w,d$}
-    \State assign $Level(t)$ to each task t in w
-    \For{$level\ l\ in\ w$}
-        \State $Tasks(l) \gets$ number of tasks in l
-        \State $Runtime(l) \gets$ runtime of tasks in l
-    \EndFor
-    \State $T \gets$ total number of tasks in w
-    \State $R \gets$ total runtime of tasks in w
-    \State $CP \gets$ critical path of w
-    \For{$level\ l\ in\ w$}
+    \State $T \gets$ total number of tasks in $w$
+    \State $R \gets$ total runtime of tasks in $w$
+    \State $CP \gets$ length of critical path of $w$
+    \State assign $Level(t)$ to each task $t$ in $w$
+    \For{level $l\ \in w$}
+        \State $Tasks(l) \gets$ number of tasks in $l$
+        \State $Runtime(l) \gets$ runtime of tasks in $l$
         \State $r \gets \alpha * (Runtime(l)/R)$
@@
     \EndFor
-    \For{$task\ t\ in\ w$}
-        \State $est \gets 0$
-        \For{p in Pred(t)}
-            \State $est \gets Deadline(p)$
-        \EndFor
-        \State $Deadline(t) \gets est + RT(t) + FT(Level(t))$
+    \For{task $t \in w$}
+        \State $LST(t) \gets \max(Deadline(p))\ for\ p \in Pred(t)$
+        \State $Deadline(t) \gets LST(t) + RT(t) + FT(Level(t))$
     \EndFor
@@
 \end{algorithm}
+
+\begin{equation}
+\label{eq:scaling}
+FT(l) = \left[{\alpha}*\frac{T_l}{t}\right] + \left[{(1 - \alpha)}*\frac{R_l}{r} \right]
+\end{equation}
 

### 2012-01-11 scripts/R/stats.R
nicer plots in PDF

@@ -16,24 +16,28 @@ dpdsStats <- function(dag, budgets, deadline, max_scaling, runIds={0}) {
 	
 	m <- array(0,  dim=c(n_runIds, n_budgets, n_deadlines))
 	mA <- array(0,  dim=c(n_runIds, n_budgets, n_deadlines))
+	mS <- array(0,  dim=c(n_runIds, n_budgets, n_deadlines))
 	
 	for (runId in runIds) {
 	
 		files <- paste(path, dag, "b", budgets, ".0h", deadline, "m", max_scaling, ".0", "run", runId, "-outputSimple.txt", sep="")
 		filesA <- paste(path, dag, "b", budgets, ".0h", deadline, "m", max_scaling, ".0", "run", runId, "-outputAware.txt", sep="")
+		filesS <- paste(path, dag, "b", budgets, ".0h", deadline, "m", max_scaling, ".0", "run", runId, "-outputSPSS.txt", sep="")
 
 		deadlines <- read.table(files[1])$V1
 		hours <- deadlines /3600
 		for (i in 1:n_budgets) {
-			#print(files[i])
-			m[runId,i,] = read.table(files[i])$V2	
-			mA[runId,i,] = read.table(filesA[i])$V2	
+			print(files[i])
+			m[runId+1,i,] = read.table(files[i])$V2	
+			mA[runId+1,i,] = read.table(filesA[i])$V2
+			mS[runId+1,i,] = read.table(filesS[i])$V2
 		}
 	}
-		#print(mA)
+		print(m)
 	
 	avg_m <- apply(m,c(2,3),mean)
 	avg_mA <- apply(mA,c(2,3),mean)
+	avg_mS <- apply(mS,c(2,3),mean)
 	
 	dag_cost <- 1/(avg_m/budgets)
 	dag_costA <- 1/(avg_mA/budgets)
@@ -41,6 +45,7 @@ dpdsStats <- function(dag, budgets, deadline, max_scaling, runIds={0}) {
 	print(avg_m[,1:5])
 	print(dag_cost[,1:5])
 	
+	print(avg_m)
 	
 	diff <- avg_mA-avg_m
 	positive <- diff>0
[…]

### 2012-01-11 scripts/ruby/plot_schedule_gantt.rb
plot new tests

@@ -127,4 +127,16 @@ require "#{File.dirname(__FILE__)}/plots.rb"
 
 
 #plot_schedule 'output/testSimpleUtilizationBasedProvisionerEnsembleDynamicSchedulerpsload_large.dagx40d252000.0b600.0m0.0'
-plot_schedule 'output/testSimpleUtilizationBasedProvisionerEnsembleDynamicSchedulerpsload_medium.dagx40d122400.0b100.0m0.0'
+#plot_schedule 'output/testSimpleUtilizationBasedProvisionerEnsembleDynamicSchedulerpsload_medium.dagx40d122400.0b100.0m0.0'
+
+
+#plot_schedule 'output/testSimpleUtilizationBasedProvisionerWorkflowAwareEnsembleSchedulerMONTAGE.n.1000.8.dagx100d68400.0b40.0m0.0'
+#plot_schedule 'output/testSimpleUtilizationBasedProvisionerEnsembleDynamicSchedulerMONTAGE.n.1000.8.dagx100d68400.0b40.0m0.0'
+#plot_schedule 'output/testSimpleUtilizationBasedProvisionerWorkflowAwareEnsembleSchedulerCYBERSHAKE.n.1000.8.dagx100d43200.0b100.0m0.0'
+
+#plot_schedule 'output/testSimpleUtilizationBasedProvisionerWorkflowAwareEnsembleSchedulerCYBERSHAKE.n.1000.8.dagx100d50400.0b100.0m0.0'
+#plot_schedule 'output/testSimpleUtilizationBasedProvisionerEnsembleDynamicSchedulerCYBERSHAKE.n.1000.8.dagx100d50400.0b100.0m0.0'
+
+plot_schedule 'output/testSPSSSPSSCYBERSHAKE.n.1000.8.dagx100d50400.0b100.0m0.0'
+plot_schedule 'output/testSimpleUtilizationBasedProvisionerWorkflowAwareEnsembleSchedulerCYBERSHAKE.n.1000.8.dagx100d50400.0b100.0m0.0'
+plot_schedule 'output/testSimpleUtilizationBasedProvisionerEnsembleDynamicSchedulerCYBERSHAKE.n.1000.8.dagx100d50400.0b100.0m0.0'

### 2012-01-11 scripts/ruby/plots.rb

@@ -96,8 +96,9 @@ def plot_schedule (filename)
       #puts "Saving plot to file: " + filename + ".pdf"
       plot.set "key right outside"
       plot.set "grid"
-      #plot.terminal "png size 1024,768"
+      plot.terminal "png size 1024,768"
       #plot.output 'output/' + dag + deadline + ".png"
+      plot.output filename + ".png"
 
       data = Array.new
 

### 2012-01-11 src/cws/core/experiment/Experiment.java
some hacks to run SPSS

@@ -11,6 +11,8 @@ import org.cloudbus.cloudsim.core.CloudSim;
 import cws.core.Cloud;
 import cws.core.DAGJob;
 import cws.core.EnsembleManager;
+import cws.core.Provisioner;
+import cws.core.SPSS;
 import cws.core.Scheduler;
 import cws.core.SimpleJobFactory;
 import cws.core.VM;
@@ -34,18 +36,28 @@ public class Experiment {
 	
 	public ExperimentResult runExperiment(ExperimentDescription param) {
 		
+		boolean useSPSS = false;
+		
 		ExperimentResult result = new ExperimentResult();
 		
-		AbstractProvisioner prov = (AbstractProvisioner)param.getProvisioner();
+		Provisioner prov = param.getProvisioner();
 		
 		Scheduler sched = param.getScheduler();
 		
+		if (sched.getClass().getName() == "cws.core.SPSS") useSPSS = true;
+		
 		CloudSim.init(1, null, false);
+
 		
 		Cloud cloud = new Cloud();
 		prov.setCloud(cloud);
 		
-		WorkflowEngine engine = new WorkflowEngine(new SimpleJobFactory(1000), prov, sched);
+		WorkflowEngine engine;
+		
+		if (useSPSS)  engine = new WorkflowEngine(new SimpleJobFactory(1), prov, sched);
+		else engine = new WorkflowEngine(new SimpleJobFactory(1000), prov, sched);
+		
+		sched.setWorkflowEngine(engine);
 		
 		List<DAG> dags = new ArrayList<DAG>();
 		for (int i = 0; i < param.getDags().length; i++) {
@@ -69,12 +81,19 @@ public class Experiment {
 		Log.printLine(CloudSim.clock() + " Estimated num of VMs " + numVMs);
 		Log.printLine(CloudSim.clock() + " Total budget " + param.getBudget());
 		
-		HashSet<VM> vms = new HashSet<VM>();
-		for (int i = 0; i < numVMs; i++) {
-			VM vm = new VM(1000, 1, 1.0, param.getPrice());
-			vms.add(vm);
-			CloudSim.send(engine.getId(), cloud.getId(), 0.0, WorkflowEvent.VM_LAUNCH, vm);
-		}	
+		if (!useSPSS) { 
+			HashSet<VM> vms = new HashSet<VM>();
+			for (int i = 0; i < numVMs; i++) {
+				VM vm = new VM(1000, 1, 1.0, param.getPrice());
+				vms.add(vm);
[…]

### 2012-01-11 src/cws/core/scheduler/WorkflowAwareEnsembleScheduler.java
fix estimator bug

@@ -130,6 +130,7 @@ public class WorkflowAwareEnsembleScheduler extends EnsembleDynamicScheduler {
 
 		// remaining budget for starting new vms 
 		double rn = engine.getBudget()-engine.getCost();
+		if (rn<0) rn=0;
 
 		// compute remaining (not consumed) budget of currently running VMs
 		double rc = 0.0;
@@ -153,6 +154,8 @@ public class WorkflowAwareEnsembleScheduler extends EnsembleDynamicScheduler {
 		
 		// we add this for safety in order not to underestimate our budget
 		double safetyMargin = 0.1;
+		
+		Log.printLine(CloudSim.clock() + " Budget for new VMs: " + rn + " Budget on running VMs: " + rc + " Remaining budget of admitted workflows: " + ra);
 		
 		return rn+rc-ra-safetyMargin;
 	}

### 2012-01-11 src/cws/core/scheduler/DAGDynamicScheduler.java
adjustments to interfaces

@@ -54,4 +54,11 @@ public class DAGDynamicScheduler implements Scheduler, WorkflowEvent {
 		}
 	}
 
+
+	@Override
+	public void setWorkflowEngine(WorkflowEngine engine) {
+		// TODO Auto-generated method stub
+		
+	}
+
 }

### 2012-01-11 src/cws/core/scheduler/DAGSchedulerFCFS.java

@@ -45,6 +45,12 @@ public class DAGSchedulerFCFS implements Scheduler, WorkflowEvent {
 				}
 		}			
 	}
+
+	@Override
+	public void setWorkflowEngine(WorkflowEngine engine) {
+		// TODO Auto-generated method stub
+		
+	}
 	
 
 }

### 2012-01-11 src/cws/core/Provisioner.java
interface adjustment

@@ -7,4 +7,5 @@ package cws.core;
  */
 public interface Provisioner {
     public void provisionResources(WorkflowEngine engine);
+    public void setCloud(Cloud cloud);
 }

### 2012-01-11 src/cws/core/Scheduler.java
interface adjustment

@@ -7,4 +7,5 @@ package cws.core;
  */
 public interface Scheduler {
     public void scheduleJobs(WorkflowEngine engine);
+    public void setWorkflowEngine(WorkflowEngine engine);
 }

### 2012-01-11 src/cws/core/SPSS.java
added priorities for plotting

@@ -33,6 +33,8 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     private double budget;
     private double ensembleDeadline;
     
+    private int priority = 0;
+    
     private Plan plan = new Plan();
     
     /** List of all dags in ensemble */
@@ -531,6 +533,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     
     private void submitDAG(DAG dag) {
         DAGJob dagJob = new DAGJob(dag, manager.getId());
+        dagJob.setPriority(priority++);
         CloudSim.send(manager.getId(), engine.getId(), 0.0, DAG_SUBMIT, dagJob);
     }
     
@@ -696,7 +699,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     }
     
     enum VMType {
-        SMALL(1, 0.10),
+        SMALL(1, 1.0),
         MEDIUM(5, 0.40),
         LARGE(10, 0.80);
         

### 2012-01-11 src/cws/core/WorkflowEngine.java
minor fix to activate provisioner

@@ -137,7 +137,8 @@ public class WorkflowEngine extends SimEntity implements WorkflowEvent {
     @Override
     public void startEntity() {
         // send the first provisioning request
-        send(this.getId(), 0.0, PROVISIONING_REQUEST);
+    	// it should be sent after dag submissions events
+        send(this.getId(), 0.0001, PROVISIONING_REQUEST);
     }
     
     @Override

### 2012-01-11 src/cws/core/experiment/ExperimentParetoTestRepeat.java
added new tests based on Pareto sampling

@@ -1,0 +1,178 @@
+package cws.core.experiment;
+
+
+import java.util.Random;
+
+import org.junit.Test;
+
+
+/**
+ * Tests series of ensembles consisting of workflows form workflow generator  
+ * https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator
+ * 
+ * From each workflow type DAGs of sizes 1000 and 900 are selected. 
+ * For smaller workflows (Montage, Cybershake) also smaller sizes (down to 500) are added to ensembles.
+ * 
+ * @author malawski
+ *
+ */
+
+public class ExperimentParetoTestRepeat {
+    
+	String dagPath;
+	String dagName;
+	double budget;
+	double price = 1.0;
+	int numDAGs = 40;
+	double max_scaling = 2.0;
+
+	
+	int N;
+	int step;
+	int start;
+	
+	/******************************
+	 * 
+	 * Tests with max scaling = 0.0
+	 * 
+	 ******************************/
+	
+	@Test
+	public void testRunExperimentMontage0() {		
+
+		dagPath = "../projects/pegasus/Montage/";
+		dagName = "MONTAGE";
+		
+		String[] dags = DAGListGenerator.generateDAGListPareto(new Random(0), dagName, 100);
+		N = 20;
+		step = 1;
+		start = 1;
+		max_scaling = 0;
+		
+		double[] budgets= {40.0, 80.0, 120.0, 160.0, 200.0};
+		
+		for (double budget : budgets) {
+			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 1);			
+		}		
+		
+	}
+
+	@Test
+	public void testRunExperimentCybershake0() {		
+
+		dagPath = "../projects/pegasus/CyberShake/";
+		dagName = "CYBERSHAKE";
+		
+		String[] dags = DAGListGenerator.generateDAGListPareto(new Random(0), dagName, 100);
+		
+		N = 20;
+		step = 1;
+		start = 1;
+		max_scaling = 0;
+
+		
+		double[] budgets= {40.0, 60.0, 80.0, 100.0, 120.0};
+		
+		for (double budget : budgets) {
+			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 1);			
+		}
+		
+	}
+
+	
+	@Test
[…]

### 2012-01-11 src/cws/core/experiment/ExperimentTestRepeat900_1000.java

@@ -51,7 +51,7 @@ public class ExperimentTestRepeat900_1000 {
 		double[] budgets= {40.0, 80.0, 120.0, 160.0, 200.0};
 		
 		for (double budget : budgets) {
-			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 10);			
+			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 20);			
 		}		
 		
 	}
@@ -73,7 +73,7 @@ public class ExperimentTestRepeat900_1000 {
 		double[] budgets= {40.0, 80.0, 120.0, 160.0, 200.0};
 		
 		for (double budget : budgets) {
-			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 10);			
+			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 20);			
 		}
 		
 	}
@@ -95,7 +95,7 @@ public class ExperimentTestRepeat900_1000 {
 		double[] budgets= {400.0, 800.0, 1200.0, 1600.0, 2000.0};
 
 		for (double budget : budgets) {
-			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 10);			
+			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 20);			
 		}
 		
 	}
@@ -118,7 +118,7 @@ public class ExperimentTestRepeat900_1000 {
 		double[] budgets= {4000.0, 8000.0, 12000.0, 16000.0, 20000.0};
 		
 		for (double budget : budgets) {
-			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 10);			
+			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 20);			
 		}
 			
 	}
@@ -140,7 +140,7 @@ public class ExperimentTestRepeat900_1000 {
 		double[] budgets= {200.0, 400.0, 600.0, 800.0, 1000.0};
 
 		for (double budget : budgets) {
-			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 10);			
+			Experiment.runSeriesRepeat(dagPath, dags, budget, price, N, step, start, max_scaling, 20);			
 		}
 		
[…]

### 2012-01-12 src/cws/core/SPSS.java
fixed priority assignment

@@ -32,9 +32,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     private Cloud cloud;
     private double budget;
     private double ensembleDeadline;
-    
-    private int priority = 0;
-    
+        
     private Plan plan = new Plan();
     
     /** List of all dags in ensemble */
@@ -533,7 +531,8 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     
     private void submitDAG(DAG dag) {
         DAGJob dagJob = new DAGJob(dag, manager.getId());
-        dagJob.setPriority(priority++);
+        int priority = allDAGs.indexOf(dag);
+        dagJob.setPriority(priority);
         CloudSim.send(manager.getId(), engine.getId(), 0.0, DAG_SUBMIT, dagJob);
     }
     

### 2012-01-12 src/cws/core/SPSS.java
Update SPSS

@@ -21,6 +21,7 @@ import cws.core.dag.Task;
 import cws.core.dag.DAGParser;
 import cws.core.dag.algorithms.CriticalPath;
 import cws.core.dag.algorithms.TopologicalOrder;
+import cws.core.log.WorkflowLog;
 
 /**
  * @author Gideon Juve <juve@usc.edu>
@@ -133,9 +134,9 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
                 if (newPlan.getCost() <= budget) {
                     admittedDAGs.add(dag);
                     plan = newPlan;
-                    System.out.println("Admitting DAG. Cost of new plan "+plan.getCost());
+                    System.out.println("Admitting DAG. Cost of new plan: "+plan.getCost());
                 } else {
-                    System.out.println("Rejecting DAG: New plan exceeds budget");
+                    System.out.println("Rejecting DAG: New plan exceeds budget: "+newPlan.getCost());
                 }
             } catch (NoFeasiblePlan m) {
                 System.out.println("Rejecting DAG: "+m.getMessage());
@@ -184,6 +185,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
         HashMap<Task, VMType> vmTypes = new HashMap<Task, VMType>();
         HashMap<Task, Double> runtimes = new HashMap<Task, Double>();
         double minCost = 0.0;
+        double totalRuntime = 0.0;
         for (Task t : order) {
             
             // Initially we assign each VM to the SMALL type
@@ -197,7 +199,11 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
             
             // Compute the minimum cost of running this workflow
             minCost += (runtime/(60*60)) * vm.price;
+            totalRuntime += runtime;
         }
+        
+        System.out.println(" Min Cost: "+minCost);
+        System.out.println(" Total Runtime: "+totalRuntime);
         
         // Check to make sure the minimum cost is within our budget
         // This is just an optimization
@@ -212,6 +218,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
         // FIXME Later we will assign each task to its fastest VM type before this
         CriticalPath path = new CriticalPath(order, runtimes);
         double criticalPath = path.getCriticalPathLength();
+        System.out.println(" Critical path: "+criticalPath);
         if (criticalPath > ensembleDeadline) {
[…]

### 2012-01-12 src/cws/core/dag/algorithms/TopologicalOrder.java
Add reverse topological sort

@@ -42,6 +42,15 @@ public class TopologicalOrder implements Iterable<Task> {
         postorder.add(task);
     }
     
+    public Iterable<Task> reverse() {
+        return new Iterable<Task>() {
+            @Override
+            public Iterator<Task> iterator() {
+                return postorder.iterator();
+            }
+        };
+    }
+    
     @Override
     public Iterator<Task> iterator() {
         return postorder.descendingIterator();

### 2012-01-13 scripts/correct_sipht.py
Add script to correct SIPHT workflows

@@ -1,0 +1,75 @@
+#!/usr/bin/env python
+
+import sys
+import os
+
+if len(sys.argv) != 3:
+    print """
+This script corrects the two problems with SIPHT workflows: 
+    1) It sets the runtime to 33.00 for Blast_synteny, and
+    2) it removes all Last_transfer jobs
+
+It works on both .dax and .dag files
+"""
+    print "Usage: python %s OLDFILE NEWFILE" % sys.argv[0]
+    exit(1)
+
+oldfile = os.path.abspath(sys.argv[1])
+newfile = os.path.abspath(sys.argv[2])
+
+if not os.path.isfile(oldfile):
+    print "%s does not exist" % olddir
+
+print oldfile
+print newfile
+
+dax = oldfile.endswith(".dax")
+
+infile = open(oldfile, "r")
+outfile = open(newfile, "w")
+
+inltjob = False
+inltchild = False
+ltids = []
+for l in infile:
+    if dax:
+        if "Blast_synteny" in l:
+            l = l.replace('runtime="0.00"','runtime="33.00"')
+        elif "Last_transfer" in l:
+            inltjob = True
+            s = l.index('"') + 1
+            e = l.index('"', s)
+            ltids.append(l[s:e])
+        
+        for ltid in ltids:
+            if '<child ref="%s"' % ltid in l:
+                inltchild = True
+        
+        if not inltjob and not inltchild:
+            outfile.write(l)
+        
+        if inltjob and "</job>" in l:
+            inltjob = False
+        
+        if inltchild and "</child>" in l:
+            inltchild = False
+    else:
+        if "Blast_synteny" in l:
+            rec = l.split()
+            rec[-1] = "33.00\n"
+            l = " ".join(rec)
+        elif "Last_transfer" in l:
+            rec = l.split()
+            ltids.append(rec[1])
+            l = None
+        else:
+            for ltid in ltids:
+                if ltid in l:
+                    l = None
+                    break
+        
+        if l is not None:
+            outfile.write(l)
+
+infile.close()
+outfile.close()

### 2012-01-13 src/cws/core/SPSS.java
Go back to old SPSS

@@ -234,7 +234,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
         
         // Get deadlines for each task (deadline distribution)
         final HashMap<Task, Double> deadlines = 
-                deadlineDistribution2(order, runtimes, this.alpha);
+                deadlineDistribution(order, runtimes, this.alpha);
         
         // Sort tasks by deadline
         LinkedList<Task> sortedTasks = new LinkedList<Task>();
@@ -801,15 +801,6 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
             if (this.resource.schedule.size() > 0 && other.resource.schedule.size() == 0)
                 return true;
             
-            
-            // New resources are better
-            /*
-            if (this.resource.schedule.size() == 0 && other.resource.schedule.size() > 0)
-                return true;
-            if (this.resource.schedule.size() > 0 && other.resource.schedule.size() == 0)
-                return false;
-            */
-            
             // Earlier starts are better
             if (this.slot.start < other.slot.start)
                 return true;
@@ -858,12 +849,12 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
         
         List<DAG> dags = new ArrayList<DAG>();
         for (int i = 0; i < 10; i++) {
-            DAG dag = DAGParser.parseDAG(new File("dags/Montage_25.dag"));
+            DAG dag = DAGParser.parseDAG(new File("dags/Sipht_1000.dag"));
             dags.add(dag);
         }
         
-        double deadline = 400;
-        double budget = 4.5;
+        double deadline = 10*3600;
+        double budget = 300;
         double alpha = 0.7;
         
         SPSS spss = new SPSS(budget, deadline, dags, alpha);
@@ -882,6 +873,12 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
         manager.addDAGJobListener(log);
         
         spss.plan();
+        
+        try {
+            //Thread.sleep(10000);
+        } catch (Exception e) {
+            
+        }
         
         CloudSim.startSimulation();
         

### 2012-01-13 src/cws/core/SPSS.java
Version of SPSS where earliest start time of first level varies

@@ -1,6 +1,9 @@
 package cws.core;
 
+import java.io.BufferedReader;
 import java.io.File;
+import java.io.FileReader;
+import java.io.InputStreamReader;
 import java.util.ArrayList;
 import java.util.Collections;
 import java.util.Comparator;
@@ -203,15 +206,6 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
         System.out.println(" Min Cost: "+minCost);
         System.out.println(" Total Runtime: "+totalRuntime);
         
-        // Check to make sure the minimum cost is within our budget
-        // This is just an optimization
-        if (currentPlan.getCost() + minCost > budget) {
-            throw new NoFeasiblePlan(
-                    "Minimum cost of workflow ("+minCost+") " +
-                    "exceeds budget ("+budget+") given " +
-                    "current plan cost ("+currentPlan.getCost()+")");
-        }
-        
         // Make sure a plan is feasible given the deadline and available VMs
         // FIXME Later we will assign each task to its fastest VM type before this
         CriticalPath path = new CriticalPath(order, runtimes);
@@ -232,9 +226,32 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
          *  }
          */
         
+        
+        Plan best = null;
+        
+        double LST = ensembleDeadline - criticalPath;
+        
+        for (double EST = 0.0; EST <= LST; EST += 3600.0) {
+            Plan newPlan = planDAG(dag, currentPlan, runtimes, vmTypes, EST);
+            if (best == null) {
+                best = newPlan;
+            } else {
+                if (newPlan.getCost() < best.getCost()) {
+                    best = newPlan;
+                }
+            }
+        }
+        
+        return best;
+    }
+    
+    Plan planDAG(DAG dag, Plan currentPlan, HashMap<Task, Double> runtimes, HashMap<Task, VMType> vmTypes, double EST) {
+        
+        TopologicalOrder order = new TopologicalOrder(dag);
+        
         // Get deadlines for each task (deadline distribution)
         final HashMap<Task, Double> deadlines = 
-                deadlineDistribution(order, runtimes, this.alpha);
+                deadlineDistributionEST(order, runtimes, this.alpha, EST);
         
         // Sort tasks by deadline
         LinkedList<Task> sortedTasks = new LinkedList<Task>();
@@ -315,7 +332,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
[…]

### 2012-01-13 src/cws/core/SPSS.java
Fix SPSS mistake

@@ -1,9 +1,6 @@
 package cws.core;
 
-import java.io.BufferedReader;
 import java.io.File;
-import java.io.FileReader;
-import java.io.InputStreamReader;
 import java.util.ArrayList;
 import java.util.Collections;
 import java.util.Comparator;
@@ -13,6 +10,7 @@ import java.util.LinkedList;
 import java.util.List;
 import java.util.HashMap;
 import java.util.Queue;
+import java.util.Random;
 import java.util.SortedSet;
 import java.util.TreeMap;
 
@@ -24,6 +22,7 @@ import cws.core.dag.Task;
 import cws.core.dag.DAGParser;
 import cws.core.dag.algorithms.CriticalPath;
 import cws.core.dag.algorithms.TopologicalOrder;
+import cws.core.experiment.DAGListGenerator;
 import cws.core.log.WorkflowLog;
 
 /**
@@ -36,7 +35,7 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
     private Cloud cloud;
     private double budget;
     private double ensembleDeadline;
-        
+    
     private Plan plan = new Plan();
     
     /** List of all dags in ensemble */
@@ -226,32 +225,16 @@ public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, 
          *  }
          */
         
-        
-        Plan best = null;
-        
-        double LST = ensembleDeadline - criticalPath;
-        
-        for (double EST = 0.0; EST <= LST; EST += 3600.0) {
-            Plan newPlan = planDAG(dag, currentPlan, runtimes, vmTypes, EST);
-            if (best == null) {
-                best = newPlan;
-            } else {
-                if (newPlan.getCost() < best.getCost()) {
-                    best = newPlan;
-                }
-            }
-        }
-        
-        return best;
+        return planDAG(dag, currentPlan, runtimes, vmTypes);
     }
     
-    Plan planDAG(DAG dag, Plan currentPlan, HashMap<Task, Double> runtimes, HashMap<Task, VMType> vmTypes, double EST) {
+    Plan planDAG(DAG dag, Plan currentPlan, HashMap<Task, Double> runtimes, HashMap<Task, VMType> vmTypes) {
         
         TopologicalOrder order = new TopologicalOrder(dag);
         
         // Get deadlines for each task (deadline distribution)
         final HashMap<Task, Double> deadlines = 
[…]

### 2012-01-16 paper/paper.tex
added sample results fig
intro and some references added

@@
 \conferenceinfo{HPDC}{'12 Delft, The Netherlands}
-%\CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden - IF NEED BE.
-%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
+% \CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden
+% - IF NEED BE. \crdata{0-12345-67-8/90/01}  % Allows default copyright data
+% (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
 
 \title{Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds}
-%\subtitle{[Extended Abstract]}
+% \subtitle{[Extended Abstract]}
 
@@
 
-Workflows are important but workflow ensembles are more important, see
-CyberShake or other examples.
+Scientific workflows, usually represented as directed acyclic graphs (DAG), are
+the important class of applications that have been studied in the context of
+resource management and scheduling on grid and utility computing systems.
+However, large computational applications are not just individual workflows but
+rather sets of inter-related workflows grouped in {\em ensembles}. All the
+workflows in an ensemble typically have a similar structure, but they differ by
+input data, number of tasks and individual tasks sizes. A good example of
+scientific workflow ensemble comes from the CyberShake
+application~\cite{Callaghan11}, which calculates seismic hazards for a given
+geographical region such as California. In order to produce a hazard map, the
+hazard curves need to be computed for a set of geographical sites, while each
+site requires running a single complex scientific workflow. Workflows for each
+site may differ not only by their parameters, but also by priority; e.g. some of
+the points on the map correspond to urban areas or strategic locations such as
+power plants, whereas others may be located in low populated regions. Another
+examples of workflow ensembles come from astronomical applications, such as
+Montage~\cite{Deelman08} or search for Earth-like planets from Kepler project
+data~\cite{vockler11}, where multiple workflows are typically required to
+process data covering different parts of the sky.
 
-Clouds are interesting and different from grids.
+Infrastructure-as-a-service (Iaas) clouds offer capabilities of creating
+(provisioning) computational resources on demand in a pay-per-use model and are regarded by
+scientific community as a potentially attractive source of computing
+resources~\cite{Ostermann10,Keahey09}. In contrast to clusters and grids which
+typically offer best-effort resource provisioning capabilities, clouds give more
+flexibility in terms of creating a controlled and managed computing environment
+with the ability of adjusting resource capacity to the computing demands, often
+called auto-scaling. Giving the users more control, however, clouds also require
+developing new methods of task scheduling and resource provisioning. The
+resource management decisions in cloud scenarios not only have to take into
+account performance related metrics such as workflow makespan or resource
+utilization, but also budget considerations, since the resources from public
+(commercial) cloud providers are usually not free~\cite{Durkee10}.
 
-Define problems: scheduling and provisioning.
+In this paper, to get insight into these resource management challenges for
+scietific workflow on clouds, we address a new important problem of maximizing
+the number of completed workflows from an ensemble under both budget and
+deadline constraints. The motivation is to answer the fundamental question of a
+researcher: how much work can be completed in a limited budget and timeframe of
+a research project. The goals of this paper are to discuss and assess possible
+static and dynamic strategies for both task scheduling and resource
+provisioning. Based on the knowledge of workflow scheduling algorithms we
+analyze strategies for on-line scheduling of individual tasks as well as static
+algorithms that rely on the information about the workflow structure (critical
+paths and workflow levels) and estimates of task runtimes. In addition, we
+analyze a hybrid workflow-aware dynamic scheduling algorithm, which uses the
+workflow structure information to estimate which workflows should be rejected
+from the ensemble due to the constraints. As methodology in the study of the
+proposed algorithms we use simulation techniques. We have developed cloud workflow simulator based on
+CloudSim~\cite{Calheiros11}, which models the infrastructure and the application. The
+algorithms are subject to evaluation on a set of scientific workflow ensembles with a broad range of
+budget and deadline parameters.
+
+The paper is organized as follows [\ldots]
 
@@
 
+\begin{figure}[htb] 
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/CYBERSHAKE-Pareto}
+\caption{Comparison of DPDS and AW-DPDS algorithms for ensembles of
+ 100 Pareto-distributed Cybershake workflows}
+\label{fig:CYBERSHAKE.n.1000.8.dagh1-20m0}
+\end{figure}
+
 \section{Conclusion and Future Work}

### 2012-01-16 scripts/copy_plots_to_paper.sh
more plots added

@@ -1,0 +1,31 @@
+#!/bin/bash
+cp MONTAGE.n.1000.8.dagh1-20m0.pdf ../paper/figures/MONTAGE-n-1000-8-dagh1-20m0.pdf
+cp CYBERSHAKE.n.1000.8.dagh1-20m0.pdf ../paper/figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf
+cp LIGO.n.1000.8.dagh1-40m0.pdf ../paper/figures/LIGO-n-1000-8-dagh1-40m0.pdf
+cp GENOME.n.1000.8.dagh100-1500m0.pdf ../paper/figures/GENOME-n-1000-8-dagh100-1500m0.pdf
+cp SIPHT.n.1000.8.dagh5-50m0.pdf ../paper/figures/SIPHT-n-1000-8-dagh5-50m0.pdf
+
+#cp score-MONTAGE.n.1000.8.dagh1-20m0.pdf ../paper/figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf
+#cp score-CYBERSHAKE.n.1000.8.dagh1-20m0.pdf ../paper/figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf
+#cp score-LIGO.n.1000.8.dagh1-40m0.pdf ../paper/figures/score-LIGO-n-1000-8-dagh1-40m0.pdf
+#cp score-GENOME.n.1000.8.dagh100-1500m0.pdf ../paper/figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf
+#cp score-SIPHT.n.1000.8.dagh5-50m0.pdf ../paper/figures/score-SIPHT-n-1000-8-dagh5-50m0.pdf
+
+cp score-MONTAGE.n.1000.8.dagh1-20m0.pdf ../paper/figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf
+cp score-CYBERSHAKE.n.1000.8.dagh1-20m0.pdf ../paper/figures/score2-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf
+cp score-LIGO.n.1000.8.dagh1-40m0.pdf ../paper/figures/score2-LIGO-n-1000-8-dagh1-40m0.pdf
+cp score-GENOME.n.1000.8.dagh100-1500m0.pdf ../paper/figures/score2-GENOME-n-1000-8-dagh100-1500m0.pdf
[…]

### 2012-01-16 src/cws/core/experiment/Experiment.java
store scores for experiment result analysis

@@ -3,6 +3,7 @@ package cws.core.experiment;
 import java.io.File;
 import java.util.ArrayList;
 import java.util.HashSet;
+import java.util.LinkedList;
 import java.util.List;
 
 import org.cloudbus.cloudsim.Log;
@@ -113,14 +114,36 @@ public class Experiment {
 		result.setNumBusyVMs(engine.getBusyVMs().size());
 		result.setNumFreeVMs(engine.getFreeVMs().size());
 		
+		List<Integer> priorities = new LinkedList<Integer>();
+		List<Double> sizes = new LinkedList<Double>();
+
+		
 		int finished = 0;
 		for (DAGJob dj : engine.getAllDags()) {
-			if (dj.isFinished()) finished++;
+			if (dj.isFinished()) {
+				finished++;
+				priorities.add(dj.getPriority());
+				sizes.add(sumRuntime(dj.getDAG()));
+			}
 		}
 		result.setNumFinishedDAGs(finished);
+		result.setPriorities(priorities);
+		result.setSizes(sizes);
 		
 		return result;
 	}
+	
+	
+	/**
+	 * @return The total runtime of all tasks in the workflow
+	 */
+	public double sumRuntime(DAG dag) {
+        double sum = 0.0;
+        for (String taskName : dag.getTasks()) {
+            sum += dag.getTask(taskName).size;
+        }
+        return sum;
+    }
 	
 	
 	/**
@@ -187,6 +210,8 @@ public class Experiment {
 			        deadline, budget, price, max_scaling));
 		}
 		
+		// write number of dags finished
+		
 		StringBuffer outSPSS = new StringBuffer();
 		StringBuffer outAware = new StringBuffer();
 		StringBuffer outSimple = new StringBuffer();
@@ -201,6 +226,39 @@ public class Experiment {
 		WorkflowLog.stringToFile(outSPSS.toString(), dags[0] + "b" + budget + "h" +start + "-" + N + "m" + max_scaling + "run" + runID + "-outputSPSS.txt");
 		WorkflowLog.stringToFile(outAware.toString(), dags[0] + "b" + budget + "h" +start + "-" + N + "m" + max_scaling + "run" + runID + "-outputAware.txt");
[…]

### 2012-01-16 src/cws/core/experiment/ExperimentResult.java

@@ -1,5 +1,7 @@
 package cws.core.experiment;
 
+import java.util.List;
+
 public class ExperimentResult {
 	
 	private double cost;
@@ -9,6 +11,10 @@ public class ExperimentResult {
 	private int numTotalVMs;
 	private int numFinishedDAGs;
 	private double deadline;
+	private List<Integer> priorities;
+	private List<Double> sizes;
+	
+
 
 	public double getCost() {
 		return cost;
@@ -52,5 +58,58 @@ public class ExperimentResult {
 	public void setDeadline(double deadline) {
 		this.deadline = deadline;
 	}
+	public List<Integer> getPriorities() {
+		return priorities;
+	}
+	public void setPriorities(List<Integer> priorities) {
+		this.priorities = priorities;
+	}
+	public List<Double> getSizes() {
+		return sizes;
+	}
+	public void setSizes(List<Double> sizes) {
+		this.sizes = sizes;
+	}
+	
+	/**
+	 * Format the list of priorities of completed DAGs as a string
+	 * @return the string containing: deadline and space separated priorities
+	 */
+	
+	public String formatPriorities() {
+		
+		StringBuffer prioritiesBuffer = new StringBuffer();
+
+		prioritiesBuffer.append(getDeadline());
+		
+		for (int priority : getPriorities()) {
+			prioritiesBuffer.append(" " + priority);
+		}
+		
+		prioritiesBuffer.append("\n");
+		
+		return prioritiesBuffer.toString();
+	}
+	
+	/**
+	 * Format the list of sizes of completed DAGs as a string
+	 * @return the string containing: deadline and space separated sizes
+	 */
+	
+	public String formatSizes() {
+		
+		StringBuffer sizesBuffer = new StringBuffer();
+
+		sizesBuffer.append(getDeadline());
+		
+		for (double size : getSizes()) {
+			sizesBuffer.append(" " + size);
+		}
+		
+		sizesBuffer.append("\n");
+		
+		return sizesBuffer.toString();
+	}
+	
 
 }

### 2012-01-16 paper/paper.tex

@@
 
-\begin{figure}[htb] 
+\begin{figure*}[htb] 
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/CYBERSHAKE-Pareto}
-\caption{Comparison of DPDS and AW-DPDS algorithms for ensembles of
- 100 Pareto-distributed Cybershake workflows}
-\label{fig:CYBERSHAKE.n.1000.8.dagh1-20m0}
-\end{figure}
+\includegraphics[width=1.0\textwidth]{figures/MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
+Pareto-distributed workflows}
+\label{fig:}
+\end{figure*}
+
+\begin{figure*}[htb] 
+\centering
+\includegraphics[width=1.0\textwidth]{figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/score-LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/score-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
+Pareto-distributed workflows. Score is computed as $\sum(1/(p+1))$ where
+priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
+priority).}
+\label{fig:}
+\end{figure*}
+
+\begin{figure*}[htb] 
+\centering
+\includegraphics[width=1.0\textwidth]{figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/score2-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/score2-LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/score2-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/score2-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
+Pareto-distributed workflows. Score is computed as $\sum(2^(-p))$ where
+priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
+priority).}
+\label{fig:}
+\end{figure*}
+
+\begin{figure*}[htb] 
+\centering
+\includegraphics[width=1.0\textwidth]{figures/size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/size-LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/size-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/size-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
+Pareto-distributed workflows. Score is equal to the total work completed,
+i.e. it is sum of runtimes (in hours) of tasks of completed workflows.}
+\label{fig:}
+\end{figure*}
+
+\begin{figure*}[htb] 
+\centering
+\includegraphics[width=1.0\textwidth]{figures/cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/cost-LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=1.0\textwidth]{figures/cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
+Pareto-distributed workflows. Plot shows effective computing cost in \$ per
+hour calclated by dividing sum of task runtimes by the total budget. Higher
+cost results from lower resource utilization.}
+\label{fig:}
+\end{figure*}
 

### 2012-01-17 paper/paper.tex

@@
 \subsection{Static Provisioning Dynamic Scheduling (SPDS)}
-
+\label{sec:spds}
 This is the simplest strategy based on static provisioning of resources and
@@
 together with the hourly price $p$ of a VM it is possible to compute the number
-of VMs $N_{VM}$ to provision so that they consume the whole budget between the
+of VMs $N_{VM}$ to provision so that they consume the whole budget before the
 deadline.
@@
 Once the resources are provisioned, the tasks are scheduled to VMs using dynamic
-priority-based algorithm shown in Algorithm~\ref{alg:ds}.
+priority-based Algorithm~\ref{alg:ds}.
 
-% This is an example of how we can do algorithms
 \begin{algorithm}
@@
 become ready. Fig.~\ref{fig:spss-example} shows the example plan generated using
-SPSS algorithm.
+SPSS algorithm. It can be seen that tasks from lower priority workflows can fill
+idle time of provisioned virtual machines.
 
@@
 \subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
+
+Static provisioning plan used in SPDS algorithm allows no changes to resource
+pool at runtime. However, such plan may not be optimal when resource demand
+changes during execution of workflows. In policy-based provisioning
+(or auto-scaling) systems the metrics such as resource
+utilization or queue length are typicaly used to estimate curret resource
+demand. Our DPDS algorithm is based on resource utilization: the provisioner
+module periodically computes utilization as the percentage of idle VMs. The
+provisioner algorithm detects which VMs are finishing their hourly billing cycle
+during the next provisioning period and decides whether they should continue
+running, should be terminated or new VMs should be provisioned: see
+Algorithm~\ref{alg:prov}.
+
+\begin{algorithm}
+\caption{Dynamic provisioner}
+\label{alg:prov}
+\begin{algorithmic}[1]
+\Require $c$: consumed budget; $b$: total budget; $d$: deadline; $p$: price;
+$t$: current time; $u_h$: upper utilization thershold; $u_l$: lower utilization
+threshold.
+\Procedure{provision}{}
+    \State $V_C\gets set\ of\ VMs\ completing\ billing\ cycle$
+    \State $n_T\gets\ 0$ \Comment{number of VMs to terminate} 
+    \If{$ b - c < p * |V_C| \vee t > d $ }
+    	\State $V_R\gets set\ of\ running\ VMs$
+    	\State $n_T\gets |V_R| - \lfloor(b-c)/p\rfloor$ 
+    	\State $V_T\gets select\ n_T VMs\ to\ terminate\ from\ V_C$
+    	\State \Call{Terminate}{$V_T$}
+    \Else 
+    	\If{$u>u_h$}
+    		\State \Call{Start}{$new\ VM$}
+    	\EndIf
+    	\If{$u<u_l$}
+    		\State $V_F\gets free\ (idle)\ VMs$
+    		\State $n_T\gets \lceil|V_F|/2\rceil$ 
+    		\State \Call{Terminate}{$V_T$}
+    	\EndIf 
+    \EndIf
+\EndProcedure
+\end{algorithmic} 
+\end{algorithm}
+
+The number of instances to terminate $n_T$ computed in line 6 is the number that
+would overrun the budget if not terminated. On the other hand $n_T$ computed in
+line 15 prevents terminating all the instances too quickly, but makes sure that
+if there is only one instance left it should be terminated. However, the
+termination process is not immediate: to make sure that VMs that have been
+already paid for are not terminated prematurely the {\em Terminate()} procedure
+terminates only these instances that are close to the end of their billing
+cycle. The remaining time threshold is calculated as a sum of provisioner
+interval and estimated termination delay. This guarantees that no VMs can
+overrun the assigned budget. To avoid uncontrolled increase of number of
+instances, which may happen in the case of highly parallel workflows,
+provisioner checks also the maximum scaling factor from the initial number of
+provisioned VMs and prohibits starting new ones when this threshold is reached.
 

### 2012-01-17 paper/paper.tex
added result discussion

@@
 \usepackage{algorithm,algpseudocode}
+\usepackage{url}
 
@@
 \section{Related Work}
-General policy and rule-based approaches to dynamic provisioning (e.g. Amazon Auto Scaling\footnote{http://aws.amazon.com/autoscaling} and RightScale\footnote{http://www.rightscale.com}).
+General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
+Auto Scaling\footnote{\url{http://aws.amazon.com/autoscaling}} and
+RightScale\footnote{\url{http://www.rightscale.com}}).
 
@@
 
-\subsection{Static Provisioning Dynamic Scheduling (SPDS)}
+\subsection{Static Provisioning Dynamic Scheduling\\
+(SPDS) }
 \label{sec:spds}
@@
 become ready. Fig.~\ref{fig:spss-example} shows the example plan generated using
-SPSS algorithm. It can be seen that tasks from lower priority workflows can fill
+SPDS algorithm. It can be seen that tasks from lower priority workflows can fill
 idle time of provisioned virtual machines.
@@
 \includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
- \caption{Example of schedule generated using SPSS algorithm: tasks are labelled
+ \caption{Example of schedule generated using SPDS algorithm: tasks are labelled
  by colors depending on the workflow. }
@@
 \subsection{Scenarios}
+\label{sec:scenarios}
 
-\subsection{Results}
+We have prepared test scenarios to evaluate the proposed algorithms under
+such a range of parameters that allows to observe the interesting properties of
+algorithms for smaller and larger constraints. For a given ensemble we can
+expect in general that the number of work done (in terms of workflows completed)
+mainly depends on the budget constraint, which determines the number of
+machine-hours that can be provisioned. The ideal but trivial case is then a
+deadline long enough to allow running all the workflow tasks sequentially: this
+eliminates all the overheads of parallel execution and leads to the highest
+resource utilization. More interesting are the cases when the deadline is
+shorter, since it forces provisioning of multiple VMs and thus increases
+overheads due to parallel execution so the budget cannot be spent so
+efficiently.
+
+We assumed that the interesting number of workflows in ensemble should be in the
+order of between 10 and 100 workflows. This is motivated by several reasons:
+first, such numbers are typical for the real applications, e.g. the number of geographical sites for
+CyberShake workflow is in the order of 100. Second, the smaller ensembles
+consisting of just a few workflows can be aggregated in a single workflow so
+there is no need to treat them as ensembles. Similarly, when the number of
+workflows is larger that 100 and each workflow has a large number of parallel
+tasks, the problem of efficiently allocating them to the resources becomes much
+easier and similar to the bag-of-tasks problem.
+
+\begin{figure}[htb] 
+\centering
+\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
+\caption{Pareto-like distribution of workflow sizes in ensemble. }
+\label{fig:ensemble-distribution}
+\end{figure}
+
+In order to evaluate the algorithms on a standard ser of workflows, we used
+ensembles constructed from the DAGs available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
+in the workflow generator gallery~\cite{Bharathi08}. The gallery contains
+workflows of five applications from Pegasus and from Indiana
+University~\cite{Ramakrishnan08}. We constructed ensembles of 100 DAGs by sampling the
+workflows of different sizes based on a given distribution. Two distributions
+are considered:
+\begin{itemize}
+  \item ensemble of workflows of equal size (constant)
+  \item Pareto-like distribution.
+\end{itemize}
+
+Pareto-like distribution gives a small number of workflows with large size and a
+large number of workflows with small size, as seen in
+Fig.~\ref{fig:ensemble-distribution}. The number of largest
+workflows (of size $\geq$ 900) is slightly increased to represent the
+``long-tail'', i.e. that there is a non-insignificant number of large workflows
+in the ensemble. The workflows are sorted according to their size and the largest workflows are
+given the highest priority.
+
+In the tests we we assumed that all the VMs are homogeneous and have the
+processing speed of 1000 MIPS (million instructions per second), the price for
+one VM-hour is \$1 and that the sizes of tasks in the workflow gallery are given
+in seconds measured on 1000 MIPS machines. In this study we do not take into
+account the heterogeneity of the infrastructure since we assume that it is
+always possible to select a VM type that has the best price to performance ratio
+for a given application.
+
+For each application type, we selected such ranges of constraints (deadline,
+budget) that it is possible to observe the possibly broad result space: from the
+tight constraints, when only a small number of workflows can finish to the
+liberal constraints when all or almost all the workflows can finish. These
+parameters vary among the applications. Montage and CyberShake benchmarks
+workflows required the budget in \$20 -- \$150 range and deadlines from 1 to
+20 hours to complete the ensemble. LIGO and SIPHT workflows have more long
+tasks, so they require budgets up to \$1,500 and deadlines extended to 40
+hours. Epigenomics workflow is even more costly and needs approximately \$12,000
+and 1500 hours to complete the full ensemble.
+
+\subsection{Metrics}
+
+There are multiple ways to evaluate the algorithms and also different metrics
+are relevant to the application users that are needed to take the planning
+decisions. The set of metrics for each ensemble run we prepared includes:
+\begin{itemize}
+  \item number of workflows completed $N_c$,
+  \item user preference score (utility function) $s$,
+  \item total amount of computing time completed $T_C$,
+  \item effective cost of computing per hour $C_{eff}$, which includes
+  overheads.
+\end{itemize}
+
+\begin{figure*}[t] 
+\centering
+\includegraphics[width=0.19\textwidth]{figures/MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
+algorithms on ensembles of 100 Pareto-distributed workflows}
+\label{fig:number-complete-pareto}
+\end{figure*}
+
+The number of workflows completed is a simplest metric and allows distinguishing
+which algorithm performs better. However, as the workflows have
+different priorities and sizes, a less efficient algorithm may schedule smaller
+low-priority workflows first, thus increasing the total number of workflows
+complete. It may also happen that too conservative admission algoritm may reject
+overestimate the costs of some large workflows, thus rejest some high-priority
+ones which actually could fit into the schedule. Therefore we defined a user
+preference score as:
+
+\begin{equation}
+\label{eq:score}
+s = \sum_{DAGs\ completed}(1/(p+1))
+\end{equation}
+
+where priorities of subsequent workflows in the ensemble are $p=0,1,2,\ldots$ (0
+means highest priority). This scoring function gives more weight to the top high
+priority workflows, whereas the low priority ones are trated almost equally.
+
+Another metric is the total actual computing time completed and it is calculated
+as the sum of all the runtimes of all the tasks of completed workflows. As in
+our study we assumed that the resource cost pre hour is equal to \$1, the total
+computing time can be easily compared to the allocated budget $b$. Their ratio
+$u = (T_C * \$1)/b$ is the resource utilization and its inverse is the effective
+resource cost: $C_{eff} = b/T_C$ in dollars per hour.
+
+\subsection{Discussion of Results}
+
+\begin{figure}[b!] 
+\centering
+%\includegraphics[width=0.19\textwidth]{figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+%\includegraphics[width=0.19\textwidth]{figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/score-LIGO-n-1000-8-dagh1-40m0.pdf}
+%\includegraphics[width=0.19\textwidth]{figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/score-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Comparison of algorithms based on score, which  is computed as
+$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
+priority).}
+\label{fig:score}
+\end{figure}
+
+We run the simulations for the constraint values ranging between the
+estimated limits as indicated in Section~\ref{sec:scenarios}, resulting in
+ca. 50-150 (budget, deadline) points for each application. The runs were
+repeated for Pareto-distributed and costant workflow sizes. From this result
+space we selected the most interesting plots which give the good overview  of
+the observed algorithm behavior.
 
@@
 \centering
-\includegraphics[width=1.0\textwidth]{figures/MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
-Pareto-distributed workflows}
-\label{fig:}
-\end{figure*}
-
-\begin{figure*}[htb] 
-\centering
-\includegraphics[width=1.0\textwidth]{figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/score-LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/score-SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
-Pareto-distributed workflows. Score is computed as $\sum(1/(p+1))$ where
-priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
-priority).}
-\label{fig:}
-\end{figure*}
-
-\begin{figure*}[htb] 
-\centering
-\includegraphics[width=1.0\textwidth]{figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/score2-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/score2-LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/score2-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/score2-SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
-Pareto-distributed workflows. Score is computed as $\sum(2^(-p))$ where
-priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
-priority).}
-\label{fig:}
-\end{figure*}
-
-\begin{figure*}[htb] 
-\centering
-\includegraphics[width=1.0\textwidth]{figures/size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/size-LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/size-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/size-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/size-LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/size-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/size-SIPHT-n-1000-8-dagh5-50m0.pdf}
 \caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
@@
 i.e. it is sum of runtimes (in hours) of tasks of completed workflows.}
-\label{fig:}
+\label{fig:total-time}
 \end{figure*}
 
-\begin{figure*}[htb] 
+\begin{figure*}[b!] 
 \centering
-\includegraphics[width=1.0\textwidth]{figures/cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/cost-LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=1.0\textwidth]{figures/cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
-Pareto-distributed workflows. Plot shows effective computing cost in \$ per
-hour calclated by dividing sum of task runtimes by the total budget. Higher
-cost results from lower resource utilization.}
+\includegraphics[width=0.19\textwidth]{figures/cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/cost-LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Effective computing cost in \$ per hour calclated by dividing sum of
+task runtimes by the total budget. Higher cost results from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
+Pareto-distributed workflows.}
+\label{fig:cost}
+\end{figure*}
+
+Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
+completed for two budget values per each application ensemble. The budget values
+selected correspond to two opposite cases: {\em small budget} (top) and {\em
+large budget} (bottom). The first one allows completing only a small number of
+workflows from the ensemble, while the latter one suffices to finish almost all
+the workflows. It can be seen that the static planning algorithm (SPSS) performs
+better for tighter constraints: when the budget is low to admit only a small
+number of workflows, the static procedure is able to more efficiently pack these
+workflow tasks onto the resources. This can be clearly seen for the LIGO,
+Epigenomics and SIPHT ensembles for small budgets and also for the same
+ensembles with large budgets but short deadlines. The exception are Montage and
+CyberShake ensembles, which consist of many fine-grained tasks that can be
+executed in parallel. For these cases, as well for the other ensembles with
+large budgets, the simple dynamic scheduling techniqies perform equally well or
+better. The reason for that behavior is that the large number of tasks
+guarantees that resources are almost never idle, while the provisioner algorithm
+makes sure that the VM instances are most efficiently utilized before shutdown.
+We can also observe that the workflow-aware dynamic algorithm (WA-DPDS) performs
+better that the workflow-unaware version. This means that for the tested
+workflows the simple admission procedure based on estimation of workflow cost
+does not degrade the solution, but rather in many cases it allows rejecting
+larger workflows that would lead to budget overrun. Thus it can save the space
+for smaller workflows that can complete.
+
+Fig.~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
+computed for two sample applications. Comparing
+Fig.~\ref{fig:number-complete-pareto} to Fig.~\ref{fig:score} we can see that
+the latter one gives more flattened shape: this means that e.g. the improvement
+of the solution by one workflow may mean adding a one low-priority workflow.
+Nevertheless we can see that the curves on both figures have a very similar
+shape, which mens that our algorithms obey the priority rules and do not try to
+admit lower-priority workflows at the cost of the more important ones.
+
+Fig.~\ref{fig:total-time} presents the same results using the total computing
+time metric $T_C$. These plots confirm our previous observations about the
+algorithms performance, but they also allow to observe a ore general property.
+For a given budget, when the deadline is tightened the number of work done
+steeply decreases. Assiging a shorter deadline means that more VMs need to be
+provisioned to complete the work, but this leads to higher parallelism and thus
+to the lower parallel efficiency. E.g. running 2 VMs for 10 hours will be more
+efficient than running 10 VMs for 2 hours. Therefore, if deadline permits and
+costs are important, parallelism should be avoided when possible.
+
+These observations are further confirmed when examining the efffective resource
+cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
+utilization and lower parallel efficiency for short deadlines lead to
+substantial increase of actual costs of computing. This increase reaches ca.
+20\%, which means that if we need to run e.g. a SIPHT workflow ensemble with a
+short deadline, ca. \$200 out of \$1000 is spent on idle resources.
+
+Another general concluson can be drawn from Fig.~\ref{fig:cost}: despite some
+random fluctuations resulting mostly from the finite result space, all the
+curves for all the applications and budgets have a similar shape. This shape
+represents a trade-off curve between two conflicting objectives: time and cost.
+We can see that the selection of deadline and cost can be formulated as
+multi-criteria optimization problem and the curves presented in
+Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
+solutions to this problem. In our case the solutions were achieved by maximizing
+the number of work done, whcich direcly results in minimization of effective cost parameter for a given
+deadline. Therefore, the obtained results can be used to assist the decison in
+resource allocation planning, when both cost and time criteria are given not as
+constraints but as objectives.
+
+% \begin{figure*}[htb] 
+% \centering
+% \includegraphics[width=0.19\textwidth]{figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+% \includegraphics[width=0.19\textwidth]{figures/score2-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+% \includegraphics[width=0.19\textwidth]{figures/score2-LIGO-n-1000-8-dagh1-40m0.pdf}
+% \includegraphics[width=0.19\textwidth]{figures/score2-GENOME-n-1000-8-dagh100-1500m0.pdf}
+% \includegraphics[width=0.19\textwidth]{figures/score2-SIPHT-n-1000-8-dagh5-50m0.pdf}
+% \caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
+% Pareto-distributed workflows. Score is computed as $\sum(2^(-p))$ where
+% priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
+% priority).}
+% \label{fig:}
+% \end{figure*}
+
+\begin{figure}[htb] 
+\centering
+\includegraphics[width=0.6\columnwidth]{figures/budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}
+\caption{Comparison of number of workflows completed for a given deadline,
+plotted versus different budget amounts.}
 \label{fig:}
-\end{figure*}
+\end{figure}
 

### 2012-01-17 scripts/copy_plots_to_paper.sh
scripts to produce nicer plots for the paper

@@ -1,9 +1,9 @@
 #!/bin/bash
-cp MONTAGE.n.1000.8.dagh1-20m0.pdf ../paper/figures/MONTAGE-n-1000-8-dagh1-20m0.pdf
-cp CYBERSHAKE.n.1000.8.dagh1-20m0.pdf ../paper/figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf
-cp LIGO.n.1000.8.dagh1-40m0.pdf ../paper/figures/LIGO-n-1000-8-dagh1-40m0.pdf
-cp GENOME.n.1000.8.dagh100-1500m0.pdf ../paper/figures/GENOME-n-1000-8-dagh100-1500m0.pdf
-cp SIPHT.n.1000.8.dagh5-50m0.pdf ../paper/figures/SIPHT-n-1000-8-dagh5-50m0.pdf
+#cp MONTAGE.n.1000.8.dagh1-20m0.pdf ../paper/figures/MONTAGE-n-1000-8-dagh1-20m0.pdf
+#cp CYBERSHAKE.n.1000.8.dagh1-20m0.pdf ../paper/figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf
+#cp LIGO.n.1000.8.dagh1-40m0.pdf ../paper/figures/LIGO-n-1000-8-dagh1-40m0.pdf
+#cp GENOME.n.1000.8.dagh100-1500m0.pdf ../paper/figures/GENOME-n-1000-8-dagh100-1500m0.pdf
+#cp SIPHT.n.1000.8.dagh5-50m0.pdf ../paper/figures/SIPHT-n-1000-8-dagh5-50m0.pdf
 
 #cp score-MONTAGE.n.1000.8.dagh1-20m0.pdf ../paper/figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf
 #cp score-CYBERSHAKE.n.1000.8.dagh1-20m0.pdf ../paper/figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf
@@ -11,11 +11,11 @@ cp SIPHT.n.1000.8.dagh5-50m0.pdf ../paper/figures/SIPHT-n-1000-8-dagh5-50m0.pdf
 #cp score-GENOME.n.1000.8.dagh100-1500m0.pdf ../paper/figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf
[…]

### 2012-01-17 scripts/ruby/plot_schedule_gantt.rb

@@ -144,3 +144,4 @@ require "#{File.dirname(__FILE__)}/plots.rb"
 plot_schedule 'output/testSPSSSPSSGENOME.n.1000.8.dagx100d3600000.0b4000.0m0.0'
 plot_schedule 'output/testSimpleUtilizationBasedProvisionerWorkflowAwareEnsembleSchedulerGENOME.n.1000.8.dagx100d3600000.0b4000.0m0.0'
 #plot_schedule 'output/testSimpleUtilizationBasedProvisionerEnsembleDynamicSchedulerGENOME.n.1000.8.dagx100d3600000.0b4000.0m0.0'
+plot_schedule 'output/testSimpleUtilizationBasedProvisionerEnsembleDynamicSchedulerSIPHT.n.1000.0.dagx100d108000.0b200.0m0.0'

### 2012-01-17 src/cws/core/experiment/DAGListGenerator.java
tests for constant workflow sizes

@@ -40,6 +40,16 @@ public class DAGListGenerator {
 		}
 		return dags;
 	}
+
+	
+	public static String[] generateDAGListConstant(String name, int size, int length) {
+		
+		String dags[] = new String[length];
+		for (int i=0; i<length; i++) {
+				dags[i] = name + ".n." + size + "." + i%20 + ".dag";
+		}
+		return dags;
+	}
 	
 	public static String[] generateDAGListPareto(Random seed, String name, int length) {
 		ArrayList<String> dags = new ArrayList<String>(length);

### 2012-01-17 src/cws/core/experiment/DAGListGeneratorTest.java

@@ -33,4 +33,12 @@ public class DAGListGeneratorTest {
 		assertEquals("SIPHT.n.50.0.dag", dags[39]);
 		assertEquals("SIPHT.n.300.1.dag", dags[7]);	
 	}
+	
+	@Test
+	public void testConstant40() {
+		String[] dags = DAGListGenerator.generateDAGListConstant("SIPHT", 1000, 40);
+		assertEquals("SIPHT.n.1000.0.dag", dags[0]);
+		assertEquals("SIPHT.n.1000.19.dag", dags[39]);
+		assertEquals("SIPHT.n.1000.7.dag", dags[7]);	
+	}
 }

### 2012-01-17 src/cws/core/experiment/Experiment.java

@@ -165,6 +165,7 @@ public class Experiment {
 	 * Uses two schedulers: workflow aware and unaware.
 	 * For each scheduler a text file is produced, containing number of workflows finished for a given deadline.
 	 * 
+	 * @param prefix prefix to append to generated output files
 	 * @param dagPath path to dags
 	 * @param dags array of file names
 	 * @param budget budget in $
@@ -176,7 +177,7 @@ public class Experiment {
 	 * @param runID id of this series
 	 */
 	
-	public static void runSeries(String dagPath, String[] dags, double budget, double price,
+	public static void runSeries(String prefix, String dagPath, String[] dags, double budget, double price,
 			int N, int step, int start, double max_scaling, int runID) {
 		
 		double deadline;
@@ -223,9 +224,9 @@ public class Experiment {
 			
 		}
 
-		WorkflowLog.stringToFile(outSPSS.toString(), dags[0] + "b" + budget + "h" +start + "-" + N + "m" + max_scaling + "run" + runID + "-outputSPSS.txt");
-		WorkflowLog.stringToFile(outAware.toString(), dags[0] + "b" + budget + "h" +start + "-" + N + "m" + max_scaling + "run" + runID + "-outputAware.txt");
-		WorkflowLog.stringToFile(outSimple.toString(), dags[0] + "b" + budget + "h" +start + "-" + N + "m" + max_scaling + "run" + runID + "-outputSimple.txt");			
+		WorkflowLog.stringToFile(outSPSS.toString(), prefix + dags[0] + "b" + budget + "h" +start + "-" + N + "m" + max_scaling + "run" + runID + "-outputSPSS.txt");
+		WorkflowLog.stringToFile(outAware.toString(), prefix + dags[0] + "b" + budget + "h" +start + "-" + N + "m" + max_scaling + "run" + runID + "-outputAware.txt");
+		WorkflowLog.stringToFile(outSimple.toString(), prefix + dags[0] + "b" + budget + "h" +start + "-" + N + "m" + max_scaling + "run" + runID + "-outputSimple.txt");			
 
 		// write priorities of dags finished
 		
@@ -239,9 +240,9 @@ public class Experiment {
 			prioritiesSimple.append(resultsSimple[i].formatPriorities());			
 		}
 
[…]

### 2012-01-17 src/cws/core/experiment/ExperimentConstantTestRepeat.java

@@ -17,7 +17,7 @@ import org.junit.Test;
  *
  */
 
-public class ExperimentParetoTestRepeat {
+public class ExperimentConstantTestRepeat {
     
 	String dagPath;
 	String dagName;
@@ -25,7 +25,7 @@ public class ExperimentParetoTestRepeat {
 	double price = 1.0;
 	int numDAGs = 40;
 	double max_scaling = 2.0;
-	String prefix = "pareto-";
+	String prefix = "constant-";
 
 	
 	int N;
@@ -44,14 +44,14 @@ public class ExperimentParetoTestRepeat {
 		dagPath = "../projects/pegasus/Montage/";
 		dagName = "MONTAGE";
 		
-		String[] dags = DAGListGenerator.generateDAGListPareto(new Random(0), dagName, 100);
+		String[] dags = DAGListGenerator.generateDAGListConstant(dagName, 1000, 100);
 		N = 20;
 		step = 1;
 		start = 1;
 		max_scaling = 0;
 		
 //		double[] budgets= {10.0, 20.0, 30.0, 40.0, 50.0, 60.0, 70.0, 80.0, 90.0, 100.0};
-		double[] budgets= {20.0, 30.0, 50.0, 60.0, 80.0};
+		double[] budgets= {40.0, 80.0, 120.0, 160.0, 200.0};
 		
 		for (double budget : budgets) {
 			Experiment.runSeriesRepeat(prefix, dagPath, dags, budget, price, N, step, start, max_scaling, 1);			
@@ -65,7 +65,7 @@ public class ExperimentParetoTestRepeat {
 		dagPath = "../projects/pegasus/CyberShake/";
 		dagName = "CYBERSHAKE";
 		
-		String[] dags = DAGListGenerator.generateDAGListPareto(new Random(0), dagName, 100);
+		String[] dags = DAGListGenerator.generateDAGListConstant(dagName, 1000, 100);
 		
 		N = 20;
 		step = 1;
@@ -74,7 +74,7 @@ public class ExperimentParetoTestRepeat {
 
 		
 //		double[] budgets= {10.0, 20.0, 30.0, 40.0, 50.0, 60.0, 80.0, 100.0, 120.0, 140.0};
-		double[] budgets= {30.0, 50.0, 80.0, 100.0, 140.0};
+		double[] budgets= {50.0, 150.0, 250.0, 350.0, 450.0};
 		
 		for (double budget : budgets) {
[…]

### 2012-01-17 src/cws/core/experiment/ExperimentTest900_1000.java

@@ -23,6 +23,7 @@ public class ExperimentTest900_1000 {
 	double price = 1.0;
 	int numDAGs = 40;
 	double max_scaling = 2.0;
+	String prefix = "test900-1000-";
 
 	
 	int N;
@@ -52,7 +53,7 @@ public class ExperimentTest900_1000 {
 		double[] budgets= {40.0, 80.0, 120.0, 160.0, 200.0};
 		
 		for (double budget : budgets) {
-			Experiment.runSeries(dagPath, dags, budget, price, N, step, start, max_scaling);			
+			Experiment.runSeries(prefix, dagPath, dags, budget, price, N, step, start, max_scaling);			
 		}		
 		
 	}
@@ -74,7 +75,7 @@ public class ExperimentTest900_1000 {
 		double[] budgets= {40.0, 80.0, 120.0, 160.0, 200.0};
 		
 		for (double budget : budgets) {
-			Experiment.runSeries(dagPath, dags, budget, price, N, step, start, max_scaling);			
+			Experiment.runSeries(prefix, dagPath, dags, budget, price, N, step, start, max_scaling);			
 		}
 		
 	}
@@ -96,7 +97,7 @@ public class ExperimentTest900_1000 {
 		double[] budgets= {400.0, 800.0, 1200.0, 1600.0, 2000.0};
 
 		for (double budget : budgets) {
-			Experiment.runSeries(dagPath, dags, budget, price, N, step, start, max_scaling);			
+			Experiment.runSeries(prefix, dagPath, dags, budget, price, N, step, start, max_scaling);			
 		}
 		
 	}
@@ -119,7 +120,7 @@ public class ExperimentTest900_1000 {
 		double[] budgets= {4000.0, 8000.0, 12000.0, 16000.0, 20000.0};
 		
 		for (double budget : budgets) {
-			Experiment.runSeries(dagPath, dags, budget, price, N, step, start, max_scaling);			
+			Experiment.runSeries(prefix, dagPath, dags, budget, price, N, step, start, max_scaling);			
 		}
 			
 	}
@@ -141,7 +142,7 @@ public class ExperimentTest900_1000 {
 		double[] budgets= {200.0, 400.0, 600.0, 800.0, 1000.0};
 
 		for (double budget : budgets) {
-			Experiment.runSeries(dagPath, dags, budget, price, N, step, start, max_scaling);			
[…]

### 2012-01-18 src/cws/core/experiment/Experiment.java
Move SPSS

@@ -13,16 +13,15 @@ import cws.core.Cloud;
 import cws.core.DAGJob;
 import cws.core.EnsembleManager;
 import cws.core.Provisioner;
-import cws.core.SPSS;
 import cws.core.Scheduler;
 import cws.core.SimpleJobFactory;
 import cws.core.VM;
 import cws.core.WorkflowEngine;
 import cws.core.WorkflowEvent;
+import cws.core.algorithms.SPSS;
 import cws.core.dag.DAG;
 import cws.core.dag.DAGParser;
 import cws.core.log.WorkflowLog;
-import cws.core.provisioner.AbstractProvisioner;
 import cws.core.provisioner.SimpleUtilizationBasedProvisioner;
 import cws.core.scheduler.EnsembleDynamicScheduler;
 import cws.core.scheduler.WorkflowAwareEnsembleScheduler;

### 2012-01-18 src/cws/core/experiment/DAGListGeneratorTest.java
Remove unused imports

@@ -2,10 +2,8 @@ package cws.core.experiment;
 
 import static org.junit.Assert.*;
 
-import java.util.HashMap;
 import java.util.Random;
 
-import org.cloudbus.cloudsim.distributions.ParetoDistr;
 import org.junit.Test;
 
 public class DAGListGeneratorTest {

### 2012-01-18 scripts/ruby/plot_gantt_chart.rb
Add new algorithms, fix SPSS bug

@@ -1,0 +1,5 @@
+require "#{File.dirname(__FILE__)}/plots.rb"
+
+ARGV.each do|a|
+  plot_schedule(a)
+end

### 2012-01-18 scripts/ruby/plots.rb

@@ -96,9 +96,9 @@ def plot_schedule (filename)
       #puts "Saving plot to file: " + filename + ".pdf"
       plot.set "key right outside"
       plot.set "grid"
-      #plot.terminal "png size 1024,768"
+      plot.terminal "png size 1024,768"
       #plot.output 'output/' + dag + deadline + ".png"
-      #plot.output filename + ".png"
+      plot.output filename + ".png"
 
       data = Array.new
 

### 2012-01-18 src/cws/core/SPSS.java

@@ -1,891 +1,0 @@
-package cws.core;
-
-import java.io.File;
-import java.util.ArrayList;
-import java.util.Collections;
-import java.util.Comparator;
-import java.util.HashSet;
-import java.util.LinkedHashSet;
-import java.util.LinkedList;
-import java.util.List;
-import java.util.HashMap;
-import java.util.Queue;
-import java.util.Random;
-import java.util.SortedSet;
-import java.util.TreeMap;
-
-import org.cloudbus.cloudsim.core.CloudSim;
-
-import cws.core.Job.Result;
-import cws.core.dag.DAG;
-import cws.core.dag.Task;
-import cws.core.dag.DAGParser;
-import cws.core.dag.algorithms.CriticalPath;
-import cws.core.dag.algorithms.TopologicalOrder;
-import cws.core.experiment.DAGListGenerator;
-import cws.core.log.WorkflowLog;
-
-/**
- * @author Gideon Juve <juve@usc.edu>
- */
-public class SPSS implements WorkflowEvent, Provisioner, Scheduler, VMListener, JobListener, DAGJobListener {
-    
-    private WorkflowEngine engine;
-    private EnsembleManager manager;
-    private Cloud cloud;
-    private double budget;
-    private double ensembleDeadline;
-    
-    private Plan plan = new Plan();
-    
-    /** List of all dags in ensemble */
-    private List<DAG> allDAGs;
-    
-    /** List of DAGs that were admitted to run */
-    private List<DAG> admittedDAGs = new LinkedList<DAG>();
-    
-    /** Set of jobs that are ready to run arranged by Task */
-    private HashMap<Task, Job> readyJobs = new HashMap<Task, Job>();
-    
-    /** Mapping of Task to the VM that will run the task */
-    private HashMap<Task, VM> taskMap = new HashMap<Task, VM>();
-    
-    /** Schedule of tasks for each VM */
-    private HashMap<VM, Queue<Task>> vmQueues = new HashMap<VM, Queue<Task>>();
-    
-    /** Set of idle VMs */
-    private HashSet<VM> idle = new HashSet<VM>();
-    
-    /** Tuning parameter for deadline distribution (low alpha = runtime, high alpha = tasks) */
-    private double alpha;
-    
-    public SPSS(double budget, double deadline, List<DAG> dags, double alpha) {
-        this.budget = budget;
-        this.ensembleDeadline = deadline;
-        this.allDAGs = dags;
-        this.alpha = alpha;
-    }
-    
[…]

### 2012-01-18 src/cws/core/algorithms/Algorithm.java

@@ -1,0 +1,89 @@
+package cws.core.algorithms;
+
+import java.util.LinkedList;
+import java.util.List;
+
+import cws.core.dag.DAG;
+
+public abstract class Algorithm {
+    private double budget;
+    private double deadline;
+    private List<DAG> dags;
+    
+    public Algorithm(double budget, double deadline, List<DAG> dags) {
+        this.budget = budget;
+        this.deadline = deadline;
+        this.dags = dags;
+    }
+    
+    public List<DAG> getDAGs() {
+        return dags;
+    }
+    
+    public double getBudget() {
+        return budget;
+    }
+    
+    public double getDeadline() {
+        return deadline;
+    }
+    
+    public String getName() {
+        return this.getClass().getSimpleName();
+    }
+    
+    abstract public void simulate(String logname);
+    
+    abstract public double getActualCost();
+    
+    abstract public double getActualFinishTime();
+    
+    abstract public List<DAG> getCompletedDAGs();
+    
+    public int numCompletedDAGs() {
+        return getCompletedDAGs().size();
+    }
+    
+    public List<Integer> completedDAGPriorities() {
+        List<DAG> dags = getDAGs();
+        List<Integer> priorities = new LinkedList<Integer>();
+        for (DAG dag : getCompletedDAGs()) {
+            int index = dags.indexOf(dag);
+            int priority = index;
+            priorities.add(priority);
+        }
+        return priorities;
+    }
+    
+    public String completedDAGPriorityString() {
+        StringBuilder b = new StringBuilder("[");
+        boolean first = true;
+        for (int priority : completedDAGPriorities()) {
+            if (!first) {
+                b.append(", ");
+            }
+            b.append(priority);
+            first = false;
+        }
+        b.append("]");
+        return b.toString();
+    }
+    
+    /** score = sum[ 1 / 2^priority ] */
+    public double getExponentialScore() {
+        double score = 0.0;
+        for (int priority : completedDAGPriorities()) {
+            score += 1.0 / Math.pow(2.0, priority);
+        }
+        return score;
+    }
+    
+    /** score = sum[ 1 / priority ] */
+    public double getLinearScore() {
+        double score = 0.0;
+        for (int priority : completedDAGPriorities()) {
+            score += 1.0 / (priority+1);
+        }
+        return score;
+    }
+}

### 2012-01-18 src/cws/core/algorithms/Backtrack.java

@@ -1,0 +1,298 @@
+package cws.core.algorithms;
+
+import java.util.List;
+import java.util.HashMap;
+
+import cws.core.dag.DAG;
+import cws.core.dag.Task;
+import cws.core.dag.algorithms.CriticalPath;
+import cws.core.dag.algorithms.TopologicalOrder;
+
+/**
+ * @author Gideon Juve <juve@usc.edu>
+ */
+public class Backtrack extends StaticAlgorithm {
+    
+    public Backtrack(double budget, double deadline, List<DAG> dags) {
+        super(budget, deadline, dags);
+    }
+    
+    /**
+     * Develop a plan for a single DAG
+     */
+    Plan planDAG(DAG dag, Plan currentPlan) throws NoFeasiblePlan {
+        TopologicalOrder order = new TopologicalOrder(dag);
+        
+        // Initial task assignment
+        HashMap<Task, VMType> vmTypes = new HashMap<Task, VMType>();
+        HashMap<Task, Double> runtimes = new HashMap<Task, Double>();
+        double minCost = 0.0;
+        double totalRuntime = 0.0;
+        for (Task t : order) {
+            
+            // Initially we assign each VM to the SMALL type
+            VMType vm = VMType.SMALL;
+            vmTypes.put(t, vm);
+            
+            // The runtime is just the size of the task (MI) divided by the
+            // MIPS of the VM
+            double runtime = t.size / vm.mips;
+            runtimes.put(t, runtime);
+            
+            // Compute the minimum cost of running this workflow
+            minCost += (runtime/(60*60)) * vm.price;
+            totalRuntime += runtime;
+        }
+        
+        System.out.println(" Min Cost: "+minCost);
+        System.out.println(" Total Runtime: "+totalRuntime);
+        
+        // Make sure a plan is feasible given the deadline and available VMs
+        // FIXME Later we will assign each task to its fastest VM type before this
+        CriticalPath path = new CriticalPath(order, runtimes);
+        double criticalPath = path.getCriticalPathLength();
+        System.out.println(" Critical path: "+criticalPath);
+        if (criticalPath > getDeadline()) {
+            throw new NoFeasiblePlan(
+                    "Best critical path ("+criticalPath+") " +
+                    "> deadline ("+getDeadline()+")");
+        }
+        
+        /* FIXME Later we will determine the best VM type for each task
+         *  assignEachTaskToCheapestResource()
+         *  criticalPath = computeCriticalPath()
[…]

### 2012-01-18 src/cws/core/algorithms/DPDS.java

@@ -1,0 +1,15 @@
+package cws.core.algorithms;
+
+import java.util.List;
+
+import cws.core.dag.DAG;
+import cws.core.provisioner.SimpleUtilizationBasedProvisioner;
+import cws.core.scheduler.EnsembleDynamicScheduler;
+
+public class DPDS extends DynamicAlgorithm {
+    public DPDS(double budget, double deadline, List<DAG> dags, double price, double maxScaling) {
+        super(budget, deadline, dags, price, 
+                new EnsembleDynamicScheduler(), 
+                new SimpleUtilizationBasedProvisioner(maxScaling));
+    }
+}

### 2012-01-18 src/cws/core/algorithms/DynamicAlgorithm.java

@@ -1,0 +1,126 @@
+package cws.core.algorithms;
+
+import java.util.HashSet;
+import java.util.LinkedList;
+import java.util.List;
+
+import org.cloudbus.cloudsim.Log;
+import org.cloudbus.cloudsim.core.CloudSim;
+
+import cws.core.Cloud;
+import cws.core.DAGJob;
+import cws.core.DAGJobListener;
+import cws.core.EnsembleManager;
+import cws.core.Provisioner;
+import cws.core.Scheduler;
+import cws.core.SimpleJobFactory;
+import cws.core.VM;
+import cws.core.WorkflowEngine;
+import cws.core.WorkflowEvent;
+import cws.core.dag.DAG;
+import cws.core.log.WorkflowLog;
+
+public class DynamicAlgorithm extends Algorithm implements DAGJobListener {
+    
+    private double price;
+    
+    private Scheduler scheduler;
+    
+    private Provisioner provisioner;
+    
+    private List<DAG> completedDAGs = new LinkedList<DAG>();
+    
+    private double actualCost = 0.0;
+    
+    private double actualFinishTime = 0.0;
+    
+    public DynamicAlgorithm(double budget, double deadline, List<DAG> dags, double price, Scheduler scheduler, Provisioner provisioner) {
+        super(budget, deadline, dags);
+        this.price = price;
+        this.provisioner = provisioner;
+        this.scheduler = scheduler;
+    }
+    
+    @Override
+    public double getActualCost() {
+        return actualCost;
+    }
+
+    @Override
+    public void dagStarted(DAGJob dagJob) {
+        /* Do nothing */
+    }
+    
+    @Override
+    public void dagFinished(DAGJob dagJob) {
+        actualFinishTime = Math.max(actualFinishTime, CloudSim.clock());
+    }
+
+    @Override
+    public double getActualFinishTime() {
+        return actualFinishTime;
+    }
+    
+    @Override
+    public List<DAG> getCompletedDAGs() {
+        return completedDAGs;
+    }
+    
+    public void simulate(String logname) {
+        CloudSim.init(1, null, false);
+        
+        Cloud cloud = new Cloud();
+        provisioner.setCloud(cloud);
+        
+        WorkflowEngine engine = new WorkflowEngine(
+                new SimpleJobFactory(1000), provisioner, scheduler);
+        engine.setDeadline(getDeadline());
+        engine.setBudget(getBudget());
+        
+        scheduler.setWorkflowEngine(engine);
[…]

### 2012-01-18 src/cws/core/algorithms/MaxMin.java

@@ -1,0 +1,28 @@
+package cws.core.algorithms;
+
+import java.util.List;
+
+import cws.core.dag.DAG;
+
+public class MaxMin extends MinMin {
+    public MaxMin(double budget, double deadline, List<DAG> dags) {
+        super(budget, deadline, dags);
+    }
+    
+    /**
+     * return the more expensive solution (MaxMin)
+     */
+    @Override
+    Solution bestSolution(Solution a, Solution b) {
+        if (b == null)
+            return a;
+        
+        if (a.cost > b.cost)
+            return a;
+        
+        if (a.cost == b.cost && a.slot.start < b.slot.start)
+            return a;
+        
+        return b;
+    }
+}

### 2012-01-18 src/cws/core/algorithms/MinMin.java

@@ -1,0 +1,341 @@
+package cws.core.algorithms;
+
+import java.util.HashSet;
+import java.util.LinkedList;
+import java.util.List;
+import java.util.HashMap;
+
+import cws.core.dag.DAG;
+import cws.core.dag.Task;
+import cws.core.dag.algorithms.CriticalPath;
+import cws.core.dag.algorithms.TopologicalOrder;
+
+/**
+ * @author Gideon Juve <juve@usc.edu>
+ */
+public class MinMin extends StaticAlgorithm {
+    
+    public MinMin(double budget, double deadline, List<DAG> dags) {
+        super(budget, deadline, dags);
+    }
+    
+    /**
+     * Develop a plan for a single DAG
+     */
+    Plan planDAG(DAG dag, Plan currentPlan) throws NoFeasiblePlan {
+        TopologicalOrder order = new TopologicalOrder(dag);
+        
+        // Initial task assignment
+        HashMap<Task, VMType> vmTypes = new HashMap<Task, VMType>();
+        HashMap<Task, Double> runtimes = new HashMap<Task, Double>();
+        double minCost = 0.0;
+        double totalRuntime = 0.0;
+        for (Task t : order) {
+            
+            // Initially we assign each VM to the SMALL type
+            VMType vm = VMType.SMALL;
+            vmTypes.put(t, vm);
+            
+            // The runtime is just the size of the task (MI) divided by the
+            // MIPS of the VM
+            double runtime = t.size / vm.mips;
+            runtimes.put(t, runtime);
+            
+            // Compute the minimum cost of running this workflow
+            minCost += (runtime/(60*60)) * vm.price;
+            totalRuntime += runtime;
+        }
+        
+        System.out.println(" Min Cost: "+minCost);
+        System.out.println(" Total Runtime: "+totalRuntime);
+        
+        // Make sure a plan is feasible given the deadline and available VMs
+        // FIXME Later we will assign each task to its fastest VM type before this
+        CriticalPath path = new CriticalPath(order, runtimes);
+        double criticalPath = path.getCriticalPathLength();
+        System.out.println(" Critical path: "+criticalPath);
+        if (criticalPath > getDeadline()) {
+            throw new NoFeasiblePlan(
+                    "Best critical path ("+criticalPath+") " +
+                    "> deadline ("+getDeadline()+")");
+        }
+        
+        /* FIXME Later we will determine the best VM type for each task
+         *  assignEachTaskToCheapestResource()
[…]

### 2012-01-18 src/cws/core/algorithms/SPSS.java

@@ -1,0 +1,295 @@
+package cws.core.algorithms;
+
+import java.util.Collections;
+import java.util.Comparator;
+import java.util.LinkedList;
+import java.util.List;
+import java.util.HashMap;
+
+import cws.core.dag.DAG;
+import cws.core.dag.Task;
+import cws.core.dag.algorithms.CriticalPath;
+import cws.core.dag.algorithms.TopologicalOrder;
+
+/**
+ * @author Gideon Juve <juve@usc.edu>
+ */
+public class SPSS extends StaticAlgorithm {
+    
+    /** Tuning parameter for deadline distribution (low alpha = runtime, high alpha = tasks) */
+    private double alpha;
+    
+    public SPSS(double budget, double deadline, List<DAG> dags, double alpha) {
+        super(budget, deadline, dags);
+        this.alpha = alpha;
+    }
+    
+    /**
+     * Develop a plan for a single DAG
+     */
+    Plan planDAG(DAG dag, Plan currentPlan) throws NoFeasiblePlan {
+        TopologicalOrder order = new TopologicalOrder(dag);
+        
+        // Initial task assignment
+        HashMap<Task, VMType> vmTypes = new HashMap<Task, VMType>();
+        HashMap<Task, Double> runtimes = new HashMap<Task, Double>();
+        double minCost = 0.0;
+        double totalRuntime = 0.0;
+        for (Task t : order) {
+            
+            // Initially we assign each VM to the SMALL type
+            VMType vm = VMType.SMALL;
+            vmTypes.put(t, vm);
+            
+            // The runtime is just the size of the task (MI) divided by the
+            // MIPS of the VM
+            double runtime = t.size / vm.mips;
+            runtimes.put(t, runtime);
+            
+            // Compute the minimum cost of running this workflow
+            minCost += (runtime/(60*60)) * vm.price;
+            totalRuntime += runtime;
+        }
+        
+        System.out.println(" Min Cost: "+minCost);
+        System.out.println(" Total Runtime: "+totalRuntime);
+        
+        // Make sure a plan is feasible given the deadline and available VMs
+        // FIXME Later we will assign each task to its fastest VM type before this
+        CriticalPath path = new CriticalPath(order, runtimes);
+        double criticalPath = path.getCriticalPathLength();
+        System.out.println(" Critical path: "+criticalPath);
+        if (criticalPath > getDeadline()) {
+            throw new NoFeasiblePlan(
+                    "Best critical path ("+criticalPath+") " +
[…]

### 2012-01-18 src/cws/core/algorithms/StaticAlgorithm.java

@@ -1,0 +1,613 @@
+package cws.core.algorithms;
+
+import java.util.HashMap;
+import java.util.HashSet;
+import java.util.LinkedHashSet;
+import java.util.LinkedList;
+import java.util.List;
+import java.util.Queue;
+import java.util.SortedSet;
+import java.util.TreeMap;
+
+import org.cloudbus.cloudsim.core.CloudSim;
+
+import cws.core.Cloud;
+import cws.core.DAGJob;
+import cws.core.DAGJobListener;
+import cws.core.EnsembleManager;
+import cws.core.Job;
+import cws.core.JobListener;
+import cws.core.Provisioner;
+import cws.core.Scheduler;
+import cws.core.VM;
+import cws.core.VMListener;
+import cws.core.WorkflowEngine;
+import cws.core.WorkflowEvent;
+import cws.core.Job.Result;
+import cws.core.dag.DAG;
+import cws.core.dag.Task;
+import cws.core.dag.algorithms.CriticalPath;
+import cws.core.dag.algorithms.TopologicalOrder;
+import cws.core.log.WorkflowLog;
+
+public abstract class StaticAlgorithm extends Algorithm implements WorkflowEvent, Provisioner, Scheduler, VMListener, JobListener, DAGJobListener {
+    
+    /** Engine that executes workflows */
+    private WorkflowEngine engine;
+    
+    /** Ensemble manager that submits DAGs */
+    private EnsembleManager manager;
+    
+    /** Cloud to provision VMs from */
+    private Cloud cloud;
+    
+    /** Plan */
+    private Plan plan = new Plan();
+    
+    /** List of DAGs that were admitted to run */
+    private List<DAG> admittedDAGs = new LinkedList<DAG>();
+    
+    /** Set of jobs that are ready to run arranged by Task */
+    private HashMap<Task, Job> readyJobs = new HashMap<Task, Job>();
+    
+    /** Mapping of Task to the VM that will run the task */
+    private HashMap<Task, VM> taskMap = new HashMap<Task, VM>();
+    
+    /** Schedule of tasks for each VM */
+    private HashMap<VM, Queue<Task>> vmQueues = new HashMap<VM, Queue<Task>>();
+    
+    /** Set of idle VMs */
+    private HashSet<VM> idle = new HashSet<VM>();
+    
+    private int dagsFinished = 0;
+    
+    private double actualFinishTime = 0.0;
+    
[…]

### 2012-01-18 src/cws/core/algorithms/TestRun.java

@@ -1,0 +1,78 @@
+package cws.core.algorithms;
+
+import java.io.File;
+import java.util.ArrayList;
+import java.util.List;
+import java.util.Random;
+
+import cws.core.dag.DAG;
+import cws.core.dag.DAGParser;
+import cws.core.experiment.DAGListGenerator;
+
+public class TestRun {
+    public static void main(String[] args) throws Exception {
+        List<DAG> dags = new ArrayList<DAG>();
+        
+        String[] names = DAGListGenerator.generateDAGListPareto(
+                new Random(0), 
+                "/Volumes/HDD/SyntheticWorkflows/CyberShake/CYBERSHAKE", 
+                100);
+        
+        for (String name : names) {
+            DAG dag = DAGParser.parseDAG(new File(name));
+            
+            // LARGER TASKS
+            /*
+            for (String tid : dag.getTasks()) {
+                Task t = dag.getTask(tid);
+                t.size = t.size * 100;
+            }
+            */
+            dags.add(dag);
+        }
+        
+        double deadline = 10*3600;
+        double budget = 50;
+        double price = 1;
+        
+        double alpha = 0.7;
+        
+        double maxScaling = 2.0;
+        
+        Algorithm[] algos = new Algorithm[]{
+                /*
+            
+            new MaxMin(budget, deadline, dags),
+            new MinMin(budget, deadline, dags),*/
+            new Wide(budget, deadline, dags),
+            new Backtrack(budget, deadline, dags),
+            new SPSS(budget, deadline, dags, alpha),
+            new DPDS(budget, deadline, dags, price, maxScaling),
+            new WADPDS(budget, deadline, dags, price, maxScaling)
+        };
+        
+        for (Algorithm a : algos) {
+            a.simulate(a.getName());
+        }
+        
+        System.out.println("DAGs Completed:");
+        for (Algorithm a : algos) {
+            System.out.printf("    %10s: %s\n", a.getName(), a.completedDAGPriorityString());
+        }
+        
+        System.out.println("Num DAGs Completed:");
+        for (Algorithm a : algos) {
+            System.out.printf("    %10s: %d\n", a.getName(), a.numCompletedDAGs());
+        }
+        
+        System.out.println("Exponential score:");
+        for (Algorithm a : algos) {
+            System.out.printf("    %10s: %.10f\n", a.getName(), a.getExponentialScore());
+        }
+        
+        System.out.println("Linear score:");
[…]

### 2012-01-18 src/cws/core/algorithms/WADPDS.java

@@ -1,0 +1,15 @@
+package cws.core.algorithms;
+
+import java.util.List;
+
+import cws.core.dag.DAG;
+import cws.core.provisioner.SimpleUtilizationBasedProvisioner;
+import cws.core.scheduler.WorkflowAwareEnsembleScheduler;
+
+public class WADPDS extends DynamicAlgorithm {
+    public WADPDS(double budget, double deadline, List<DAG> dags, double price, double maxScaling) {
+        super(budget, deadline, dags, price, 
+                new WorkflowAwareEnsembleScheduler(), 
+                new SimpleUtilizationBasedProvisioner(maxScaling));
+    }
+}

### 2012-01-18 src/cws/core/algorithms/Wide.java

@@ -1,0 +1,30 @@
+package cws.core.algorithms;
+
+import java.util.List;
+
+import cws.core.dag.DAG;
+
+/**
+ * @author Gideon Juve <juve@usc.edu>
+ */
+public class Wide extends Backtrack {
+    
+    public Wide(double budget, double deadline, List<DAG> dags) {
+        super(budget, deadline, dags);
+    }
+    
+    public void plan() {
+        // Estimate number of nodes we can use
+        double hours = getDeadline() / (60*60);
+        double nodeHours = getBudget() / VMType.SMALL.price;
+        int N = (int) Math.floor(nodeHours/hours);
+        
+        // Add them to the initial plan
+        Plan plan = getPlan();
+        for (int i=0; i<N; i++) {
+            plan.resources.add(new Resource(VMType.SMALL));
+        }
+        
+        super.plan();
+    }
+}

### 2012-01-18 src/cws/core/algorithms/StaticAlgorithm.java
Existing resources before earlier starts

@@ -555,14 +555,6 @@ public abstract class StaticAlgorithm extends Algorithm implements WorkflowEvent
                 return false;
             }
             
-            // Earlier starts are better
-            if (this.slot.start < other.slot.start) {
-                return true;
-            }
-            if (this.slot.start > other.slot.start) {
-                return false;
-            }
-            
             // Existing resources are better
             if (!this.newresource && other.newresource) {
                 return true;
@@ -571,7 +563,14 @@ public abstract class StaticAlgorithm extends Algorithm implements WorkflowEvent
                 return false;
             }
             
-            //throw new RuntimeException("Here");
+            // Earlier starts are better
+            if (this.slot.start < other.slot.start) {
+                return true;
+            }
+            if (this.slot.start > other.slot.start) {
+                return false;
+            }
+            
             return true;
         }
         

### 2012-01-18 paper/paper.tex
updated paper with new SPSS

@@
 low-priority tasks may delay execution of high-priority tasks when such ones
-become ready. Fig.~\ref{fig:spss-example} shows the example plan generated using
+become ready. Fig.~\ref{fig:spds-example} shows the example plan generated using
 SPDS algorithm. It can be seen that tasks from lower priority workflows can fill
@@
  by colors depending on the workflow. }
-\label{fig:spss-example}
+\label{fig:spds-example}
 \end{figure}
@@
 \subsection{Workflow-Aware DPDS (WA-DPDS)}
+
+DPDS algorithm does not use any information about workflow structure: it deals
+only with scheduling of individual released/ready tasks using their
+workflow priorities. If there are no high priority tasks in the ready queue,
+then lower priority tasks are submitted to idle resources. This may lead to the
+deficiency in the following scenario: if a long-running task of low
+priority is submitted to the idle machine and then several high
+priority tasks are released to the ready queue, they become delayed by the
+execution of the low priority task. This usually does not cause a problem when
+the number of tasks and resources is high, but when the execution approaches the
+budget and deadline constraint, we have observed that tasks of low-priority
+workflows may prevent higher priority workflows from completion. This
+observation leads to the improvemed algorithm which we call workflow-aware DPDS
+(WA-DPDS).
+
+WA-DPDS extends DPDS by introducing a workflow admission procedure, which is
+invoked each time a task of a new workflow is at the head of priority queue,
+i.e. just before it is submitted. The procedure estimates if there is enough
+budget remaining to admit a new workflow: if not, all its tasks are rejected.
+Algorithm~\ref{alg:wa-dpds} estimates the cost of admitting a new workflow and
+compares it to the current cost (consumed budget) and remaining budget, taking
+into account that started VMs have been already paid for. It adds a small safety
+margin choosen to avoid overestimation of a budget.
+
+\begin{algorithm}
+\caption{Workflow admission algorithm for WA-DPDS}
+\label{alg:wa-dpds}
+\begin{algorithmic}[1]
+\Require $w$: workflow; $b$: budget; $c$: current cost
+\Procedure{Admit}{$w,b,c$}
+    \State $r_n\gets b-c$ \Comment{Remaining budget for new VMs}
+    \State $r_c\gets remaining\ cost\ on\ running\ VMs$
+    \State $r_a\gets remaining\ cost\ of\ admitted\ workflows$
+	\State $r_m\gets 0.1$ \Comment{Safety margin}
+	\State $r_b\gets r_n+r_c-r_a-r_m$ \Comment{Budget remaining}
+	\State $c_w\gets estimate\ cost(w)$
+    \If{$c_w<r_b$}
+    	\State \textbf{return} $TRUE$
+    \Else
+    	\State \textbf{return} $FALSE$
+	\EndIf    	 
+\EndProcedure
+\end{algorithmic} 
+\end{algorithm}
+
+Admission procedure relies only on the total estimated resource consumption and
+compares it to the budget remaining. We found that this estimation is useful not
+only to prevent low-priority workflows from delaying ligh-priority ones, but
+also in the case of ensembles consisting of workflows of non-uniform sizes, it
+allows to reject big and costly workflows that would overrun the budget and
+instead schedule some smaller ones to efficiently utilize idle resources. It
+would be also possible to extend the admission procedure to check other
+constraints, such as whether estimated critical path exceeds the deadline.
 
@@
 
+\begin{figure}[htb] 
+\centering
+\includegraphics[width=1.0\columnwidth]{figures/spss-gantt}
+ \caption{Example of schedule generated using SPSS algorithm: tasks are labelled
+ by colors depending on the workflow. }
+\label{fig:spss-example}
+\end{figure}
+
 \section{Performance Evaluation}
@@
 
+\begin{figure}[tb] 
+\centering
+\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
+\caption{Pareto-like distribution of workflow sizes in ensemble. }
+\label{fig:ensemble-distribution}
+\end{figure}
+
 We assumed that the interesting number of workflows in ensemble should be in the
@@
 
-\begin{figure}[htb] 
+\begin{figure*}[t] 
 \centering
-\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
-\caption{Pareto-like distribution of workflow sizes in ensemble. }
-\label{fig:ensemble-distribution}
+\includegraphics[width=0.19\textwidth]{figures/pareto-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
+algorithms on ensembles of 100 Pareto-distributed workflows}
+\label{fig:number-complete-pareto}
+\end{figure*}
+
+\begin{figure}[b!] 
+\centering
+%\includegraphics[width=0.19\textwidth]{figures/pareto-score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+%\includegraphics[width=0.19\textwidth]{figures/pareto-score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-score-LIGO-n-1000-8-dagh1-40m0.pdf}
+%\includegraphics[width=0.19\textwidth]{figures/pareto-score-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-score-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{Comparison of algorithms based on score, which  is computed as
+$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
+priority).}
+\label{fig:score}
 \end{figure}
@@
 
+\begin{figure*}[t]  
+\centering
+\includegraphics[width=0.19\textwidth]{figures/pareto-size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-size-LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-size-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-size-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\caption{ Total work completed expressed as the sum of runtimes (in hours) of
+tasks of completed workflows.Data computed for DPDS, AW-DPDS and SPSS
+algorithms on ensembles of 100 Pareto-distributed workflows}
+\label{fig:total-time}
+\end{figure*}
+
 In the tests we we assumed that all the VMs are homogeneous and have the
@@
 
+We show the results of three algorithms (TODO: ADD MORE): DPDS, WA-DPDS and
+SPSS. These tests were run with maximum autoscaling factor set to 1.0 for
+DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that due to
+high parallelism of workflows the resource utilization remains high most of the time, 
+which leads to rapid provisioning of new resources at the beginning of ensemble
+execution. This in turn results in less efficient total utilization due to high
+level of parallelism. Setting the scaling limit to 1.0, ensures that the shape
+of resource provisioning plan is closest to a rectangle (see
+Fig.~\ref{fig:spds-example}) and the VM termination procedure
+(Algorithm~\ref{alg:prov}) allows maximum utilization of VMs before shutdown.
+
 \subsection{Metrics}
@@
 \end{itemize}
-
-\begin{figure*}[t] 
-\centering
-\includegraphics[width=0.19\textwidth]{figures/MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
-algorithms on ensembles of 100 Pareto-distributed workflows}
-\label{fig:number-complete-pareto}
-\end{figure*}
 
@@
 
-\begin{figure}[b!] 
-\centering
-%\includegraphics[width=0.19\textwidth]{figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-%\includegraphics[width=0.19\textwidth]{figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/score-LIGO-n-1000-8-dagh1-40m0.pdf}
-%\includegraphics[width=0.19\textwidth]{figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/score-SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Comparison of algorithms based on score, which  is computed as
-$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
-priority).}
-\label{fig:score}
-\end{figure}
-
 We run the simulations for the constraint values ranging between the
@@
 \centering
-\includegraphics[width=0.19\textwidth]{figures/size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/size-LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/size-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/size-SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
-Pareto-distributed workflows. Score is equal to the total work completed,
-i.e. it is sum of runtimes (in hours) of tasks of completed workflows.}
-\label{fig:total-time}
-\end{figure*}
-
-\begin{figure*}[b!] 
-\centering
-\includegraphics[width=0.19\textwidth]{figures/cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/cost-LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-cost-LIGO-n-1000-8-dagh1-40m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
+\includegraphics[width=0.19\textwidth]{figures/pareto-cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
 \caption{Effective computing cost in \$ per hour calclated by dividing sum of
@@
 admit lower-priority workflows at the cost of the more important ones.
+
+\begin{figure}[htb] 
+\centering
+\includegraphics[width=0.6\columnwidth]{figures/pareto-budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}
+\caption{Comparison of number of workflows completed for a given deadline,
+plotted versus different budget amounts.}
+\label{fig:budgets}
+\end{figure}
 
@@
 
-\begin{figure}[htb] 
-\centering
-\includegraphics[width=0.6\columnwidth]{figures/budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}
-\caption{Comparison of number of workflows completed for a given deadline,
-plotted versus different budget amounts.}
-\label{fig:}
-\end{figure}
+Fig.~\ref{fig:budgets} whows the same data for SIPHT ensemble from another
+perspective. It helps answer the question: given a deadline, how many workflows
+can be finished when we increse the budget. In Fig.~\ref{fig:budgets} the growth
+is super-linear, what results from the distribution of workflow sizes in the
+ensemble (see Fig.~\ref{fig:ensemble-distribution}). For ensemble of equal size
+workflows (see Fig. TODO: ADD FIGURE FROM CONSTANT DISTRIBUTION) the growth is
+nearly linear. It can be also observed that staitc algorithms give better
+results for smaller budgets, while dynamic algorithms for larger ones.
 
 \section{Conclusion and Future Work}
+
+Summary:
+
+ - new interesting problem was addressed and analyzed
+
+ - dynamic and static strtegies were evaluated
+
+ - new algorithms were developed
+
+ - simulation studies were done
+
+Lessons learned:
+
+ - pros and cons of dynamic/static methods
+
+ - problems faced and solved
+
+ - observations about parallelization level, shape of the plan
+
+Future work:
+
+ - deal with uncertaineties of predictions
+
+ - deal with heterogeneity and multiple clouds (ref to Gideon's and Jens'
+ experiments on EC2 and FutureGrid)
+
+ - add data transfers to cost models (ref to paper on data sharing)
 

### 2012-01-18 scripts/R/stats.R
adjusted scripts to new tests

@@ -1,42 +1,3 @@
-#path <- "/home/malawski/cloudworkflowsimulator/output/"
-
-path <- ""
-prefix <- ""
-
-dpdsStats("Montage", prefix, "MONTAGE.n.1000.8.dag", c(20.0, 80.0), '1-20', 0)
-#dpdsStats("Montage", prefix, "MONTAGE.n.1000.8.dag", c(20.0, 30.0, 50.0, 60.0, 80.0), '1-20', 0)
-#dpdsStats("MONTAGE.n.1000.8.dag", c(10.0, 20.0, 30.0, 40.0, 50.0, 60.0, 70.0, 80.0, 90.0, 100.0), '1-20', 0)
-#dpdsStats("MONTAGE.n.1000.0.dag", c(40.0, 80.0, 120.0, 160.0, 200.0), '1-20', 2, 0:9)
-
-dpdsStats("CyberShake", prefix, "CYBERSHAKE.n.1000.8.dag", c(50.0, 100.0), '1-20', 0)
-#dpdsStats("CyberShake", prefix, "CYBERSHAKE.n.1000.8.dag", c(30.0, 50.0, 80.0, 100.0, 140.0), '1-20', 0)
-#dpdsStats("CYBERSHAKE.n.1000.8.dag", c(10.0, 20.0, 30.0, 40.0, 50.0, 60.0, 80.0, 100.0, 120.0, 140.0), '1-20', 0)
-#dpdsStats("CYBERSHAKE.n.1000.8.dag", c(40.0, 80.0, 120.0, 160.0, 200.0), '1-20', 0)
-#dpdsStats("CYBERSHAKE.n.1000.0.dag", c(40.0, 80.0, 120.0, 160.0, 200.0), '1-20', 2, 0:9)
-
-dpdsStats("LIGO", prefix, "LIGO.n.1000.8.dag", c(600.0, 1200.0), '1-40', 0)
-#dpdsStats("LIGO", prefix, "LIGO.n.1000.8.dag", c(400.0, 600.0, 800.0, 1000.0, 1200.0), '1-40', 0)
-#dpdsStats("LIGO.n.1000.8.dag", c(200.0, 400.0, 600.0, 800.0, 1000.0, 1200.0, 1400.0, 1600.0, 1800.0, 2000.0), '1-40', 0)
[…]

### 2012-01-18 scripts/copy_plots_to_paper.sh

@@ -1,15 +1,15 @@
 #!/bin/bash
-#cp MONTAGE.n.1000.8.dagh1-20m0.pdf ../paper/figures/MONTAGE-n-1000-8-dagh1-20m0.pdf
-#cp CYBERSHAKE.n.1000.8.dagh1-20m0.pdf ../paper/figures/CYBERSHAKE-n-1000-8-dagh1-20m0.pdf
-#cp LIGO.n.1000.8.dagh1-40m0.pdf ../paper/figures/LIGO-n-1000-8-dagh1-40m0.pdf
-#cp GENOME.n.1000.8.dagh100-1500m0.pdf ../paper/figures/GENOME-n-1000-8-dagh100-1500m0.pdf
-#cp SIPHT.n.1000.8.dagh5-50m0.pdf ../paper/figures/SIPHT-n-1000-8-dagh5-50m0.pdf
+#cp pareto-MONTAGE.n.1000.8.dagh1-20m0.pdf ../paper/figures/pareto-MONTAGE-n-1000-8-dagh1-20m0.pdf
+#cp pareto-CYBERSHAKE.n.1000.8.dagh1-20m0.pdf ../paper/figures/pareto-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf
+#cp pareto-LIGO.n.1000.8.dagh1-40m0.pdf ../paper/figures/pareto-LIGO-n-1000-8-dagh1-40m0.pdf
+#cp pareto-GENOME.n.1000.8.dagh100-1500m0.pdf ../paper/figures/pareto-GENOME-n-1000-8-dagh100-1500m0.pdf
+#cp pareto-SIPHT.n.1000.8.dagh5-50m0.pdf ../paper/figures/pareto-SIPHT-n-1000-8-dagh5-50m0.pdf
 
-#cp score-MONTAGE.n.1000.8.dagh1-20m0.pdf ../paper/figures/score-MONTAGE-n-1000-8-dagh1-20m0.pdf
-#cp score-CYBERSHAKE.n.1000.8.dagh1-20m0.pdf ../paper/figures/score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf
-#cp score-LIGO.n.1000.8.dagh1-40m0.pdf ../paper/figures/score-LIGO-n-1000-8-dagh1-40m0.pdf
-#cp score-GENOME.n.1000.8.dagh100-1500m0.pdf ../paper/figures/score-GENOME-n-1000-8-dagh100-1500m0.pdf
[…]

### 2012-01-18 src/cws/core/experiment/Experiment.java
small fix to adjust to new packages

@@ -44,7 +44,7 @@ public class Experiment {
 		
 		Scheduler sched = param.getScheduler();
 		
-		if (sched.getClass().getName() == "cws.core.SPSS") useSPSS = true;
+		if (sched.getClass().getName() == "cws.core.algorithms.SPSS") useSPSS = true;
 		
 		CloudSim.init(1, null, false);
 

### 2012-01-19 scripts/R/stats.R
count rankings

@@ -1,280 +1,337 @@
 
+global_titles <<- ""
+global_rankings <<- cbind(c(0,0,0))
+global_rankings_scores <<- cbind(c(0,0,0))
 
 
+score_weight <- function(p) {
+        return (1/(p+1))
+}
 
 readScores <- function(prefix, dag, budgets, runIds, deadline, n_deadlines, max_scaling, suffix) {
-	n_budgets = length(budgets)
-	n_runIds = length(runIds)
-	
-	
-	s <- array(0,  dim=c(length(runIds), n_budgets, n_deadlines))
-	
-	for (runId in runIds) {
-		score_files <- paste(path, prefix, dag, "b", budgets, ".0h", deadline, "m", max_scaling, ".0", "run", runId, suffix, sep="")
-		
-		for (i in 1:n_budgets) {
-			lines = readLines(score_files[i], n=-1)
-			#print(lines)
-			deadline_id = 0
-			for (line in lines) {
-				deadline_id = deadline_id +1
-				scores = as.numeric(unlist(strsplit(line, " ")))
-				n_scores = length(scores)
-				score = 0.0
-				# we skip the first value since it contains a deadline
-				for (p in scores[2:n_scores]) {
-					#score = score + 1/(p+1)
-					#score = score + 2^(-p)
-					score = score + p/3600.0
-				}
-				#print(score)
-				#s[runId+1,i,deadline_id] = score
-				s[runId+1,i,deadline_id] = budgets[i]/score
-				
-			}
-		}
-	}
-	return(s)
+        n_budgets = length(budgets)
+        n_runIds = length(runIds)
+
+
+        s <- array(0,  dim=c(length(runIds), n_budgets, n_deadlines))
+
+        for (runId in runIds) {
+                score_files <- paste(path, prefix, dag, "b", budgets, ".0h", deadline, "m", max_scaling, ".0", "run", runId, suffix, sep="")
+
+                for (i in 1:n_budgets) {
+                        lines = readLines(score_files[i], n=-1)
+                        #print(lines)
+                        deadline_id = 0
+                        for (line in lines) {
+                                deadline_id = deadline_id +1
+                                scores = as.numeric(unlist(strsplit(line, " ")))
+                                n_scores = length(scores)
+                                score = 0.0
[…]

### 2012-01-19 paper/paper.tex
Revised section 5

@@
 \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
-\caption{Pareto-like distribution of workflow sizes in ensemble. }
+\caption{Pareto-like distribution of workflow sizes in ensemble. Size is
+measured in number of tasks.}
 \label{fig:ensemble-distribution}
@@
 \caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
-algorithms on ensembles of 100 Pareto-distributed workflows}
+algorithms on ensembles of 100 Pareto-distributed workflows, {\em small budget}
+(top) and {\em large budget} (bottom).}
 \label{fig:number-complete-pareto}
@@
 
-\begin{figure}[b!] 
-\centering
+%\begin{figure}[b!] 
+%\centering
 %\includegraphics[width=0.19\textwidth]{figures/pareto-score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
 %\includegraphics[width=0.19\textwidth]{figures/pareto-score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-score-LIGO-n-1000-8-dagh1-40m0.pdf}
+%\includegraphics[width=0.19\textwidth]{figures/pareto-score-LIGO-n-1000-8-dagh1-40m0.pdf}
 %\includegraphics[width=0.19\textwidth]{figures/pareto-score-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-score-SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Comparison of algorithms based on score, which  is computed as
-$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
-priority).}
-\label{fig:score}
-\end{figure}
+%\includegraphics[width=0.19\textwidth]{figures/pareto-score-SIPHT-n-1000-8-dagh5-50m0.pdf}
+%\caption{Comparison of algorithms based on score, which  is computed as
+%$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$
+%(0 means highest priority).}
+%\label{fig:score}
+%\end{figure}
 
@@
 University~\cite{Ramakrishnan08}. We constructed ensembles of 100 DAGs by sampling the
-workflows of different sizes based on a given distribution. Two distributions
-are considered:
+workflows of different sizes based on a given distribution. By the size of the
+workflow means the number of jobs. Two distributions are considered:
 \begin{itemize}
@@
 \caption{ Total work completed expressed as the sum of runtimes (in hours) of
-tasks of completed workflows.Data computed for DPDS, AW-DPDS and SPSS
+tasks of completed workflows. Data computed for DPDS, AW-DPDS and SPSS
 algorithms on ensembles of 100 Pareto-distributed workflows}
@@
 
+%\begin{equation}
+%\label{eq:score}
+%s = \sum_{DAGs\ completed}(1/(p+1))
+%\end{equation}
+
 \begin{equation}
 \label{eq:score}
-s = \sum_{DAGs\ completed}(1/(p+1))
+s = \sum_{DAGs\ completed}2^{-p}
 \end{equation}
@@
 where priorities of subsequent workflows in the ensemble are $p=0,1,2,\ldots$ (0
-means highest priority). This scoring function gives more weight to the top high
-priority workflows, whereas the low priority ones are trated almost equally.
+means highest priority). This exponential scoring function gives a strong weight
+to the top priority workflows, since it has a property that a score for
+completing a single workflow with priority $p$ is higher that the score for
+completing all lower priority ones: 
+\begin{equation}
+\label{eq:score-property}
+2^{-p} > \sum_{i=p+1,\ldots}2^{-i}
+\end{equation}
 
@@
 \includegraphics[width=0.19\textwidth]{figures/pareto-cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Effective computing cost in \$ per hour calclated by dividing sum of
+\caption{Effective computing cost in \$ per hour calculated by dividing sum of
 task runtimes by the total budget. Higher cost results from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
@@
 executed in parallel. For these cases, as well for the other ensembles with
-large budgets, the simple dynamic scheduling techniqies perform equally well or
+large budgets, the simple dynamic scheduling techniques perform equally well or
 better. The reason for that behavior is that the large number of tasks
@@
 We can also observe that the workflow-aware dynamic algorithm (WA-DPDS) performs
-better that the workflow-unaware version. This means that for the tested
+better than the workflow-unaware version. This means that for the tested
 workflows the simple admission procedure based on estimation of workflow cost
@@
 
-Fig.~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
-computed for two sample applications. Comparing
-Fig.~\ref{fig:number-complete-pareto} to Fig.~\ref{fig:score} we can see that
-the latter one gives more flattened shape: this means that e.g. the improvement
-of the solution by one workflow may mean adding a one low-priority workflow.
-Nevertheless we can see that the curves on both figures have a very similar
-shape, which mens that our algorithms obey the priority rules and do not try to
-admit lower-priority workflows at the cost of the more important ones.
+%Fig.~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
+%computed for two sample applications. Comparing
+%Fig.~\ref{fig:number-complete-pareto} to Fig.~\ref{fig:score} we can see that
+%the latter one gives more flattened shape: this means that e.g. the improvement
+%of the solution by one workflow may mean adding a one low-priority workflow.
+%Nevertheless we can see that the curves on both figures have a very similar
+%shape, which mens that our algorithms obey the priority rules and do not try to
+%admit lower-priority workflows at the cost of the more important ones.
 
@@
 
+\begin{table}[tb]
+\centering
+
+(a) Number of workflows completed $N_c$:
+\medskip
+\begin{tabular}{r|cccccc}
+ & M & C & L & E & S & TOTAL\tabularnewline
+\hline
+DPDS      &   57  & 45 &  18 &  14  &  0 & 134\tabularnewline
+WA-DPDS   &    96 &  87  & 48  & 34  &  0 & 265\tabularnewline
+SPSS     &    29  & 42  & 192  & 74  & 50 & 387\tabularnewline
+\end{tabular}
+\medskip
+
+(b) Total amount of computing time completed $T_C$:
+\medskip
+\begin{tabular}{r|cccccc}
+ & M & C & L & E & S & TOTAL\tabularnewline
+\hline
+DPDS      &   58  & 45 &  19 &  14  &  0 & 136\tabularnewline
+WA-DPDS   &    90  & 96  & 57  & 27 &   8 & 278\tabularnewline
+SPSS     &    31 &  18 &  170  & 75 &  42 & 336\tabularnewline
+\end{tabular}
+\medskip
+
+(c) Exponential score $s$:
+\medskip
+\begin{tabular}{r|cccccc}
+ & M & C & L & E & S & TOTAL\tabularnewline
+\hline
+DPDS      &   58 &  45 &  46 &  15  &  9 & 173\tabularnewline
+WA-DPDS   &    90 &  96  & 84  & 29 &  17 & 316\tabularnewline
+SPSS     &    48 &  27 & 173 &  75 &  42 & 365\tabularnewline
+\end{tabular}
+\medskip
+
+M -- Montage; C -- CyberShake; L -- LIGO; E -- Epigenomics; S --SIPHT
+
+\label{tab:num-dags-pareto}
+\caption{Number of times each algorithm achieved the highest score for each
+application (over all budgets and deadlines). Data based on total 525 runs on Pareto-distributed ensembles.}
+\end{table}
+
 Fig.~\ref{fig:total-time} presents the same results using the total computing
 time metric $T_C$. These plots confirm our previous observations about the
-algorithms performance, but they also allow to observe a ore general property.
+algorithms performance, but they also allow to observe a more general property.
 For a given budget, when the deadline is tightened the number of work done
@@
 utilization and lower parallel efficiency for short deadlines lead to
-substantial increase of actual costs of computing. This increase reaches ca.
-20\%, which means that if we need to run e.g. a SIPHT workflow ensemble with a
-short deadline, ca. \$200 out of \$1000 is spent on idle resources.
+substantial increase of actual costs of computing. This increase reaches
+$\sim$20\%, which means that if we need to run e.g. a SIPHT workflow ensemble
+with a short deadline, $\sim$\$200 out of \$1000 is spent on idle resources.
 
-Another general concluson can be drawn from Fig.~\ref{fig:cost}: despite some
+Another general conclusion can be drawn from Fig.~\ref{fig:cost}: despite some
 random fluctuations resulting mostly from the finite result space, all the
@@
 solutions to this problem. In our case the solutions were achieved by maximizing
-the number of work done, whcich direcly results in minimization of effective cost parameter for a given
+the number of work done, which direcly results in minimization of effective cost parameter for a given
 deadline. Therefore, the obtained results can be used to assist the decison in
@@
 
-% \begin{figure*}[htb] 
-% \centering
+% \begin{figure*}[htb] \centering
 % \includegraphics[width=0.19\textwidth]{figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf}
@@
 % priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
-% priority).}
-% \label{fig:}
-% \end{figure*}
+% priority).} \label{fig:} \end{figure*}
+
+In order to compare efficiency of algorithms, we counted the number of times
+each algorithm achieved the best results for all tested applications, deadlines
+and budgets. If two or more algorithms achieve the same highest results, all are
+counted as highest. Table~\ref{tab:num-dags-pareto} shows these summary values.
+These results depend on the metric: number of workflows completed does not count
+the priorities or sizes of workflows, which means that it happens more often
+that all the algorithms achieve the same (best) result. However, it does not
+mean that all algorithms complete the same workflows: that is why values
+in the tables (a), (b) and (c) are different. it can be observed that based on
+the sampled problem space the workflow-aware algorithms achieve highest score
+in $\sim$2--3 times more cases than the simple DPDS which does not take into
+account workflow structure or size. In average, SPSS performs better in
+$\sim$20--50\% more times than WA-DPDS. We have to note, that the sampled
+paramter space included the Montage and CyberShake workflows consisting of many
+small tasks, for which cases SPSS does not perform equally well.
 

### 2012-01-19 paper/paper.tex
conclusions added

@@
 
-Summary:
+In this paper we addressed a new interesting and important problem of scheduling
+and resource provisioning for scientific workflow ensembles on IaaS cloud
+infrastructures. The problem is different from previous work on grid or utility
+grid scheduling that cloud infrastructure gives more control over the resources,
+so the resource provisioning plan can be adjusted to the application
+requirements. Therefore the problem space becomes larger: it requires not
+only to select the best scheduling of tasks to available resources, but also
+select the best shape of resource provisioning plan. The formulation of a
+problem as maximization of the number of prioritized workflows completed from
+the ensemble requires also to take the decisions on admitting or rejecting workflows
+based on their estimated resource demands. We believe that such formulation of
+the bi-constrained problem is highly relevant since such
+constraints are typically imposed on many real-world projects.
 
- - new interesting problem was addressed and analyzed
+We have analyzed both static and dynamic scheduling approaches and
+developed the SPDS, DPDS, WA-DPDS and SPSS algorithms realizing these
+strategies. The algorithms were evaluated on ensembles of benchmark
+workflows which represent typical real scientific applications. For the
+purpose of evaluation we have developed a simulator that models the cloud
+infrastructure and the tightly-coupled scheduler and provisioner modules of
+workflow engine. The results of simulation studies indicate that the two algorithms (WA-DPDS and
+SPSD) that take into account the information about the workflow structure and
+estimation of task runtimes yield better results than the simple on-line
+priority-based scheduling strategy with static (constant) resource provisioning
+plan.
 
- - dynamic and static strtegies were evaluated
+During our work, several interesting lessons were learned about the workflow
+ensembles on cloud infrastructures. One result is that the simple strategy of
+SPDS performs relatively well in terms of number of workflows completed and the
+high resource utilization it can achieve. However, this method does not allow to
+give any advance prediction about the result, so to see its performance the
+whole ensemble needs to be simulated. On the other hand, the method may be
+the only choice in the case no estimates about the task runtimes are available
+in advance. It can be seen that the workflow-aware and static algorithms are
+able to find better quality solutions in terms of the defined metrics: this the
+result mainly of the admission algorithm implemented, which allow rejecting too
+large workflows and thus saving resources for completing the less costly ones.
 
- - new algorithms were developed
+Another general observation is that although clouds provide such capabilities as
+dynamic adding and removing of resources at runtime and high parallelization due
+to ilusion of unlimited resources, it turns out that is recommended to avoid
+using these capabilities when possible, in order to achieve high resource
+utilization and cost effectiveness. This observation comes from the fact that
+starting and stopping more VMs incurres a cost which can be amortized by running
+each started VM as long as possible before shutdown. This effect was confirmed
+in the case of SPSS algorithm for Montage and CyberShake workflows with small
+budgets. Similarly, high degree of parallelization (large number of nodes) leads
+also to less efficient resource utilization and in the case of clouds the drop
+in parallel efficiency means also the increase of cost. This result, clearly
+visible in Fig~\ref{fig:cost}, leads to multi-criteria decision problems and
+related trade-offs. E.g. our simulation results can be used as hints to answer
+the question how much time and money is required to actually complete all the
+workflows from the given ansemble.
 
- - simulation studies were done
+Another lesson learned from our studies is related to the shapes and granularity
+of the tasks of workflows in the ensembles. The benchmarks workflows that we
+tested have relatively large number of tasks (50--1000). Montage and CyberShake
+workflows have fine-grained tasks of short execution times, in the order of
+seconds and minutes. LIGO, Epigenomics and SIPHT have longer tasks: from minutes to hours.
+Therefore the dynamic algorithms achieved better results on the fine grained
+workflows, since the simple random-based allocation strategy leads to good
+load-balancing and high resource utilization. On the other hand, when tasks are
+longer, it becomes more beneficial to apply a static planning strategy of SPSS.
 
-Lessons learned:
-
- - pros and cons of dynamic/static methods
-
- - problems faced and solved
-
- - observations about parallelization level, shape of the plan
-
-Future work:
-
- - deal with uncertaineties of predictions
-
- - deal with heterogeneity and multiple clouds (ref to Gideon's and Jens'
- experiments on EC2 and FutureGrid)
-
- - add data transfers to cost models (ref to paper on data sharing)
+Our current study reveals multiple paths for future work. One interesting
+question would be to answer how the static and dynamic algorithms perform in the
+cases of increased uncertainities and failures, that have important influence
+on overall system performance~\cite{Sakellariou2010,Dongarra2007}. Second
+important goal would be to extend our application and infrastructure model to
+include various data storage options available on clouds. The experimental study
+on that subject~\cite{Juve2010} suggests that data demands of scientific
+workflows have a high impact not only on execution time but also on costs on
+commercial clouds. Finally, we would like to address the heterogeneity of the
+infrastructure, as multiple instance types and providers, including private
+and community clouds will make the problem more complex and
+challenging~\cite{Marshall2010,vockler11,Juve2010}.
 

### 2012-01-19 paper/paper.tex
Update SPDS and DPDS descriptions

--- 2012-01-19 paper/paper.tex
@@
 
-\subsection{Static Provisioning Dynamic Scheduling\\
+In this section we describe several algorithms we developed to execute ensembles of
+workflows on the cloud under budget and deadline constraints.
+
+\subsection{Static Provisioning Dynamic Scheduling \\
 (SPDS) }
 \label{sec:spds}
-This is the simplest strategy based on static provisioning of resources and
-on-line scheduling of workflow tasks. Given the budget $B$ and deadline $d$
-together with the hourly price $p$ of a VM it is possible to compute the number
-of VMs $N_{VM}$ to provision so that they consume the whole budget before the
-deadline.
+
+The simplest strategy for executing ensembles of workflows in the cloud is to use
+provision resources statically and schedule workflow tasks dynamically. Given a 
+budget $b$, deadline $d$, and the hourly price $p$ of a VM it is easy to calculate
+the number of VMs, $N_{VM}$, to provision so that the entire budget is consumed before
+the deadline:
 
@@
 \label{eq:static-plan}
-N_{VM} = \lceil B / d / p \rceil
+N_{VM} = \lceil b / d / p \rceil
 \end{equation}
 
-Although simple, such provisioning plan has the advantage that it
-minimizes the number of provisioning and deprovisioning requests.
-Once the resources are provisioned, the tasks are scheduled to VMs using dynamic
-priority-based Algorithm~\ref{alg:ds}.
+The SPDS algorithm statically provisions VMs at the start of the ensemble execution 
+and keeps them running until the deadline is reached, or the budget runs out.
+This provisioning plan has the advantage that it minimizes the number of provisioning 
+and de-provisioning requests.
+
+Once the VMs are provisioned, the tasks are mapped onto idle VMs using the dynamic
+priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
 
 \begin{algorithm}
-\caption{Priority-based scheduling}
+\caption{Priority-based scheduling algorithm for SPDS}
 \label{alg:ds}
 \begin{algorithmic}[1]
-\Procedure{schedule}{}
-    \State $Q\gets set\ of\ released\ tasks$
-    \State $P\gets empty\ priority\ queue$
-    \While{$Q \neq \emptyset$}
-    	\For{$task\ t\in Q$}
+\Procedure{Schedule}{}
+    \State $P\gets$ empty priority queue
+	\State $IdleVMs\gets$ set of idle VMs
+	\For{root task $t$ in all workflows}
     	\State \Call{Insert}{$t,P$} 
     \EndFor
-    	\While{$FreeVMs \neq \emptyset \wedge P \neq \emptyset $}
-    		\State $VM\gets$ \Call{SelectRandom}{$FreeVMs$}
+    \While{deadline not reached}
+    	\While{$IdleVMs \neq \emptyset\ and\ P \neq \emptyset$}
+    		\State $v\gets$ \Call{SelectRandom}{$IdleVMs$}
     		\State $t\gets$ \Call{Pop}{$P$}
-    		\State \Call{Submit}{$t,VM$}
+    		\State \Call{Submit}{$t,v$}
     	\EndWhile
-    	\State \Call{Wait}{$task\ finished$}
-    	\State \Call{Update}{$Q$}
+    	\State Wait for task $t$ to finish on VM $v$
+    	\State Update $P$ with ready children of $t$
+		\State \Call{Insert}{$v,IdleVMs$}
     \EndWhile
@@
 
-Initially all the
-released tasks from all the workflows are submitted to the queue $Q$. Then, they
-are inserted into priority queue $P$ based on the priority of workflows. If
-there are free VMs available, the tasks from the priority queue are submitted to
-the randomly choosen VMs. The process is repeated each time a task finishes.
-Using the priority queue guarantees that tasks from lower priority workflows are
-always submitted later than the higher priority ones and that the lower priority
-tasks can occupy idle VMs. However, as there is no preemption, long-running
-low-priority tasks may delay execution of high-priority tasks when such ones
-become ready. Fig.~\ref{fig:spds-example} shows the example plan generated using
-SPDS algorithm. It can be seen that tasks from lower priority workflows can fill
-idle time of provisioned virtual machines.
+Initially, the ready tasks from all workflows in the ensemble are added to a 
+priority queue based on the priority of the workflow to which they belong. If
+there are idle VMs available, and the priority queue is not empty, the next task
+from the priority queue is submitted to a randomly chosen idle VM. The process is 
+repeated until there are no idle VMs or the priority queue is empty. The scheduler
+then waits for a task to finish, adds its ready children to the priority queue,
+marks the VM as idle, and the entire process repeats until the deadline is reached.
+
+This algorithm guarantees that tasks from lower priority workflows are
+always deferred when higher-priority tasks are available, but lower-priority
+tasks can still occupy idle VMs when higher-priority tasks are not available. 
+However, because there is no preemption, long-running low-priority tasks may delay 
+the execution of higher-priority tasks. In addition, tasks from low priority 
+workflows may be executed even though there is no chance that those workflows 
+will be completed within the current budget and deadline. Fig.~\ref{fig:spds-example} 
+shows an example schedule generated using the SPDS algorithm. The figure illustrates 
+how tasks from lower priority workflows backfill idle VMs when tasks from higher 
+priority workflows are not available.
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
- \caption{Example of schedule generated using SPDS algorithm: tasks are labelled
- by colors depending on the workflow. }
+ \caption{Example schedule generated using SPDS algorithm. Tasks are colored by workflow. }
 \label{fig:spds-example}
@@
 
-Static provisioning plan used in SPDS algorithm allows no changes to resource
-pool at runtime. However, such plan may not be optimal when resource demand
-changes during execution of workflows. In policy-based provisioning
-(or auto-scaling) systems the metrics such as resource
-utilization or queue length are typicaly used to estimate curret resource
-demand. Our DPDS algorithm is based on resource utilization: the provisioner
+The static provisioning approach of SPDS may not be optimal when resource demand
+changes during ensemble execution. In those cases, SPDS may either over-provision, 
+wasting the budget on idle VMs, or under-provision, using up the deadline while starving 
+the application of usable VMs. One of the advantages of elasticity in cloud computing
+is that the number of resources can be changed in response to application demand. 
+In policy-based provisioning (or auto-scaling) systems, the number of provisioned VMs 
+is adjusted based on application demand as measured by metrics such as resource 
+utilization and queue length. The DPDS algorithm uses resource utilization
+to determine when to provision and de-provision VMs. The provisioner
 module periodically computes utilization as the percentage of idle VMs. The
@@
 
+Because it is assumed that VMs are billed by the hour, DPDS only considers VMs that are
+approaching their hourly billing cycle when deciding which VMs to terminate.
+
 \begin{algorithm}
-\caption{Dynamic provisioner}
+\caption{Dynamic provisioning algorithm for DPDS}
 \label{alg:prov}
@@
 \Require $c$: consumed budget; $b$: total budget; $d$: deadline; $p$: price;
-$t$: current time; $u_h$: upper utilization thershold; $u_l$: lower utilization
+$t$: current time; $u_h$: upper utilization threshold; $u_l$: lower utilization
 threshold.
-\Procedure{provision}{}
+\Procedure{Provision}{}
     \State $V_C\gets set\ of\ VMs\ completing\ billing\ cycle$
@@
 termination process is not immediate: to make sure that VMs that have been
-already paid for are not terminated prematurely the {\em Terminate()} procedure
+already paid for are not terminated prematurely the \textsc{Terminate} procedure
 terminates only these instances that are close to the end of their billing
@@
 workflows may prevent higher priority workflows from completion. This
-observation leads to the improvemed algorithm which we call workflow-aware DPDS
+observation leads to the improved algorithm which we call Workflow-Aware DPDS
 (WA-DPDS).

### 2012-01-19 paper/paper.tex
Update DPDS description

@@
 the application of usable VMs. One of the advantages of elasticity in cloud computing 
-is that the number of resources can be changed in response to application demand. 
-In policy-based provisioning (or auto-scaling) systems, the number of provisioned VMs 
-is adjusted based on application demand as measured by metrics such as resource 
-utilization and queue length. The DPDS algorithm uses resource utilization
-to determine when to provision and de-provision VMs. The provisioner
-module periodically computes utilization as the percentage of idle VMs. The
-provisioner algorithm detects which VMs are finishing their hourly billing cycle
-during the next provisioning period and decides whether they should continue
-running, should be terminated or new VMs should be provisioned: see
-Algorithm~\ref{alg:prov}.
-
+is that the number of resources can be changed to meet the demands of an application.
+In policy-based provisioning (or auto-scaling) systems, metrics such as resource 
+utilization and queue length are monitored to estimate application demand. The DPDS 
+algorithm uses this approach to determine when to provision and de-provision VMs. The 
+algorithm periodically computes resource utilization as the percentage of idle VMs, and
+adjusts the number of VMs if the utilization is above or below given thresholds.
 Because it is assumed that VMs are billed by the hour, DPDS only considers VMs that are
-approaching their hourly billing cycle when deciding which VMs to terminate.
+approaching their hourly billing cycle when deciding which VMs to terminate. This process
+is described in Algorithm~\ref{alg:prov}.
 
@@
 $t$: current time; $u_h$: upper utilization threshold; $u_l$: lower utilization
-threshold.
+threshold; $v_{max}$: maximum number of VMs
 \Procedure{Provision}{}
-    \State $V_C\gets set\ of\ VMs\ completing\ billing\ cycle$
-    \State $n_T\gets\ 0$ \Comment{number of VMs to terminate} 
-    \If{$ b - c < p * |V_C| \vee t > d $ }
-    	\State $V_R\gets set\ of\ running\ VMs$
+	\State $V_R\gets$ set of running VMs
+    \State $V_C\gets$ set of VMs completing billing cycle
+    \If{$ b - c < |V_C| * p\ or\ t > d $ }
     	\State $n_T\gets |V_R| - \lfloor(b-c)/p\rfloor$
-    	\State $V_T\gets select\ n_T VMs\ to\ terminate\ from\ V_C$
-    	\State \Call{Terminate}{$V_T$}
+    	\State $V_T\gets$ select $n_T$ VMs to terminate from $V_C$
+    	\State \Call{Terminate}{$V_T$} \label{l:terminate1}
     \Else 
-    	\If{$u>u_h$}
+		\State $u\gets$ current VM utilization
+    	\If{$u>u_h$ and $|V_R| < v_{max}$}
     		\State \Call{Start}{$new\ VM$}
-    	\EndIf
-    	\If{$u<u_l$}
-    		\State $V_F\gets free\ (idle)\ VMs$
-    		\State $n_T\gets \lceil|V_F|/2\rceil$ 
-    		\State \Call{Terminate}{$V_T$}
+    	\ElsIf{$u<u_l$}
+    		\State $V_I\gets$ set of idle VMs
+    		\State $n_T\gets \lceil|V_I|/2\rceil$ \label{l:nT2}
+			\State $V_T\gets$ select $n_T$ VMs to terminate from $V_I$
+    		\State \Call{Terminate}{$V_T$} \label{l:terminate2}
     	\EndIf 
@@
 
-The number of instances to terminate $n_T$ computed in line 6 is the number that
-would overrun the budget if not terminated. On the other hand $n_T$ computed in
-line 15 prevents terminating all the instances too quickly, but makes sure that
-if there is only one instance left it should be terminated. However, the
-termination process is not immediate: to make sure that VMs that have been
-already paid for are not terminated prematurely the \textsc{Terminate} procedure
-terminates only these instances that are close to the end of their billing
-cycle. The remaining time threshold is calculated as a sum of provisioner
-interval and estimated termination delay. This guarantees that no VMs can
-overrun the assigned budget. To avoid uncontrolled increase of number of
-instances, which may happen in the case of highly parallel workflows,
-provisioner checks also the maximum scaling factor from the initial number of
-provisioned VMs and prohibits starting new ones when this threshold is reached.
+The set of VMs completing their billing cycle is determined by considering both the
+provisioner interval, and the termination delay of the provider. This guarantees 
+that VMs can be terminated before they start the next billing cycle and prevents the
+budget from being overrun. The VMs terminated in line \ref{l:terminate1} are the ones 
+that would overrun the budget if not terminated in the current provisioning cycle. 
+The VMs terminated on line \ref{l:nT2} are chosen to increase the resource utilization
+to the desired threshold. In order to prevent instances from being
+terminated too quickly, potentially wasting resources that have already been paid for
+but could be used later, no more than half of the idle resources are terminated during
+each provisioning cycle. To avoid an uncontrolled increase in the number of
+instances, which may happen in the case of highly parallel workflows, the provisioner 
+will not start a new VM if the user-defined scaling parameter, $v_{max}$, has
+been reached.
 

### 2012-01-19 paper/paper.tex
Updates to WA-DPDS

@@
 
-DPDS algorithm does not use any information about workflow structure: it deals
-only with scheduling of individual released/ready tasks using their
-workflow priorities. If there are no high priority tasks in the ready queue,
-then lower priority tasks are submitted to idle resources. This may lead to the
-deficiency in the following scenario: if a long-running task of low
-priority is submitted to the idle machine and then several high
-priority tasks are released to the ready queue, they become delayed by the
-execution of the low priority task. This usually does not cause a problem when
-the number of tasks and resources is high, but when the execution approaches the
-budget and deadline constraint, we have observed that tasks of low-priority
-workflows may prevent higher priority workflows from completion. This
-observation leads to the improved algorithm which we call Workflow-Aware DPDS
-(WA-DPDS).
+The DPDS algorithm does not use any information about the structure of the workflows
+in the ensemble when scheduling tasks: it looks only at the priorities of the ready 
+tasks when deciding which task to schedule next. It does not consider whether a lower 
+priority task belongs to a workflow that it will never be able to complete given the 
+current budget and deadline. As a result, DPDS may execute lower priority tasks just 
+to keep VMs busy that will end up delaying higher priority tasks, making it less likely
+that higher priority workflows will be able to finish.
 
-WA-DPDS extends DPDS by introducing a workflow admission procedure, which is
-invoked each time a task of a new workflow is at the head of priority queue,
-i.e. just before it is submitted. The procedure estimates if there is enough
-budget remaining to admit a new workflow: if not, all its tasks are rejected.
+In order to address this issue, the Workflow-Aware DPDS (WA-DPDS) algorithm extends 
+DPDS by introducing a workflow admission procedure. The admission procedure is
+invoked whenever WA-DPDS sees the first task of a new workflow at the head of the 
+priority queue (i.e. no other tasks from that workflow have been scheduled yet). 
+The procedure estimates if there is enough
+budget remaining to admit the new workflow: if not, all its tasks are rejected.
 Algorithm~\ref{alg:wa-dpds} estimates the cost of admitting a new workflow and

### 2012-01-19 paper/paper.tex
Update exponential scores table

@@
 invoked whenever WA-DPDS sees the first task of a new workflow at the head of the 
-priority queue (i.e. no other tasks from that workflow have been scheduled yet). 
-The procedure estimates if there is enough
-budget remaining to admit the new workflow: if not, all its tasks are rejected.
+priority queue (i.e. no other tasks from the new workflow have been scheduled yet). 
+The admission procedure estimates if there is enough budget remaining to admit the 
+new workflow: if not, all its tasks are rejected.
 Algorithm~\ref{alg:wa-dpds} estimates the cost of admitting a new workflow and
@@
 into account that started VMs have been already paid for. It adds a small safety
-margin choosen to avoid overestimation of a budget.
+margin chosen to avoid overestimation of a budget.
 
@@
 \hline
-DPDS      &   58 &  45 &  46 &  15  &  9 & 173\tabularnewline
-WA-DPDS   &    90 &  96  & 84  & 29 &  17 & 316\tabularnewline
-SPSS     &    48 &  27 & 173 &  75 &  42 & 365\tabularnewline
+DPDS      &   58  & 45 &  19 &  14  &  0 & 136\tabularnewline
+WA-DPDS   &    90  & 96  & 57  & 27 &   8 & 278\tabularnewline
+SPSS     &    31 &  18 &  170  & 75 &  42 & 336\tabularnewline
 \end{tabular}

### 2012-01-19 paper/paper.tex
extended sec 2 and added sec 5.1

--- 2012-01-19 paper/paper.tex
@@
 Auto Scaling\footnote{\url{http://aws.amazon.com/autoscaling}} and
-RightScale\footnote{\url{http://www.rightscale.com}}).
+RightScale\footnote{\url{http://www.rightscale.com}}) allow adjusting the size
+of resource pool based on metrics related to infrastructure and application.
+Typical infrastructure-specific metric is system load, whereas
+application-specific metrics include response time and queue length. It is
+possible to set thresholds and limits to tune the behavior of these autoscaling
+systems, but no support for complex applications is provided.
 
-Policy-based approaches for scientific workloads (e.g.
-\cite{Marshall2010, Kim2011}).
-Our approach is different in that we consider workflows, while policy based approaches typically consider bags of independent tasks or unpredictable batch workloads.
-This enables us to take advantage of scheduling heuristics that cannot be applied to independent tasks.
+Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
+Kim2011}) also allow to scale the the cloud resource pool or to extend the
+capabilities of clusters using cloud-burst techniqies. Our approach is different
+in that we consider workflows, while policy based approaches typically consider
+bags of independent tasks or unpredictable batch workloads. This enables us to
+take advantage of scheduling heuristics that cannot be applied to independent
+tasks.
 
-Deadline-constrained cost-minimization workflow scheduling.
-Our work is different from \cite{Yu2005, Abrishami2010} in that we consider ensembles of workflows in IaaS clouds, which allow one to provision resources on a per-hour billing model, rather than utility grids, which allow one to choose from a pool of existing resources with a per-job billing model.
-Our work is different from \cite{Mao2011} in that we consider ensembles of workflows rather than unpredictable workloads containing workflows.
-We also have budget constraints rather than cost minimization as a goal.
-In other words, we assume that there is more work to be done than the available budget, so some work must be rejected.
+Our work is related to the strategies for deadline-constrained cost-minimization
+workflow scheduling, developed for utility grid systems. However, our problem is
+different from \cite{Yu2005, Abrishami2010} in that we consider ensembles of workflows in
+IaaS clouds, which allow one to provision resources on a per-hour billing model,
+rather than utility grids, which allow one to choose from a pool of existing
+resources with a per-job billing model. Our work is also different from
+cloud-targeted autoscaling solution~\cite{Mao2011} in that we consider ensembles
+of workflows rather than unpredictable workloads containing workflows. We also have budget constraints
+rather than cost minimization as a goal. In other words, we assume that there is
+more work to be done than the available budget, so some work must be rejected.
 Therefore cost is not something we optimize, but rather a constraint.
 
-Budget-constrained workflow scheduling \cite{Sakellariou2007}.
-
-Bi-criteria scheduling and multi-criteria scheduling.
-These approaches are similar to ours in that we have two scheduling criteria: cost and makespan.
-The challenge in multi-criteria scheduling is to derive an objective function that takes into account all of the criteria.
-These approaches typically use metaheuristics that run for a long time before producing good results.
+We should refer to bi-criteria scheduling and multi-criteria scheduling of
+workflows~\cite{Wieczorek2009,Prodan10,Dongarra2007}. These approaches are
+similar to ours in that we have two scheduling criteria: cost and makespan. The
+challenge in multi-criteria scheduling is to derive an objective function that
+takes into account all of the criteria. In our case one objective (amount
+of work completed) is subject to optimization, whereas time and cost are
+trated as constraints. Other approaches~\cite{Talukder2009,Pandey2010} use
+metaheuristics that usually run for a long time before producing good results,
+which makes tham less useful in practical applications. Our work can be also
+regarded as extension of the budget-constrained workflow scheduling
+\cite{Sakellariou2007} in the sense that we are dealing with workflow ensembles
+and the deadline constraint is added.
 
@@
 
-In this section we describe several algorithms we developed to execute ensembles of
-workflows on the cloud under budget and deadline constraints.
-
-\subsection{Static Provisioning Dynamic Scheduling \\
+\subsection{Static Provisioning Dynamic Scheduling\\
 (SPDS) }
 \label{sec:spds}
-
-The simplest strategy for executing ensembles of workflows in the cloud is to use
-provision resources statically and schedule workflow tasks dynamically. Given a 
-budget $b$, deadline $d$, and the hourly price $p$ of a VM it is easy to calculate
-the number of VMs, $N_{VM}$, to provision so that the entire budget is consumed before
-the deadline:
+This is the simplest strategy based on static provisioning of resources and
+on-line scheduling of workflow tasks. Given the budget $B$ and deadline $d$
+together with the hourly price $p$ of a VM it is possible to compute the number
+of VMs $N_{VM}$ to provision so that they consume the whole budget before the
+deadline.
 
@@
 \label{eq:static-plan}
-N_{VM} = \lceil b / d / p \rceil
+N_{VM} = \lceil B / d / p \rceil
 \end{equation}
 
-The SPDS algorithm statically provisions VMs at the start of the ensemble execution 
-and keeps them running until the deadline is reached, or the budget runs out.
-This provisioning plan has the advantage that it minimizes the number of provisioning 
-and de-provisioning requests.
-
-Once the VMs are provisioned, the tasks are mapped onto idle VMs using the dynamic
-priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
+Although simple, such provisioning plan has the advantage that it
+minimizes the number of provisioning and deprovisioning requests.
+Once the resources are provisioned, the tasks are scheduled to VMs using dynamic
+priority-based Algorithm~\ref{alg:ds}.
 
 \begin{algorithm}
-\caption{Priority-based scheduling algorithm for SPDS}
+\caption{Priority-based scheduling}
 \label{alg:ds}
 \begin{algorithmic}[1]
-\Procedure{Schedule}{}
-    \State $P\gets$ empty priority queue
-	\State $IdleVMs\gets$ set of idle VMs
-	\For{root task $t$ in all workflows}
+\Procedure{schedule}{}
+    \State $Q\gets set\ of\ released\ tasks$
+    \State $P\gets empty\ priority\ queue$
+    \While{$Q \neq \emptyset$}
+    	\For{$task\ t\in Q$}
     		\State \Call{Insert}{$t,P$} 
     	\EndFor
-    \While{deadline not reached}
-    	\While{$IdleVMs \neq \emptyset\ and\ P \neq \emptyset$}
-    		\State $v\gets$ \Call{SelectRandom}{$IdleVMs$}
+    	\While{$FreeVMs \neq \emptyset \wedge P \neq \emptyset $}
+    		\State $VM\gets$ \Call{SelectRandom}{$FreeVMs$}
     		\State $t\gets$ \Call{Pop}{$P$}
-    		\State \Call{Submit}{$t,v$}
+    		\State \Call{Submit}{$t,VM$}
     	\EndWhile
-    	\State Wait for task $t$ to finish on VM $v$
-    	\State Update $P$ with ready children of $t$
-		\State \Call{Insert}{$v,IdleVMs$}
+    	\State \Call{Wait}{$task\ finished$}
+    	\State \Call{Update}{$Q$}
     \EndWhile
@@
 
-Initially, the ready tasks from all workflows in the ensemble are added to a 
-priority queue based on the priority of the workflow to which they belong. If
-there are idle VMs available, and the priority queue is not empty, the next task
-from the priority queue is submitted to a randomly chosen idle VM. The process is 
-repeated until there are no idle VMs or the priority queue is empty. The scheduler
-then waits for a task to finish, adds its ready children to the priority queue,
-marks the VM as idle, and the entire process repeats until the deadline is reached.
-
-This algorithm guarantees that tasks from lower priority workflows are
-always deferred when higher-priority tasks are available, but lower-priority
-tasks can still occupy idle VMs when higher-priority tasks are not available. 
-However, because there is no preemption, long-running low-priority tasks may delay 
-the execution of higher-priority tasks. In addition, tasks from low priority 
-workflows may be executed even though there is no chance that those workflows 
-will be completed within the current budget and deadline. Fig.~\ref{fig:spds-example} 
-shows an example schedule generated using the SPDS algorithm. The figure illustrates 
-how tasks from lower priority workflows backfill idle VMs when tasks from higher 
-priority workflows are not available.
+Initially all the
+released tasks from all the workflows are submitted to the queue $Q$. Then, they
+are inserted into priority queue $P$ based on the priority of workflows. If
+there are free VMs available, the tasks from the priority queue are submitted to
+the randomly choosen VMs. The process is repeated each time a task finishes.
+Using the priority queue guarantees that tasks from lower priority workflows are
+always submitted later than the higher priority ones and that the lower priority
+tasks can occupy idle VMs. However, as there is no preemption, long-running
+low-priority tasks may delay execution of high-priority tasks when such ones
+become ready. Fig.~\ref{fig:spds-example} shows the example plan generated using
+SPDS algorithm. It can be seen that tasks from lower priority workflows can fill
+idle time of provisioned virtual machines.
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
- \caption{Example schedule generated using SPDS algorithm. Tasks are colored by workflow. }
+ \caption{Example of schedule generated using SPDS algorithm: tasks are labelled
+ by colors depending on the workflow. }
 \label{fig:spds-example}
@@
 
-The static provisioning approach of SPDS may not be optimal when resource demand
-changes during ensemble execution. In those cases, SPDS may either over-provision, 
-wasting the budget on idle VMs, or under-provision, using up the deadline while starving 
-the application of usable VMs. One of the advantages of elasticity in cloud computing 
-is that the number of resources can be changed to meet the demands of an application.
-In policy-based provisioning (or auto-scaling) systems, metrics such as resource 
-utilization and queue length are monitored to estimate application demand. The DPDS 
-algorithm uses this approach to determine when to provision and de-provision VMs. The 
-algorithm periodically computes resource utilization as the percentage of idle VMs, and
-adjusts the number of VMs if the utilization is above or below given thresholds.
-Because it is assumed that VMs are billed by the hour, DPDS only considers VMs that are
-approaching their hourly billing cycle when deciding which VMs to terminate. This process
-is described in Algorithm~\ref{alg:prov}.
+Static provisioning plan used in SPDS algorithm allows no changes to resource
+pool at runtime. However, such plan may not be optimal when resource demand
+changes during execution of workflows. In policy-based provisioning
+(or auto-scaling) systems the metrics such as resource
+utilization or queue length are typicaly used to estimate curret resource
+demand. Our DPDS algorithm is based on resource utilization: the provisioner
+module periodically computes utilization as the percentage of idle VMs. The
+provisioner algorithm detects which VMs are finishing their hourly billing cycle
+during the next provisioning period and decides whether they should continue
+running, should be terminated or new VMs should be provisioned: see
+Algorithm~\ref{alg:prov}.
 
 \begin{algorithm}
-\caption{Dynamic provisioning algorithm for DPDS}
+\caption{Dynamic provisioner}
 \label{alg:prov}
@@
 \Require $c$: consumed budget; $b$: total budget; $d$: deadline; $p$: price;
-$t$: current time; $u_h$: upper utilization threshold; $u_l$: lower utilization
-threshold; $v_{max}$: maximum number of VMs
-\Procedure{Provision}{}
-	\State $V_R\gets$ set of running VMs
-    \State $V_C\gets$ set of VMs completing billing cycle
-    \If{$ b - c < |V_C| * p\ or\ t > d $ }
+$t$: current time; $u_h$: upper utilization thershold; $u_l$: lower utilization
+threshold.
+\Procedure{provision}{}
+    \State $V_C\gets set\ of\ VMs\ completing\ billing\ cycle$
+    \State $n_T\gets\ 0$ \Comment{number of VMs to terminate} 
+    \If{$ b - c < p * |V_C| \vee t > d $ }
+    	\State $V_R\gets set\ of\ running\ VMs$
     	\State $n_T\gets |V_R| - \lfloor(b-c)/p\rfloor$ 
-    	\State $V_T\gets$ select $n_T$ VMs to terminate from $V_C$
-    	\State \Call{Terminate}{$V_T$} \label{l:terminate1}
+    	\State $V_T\gets select\ n_T VMs\ to\ terminate\ from\ V_C$
+    	\State \Call{Terminate}{$V_T$}
     \Else 
-		\State $u\gets$ current VM utilization
-    	\If{$u>u_h$ and $|V_R| < v_{max}$}
+    	\If{$u>u_h$}
     		\State \Call{Start}{$new\ VM$}
-    	\ElsIf{$u<u_l$}
-    		\State $V_I\gets$ set of idle VMs
-    		\State $n_T\gets \lceil|V_I|/2\rceil$ \label{l:nT2}
-			\State $V_T\gets$ select $n_T$ VMs to terminate from $V_I$
-    		\State \Call{Terminate}{$V_T$} \label{l:terminate2}
+    	\EndIf
+    	\If{$u<u_l$}
+    		\State $V_F\gets free\ (idle)\ VMs$
+    		\State $n_T\gets \lceil|V_F|/2\rceil$ 
+    		\State \Call{Terminate}{$V_T$}
     	\EndIf 
@@
 
-The set of VMs completing their billing cycle is determined by considering both the
-provisioner interval, and the termination delay of the provider. This guarantees 
-that VMs can be terminated before they start the next billing cycle and prevents the
-budget from being overrun. The VMs terminated in line \ref{l:terminate1} are the ones 
-that would overrun the budget if not terminated in the current provisioning cycle. 
-The VMs terminated on line \ref{l:nT2} are chosen to increase the resource utilization
-to the desired threshold. In order to prevent instances from being
-terminated too quickly, potentially wasting resources that have already been paid for
-but could be used later, no more than half of the idle resources are terminated during
-each provisioning cycle. To avoid an uncontrolled increase in the number of
-instances, which may happen in the case of highly parallel workflows, the provisioner 
-will not start a new VM if the user-defined scaling parameter, $v_{max}$, has
-been reached.
+The number of instances to terminate $n_T$ computed in line 6 is the number that
+would overrun the budget if not terminated. On the other hand $n_T$ computed in
+line 15 prevents terminating all the instances too quickly, but makes sure that
+if there is only one instance left it should be terminated. However, the
+termination process is not immediate: to make sure that VMs that have been
+already paid for are not terminated prematurely the {\em Terminate()} procedure
+terminates only these instances that are close to the end of their billing
+cycle. The remaining time threshold is calculated as a sum of provisioner
+interval and estimated termination delay. This guarantees that no VMs can
+overrun the assigned budget. To avoid uncontrolled increase of number of
+instances, which may happen in the case of highly parallel workflows,
+provisioner checks also the maximum scaling factor from the initial number of
+provisioned VMs and prohibits starting new ones when this threshold is reached.
 
@@
 
-The DPDS algorithm does not use any information about the structure of the workflows
-in the ensemble when scheduling tasks: it looks only at the priorities of the ready 
-tasks when deciding which task to schedule next. It does not consider whether a lower 
-priority task belongs to a workflow that it will never be able to complete given the 
-current budget and deadline. As a result, DPDS may execute lower priority tasks just 
-to keep VMs busy that will end up delaying higher priority tasks, making it less likely
-that higher priority workflows will be able to finish.
+DPDS algorithm does not use any information about workflow structure: it deals
+only with scheduling of individual released/ready tasks using their
+workflow priorities. If there are no high priority tasks in the ready queue,
+then lower priority tasks are submitted to idle resources. This may lead to the
+deficiency in the following scenario: if a long-running task of low
+priority is submitted to the idle machine and then several high
+priority tasks are released to the ready queue, they become delayed by the
+execution of the low priority task. This usually does not cause a problem when
+the number of tasks and resources is high, but when the execution approaches the
+budget and deadline constraint, we have observed that tasks of low-priority
+workflows may prevent higher priority workflows from completion. This
+observation leads to the improvemed algorithm which we call workflow-aware DPDS
+(WA-DPDS).
 
-In order to address this issue, the Workflow-Aware DPDS (WA-DPDS) algorithm extends 
-DPDS by introducing a workflow admission procedure. The admission procedure is
-invoked whenever WA-DPDS sees the first task of a new workflow at the head of the 
-priority queue (i.e. no other tasks from that workflow have been scheduled yet). 
-The procedure estimates if there is enough
-budget remaining to admit the new workflow: if not, all its tasks are rejected.
+WA-DPDS extends DPDS by introducing a workflow admission procedure, which is
+invoked each time a task of a new workflow is at the head of priority queue,
+i.e. just before it is submitted. The procedure estimates if there is enough
+budget remaining to admit a new workflow: if not, all its tasks are rejected.
 Algorithm~\ref{alg:wa-dpds} estimates the cost of admitting a new workflow and
@@
 \subsection{Simulator}
+
+To evaluate our algorithms, we developed a cloud workflows simulator based on
+CloudSim~\cite{Calheiros11}. To represent our infrastructure model, we defined
+Cloud, VM and WorkflowEngine simulation entities. Cloud object is responsible
+for staring and terminating VMs using API similar to Amazon EC2. VM object is
+responsible for execution of individual tasks, while WorkflowEngine deals with
+management of ensembles and workflows. WorkflowEngine uses separate, although
+tightly coupled Scheduler and Provisioner modules. We assume that a single core
+of a VM executed submitted tasks sequentially and that a task is given exclusive
+access to allocated VM. Although CloudSim provides a more advanced
+infrastructure model which includes time-sharing and space-sharing policies and
+power consumption in a datacenter, we do not use these features, since we are
+interested mainly in execution of tasks on VMs and high-level worklow scheduling
+on top of that. The simulator relies on DAG and DAX format used by Pegasus
+project and workflow gallery.
 

### 2012-01-19 scripts/R/stats.R
fixed exponential cancellation bug in scores

@@ -1,47 +1,89 @@
+library(gmp)
 
 global_titles <<- ""
 global_rankings <<- cbind(c(0,0,0))
 global_rankings_scores <<- cbind(c(0,0,0))
+global_rankings_escores <<- cbind(c(0,0,0))
 
+global_totals <<- cbind(c(0,0,0))
+global_stotals <<- cbind(c(0,0,0))
 
 score_weight <- function(p) {
         return (1/(p+1))
 }
 
-readScores <- function(prefix, dag, budgets, runIds, deadline, n_deadlines, max_scaling, suffix) {
-        n_budgets = length(budgets)
-        n_runIds = length(runIds)
-
-
-        s <- array(0,  dim=c(length(runIds), n_budgets, n_deadlines))
-
-        for (runId in runIds) {
-                score_files <- paste(path, prefix, dag, "b", budgets, ".0h", deadline, "m", max_scaling, ".0", "run", runId, suffix, sep="")
-
-                for (i in 1:n_budgets) {
-                        lines = readLines(score_files[i], n=-1)
-                        #print(lines)
-                        deadline_id = 0
-                        for (line in lines) {
-                                deadline_id = deadline_id +1
-                                scores = as.numeric(unlist(strsplit(line, " ")))
-                                n_scores = length(scores)
-                                score = 0.0
-                                # we skip the first value since it contains a deadline
-                                for (p in scores[2:n_scores]) {
-                                        #score = score + 1/(p+1)                                                                                                            
-                                        #score = score + 2^(-p)                                                                                                              
-                                        score = score + p/3600.0                                                                                                           
-                                }                                                                                                                                           
[…]

### 2012-01-19 paper/paper.tex
Update score table

--- 2012-01-19 paper/paper.tex
@@
 \begin{tabular}{r|cccccc}
- & M & C & L & E & S & TOTAL\tabularnewline
+ & M & C & L & E & S & All\tabularnewline
 \hline
@@
 
-M -- Montage; C -- CyberShake; L -- LIGO; E -- Epigenomics; S --SIPHT
-
 \label{tab:num-dags-pareto}
-\caption{Number of times each algorithm achieved the highest score for each
-application (over all budgets and deadlines). Data based on total 525 runs on Pareto-distributed ensembles.}
+\caption{Number of times each algorithm achieved the highest score 
+over all budgets and deadlines. Data from a total of 525 runs on 
+Pareto-distributed ensembles for Montage (M), CyberShake (C), LIGO (L), 
+Epigenomics (E), SIPHT (S).}
 \end{table}

### 2012-01-20 paper/paper.tex
Update SPSS

@@
 invoked whenever WA-DPDS sees the first task of a new workflow at the head of the 
-priority queue (i.e. no other tasks from the new workflow have been scheduled yet). 
-The admission procedure estimates if there is enough budget remaining to admit the 
-new workflow: if not, all its tasks are rejected.
-Algorithm~\ref{alg:wa-dpds} estimates the cost of admitting a new workflow and
-compares it to the current cost (consumed budget) and remaining budget, taking
-into account that started VMs have been already paid for. It adds a small safety
-margin chosen to avoid overestimation of a budget.
+priority queue (i.e. when no other tasks from the workflow have been scheduled yet). 
+The admission procedure---shown in Algorithm~\ref{alg:wa-dpds}---estimates if there is 
+enough budget remaining to admit the new workflow; if there is not, then the 
+workflow is rejected and its tasks are removed from the queue. This algorithm 
+compares the current cost (consumed budget) and remaining budget, taking
+into account the cost of currently running VMs, and the cost of workflows 
+that have already been admitted. It adds a small safety margin avoid going over
+the budget.
 
@@
 \Procedure{Admit}{$w,b,c$}
-    \State $r_n\gets b-c$ \Comment{Remaining budget for new VMs}
-    \State $r_c\gets remaining\ cost\ on\ running\ VMs$
-    \State $r_a\gets remaining\ cost\ of\ admitted\ workflows$
+    \State $r_n\gets b-c$ \Comment{Budget remaining for new VMs}
+    \State $r_c\gets $cost committed to VMs that are running
+    \State $r_a\gets $cost to complete workflows previously admitted
 	\State $r_m\gets 0.1$ \Comment{Safety margin}
 	\State $r_b\gets r_n+r_c-r_a-r_m$ \Comment{Budget remaining}
-	\State $c_w\gets estimate\ cost(w)$
+	\State $c_w\gets$ \Call{EstimateCost}{w}
     \If{$c_w<r_b$}
@@
 
-Admission procedure relies only on the total estimated resource consumption and
+This admission procedure relies only on the total estimated resource consumption and
 compares it to the budget remaining. We found that this estimation is useful not
-only to prevent low-priority workflows from delaying ligh-priority ones, but
-also in the case of ensembles consisting of workflows of non-uniform sizes, it
-allows to reject big and costly workflows that would overrun the budget and
-instead schedule some smaller ones to efficiently utilize idle resources. It
-would be also possible to extend the admission procedure to check other
-constraints, such as whether estimated critical path exceeds the deadline.
+only to prevent low-priority workflows from delaying high-priority ones, but
+also to reject large and costly workflows that would overrun the budget and
+admit smaller workflows that can efficiently utilize idle resources in ensembles 
+containing workflows of non-uniform sizes. It would also be possible to extend this 
+admission procedure to check other constraints, such as whether the estimated 
+critical path of the new workflow exceeds the time remaining until the deadline.
 
@@
 
-We assume that the workflows are very wide such that the critical path of any given workflow is much less than the deadline.
+The previous algorithms are all online algorithms that make provisioning 
+and scheduling decisions at runtime. As such, they are not able to analyze 
+alternative decisions in order to optimize the schedules.
 
-The approach is to schedule each workflow in the ensemble in priority order, rejecting any workflows that cannot be completed by the deadline or that cause the cost of the plan to exceed the budget.
-Once the plan is complete, then the workflows are executed according to the plan.
+The SPSS algorithm creates a provisioning and scheduling plan before running 
+any workflow tasks.
+
+The approach is to schedule each workflow in the ensemble in priority order, 
+rejecting any workflows that cannot be completed by the deadline or that cause 
+the cost of the plan to exceed the budget. Once the plan is complete, then the 
+workflows are executed according to the plan.
+
+The disadvantage of the static planning approach used by SPSS is that it is 
+sensitive to dynamic changes in the environment and the application that may 
+disrupt the desired plan.
+
+Algorithm~\ref{alg:admit} shows how ensembles are planned in SPSS.
 
 \begin{algorithm}
-\caption{Ensemble planning algorithm}
+\caption{Ensemble planning algorithm for SPSS}
 \label{alg:admit}
@@
 
+The intuition behind this algorithm is that more workflows can be admitted
+if each workflow is able to achieve the lowest-cost schedule that still
+meets the ensemble deadline.
+
+In order to create a plan for a single workflow, the SPSS algorithm assigns 
+sub-deadlines to each individual task in the workflow, and then schedules 
+each task so as to minimize the cost of the task while still meeting its 
+assigned sub-deadline. The idea is that if each task can be completed by 
+its deadline in the least expensive manner possible, then the cost of the
+entire workflow is being minimized without exceeding the ensemble deadline.
+
+SPSS assigns sub-deadlines to each task based on the \emph{float time}
+of the workflow. The float time is the amount of extra time that a workflow
+can extend its critical path and still be completed by the ensemble deadline.
+For a workflow $w$, the float time of $w$ is:
+$$
+FT(w) = d - CP(w)
+$$
+where $d$ is the ensemble deadline and $CP(w)$ is the critical path of $w$.
+Note that we assume that $CP(w) \leq d$, otherwise the workflow cannot be
+completed by the deadline and must be rejected.
+
+SPSS distributes the float time of the workflow by level, so that each level
+of the workflow gets a portion of the workflow's float time proportional to
+the number of tasks in the level, and the total runtime of tasks in the level.
+The idea is that workflows with many tasks and large runtimes should be given
+a larger portion of the float time so that those levels may be serialized.
+Otherwise, many resources need to be allocated to run all of the tasks in 
+parallel, which may be more costly.
+
+The float time of a level $l$ in workflow $w$ is given by:
+
+$$
+FT(l) = FT(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
+$$
+
+where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks 
+in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is
+the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between
+0 and 1 that causes more float time to be given to levels with more tasks (large 
+$\alpha$) or more runtime (small $\alpha$). We set $\alpha$ to be around .7.
+
+The deadline of a task $t$ is then:
+$$
+DL(t) = LST(t) + RT(t) + FT(Level(t))
+$$
+where $Level(t)$ is the level of task t, $RT(t)$ is the runtime of $t$, and $LST(t)$ 
+is the latest start time of $t$ determined by:
+$$
+LST(t) = \begin{cases}
+0,&$if t is a root task$\cr
+\max_{p \in Pred(t)} DL(p),&$otherwise.$\cr
+\end{cases}
+$$
+
+For each task in the workflow the \textsc{PlanDAG} procedure chooses the 
+least expensive scheduling solution that is able to complete the task by 
+its deadline. The algorithm allocates VMs in blocks of one billing cycle 
+(one hour) regardless of the size of the task. When computing the cost 
+of scheduling a task on a given VM it considers idle time from blocks that 
+were allocated for previous tasks to be free, while new blocks cost full
+price. For example, if a task has a runtime of 10 minutes and the price of
+a one-hour billing cycle is \$1, then the algorithm will either schedule 
+the task on an existing VM that has 10 minutes of idle time for a cost of 
+\$0, or it will allocate another block on an existing or new VM for a cost
+of \$1. In addition, the algorithm always prefers to extend the runtime of 
+existing VMs rather than allocating new VMs, and if the cost of scheduling 
+the task on two different VMs is equal, then the solution with the earliest 
+start time is preferred. The result of this is that the algorithm will only
+allocate a new block if there are no existing blocks with enough idle time
+to complete the task by its deadline, and it will only allocate a new VM if
+it cannot add a block to the start or end of an existing VM to complete the
+task by its deadline.
+
 \begin{algorithm}
-\caption{DAG planning algorithm}
+\caption{DAG planning algorithm for SPSS}
 \label{alg:plandag}
@@
 
-\begin{algorithm}
-\caption{Deadline distribution algorithm}
-\label{alg:deadlinedistribution}
-\begin{algorithmic}[1]
-\Require $w$: DAG; $d$: global deadline
-\Ensure Assign sub-deadlines to each task in w
-\Procedure{DeadlineDistribution}{$w,d$}
-    \State $T \gets$ total number of tasks in $w$
-    \State $R \gets$ total runtime of tasks in $w$
-    \State $CP \gets$ length of critical path of $w$
-    \State assign $Level(t)$ to each task $t$ in $w$
-    \For{level $l\ \in w$}
-        \State $Tasks(l) \gets$ number of tasks in $l$
-        \State $Runtime(l) \gets$ runtime of tasks in $l$
-        \State $r \gets \alpha * (Runtime(l)/R)$
-        \State $t \gets (1-\alpha) * (Tasks(l)/T)$
-        \State $FT(l) \gets (r + t) * (d - CP)$
-    \EndFor
-    \For{task $t \in w$}
-        \State $LST(t) \gets \max(Deadline(p))\ for\ p \in Pred(t)$
-        \State $Deadline(t) \gets LST(t) + RT(t) + FT(Level(t))$
-    \EndFor
-\EndProcedure
-\end{algorithmic} 
-\end{algorithm}
-
-\begin{equation}
-\label{eq:scaling}
-FT(l) = \left[{\alpha}*\frac{T_l}{t}\right] + \left[{(1 - \alpha)}*\frac{R_l}{r} \right]
-\end{equation}
+An example schedule generated by SPSS is shown in Fig.~\ref{fig:spss-example}.
+This example shows how SPSS tends to start many workflows in parallel, running
+each workflow on only a few VMs over a longer period of time. In comparison, 
+the dynamic algorithms tend to run one workflow at a time across many VMs.
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/spss-gantt}
- \caption{Example of schedule generated using SPSS algorithm: tasks are labelled
- by colors depending on the workflow. }
+ \caption{Example of schedule generated using SPSS algorithm. Tasks are colored by workflow. }
 \label{fig:spss-example}

### 2012-01-20 paper/paper.tex
Update SPSS description

@@
 
-The simplest strategy for executing ensembles of workflows in the cloud is to use
+The simplest strategy for executing ensembles of workflows in the cloud is to
 provision resources statically and schedule workflow tasks dynamically. Given a 
-budget $b$, deadline $d$, and the hourly price $p$ of a VM it is easy to calculate
+budget $b$, deadline $d$, and the hourly price $p$ of a VM, it is easy to calculate
 the number of VMs, $N_{VM}$, to provision so that the entire budget is consumed before
@@
 
-The SPDS algorithm statically provisions VMs at the start of the ensemble execution 
+The SPDS algorithm statically provisions $N_{VM}$ VMs at the start of the ensemble execution 
 and keeps them running until the deadline is reached, or the budget runs out.
@@
 
-\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
+\subsection{Dynamic Provisioning Dynamic \\Scheduling (DPDS)}
 
@@
 
-\subsection{Static Provisioning Static Scheduling (SPSS)}
+\subsection{Static Provisioning Static Scheduling \\(SPSS)}
 
 The previous algorithms are all online algorithms that make provisioning 
-and scheduling decisions at runtime. As such, they are not able to analyze 
-alternative decisions in order to optimize the schedules.
+and scheduling decisions at runtime. In comparison, the SPSS algorithm 
+creates a provisioning and scheduling plan before running any workflow tasks.
+This enables SPSS to start only those workflows that it knows can be completed
+given the deadline and budget constraints, and eliminates any waste that may
+be allowed by the dynamic algorithms.
 
-The SPSS algorithm creates a provisioning and scheduling plan before running 
-any workflow tasks.
-
-The approach is to schedule each workflow in the ensemble in priority order, 
-rejecting any workflows that cannot be completed by the deadline or that cause 
-the cost of the plan to exceed the budget. Once the plan is complete, then the 
-workflows are executed according to the plan.
+The approach used by SPSS is to plan each workflow in the ensemble in priority 
+order, rejecting any workflows that cannot be completed by the deadline or that 
+cause the cost of the plan to exceed the budget. Once the plan is complete, then 
+the VMs are provisioned and tasks are executed according to the schedules given
+by the plan.
 
@@
 sensitive to dynamic changes in the environment and the application that may 
-disrupt the desired plan.
+disrupt the carefully constructed plan.
 
-Algorithm~\ref{alg:admit} shows how ensembles are planned in SPSS.
+Algorithm~\ref{alg:admit} shows how ensembles are planned in SPSS. The algorithm
+starts with an empty plan containing no VMs and no scheduled tasks. Workflows
+from the ensemble are considered in priority order. For each workflow, SPSS
+attempts to build on top of the current plan by provisioning VMs to schedule 
+the tasks of the workflow so that it finishes before the deadline with the least
+possible cost. If the cost of the new plan is less than the budget, then the 
+new plan is accepted and the workflow is admitted. If not, then the new plan is 
+rejected and the process continues with the next workflow in the ensemble. The 
+idea behind this algorithm is that, if each workflow can be completed by the 
+deadline with the lowest possible cost, then the number of workflows that can 
+be completed within the given budget will be maximized.
 
@@
 \Procedure{PlanEnsemble}{$W,b,d$}
-    \State $P\gets empty\ plan$
+    \State $P\gets \emptyset$ \Comment{Current plan}
     \State $A\gets \emptyset$ \Comment{Set of admitted DAGs}
     \For{$w\ in\ W$}
-        \State $P^\prime \gets$\ \Call{PlanDAG}{$w,P,d$}
-        \If{$P^\prime\ is\ a\ valid\ plan$}
+        \State $P^\prime \gets$\ \Call{PlanWorkflow}{$w,P,d$}
         \If{$Cost(P^\prime) \le b$}
@@
             \State $A \gets A\ +\ w$ \Comment{Admit w}
-            \EndIf
         \EndIf
@@
 
-The intuition behind this algorithm is that more workflows can be admitted
-if each workflow is able to achieve the lowest-cost schedule that still
-meets the ensemble deadline.
-
-In order to create a plan for a single workflow, the SPSS algorithm assigns 
+In order to plan a workflow, the SPSS algorithm assigns 
 sub-deadlines to each individual task in the workflow, and then schedules 
@@
 its deadline in the least expensive manner possible, then the cost of the
-entire workflow is being minimized without exceeding the ensemble deadline.
+entire workflow can be minimized without exceeding the ensemble deadline.
 
@@
 For a workflow $w$, the float time of $w$ is:
+
 $$
@@
 $$
+
 where $d$ is the ensemble deadline and $CP(w)$ is the critical path of $w$.
-Note that we assume that $CP(w) \leq d$, otherwise the workflow cannot be
-completed by the deadline and must be rejected.
+We assume that $CP(w) \leq d$, otherwise the workflow cannot be completed 
+by the deadline and must be rejected. For large ensembles we expect the 
+critical path of any individual workflow to be much less than the deadline.
 
@@
 of the workflow gets a portion of the workflow's float time proportional to
-the number of tasks in the level, and the total runtime of tasks in the level.
-The idea is that workflows with many tasks and large runtimes should be given
-a larger portion of the float time so that those levels may be serialized.
+the number of tasks in the level and the total runtime of tasks in the level.
+The idea is that levels containing many tasks and large runtimes should be given
+a larger portion of the float time so that tasks in those levels may be serialized.
 Otherwise, many resources need to be allocated to run all of the tasks in 
@@
 0 and 1 that causes more float time to be given to levels with more tasks (large 
-$\alpha$) or more runtime (small $\alpha$). We set $\alpha$ to be around .7.
+$\alpha$) or more runtime (small $\alpha$). We set $\alpha$ to be 0.7.
 
 The deadline of a task $t$ is then:
+
 $$
@@
 $$
-where $Level(t)$ is the level of task t, $RT(t)$ is the runtime of $t$, and $LST(t)$ 
+
+where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and $LST(t)$ 
 is the latest start time of $t$ determined by:
+
 $$
@@
 
-For each task in the workflow the \textsc{PlanDAG} procedure chooses the 
-least expensive scheduling solution that is able to complete the task by 
-its deadline. The algorithm allocates VMs in blocks of one billing cycle 
-(one hour) regardless of the size of the task. When computing the cost 
-of scheduling a task on a given VM it considers idle time from blocks that 
-were allocated for previous tasks to be free, while new blocks cost full
-price. For example, if a task has a runtime of 10 minutes and the price of
-a one-hour billing cycle is \$1, then the algorithm will either schedule 
-the task on an existing VM that has 10 minutes of idle time for a cost of 
-\$0, or it will allocate another block on an existing or new VM for a cost
-of \$1. In addition, the algorithm always prefers to extend the runtime of 
-existing VMs rather than allocating new VMs, and if the cost of scheduling 
-the task on two different VMs is equal, then the solution with the earliest 
-start time is preferred. The result of this is that the algorithm will only
-allocate a new block if there are no existing blocks with enough idle time
-to complete the task by its deadline, and it will only allocate a new VM if
-it cannot add a block to the start or end of an existing VM to complete the
-task by its deadline.
+Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for
+each workflow. The \textsc{PlanWorkflow} procedure first calls 
+\textsc{DeadlineDistribution} to assign sub-deadlines to tasks according
+to the procedure described above. Then, the \textsc{PlanWorkflow} procedure
+schedules tasks onto VMs, allocating new VMs when necessary. For each task 
+in the workflow, the the least expensive slot is chosen to schedule the 
+task so that it can be completed by  its deadline. VMs are allocated in 
+blocks of one billing cycle (one hour) regardless of the size of the task. 
+When computing the cost of scheduling a task on a given VM, the algorithm 
+considers idle slots in blocks that were allocated for previous tasks to be 
+free, while slots in new blocks cost the full price of the new block. For 
+example, if a task has a runtime of 10 minutes, and the price of a one-hour 
+billing cycle is \$1, then the algorithm will either schedule the task on 
+an existing VM that has an idle slot larger than 10 minutes for a cost of \$0, 
+or it will allocate a new block on an existing VM, or provision a new VM, for 
+a cost of \$1. If the cost of slots on two different VMs is equal, then the 
+slot with the earliest start time is chosen. To prevent too many VMs from 
+being provisioned, the algorithm always prefers to extend the runtime of 
+existing VMs before allocating new VMs. The result of this is that the 
+algorithm will only allocate a new block if there are no idle slots on 
+existing blocks large enough or early enough to complete the 
+task by its deadline, and it will only allocate a new VM if it cannot add 
+a block to an existing VM to complete the task by its deadline.
 
 \begin{algorithm}
-\caption{DAG planning algorithm for SPSS}
-\label{alg:plandag}
+\caption{Workflow planning algorithm for SPSS}
+\label{alg:planworkflow}
 \begin{algorithmic}[1]
-\Require $w$: DAG; $P$: current plan; $d$: deadline
+\Require $w$: workflow; $P$: current plan; $d$: deadline
 \Ensure Create plan for $w$ that minimizes cost and meets deadline $d$
-\Procedure{PlanDAG}{$w,P,d$}
+\Procedure{PlanWorkflow}{$w,P,d$}
     \State $P^\prime\gets$ copy of $P$
     \State \Call{DeadlineDistribution}{w,d}
-    \For{$t\ in\ w\ sorted\ by\ Deadline(t)$}
-        \State Choose resource r in $P^\prime$ that minimizes Cost(t)
-        \If{$AFT(t,r) < Deadline(t)$}
-            \State Schedule(t,r)
+    \For{$t\ in\ w\ sorted\ by\ DL(t)$}
+        \State $v \gets$ VM that minimizes cost and start time of t
+        \If{$FinishTime(t,v) < DL(t)$}
+            \State Schedule(t,v)
         \Else
-            \State Provision a new resource r
-            \State Schedule(t,r)
+            \State Provision a new VM v
+            \State Schedule(t,v)
         \EndIf
@@
 This example shows how SPSS tends to start many workflows in parallel, running
-each workflow on only a few VMs over a longer period of time. In comparison, 
-the dynamic algorithms tend to run one workflow at a time across many VMs.
+each workflow over a longer period of time on only a few VMs to minimize cost. 
+In comparison, the dynamic algorithms tend to run one workflow at a time across 
+many VMs in parallel.
 

### 2012-01-20 paper/paper.tex
Some minor updates to the introduction

@@
 resource management and scheduling on grid and utility computing systems.
-However, large computational applications are not just individual workflows but
-rather sets of inter-related workflows grouped in {\em ensembles}. All the
+However, large computational applications are often composed of several 
+inter-related workflows grouped into {\em ensembles}. All the
 workflows in an ensemble typically have a similar structure, but they differ by
@@
 the points on the map correspond to urban areas or strategic locations such as
-power plants, whereas others may be located in low populated regions. Another
-examples of workflow ensembles come from astronomical applications, such as
+power plants, whereas others may be located in less populated regions. Another
+example of workflow ensembles comes from astronomical applications, such as
 Montage~\cite{Deelman08} or search for Earth-like planets from Kepler project
-data~\cite{vockler11}, where multiple workflows are typically required to
-process data covering different parts of the sky.
+data~\cite{vockler11}, where multiple, different workflows are used to
+process data for different areas of the sky.
 
-Infrastructure-as-a-service (Iaas) clouds offer capabilities of creating
-(provisioning) computational resources on demand in a pay-per-use model and are regarded by
-scientific community as a potentially attractive source of computing
-resources~\cite{Ostermann10,Keahey09}. In contrast to clusters and grids which
-typically offer best-effort resource provisioning capabilities, clouds give more
-flexibility in terms of creating a controlled and managed computing environment
-with the ability of adjusting resource capacity to the computing demands, often
-called auto-scaling. Giving the users more control, however, clouds also require
-developing new methods of task scheduling and resource provisioning. The
-resource management decisions in cloud scenarios not only have to take into
-account performance related metrics such as workflow makespan or resource
-utilization, but also budget considerations, since the resources from public
-(commercial) cloud providers are usually not free~\cite{Durkee10}.
+Infrastructure-as-a-Service (IaaS) clouds offer the ability to provision 
+computational resources on-demand according to a pay-per-use model. These systems 
+are regarded by the scientific community as a potentially attractive source of 
+low-cost, on-demand computing resources~\cite{Ostermann10,Keahey09}. In contrast 
+to clusters and grids, which typically offer best-effort quality of service, clouds 
+give users more flexibility to creating a controlled and managed computing environment
+with the ability to adjust resource capacity according to the dynamically changing
+computing demands of the application, often called auto-scaling. Giving the users 
+more control, however, also requires the development of new methods for task 
+scheduling and resource provisioning. The resource management decisions required 
+in cloud scenarios not only have to take into account traditional performance-related 
+metrics such as workflow makespan or resource utilization, but must also now consider
+budget constraints, since the resources from public (commercial) cloud providers 
+are usually not free~\cite{Durkee10}.
 
 In this paper, to get insight into these resource management challenges for
-scietific workflow on clouds, we address a new important problem of maximizing
+scientific workflows on clouds, we address a new important problem of maximizing
 the number of completed workflows from an ensemble under both budget and
 deadline constraints. The motivation is to answer the fundamental question of a
-researcher: how much work can be completed in a limited budget and timeframe of
-a research project. The goals of this paper are to discuss and assess possible
+researcher: How much computation can be completed given the limited budget and 
+timeframe of a research project? The goals of this paper are to discuss and assess possible
 static and dynamic strategies for both task scheduling and resource
@@
 		\State $u\gets$ current VM utilization
-    	\If{$u>u_h$ and $|V_R| < v_{max}$}
+    	\If{$u>u_h$ and $|V_R| < v_{max}*N_{VM}$}
     		\State \Call{Start}{$new\ VM$}
@@
 instances, which may happen in the case of highly parallel workflows, the provisioner 
-will not start a new VM if the user-defined scaling parameter, $v_{max}$, has
-been reached.
+will not start a new VM if the number of running VMs is greater than the product of
+$N_{VM}$ and an autoscaling factor, $v_{max}$.
 
@@
 power consumption in a datacenter, we do not use these features, since we are
-interested mainly in execution of tasks on VMs and high-level worklow scheduling
+interested mainly in execution of tasks on VMs and high-level workflow scheduling
 on top of that. The simulator relies on DAG and DAX format used by Pegasus
@@
 
-In order to evaluate the algorithms on a standard ser of workflows, we used
+In order to evaluate the algorithms on a standard set of workflows, we used
 ensembles constructed from the DAGs available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
 in the workflow generator gallery~\cite{Bharathi08}. The gallery contains
-workflows of five applications from Pegasus and from Indiana
-University~\cite{Ramakrishnan08}. We constructed ensembles of 100 DAGs by sampling the
+workflows of five applications from Pegasus as well as several other applications 
+from~\cite{Ramakrishnan08}. We constructed ensembles of 100 DAGs by sampling the
 workflows of different sizes based on a given distribution. By the size of the
@@
 tasks, so they require budgets up to \$1,500 and deadlines extended to 40
-hours. Epigenomics workflow is even more costly and needs approximately \$12,000
+hours. The Epigenomics workflow is even more costly and needs approximately \$12,000
 and 1500 hours to complete the full ensemble.
 
-We show the results of three algorithms (TODO: ADD MORE): DPDS, WA-DPDS and
+We show the results for three algorithms: DPDS, WA-DPDS and
 SPSS. These tests were run with maximum autoscaling factor set to 1.0 for

### 2012-01-20 paper/paper.tex
Changes to section 3

@@
 \section{Problem Description}
+
 \subsection{Resource Model}
-Describe the Amazon EC2 resource model.
-Instances are requested on-demand.
-Instance pricing is per-hour.
-Multiple VM types are available, but for this paper we focus on a single VM type because we assume that there will typically be one VM type with the best price/performance ratio for the application.
+
+We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2).
+In this model, virtual machine (VM) instances are requested on-demand.
+The VMs are billed by the hour, with partial hours being rounded up.
+Although there are multiple VM types with different amounts of CPU, memory, disk space, and I/O, for this paper we focus on a single VM type because we assume that for most applications there will typically be one VM type with the best price/performance ratio for the application.
 
 \subsection{Application Model}
-The target applications for this paper are scientific workflows that can be modeled as Directed Acyclic Graphs (DAGs).
+The target applications for this paper are scientific workflows that can be modeled as Directed Acyclic Graphs (DAGs), where the nodes in the graph represent computational tasks, and the edges represent data- or control-flow dependencies between the tasks.
 
-We assume that we have runtime estimates for each task in the workflow.
+We assume that we have runtime estimates for each task in the workflow based on either a performance model of the application, or historical data that can be mined.
 
-Currently we do not consider data sizes, but rather assume that data is stored in a shared cloud storage system such as Amazon S3 and that data transfer times are included in task runtime estimates.
-Further, data transfer times are equal across the VMs.
+Although scientific workflows are often data-intensive, we do not currently consider the size of input and output data when scheduling tasks.
+Instead we assume that all workflow data is stored in a shared cloud storage system, such as Amazon S3, and that data transfer times are included in task runtime estimates.
+We also assume that data transfer times between the shared storage and the VMs are equal for different VMs so that task placement decisions do not have an impact on the runtime of the tasks.
 
-An ensemble consists of several related workflows. Each workflow is given a priority.
+This paper focuses on scheduling and provisioning for workflow ensembles.
+An ensemble consists of several related workflows that a user wishes to run.
+Each workflow in an ensemble is given a priority that indicates how important the workflow is to the user.
 
-The goal is to complete as many workflows as possible given a budget and a deadline.
+The goal of this workflow ensemble scheduling and cloud provisioning problem is to complete as many high-priority workflows as possible given a fixed budget and deadline.
 

### 2012-01-20 paper/paper.tex
updated results section

@@
 
-The paper is organized as follows [\ldots]
+The paper is organized as follows: after analysis of related work in
+Section~\ref{sec:related}, Section~\ref{sec:problem} gives the details about the
+infrastructure and application model we are targeting in this paper.
+Subsequently, in Section~\ref{sec:algorithms} we describe the dynamic and
+static algorithms we developed. Section~\ref{sec:performance} presents test
+scenarios, metrics and results of simulation studies of these algorithms.
+Finally, general conslusions, lessons learned and future work is outlined in
+Section~\ref{sec:conclusions}.
 
 \section{Related Work}
+\label{sec:related}
 General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
@@
 \section{Problem Description}
+\label{sec:problem}
 
@@
 \section{Algorithms}
+\label{sec:algorithms}
 
@@
 \section{Performance Evaluation}
+\label{sec:performance}
 
@@
 
-\begin{figure}[htb] 
-\centering
-\includegraphics[width=0.6\columnwidth]{figures/pareto-budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}
-\caption{Comparison of number of workflows completed for a given deadline,
-plotted versus different budget amounts.}
-\label{fig:budgets}
-\end{figure}
-
 \begin{table}[tb]
@@
 
-\label{tab:num-dags-pareto}
-\caption{Number of times each algorithm achieved the highest score 
-over all budgets and deadlines. Data from a total of 525 runs on 
+\caption{Summary table for Pareto-distributed workflows showing number of times
+each algorithm achieved the highest score over all budgets and deadlines. Data from a total of 525 runs on 
 Pareto-distributed ensembles for Montage (M), CyberShake (C), LIGO (L), 
-Epigenomics (E), SIPHT (S).}
+Epigenomics (E), SIPHT (S).
+\label{tab:num-dags-pareto}}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[tb]
+\centering
+
+(a) Number of workflows completed $N_c$:
+\medskip
+\begin{tabular}{r|cccccc}
+ & M & C & L & E & S & All\tabularnewline
+\hline
+DPDS      &   78 &  93 & 161 &  24  & 30 & 386\tabularnewline
+WA-DPDS   &    78  & 97 & 179  & 58  & 28 & 440\tabularnewline
+SPSS     &    78  & 79 & 195 &  75 &  50 & 477\tabularnewline
+\end{tabular}
+\medskip
+
+(b) Total amount of computing time completed $T_C$:
+\medskip
+\begin{tabular}{r|cccccc}
+ & M & C & L & E & S & All\tabularnewline
+\hline
+DPDS      &   78 &  93 & 161 &  24 &  30 & 386\tabularnewline
+WA-DPDS   &    78  & 97 & 179 &  46  & 28 & 428\tabularnewline
+SPSS     &    78 &  77 & 195  & 75 &  50 & 475\tabularnewline
+\end{tabular}
+\medskip
+
+(c) Exponential score $s$:
+\medskip
+\begin{tabular}{r|cccccc}
+ & M & C & L & E & S & All\tabularnewline
+\hline
+DPDS      &   78  & 93 & 161 &  24 &  30 & 386\tabularnewline
+WA-DPDS   &    78 &  97 & 179  & 46  & 28 & 428\tabularnewline
+SPSS     &    78 &  77 & 195 &  75  & 50 & 475\tabularnewline
+\end{tabular}
+\medskip
+
+\caption{Summary table for equal size workflows whowing number of times each
+algorithm achieved the highest score over all budgets and deadlines. Data from a total of 525 runs on 
+ensembles of equal size workflows for Montage (M), CyberShake (C), LIGO (L), 
+Epigenomics (E), SIPHT (S).
+\label{tab:num-dags-equal}}
 \end{table}
@@
 
+\begin{figure}[htb] \centering
+\includegraphics[width=0.7\columnwidth]{figures/pareto-budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}\\
+(a) ensemble of Pareto-distributed workflows
+\includegraphics[width=0.7\columnwidth]{figures/constant-budget-SIPHT-n-1000-0-dagh5-50h4m0.pdf}\\
+(b) ensemble of workflows of equal size \caption{Comparison of number of
+workflows completed for a given deadline, plotted versus different budget
+amounts.}
+\label{fig:budgets}
+\end{figure}
+
 In order to compare efficiency of algorithms, we counted the number of times
@@
 and budgets. If two or more algorithms achieve the same highest results, all are
-counted as highest. Table~\ref{tab:num-dags-pareto} shows these summary values.
-These results depend on the metric: number of workflows completed does not count
-the priorities or sizes of workflows, which means that it happens more often
-that all the algorithms achieve the same (best) result. However, it does not
-mean that all algorithms complete the same workflows: that is why values
-in the tables (a), (b) and (c) are different. it can be observed that based on
-the sampled problem space the workflow-aware algorithms achieve highest score
-in $\sim$2--3 times more cases than the simple DPDS which does not take into
-account workflow structure or size. In average, SPSS performs better in
-$\sim$20--50\% more times than WA-DPDS. We have to note, that the sampled
-paramter space included the Montage and CyberShake workflows consisting of many
-small tasks, for which cases SPSS does not perform equally well.
-
-Fig.~\ref{fig:budgets} whows the same data for SIPHT ensemble from another
+counted as highest. Table~\ref{tab:num-dags-pareto} shows these summary values
+for Pareto-distributed workflows. These results depend on the metric: number of
+workflows completed does not count the priorities or sizes of workflows, which
+means that it happens more often that all the algorithms achieve the same (best)
+result. However, it does not mean that all algorithms complete the same
+workflows: that is why values in the tables (a), (b) and (c) are different.
+Exponential score and total amount of computing time complete give the same
+rankings, since in our test ensembles larger workflows were given higher
+priorities. It can be observed that based on the sampled problem space the
+workflow-aware algorithms achieve highest score in $\sim$2--3 times more cases
+than the simple DPDS which does not take into account workflow structure or
+size. In average, SPSS performs better in $\sim$20--50\% more times than
+WA-DPDS. We have to note, that the sampled paramter space included the Montage
+and CyberShake workflows consisting of many small tasks, for which cases SPSS
+does not perform equally well.
+
+Similar results were achieved for ensembles consisting of workflows of equal
+size. Summary of these results is shown in Table~\ref{tab:num-dags-equal}.
+Again, we can see that static SPSS algorithm performed better for coarse-grained
+workflows (LIGO, Epigenomics, SIPHT). The differences between algorithm scores
+are smaller since the workflows have equal size thus rejecting one workflow
+does not help to complete more work in any case.
+
+Fig.~\ref{fig:budgets} shows the same data for SIPHT ensemble from another
 perspective. It helps answer the question: given a deadline, how many workflows
-can be finished when we increse the budget. In Fig.~\ref{fig:budgets} the growth
-is super-linear, what results from the distribution of workflow sizes in the
-ensemble (see Fig.~\ref{fig:ensemble-distribution}). For ensemble of equal size
-workflows (see Fig. TODO: ADD FIGURE FROM CONSTANT DISTRIBUTION) the growth is
-nearly linear. It can be also observed that staitc algorithms give better
-results for smaller budgets, while dynamic algorithms for larger ones.
+can be finished when we increse the budget. In Fig.~\ref{fig:budgets}(a) the
+growth is super-linear, what results from the distribution of workflow sizes in
+the ensemble (see Fig.~\ref{fig:ensemble-distribution}). For ensemble of equal
+size workflows (see Fig.~\ref{fig:budgets}(b)) the growth is nearly linear. It
+can be also observed that staitc algorithms give better results for smaller
+budgets, while dynamic algorithms for larger ones.
 
 \section{Conclusion and Future Work}
+\label{sec:conclusions}
 

### 2012-01-20 paper/paper.tex
Spelling corrections

@@
 low-priority workflows first, thus increasing the total number of workflows
-complete. It may also happen that too conservative admission algoritm may reject
-overestimate the costs of some large workflows, thus rejest some high-priority
+complete. It may also happen that too conservative admission algorithm may reject
+overestimate the costs of some large workflows, thus reject some high-priority
 ones which actually could fit into the schedule. Therefore we defined a user
@@
 ca. 50-150 (budget, deadline) points for each application. The runs were
-repeated for Pareto-distributed and costant workflow sizes. From this result
+repeated for Pareto-distributed and constant workflow sizes. From this result
 space we selected the most interesting plots which give the good overview  of
@@
 
-\caption{Summary table for equal size workflows whowing number of times each
+\caption{Summary table for equal size workflows showing number of times each
 algorithm achieved the highest score over all budgets and deadlines. Data from a total of 525 runs on 
@@
 For a given budget, when the deadline is tightened the number of work done
-steeply decreases. Assiging a shorter deadline means that more VMs need to be
+steeply decreases. Assigning a shorter deadline means that more VMs need to be
 provisioned to complete the work, but this leads to higher parallelism and thus
@@
 
-These observations are further confirmed when examining the efffective resource
+These observations are further confirmed when examining the effective resource
 cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
@@
 solutions to this problem. In our case the solutions were achieved by maximizing
-the number of work done, which direcly results in minimization of effective cost parameter for a given
-deadline. Therefore, the obtained results can be used to assist the decison in
+the number of work done, which directly results in minimization of effective cost parameter for a given
+deadline. Therefore, the obtained results can be used to assist the decision in
 resource allocation planning, when both cost and time criteria are given not as
@@
 size. In average, SPSS performs better in $\sim$20--50\% more times than
-WA-DPDS. We have to note, that the sampled paramter space included the Montage
+WA-DPDS. We have to note, that the sampled parameter space included the Montage
 and CyberShake workflows consisting of many small tasks, for which cases SPSS
@@
 perspective. It helps answer the question: given a deadline, how many workflows
-can be finished when we increse the budget. In Fig.~\ref{fig:budgets}(a) the
+can be finished when we increase the budget. In Fig.~\ref{fig:budgets}(a) the
 growth is super-linear, what results from the distribution of workflow sizes in
@@
 size workflows (see Fig.~\ref{fig:budgets}(b)) the growth is nearly linear. It
-can be also observed that staitc algorithms give better results for smaller
+can be also observed that static algorithms give better results for smaller
 budgets, while dynamic algorithms for larger ones.
@@
 dynamic adding and removing of resources at runtime and high parallelization due
-to ilusion of unlimited resources, it turns out that is recommended to avoid
+to illusion of unlimited resources, it turns out that is recommended to avoid
 using these capabilities when possible, in order to achieve high resource
 utilization and cost effectiveness. This observation comes from the fact that
-starting and stopping more VMs incurres a cost which can be amortized by running
+starting and stopping more VMs incurs a cost which can be amortized by running
 each started VM as long as possible before shutdown. This effect was confirmed
@@
 the question how much time and money is required to actually complete all the
-workflows from the given ansemble.
+workflows from the given ensemble.
 
@@
 question would be to answer how the static and dynamic algorithms perform in the
-cases of increased uncertainities and failures, that have important influence
+cases of increased uncertainties and failures, that have important influence
 on overall system performance~\cite{Sakellariou2010,Dongarra2007}. Second

### 2012-01-20 paper/paper.tex
Changes to sections 5.1 and 5.2

@@
 
-To evaluate our algorithms, we developed a cloud workflows simulator based on
-CloudSim~\cite{Calheiros11}. To represent our infrastructure model, we defined
-Cloud, VM and WorkflowEngine simulation entities. Cloud object is responsible
-for staring and terminating VMs using API similar to Amazon EC2. VM object is
-responsible for execution of individual tasks, while WorkflowEngine deals with
-management of ensembles and workflows. WorkflowEngine uses separate, although
-tightly coupled Scheduler and Provisioner modules. We assume that a single core
-of a VM executed submitted tasks sequentially and that a task is given exclusive
-access to allocated VM. Although CloudSim provides a more advanced
-infrastructure model which includes time-sharing and space-sharing policies and
-power consumption in a datacenter, we do not use these features, since we are
-interested mainly in execution of tasks on VMs and high-level workflow scheduling
-on top of that. The simulator relies on DAG and DAX format used by Pegasus
-project and workflow gallery.
+To evaluate and compare our algorithms, we developed a cloud workflow simulator based on
+CloudSim~\cite{Calheiros11}. Our simulation model consists of Cloud, VM and WorkflowEngine 
+entities. The Cloud entity is responsible for staring and terminating VM entities using 
+an API similar to Amazon EC2. VM entities simulate the execution of individual tasks,
+and the WorkflowEngine entity manages the scheduling of tasks and the provisioning of
+VMs. We assume that the VMs have a single core and execute tasks sequentially. Although 
+CloudSim provides a more advanced infrastructure model which includes time-sharing and 
+space-sharing policies, and power consumption in a datacenter, we do not use these 
+features, since we are interested mainly in the execution of tasks on VMs and high-level 
+workflow scheduling and provisioning. The simulator reads workflows in a modified version of
+the DAX format used by the Pegasus Workflow Management System~\cite{Deelman2005} that was
+created for the synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi08}.
 
@@
 
-We have prepared test scenarios to evaluate the proposed algorithms under
-such a range of parameters that allows to observe the interesting properties of
-algorithms for smaller and larger constraints. For a given ensemble we can
-expect in general that the number of work done (in terms of workflows completed)
-mainly depends on the budget constraint, which determines the number of
-machine-hours that can be provisioned. The ideal but trivial case is then a
-deadline long enough to allow running all the workflow tasks sequentially: this
+We have prepared test scenarios to evaluate the proposed algorithms under a 
+range of parameters that enables us to observe the interesting characteristics 
+of the different algorithms for smaller and larger constraints. For a given 
+ensemble we can expect, in general, that the amount of work done (in terms 
+of workflows completed) depends mainly on the budget constraint, which 
+determines the number of machine-hours that can be provisioned. The deadline, 
+although important in terms of scheduling, is assumed to be greater than the 
+critical path of every workflow. The ideal but trivial case, then, is a deadline 
+long enough to allow all the workflow tasks to be executed sequentially: this 
 eliminates all the overheads of parallel execution and leads to the highest 
-resource utilization. More interesting are the cases when the deadline is
-shorter, since it forces provisioning of multiple VMs and thus increases
-overheads due to parallel execution so the budget cannot be spent so
-efficiently.
+resource utilization. More interesting, however, are the cases where the deadline 
+is shorter than the serial schedule, since it forces the system to provision 
+multiple VMs to execute workflows in parallel.
 
@@
 
-We assumed that the interesting number of workflows in ensemble should be in the
-order of between 10 and 100 workflows. This is motivated by several reasons:
-first, such numbers are typical for the real applications, e.g. the number of geographical sites for
-CyberShake workflow is in the order of 100. Second, the smaller ensembles
-consisting of just a few workflows can be aggregated in a single workflow so
-there is no need to treat them as ensembles. Similarly, when the number of
-workflows is larger that 100 and each workflow has a large number of parallel
-tasks, the problem of efficiently allocating them to the resources becomes much
-easier and similar to the bag-of-tasks problem.
+The number of workflows in an ensemble depends on the user's requirements, but
+we assume that ensemble sizes are on the order of between 10 and 100 workflows.
+This range is motivated by several reasons: First, such sizes are typical of the
+real applications we have examined. For example, the number of geographical 
+sites of interest to the users of the CyberShake application is on the order 
+of 100. Second, smaller ensembles consisting of just a few workflows can be 
+aggregated into a single workflow, so there is no need to treat them as an
+ensemble. Similarly, when the number of workflows is larger than 100, and each 
+workflow has a large number of tasks, either the constraints are low enough to
+prevent many of the workflows from running, or the problem of efficiently 
+allocating them to the resources becomes similar to the bag-of-tasks problem, 
+which is easier to solve efficiently.
 
@@
 In order to evaluate the algorithms on a standard set of workflows, we used
-ensembles constructed from the DAGs available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
-in the workflow generator gallery~\cite{Bharathi08}. The gallery contains
-workflows of five applications from Pegasus as well as several other applications 
-from~\cite{Ramakrishnan08}. We constructed ensembles of 100 DAGs by sampling the
-workflows of different sizes based on a given distribution. By the size of the
-workflow means the number of jobs. Two distributions are considered:
+ensembles constructed from the workflows available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
+in the workflow generator gallery~\cite{Bharathi08}. The gallery contains synthetic
+workflows modeled on real applications. We used workflows from five applications 
+that were developed for the Pegasus Workflow Management System. We constructed 
+ensembles of 100 workflows by sampling collections of workflows of various sizes
+(number of tasks) based on a given distribution. Two distributions are considered:
+
 \begin{itemize}
-  \item ensemble of workflows of equal size (constant)
+  \item constant distribution
   \item Pareto-like distribution.
@@
 
-Pareto-like distribution gives a small number of workflows with large size and a
-large number of workflows with small size, as seen in
-Fig.~\ref{fig:ensemble-distribution}. The number of largest
-workflows (of size $\geq$ 900) is slightly increased to represent the
-``long-tail'', i.e. that there is a non-insignificant number of large workflows
-in the ensemble. The workflows are sorted according to their size and the largest workflows are
-given the highest priority.
+The constant distribution produces ensembles containing of workflows of equal size, 
+while the Pareto-like distribution produces ensembles with a small number of larger 
+workflows and a large number of smaller workflows, as seen in 
+Fig.~\ref{fig:ensemble-distribution}. The number of large workflows (of size $\geq$ 900) 
+is slightly increased to represent the ``heavy-tail'', i.e. that there is a 
+not-insignificant number of large workflows in the ensemble. We assume that large
+workflows are more important to users, so in the Pareto ensembles the workflows are 
+assigned priorities according to their size so that the largest workflows are given 
+the highest priority. In future work we plan to explore other ensemble types such as
+uniform distributions.
 
@@
 
-In the tests we we assumed that all the VMs are homogeneous and have the
-processing speed of 1000 MIPS (million instructions per second), the price for
-one VM-hour is \$1 and that the sizes of tasks in the workflow gallery are given
-in seconds measured on 1000 MIPS machines. In this study we do not take into
-account the heterogeneity of the infrastructure since we assume that it is
-always possible to select a VM type that has the best price to performance ratio
-for a given application.
-
-For each application type, we selected such ranges of constraints (deadline,
-budget) that it is possible to observe the possibly broad result space: from the
-tight constraints, when only a small number of workflows can finish to the
-liberal constraints when all or almost all the workflows can finish. These
-parameters vary among the applications. Montage and CyberShake benchmarks
-workflows required the budget in \$20 -- \$150 range and deadlines from 1 to
-20 hours to complete the ensemble. LIGO and SIPHT workflows have more long
-tasks, so they require budgets up to \$1,500 and deadlines extended to 40
-hours. The Epigenomics workflow is even more costly and needs approximately \$12,000
-and 1500 hours to complete the full ensemble.
-
-We show the results for three algorithms: DPDS, WA-DPDS and
-SPSS. These tests were run with maximum autoscaling factor set to 1.0 for
-DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that due to
-high parallelism of workflows the resource utilization remains high most of the time, 
-which leads to rapid provisioning of new resources at the beginning of ensemble
-execution. This in turn results in less efficient total utilization due to high
-level of parallelism. Setting the scaling limit to 1.0, ensures that the shape
-of resource provisioning plan is closest to a rectangle (see
-Fig.~\ref{fig:spds-example}) and the VM termination procedure
-(Algorithm~\ref{alg:prov}) allows maximum utilization of VMs before shutdown.
+In the experiments we assumed that all the VMs are homogeneous, have a
+processing speed of 1000 MIPS (million instructions per second), and a price of
+\$1 per VM-hour. We also assume that the sizes of tasks in the workflow gallery 
+are given in seconds measured on a 1000 MIPS machine. In this study we do not 
+take into account the heterogeneity of the infrastructure since we assume that 
+it is always possible to select a VM type that has the best price to performance 
+ratio for a given application.
+
+For each application, we selected ranges of constraints (deadline and budget) so
+that it is possible to observe the possibly broad result space: from tight 
+constraints, where only a small number of workflows can be completed, to 
+more liberal constraints where all, or almost all, of the workflows can be 
+completed. These parameters vary among the applications. Montage and CyberShake 
+ensembles required budgets in the \$20 -- \$150 range and deadlines from 1 to
+20 hours to be completed. The LIGO and SIPHT workflows have more long-running
+tasks, so their ensembles required budgets up to \$1,500 and deadlines up to 
+40 and 60 hours. The Epigenomics workflow is even more costly, requiring 
+approximately \$12,000 and 1500 hours to complete a full ensemble.
+
+We show the results for three algorithms: DPDS, WA-DPDS and SPSS. These 
+experiments were run with maximum autoscaling factor set to 1.0 for
+DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that, 
+due to the high parallelism of workflows, the resource utilization remains 
+high most of the time, which leads to the rapid provisioning of new resources 
+at the beginning of ensemble execution. This, in turn, results in less 
+efficient total utilization due to high levels of parallelism. Setting the 
+scaling limit to 1.0, ensures that the shape of the resource provisioning 
+plan  is close to a rectangle (see Fig.~\ref{fig:spds-example}) and the VM 
+termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
+of VMs before shutdown.
 

### 2012-01-20 paper/paper.tex
Changes to section 5.3

@@
 termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
-of VMs before shutdown.
+of VMs before shutdown. For the SPSS algorithm we set the $\alpha$ parameter
+for deadline distribution to be 0.7.
 
@@
 
-There are multiple ways to evaluate the algorithms and also different metrics
-are relevant to the application users that are needed to take the planning
-decisions. The set of metrics for each ensemble run we prepared includes:
+There are many different ways to evaluate and compare the performance of 
+the algorithms. The set of ensemble performance metrics we have used to
+compare our algorithms includes:
+
 \begin{itemize}
-  \item number of workflows completed $N_c$,
-  \item user preference score (utility function) $s$,
-  \item total amount of computing time completed $T_C$,
-  \item effective cost of computing per hour $C_{eff}$, which includes
+  \item $N_c$: number of workflows completed,
+  \item $s$: user preference score (utility function),
+  \item $T_C$: total amount of computing time completed,
+  \item $C_{eff}$: effective cost of computing per hour, which includes
   overheads.
@@
 
-The number of workflows completed is a simplest metric and allows distinguishing
-which algorithm performs better. However, as the workflows have
-different priorities and sizes, a less efficient algorithm may schedule smaller
-low-priority workflows first, thus increasing the total number of workflows
-complete. It may also happen that too conservative admission algorithm may reject
-overestimate the costs of some large workflows, thus reject some high-priority
-ones which actually could fit into the schedule. Therefore we defined a user
-preference score as:
-
-%\begin{equation}
-%\label{eq:score}
-%s = \sum_{DAGs\ completed}(1/(p+1))
-%\end{equation}
+The number of workflows completed is a simple metric that can be used for
+comparison. However, because our ensembles contain workflows with different 
+priorities and sizes, a less efficient algorithm may schedule smaller
+workflows first, completing a large number of low-priority workflows. It may 
+also happen that a conservative admission algorithm may overestimate the cost 
+of a large workflow, thus rejecting high-priority workflows which actually 
+could fit into the schedule. Therefore, we defined the user preference score as:
 
@@
 \label{eq:score}
-s = \sum_{DAGs\ completed}2^{-p}
+s = \sum_{workflows\ completed}{2^{-d}}
 \end{equation}
 
-where priorities of subsequent workflows in the ensemble are $p=0,1,2,\ldots$ (0
-means highest priority). This exponential scoring function gives a strong weight
-to the top priority workflows, since it has a property that a score for
-completing a single workflow with priority $p$ is higher that the score for
-completing all lower priority ones: 
+where $p$ is the priority of the workflow such that the highest-priority workflow 
+has $p=0$, the next highest workflow has $p=1$, and so on. This exponential scoring 
+function gives the highest priority workflow a score that is higher than all the 
+lower-priority workflows combined:
+
 \begin{equation}
@@
 
-Another metric is the total actual computing time completed and it is calculated
-as the sum of all the runtimes of all the tasks of completed workflows. As in
-our study we assumed that the resource cost pre hour is equal to \$1, the total
-computing time can be easily compared to the allocated budget $b$. Their ratio
-$u = (T_C * \$1)/b$ is the resource utilization and its inverse is the effective
-resource cost: $C_{eff} = b/T_C$ in dollars per hour.
+This scoring is consistent with our definition of the problem, which is to 
+complete as many workflows as possible, according to their priorities, given 
+a budget and deadline.
+
+The total amount computing time completed, $T_C$, is calculated as the sum 
+of the runtimes of all tasks in all completed workflows. Because we assumed 
+that the resource cost per hour is equal to \$1, the total computing time 
+can be easily compared to the allocated budget $b$. Their ratio: 
+$u = (T_C * \$1)/b$, is the resource utilization, and its inverse: 
+$C_{eff} = b/T_C$, is the effective resource cost in dollars per hour.
 

### 2012-01-20 paper/paper.tex
applied Ewa's comments

@@
 the important class of applications that have been studied in the context of
-resource management and scheduling on grid and utility computing systems.
+resource management and task scheduling on grid and utility computing systems.
 However, large computational applications are often composed of several 
@@
 workflows in an ensemble typically have a similar structure, but they differ by
-input data, number of tasks and individual tasks sizes. A good example of
+input data, number of tasks and individual task sizes. A good example of
 scientific workflow ensemble comes from the CyberShake
@@
 site requires running a single complex scientific workflow. Workflows for each
-site may differ not only by their parameters, but also by priority; e.g. some of
+site may differ not only in their parameters, but also in their priority; e.g. some of
 the points on the map correspond to urban areas or strategic locations such as
@@
 example of workflow ensembles comes from astronomical applications, such as
-Montage~\cite{Deelman08} or search for Earth-like planets from Kepler project
+Montage~\cite{Deelman08} or search for exo-solar planets from Kepler project
 data~\cite{vockler11}, where multiple, different workflows are used to
@@
 to clusters and grids, which typically offer best-effort quality of service, clouds 
-give users more flexibility to creating a controlled and managed computing environment
-with the ability to adjust resource capacity according to the dynamically changing
-computing demands of the application, often called auto-scaling. Giving the users 
-more control, however, also requires the development of new methods for task 
+give users more flexibility to creating a controlled and managed computing environment.
+Clouds provide the ability to adjust resource capacity according to the dynamically changing
+computing demands of the application, often called auto-scaling. However, giving users 
+more control over the resources also requires the development of new methods for task 
 scheduling and resource provisioning. The resource management decisions required 
 in cloud scenarios not only have to take into account traditional performance-related 
-metrics such as workflow makespan or resource utilization, but must also now consider
+metrics such as the workflow makespan or the resource utilization, but must also now consider
 budget constraints, since the resources from public (commercial) cloud providers 
-are usually not free~\cite{Durkee10}.
-
-In this paper, to get insight into these resource management challenges for
-scientific workflows on clouds, we address a new important problem of maximizing
+usually have monetary costs associated with them~\cite{Durkee10}.
+
+In this paper, we aim to gain insight into resource management challenges when executing 
+scientific workflow ensembles on clouds. In particular we address a new important problem of maximizing
 the number of completed workflows from an ensemble under both budget and
@@
 workflow structure information to estimate which workflows should be rejected
-from the ensemble due to the constraints. As methodology in the study of the
-proposed algorithms we use simulation techniques. We have developed cloud workflow simulator based on
+from the ensemble due to the constraints. We evaluate the algorithms via simulation. 
+We have developed cloud workflow simulator based on
 CloudSim~\cite{Calheiros11}, which models the infrastructure and the application. The
@@
 infrastructure and application model we are targeting in this paper.
-Subsequently, in Section~\ref{sec:algorithms} we describe the dynamic and
+In Section~\ref{sec:algorithms} we describe the dynamic and
 static algorithms we developed. Section~\ref{sec:performance} presents test
 scenarios, metrics and results of simulation studies of these algorithms.
-Finally, general conslusions, lessons learned and future work is outlined in
+Finally, general conclusions, lessons learned and future work are outlined in
 Section~\ref{sec:conclusions}.
@@
 of resource pool based on metrics related to infrastructure and application.
-Typical infrastructure-specific metric is system load, whereas
-application-specific metrics include response time and queue length. It is
-possible to set thresholds and limits to tune the behavior of these autoscaling
-systems, but no support for complex applications is provided.
+A typical infrastructure-specific metric is system load, whereas
+application-specific metrics include response time and length of task or
+request queue. It is possible to set thresholds and limits to tune the behavior
+of these autoscaling systems, but no support for complex applications is provided.
 
@@
 bags of independent tasks or unpredictable batch workloads. This enables us to
-take advantage of scheduling heuristics that cannot be applied to independent
-tasks.
+take advantage of workflow-aware scheduling heuristics that cannot be applied to
+independent tasks.
 
@@
 
-We should refer to bi-criteria scheduling and multi-criteria scheduling of
+This work is related to bi-criteria scheduling and multi-criteria scheduling of
 workflows~\cite{Wieczorek2009,Prodan10,Dongarra2007}. These approaches are
@@
 of work completed) is subject to optimization, whereas time and cost are
-trated as constraints. Other approaches~\cite{Talukder2009,Pandey2010} use
+treated as constraints. Other approaches~\cite{Talukder2009,Pandey2010} use
 metaheuristics that usually run for a long time before producing good results,
-which makes tham less useful in practical applications. Our work can be also
-regarded as extension of the budget-constrained workflow scheduling
+which makes them less useful in practice. Our work can be also
+regarded as an extension of the budget-constrained workflow scheduling
 \cite{Sakellariou2007} in the sense that we are dealing with workflow ensembles
@@
 
-We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2).
-In this model, virtual machine (VM) instances are requested on-demand.
-The VMs are billed by the hour, with partial hours being rounded up.
-Although there are multiple VM types with different amounts of CPU, memory, disk space, and I/O, for this paper we focus on a single VM type because we assume that for most applications there will typically be one VM type with the best price/performance ratio for the application.
+We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2). In
+this model, virtual machine (VM) instances are requested on-demand. The VMs are
+billed by the hour, with partial hours being rounded up. Although there are
+multiple VM types with different amounts of CPU, memory, disk space, and I/O,
+for this paper we focus on a single VM type because we assume that for most
+applications there will typically be one VM type with the best price/performance
+ratio for the application~\cite{Juve2009}. We also assume that a submitted task
+has exclusive access to a VM instance and that there is no preemption.
 
 \subsection{Application Model}
-The target applications for this paper are scientific workflows that can be modeled as Directed Acyclic Graphs (DAGs), where the nodes in the graph represent computational tasks, and the edges represent data- or control-flow dependencies between the tasks.
-
-We assume that we have runtime estimates for each task in the workflow based on either a performance model of the application, or historical data that can be mined.
-
-Although scientific workflows are often data-intensive, we do not currently consider the size of input and output data when scheduling tasks.
-Instead we assume that all workflow data is stored in a shared cloud storage system, such as Amazon S3, and that data transfer times are included in task runtime estimates.
-We also assume that data transfer times between the shared storage and the VMs are equal for different VMs so that task placement decisions do not have an impact on the runtime of the tasks.
-
-This paper focuses on scheduling and provisioning for workflow ensembles.
-An ensemble consists of several related workflows that a user wishes to run.
-Each workflow in an ensemble is given a priority that indicates how important the workflow is to the user.
-
-The goal of this workflow ensemble scheduling and cloud provisioning problem is to complete as many high-priority workflows as possible given a fixed budget and deadline.
+The target applications for this paper are scientific workflows that can be
+modeled as Directed Acyclic Graphs (DAGs), where the nodes in the graph
+represent computational tasks, and the edges represent data- or control-flow
+dependencies between the tasks.
+
+We assume that we have runtime estimates for each task in the workflow based on
+either a performance model of the application, or historical data that can be
+mined.
+
+Although scientific workflows are often data-intensive, we do not currently
+consider the size of input and output data when scheduling tasks. Instead we
+assume that all workflow data is stored in a shared cloud storage system, such
+as Amazon S3, and that data transfer times are included in task runtime
+estimates. We also assume that data transfer times between the shared storage
+and the VMs are equal for different VMs so that task placement decisions do not
+have an impact on the runtime of the tasks.
+
+This paper focuses on scheduling and provisioning for workflow ensembles. A
+workflow ensemble consists of several related workflows that a user wishes to
+run. Each workflow in an ensemble is given a priority that indicates how
+important the workflow is to the user.
+
+The goal of this workflow ensemble scheduling and cloud provisioning problem is
+to complete as many high-priority workflows as possible given a fixed budget and
+deadline.
 
@@
 \includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
- \caption{Example schedule generated using SPDS algorithm. Tasks are colored by workflow. }
+ \caption{Example schedule generated using SPDS algorithm. The colors indicate how various tasks from different workflows are scheduled onto Virtual Machines. }
 \label{fig:spds-example}
@@
 In policy-based provisioning (or auto-scaling) systems, metrics such as resource 
-utilization and queue length are monitored to estimate application demand. The DPDS 
-algorithm uses this approach to determine when to provision and de-provision VMs. The 
+utilization or task queue length are monitored to estimate application demand.
+The DPDS algorithm uses this approach to determine when to provision and de-provision VMs. The 
 algorithm periodically computes resource utilization as the percentage of idle VMs, and
@@
 that VMs can be terminated before they start the next billing cycle and prevents the
-budget from being overrun. The VMs terminated in line \ref{l:terminate1} are the ones 
+budget from being overrun. The VMs terminated in line \ref{l:terminate1} of Algorithm~\ref{alg:prov} are the ones 
 that would overrun the budget if not terminated in the current provisioning cycle. 
@@
 will not start a new VM if the number of running VMs is greater than the product of
-$N_{VM}$ and an autoscaling factor, $v_{max}$.
+$N_{VM}$ and an autoscaling factor, $v_{max}$ that has to be set manually.
 
@@
 tasks when deciding which task to schedule next. It does not consider whether a lower 
-priority task belongs to a workflow that it will never be able to complete given the 
+priority task belongs to a workflow that will never be able to complete given the 
 current budget and deadline. As a result, DPDS may execute lower priority tasks just 
-to keep VMs busy that will end up delaying higher priority tasks, making it less likely
+to keep VMs busy and that will end up delaying higher priority tasks, making it less likely
 that higher priority workflows will be able to finish.
@@
 priority queue (i.e. when no other tasks from the workflow have been scheduled yet). 
-The admission procedure---shown in Algorithm~\ref{alg:wa-dpds}---estimates if there is 
+The admission procedure---shown in Algorithm~\ref{alg:wa-dpds}---estimates whether there is 
 enough budget remaining to admit the new workflow; if there is not, then the 
@@
 into account the cost of currently running VMs, and the cost of workflows 
-that have already been admitted. It adds a small safety margin avoid going over
+that have already been admitted. It adds a small safety margin to  avoid going over
 the budget.
@@
 This admission procedure relies only on the total estimated resource consumption and
-compares it to the budget remaining. We found that this estimation is useful not
+compares it to the remaining budget. We found that this estimation is useful not
 only to prevent low-priority workflows from delaying high-priority ones, but
@@
 the VMs are provisioned and tasks are executed according to the schedules given
-by the plan.
+by the plan. However, in this paper we only consider a deterministic environment and task times.
 

### 2012-01-20 paper/paper.tex
Ewa's changes

--- 2012-01-20 paper/paper.tex
@@
 
-SPSS assigns sub-deadlines to each task based on the \emph{float time}
-of the workflow. The float time is the amount of extra time that a workflow
+SPSS assigns sub-deadlines to each task based on the slack time of the
+workflow, which is defined as the amount of extra time that a workflow
 can extend its critical path and still be completed by the ensemble deadline.
-For a workflow $w$, the float time of $w$ is:
+For a workflow $w$, the slack time of $w$ is:
 
 $$
-FT(w) = d - CP(w)
+ST(w) = d - CP(w)
 $$
@@
 
-SPSS distributes the float time of the workflow by level, so that each level
-of the workflow gets a portion of the workflow's float time proportional to
+A task's level is the length of the longest path between the task and an entry
+task of the workflow:
+
+$$
+Level(t) = \begin{cases}
+0,&$if~$Pred(t) = \emptyset\cr
+\max_{p \in Pred(t)}Level(p)+1,&$otherwise.$
+\end{cases}
+$$
+
+SPSS distributes the slack time of the workflow by level, so that each level
+of the workflow gets a portion of the workflow's slack time proportional to
 the number of tasks in the level and the total runtime of tasks in the level.
 The idea is that levels containing many tasks and large runtimes should be given
-a larger portion of the float time so that tasks in those levels may be serialized.
+a larger portion of the slack time so that tasks in those levels may be serialized.
 Otherwise, many resources need to be allocated to run all of the tasks in 
@@
 
-The float time of a level $l$ in workflow $w$ is given by:
+The slack time of a level $l$ in workflow $w$ is given by:
 
 $$
-FT(l) = FT(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
+ST(l) = ST(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
 $$
@@
 the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between
-0 and 1 that causes more float time to be given to levels with more tasks (large 
-$\alpha$) or more runtime (small $\alpha$). We set $\alpha$ to be 0.7.
+0 and 1 that causes more slack time to be given to levels with more tasks (large 
+$\alpha$) or more runtime (small $\alpha$).
 
@@
 $$
-DL(t) = LST(t) + RT(t) + FT(Level(t))
+DL(t) = LST(t) + RT(t) + ST(Level(t))
 $$
@@
 LST(t) = \begin{cases}
-0,&$if t is a root task$\cr
-\max_{p \in Pred(t)} DL(p),&$otherwise.$\cr
+0,&$if~$Pred(t) = \emptyset\cr
+\max_{p \in Pred(t)} DL(p),&$otherwise.$
 \end{cases}
@@
 considers idle slots in blocks that were allocated for previous tasks to be 
-free, while slots in new blocks cost the full price of the new block. For 
-example, if a task has a runtime of 10 minutes, and the price of a one-hour 
-billing cycle is \$1, then the algorithm will either schedule the task on 
+free, while slots in new blocks cost the full price of a billing cycle. For 
+example, if a task has a runtime of 10 minutes, and the price of a block is 
+\$1, then the algorithm will either schedule the task on 
 an existing VM that has an idle slot larger than 10 minutes for a cost of \$0, 
@@
 termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
-of VMs before shutdown. For the SPSS algorithm we set the $\alpha$ parameter
-for deadline distribution to be 0.7.
+of VMs before shutdown. Based on experiments with different applications, 
+we set the $\alpha$ parameter for deadline distribution in SPSS to be 0.7.
 
@@
 estimated limits as indicated in Section~\ref{sec:scenarios}, resulting in
-ca. 50-150 (budget, deadline) points for each application. The runs were
+approximately 50-150 (budget, deadline) points for each application. The runs were
 repeated for Pareto-distributed and constant workflow sizes. From this result

### 2012-01-20 paper/paper.tex
Changes to section 5.4

@@
 \label{eq:score}
-s = \sum_{workflows\ completed}{2^{-d}}
+s = \sum_{workflows\ completed}{2^{-p}}
 \end{equation}
@@
 
-We run the simulations for the constraint values ranging between the
-estimated limits as indicated in Section~\ref{sec:scenarios}, resulting in
-approximately 50-150 (budget, deadline) points for each application. The runs were
-repeated for Pareto-distributed and constant workflow sizes. From this result
-space we selected the most interesting plots which give the good overview  of
-the observed algorithm behavior.
+We ran simulations for constraint values ranging between the estimated 
+limits indicated in Section~\ref{sec:scenarios}, resulting in approximately 
+50-150 (budget, deadline) data points for each application. The runs were
+repeated for Pareto-distributed and constant workflow ensembles. From this result
+space we selected the most interesting plots that characterize the observed 
+behavior of the algorithms best.
 
@@
 Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
-completed for two budget values per each application ensemble. The budget values
+completed for two budget values for each application ensemble. The budget values
 selected correspond to two opposite cases: {\em small budget} (top) and {\em
-large budget} (bottom). The first one allows completing only a small number of
-workflows from the ensemble, while the latter one suffices to finish almost all
-the workflows. It can be seen that the static planning algorithm (SPSS) performs
-better for tighter constraints: when the budget is low to admit only a small
-number of workflows, the static procedure is able to more efficiently pack these
-workflow tasks onto the resources. This can be clearly seen for the LIGO,
-Epigenomics and SIPHT ensembles for small budgets and also for the same
-ensembles with large budgets but short deadlines. The exception are Montage and
-CyberShake ensembles, which consist of many fine-grained tasks that can be
-executed in parallel. For these cases, as well for the other ensembles with
-large budgets, the simple dynamic scheduling techniques perform equally well or
-better. The reason for that behavior is that the large number of tasks
-guarantees that resources are almost never idle, while the provisioner algorithm
-makes sure that the VM instances are most efficiently utilized before shutdown.
-We can also observe that the workflow-aware dynamic algorithm (WA-DPDS) performs
-better than the workflow-unaware version. This means that for the tested
-workflows the simple admission procedure based on estimation of workflow cost
-does not degrade the solution, but rather in many cases it allows rejecting
-larger workflows that would lead to budget overrun. Thus it can save the space
-for smaller workflows that can complete.
+large budget} (bottom). The small budgets permit the algorithms to complete 
+only a small number of workflows from the ensemble, while the larger budgets 
+enables the algorithms to finish almost all the workflows. In three of five 
+cases the SPSS algorithm performs better for tighter constraints: when the 
+budget is low, or the deadline is short, the static procedure is able to more 
+efficiently pack workflow tasks onto limited resources. This is the case for 
+the LIGO, Epigenomics and SIPHT ensembles for small budgets, and for large
+budgets with short deadlines. The exceptions are the Montage and CyberShake 
+ensembles,  which consist of many fine-grained tasks. In these applications, 
+as well for the other ensembles with large budgets and deadlines, the simple 
+dynamic scheduling techniques perform equally well or better. The reason for 
+this behavior is that the large number of tasks guarantees that resources 
+are almost never idle regardless of how the tasks are scheduled. We also 
+observe that WA-DPDS performs better than the workflow-unaware version, DPDS. 
+This means that for the tested workflows the simple admission procedure based 
+on estimation of workflow cost does not degrade the solution, but rather in 
+many cases it enables larger workflows that would lead to budget overrun to
+be rejected. Thus it can save the space for smaller workflows to complete.
 
@@
 Fig.~\ref{fig:total-time} presents the same results using the total computing
-time metric $T_C$. These plots confirm our previous observations about the
-algorithms performance, but they also allow to observe a more general property.
-For a given budget, when the deadline is tightened the number of work done
+time metric, $T_C$. These plots confirm our previous observations about the
+algorithms' performance, but they also allow us observe a more general property:
+for a given budget, when the deadline is tightened the amount of work done
 steeply decreases. Assigning a shorter deadline means that more VMs need to be
 provisioned to complete the work, but this leads to higher parallelism and thus
-to the lower parallel efficiency. E.g. running 2 VMs for 10 hours will be more
-efficient than running 10 VMs for 2 hours. Therefore, if deadline permits and
-costs are important, parallelism should be avoided when possible.
-
-These observations are further confirmed when examining the effective resource
+to lower parallel efficiency. For example, running 2 VMs for 10 hours will tend
+to be more efficient than running 10 VMs for 2 hours. Similarly, the only way 
+to reduce cost when adding VMs is to achieve a super-linear speedup, which 
+is unlikely in practice. Therefore, if the deadline permits, and cost is 
+important, parallelism should be avoided when possible.
+
+These observations are further confirmed when examining the effective computing
 cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
 utilization and lower parallel efficiency for short deadlines lead to
-substantial increase of actual costs of computing. This increase reaches
-$\sim$20\%, which means that if we need to run e.g. a SIPHT workflow ensemble
-with a short deadline, $\sim$\$200 out of \$1000 is spent on idle resources.
+substantial increases in the effective computing cost. This increase can be 
+significant. For example, a SIPHT ensemble with short deadlines and a budget
+of \$1,000 has an effective computing cost $\sim$8--20\% higher than the 
+normal cost, depending on the algorithm used. That means that out of the 
+\$1,000 budget, \$80--\$200 is spent on idle resources.
 
@@
 represents a trade-off curve between two conflicting objectives: time and cost.
-We can see that the selection of deadline and cost can be formulated as
-multi-criteria optimization problem and the curves presented in
+We observe that the selection of deadline and cost can be formulated as a
+multi-criteria optimization problem, and that the curves presented in 
 Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
 solutions to this problem. In our case the solutions were achieved by maximizing
-the number of work done, which directly results in minimization of effective cost parameter for a given
-deadline. Therefore, the obtained results can be used to assist the decision in
-resource allocation planning, when both cost and time criteria are given not as
-constraints but as objectives.
+the amount of work done, which directly results in minimization of effective 
+computing cost metric for a given deadline. Therefore, the obtained results 
+can be used to aid in resource allocation planning, when both cost and time 
+criteria are given not as constraints but as objectives.
 
@@
 
-In order to compare efficiency of algorithms, we counted the number of times
-each algorithm achieved the best results for all tested applications, deadlines
-and budgets. If two or more algorithms achieve the same highest results, all are
-counted as highest. Table~\ref{tab:num-dags-pareto} shows these summary values
-for Pareto-distributed workflows. These results depend on the metric: number of
+In order to compare the efficiency of our algorithms, we counted the number of times
+each algorithm achieved the best result for all simulated applications, deadlines
+and budgets. Table~\ref{tab:num-dags-pareto} shows these summary values for 
+Pareto-distributed workflows. Note that all algorithms that achieve the best 
+result are credited, so the sum for a given application can be greater than the
+total number of scenarios. These results depend on the metric: number of
 workflows completed does not count the priorities or sizes of workflows, which
-means that it happens more often that all the algorithms achieve the same (best)
+means that it happens more often that all the algorithms achieve the same best
 result. However, it does not mean that all algorithms complete the same
-workflows: that is why values in the tables (a), (b) and (c) are different.
-Exponential score and total amount of computing time complete give the same
+workflows: that is why the values in (a), (b) and (c) are different. The 
+exponential score and total amount of computing time completed give the same
 rankings, since in our test ensembles larger workflows were given higher
-priorities. It can be observed that based on the sampled problem space the
+priorities. It can be observed based on the sampled problem space that the
 workflow-aware algorithms achieve highest score in $\sim$2--3 times more cases
-than the simple DPDS which does not take into account workflow structure or
-size. In average, SPSS performs better in $\sim$20--50\% more times than
-WA-DPDS. We have to note, that the sampled parameter space included the Montage
-and CyberShake workflows consisting of many small tasks, for which cases SPSS
-does not perform equally well.
+than the simple DPDS algorithm which does not take into account workflow 
+structure or size. SPSS performs the best in $\sim$20\% more cases than WA-DPDS,
+including the Montage and CyberShake ensembles, for which SPSS does not 
+perform as well as DPDS and WA-DPDS.
 

### 2012-01-21 paper/paper.tex
More changes to 5.4

@@
 priorities. It can be observed based on the sampled problem space that the
-workflow-aware algorithms achieve highest score in $\sim$2--3 times more cases
-than the simple DPDS algorithm which does not take into account workflow 
+workflow-aware algorithms achieve the highest score in $\sim$2--3 times more 
+cases than the simple DPDS algorithm, which does not take into account workflow 
 structure or size. SPSS performs the best in $\sim$20\% more cases than WA-DPDS,
 including the Montage and CyberShake ensembles, for which SPSS does not 
-perform as well as DPDS and WA-DPDS.
+perform as well as WA-DPDS.
 
 Similar results were achieved for ensembles consisting of workflows of equal
-size. Summary of these results is shown in Table~\ref{tab:num-dags-equal}.
-Again, we can see that static SPSS algorithm performed better for coarse-grained
+size. A summary of these results is shown in Table~\ref{tab:num-dags-equal}.
+Again, we can see that the static SPSS algorithm performed better for coarse-grained
 workflows (LIGO, Epigenomics, SIPHT). The differences between algorithm scores
-are smaller since the workflows have equal size thus rejecting one workflow
-does not help to complete more work in any case.
-
-Fig.~\ref{fig:budgets} shows the same data for SIPHT ensemble from another
-perspective. It helps answer the question: given a deadline, how many workflows
-can be finished when we increase the budget. In Fig.~\ref{fig:budgets}(a) the
-growth is super-linear, what results from the distribution of workflow sizes in
-the ensemble (see Fig.~\ref{fig:ensemble-distribution}). For ensemble of equal
-size workflows (see Fig.~\ref{fig:budgets}(b)) the growth is nearly linear. It
-can be also observed that static algorithms give better results for smaller
-budgets, while dynamic algorithms for larger ones.
+are smaller since the workflows have equal size, therefore rejecting a particular
+workflow does not enable the algorithms to complete any other workflows.
+
+Fig.~\ref{fig:budgets} shows the same data for the SIPHT ensemble from another
+perspective. It helps answer the question: Given a deadline, how many workflows
+can be finished when we increase the budget? In Fig.~\ref{fig:budgets}(a) the
+growth is super-linear, which results from the Pareto-like distribution of 
+workflow sizes in the ensemble (see Fig.~\ref{fig:ensemble-distribution}). 
+For ensembles of constant size workflows (see Fig.~\ref{fig:budgets}(b)) 
+the growth is nearly linear.
 

### 2012-01-21 paper/paper.tex
Ewa's comments

@@
 
-\keywords{Scientific workflows, DAG scheduling, simulation}
+\keywords{Scientific workflows, cloud computing, DAG scheduling, simulation}
 
@@
 either a performance model of the application, or historical data that can be
-mined.
+mined. In particular, in this study we use workflows from the synthetic workflow
+generator~\cite{Bharathi08}.
 
@@
 workflow scheduling and provisioning. The simulator reads workflows in a modified version of
-the DAX format used by the Pegasus Workflow Management System~\cite{Deelman2005} that was
+the DAX format used by the Pegasus Workflow Management System~\cite{Deelman2005}. The modified format was
 created for the synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi08}.
@@
 of the different algorithms for smaller and larger constraints. For a given
-ensemble we can expect, in general, that the amount of work done (in terms 
-of workflows completed) depends mainly on the budget constraint, which 
-determines the number of machine-hours that can be provisioned. The deadline, 
-although important in terms of scheduling, is assumed to be greater than the 
-critical path of every workflow. The ideal but trivial case, then, is a deadline 
-long enough to allow all the workflow tasks to be executed sequentially: this 
-eliminates all the overheads of parallel execution and leads to the highest 
-resource utilization. More interesting, however, are the cases where the deadline 
-is shorter than the serial schedule, since it forces the system to provision 
-multiple VMs to execute workflows in parallel.
+ensemble we can expect, in general, that the amount of work done (in terms of
+workflows completed) depends mainly on the budget constraint, which determines
+the number of machine-hours that can be provisioned. The deadline, although
+important in terms of scheduling, is assumed to be greater than the critical
+path of every workflow. We can then have a trivial but unrealistic solution
+where the deadline  is long enough to allow all the workflow tasks to be
+executed sequentially: this eliminates all the overheads of parallel execution
+and leads to the highest resource utilization. More interesting and more
+realistic, however, are the cases where the deadline is shorter than the serial
+schedule, since it forces the system to provision multiple VMs to execute
+workflow tasks in parallel.
 
@@
 
-The number of workflows in an ensemble depends on the user's requirements, but
+The number of workflows in an ensemble depends on the particular application, but
 we assume that ensemble sizes are on the order of between 10 and 100 workflows.
@@
 workflows modeled on real applications. We used workflows from five applications 
-that were developed for the Pegasus Workflow Management System. We constructed 
-ensembles of 100 workflows by sampling collections of workflows of various sizes
-(number of tasks) based on a given distribution. Two distributions are considered:
+that were managed by the Pegasus Workflow Management System. For each
+application we constructed ensembles of 100 workflows by sampling collections of
+workflows of various sizes (number of tasks) based on a given distribution.
+Two distributions are considered:
 
@@
 
-The constant distribution produces ensembles containing of workflows of equal size, 
+The constant distribution produces ensembles containing of workflows with the same number of tasks per workflow, 
 while the Pareto-like distribution produces ensembles with a small number of larger 
@@
 \$1 per VM-hour. We also assume that the sizes of tasks in the workflow gallery
-are given in seconds measured on a 1000 MIPS machine. In this study we do not 
-take into account the heterogeneity of the infrastructure since we assume that 
-it is always possible to select a VM type that has the best price to performance 
-ratio for a given application.
+are given in seconds measured on a 1000 MIPS machine. These values were chosen
+to approximate realistic numbers, and also to simplify interpretation of
+results. Moreover, the absolute values of these parameters do not affect the
+relative comparison between the algorithms. In this study we do not take into
+account the heterogeneity of the infrastructure since we assume that it is
+always possible to select a VM type that has the best price to performance ratio
+for a given application.
 
@@
 40 and 60 hours. The Epigenomics workflow is even more costly, requiring 
-approximately \$12,000 and 1500 hours to complete a full ensemble.
+approximately \$12,000 and 1500 hours to complete a full ensemble. These
+differences in total runtime and cost come from the fact that we used the
+same workflow size distribution for all five applications, but
+the sizes of tasks vary between applications.
 
@@
   \item $C_{eff}$: effective cost of computing per hour, which includes
-  overheads.
+  overheads due to cost of idle resources.
 \end{itemize}
@@
 \label{eq:score}
-s = \sum_{workflows\ completed}{2^{-p}}
+s = \sum_{workflows\ completed}{2^{-d}}
 \end{equation}
@@
 
-The total amount computing time completed, $T_C$, is calculated as the sum 
+The total amount of computing time completed, $T_C$, is calculated as the sum 
 of the runtimes of all tasks in all completed workflows. Because we assumed 
@@
 $u = (T_C * \$1)/b$, is the resource utilization, and its inverse: 
-$C_{eff} = b/T_C$, is the effective resource cost in dollars per hour.
+$C_{eff} = b/T_C$, is the effective resource cost in dollars per hour. To
+calculate $C_{eff}$ and $u$ we use the value of allocated budget $b$, not the
+actual total cost, since we are interested in the scenarios when the budget is
+limited and does not allow completing all the workflows from the ensemble.
 

### 2012-01-21 paper/paper.tex
Update section 6

@@
 as well for the other ensembles with large budgets and deadlines, the simple 
-dynamic scheduling techniques perform equally well or better. The reason for 
-this behavior is that the large number of tasks guarantees that resources 
+dynamic scheduling techniques perform equally well or better. For these 
+applications the large number of small tasks guarantees that resources 
 are almost never idle regardless of how the tasks are scheduled. We also 
@@
 
-In this paper we addressed a new interesting and important problem of scheduling
-and resource provisioning for scientific workflow ensembles on IaaS cloud
-infrastructures. The problem is different from previous work on grid or utility
-grid scheduling that cloud infrastructure gives more control over the resources,
-so the resource provisioning plan can be adjusted to the application
-requirements. Therefore the problem space becomes larger: it requires not
-only to select the best scheduling of tasks to available resources, but also
-select the best shape of resource provisioning plan. The formulation of a
-problem as maximization of the number of prioritized workflows completed from
-the ensemble requires also to take the decisions on admitting or rejecting workflows
-based on their estimated resource demands. We believe that such formulation of
-the bi-constrained problem is highly relevant since such
-constraints are typically imposed on many real-world projects.
-
-We have analyzed both static and dynamic scheduling approaches and
-developed the SPDS, DPDS, WA-DPDS and SPSS algorithms realizing these
-strategies. The algorithms were evaluated on ensembles of benchmark
-workflows which represent typical real scientific applications. For the
-purpose of evaluation we have developed a simulator that models the cloud
-infrastructure and the tightly-coupled scheduler and provisioner modules of
-workflow engine. The results of simulation studies indicate that the two algorithms (WA-DPDS and
-SPSD) that take into account the information about the workflow structure and
-estimation of task runtimes yield better results than the simple on-line
-priority-based scheduling strategy with static (constant) resource provisioning
-plan.
-
-During our work, several interesting lessons were learned about the workflow
-ensembles on cloud infrastructures. One result is that the simple strategy of
-SPDS performs relatively well in terms of number of workflows completed and the
-high resource utilization it can achieve. However, this method does not allow to
-give any advance prediction about the result, so to see its performance the
-whole ensemble needs to be simulated. On the other hand, the method may be
-the only choice in the case no estimates about the task runtimes are available
-in advance. It can be seen that the workflow-aware and static algorithms are
-able to find better quality solutions in terms of the defined metrics: this the
-result mainly of the admission algorithm implemented, which allow rejecting too
-large workflows and thus saving resources for completing the less costly ones.
-
-Another general observation is that although clouds provide such capabilities as
-dynamic adding and removing of resources at runtime and high parallelization due
-to illusion of unlimited resources, it turns out that is recommended to avoid
-using these capabilities when possible, in order to achieve high resource
-utilization and cost effectiveness. This observation comes from the fact that
-starting and stopping more VMs incurs a cost which can be amortized by running
-each started VM as long as possible before shutdown. This effect was confirmed
-in the case of SPSS algorithm for Montage and CyberShake workflows with small
-budgets. Similarly, high degree of parallelization (large number of nodes) leads
-also to less efficient resource utilization and in the case of clouds the drop
-in parallel efficiency means also the increase of cost. This result, clearly
-visible in Fig~\ref{fig:cost}, leads to multi-criteria decision problems and
-related trade-offs. E.g. our simulation results can be used as hints to answer
+In this paper we addressed the interesting and important new problem of 
+scheduling and resource provisioning for scientific workflow ensembles 
+on IaaS clouds. The problem is different from previous work on grid and 
+utility grid scheduling in that cloud infrastructures provide more control 
+over the resources, so the resource provisioning plan can be adjusted 
+according to the requirements of the application. Therefore the problem 
+space becomes larger: it requires not only an efficient mapping of tasks
+to available resources, but also the selection of the best resource 
+provisioning plan. Formulating the problem as a maximization of the
+number of prioritized workflows completed from the ensemble also 
+requires workflows to be admitted or rejected based on their estimated 
+resource demands. We believe that this bi-constrained problem is highly 
+relevant because such constraints are commonly imposed on many real-world 
+projects.
+
+We have considered both static and dynamic scheduling approaches and 
+developed the SPDS, DPDS, WA-DPDS and SPSS algorithms based on these 
+strategies. The algorithms were evaluated on ensembles of synthetic 
+workflows, which were generated based on real scientific applications. For 
+the purposes of evaluation we have developed a simulator that models: the 
+cloud infrastructure, and a workflow engine with tightly-coupled scheduler 
+and provisioner modules. The results of our simulation studies indicate 
+that the two algorithms that take into account information about the 
+structure of the workflow and task runtime estimates (WA-DPDS and SPSS) 
+yield better results than the simple on-line priority-based scheduling 
+strategy (DPDS).
+
+During our work, several interesting lessons were learned about running 
+workflow ensembles on cloud infrastructures. One result is that the simple 
+strategy used by DPDS performs relatively well in terms of number of 
+workflows completed and the high resource utilization it can achieve. However,
+the extra workflows may be of lower-priority, and the high utilization
+may be achieved by running tasks from workflows that cannot be completed
+given the constraints. On the other hand, DPDS may be the only choice when 
+estimates of task runtimes are not available in advance. We also observed
+that the WA-DPDS and SPSS algorithms are able to find better quality 
+solutions in terms of the target metrics: this is primarily a result of the 
+admission algorithms rejecting workflows that are too large to be completed 
+given the constraints, and accepting more smaller workflows that can be 
+completed.
+
+Another general observation is that although IaaS clouds provide elasticity
+and the illusion of unlimited resources, which makes it possible to provision
+more VMs to run workflows in parallel, it turns out that higher resource
+utilization and better cost effectiveness can be achieved by limiting 
+parallelism as much as possible. This observation comes from the fact that 
+starting and stopping VMs incurs overheads which can be amortized by running
+each VM for as long as possible. This effect was confirmed in the case of 
+the SPSS algorithm for Montage and CyberShake workflows with small budgets
+where SPSS opted to provision many VMs in parallel, and shut them down early,
+so that tasks early in the ensemble workflows could be completed by their 
+deadlines. Similarly, a high degree of parallelization leads to lower 
+resource utilization and increased cost. This result, visible in 
+Fig~\ref{fig:cost}, leads to multi-criteria decision problems and related 
+trade-offs. E.g. our simulation results can be used as hints to answer
 the question how much time and money is required to actually complete all the
@@
 
-Another lesson learned from our studies is related to the shapes and granularity
-of the tasks of workflows in the ensembles. The benchmarks workflows that we
-tested have relatively large number of tasks (50--1000). Montage and CyberShake
-workflows have fine-grained tasks of short execution times, in the order of
-seconds and minutes. LIGO, Epigenomics and SIPHT have longer tasks: from minutes to hours.
-Therefore the dynamic algorithms achieved better results on the fine grained
-workflows, since the simple random-based allocation strategy leads to good
-load-balancing and high resource utilization. On the other hand, when tasks are
-longer, it becomes more beneficial to apply a static planning strategy of SPSS.
-
-Our current study reveals multiple paths for future work. One interesting
-question would be to answer how the static and dynamic algorithms perform in the
-cases of increased uncertainties and failures, that have important influence
-on overall system performance~\cite{Sakellariou2010,Dongarra2007}. Second
-important goal would be to extend our application and infrastructure model to
-include various data storage options available on clouds. The experimental study
-on that subject~\cite{Juve2010} suggests that data demands of scientific
-workflows have a high impact not only on execution time but also on costs on
-commercial clouds. Finally, we would like to address the heterogeneity of the
-infrastructure, as multiple instance types and providers, including private
-and community clouds will make the problem more complex and
-challenging~\cite{Marshall2010,vockler11,Juve2010}.
+Another lesson learned from our studies is the effect of task granularity on 
+workflow ensembles. The benchmark workflows that we tested have relatively 
+large number of tasks (50--1000), but the Montage and CyberShake workflows have 
+fine-grained tasks with execution times on the order of seconds to minutes, 
+while LIGO, Epigenomics and SIPHT have coarser-grained tasks with runtimes 
+ranging from several minutes to hours. We observed that the dynamic 
+algorithms achieved better results on the fine grained workflows since the 
+simple random scheduling strategy leads to good load-balancing and high 
+resource utilization for these ensembles. On the other hand, when tasks are
+longer, it becomes more beneficial to apply the more sophisticated static 
+scheduling strategy of SPSS.
+
+Our current study suggests many areas for future work. One interesting
+issue that we plan to investigate is how the static and dynamic algorithms 
+perform with uncertain task runtimes and failures, which are known to have 
+an important influence on overall system performance~\cite{Sakellariou2010,Dongarra2007}. 
+We also plan to extend our application and infrastructure model to 
+include the various data storage options available on clouds. A previous 
+experimental study on that subject~\cite{Juve2010} suggests that the data 
+demands of scientific workflows have a large impact on not only the 
+execution time, but also on the cost of workflows in commercial clouds. 
+Finally, we plan to investigate heterogeneous environments that include 
+multiple VM types and cloud providers, including private and community clouds, 
+which will make the problem even more complex and challenging~\cite{Marshall2010,vockler11,Juve2010}.
 

### 2012-01-21 paper/paper.tex
Add Ewa's changes to 5.4

@@
 50-150 (budget, deadline) data points for each application. The runs were
-repeated for Pareto-distributed and constant workflow ensembles. From this result
-space we selected the most interesting plots that characterize the observed 
-behavior of the algorithms best.
+performed for Pareto-distributed and constant workflow ensembles. From this result
+space we selected representative plots.
 
@@
 \includegraphics[width=0.19\textwidth]{figures/pareto-cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
-\caption{Effective computing cost in \$ per hour calculated by dividing sum of
-task runtimes by the total budget. Higher cost results from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
+\caption{Effective computing cost in \$ per hour calculated by dividing the sum of
+task runtimes by the total budget. Higher costs result from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
 Pareto-distributed workflows.}
@@
 completed for two budget values for each application ensemble. The budget values
-selected correspond to two opposite cases: {\em small budget} (top) and {\em
+selected correspond to two extreme cases: {\em small budget} (top) and {\em
 large budget} (bottom). The small budgets permit the algorithms to complete 
 only a small number of workflows from the ensemble, while the larger budgets 
-enables the algorithms to finish almost all the workflows. In three of five 
-cases the SPSS algorithm performs better for tighter constraints: when the 
+enable the algorithms to finish almost all the workflows. In three of five 
+cases the SPSS algorithm performs better for tighter constraints. When the 
 budget is low, or the deadline is short, the static procedure is able to more 
@@
 as well for the other ensembles with large budgets and deadlines, the simple 
-dynamic scheduling techniques perform equally well or better. For these 
-applications the large number of small tasks guarantees that resources 
+dynamic scheduling techniques perform equally well or better. The reason for 
+this behavior is that the large number of tasks guarantees that resources 
 are almost never idle regardless of how the tasks are scheduled. We also 
 observe that WA-DPDS performs better than the workflow-unaware version, DPDS. 
-This means that for the tested workflows the simple admission procedure based 
+This means that for the workflows we tested the simple admission procedure based 
 on estimation of workflow cost does not degrade the solution, but rather in 
@@
 steeply decreases. Assigning a shorter deadline means that more VMs need to be
-provisioned to complete the work, but this leads to higher parallelism and thus
-to lower parallel efficiency. For example, running 2 VMs for 10 hours will tend
-to be more efficient than running 10 VMs for 2 hours. Similarly, the only way 
-to reduce cost when adding VMs is to achieve a super-linear speedup, which 
-is unlikely in practice. Therefore, if the deadline permits, and cost is 
-important, parallelism should be avoided when possible.
+provisioned to complete the work. However, because of the dependencies in the 
+workflows, many VMs will either be idle, or will be backfilled with tasks from 
+low-priority workflows that never finish, resulting in reduced resource 
+utilization. Similarly, the only way to reduce cost when adding VMs is to 
+achieve a super-linear speedup, which is unlikely in practice. Therefore, if 
+the deadline permits, and cost is important, parallelism should be kept to a 
+minimum.
 

### 2012-01-21 paper/paper.tex
Add Ewa's changest for section 6

@@
 provisioning plan. Formulating the problem as a maximization of the
-number of prioritized workflows completed from the ensemble also 
+number of prioritized workflows completed from the ensemble is also novel and  
 requires workflows to be admitted or rejected based on their estimated 
@@
 relevant because such constraints are commonly imposed on many real-world 
-projects.
+projects. The approach is also directly applicable to grid environments 
+that provide resource reservations and charge service units for resource use.
 
 We have considered both static and dynamic scheduling approaches and 
-developed the SPDS, DPDS, WA-DPDS and SPSS algorithms based on these 
+developed the DPDS, WA-DPDS and SPSS algorithms based on these 
 strategies. The algorithms were evaluated on ensembles of synthetic 
@@
 workflow ensembles on cloud infrastructures. One result is that the simple 
-strategy used by DPDS performs relatively well in terms of number of 
+strategy used by DPDS performs relatively well in terms of the number of 
 workflows completed and the high resource utilization it can achieve. However,
@@
 utilization and better cost effectiveness can be achieved by limiting 
-parallelism as much as possible. This observation comes from the fact that 
+parallelism. This observation comes from the fact that 
 starting and stopping VMs incurs overheads which can be amortized by running
 each VM for as long as possible. This effect was confirmed in the case of 
-the SPSS algorithm for Montage and CyberShake workflows with small budgets
-where SPSS opted to provision many VMs in parallel, and shut them down early,
+the SPSS algorithm for Montage and CyberShake workflows with small budgets.
+In these cases SPSS decided to provision many VMs in parallel, and shut them down early,
 so that tasks early in the ensemble workflows could be completed by their 
@@
 Another lesson learned from our studies is the effect of task granularity on 
-workflow ensembles. The benchmark workflows that we tested have relatively 
-large number of tasks (50--1000), but the Montage and CyberShake workflows have 
+workflow ensembles. All the benchmark workflows that we tested have an average 
+number of tasks (50--1000), but the Montage and CyberShake workflows have 
 fine-grained tasks with execution times on the order of seconds to minutes, 
@@
 an important influence on overall system performance~\cite{Sakellariou2010,Dongarra2007}. 
-We also plan to extend our application and infrastructure model to 
-include the various data storage options available on clouds. A previous 
+We also plan to extend our application and infrastructure model to explicitly 
+include the various data storage options available on clouds (right now we only 
+model data access as part of the task time and do not include data costs). A previous 
 experimental study on that subject~\cite{Juve2010} suggests that the data 

### 2012-01-21 paper/paper.tex
Addressing some of Ewa's comments from 5.4

@@
 either a performance model of the application, or historical data that can be
-mined. In particular, in this study we use workflows from the synthetic workflow
-generator~\cite{Bharathi08}.
+mined. In this study we use synthetic workflows that were generated using 
+historical data from real applications~\cite{Bharathi08}. These applications 
+come from a wide variety of domains including: bioinformatics (Epigenomics, 
+SIPHT), astronomy (Montage), earthquake science (CyberShake), and physics 
+(LIGO).
 
@@
 run. Each workflow in an ensemble is given a priority that indicates how
-important the workflow is to the user.
+important the workflow is to the user. These priorities are absolute in the 
+sense that completing a high priority workflow is more valuable than 
+completing all lower priority workflows combined.
 
@@
 to complete as many high-priority workflows as possible given a fixed budget and
-deadline.
+deadline. Only workflows for which all tasks are finished by the deadline are 
+considered to be complete---partial results are not usable in this model.
 
@@
 of \$1,000 has an effective computing cost $\sim$8--20\% higher than the 
-normal cost, depending on the algorithm used. That means that out of the 
+minimum effective cost of \$1/hour, depending on the algorithm used. That means that out of the 
 \$1,000 budget, \$80--\$200 is spent on idle resources.

### 2012-01-22 paper/paper.tex
Updates to the conclusions

@@
 for a given budget, when the deadline is tightened the amount of work done
-steeply decreases. Assigning a shorter deadline means that more VMs need to be
+steeply decreases. This is partially due to the fact that fewer workflows can be
+completed given a short deadline because their critical paths are simply too long
+and they are squeezed out of early start times by high priority workflows.
+In addition, assigning a shorter deadline means that more VMs need to be
 provisioned to complete the work. However, because of the dependencies in the 
-workflows, many VMs will either be idle, or will be backfilled with tasks from 
-low-priority workflows that never finish, resulting in reduced resource 
-utilization. Similarly, the only way to reduce cost when adding VMs is to 
-achieve a super-linear speedup, which is unlikely in practice. Therefore, if 
-the deadline permits, and cost is important, parallelism should be kept to a 
-minimum.
+workflows, many of these VMs will either be idle, or will be backfilled with 
+tasks from low-priority workflows that never finish, resulting, effectively, 
+in reduced resource utilization and fewer completed workflows.
 
@@
 of \$1,000 has an effective computing cost $\sim$8--20\% higher than the 
-minimum effective cost of \$1/hour, depending on the algorithm used. That means that out of the 
-\$1,000 budget, \$80--\$200 is spent on idle resources.
+minimum effective cost of \$1/hour, depending on the algorithm used. This means 
+that out of the \$1,000 budget, \$80--\$200 is spent on idle resources.
 
@@
 solutions to this problem. In our case the solutions were achieved by maximizing
-the amount of work done, which directly results in minimization of effective 
+the amount of work done, which is the inverse of minimizing the effective 
 computing cost metric for a given deadline. Therefore, the obtained results 
@@
 are smaller since the workflows have equal size, therefore rejecting a particular
-workflow does not enable the algorithms to complete any other workflows.
+workflow does not make it any easier for the algorithms to complete any other 
+workflow.
 
@@
 For ensembles of constant size workflows (see Fig.~\ref{fig:budgets}(b)) 
-the growth is nearly linear.
+the growth is linear.
 
@@
 
-Another general observation is that although IaaS clouds provide elasticity
-and the illusion of unlimited resources, which makes it possible to provision
-more VMs to run workflows in parallel, it turns out that higher resource
-utilization and better cost effectiveness can be achieved by limiting 
-parallelism. This observation comes from the fact that 
-starting and stopping VMs incurs overheads which can be amortized by running
-each VM for as long as possible. This effect was confirmed in the case of 
-the SPSS algorithm for Montage and CyberShake workflows with small budgets.
-In these cases SPSS decided to provision many VMs in parallel, and shut them down early,
-so that tasks early in the ensemble workflows could be completed by their 
-deadlines. Similarly, a high degree of parallelization leads to lower 
-resource utilization and increased cost. This result, visible in 
-Fig~\ref{fig:cost}, leads to multi-criteria decision problems and related 
-trade-offs. E.g. our simulation results can be used as hints to answer
-the question how much time and money is required to actually complete all the
-workflows from the given ensemble.
-
 Another lesson learned from our studies is the effect of task granularity on 

### 2012-01-22 paper/paper.tex
Minor changes

@@
 workflows completed and the high resource utilization it can achieve. However,
-the extra workflows may be of lower-priority, and the high utilization
+some of these workflows may be of lower priority, and the high utilization
 may be achieved by running tasks from workflows that cannot be completed
@@
 solutions in terms of the target metrics: this is primarily a result of the 
-admission algorithms rejecting workflows that are too large to be completed 
-given the constraints, and accepting more smaller workflows that can be 
-completed.
+admission algorithms, which reject workflows that are too large to be 
+completed given the constraints, and accept more smaller workflows that 
+can be completed.
 
@@
 resource utilization for these ensembles. On the other hand, when tasks are
-longer, it becomes more beneficial to apply the more sophisticated static 
-scheduling strategy of SPSS.
+longer, the dynamic algorithms tend to use up the budget by running tasks 
+from low-priority workflows while the more sophisticated static scheduling 
+strategy of SPSS is more circumspect. In the future we plan to investigate 
+the effects of task granularity on ensemble scheduling in a more systematic
+fashion.
 

### 2012-01-22 paper/paper.tex
addressed Ewa's comments

@@
 ratio for the application~\cite{Juve2009}. We also assume that a submitted task
-has exclusive access to a VM instance and that there is no preemption.
+has exclusive access to a VM instance and that there is no preemption. For this
+study we assumed that there are no delays in VM start and terminate requests.
 
@@
 algorithms on ensembles of 100 Pareto-distributed workflows, {\em small budget}
-(top) and {\em large budget} (bottom).}
+(top) and {\em large budget} (bottom). The ranges on $y$ axis are chosen to 
+emphasize the differences the performance of the algorithms.}
 \label{fig:number-complete-pareto}
@@
 assigned priorities according to their size so that the largest workflows are given 
-the highest priority. In future work we plan to explore other ensemble types such as
-uniform distributions.
+the highest priority. In
+future work we plan to explore other ensemble types such as uniform distributions.
 
@@
 tasks of completed workflows. Data computed for DPDS, AW-DPDS and SPSS
-algorithms on ensembles of 100 Pareto-distributed workflows}
+algorithms on ensembles of 100 Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
+emphasize the differences between the performance of the algorithms.}
 \label{fig:total-time}
@@
 task runtimes by the total budget. Higher costs result from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
-Pareto-distributed workflows.}
+Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
+emphasize the differences the performance of the algorithms.}
 \label{fig:cost}
@@
 only a small number of workflows from the ensemble, while the larger budgets 
-enable the algorithms to finish almost all the workflows. In three of five 
-cases the SPSS algorithm performs better for tighter constraints. When the 
+enable the algorithms to finish almost all the workflows. The rangesIn three of
+five cases the SPSS algorithm performs better for tighter constraints. When the 
 budget is low, or the deadline is short, the static procedure is able to more 
@@
 \begin{figure}[htb] \centering
-\includegraphics[width=0.7\columnwidth]{figures/pareto-budget-SIPHT-n-1000-8-dagh5-50h10m0.pdf}\\
+\includegraphics[width=0.7\columnwidth]{figures/pareto-budget-SIPHT-n-1000-8-dagh5-50h4m0.pdf}\\
 (a) ensemble of Pareto-distributed workflows
@@
 Similar results were achieved for ensembles consisting of workflows of equal
-size. A summary of these results is shown in Table~\ref{tab:num-dags-equal}.
-Again, we can see that the static SPSS algorithm performed better for coarse-grained
+size set to 1000. A summary of these results is shown in
+Table~\ref{tab:num-dags-equal}. Again, we can see that the static SPSS algorithm performed better for coarse-grained
 workflows (LIGO, Epigenomics, SIPHT). The differences between algorithm scores
@@
 For ensembles of constant size workflows (see Fig.~\ref{fig:budgets}(b)) 
-the growth is linear.
+the growth is linear and the costs are higher, since all the worklfows in the
+test consisted of 1000 tasks.
 

### 2012-01-22 src/cws/core/experiment/Experiment.java
set provisioning and deprovisioning delays to 0

@@ -85,6 +85,8 @@ public class Experiment {
 			HashSet<VM> vms = new HashSet<VM>();
 			for (int i = 0; i < numVMs; i++) {
 				VM vm = new VM(1000, 1, 1.0, param.getPrice());
+	            vm.setProvisioningDelay(0.0);
+	            vm.setDeprovisioningDelay(0.0);
 				vms.add(vm);
 				CloudSim.send(engine.getId(), cloud.getId(), 0.0, WorkflowEvent.VM_LAUNCH, vm);
 			}	

### 2012-01-22 src/cws/core/provisioner/SimpleUtilizationBasedProvisioner.java

@@ -65,8 +65,8 @@ public class SimpleUtilizationBasedProvisioner extends AbstractProvisioner imple
 			// seconds till next full hour
 			double secondsRemaining = vmHours*3600.0-vmRuntime;
 			
-			// we add "magic number" 11.0 to include also the deprovisioning time
-			if (secondsRemaining<PROVISIONER_INTERVAL+11.0) {
+			// we add "magic number" 1.0 to include also the deprovisioning time
+			if (secondsRemaining<PROVISIONER_INTERVAL+1.0) {
 				completingVMs.add(vm);
 			}
 		}
@@ -147,6 +147,8 @@ public class SimpleUtilizationBasedProvisioner extends AbstractProvisioner imple
 		if (! finishing_phase && utilization > UPPER_THRESHOLD && numBusyVMs+numFreeVMS <= getMaxScaling() * initialNumVMs && budget - cost >= vmPrice) {
 			
 			VM vm = new VM(1000, 1, 1.0, 1.0);
+            vm.setProvisioningDelay(0.0);
+            vm.setDeprovisioningDelay(0.0);
 			Log.printLine(CloudSim.clock() + " Starting VM: " + vm.getId());
 			CloudSim.send(engine.getId(), cloud.getId(), 0.0, VM_LAUNCH, vm);
 			
@@ -202,7 +204,7 @@ public class SimpleUtilizationBasedProvisioner extends AbstractProvisioner imple
 			double secondsRemaining = vmHours*3600.0-vmRuntime;
 			
 			//terminate only vms that have less seconds remaining than a defined threshold
-			if (secondsRemaining<PROVISIONER_INTERVAL+11.0) {
+			if (secondsRemaining<PROVISIONER_INTERVAL+1.0) {
 				vmIt.remove();
 				removed.add(vm);
 				Log.printLine(CloudSim.clock() + " Terminating VM: " + vm.getId());

### 2012-01-22 paper/paper.tex
Updated figures for pareto ensembles, cleanup unused figures
added abstract, related work and affiliation

@@
 
-\numberofauthors{2}
+\numberofauthors{3}
 \author{
-    \alignauthor Maciej Malawski and Jarek Nabrzyski\\
+    \alignauthor Maciej Malawski\\
+       \affaddr{AGH University of Science and Technology}\\
+       \affaddr{Dept. of Computer Science}\\
+       \affaddr{Krakow, Poland; and}\\
        \affaddr{University of Notre Dame}\\
        \affaddr{Center for Research Computing}\\
-       \affaddr{111 Information Technology Center}\\
-       \affaddr{Notre Dame, IN, USA}\\
-       \email{\{mmalawsk,naber\}@nd.edu}
+       \affaddr{111 ITC, Notre Dame, IN, USA}\\
+       \email{mmalawsk@nd.edu}
+    \alignauthor Jarek Nabrzyski\\
+       \affaddr{University of Notre Dame}\\
+       \affaddr{Center for Research Computing}\\
+       \affaddr{111 ITC, Notre Dame, IN, USA}\\
+       \email{naber@nd.edu}
     \alignauthor Gideon Juve and Ewa Deelman\\
@@
 \begin{abstract}
-Abstract goes here.
+
+Large-scale computational applications represented as scientific workflows, are
+often grouped into ensembles composed of several inter-related workflows. In
+this paper, we address a new important problem of maximizing the number of
+completed workflows from an ensemble under both budget and deadline constraints
+on the Infrastructure-as-a-Service (IaaS) clouds, characterized by dynamic
+resource provisioning capabilities and pay-per-use model.  We discuss, develop
+and assess algorithms based on static and dynamic strategies for both task
+scheduling and resource provisioning. We evaluate the algorithms via simulation
+using a set of scientific workflow ensembles with a broad range of budget and
+deadline parameters. This allows to identify the application scenarios and
+constraint ranges for which the algorithms perform better. We found that the key
+for algorithm performance is the decision which workflows should be admitted or
+rejected from the ensemble. The results show that the admission procedure based
+on worklfow structure and estimates of task runtimes can significantly improve
+the quality of solutions.
+
 \end{abstract}
@@
 budget and deadline parameters.
+
+The main contributions of this paper are:
+\begin{itemize}
+  \item definition of the scheduling problem of priorotized workflows grouped in
+  ensembles, under both budget and deadline constraints,
+  \item analysis of static and dynamic strategies for task scheduling and
+  resource provisioning on clouds, together with the admission procedure to
+  predict which workflows from the ensemble can be completed,
+  \item development of static and dynamic algorithms and of the simulator based
+  on the defined application and infrastructure model,
+  \item performance evaluation of algorithms on the set of synthetic benchmark
+  workflow ensembles representing important real scientific applications,
+  \item analysis of results, which allows to assess the advantages and
+  disadvantages of proposed algorithms, depending on application scenarios and
+  varying constraint values.
+\end{itemize}
 
@@
 \hline
-DPDS      &   57  & 45 &  18 &  14  &  0 & 134\tabularnewline
-WA-DPDS   &    96 &  87  & 48  & 34  &  0 & 265\tabularnewline
-SPSS     &    29  & 42  & 192  & 74  & 50 & 387\tabularnewline
+DPDS      &    55 &  47  & 23  & 14  &  0 & 139\tabularnewline 
+WA-DPDS   &    97 &  89  & 50  & 34  &  1 & 271\tabularnewline 
+SPSS      &    38 &  37  & 190 & 74  & 50 & 389\tabularnewline
 \end{tabular}
@@
 \hline
-DPDS      &   58  & 45 &  19 &  14  &  0 & 136\tabularnewline
-WA-DPDS   &    90  & 96  & 57  & 27 &   8 & 278\tabularnewline
-SPSS     &    31 &  18 &  170  & 75 &  42 & 336\tabularnewline
+DPDS      &   55  & 47  &  26  & 14  &  0 & 142\tabularnewline
+WA-DPDS   &   96  & 98  &  59  & 26  & 10 & 289\tabularnewline
+SPSS      &   32  & 12  & 165  & 75  & 41 & 325\tabularnewline
 \end{tabular}

### 2012-01-22 paper/paper.tex
Ewa's edits to abstract and contributions

@@
 often grouped into ensembles composed of several inter-related workflows. In
-this paper, we address a new important problem of maximizing the number of
-completed workflows from an ensemble under both budget and deadline constraints
-on the Infrastructure-as-a-Service (IaaS) clouds, characterized by dynamic
-resource provisioning capabilities and pay-per-use model.  We discuss, develop
-and assess algorithms based on static and dynamic strategies for both task
-scheduling and resource provisioning. We evaluate the algorithms via simulation
-using a set of scientific workflow ensembles with a broad range of budget and
-deadline parameters. This allows to identify the application scenarios and
-constraint ranges for which the algorithms perform better. We found that the key
-for algorithm performance is the decision which workflows should be admitted or
-rejected from the ensemble. The results show that the admission procedure based
-on worklfow structure and estimates of task runtimes can significantly improve
-the quality of solutions.
+this paper, we address a new important problem of efficiently managing such
+ensembles under both budget and deadline constraints on the
+Infrastructure-as-a-Service (IaaS) clouds. The latter are characterized by
+dynamic resource provisioning capabilities and a pay-per-use model.  We discuss,
+develop and assess algorithms based on static and dynamic strategies for both
+task scheduling and resource provisioning. We assume that the workflows within
+an ensemble are prioritized by the user. We evaluate the algorithms via
+simulation using a set of scientific workflow ensembles with a broad range of
+budget and deadline parameters. This allows to identify the application
+scenarios and constraint ranges for which the algorithms perform better. We
+found that the key for algorithm performance is the decision which workflows
+within an ensemble should be admitted or rejected for execution. The results
+show that the admission procedure based on workflow structure and estimates of
+task runtimes can significantly improve the quality of solutions.
 
@@
 \begin{itemize}
-  \item definition of the scheduling problem of priorotized workflows grouped in
+  \item the definition of the scheduling problem of prioritized workflows grouped in
   ensembles, under both budget and deadline constraints,
-  \item analysis of static and dynamic strategies for task scheduling and
+  \item an analysis of static and dynamic strategies for task scheduling and
   resource provisioning on clouds, together with the admission procedure to
   predict which workflows from the ensemble can be completed,
-  \item development of static and dynamic algorithms and of the simulator based
+  \item the development of static and dynamic algorithms and of the simulator based
   on the defined application and infrastructure model,

### 2012-01-23 paper/paper.tex
Edits to abstract and introduction

--- 2012-01-22 paper/paper.tex
@@
 \begin{abstract}
-Abstract goes here.
+
+Large-scale computational applications represented as scientific workflows, are
+often grouped into ensembles composed of several inter-related workflows. In
+this paper, we address a new and important problem concerning the efficient 
+management of such ensembles under both budget and deadline constraints on 
+Infrastructure-as-a-Service (IaaS) clouds. IaaS clouds are characterized by
+on-demand resource provisioning capabilities and a pay-per-use model.  We discuss,
+develop, and assess algorithms based on static and dynamic strategies for both
+task scheduling and resource provisioning. We assume that the workflows within
+an ensemble are prioritized by the user. We evaluate the algorithms via
+simulation using a set of scientific workflow ensembles with a broad range of
+budget and deadline parameters, identifying which algorithm performs best for 
+different application scenarios and constraint ranges. We found that the key 
+factor determining the performance of an algorithm is its ability to decide which
+workflows in an ensemble to admit or reject for execution. The results
+show that the admission procedure based on workflow structure and estimates of
+task runtimes can significantly improve the quality of solutions.
+
 \end{abstract}
@@
 
-Scientific workflows, usually represented as directed acyclic graphs (DAG), are
-the important class of applications that have been studied in the context of
+Scientific workflows, usually represented as directed acyclic graphs (DAGs), are
+an important class of distributed applications that have been studied in the context of
 resource management and task scheduling on grid and utility computing systems.
-However, large computational applications are often composed of several 
-inter-related workflows grouped into {\em ensembles}. All the
-workflows in an ensemble typically have a similar structure, but they differ by
-input data, number of tasks and individual task sizes. A good example of
-scientific workflow ensemble comes from the CyberShake
-application~\cite{Callaghan11}, which calculates seismic hazards for a given
-geographical region such as California. In order to produce a hazard map, the
-hazard curves need to be computed for a set of geographical sites, while each
-site requires running a single complex scientific workflow. Workflows for each
-site may differ not only in their parameters, but also in their priority; e.g. some of
-the points on the map correspond to urban areas or strategic locations such as
-power plants, whereas others may be located in less populated regions. Another
-example of workflow ensembles comes from astronomical applications, such as
-Montage~\cite{Deelman08} or search for exo-solar planets from Kepler project
-data~\cite{vockler11}, where multiple, different workflows are used to
+Workflow applications for large computational problems are often composed of 
+several inter-related workflows grouped into {\em ensembles}. All the
+workflows in an ensemble typically have a similar structure, but they differ in their
+input data, number of tasks, and individual task sizes. The CyberShake 
+application~\cite{Callaghan11} is a good example of an application that requires 
+scientific workflow ensembles. CyberShake computes seismic hazard curves for a given
+geographical location. In order to produce a hazard map for a region such as 
+Southern California, many hazard curves need to be computed for a large number of 
+geographical sites, where each site requires running a single, complex scientific 
+workflow. Workflows for individual sites may differ not only in their parameters, 
+but also in their priority. For example, some sites may be heavily populated areas 
+or in strategic locations such as power plants, while others may be less important. 
+Other examples of workflow ensembles come from astronomical applications that create
+science-grade image mosaics~\cite{Deelman08} or search for extrasolar 
+planets~\cite{vockler11}, where multiple, different workflows are used to
 process data for different areas of the sky.
@@
 to clusters and grids, which typically offer best-effort quality of service, clouds 
-give users more flexibility to creating a controlled and managed computing environment.
+give users more flexibility in creating a controlled and managed computing environment.
 Clouds provide the ability to adjust resource capacity according to the dynamically changing
@@
 in cloud scenarios not only have to take into account traditional performance-related 
-metrics such as the workflow makespan or the resource utilization, but must also now consider
+metrics such as workflow makespan or resource utilization, but must also now consider
 budget constraints, since the resources from public (commercial) cloud providers 
@@
 In this paper, we aim to gain insight into resource management challenges when executing 
-scientific workflow ensembles on clouds. In particular we address a new important problem of maximizing
-the number of completed workflows from an ensemble under both budget and
-deadline constraints. The motivation is to answer the fundamental question of a
-researcher: How much computation can be completed given the limited budget and 
+scientific workflow ensembles on clouds. In particular, we address a new and important 
+problem of maximizing the number of completed workflows from an ensemble under both budget and
+deadline constraints. The motivation for this work is to answer the fundamental question of 
+concern to a researcher: How much computation can be completed given the limited budget and 
 timeframe of a research project? The goals of this paper are to discuss and assess possible
@@
 workflow structure information to estimate which workflows should be rejected
-from the ensemble due to the constraints. We evaluate the algorithms via simulation. 
-We have developed cloud workflow simulator based on
-CloudSim~\cite{Calheiros11}, which models the infrastructure and the application. The
-algorithms are subject to evaluation on a set of scientific workflow ensembles with a broad range of
-budget and deadline parameters.
-
-The paper is organized as follows: after analysis of related work in
-Section~\ref{sec:related}, Section~\ref{sec:problem} gives the details about the
+from the ensemble based on the constraints. We evaluate these algorithms using a
+simulator we have developed using CloudSim~\cite{Calheiros11}, which models the 
+infrastructure and the application. The algorithms were evaluated on a set 
+of scientific workflow ensembles from several different applications using a 
+broad range of budget and deadline parameters.
+
+The main contributions of this paper are:
+\begin{itemize}
+  \item a definition of the problem of scheduling prioritized workflows grouped into
+  ensembles, under both budget and deadline constraints,
+  \item an analysis of static and dynamic strategies for task scheduling and
+  resource provisioning on clouds, together with admission procedures to
+  predict which workflows from the ensemble can be completed,
+  \item the development of several static and dynamic algorithms and of a simulator based
+  on the defined application and infrastructure model,
+  \item a performance evaluation of the algorithms on a set of synthetic benchmark
+  workflow ensembles based on important, real scientific applications,
+  \item an analysis of performance results that assesses the advantages and
+  disadvantages of the proposed algorithms on different application scenarios and
+  varying constraint values.
+\end{itemize}
+
+The paper is organized as follows: after an analysis of related work in
+Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
 infrastructure and application model we are targeting in this paper.
@@
 static algorithms we developed. Section~\ref{sec:performance} presents test
-scenarios, metrics and results of simulation studies of these algorithms.
-Finally, general conclusions, lessons learned and future work are outlined in
+scenarios, performance metrics, and the results of simulation studies of these algorithms.
+Finally, general conclusions, lessons learned, and future work are outlined in
 Section~\ref{sec:conclusions}.

### 2012-01-23 paper/paper.tex
Update paper/paper.tex

@@
 simulator we have developed using CloudSim~\cite{Calheiros11}, which models the 
-infrastructure and the application. The algorithms were evaluated on a set 
+infrastructure and the application. The simulator includes also tightly-coupled 
+scheduler and provisioner modules. The algorithms were evaluated on a set 
 of scientific workflow ensembles from several different applications using a 
@@
 Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
-Kim2011}) also allow to scale the the cloud resource pool or to extend the
-capabilities of clusters using cloud-burst techniqies. Our approach is different
+Kim2011}) also allow to scale the cloud resource pool or to extend the
+capabilities of clusters using cloud-burst techniques. Our approach is different
 in that we consider workflows, while policy based approaches typically consider
@@
 cloud-targeted autoscaling solution~\cite{Mao2011} in that we consider ensembles
-of workflows rather than unpredictable workloads containing workflows. We also have budget constraints
+of workflows rather than unpredictable workloads containing workflows. We also consider budget constraints
 rather than cost minimization as a goal. In other words, we assume that there is
 more work to be done than the available budget, so some work must be rejected.
-Therefore cost is not something we optimize, but rather a constraint.
+Therefore cost is not something we optimize (i.e. an objective), but rather a constraint.
 
@@
 metaheuristics that usually run for a long time before producing good results,
-which makes them less useful in practice. Our work can be also
+which makes them less useful in the scenarios we consider in this paper. Our work can be also
 regarded as an extension of the budget-constrained workflow scheduling
@@
 come from a wide variety of domains including: bioinformatics (Epigenomics, 
-SIPHT), astronomy (Montage), earthquake science (CyberShake), and physics 
+SIPHT: sRNA identification protocol using high-throughput
+technology), astronomy (Montage), earthquake science (CyberShake), and physics 
 (LIGO).
@@
 workflows may be executed even though there is no chance that those workflows 
-will be completed within the current budget and deadline. Fig.~\ref{fig:spds-example} 
+will be completed within the current budget and deadline. Figure~\ref{fig:spds-example} 
 shows an example schedule generated using the SPDS algorithm. The figure illustrates 
@@
 
-An example schedule generated by SPSS is shown in Fig.~\ref{fig:spss-example}.
+An example schedule generated by SPSS is shown in Figure.~\ref{fig:spss-example}.
 This example shows how SPSS tends to start many workflows in parallel, running
@@
 workflows and a large number of smaller workflows, as seen in 
-Fig.~\ref{fig:ensemble-distribution}. The number of large workflows (of size $\geq$ 900) 
+Figure~\ref{fig:ensemble-distribution}. The number of large workflows (of size $\geq$ 900) 
 is slightly increased to represent the ``heavy-tail'', i.e. that there is a 
@@
 scaling limit to 1.0, ensures that the shape of the resource provisioning 
-plan  is close to a rectangle (see Fig.~\ref{fig:spds-example}) and the VM 
+plan  is close to a rectangle (see Figure~\ref{fig:spds-example}) and the VM 
 termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
@@
 
-Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
+Figure~\ref{fig:number-complete-pareto} shows the number of the workflows
 completed for two budget values for each application ensemble. The budget values
@@
 
-%Fig.~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
+%Figure~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
 %computed for two sample applications. Comparing
-%Fig.~\ref{fig:number-complete-pareto} to Fig.~\ref{fig:score} we can see that
+%Figure~\ref{fig:number-complete-pareto} to Figure~\ref{fig:score} we can see that
 %the latter one gives more flattened shape: this means that e.g. the improvement
@@
 
-Fig.~\ref{fig:total-time} presents the same results using the total computing
+Figure~\ref{fig:total-time} presents the same results using the total computing
 time metric, $T_C$. These plots confirm our previous observations about the
@@
 These observations are further confirmed when examining the effective computing
-cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
+cost, as shown in Figure~\ref{fig:cost}. It can be seen that lower resource
 utilization and lower parallel efficiency for short deadlines lead to
@@
 
-Another general conclusion can be drawn from Fig.~\ref{fig:cost}: despite some
+Another general conclusion can be drawn from Figure~\ref{fig:cost}: despite some
 random fluctuations resulting mostly from the finite result space, all the
@@
 multi-criteria optimization problem, and that the curves presented in 
-Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
+Figure~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
 solutions to this problem. In our case the solutions were achieved by maximizing
@@
 
-Fig.~\ref{fig:budgets} shows the same data for the SIPHT ensemble from another
+Figure~\ref{fig:budgets} shows the same data for the SIPHT ensemble from another
 perspective. It helps answer the question: Given a deadline, how many workflows
-can be finished when we increase the budget? In Fig.~\ref{fig:budgets}(a) the
+can be finished when we increase the budget? In Figure~\ref{fig:budgets}(a) the
 growth is super-linear, which results from the Pareto-like distribution of 
-workflow sizes in the ensemble (see Fig.~\ref{fig:ensemble-distribution}). 
-For ensembles of constant size workflows (see Fig.~\ref{fig:budgets}(b)) 
-the growth is linear and the costs are higher, since all the worklfows in the
+workflow sizes in the ensemble (see Figure~\ref{fig:ensemble-distribution}). 
+For ensembles of constant size workflows (see Figure~\ref{fig:budgets}(b)) 
+the growth is linear and the costs are higher, since all the workflows in the
 test consisted of 1000 tasks.

### 2012-01-23 paper/paper.tex
change ref to line in alg 2
updated data in table 2
added last paragraph

@@
 of resource pool based on metrics related to infrastructure and application.
-A typical infrastructure-specific metric is system load, whereas
-application-specific metrics include response time and length of task or
+A typical infrastruc\-ture-specific metric is system load, whereas
+application-speci\-fic metrics include response time and length of task or
 request queue. It is possible to set thresholds and limits to tune the behavior
@@
 
-Our work is related to the strategies for deadline-constrained cost-minimization
-workflow scheduling, developed for utility grid systems. However, our problem is
+Our work is related to the strategies for deadline-con\-strained
+cost-minimization workflow scheduling, developed for utility grid systems. However, our problem is
 different from \cite{Yu2005, Abrishami2010} in that we consider ensembles of workflows in
@@
 
-\begin{algorithm}
+\begin{algorithm}[tb]
 \caption{Priority-based scheduling algorithm for SPDS}
@@
 
-\begin{algorithm}
+\begin{algorithm}[tb]
 \caption{Dynamic provisioning algorithm for DPDS}
@@
 that would overrun the budget if not terminated in the current provisioning cycle. 
-The VMs terminated on line \ref{l:nT2} are chosen to increase the resource utilization
-to the desired threshold. In order to prevent instances from being
+The VMs terminated in line \ref{l:terminate2} are chosen to increase the resource
+utilization to the desired threshold. In order to prevent instances from being
 terminated too quickly, potentially wasting resources that have already been paid for
@@
 priority queue (i.e. when no other tasks from the workflow have been scheduled yet). 
-The admission procedure---shown in Algorithm~\ref{alg:wa-dpds}---estimates whether there is 
-enough budget remaining to admit the new workflow; if there is not, then the 
+The admission procedure -- shown in Algorithm~\ref{alg:wa-dpds} -- estimates
+whether there is enough budget remaining to admit the new workflow; if there is not, then the 
 workflow is rejected and its tasks are removed from the queue. This algorithm 
@@
 
-\begin{algorithm}
+\begin{algorithm}[tb]
 \caption{Workflow admission algorithm for WA-DPDS}
@@
 
-\begin{algorithm}
+\begin{algorithm}[tb]
 \caption{Ensemble planning algorithm for SPSS}
@@
 $$
+
+\begin{algorithm}[tb]
+\caption{Workflow planning algorithm for SPSS}
+\label{alg:planworkflow}
+\begin{algorithmic}[1]
+\Require $w$: workflow; $P$: current plan; $d$: deadline
+\Ensure Create plan for $w$ that minimizes cost and meets deadline $d$
+\Procedure{PlanWorkflow}{$w,P,d$}
+    \State $P^\prime\gets$ copy of $P$
+    \State \Call{DeadlineDistribution}{w,d}
+    \For{$t\ in\ w\ sorted\ by\ DL(t)$}
+        \State $v \gets$ VM that minimizes cost and start time of t
+        \If{$FinishTime(t,v) < DL(t)$}
+            \State Schedule(t,v)
+        \Else
+            \State Provision a new VM v
+            \State Schedule(t,v)
+        \EndIf
+    \EndFor
+    \State \textbf{return} $P^\prime$
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
@@
 
-\begin{algorithm}
-\caption{Workflow planning algorithm for SPSS}
-\label{alg:planworkflow}
-\begin{algorithmic}[1]
-\Require $w$: workflow; $P$: current plan; $d$: deadline
-\Ensure Create plan for $w$ that minimizes cost and meets deadline $d$
-\Procedure{PlanWorkflow}{$w,P,d$}
-    \State $P^\prime\gets$ copy of $P$
-    \State \Call{DeadlineDistribution}{w,d}
-    \For{$t\ in\ w\ sorted\ by\ DL(t)$}
-        \State $v \gets$ VM that minimizes cost and start time of t
-        \If{$FinishTime(t,v) < DL(t)$}
-            \State Schedule(t,v)
-        \Else
-            \State Provision a new VM v
-            \State Schedule(t,v)
-        \EndIf
-    \EndFor
-    \State \textbf{return} $P^\prime$
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
 An example schedule generated by SPSS is shown in Figure.~\ref{fig:spss-example}.
@@
 
-\begin{figure}[htb] 
+\begin{figure}[tb] 
 \centering
@@
 only a small number of workflows from the ensemble, while the larger budgets 
-enable the algorithms to finish almost all the workflows. The rangesIn three of
+enable the algorithms to finish almost all the workflows. In three of
 five cases the SPSS algorithm performs better for tighter constraints. When the 
@@
 \hline
-DPDS      &   78 &  93 & 161 &  24  & 30 & 386\tabularnewline
-WA-DPDS   &    78  & 97 & 179  & 58  & 28 & 440\tabularnewline
-SPSS     &    78  & 79 & 195 &  75 &  50 & 477\tabularnewline
+DPDS      &   79  & 92  & 170 &  24 &  29 & 394\tabularnewline
+WA-DPDS   &   82  & 100 & 180 &  58 &  27 & 447\tabularnewline
+SPSS      &   72  & 77  & 195 &  75 &  50 & 469 \tabularnewline
 \end{tabular}
@@
 \hline
-DPDS      &   78 &  93 & 161 &  24 &  30 & 386\tabularnewline
-WA-DPDS   &    78  & 97 & 179 &  46  & 28 & 428\tabularnewline
-SPSS     &    78 &  77 & 195  & 75 &  50 & 475\tabularnewline
+DPDS      &   79  & 92  & 170 &  24 &  29 & 394\tabularnewline
+WA-DPDS   &   82  & 100 & 180 &  46 &  27 & 435\tabularnewline
+SPSS      &   72  & 75  & 195 &  75 &  50 & 467 \tabularnewline
 \end{tabular}
@@
 \hline
-DPDS      &   78  & 93 & 161 &  24 &  30 & 386\tabularnewline
-WA-DPDS   &    78 &  97 & 179  & 46  & 28 & 428\tabularnewline
-SPSS     &    78 &  77 & 195 &  75  & 50 & 475\tabularnewline
+DPDS      &   79  & 92  & 170 &  24 &  29 & 394\tabularnewline
+WA-DPDS   &   82  & 100 & 180 &  46 &  27 & 435\tabularnewline
+SPSS      &   72  & 75  & 195 &  75 &  50 & 467 \tabularnewline
 \end{tabular}
@@
 and budgets. Table~\ref{tab:num-dags-pareto} shows these summary values for 
-Pareto-distributed workflows. Note that all algorithms that achieve the best 
+Pareto-distri\-buted workflows. Note that all algorithms that achieve the best 
 result are credited, so the sum for a given application can be greater than the
@@
 
-\section{Conclusion and Future Work}
+\section{Conclusions and Future Work}
 \label{sec:conclusions}
@@
 
+We believe that the results of our study can be applied in practice to develop 
+tools that assist researchers in planning their large-scale computational
+experiments. The estimates of costs, runtimes and number of workflows completed
+that can be obtained from both the static algorithms and from the simulation
+runs, constitute valuable hints for the planning decisions and assessing the
+associated trade-offs. For this purpose our future work includes also further
+extension of our experimental studies~\cite{Juve2009,vockler11} of workflows on
+available cloud infrastructures to develop more accurate performance and cost
+models of the applications. This should lead to better understanding of the
+problems and costs of doing science in the cloud.
+
 %\section{Acknowledgments}

### 2012-01-23 paper/paper.tex
added categories and moved Jarek to the right

@@
        \email{mmalawsk@nd.edu}
+    \alignauthor Gideon Juve and Ewa Deelman\\
+       \affaddr{USC Information Sciences Institute}\\
+       \affaddr{4676 Admiralty Way}\\
+       \affaddr{Marina del Rey, CA, USA}\\
+       \email{\{gideon,deelman\}@isi.edu}
     \alignauthor Jarek Nabrzyski\\
@@
        \email{naber@nd.edu}
-    \alignauthor Gideon Juve and Ewa Deelman\\
-       \affaddr{USC Information Sciences Institute}\\
-       \affaddr{4676 Admiralty Way}\\
-       \affaddr{Marina del Rey, CA, USA}\\
-       \email{\{gideon,deelman\}@isi.edu}
 }
@@
 
-%\category{H.4}{Information Systems Applications}{Miscellaneous}
-%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
-
-%\terms{Theory}
-
-\keywords{Scientific workflows, cloud computing, DAG scheduling, simulation}
+\category{H.3.4}{Systems and Software}{Distributed systems}
+\category{D.4.8}{Performance}{Simulation}
+
+\terms{DAG, Scheduling, Simulation}
+
+\keywords{Scientific workflows, Workflow ensembles,
+Cloud computing, Infrastructure-as-a-Service}
 
@@
 
-SPSS assigns sub-deadlines to each task based on the slack time of the
-workflow, which is defined as the amount of extra time that a workflow
-can extend its critical path and still be completed by the ensemble deadline.
-For a workflow $w$, the slack time of $w$ is:
-
-$$
-ST(w) = d - CP(w)
-$$
-
-where $d$ is the ensemble deadline and $CP(w)$ is the critical path of $w$.
-We assume that $CP(w) \leq d$, otherwise the workflow cannot be completed 
-by the deadline and must be rejected. For large ensembles we expect the 
-critical path of any individual workflow to be much less than the deadline.
-
-A task's level is the length of the longest path between the task and an entry
-task of the workflow:
-
-$$
-Level(t) = \begin{cases}
-0,&$if~$Pred(t) = \emptyset\cr
-\max_{p \in Pred(t)}Level(p)+1,&$otherwise.$
-\end{cases}
-$$
-
-SPSS distributes the slack time of the workflow by level, so that each level
-of the workflow gets a portion of the workflow's slack time proportional to
-the number of tasks in the level and the total runtime of tasks in the level.
-The idea is that levels containing many tasks and large runtimes should be given
-a larger portion of the slack time so that tasks in those levels may be serialized.
-Otherwise, many resources need to be allocated to run all of the tasks in 
-parallel, which may be more costly.
-
-The slack time of a level $l$ in workflow $w$ is given by:
-
-$$
-ST(l) = ST(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
-$$
-
-where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks 
-in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is
-the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between
-0 and 1 that causes more slack time to be given to levels with more tasks (large 
-$\alpha$) or more runtime (small $\alpha$).
-
-The deadline of a task $t$ is then:
-
-$$
-DL(t) = LST(t) + RT(t) + ST(Level(t))
-$$
-
-where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and $LST(t)$ 
-is the latest start time of $t$ determined by:
-
-$$
-LST(t) = \begin{cases}
-0,&$if~$Pred(t) = \emptyset\cr
-\max_{p \in Pred(t)} DL(p),&$otherwise.$
-\end{cases}
-$$
-
 \begin{algorithm}[tb]
@@
 
+SPSS assigns sub-deadlines to each task based on the slack time of the
+workflow, which is defined as the amount of extra time that a workflow
+can extend its critical path and still be completed by the ensemble deadline.
+For a workflow $w$, the slack time of $w$ is:
+
+$$
+ST(w) = d - CP(w)
+$$
+
+where $d$ is the ensemble deadline and $CP(w)$ is the critical path of $w$.
+We assume that $CP(w) \leq d$, otherwise the workflow cannot be completed 
+by the deadline and must be rejected. For large ensembles we expect the 
+critical path of any individual workflow to be much less than the deadline.
+
+A task's level is the length of the longest path between the task and an entry
+task of the workflow:
+
+$$
+Level(t) = \begin{cases}
+0,&$if~$Pred(t) = \emptyset\cr
+\max_{p \in Pred(t)}Level(p)+1,&$otherwise.$
+\end{cases}
+$$
+
+SPSS distributes the slack time of the workflow by level, so that each level
+of the workflow gets a portion of the workflow's slack time proportional to
+the number of tasks in the level and the total runtime of tasks in the level.
+The idea is that levels containing many tasks and large runtimes should be given
+a larger portion of the slack time so that tasks in those levels may be serialized.
+Otherwise, many resources need to be allocated to run all of the tasks in 
+parallel, which may be more costly.
+
+The slack time of a level $l$ in workflow $w$ is given by:
+
+$$
+ST(l) = ST(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
+$$
+
+where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks 
+in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is
+the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between
+0 and 1 that causes more slack time to be given to levels with more tasks (large 
+$\alpha$) or more runtime (small $\alpha$).
+
+The deadline of a task $t$ is then:
+
+$$
+DL(t) = LST(t) + RT(t) + ST(Level(t))
+$$
+
+where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and $LST(t)$ 
+is the latest start time of $t$ determined by:
+
+$$
+LST(t) = \begin{cases}
+0,&$if~$Pred(t) = \emptyset\cr
+\max_{p \in Pred(t)} DL(p),&$otherwise.$
+\end{cases}
+$$
+
 Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for
@@
 schedules tasks onto VMs, allocating new VMs when necessary. For each task 
-in the workflow, the the least expensive slot is chosen to schedule the 
+in the workflow, the least expensive slot is chosen to schedule the 
 task so that it can be completed by  its deadline. VMs are allocated in 
@@
 
+\begin{figure}[tb] 
+\centering
+\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
+\caption{Pareto-like distribution of workflow sizes in ensemble. Size is
+measured in number of tasks.}
+\label{fig:ensemble-distribution}
+\end{figure}
+
 We have prepared test scenarios to evaluate the proposed algorithms under a
@@
 
-\begin{figure}[tb] 
-\centering
-\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
-\caption{Pareto-like distribution of workflow sizes in ensemble. Size is
-measured in number of tasks.}
-\label{fig:ensemble-distribution}
-\end{figure}
-
-The number of workflows in an ensemble depends on the particular application, but
-we assume that ensemble sizes are on the order of between 10 and 100 workflows.
-This range is motivated by several reasons: First, such sizes are typical of the
-real applications we have examined. For example, the number of geographical 
-sites of interest to the users of the CyberShake application is on the order 
-of 100. Second, smaller ensembles consisting of just a few workflows can be 
-aggregated into a single workflow, so there is no need to treat them as an
-ensemble. Similarly, when the number of workflows is larger than 100, and each 
-workflow has a large number of tasks, either the constraints are low enough to
-prevent many of the workflows from running, or the problem of efficiently 
-allocating them to the resources becomes similar to the bag-of-tasks problem, 
-which is easier to solve efficiently.
-
 \begin{figure*}[t] 
@@
 
-%\begin{figure}[b!] 
-%\centering
-%\includegraphics[width=0.19\textwidth]{figures/pareto-score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-%\includegraphics[width=0.19\textwidth]{figures/pareto-score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-%\includegraphics[width=0.19\textwidth]{figures/pareto-score-LIGO-n-1000-8-dagh1-40m0.pdf}
-%\includegraphics[width=0.19\textwidth]{figures/pareto-score-GENOME-n-1000-8-dagh100-1500m0.pdf}
-%\includegraphics[width=0.19\textwidth]{figures/pareto-score-SIPHT-n-1000-8-dagh5-50m0.pdf}
-%\caption{Comparison of algorithms based on score, which  is computed as
-%$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$
-%(0 means highest priority).}
-%\label{fig:score}
-%\end{figure}
+The number of workflows in an ensemble depends on the particular application, but
+we assume that ensemble sizes are on the order of between 10 and 100 workflows.
+This range is motivated by several reasons: First, such sizes are typical of the
+real applications we have examined. For example, the number of geographical 
+sites of interest to the users of the CyberShake application is on the order 
+of 100. Second, smaller ensembles consisting of just a few workflows can be 
+aggregated into a single workflow, so there is no need to treat them as an
+ensemble. Similarly, when the number of workflows is larger than 100, and each 
+workflow has a large number of tasks, either the constraints are low enough to
+prevent many of the workflows from running, or the problem of efficiently 
+allocating them to the resources becomes similar to the bag-of-tasks problem, 
+which is easier to solve efficiently.
 
@@
 Another lesson learned from our studies is the effect of task granularity on
-workflow ensembles. All the benchmark workflows that we tested have an average 
-number of tasks (50--1000), but the Montage and CyberShake workflows have 
-fine-grained tasks with execution times on the order of seconds to minutes, 
-while LIGO, Epigenomics and SIPHT have coarser-grained tasks with runtimes 
-ranging from several minutes to hours. We observed that the dynamic 
-algorithms achieved better results on the fine grained workflows since the 
-simple random scheduling strategy leads to good load-balancing and high 
-resource utilization for these ensembles. On the other hand, when tasks are
-longer, the dynamic algorithms tend to use up the budget by running tasks 
-from low-priority workflows while the more sophisticated static scheduling 
-strategy of SPSS is more circumspect. In the future we plan to investigate 
-the effects of task granularity on ensemble scheduling in a more systematic
-fashion.
+workflow ensembles. The Montage and CyberShake workflows have fine-grained tasks
+with execution times on the order of seconds to minutes, while LIGO, Epigenomics
+and SIPHT have coarser-grained tasks with runtimes ranging from several minutes
+to hours. We observed that the dynamic algorithms achieved better results on the
+fine grained workflows since the simple random scheduling strategy leads to good
+load-balancing and high resource utilization for these ensembles. On the other
+hand, when tasks are longer, the dynamic algorithms tend to use up the budget by
+running tasks from low-priority workflows while the more sophisticated static
+scheduling strategy of SPSS is more circumspect. In the future we plan to
+investigate the effects of task granularity on ensemble scheduling in a more
+systematic fashion.
 

### 2012-01-23 paper/paper.tex
applied Ewa's edits and comments

@@
 simulation using a set of scientific workflow ensembles with a broad range of
-budget and deadline parameters, identifying which algorithm performs best for 
-different application scenarios and constraint ranges. We found that the key 
+budget and deadline parameters, identify which algorithm performs best for 
+different application scenarios and constraint ranges. We find that the key 
 factor determining the performance of an algorithm is its ability to decide which
@@
 workflow. Workflows for individual sites may differ not only in their parameters, 
-but also in their priority. For example, some sites may be heavily populated areas 
+but also in their priority. For example, some sites may be in heavily populated areas 
 or in strategic locations such as power plants, while others may be less important. 
@@
 A typical infrastruc\-ture-specific metric is system load, whereas
-application-speci\-fic metrics include response time and length of task or
-request queue. It is possible to set thresholds and limits to tune the behavior
+application-speci\-fic metrics include response time and length of a task or
+of a request queue. It is possible to set thresholds and limits to tune the behavior
 of these autoscaling systems, but no support for complex applications is provided.
@@
 in that we consider workflows, while policy based approaches typically consider
-bags of independent tasks or unpredictable batch workloads. This enables us to
+bags of independent tasks or unpredictable batch workloads. Our approach enables us to
 take advantage of workflow-aware scheduling heuristics that cannot be applied to
@@
 has exclusive access to a VM instance and that there is no preemption. For this
-study we assumed that there are no delays in VM start and terminate requests.
+study we assumed that the delays in VM start and terminate requests are negligible.
 
@@
 \label{eq:static-plan}
-N_{VM} = \lceil b / d / p \rceil
+N_{VM} = \lceil b / (d * p) \rceil
 \end{equation}
@@
 
-The static provisioning approach of SPDS may not be optimal when resource demand
+The static provisioning approach of SPDS may perform poorly when resource demand
 changes during ensemble execution. In those cases, SPDS may either over-provision, 
@@
     \State $V_C\gets$ set of VMs completing billing cycle
+    \State $V_T\gets \emptyset$ \Comment{set of VMs to terminate}
+    \State $n_T\gets 0$ \Comment{number of VMs to terminate}
     \If{$ b - c < |V_C| * p\ or\ t > d $ }
@@
 
+\medskip
+\medskip
+
 \subsection{Static Provisioning Static Scheduling \\(SPSS)}
@@
 order, rejecting any workflows that cannot be completed by the deadline or that 
-cause the cost of the plan to exceed the budget. Once the plan is complete, then 
+cause the cost of the plan to exceed the budget. Once the plan is complete, 
 the VMs are provisioned and tasks are executed according to the schedules given
@@
 the number of machine-hours that can be provisioned. The deadline, although
-important in terms of scheduling, is assumed to be greater than the critical
-path of every workflow. We can then have a trivial but unrealistic solution
+important in terms of scheduling, is assumed to be greater than the longest critical
+path of all workflows in the ensemble. We can then have a trivial but unrealistic solution
 where the deadline  is long enough to allow all the workflow tasks to be
@@
 aggregated into a single workflow, so there is no need to treat them as an
-ensemble. Similarly, when the number of workflows is larger than 100, and each 
+ensemble. Similarly, when the number of workflows grows, and each 
 workflow has a large number of tasks, either the constraints are low enough to
@@
 
-The constant distribution produces ensembles containing of workflows with the same number of tasks per workflow, 
+The constant distribution produces ensembles containing workflows with the same number of tasks per workflow, 
 while the Pareto-like distribution produces ensembles with a small number of larger 
@@
 is slightly increased to represent the ``heavy-tail'', i.e. that there is a 
-not-insignificant number of large workflows in the ensemble. We assume that large
+significant number of large workflows in the ensemble. We assume that large
 workflows are more important to users, so in the Pareto ensembles the workflows are 
@@
 40 and 60 hours. The Epigenomics workflow is even more costly, requiring 
-approximately \$12,000 and 1500 hours to complete a full ensemble. These
+approximately \$12,000 and 1,500 hours to complete a full ensemble. These
 differences in total runtime and cost come from the fact that we used the
@@
   \item $C_{eff}$: effective cost of computing per hour, which includes
-  overheads due to cost of idle resources.
+  overheads due to the cost of idle resources.
 \end{itemize}
@@
 
-During our work, several interesting lessons were learned about running 
+During our work, several interesting lessons were learned from simulating the execution of
 workflow ensembles on cloud infrastructures. One result is that the simple 
@@
 scheduling strategy of SPSS is more circumspect. In the future we plan to
-investigate the effects of task granularity on ensemble scheduling in a more
+investigate further the effects of task granularity on ensemble scheduling in a more
 systematic fashion.
@@
 available cloud infrastructures to develop more accurate performance and cost
-models of the applications. This should lead to better understanding of the
-problems and costs of doing science in the cloud.
+models of the applications. This should lead to better understanding of theproblems and costs of doing science in the cloud.
 

### 2012-01-23 paper/paper.tex
Jarek's final edits. ready for submission

@@
 
-The goal of this workflow ensemble scheduling and cloud provisioning problem is
-to complete as many high-priority workflows as possible given a fixed budget and
-deadline. Only workflows for which all tasks are finished by the deadline are 
-considered to be complete---partial results are not usable in this model.
+The goal of the workflow ensemble scheduling and cloud provisioning problem
+considered in this paper is to complete as many high-priority workflows as
+possible given a fixed budget and deadline. Only workflows for which all tasks
+are finished by the deadline are considered to be complete---partial results are
+not usable in this model.
 
@@
 (top) and {\em large budget} (bottom). The ranges on $y$ axis are chosen to 
-emphasize the differences the performance of the algorithms.}
+emphasize the differences between the performance of the algorithms.}
 \label{fig:number-complete-pareto}
@@
 Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
-emphasize the differences the performance of the algorithms.}
+emphasize the differences  between  the performance of the algorithms.}
 \label{fig:cost}
@@
 available cloud infrastructures to develop more accurate performance and cost
-models of the applications. This should lead to better understanding of theproblems and costs of doing science in the cloud.
+models of the applications. This should lead to better understanding of the
+problems and costs of doing science in the cloud.
 

### 2012-04-19 paper/paper.tex
Convert to IEEE format

@@
-\documentclass{sig-alternate}
+\documentclass[conference]{IEEEtran}
+
+\usepackage[pdftex]{graphicx}
+% declare the path(s) where your graphic files are
+\graphicspath{{./figures/}}
+% and their extensions so you won't have to specify these with
+% every instance of \includegraphics
+\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\usepackage[cmex10]{amsmath}
 \usepackage{algorithm,algpseudocode}
@@
 
-\conferenceinfo{HPDC}{'12 Delft, The Netherlands}
-% \CopyrightYear{2007} % Allows default copyright year (20XX) to be over-ridden
-% - IF NEED BE. \crdata{0-12345-67-8/90/01}  % Allows default copyright data
-% (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
-
+% can use linebreaks \\ within to get better formatting as desired
 \title{Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds}
-% \subtitle{[Extended Abstract]}
-
-\numberofauthors{3}
-\author{
-    \alignauthor Maciej Malawski\\
-       \affaddr{AGH University of Science and Technology}\\
-       \affaddr{Dept. of Computer Science}\\
-       \affaddr{Krakow, Poland; and}\\
-       \affaddr{University of Notre Dame}\\
-       \affaddr{Center for Research Computing}\\
-       \affaddr{111 ITC, Notre Dame, IN, USA}\\
-       \email{mmalawsk@nd.edu}
-    \alignauthor Gideon Juve and Ewa Deelman\\
-       \affaddr{USC Information Sciences Institute}\\
-       \affaddr{4676 Admiralty Way}\\
-       \affaddr{Marina del Rey, CA, USA}\\
-       \email{\{gideon,deelman\}@isi.edu}
-    \alignauthor Jarek Nabrzyski\\
-       \affaddr{University of Notre Dame}\\
-       \affaddr{Center for Research Computing}\\
-       \affaddr{111 ITC, Notre Dame, IN, USA}\\
-       \email{naber@nd.edu}
-}
+
+% author names and affiliations
+% use a multiple column layout for up to three different
+% affiliations
+\author{\IEEEauthorblockN{Michael Shell}
+  \IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
+Georgia Institute of Technology\\
+Atlanta, Georgia 30332--0250\\
+Email: http://www.michaelshell.org/contact.html}
+\and
+\IEEEauthorblockN{Homer Simpson}
+\IEEEauthorblockA{Twentieth Century Fox\\
+Springfield, USA\\
+Email: homer@thesimpsons.com}
+\and
+\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
+\IEEEauthorblockA{Starfleet Academy\\
+San Francisco, California 96678-2391\\
+Telephone: (800) 555--1212\\
+Fax: (888) 555--1212}}
+
+% ACM Format
+%\numberofauthors{3}
+%\author{
+%    \alignauthor Maciej Malawski\\
+%       \affaddr{AGH University of Science and Technology}\\
+%       \affaddr{Dept. of Computer Science}\\
+%       \affaddr{Krakow, Poland; and}\\
+%       \affaddr{University of Notre Dame}\\
+%       \affaddr{Center for Research Computing}\\
+%       \affaddr{111 ITC, Notre Dame, IN, USA}\\
+%       \email{mmalawsk@nd.edu}
+%    \alignauthor Gideon Juve and Ewa Deelman\\
+%       \affaddr{USC Information Sciences Institute}\\
+%       \affaddr{4676 Admiralty Way}\\
+%       \affaddr{Marina del Rey, CA, USA}\\
+%       \email{\{gideon,deelman\}@isi.edu}
+%    \alignauthor Jarek Nabrzyski\\
+%       \affaddr{University of Notre Dame}\\
+%       \affaddr{Center for Research Computing}\\
+%       \affaddr{111 ITC, Notre Dame, IN, USA}\\
+%       \email{naber@nd.edu}
+%}
 
@@
 
-\category{H.3.4}{Systems and Software}{Distributed systems}
-\category{D.4.8}{Performance}{Simulation}
-
-\terms{DAG, Scheduling, Simulation}
-
-\keywords{Scientific workflows, Workflow ensembles,
-Cloud computing, Infrastructure-as-a-Service}
+% IEEE does not have these
+%\category{H.3.4}{Systems and Software}{Distributed systems}
+%\category{D.4.8}{Performance}{Simulation}
+%\terms{DAG, Scheduling, Simulation}
+%\keywords{Scientific workflows, Workflow ensembles, Cloud computing, Infrastructure-as-a-Service}
 
@@
 
-%\section{Acknowledgments}
+%\section*{Acknowledgments}
 %Acknowledgements go here.
 
-\bibliographystyle{abbrv}
+% ACM style
+%\bibliographystyle{abbrv}
+% IEEE style
+\bibliographystyle{IEEEtran}
+
 \bibliography{paper}

### 2012-04-19 paper/paper.tex
Update authors

@@
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{\IEEEauthorblockN{Michael Shell}
-  \IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
-Georgia Institute of Technology\\
-Atlanta, Georgia 30332--0250\\
-Email: http://www.michaelshell.org/contact.html}
+% IEE Format
+\author{
+  \IEEEauthorblockN{Maciej Malawski}
+  \IEEEauthorblockA{AGH University of Science and Technology\\
+Dept. of Computer Science\\
+Krakow, Poland; and\\
+University of Notre Dame\\
+Center for Research Computing\\
+111 ITC, Notre Dame, IN, USA\\
+Email: mmalawsk@nd.edu}
   \and
-\IEEEauthorblockN{Homer Simpson}
-\IEEEauthorblockA{Twentieth Century Fox\\
-Springfield, USA\\
-Email: homer@thesimpsons.com}
+  \IEEEauthorblockN{Gideon Juve and Ewa Deelman}
+  \IEEEauthorblockA{USC Information Sciences Institute\\
+4676 Admiralty Way\\
+Marina del Rey, CA, USA\\
+Email: \{gideon,deelman\}@isi.edu}
   \and
-\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
-\IEEEauthorblockA{Starfleet Academy\\
-San Francisco, California 96678-2391\\
-Telephone: (800) 555--1212\\
-Fax: (888) 555--1212}}
+  \IEEEauthorblockN{Jarek Nabrzyski}
+  \IEEEauthorblockA{University of Notre Dame\\
+Center for Research Computing\\
+111 ITC, Notre Dame, IN, USA\\
+Email: naber@nd.edu}}
 

### 2012-04-19 paper/paper.tex
Add hyperref package
Title
Clean up figures

@@
 \usepackage[pdftex]{graphicx}
-% declare the path(s) where your graphic files are
 \graphicspath{{./figures/}}
-% and their extensions so you won't have to specify these with
-% every instance of \includegraphics
 \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
@@
 \usepackage{algorithm,algpseudocode}
+\usepackage{hyperref} % must come before url
 \usepackage{url}
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds}
-
-% IEE Format
+\title{Cost- and Deadline-Constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds}
+
+% IEEE Format
 \author{
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/spds-gantt}
+\includegraphics[width=1.0\columnwidth]{spds-gantt}
  \caption{Example schedule generated using SPDS algorithm. The colors indicate how various tasks from different workflows are scheduled onto Virtual Machines. }
@@
 \centering
-\includegraphics[width=1.0\columnwidth]{figures/spss-gantt}
+\includegraphics[width=1.0\columnwidth]{spss-gantt}
  \caption{Example of schedule generated using SPSS algorithm. Tasks are colored by workflow. }
@@
 \centering
-\includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
+\includegraphics[width=0.6\columnwidth]{ensemble-pareto}
 \caption{Pareto-like distribution of workflow sizes in ensemble. Size is
@@
 \centering
-\includegraphics[width=0.19\textwidth]{figures/pareto-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\includegraphics[width=0.19\textwidth]{pareto-MONTAGE-n-1000-8-dagh1-20m0}
+\includegraphics[width=0.19\textwidth]{pareto-CYBERSHAKE-n-1000-8-dagh1-20m0}
+\includegraphics[width=0.19\textwidth]{pareto-LIGO-n-1000-8-dagh1-40m0}
+\includegraphics[width=0.19\textwidth]{pareto-GENOME-n-1000-8-dagh100-1500m0}
+\includegraphics[width=0.19\textwidth]{pareto-SIPHT-n-1000-8-dagh5-50m0}
 \caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
@@
 \centering
-\includegraphics[width=0.19\textwidth]{figures/pareto-size-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-size-LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-size-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-size-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\includegraphics[width=0.19\textwidth]{pareto-size-MONTAGE-n-1000-8-dagh1-20m0}
+\includegraphics[width=0.19\textwidth]{pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0}
+\includegraphics[width=0.19\textwidth]{pareto-size-LIGO-n-1000-8-dagh1-40m0}
+\includegraphics[width=0.19\textwidth]{pareto-size-GENOME-n-1000-8-dagh100-1500m0}
+\includegraphics[width=0.19\textwidth]{pareto-size-SIPHT-n-1000-8-dagh5-50m0}
 \caption{ Total work completed expressed as the sum of runtimes (in hours) of
@@
 \centering
-\includegraphics[width=0.19\textwidth]{figures/pareto-cost-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-cost-LIGO-n-1000-8-dagh1-40m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-cost-GENOME-n-1000-8-dagh100-1500m0.pdf}
-\includegraphics[width=0.19\textwidth]{figures/pareto-cost-SIPHT-n-1000-8-dagh5-50m0.pdf}
+\includegraphics[width=0.19\textwidth]{pareto-cost-MONTAGE-n-1000-8-dagh1-20m0}
+\includegraphics[width=0.19\textwidth]{pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0}
+\includegraphics[width=0.19\textwidth]{pareto-cost-LIGO-n-1000-8-dagh1-40m0}
+\includegraphics[width=0.19\textwidth]{pareto-cost-GENOME-n-1000-8-dagh100-1500m0}
+\includegraphics[width=0.19\textwidth]{pareto-cost-SIPHT-n-1000-8-dagh5-50m0}
 \caption{Effective computing cost in \$ per hour calculated by dividing the sum of
@@
 % \begin{figure*}[htb] \centering
-% \includegraphics[width=0.19\textwidth]{figures/score2-MONTAGE-n-1000-8-dagh1-20m0.pdf}
-% \includegraphics[width=0.19\textwidth]{figures/score2-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
-% \includegraphics[width=0.19\textwidth]{figures/score2-LIGO-n-1000-8-dagh1-40m0.pdf}
-% \includegraphics[width=0.19\textwidth]{figures/score2-GENOME-n-1000-8-dagh100-1500m0.pdf}
-% \includegraphics[width=0.19\textwidth]{figures/score2-SIPHT-n-1000-8-dagh5-50m0.pdf}
+% \includegraphics[width=0.19\textwidth]{score2-MONTAGE-n-1000-8-dagh1-20m0}
+% \includegraphics[width=0.19\textwidth]{score2-CYBERSHAKE-n-1000-8-dagh1-20m0}
+% \includegraphics[width=0.19\textwidth]{score2-LIGO-n-1000-8-dagh1-40m0}
+% \includegraphics[width=0.19\textwidth]{score2-GENOME-n-1000-8-dagh100-1500m0}
+% \includegraphics[width=0.19\textwidth]{score2-SIPHT-n-1000-8-dagh5-50m0}
 % \caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
@@
 \begin{figure}[htb] \centering
-\includegraphics[width=0.7\columnwidth]{figures/pareto-budget-SIPHT-n-1000-8-dagh5-50h4m0.pdf}\\
+\includegraphics[width=0.7\columnwidth]{pareto-budget-SIPHT-n-1000-8-dagh5-50h4m0}\\
 (a) ensemble of Pareto-distributed workflows
-\includegraphics[width=0.7\columnwidth]{figures/constant-budget-SIPHT-n-1000-0-dagh5-50h4m0.pdf}\\
+\includegraphics[width=0.7\columnwidth]{constant-budget-SIPHT-n-1000-0-dagh5-50h4m0}\\
 (b) ensemble of workflows of equal size \caption{Comparison of number of

### 2012-04-19 paper/paper.tex
Add text from Gideon's thesis

@@
 \usepackage{url}
+\usepackage{color} % for \textcolor
+\usepackage{subfig} % for \subfloat
+
+\newcommand{\TODO}[1]{
+  {\Large \textcolor{red}{\textbf{TODO: }#1}}
+}
 
@@
 input data, number of tasks, and individual task sizes. The CyberShake 
-application~\cite{Callaghan11} is a good example of an application that requires 
+application~\cite{Callaghan2011} is a good example of an application that requires 
 scientific workflow ensembles. CyberShake computes seismic hazard curves for a given
@@
 Other examples of workflow ensembles come from astronomical applications that create
-science-grade image mosaics~\cite{Deelman08} or search for extrasolar 
-planets~\cite{vockler11}, where multiple, different workflows are used to
+science-grade image mosaics~\cite{Deelman2008} or search for extrasolar 
+planets~\cite{Vockler2011}, where multiple, different workflows are used to
 process data for different areas of the sky.
@@
 are regarded by the scientific community as a potentially attractive source of 
-low-cost, on-demand computing resources~\cite{Ostermann10,Keahey09}. In contrast 
+low-cost, on-demand computing resources~\cite{Ostermann2010, Keahey2009}. In contrast 
 to clusters and grids, which typically offer best-effort quality of service, clouds 
@@
 budget constraints, since the resources from public (commercial) cloud providers 
-usually have monetary costs associated with them~\cite{Durkee10}.
+usually have monetary costs associated with them~\cite{Durkee2010}.
 
@@
 from the ensemble based on the constraints. We evaluate these algorithms using a
-simulator we have developed using CloudSim~\cite{Calheiros11}, which models the 
+simulator we have developed using CloudSim~\cite{Calheiros2011}, which models the 
 infrastructure and the application. The simulator includes also tightly-coupled 
@@
 This work is related to bi-criteria scheduling and multi-criteria scheduling of
-workflows~\cite{Wieczorek2009,Prodan10,Dongarra2007}. These approaches are
+workflows~\cite{Wieczorek2009,Prodan2010,Dongarra2007}. These approaches are
 similar to ours in that we have two scheduling criteria: cost and makespan. The
@@
 
-We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2). In
-this model, virtual machine (VM) instances are requested on-demand. The VMs are
-billed by the hour, with partial hours being rounded up. Although there are
-multiple VM types with different amounts of CPU, memory, disk space, and I/O,
-for this paper we focus on a single VM type because we assume that for most
-applications there will typically be one VM type with the best price/performance
-ratio for the application~\cite{Juve2009}. We also assume that a submitted task
-has exclusive access to a VM instance and that there is no preemption. For this
-study we assumed that the delays in VM start and terminate requests are negligible.
+We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2).
+In this model, virtual machine (VM) instances are requested on-demand.
+The VMs are billed by the hour, with partial hours being rounded up.
+Although there are multiple VM types with different amounts of CPU, memory, disk space, and I/O, for this thesis we focus on a single VM type because we assume that for most applications there will typically be only one or two VM types with the best price/performance ratio for the application~\cite{Juve2009}.
+We also assume that a submitted task has exclusive access to a VM instance and that there is no preemption.
+
+We assume that there is a delay between the time that a new VM instance is requested and when it becomes available to execute tasks.
+In practice this delay is typically on the order of tens of seconds to a few minutes, and is highly dependent upon the cloud platform and the VM image size~\cite{Nurmi2008b}.
 
 \subsection{Application Model}
-The target applications for this paper are scientific workflows that can be
-modeled as Directed Acyclic Graphs (DAGs), where the nodes in the graph
-represent computational tasks, and the edges represent data- or control-flow
-dependencies between the tasks.
-
-We assume that we have runtime estimates for each task in the workflow based on
-either a performance model of the application, or historical data that can be
-mined. In this study we use synthetic workflows that were generated using 
-historical data from real applications~\cite{Bharathi08}. These applications 
-come from a wide variety of domains including: bioinformatics (Epigenomics, 
-SIPHT: sRNA identification protocol using high-throughput
-technology), astronomy (Montage), earthquake science (CyberShake), and physics 
-(LIGO).
-
-Although scientific workflows are often data-intensive, we do not currently
-consider the size of input and output data when scheduling tasks. Instead we
-assume that all workflow data is stored in a shared cloud storage system, such
-as Amazon S3, and that data transfer times are included in task runtime
-estimates. We also assume that data transfer times between the shared storage
-and the VMs are equal for different VMs so that task placement decisions do not
-have an impact on the runtime of the tasks.
-
-This paper focuses on scheduling and provisioning for workflow ensembles. A
-workflow ensemble consists of several related workflows that a user wishes to
-run. Each workflow in an ensemble is given a priority that indicates how
-important the workflow is to the user. These priorities are absolute in the 
-sense that completing a high priority workflow is more valuable than 
-completing all lower priority workflows combined.
-
-The goal of the workflow ensemble scheduling and cloud provisioning problem
-considered in this paper is to complete as many high-priority workflows as
-possible given a fixed budget and deadline. Only workflows for which all tasks
-are finished by the deadline are considered to be complete---partial results are
-not usable in this model.
+
+The target applications are ensembles of scientific workflows that can be modeled as Directed Acyclic Graphs (DAGs), where the nodes in the graph represent computational tasks, and the edges represent data- or control-flow dependencies between the tasks.
+
+We assume that there are runtime estimates for each task in the workflow based on either a performance model of the application, or historical data that can be mined.
+Additionally, we assume that the runtime estimates for individual workflow tasks are not perfect, but may vary based on a uniform distribution of $\pm p\%$.
+
+This study uses synthetic workflows that were generated using historical data from real applications~\cite{Bharathi2008}.
+The applications come from a wide variety of domains including: bioinformatics (Epigenomics, SIPHT: sRNA identification protocol using high-throughput technology), astronomy (Montage), earthquake science (CyberShake), and physics (LIGO).
+The runtimes used to generate the synthetic workflows are based on distributions gathered from running real workflows.
+The synthetic workflows were generated using code developed by Bharathi, et al~\cite{WorkflowGenerator}.
+
+Although scientific workflows are often data-intensive, the algorithms described in the next section do not currently consider the size of input and output data when scheduling tasks.
+Instead it is assumed that all workflow data is stored in a shared cloud storage system, such as Amazon S3, and that intermediate data transfer times are included in task runtime estimates.
+It is also assumed that data transfer times between the shared storage and the VMs are equal for different VMs so that task placement decisions do not have an impact on the runtime of the tasks.
+
+This work focuses on scheduling and provisioning for workflow ensembles.
+A workflow ensemble consists of several related workflows that a user wishes to run.
+Each workflow in an ensemble is given a numeric priority that indicates how important the workflow is to the user.
+As such, the priorities indicate the utility function of the user.
+These priorities are absolute in the sense that completing a workflow with a given priority is more valuable (gives the user more utility) than completing all other workflows in the ensemble with lower priorities combined.
+
+The goal of the workflow ensemble scheduling and cloud provisioning problem is to complete as many high-priority workflows as possible given a fixed budget and deadline.
+Only workflows for which all tasks are finished by the deadline are considered to be complete---partial results are not usable in this model.
 
@@
 
-In this section we describe several algorithms we developed to execute ensembles of
-workflows on the cloud under budget and deadline constraints.
-
-\subsection{Static Provisioning Dynamic Scheduling \\
-(SPDS) }
+This section describes several algorithms were developed to schedule and provision resources for ensembles of workflows on the cloud under budget and deadline constraints.
+
+\subsection{Static Provisioning Dynamic Scheduling (SPDS) }
 \label{sec:spds}
 
-The simplest strategy for executing ensembles of workflows in the cloud is to
-provision resources statically and schedule workflow tasks dynamically. Given a 
-budget $b$, deadline $d$, and the hourly price $p$ of a VM, it is easy to calculate
-the number of VMs, $N_{VM}$, to provision so that the entire budget is consumed before
-the deadline:
-
+The simplest strategy for executing ensembles of workflows in the cloud is to provision resources statically and schedule workflow tasks dynamically.
+Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a VM in dollars $p$, it is easy to calculate the number of VMs, $N_{VM}$, to provision so that the entire budget is consumed before the deadline:
+%
 \begin{equation}
@@
 \end{equation}
-
-The SPDS algorithm statically provisions $N_{VM}$ VMs at the start of the ensemble execution 
-and keeps them running until the deadline is reached, or the budget runs out.
-This provisioning plan has the advantage that it minimizes the number of provisioning 
-and de-provisioning requests.
-
-Once the VMs are provisioned, the tasks are mapped onto idle VMs using the dynamic
-priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
+%
+The SPDS algorithm statically provisions $N_{VM}$ VMs at the start of the ensemble execution and keeps them running until the deadline is reached, or the budget runs out.
+This provisioning plan has the advantage that it minimizes the number of provisioning and de-provisioning requests.
+
+Once the VMs are provisioned, the tasks are mapped onto idle VMs using the dynamic priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
 
@@
 
-Initially, the ready tasks from all workflows in the ensemble are added to a 
-priority queue based on the priority of the workflow to which they belong. If
-there are idle VMs available, and the priority queue is not empty, the next task
-from the priority queue is submitted to a randomly chosen idle VM. The process is 
-repeated until there are no idle VMs or the priority queue is empty. The scheduler
-then waits for a task to finish, adds its ready children to the priority queue,
-marks the VM as idle, and the entire process repeats until the deadline is reached.
-
-This algorithm guarantees that tasks from lower priority workflows are
-always deferred when higher-priority tasks are available, but lower-priority
-tasks can still occupy idle VMs when higher-priority tasks are not available. 
-However, because there is no preemption, long-running low-priority tasks may delay 
-the execution of higher-priority tasks. In addition, tasks from low priority 
-workflows may be executed even though there is no chance that those workflows 
-will be completed within the current budget and deadline. Figure~\ref{fig:spds-example} 
-shows an example schedule generated using the SPDS algorithm. The figure illustrates 
-how tasks from lower priority workflows backfill idle VMs when tasks from higher 
-priority workflows are not available.
-
-\begin{figure}[htb] 
+Initially, the ready tasks from all workflows in the ensemble are added to a priority queue based on the priority of the workflow to which they belong.
+If there are idle VMs available, and the priority queue is not empty, the next task from the priority queue is submitted to an arbitrarily chosen idle VM.
+The process is repeated until there are no idle VMs or the priority queue is empty.
+The scheduler then waits for a task to finish, adds its ready children to the priority queue, marks the VM as idle, and the entire process repeats until the deadline is reached.
+
+This algorithm guarantees that tasks from lower priority workflows are always deferred when higher-priority tasks are available, but lower-priority tasks can still occupy idle VMs when higher-priority tasks are not available.
+However, because there is no preemption, long-running low-priority tasks may delay the execution of higher-priority tasks.
+In addition, tasks from low priority workflows may be executed even though there is no chance that those workflows will be completed within the current budget and deadline.
+Figure~\ref{fig:spds-example} shows an example schedule generated using the SPDS algorithm.
+The figure illustrates how tasks from lower priority workflows backfill idle VMs when tasks from higher priority workflows are not available.
+
+\begin{figure}[tb]
   \centering
-\includegraphics[width=1.0\columnwidth]{spds-gantt}
- \caption{Example schedule generated using SPDS algorithm. The colors indicate how various tasks from different workflows are scheduled onto Virtual Machines. }
+  \includegraphics[width=0.5\textwidth]{figures/spds-gantt}
+  \caption[Example schedule generated by the SPDS algorithm.]{Example schedule generated by the SPDS algorithm. Each row is a different VM. Tasks are represented as boxes and are colored by workflow.}
   \label{fig:spds-example}
@@
 
-\subsection{Dynamic Provisioning Dynamic \\Scheduling (DPDS)}
-
-The static provisioning approach of SPDS may perform poorly when resource demand
-changes during ensemble execution. In those cases, SPDS may either over-provision, 
-wasting the budget on idle VMs, or under-provision, using up the deadline while starving 
-the application of usable VMs. One of the advantages of elasticity in cloud computing 
-is that the number of resources can be changed to meet the demands of an application.
-In policy-based provisioning (or auto-scaling) systems, metrics such as resource 
-utilization or task queue length are monitored to estimate application demand.
-The DPDS algorithm uses this approach to determine when to provision and de-provision VMs. The 
-algorithm periodically computes resource utilization as the percentage of idle VMs, and
-adjusts the number of VMs if the utilization is above or below given thresholds.
-Because it is assumed that VMs are billed by the hour, DPDS only considers VMs that are
-approaching their hourly billing cycle when deciding which VMs to terminate. This process
-is described in Algorithm~\ref{alg:prov}.
+\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
+
+The static provisioning approach of SPDS may perform poorly when resource demand changes during ensemble execution.
+In those cases, SPDS may either over-provision, wasting the budget on idle VMs, or under-provision, using up the deadline while starving the application of usable VMs.
+One of the advantages of elasticity in cloud computing is that the number of resources can be changed to meet the demands of an application.
+In policy-based provisioning (or auto-scaling) systems, metrics such as resource utilization or task queue length are monitored to estimate application demand.
+This is the approach used by the Dynamic Provisioning Dynamic Scheduling (DPDS) algorithm to determine when to provision and de-provision VMs.
+The algorithm periodically computes resource utilization using the percentage of idle VMs over time, and adjusts the number of VMs if the utilization is above or below given thresholds.
+Because it is assumed that VMs are billed by the hour, DPDS only considers VMs that are approaching their hourly billing cycle when deciding which VMs to terminate.
+The DPDS algorithm is shown in Algorithm~\ref{alg:prov}.
 
@@
 
-The set of VMs completing their billing cycle is determined by considering both the
-provisioner interval, and the termination delay of the provider. This guarantees 
-that VMs can be terminated before they start the next billing cycle and prevents the
-budget from being overrun. The VMs terminated in line \ref{l:terminate1} of Algorithm~\ref{alg:prov} are the ones 
-that would overrun the budget if not terminated in the current provisioning cycle. 
-The VMs terminated in line \ref{l:terminate2} are chosen to increase the resource
-utilization to the desired threshold. In order to prevent instances from being
-terminated too quickly, potentially wasting resources that have already been paid for
-but could be used later, no more than half of the idle resources are terminated during
-each provisioning cycle. To avoid an uncontrolled increase in the number of
-instances, which may happen in the case of highly parallel workflows, the provisioner 
-will not start a new VM if the number of running VMs is greater than the product of
-$N_{VM}$ and an autoscaling factor, $v_{max}$ that has to be set manually.
+The set of VMs completing their billing cycle is determined by considering both the provisioner interval, and the termination delay of the provider.
+This guarantees that VMs can be terminated before they start the next billing cycle and prevents the budget from being overrun.
+The VMs terminated in line \ref{l:terminate1} of Algorithm~\ref{alg:prov} are the ones that would overrun the budget if not terminated in the current provisioning cycle.
+The VMs terminated in line \ref{l:terminate2} are chosen to increase the resource utilization to the desired threshold.
+In order to prevent instances from being terminated too quickly, potentially wasting resources that have already been paid for but could be used later, no more than half of the idle resources are terminated during each provisioning cycle.
+To avoid an uncontrolled increase in the number of instances, which may happen in the case of highly parallel workflows, the provisioner will not start a new VM if the number of running VMs is greater than the product of $N_{VM}$ (from Equation~\ref{eq:static-plan}) and an autoscaling parameter, $v_{max}$.
+Unless otherwise specified, $v_{max}$ is assumed to be 1.
 
@@
 
-The DPDS algorithm does not use any information about the structure of the workflows
-in the ensemble when scheduling tasks: it looks only at the priorities of the ready 
-tasks when deciding which task to schedule next. It does not consider whether a lower 
-priority task belongs to a workflow that will never be able to complete given the 
-current budget and deadline. As a result, DPDS may execute lower priority tasks just 
-to keep VMs busy and that will end up delaying higher priority tasks, making it less likely
-that higher priority workflows will be able to finish.
-
-In order to address this issue, the Workflow-Aware DPDS (WA-DPDS) algorithm extends 
-DPDS by introducing a workflow admission procedure. The admission procedure is
-invoked whenever WA-DPDS sees the first task of a new workflow at the head of the 
-priority queue (i.e. when no other tasks from the workflow have been scheduled yet). 
-The admission procedure -- shown in Algorithm~\ref{alg:wa-dpds} -- estimates
-whether there is enough budget remaining to admit the new workflow; if there is not, then the 
-workflow is rejected and its tasks are removed from the queue. This algorithm 
-compares the current cost (consumed budget) and remaining budget, taking
-into account the cost of currently running VMs, and the cost of workflows 
-that have already been admitted. It adds a small safety margin to  avoid going over
-the budget.
+The DPDS algorithm does not use any information about the structure of the workflows in the ensemble when scheduling tasks: it looks only at the priorities of the ready tasks when deciding which task to schedule next.
+It does not consider whether a lower priority task belongs to a workflow that will never be able to complete given the current budget and deadline.
+As a result, DPDS may start lower priority tasks just to keep VMs busy that will end up delaying higher priority tasks later on, making it less likely that higher priority workflows will be able to finish.
+
+In order to address this issue, the Workflow-Aware DPDS (WA-DPDS) algorithm extends DPDS by introducing a workflow admission procedure.
+The admission procedure is invoked whenever WA-DPDS sees the first task of a new workflow at the head of the priority queue (i.e.
+when no other tasks from the workflow have been scheduled yet).
+The admission procedure---shown in Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough budget remaining to admit the new workflow; if there is not, then the workflow is rejected and its tasks are removed from the queue.
+This algorithm compares the current cost (consumed budget) and remaining budget, taking into account the cost of currently running VMs, and the cost of workflows that have already been admitted.
+In addition, it adds a small safety margin of \$0.10 to avoid going over the budget.
 
@@
 
-This admission procedure relies only on the total estimated resource consumption and
-compares it to the remaining budget. We found that this estimation is useful not
-only to prevent low-priority workflows from delaying high-priority ones, but
-also to reject large and costly workflows that would overrun the budget and
-admit smaller workflows that can efficiently utilize idle resources in ensembles 
-containing workflows of non-uniform sizes. It would also be possible to extend this 
-admission procedure to check other constraints, such as whether the estimated 
-critical path of the new workflow exceeds the time remaining until the deadline.
-
-\medskip
-\medskip
-
-\subsection{Static Provisioning Static Scheduling \\(SPSS)}
-
-The previous algorithms are all online algorithms that make provisioning 
-and scheduling decisions at runtime. In comparison, the SPSS algorithm 
-creates a provisioning and scheduling plan before running any workflow tasks.
-This enables SPSS to start only those workflows that it knows can be completed
-given the deadline and budget constraints, and eliminates any waste that may
-be allowed by the dynamic algorithms.
-
-The approach used by SPSS is to plan each workflow in the ensemble in priority 
-order, rejecting any workflows that cannot be completed by the deadline or that 
-cause the cost of the plan to exceed the budget. Once the plan is complete, 
-the VMs are provisioned and tasks are executed according to the schedules given
-by the plan. However, in this paper we only consider a deterministic environment and task times.
-
-The disadvantage of the static planning approach used by SPSS is that it is 
-sensitive to dynamic changes in the environment and the application that may 
-disrupt the carefully constructed plan.
-
-Algorithm~\ref{alg:admit} shows how ensembles are planned in SPSS. The algorithm
-starts with an empty plan containing no VMs and no scheduled tasks. Workflows
-from the ensemble are considered in priority order. For each workflow, SPSS
-attempts to build on top of the current plan by provisioning VMs to schedule 
-the tasks of the workflow so that it finishes before the deadline with the least
-possible cost. If the cost of the new plan is less than the budget, then the 
-new plan is accepted and the workflow is admitted. If not, then the new plan is 
-rejected and the process continues with the next workflow in the ensemble. The 
-idea behind this algorithm is that, if each workflow can be completed by the 
-deadline with the lowest possible cost, then the number of workflows that can 
-be completed within the given budget will be maximized.
+This admission procedure relies only on the total estimated resource consumption and compares it to the remaining budget.
+We found that this estimation is useful not only to prevent low-priority workflows from delaying high-priority ones, but also to reject large and costly workflows that would overrun the budget and admit smaller workflows that can efficiently utilize idle resources in ensembles containing workflows of non-uniform sizes.
+It would also be possible to extend this admission procedure to check other constraints, such as whether the estimated critical path of the new workflow exceeds the time remaining until the deadline.
+
+\subsection{Static Provisioning Static Scheduling (SPSS)}
+
+The previous algorithms are all online algorithms that make provisioning and scheduling decisions at runtime.
+In comparison, the SPSS algorithm creates a provisioning and scheduling plan before running any workflow tasks.
+This enables SPSS to start only those workflows that it knows can be completed given the deadline and budget constraints, and eliminates any waste that may be allowed by the dynamic algorithms.
+
+The approach used by SPSS is to plan each workflow in the ensemble in priority order, rejecting any workflows that cannot be completed by the deadline or that cause the cost of the plan to exceed the budget.
+Once the plan is complete, the VMs are provisioned and tasks are executed according to the schedules given by the plan.
+
+The disadvantage of the static planning approach used by SPSS is that it is sensitive to dynamic changes in the environment and the application that may disrupt its carefully constructed plan.
+For example, if there are provisioning delays, or if the runtime estimates for the tasks are inaccurate, then workflow execution may diverge from the plan.
+This issue will be discussed further in Sections~\ref{sec:delays} and \ref{sec:variances}.
+
+Algorithm~\ref{alg:admit} shows how ensembles are planned in SPSS.
+The algorithm starts with an empty plan containing no VMs and no scheduled tasks.
+Workflows from the ensemble are considered in priority order.
+For each workflow, SPSS attempts to build on top of the current plan by provisioning VMs to schedule the tasks of the workflow so that it finishes before the deadline with the least possible cost.
+If the cost of the new plan is less than the budget, then the new plan is accepted and the workflow is admitted.
+If not, then the new plan is rejected and the process continues with the next workflow in the ensemble.
+The idea behind this algorithm is that, if each workflow can be completed by the deadline with the lowest possible cost, then the number of workflows that can be completed within the given budget will be maximized.
 
@@
 
-In order to plan a workflow, the SPSS algorithm assigns 
-sub-deadlines to each individual task in the workflow, and then schedules 
-each task so as to minimize the cost of the task while still meeting its 
-assigned sub-deadline. The idea is that if each task can be completed by 
-its deadline in the least expensive manner possible, then the cost of the
-entire workflow can be minimized without exceeding the ensemble deadline.
+In order to plan a workflow, the SPSS algorithm assigns sub-deadlines to each individual task in the workflow, and then schedules each task so as to minimize the cost of the task while still meeting its assigned sub-deadline.
+The idea is that if each task can be completed by its deadline in the least expensive manner possible, then the cost of the entire workflow can be minimized without exceeding the ensemble deadline.
 
@@
 
-SPSS assigns sub-deadlines to each task based on the slack time of the
-workflow, which is defined as the amount of extra time that a workflow
-can extend its critical path and still be completed by the ensemble deadline.
+SPSS assigns sub-deadlines to each task based on the slack time of the workflow, which is defined as the amount of extra time that a workflow can extend its critical path and still be completed by the ensemble deadline.
 For a workflow $w$, the slack time of $w$ is:
-
+%
 $$
@@
 $$
-
+%
 where $d$ is the ensemble deadline and $CP(w)$ is the critical path of $w$.
-We assume that $CP(w) \leq d$, otherwise the workflow cannot be completed 
-by the deadline and must be rejected. For large ensembles we expect the 
-critical path of any individual workflow to be much less than the deadline.
-
-A task's level is the length of the longest path between the task and an entry
-task of the workflow:
-
+We assume that $CP(w) \leq d$, otherwise the workflow cannot be completed by the deadline and must be rejected.
+For large ensembles we expect the critical path of any individual workflow to be much less than the deadline.
+
+A task's level is the length of the longest path between the task and an entry task of the workflow:
+%
 $$
@@
 $$
-
-SPSS distributes the slack time of the workflow by level, so that each level
-of the workflow gets a portion of the workflow's slack time proportional to
-the number of tasks in the level and the total runtime of tasks in the level.
-The idea is that levels containing many tasks and large runtimes should be given
-a larger portion of the slack time so that tasks in those levels may be serialized.
-Otherwise, many resources need to be allocated to run all of the tasks in 
-parallel, which may be more costly.
+%
+SPSS distributes the slack time of the workflow by level, so that each level of the workflow gets a portion of the workflow's slack time proportional to the number of tasks in the level and the total runtime of tasks in the level.
+The idea is that levels containing many tasks and large runtimes should be given a larger portion of the slack time so that tasks in those levels may be serialized.
+Otherwise, many resources need to be allocated to run all of the tasks in parallel, which may be more costly.
 
 The slack time of a level $l$ in workflow $w$ is given by:
-
+%
 $$
@@
 $$
-
-where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks 
-in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is
-the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between
-0 and 1 that causes more slack time to be given to levels with more tasks (large 
-$\alpha$) or more runtime (small $\alpha$).
+%
+where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between 0 and 1 that causes more slack time to be given to levels with more tasks (large $\alpha$) or more runtime (small $\alpha$).
 
 The deadline of a task $t$ is then:
-
+%
 $$
@@
 $$
-
-where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and $LST(t)$ 
-is the latest start time of $t$ determined by:
-
+%
+where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and $LST(t)$ is the latest start time of $t$ determined by:
+%
 $$
@@
 
-Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for
-each workflow. The \textsc{PlanWorkflow} procedure first calls 
-\textsc{DeadlineDistribution} to assign sub-deadlines to tasks according
-to the procedure described above. Then, the \textsc{PlanWorkflow} procedure
-schedules tasks onto VMs, allocating new VMs when necessary. For each task 
-in the workflow, the least expensive slot is chosen to schedule the 
-task so that it can be completed by  its deadline. VMs are allocated in 
-blocks of one billing cycle (one hour) regardless of the size of the task. 
-When computing the cost of scheduling a task on a given VM, the algorithm 
-considers idle slots in blocks that were allocated for previous tasks to be 
-free, while slots in new blocks cost the full price of a billing cycle. For 
-example, if a task has a runtime of 10 minutes, and the price of a block is 
-\$1, then the algorithm will either schedule the task on 
-an existing VM that has an idle slot larger than 10 minutes for a cost of \$0, 
-or it will allocate a new block on an existing VM, or provision a new VM, for 
-a cost of \$1. If the cost of slots on two different VMs is equal, then the 
-slot with the earliest start time is chosen. To prevent too many VMs from 
-being provisioned, the algorithm always prefers to extend the runtime of 
-existing VMs before allocating new VMs. The result of this is that the 
-algorithm will only allocate a new block if there are no idle slots on 
-existing blocks large enough or early enough to complete the 
-task by its deadline, and it will only allocate a new VM if it cannot add 
-a block to an existing VM to complete the task by its deadline.
+Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for each workflow.
+The \textsc{PlanWorkflow} procedure first calls \textsc{DeadlineDistribution} to assign sub-deadlines to tasks according to the procedure described above.
+Then, the \textsc{PlanWorkflow} procedure schedules tasks onto VMs, allocating new VMs when necessary.
+For each task in the workflow, the least expensive slot is chosen to schedule the task so that it can be completed by  its deadline.
+VMs are allocated in blocks of one billing cycle (one hour) regardless of the size of the task.
+When computing the cost of scheduling a task on a given VM, the algorithm considers idle slots in blocks that were allocated for previous tasks to be free, while slots in new blocks cost the full price of a billing cycle.
+For example, if a task has a runtime of 10 minutes, and the price of a block is \$1, then the algorithm will either schedule the task on an existing VM that has an idle slot larger than 10 minutes for a cost of \$0, or it will allocate a new block on an existing VM, or provision a new VM, for a cost of \$1.
+If the cost of slots on two different VMs is equal, then the slot with the earliest start time is chosen.
+To prevent too many VMs from being provisioned, the algorithm always prefers to extend the runtime of existing VMs before allocating new VMs.
+The result of this is that the algorithm will only allocate a new block if there are no idle slots on existing blocks large enough or early enough to complete the task by its deadline, and it will only allocate a new VM if it cannot add a block to an existing VM to complete the task by its deadline.
 
 An example schedule generated by SPSS is shown in Figure.~\ref{fig:spss-example}.
-This example shows how SPSS tends to start many workflows in parallel, running
-each workflow over a longer period of time on only a few VMs to minimize cost. 
-In comparison, the dynamic algorithms tend to run one workflow at a time across 
-many VMs in parallel.
+This example shows how SPSS tends to start many workflows in parallel, running each workflow over a longer period of time on only a few VMs to minimize cost.
+In comparison, the dynamic algorithms tend to run one workflow at a time across many VMs in parallel.
 
@@
   \centering
-\includegraphics[width=1.0\columnwidth]{spss-gantt}
- \caption{Example of schedule generated using SPSS algorithm. Tasks are colored by workflow. }
+  \includegraphics[width=\textwidth]{figures/spss-gantt}
+  \caption[Example schedule generated by the SPSS algorithm]{Example schedule generated by the SPSS algorithm. Each row is a different VM. Tasks are represented as boxes and are colored by workflow.}
   \label{fig:spss-example}
@@
 
-\section{Performance Evaluation}
+\section{Evaluation}
 \label{sec:performance}
 
+This section describes a simulation study of the relative performance of three algorithms: DPDS, WA-DPDS, and SPSS.
+SPDS is not evaluated because, being a simplified version of DPDS and WA-DPDS, it will always produce results worse than the other algorithms.
+
 \subsection{Simulator}
 
-To evaluate and compare our algorithms, we developed a cloud workflow simulator based on
-CloudSim~\cite{Calheiros11}. Our simulation model consists of Cloud, VM and WorkflowEngine 
-entities. The Cloud entity is responsible for staring and terminating VM entities using 
-an API similar to Amazon EC2. VM entities simulate the execution of individual tasks,
-and the WorkflowEngine entity manages the scheduling of tasks and the provisioning of
-VMs. We assume that the VMs have a single core and execute tasks sequentially. Although 
-CloudSim provides a more advanced infrastructure model which includes time-sharing and 
-space-sharing policies, and power consumption in a datacenter, we do not use these 
-features, since we are interested mainly in the execution of tasks on VMs and high-level 
-workflow scheduling and provisioning. The simulator reads workflows in a modified version of
-the DAX format used by the Pegasus Workflow Management System~\cite{Deelman2005}. The modified format was
-created for the synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi08}.
+To evaluate and compare the algorithms, we developed a cloud workflow simulator based on CloudSim~\cite{Calheiros2011}.
+Our simulation model consists of Cloud, VM and WorkflowEngine entities.
+The Cloud entity is responsible for staring and terminating VM entities using an API similar to that used by Amazon EC2.
+VM entities simulate the execution of individual tasks, including randomized variations in runtime.
+And the WorkflowEngine entity manages the scheduling of tasks and the provisioning of VMs according to the rules established by the different algorithms.
+We assume that the VMs have a single core and execute tasks sequentially.
+Although CloudSim provides a more advanced infrastructure model which includes time-sharing and space-sharing policies, we do not use these features since we are interested mainly in the execution of tasks on VMs and high-level workflow scheduling and provisioning.
+The simulator reads workflow description files in a modified version of the DAX format used by the Pegasus Workflow Management System~\cite{Deelman2005}.
+The modified file format was created for the synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.
+
+\subsection{Workflow Ensembles}
+\label{sec:ensembles}
+
+In order to evaluate the algorithms on a standard set of workflows, we created randomized ensembles using workflows available from the workflow generator gallery published online by Bharathi, et al~\cite{WorkflowGenerator}.
+The gallery contains synthetic workflows modeled using structures and parameters that were taken from real applications.
+Ensembles were created using synthetic workflows from five real applications: SIPHT, a bioinformatics application, LIGO, a physics application, Epigenomics, another bioinformatics application, Montage, an astronomy application, and CyberShake, an earthquake science application.
+For each of these applications, workflows with 50, 100, 200, 300, 400, 500, 600, 700, 800, 900 and 1000 tasks were generated.
+For each workflow size, 20 different workflow instances were generated using parameters and task runtime distributions from traces of real workflows.
+The total collection of synthetic workflows contains 5 applications, 11 different workflow sizes, and 20 workflow instances, for a total of 1100 synthetic workflows.
+
+Using this collection of workflows, we constructed five different ensemble types:
+
+\begin{itemize}
+  \item constant
+  \item uniform, sorted
+  \item uniform, unsorted
+  \item Pareto, sorted
+  \item Pareto, unsorted
+\end{itemize}
+
+In the unsorted ensembles, workflows of different sizes are mixed together, and the priorities are assigned randomly so that there is no relationship between priority and workflow size.
+For many applications, however, large workflows are more important to users than small workflows because the represent more significant computations.
+To model this, the sorted ensembles are sorted by size, so that the largest workflows are given the highest priority.
+
+Constant ensembles contain workflows that all have the same number of tasks.
+For each constant ensemble, the number of tasks is chosen randomly from the set of possible workflow sizes.
+Once the size is determined, then N workflows of that size are chosen randomly for the ensemble from the set of synthetic workflows for the given application.
+
+Uniform ensembles contain workflows with sizes that are uniformly distributed among the set of possible sizes.
+Each workflow in a uniform ensemble is selected by first randomly choosing the size the workflow according to a uniform distribution, and then randomly choosing a workflow of that size from the set of synthetic workflows for the given application.
+
+Pareto ensembles contain a small number of larger workflows and a large number of smaller workflows.
+The sizes of the workflows in a Pareto ensemble are chosen according to a Pareto distribution.
+The distribution was modified so that the number of large workflows (of size $\geq$ 900) is increased by a small amount to produce a ``heavy-tail''.
+This causes Pareto ensembles to have a slightly larger number of large workflows, which reflects behavior commonly observed in many computational workloads.
+An example of the distribution of workflow sizes that occurs in a Pareto ensemble is shown in Figure~\ref{fig:ensemble-pareto}.
+
+\begin{figure}[ht] 
+    \centering
+    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
+    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of workflow sizes in Pareto ensembles. Workflow size is measured in number of tasks.}
+    \label{fig:ensemble-pareto}
+\end{figure}
+
+The number of workflows in an ensemble depends on the particular application, but we assume that ensemble sizes are on the order of between 10 and 100 workflows.
+This range is motivated by several reasons: First, such sizes are typical of the real applications we have examined.
+For example, the number of geographical sites of interest to the users of the CyberShake application in the past has been on the order of 100.
+Second, smaller ensembles consisting of just a few workflows can be aggregated into a single workflow, so there is no need to treat them as an ensemble.
+Similarly, when the number of workflows grows, and each workflow has a large number of tasks, either the deadline and budget constraints are low enough to prevent many of the workflows from running, or the problem of efficiently allocating them to the resources becomes similar to a bag-of-tasks problem, which is easier to solve efficiently.
 
@@
 
-\begin{figure}[tb] 
-\centering
-\includegraphics[width=0.6\columnwidth]{ensemble-pareto}
-\caption{Pareto-like distribution of workflow sizes in ensemble. Size is
-measured in number of tasks.}
-\label{fig:ensemble-distribution}
-\end{figure}
+\TODO{Edit this to remove duplication with section \ref{sec:ensembles}}
 
@@
 ensembles constructed from the workflows available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
-in the workflow generator gallery~\cite{Bharathi08}. The gallery contains synthetic
+in the workflow generator gallery~\cite{Bharathi2008}. The gallery contains synthetic
 workflows modeled on real applications. We used workflows from five applications 
@@
 workflows and a large number of smaller workflows, as seen in 
-Figure~\ref{fig:ensemble-distribution}. The number of large workflows (of size $\geq$ 900) 
+Figure~\ref{fig:ensemble-pareto}. The number of large workflows (of size $\geq$ 900) 
 is slightly increased to represent the ``heavy-tail'', i.e. that there is a 
@@
 
+\subsection{Performance Metric}
+\label{sec:perf_metric}
+
+In order to assess the relative performance of our algorithms it is necessary to define a metric that can be used to score the performance of the different algorithms on a given problem (ensemble, budget, and deadline).
+The simplest approach is to simply count the number of workflows in the ensemble that each algorithm is able to complete within the budget before the deadline.
+The problem with this approach is that it does not account for the priority-based utility function specified by the user.
+Using the counting approach, a less efficient algorithm may be able to complete a large number of low-priority workflows by executing the smallest workflows first.
+
+In order to account for the user's priority, we used an exponential scoring approach where the score for an ensemble $e$ is:
+%
+$$
+Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
+$$
+%
+where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that the highest-priority workflow has $Priority(w)=0$, the next highest workflow has $Priority(w)=1$, and so on.
+This exponential scoring function gives the highest priority workflow a score that is higher than all the lower-priority workflows combined:
+%
+$$
+2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i}
+$$
+%
+This scoring is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.
+
 \subsection{Metrics}
+
+\TODO{Merge this with Section \ref{sec:perf_metric}}
 
@@
 limited and does not allow completing all the workflows from the ensemble.
+
+\subsection{Experimental Parameters}
+
+For each application, we selected ranges of constraints (deadline and budget) to cover a broad parameter space: from tight constraints, where only a small number of workflows can be completed, to more liberal constraints where all, or almost all, of the workflows can be completed.
+These parameters vary among the applications and ensemble types and sizes.
+The budget constraints for each ensemble are calculated by identifying the smallest budget required to execute one of the workflows in the ensemble (MinBudget), and the smallest budget required to execute all workflows in the ensemble (MaxBudget):
+%
+$$
+MinBudget = \min_{w\ \in\ e}{Cost(w)}
+$$
+$$
+MaxBudget = \sum_{w\ \in\ e}{Cost(w)}
+$$
+%
+This range---$[MinBudget, MaxBudget]$---is then divided into equal intervals to determine the budgets to use in each experiment.
+Similarly, the deadline constraints for each ensemble are calculated by identifying the smallest amount of time required to execute a single workflow in the ensemble (MinDeadline), which is the length of the critical path for the workflow with the shortest critical path, and by identifying the smallest amount of time required to execute all workflows (MaxDeadline), which is the sum of the critical paths of all the workflows:
+%
+$$
+MinDeadline = \min_{w\ \in\ e}{CriticalPath(w)}
+$$
+$$
+MaxDeadline = \sum_{w\ \in\ e}{CriticalPath(w)}
+$$
+%
+This range---$[MinDeadline, MaxDeadline]$---is then divided into equal intervals to determine the deadlines to use in each experiment.
+By computing the budget and deadline constraits in this way we ensure that the experiments for each ensemble cover the most interesting area of the parameter space for the ensemble.
+
+In all the experiments we assumed that the VMs have a price of \$1 per VM-hour.
+This price was chosen to simplify interpretation of results and should not affect the relative performance of the different algorithms.
+In this study we do not take into account the heterogeneity of the infrastructure since we assume that it is always possible to select a VM type that has the best price to performance ratio for a given application.
+
+All the experiments were run with maximum autoscaling factor ($v_{max}$) set to 1.0 for DPDS and WA-DPDS.
+After experimenting with DPDS and WA-DPDS we found that, due to the high parallelism of workflows used, the resource utilization remains high enough without adjusting the autoscaling rate.
+Based on experiments with the target applications, we set the SPSS $\alpha$ parameter for deadline distribution to be 0.7, which allocates slightly more time to levels with many tasks.
 
@@
 
+\subsection{Relative Performance of Algorithms}
+
+The goal of the first experiment is to characterize the relative performance of the DPDS, WA-DPDS, and SPSS algorithms.
+This was done by simulating the algorithms on many different ensembles and comparing the scores computed using the technique described in Section~\ref{sec:perf_metric}.
+The algorithm that achieves the largest number of high scores is judged to have the best performance.
+
+Figure~\ref{fig:distributions} shows the percentage of simulations for which each algorithm achieved the highest score for a given ensemble type.
+This experiment was conducted using all five applications, with all five types of ensembles.
+For each application, ensembles of 50 workflows were created using 10 different random seeds.
+Each ensemble was simulated with all three algorithms using 10 budgets and 10 deadlines.
+The total number of simulations to produce each bar in the figure is: $10 * 10 * 10 = 1000$.
+The best scores percentage is computed by counting the number of times that a given algorithm achieved the highest score and dividing by 1000.
+Note that it is possible for multiple algorithms to get the same high score (to tie), so the numbers do not necessarily add up to 100\%.
+The sum is much higher than 100\% in cases where the dynamic algorithms perform relatively well because DPDS and WA-DPDS, which are very similar algorithms, often get the same high score.
+
+\begin{figure*}[ht]
+    \centering
+    \includegraphics[width=0.75\textwidth]{distributions}
+    \caption[Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.]{Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.
+C = Constant ensembles, PS = Pareto Sorted ensembles, PU = Pareto Unsorted ensembles, US = Uniform Sorted ensembles, and UU = Uniform Unsorted ensembles.}
+    \label{fig:distributions}
+\end{figure*}
+
+There are several interesting things to notice about Figure~\ref{fig:distributions}.
+The first is that, in most cases, SPSS significantly outperforms the dynamic algorithms (DPDS and WA-DPDS).
+This is attributed to the fact that SPSS is able to make more intelligent scheduling and provisioning decisions because it has the opportunity to compare different decisions and choose the one that results in the best outcome.
+In comparison, the dynamic algorithms are online algorithms and are not able to project into the future to weigh the outcomes of their choices.
+
+The second thing to notice about the figure is that, for constant ensembles, the dynamic algorithms perform significantly better relative to SPSS than for other ensemble types.
+This is a result of the fact that, since all of the workflows are of approximately the same size and shape, the choice of which workflow to execute next is arbitrary.
+
+Another thing to notice about these results is that the workflow-aware algorithms (WA-DPDS and SPSS) both perform better in most cases than the simple online algorithm that uses resource utilization alone to make provisioning decisions.
+This suggests that there is a significant value in having information about the structure and estimated runtime of a workflow when making scheduling and provisioning decisions.
+
+Finally, it is interesting to see that, for Montage and CyberShake, the relative performance of SPSS compared to the dynamic algorithms is significantly less than it is for other applications.
+We attribute this to the structure of Montage and CyberShake.
+The workflows for both applications are very wide relative to their height, and both have very short-running tasks.
+As a result of these two characteristics, the Montage and CyberShake applications have very short critical paths and look more like bag-of-tasks applications, which are easier to execute than more structured applications.
+DPDS and WA-DPDS are able to pack more of the tasks into the available budget and deadline because there are a) more choices for where to place the tasks, and b) the different choices have a smaller impact on the algorithms' ability to execute the workflow within the constraints.
+In addition, the short critical paths put SPSS at a disadvantage.
+Because of the way SPSS assigns deadlines to individual tasks, it is prevented from starting workflows late, which prevents it from packing tasks into the idle VM slots at the end of the schedule.
+
+\subsection{Task Granularity}
+
+In the last experiment we noted that, for Montage and CyberShake, the short runtimes of their tasks made it possible for the dynamic algorithms to perform better relative to SPSS when compared with other applications.
+In order to test this theory we adjusted the granularity of the tasks in several Montage and CyberShake ensembles to see how this would affect the relative performance of the algorithms.
+The granularity adjustment was achieved by multiplying the runtime of each task by a fixed scaling factor.
+
+Figure~\ref{fig:stretching} shows the relative performance of the algorithms as the scaling factor is increased from 1 to 16.
+Each data point represents the result of 500 simulations (5 random seeds for 10 budgets and 10 deadlines).
+The best scores percent was calculated the same way as it was for Figure~\ref{fig:distributions}.
+
+The figure clearly shows that, as the scaling factor increases, the relative performance of SPSS increases as well.
+In all cases, the performance of SPSS is superior to the dynamic algorithms for scaling factors greater than about 8.
+This suggests that, in general, for fine-grained workflows, the dynamic algorithms will produce better scores, and for coarse-grained workflows SPSS will produce better scores.
+
+\begin{figure*}[p]
+    \centering
+    \subfloat[CyberShake]{
+        \includegraphics[width=0.75\textwidth]{stretching_cybershake}
+    }\\
+    \subfloat[Montage]{
+        \includegraphics[width=0.75\textwidth]{stretching_montage}
+    }
+    \caption[Percentage of high scores achieved by each algorithm for Montage and Cybershake ensembles when task runtime is stretched]{Percentage of high scores achieved by each algorithm for Montage and Cybershake ensembles when task runtime is stretched.
+A stretching factor of $x$ means that the runtime of tasks in each workflow was multiplied by $x$.}
+    \label{fig:stretching}
+\end{figure*}
+
+\subsection{Inaccurate Task Runtime Estimates}
+\label{sec:variances}
+
+Both of the workflow-aware algorithms rely on estimates of task runtimes to make better scheduling and provisioning decisions.
+However, these estimates can be inaccurate.
+It is typically not possible to predict exactly how long a task will take to execute on a given machine at a given time.
+The actual runtime depends on many factors that are beyond the control of the application, such as network or disk contention, background processing, etc.
+
+Given that task runtime estimates are bound to be inaccurate, the question is: How does the error in the task runtime estimate impact the performance of our scheduling and provisioning algorithms?
+To test this we created an experiment to measure how introducing uniform errors in the runtime estimates affects the ability of the algorithms to achieve the budget and deadline constraints.
+This experiment used a single application, Montage, and a single ensemble type, uniform unsorted.
+This application and ensemble type was chosen arbitrarily for simplicity.
+It is expected that the application and ensemble type have little impact on how well the algorithms are able to stay within constraints.
+
+The estimation errors are generated by randomly sampling a uniform distribution between $-p\%$ and $+p\%$ of the runtime of a task.
+So, for example, if the value of $p$ is $10\%$, then the actual runtime of a task ($r_a$) is set in the simulation to be: $r_a = r_e * (1.0 + q)$, where $r_e$ is the estimated runtime of the task and $q$ is a uniform random number beween $-0.1$ and $+0.1$.
+Since the sampling is done uniformly, we expect to get just as many overestimates as underestimates in any given simulation.
+
+Note that the algorithms have \textbf{not} been changed to account for inaccurate runtime estimates.
+It is likely that better performance could be achieved if the algorithms were given some hint as to the accuracy of the task runtimes, but investigating that optimization is left for future work.
+The goal here is to determine how well the current algorithms are able to stay within the constraints given inaccurate task runtime estimates.
+
+Figure~\ref{fig:variances} shows the results of simulations for errors ranging from $0\%$ to $50\%$.
+These simulations were done for ensembles of 50 workflows, using 10 random seeds, 10 budgets, and 10 deadlines for each error value and algorithm (3000 simulations per error value).
+Each plot in the figure shows box plots of the ratio of actual cost to budget, or actual makespan to deadline.
+The actual cost and makespan are the real cost and makespan of the workflow calculated after the simulation completed.
+The ratio of actual values to constraints indicates whether the actual value exeeded the constraint.
+Values equal to 1 mean that the actual value was equal to the constraint, values less than 1 indicate that the actual value was within the constraint, and values greater than 1 mean that the actual value exceeded the constraint.
+
+Figure~\ref{fig:variances}.a shows the ratio of actual cost to budget.
+These results show two important characteristics of algorithm behavior.
+First, both dynamic algorithms very rarely exceeded the budget, even with very large errors.
+This is indicative of the fact that the dynamic algorithms are easily able to adapt to changes at runtime to ensure good results without exceeding the constraints, even when the quality of information available to them is low.
+Second, unlike the dynamic algorithms, the static algorithm frequently exceeded the budget constraint when the quality of runtime estimates was low.
+For small errors the budget was exceeded infrequently by small amounts, and for large errors (more than 10\%) the budget exceeded frequently by large amounts.
+This is a result of the fact that the static algorithm makes all of its decisions before execution and is not able to adapt to changing circumstances.
+The SPSS plan describes what tasks to execute on which VMs, but not when to execute them.
+At runtime, the worklow engine is forced to spend extra money to extend the runtime of some VMs so that all of the tasks assigned to that VM can be completed.
+At the same time, other VMs can be terminated early because the tasks they were assigned finished earlier than expected.
+However, because of dependencies in the workflow, the latter case is less likely to happen, which causes gaps in the schedule.
+So the net result is that the overall cost is increased.
+
+Figure~\ref{fig:variances}.b shows the ratio of actual makespan to budget.
+The interesting thing to notice about this figure is that, for all three algorithms, the deadline constraint is rarely exceeded, even in cases with very poor quality estimates.
+For the dynamic algorithms this is a result of the fact that they can adapt at runtime to changing circumstances to avoid exceeding the deadline.
+For the static algorithm this is a result of the way that SPSS schedules workflows and not of a particularly clever optimization.
+The SPSS algorithm tends to schedule workflows early, using up the budget long before the deadline is reached.
+This is a consequence of the deadline distribution function in SPSS, which prevents workflows from starting late.
+This is shown by the gantt chart in Figure~\ref{fig:spss-example}, which illustrates how SPSS tends to pile up workflows at the beginning of the timeline.
+In comparison, the dynamic algorithms tend to spread out workflow starts over the entire duration as shown in Figure~\ref{fig:spds-example}.
+The consequence of this behavior is that SPSS plans tend to use up the budget but leave plenty of time before the deadline.
+So when the runtime of the plan is increased by introducing errors, the SPSS plan has some room to expand without exceeding the deadline.
+
+Figure~\ref{fig:variances_scores} shows relative performance as a percentage of the number of high scores achieved by each algorithm as the error increases.
+The figure shows that, for this experiment, the scores remain the same as the error increases.
+This is a result of the way the error is applied.
+Since the error is selected uniformly from $-p\%$ to $+p\%$ it is equally likely that the error will be positive as negative.
+In other words, it is equally likely that a task will take longer as it is that the task will be shorter.
+Because the application used in this experiment, Montage, has a low task runtime variance, applying the error leaves the total runtime of the tasks in the workflow unchanged---the total number of CPU hours in each workflow remains the same regardless of the error.
+The end result is that the dynamic algorithms are able to achieve the same score without exceeding the constraints by using a different schedule that finishes the same total amount of work using the same budget and deadline.
+If we had applied a skewed error, for example, an error that was more likely to be positive than negative, then the scores for the dynamic algorithms would decrease relative to the static algorithm as the error increased.
+
+The behavior of these algorithms indicate that, when the quality of runtime estimates is low, the dynamic algorithms do a much better job of adapting to the changing circumstances to meet their obligations.
+In comparison, the static algorithm performs better with good information, but fails to meet the constraints when the quality of information is poor.
+It is possible that the static algorithm could be modified to add more breathing room in the schedule to account for situations where the runtime estimates may be inaccurate.
+Such a modification would result in slightly worse scores when the estimates are good, but would improve scores when the estimates are bad.
+Alternatively, the SPSS algorithm could be modified to generate a plan that specifies when to provision and deprovision each VM.
+In that case the workflow engine could be prevented from exceeding the budget and deadline constraints, but would prevent some workflows from finishing, which may significantly decrease the score.
+These two modifications are subjects for future work.
+
+\begin{figure*}[p]
+    \subfloat[Ratio of Actual Cost to Budget]{
+        \includegraphics[width=\textwidth]{variances_budget}
+    }\\
+    \subfloat[Ratio of Actual Makespan to Deadline]{
+        \includegraphics[width=\textwidth]{variances_deadline}
+    }
+    \caption[Boxplots for budget and deadline ratios when runtime estimates are inaccurate]{Boxplots for budget and deadline ratios when runtime estimate error varies from $\pm$0\% to $\pm$50\% for all three algorithms.
+A value of 1 indicates that the actual cost is equal to the budget, or the actual makespan is equal to the deadline, respectively.
+Values greater than 1 indicate that the budget/deadline has been exceeded.}
+    \label{fig:variances}
+\end{figure*}
+
+\subsection{Provisioning Delays}
+\label{sec:delays}
+
+One important issue to consider when provisioning resources in the cloud is the amount of time between when a resource is requested, and when it becomes available to the application.
+Typically these provisioning delays are on the order of a few minutes, and are highly dependent upon the architecture of the cloud system and/or the size of the user's VM image~\cite{Nurmi2008b}.
+
+Recall that it is assumed that resources are billed from the minute that they are requested until they are terminated, and not from the moment that they become available.
+As a result, provisioning delays have an impact on both the cost (budget constraint) and makespan (deadline constraint) of the ensemble.
+
+Figure~\ref{fig:delays} shows the ratios of actual values to constraints when the provisioning delay is increased from 0 seconds up to 15 minutes.
+The simulations were done using uniform unsorted ensembles of 50 workflows from the Montage application.
+Each plot in the figure contains three boxes: one for each algorithm.
+Each box summarizes the results of 10 budgets, 10 deadlines, and 10 random seeds for a total of 1000 simulations per algorithm and provisioning delay (3000 simulations per plot).
+As in Figure~\ref{fig:variances}, the y-axis in each plot represents the ratio of the actual simulated value to the constraint value for the budget constraint (actual cost) or the deadline constraint (actual makespan).
+These ratios indicate how frequently and by how much the actual simulated values exceeded (or did not exceed) the constraints.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.5\textwidth]{variances_scores}
+  \caption{Percentage of high scores achieved by each algorithm on Montage with uniform unsorted ensembles when runtime estimate error varies from $\pm$0\% to $\pm$50\%.}
+  \label{fig:variances_scores}
+\end{figure}
+
+\begin{figure*}[p]
+    \subfloat[Ratio of Actual Cost to Budget]{
+        \includegraphics[width=\textwidth]{delays_budget}
+    }\\
+    \subfloat[Ratio of Actual Makespan to Deadline]{
+        \includegraphics[width=\textwidth]{delays_deadline}
+    }
+    \caption[Boxplots for budget and deadline ratios when provisioning delays occur]{Boxplots for budget and deadline ratios when provisioning delay varies from 0 seconds to 15 minutes for all three algorithms.
+A value of 1 indicates that the actual cost is equal to the budget, or the actual makespan is equal to the deadline, respectively.
+Values greater than 1 indicate that the budget/deadline has been exceeded.}
+    \label{fig:delays}
+\end{figure*}
+
+The effect of provisioning delays on workflow performance is similar to that of inaccurate runtime estimates: when the delays are small, all algorithms are able to produce results within the constraints, but when the delays are larger, the dynamic algorithms are able to adapt to avoid exceeding the constraints while the static algorithm is not.
+The explanation for this behavior is also similar.
+The difference is that, a very small error in runtime estimates has little impact on the budget, while a very small provisioning delay has a large impact on the budget.
+Figure~\ref{fig:delays}.a shows that even for an impossibly short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget.
+In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011b} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget.
+With the exception of a few outliers for 10 and 15 minute delays, the dynamic algorithms consistently met the budget and deadline constraints.
+
+Figure~\ref{fig:delays_scores} shows the relative performance as a percentage of the number of high scores achieved by each algorithm as the provisioning delay increases.
+This figure shows that, as the delay increases, the relative performance of the static algorithm increases as well.
+This behavior is a result of the fact that the static algorithm is, in essence, cheating by using more time and money than the dynamic algorithms.
+In the previous section we showed how the relative performance of the algorithms remains the same when a uniform error is applied to the runtime constraints.
+In that case, the dynamic algorithms adapted to the error by rearranging the schedule to accomplish the same amount of work given the same deadline and budget.
+In this case, the dynamic algorithms adapt by performing less work (executing fewer workflows) to remain within the constraints while accounting for the delays.
+In both cases, the static algorithm completed the same amount of work, but did so by exceeding the constraints.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.5\textwidth]{delays_scores}
+  \caption{Percentage of high scores achieved by each algorithm on Montage with uniform unsorted ensembles when provisioning delay varies from 0 seconds to 15 minutes.}
+  \label{fig:delays_scores}
+\end{figure}
+
+This behavior suggests that SPSS is far too sensitive to provisioning delays in its current form to be of practical use in real systems.
+It is likely, as was the case for inaccurate runtime estimates, that modifying the SPSS algorithm would improve its performance for provisioning delays.
+In fact, since provisioning operations are infrequent (because all the algorithms tend to provision resources for a long time), it is likely that the performance of SPSS could be improved significantly by simply adding an estimate of the provisioning delay to its scheduling function.
+Such an estimate may not have to be particularly accurate to get good results, and developing an estimate from historical data should be relatively simple.
+Testing this idea is left for future work.
+
+\subsection{SPSS Planning Time}
+
+Unlike the dynamic algorithms, which execute their scheduling and provisioning logic at runtime, the SPSS algorithm plans out its provisioning and scheduling decisions ahead of workflow execution.
+Because SPSS involves more complicated logic than the dynamic algorithms, it is important to understand what impact this has on execution time.
+
+Figure~\ref{fig:spss_planning_time} shows the SPSS planning time for three different ensemble sizes (50, 100, 200), and five different workflow sizes (50, 200, 400, 600, 800).
+The ensembles were generated using a constant distribution equal to the workflow size desired.
+Two different applications were used: SIPHT and CyberShake.
+Each box summarizes the results of 2 applications, 10 random seeds, 10 budgets and 10 deadlines, or 2000 simulations per box.
+For ensemble size 200, the results for workflow size 800 are omitted because the simulations took too long to get a representative sample.
+
+Figure~\ref{fig:spss_planning_time} shows that, for small ensembles and small workflows, the SPSS planning time is reasonable, taking on the order of tens of seconds to a few minutes.
+For larger ensembles of large workflows, however, the SPSS planning time can easily reach 10-15 minutes.
+Considering that the largest workflows used in this experiment are rather small (maximum of 800 tasks), and that real workflows are often much larger (workflows with tens of thousands of tasks are common, and workflows with hundreds of thousands or millions of tasks are rare but do exist), it is unlikely that SPSS will be practical for ensembles of very large workflows.
+
+\begin{figure*}[t]
+    \includegraphics[width=\textwidth]{spss_planning_time}
+    \caption{Planning time of SPSS algorithm for different ensemble and workflow sizes.}
+    \label{fig:spss_planning_time}
+\end{figure*}
+
+The algorithmic complexity of SPSS is $O(n^2)$, where $n$ is the number of tasks in the ensemble.
+This derives from the fact that, for each task, SPSS considers scheduling the task on the cheapest available slot, which involves scanning all of the available slots on all the VMs.
+Since the number of available slots, in the worst case, is proportional to the number of tasks scheduled (because scheduling a task splits an existing slot into at most two slots), the complexity of SPSS is $O(n^2)$.
+In comparison, the dynamic algorithms all have a more scalable complexity of $O(n)$.
+DPDS only examines the tasks in the workflow once when they are scheduled, and WA-DPDS examines them twice: once in the admissions algorithm, and once when they are scheduled.
+This makes the dynamic algorithms a better fit for larger workflows and ensembles even though they may not produce results that are as good as SPSS.
+
+It may be possible to optimize SPSS to reduce its runtime by, for example, clustering the workflow to increase task granularity, which would decrease the ratio of planning time to ensemble makespan.
+It may also be possible to reduce the complexity of SPSS by employing more sophisticated data structures to store the available slots.
+Investigating these topics is left for future work.
+
 \section{Conclusions and Future Work}
@@
 multiple VM types and cloud providers, including private and community clouds, 
-which will make the problem even more complex and challenging~\cite{Marshall2010,vockler11,Juve2010}.
+which will make the problem even more complex and challenging~\cite{Marshall2010, Vockler2011, Juve2010}.
 
@@
 associated trade-offs. For this purpose our future work includes also further
-extension of our experimental studies~\cite{Juve2009,vockler11} of workflows on
+extension of our experimental studies~\cite{Juve2009, Vockler2011} of workflows on
 available cloud infrastructures to develop more accurate performance and cost

### 2012-04-19 paper/paper.tex
Fix missing references

@@
 growth is super-linear, which results from the Pareto-like distribution of 
-workflow sizes in the ensemble (see Figure~\ref{fig:ensemble-distribution}). 
+workflow sizes in the ensemble (see Figure~\ref{fig:ensemble-pareto}). 
 For ensembles of constant size workflows (see Figure~\ref{fig:budgets}(b)) 
@@
 Figure~\ref{fig:delays}.a shows that even for an impossibly short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget.
-In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011b} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget.
+In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget.
 With the exception of a few outliers for 10 and 15 minute delays, the dynamic algorithms consistently met the budget and deadline constraints.

### 2012-04-20 paper/paper.tex
updated figures

@@
 
+\begin{figure}[htb]
+    \centering
+    \subfloat[size]{
+        \includegraphics[width=0.75\columnwidth]{size-selected-LIGO}
+        }\\
+    \subfloat[cost]{
+        \includegraphics[width=0.75\columnwidth]{cost-selectedLIGO}
+        }
+    \caption{Example performance of algorithms expressed in
+    sum of dag sizes and effective cost, data for LIGO application, Pareto
+    Sorted ensemble of 100 workflows.}
+    \label{fig:size-cost}
+\end{figure}
+
 \subsection{Relative Performance of Algorithms}
@@
     \centering
-    \includegraphics[width=0.75\textwidth]{distributions}
+    \includegraphics[width=0.75\textwidth]{run-finish-variations-test-0-output-distributions}
     \caption[Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.]{Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.
@@
 
-Figure~\ref{fig:variances_scores} shows relative performance as a percentage of the number of high scores achieved by each algorithm as the error increases.
+% Figure~\ref{fig:variances_scores} shows relative performance as a percentage of the number of high scores achieved by each algorithm as the error increases.
 The figure shows that, for this experiment, the scores remain the same as the error increases.
@@
 
-\begin{figure*}[p]
+\begin{figure*}[tb]
     \subfloat[Ratio of Actual Cost to Budget]{
-        \includegraphics[width=\textwidth]{variances_budget}
-    }\\
+        \includegraphics[width=0.45\textwidth]{run-finish-variations-test-0-output-cost-all}
+        }
     \subfloat[Ratio of Actual Makespan to Deadline]{
-        \includegraphics[width=\textwidth]{variances_deadline}
+        \includegraphics[width=0.45\textwidth]{run-finish-variations-test-0-output-makespan-all}
         }
@@
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.5\textwidth]{variances_scores}
-  \caption{Percentage of high scores achieved by each algorithm on Montage with uniform unsorted ensembles when runtime estimate error varies from $\pm$0\% to $\pm$50\%.}
-  \label{fig:variances_scores}
-\end{figure}
-
-\begin{figure*}[p]
+%\begin{figure}[ht]
+%  \centering
+%  \includegraphics[width=0.5\textwidth]{variances_scores}
+%  \caption{Percentage of high scores achieved by each algorithm on Montage with
+%  uniform unsorted ensembles when runtime estimate error varies from $\pm$0\%
+%  to $\pm$50\%.}
+%  \label{fig:variances_scores}
+%\end{figure}
+
+\begin{figure*}[tb]
     \subfloat[Ratio of Actual Cost to Budget]{
-        \includegraphics[width=\textwidth]{delays_budget}
-    }\\
+        \includegraphics[width=0.45\textwidth]{run-finish-delays-test-0-output-cost-montage-uniform_unsorted}
+        }
     \subfloat[Ratio of Actual Makespan to Deadline]{
-        \includegraphics[width=\textwidth]{delays_deadline}
+        \includegraphics[width=0.45\textwidth]{run-finish-delays-test-0-output-dagfinish-montage-uniform_unsorted}
         }
@@
 
-Figure~\ref{fig:delays_scores} shows the relative performance as a percentage of the number of high scores achieved by each algorithm as the provisioning delay increases.
+% Figure~\ref{fig:delays_scores} shows the relative performance as a percentage of the number of high scores achieved by each algorithm as the provisioning delay increases.
 This figure shows that, as the delay increases, the relative performance of the static algorithm increases as well.
@@
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.5\textwidth]{delays_scores}
-  \caption{Percentage of high scores achieved by each algorithm on Montage with uniform unsorted ensembles when provisioning delay varies from 0 seconds to 15 minutes.}
-  \label{fig:delays_scores}
-\end{figure}
+% \begin{figure}[ht]
+%   \centering
+%   \includegraphics[width=0.5\textwidth]{delays_scores}
+%   \caption{Percentage of high scores achieved by each algorithm on Montage with uniform unsorted ensembles when provisioning delay varies from 0 seconds to 15 minutes.}
+%   \label{fig:delays_scores}
+% \end{figure}
 

### 2012-04-23 paper/paper.tex
Remove paragraphs that were eliminated

@@
   \centering
-  \includegraphics[width=\textwidth]{figures/spss-gantt}
+  \includegraphics[width=0.5\textwidth]{figures/spss-gantt}
   \caption[Example schedule generated by the SPSS algorithm]{Example schedule generated by the SPSS algorithm. Each row is a different VM. Tasks are represented as boxes and are colored by workflow.}
@@
 
-\subsection{Scenarios}
-\label{sec:scenarios}
-
-\TODO{Edit this to remove duplication with section \ref{sec:ensembles}}
+\subsection{Performance Metrics}
+\label{sec:perf_metric}
+
+In order to assess the relative performance of our algorithms it is necessary to define a metric that can be used to score the performance of the different algorithms on a given problem (ensemble, budget, and deadline).
+The simplest approach is to simply count the number of workflows in the ensemble that each algorithm is able to complete within the budget before the deadline.
+The problem with this approach is that it does not account for the priority-based utility function specified by the user.
+Using the counting approach, a less efficient algorithm may be able to complete a large number of low-priority workflows by executing the smallest workflows first.
+
+In order to account for the user's priority, we used an exponential scoring approach where the score for an ensemble $e$ is:
+%
+$$
+Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
+$$
+%
+where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that the highest-priority workflow has $Priority(w)=0$, the next highest workflow has $Priority(w)=1$, and so on.
+This exponential scoring function gives the highest priority workflow a score that is higher than all the lower-priority workflows combined:
+%
+$$
+2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i}
+$$
+%
+This scoring is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.
+
+\subsection{Metrics}
+
+\TODO{Merge this with Section \ref{sec:perf_metric}}
+
+\begin{itemize}
+  \item $N_c$: number of workflows completed,
+  \item $s$: user preference score (utility function),
+  \item $T_C$: total amount of computing time completed,
+  \item $C_{eff}$: effective cost of computing per hour, which includes
+  overheads due to the cost of idle resources.
+\end{itemize}
+
+The total amount of computing time completed, $T_C$, is calculated as the sum 
+of the runtimes of all tasks in all completed workflows. Because we assumed 
+that the resource cost per hour is equal to \$1, the total computing time 
+can be easily compared to the allocated budget $b$. Their ratio: 
+$u = (T_C * \$1)/b$, is the resource utilization, and its inverse: 
+$C_{eff} = b/T_C$, is the effective resource cost in dollars per hour. To
+calculate $C_{eff}$ and $u$ we use the value of allocated budget $b$, not the
+actual total cost, since we are interested in the scenarios when the budget is
+limited and does not allow completing all the workflows from the ensemble.
+
+\subsection{Experimental Parameters}
+
+\TODO{Edit this first paragraph, which was moved from the old Scenarios section}
 
@@
 
-\begin{figure*}[t] 
-\centering
-\includegraphics[width=0.19\textwidth]{pareto-MONTAGE-n-1000-8-dagh1-20m0}
-\includegraphics[width=0.19\textwidth]{pareto-CYBERSHAKE-n-1000-8-dagh1-20m0}
-\includegraphics[width=0.19\textwidth]{pareto-LIGO-n-1000-8-dagh1-40m0}
-\includegraphics[width=0.19\textwidth]{pareto-GENOME-n-1000-8-dagh100-1500m0}
-\includegraphics[width=0.19\textwidth]{pareto-SIPHT-n-1000-8-dagh5-50m0}
-\caption{Number of workflows completed for DPDS, AW-DPDS and SPSS
-algorithms on ensembles of 100 Pareto-distributed workflows, {\em small budget}
-(top) and {\em large budget} (bottom). The ranges on $y$ axis are chosen to 
-emphasize the differences between the performance of the algorithms.}
-\label{fig:number-complete-pareto}
-\end{figure*}
-
-The number of workflows in an ensemble depends on the particular application, but
-we assume that ensemble sizes are on the order of between 10 and 100 workflows.
-This range is motivated by several reasons: First, such sizes are typical of the
-real applications we have examined. For example, the number of geographical 
-sites of interest to the users of the CyberShake application is on the order 
-of 100. Second, smaller ensembles consisting of just a few workflows can be 
-aggregated into a single workflow, so there is no need to treat them as an
-ensemble. Similarly, when the number of workflows grows, and each 
-workflow has a large number of tasks, either the constraints are low enough to
-prevent many of the workflows from running, or the problem of efficiently 
-allocating them to the resources becomes similar to the bag-of-tasks problem, 
-which is easier to solve efficiently.
-
-In order to evaluate the algorithms on a standard set of workflows, we used
-ensembles constructed from the workflows available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
-in the workflow generator gallery~\cite{Bharathi2008}. The gallery contains synthetic
-workflows modeled on real applications. We used workflows from five applications 
-that were managed by the Pegasus Workflow Management System. For each
-application we constructed ensembles of 100 workflows by sampling collections of
-workflows of various sizes (number of tasks) based on a given distribution.
-Two distributions are considered:
-
-\begin{itemize}
-  \item constant distribution
-  \item Pareto-like distribution.
-\end{itemize}
-
-The constant distribution produces ensembles containing workflows with the same number of tasks per workflow, 
-while the Pareto-like distribution produces ensembles with a small number of larger 
-workflows and a large number of smaller workflows, as seen in 
-Figure~\ref{fig:ensemble-pareto}. The number of large workflows (of size $\geq$ 900) 
-is slightly increased to represent the ``heavy-tail'', i.e. that there is a 
-significant number of large workflows in the ensemble. We assume that large
-workflows are more important to users, so in the Pareto ensembles the workflows are 
-assigned priorities according to their size so that the largest workflows are given 
-the highest priority. In
-future work we plan to explore other ensemble types such as uniform distributions.
-
-\begin{figure*}[t]  
-\centering
-\includegraphics[width=0.19\textwidth]{pareto-size-MONTAGE-n-1000-8-dagh1-20m0}
-\includegraphics[width=0.19\textwidth]{pareto-size-CYBERSHAKE-n-1000-8-dagh1-20m0}
-\includegraphics[width=0.19\textwidth]{pareto-size-LIGO-n-1000-8-dagh1-40m0}
-\includegraphics[width=0.19\textwidth]{pareto-size-GENOME-n-1000-8-dagh100-1500m0}
-\includegraphics[width=0.19\textwidth]{pareto-size-SIPHT-n-1000-8-dagh5-50m0}
-\caption{ Total work completed expressed as the sum of runtimes (in hours) of
-tasks of completed workflows. Data computed for DPDS, AW-DPDS and SPSS
-algorithms on ensembles of 100 Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
-emphasize the differences between the performance of the algorithms.}
-\label{fig:total-time}
-\end{figure*}
-
-In the experiments we assumed that all the VMs are homogeneous, have a
-processing speed of 1000 MIPS (million instructions per second), and a price of
-\$1 per VM-hour. We also assume that the sizes of tasks in the workflow gallery
-are given in seconds measured on a 1000 MIPS machine. These values were chosen
-to approximate realistic numbers, and also to simplify interpretation of
-results. Moreover, the absolute values of these parameters do not affect the
-relative comparison between the algorithms. In this study we do not take into
-account the heterogeneity of the infrastructure since we assume that it is
-always possible to select a VM type that has the best price to performance ratio
-for a given application.
-
-For each application, we selected ranges of constraints (deadline and budget) so
-that it is possible to observe the possibly broad result space: from tight 
-constraints, where only a small number of workflows can be completed, to 
-more liberal constraints where all, or almost all, of the workflows can be 
-completed. These parameters vary among the applications. Montage and CyberShake 
-ensembles required budgets in the \$20 -- \$150 range and deadlines from 1 to
-20 hours to be completed. The LIGO and SIPHT workflows have more long-running
-tasks, so their ensembles required budgets up to \$1,500 and deadlines up to 
-40 and 60 hours. The Epigenomics workflow is even more costly, requiring 
-approximately \$12,000 and 1,500 hours to complete a full ensemble. These
-differences in total runtime and cost come from the fact that we used the
-same workflow size distribution for all five applications, but
-the sizes of tasks vary between applications.
-
-We show the results for three algorithms: DPDS, WA-DPDS and SPSS. These 
-experiments were run with maximum autoscaling factor set to 1.0 for
-DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that, 
-due to the high parallelism of workflows, the resource utilization remains 
-high most of the time, which leads to the rapid provisioning of new resources 
-at the beginning of ensemble execution. This, in turn, results in less 
-efficient total utilization due to high levels of parallelism. Setting the 
-scaling limit to 1.0, ensures that the shape of the resource provisioning 
-plan  is close to a rectangle (see Figure~\ref{fig:spds-example}) and the VM 
-termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
-of VMs before shutdown. Based on experiments with different applications, 
-we set the $\alpha$ parameter for deadline distribution in SPSS to be 0.7.
-
-\subsection{Performance Metric}
-\label{sec:perf_metric}
-
-In order to assess the relative performance of our algorithms it is necessary to define a metric that can be used to score the performance of the different algorithms on a given problem (ensemble, budget, and deadline).
-The simplest approach is to simply count the number of workflows in the ensemble that each algorithm is able to complete within the budget before the deadline.
-The problem with this approach is that it does not account for the priority-based utility function specified by the user.
-Using the counting approach, a less efficient algorithm may be able to complete a large number of low-priority workflows by executing the smallest workflows first.
-
-In order to account for the user's priority, we used an exponential scoring approach where the score for an ensemble $e$ is:
-%
-$$
-Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
-$$
-%
-where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that the highest-priority workflow has $Priority(w)=0$, the next highest workflow has $Priority(w)=1$, and so on.
-This exponential scoring function gives the highest priority workflow a score that is higher than all the lower-priority workflows combined:
-%
-$$
-2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i}
-$$
-%
-This scoring is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.
-
-\subsection{Metrics}
-
-\TODO{Merge this with Section \ref{sec:perf_metric}}
-
-There are many different ways to evaluate and compare the performance of 
-the algorithms. The set of ensemble performance metrics we have used to
-compare our algorithms includes:
-
-\begin{itemize}
-  \item $N_c$: number of workflows completed,
-  \item $s$: user preference score (utility function),
-  \item $T_C$: total amount of computing time completed,
-  \item $C_{eff}$: effective cost of computing per hour, which includes
-  overheads due to the cost of idle resources.
-\end{itemize}
-
-The number of workflows completed is a simple metric that can be used for
-comparison. However, because our ensembles contain workflows with different 
-priorities and sizes, a less efficient algorithm may schedule smaller
-workflows first, completing a large number of low-priority workflows. It may 
-also happen that a conservative admission algorithm may overestimate the cost 
-of a large workflow, thus rejecting high-priority workflows which actually 
-could fit into the schedule. Therefore, we defined the user preference score as:
-
-\begin{equation}
-\label{eq:score}
-s = \sum_{workflows\ completed}{2^{-p}}
-\end{equation}
-
-where $p$ is the priority of the workflow such that the highest-priority workflow 
-has $p=0$, the next highest workflow has $p=1$, and so on. This exponential scoring 
-function gives the highest priority workflow a score that is higher than all the 
-lower-priority workflows combined:
-
-\begin{equation}
-\label{eq:score-property}
-2^{-p} > \sum_{i=p+1,\ldots}2^{-i}
-\end{equation}
-
-This scoring is consistent with our definition of the problem, which is to 
-complete as many workflows as possible, according to their priorities, given 
-a budget and deadline.
-
-The total amount of computing time completed, $T_C$, is calculated as the sum 
-of the runtimes of all tasks in all completed workflows. Because we assumed 
-that the resource cost per hour is equal to \$1, the total computing time 
-can be easily compared to the allocated budget $b$. Their ratio: 
-$u = (T_C * \$1)/b$, is the resource utilization, and its inverse: 
-$C_{eff} = b/T_C$, is the effective resource cost in dollars per hour. To
-calculate $C_{eff}$ and $u$ we use the value of allocated budget $b$, not the
-actual total cost, since we are interested in the scenarios when the budget is
-limited and does not allow completing all the workflows from the ensemble.
-
-\subsection{Experimental Parameters}
-
 For each application, we selected ranges of constraints (deadline and budget) to cover a broad parameter space: from tight constraints, where only a small number of workflows can be completed, to more liberal constraints where all, or almost all, of the workflows can be completed.
@@
 
-We ran simulations for constraint values ranging between the estimated 
-limits indicated in Section~\ref{sec:scenarios}, resulting in approximately 
-50-150 (budget, deadline) data points for each application. The runs were
-performed for Pareto-distributed and constant workflow ensembles. From this result
-space we selected representative plots.
-
 \begin{figure*}[htb] 
+\TODO{Cut this figure down}\\
 \centering
@@
 
-Figure~\ref{fig:number-complete-pareto} shows the number of the workflows
-completed for two budget values for each application ensemble. The budget values
-selected correspond to two extreme cases: {\em small budget} (top) and {\em
-large budget} (bottom). The small budgets permit the algorithms to complete 
-only a small number of workflows from the ensemble, while the larger budgets 
-enable the algorithms to finish almost all the workflows. In three of
-five cases the SPSS algorithm performs better for tighter constraints. When the 
-budget is low, or the deadline is short, the static procedure is able to more 
-efficiently pack workflow tasks onto limited resources. This is the case for 
-the LIGO, Epigenomics and SIPHT ensembles for small budgets, and for large
-budgets with short deadlines. The exceptions are the Montage and CyberShake 
-ensembles,  which consist of many fine-grained tasks. In these applications, 
-as well for the other ensembles with large budgets and deadlines, the simple 
-dynamic scheduling techniques perform equally well or better. The reason for 
-this behavior is that the large number of tasks guarantees that resources 
-are almost never idle regardless of how the tasks are scheduled. We also 
-observe that WA-DPDS performs better than the workflow-unaware version, DPDS. 
-This means that for the workflows we tested the simple admission procedure based 
-on estimation of workflow cost does not degrade the solution, but rather in 
-many cases it enables larger workflows that would lead to budget overrun to
-be rejected. Thus it can save the space for smaller workflows to complete.
-
-%Figure~\ref{fig:score} shows the score $s$ as defined in Eq.~\ref{eq:score}
-%computed for two sample applications. Comparing
-%Figure~\ref{fig:number-complete-pareto} to Figure~\ref{fig:score} we can see that
-%the latter one gives more flattened shape: this means that e.g. the improvement
-%of the solution by one workflow may mean adding a one low-priority workflow.
-%Nevertheless we can see that the curves on both figures have a very similar
-%shape, which mens that our algorithms obey the priority rules and do not try to
-%admit lower-priority workflows at the cost of the more important ones.
-
-\begin{table}[tb]
-\centering
-
-(a) Number of workflows completed $N_c$:
-\medskip
-\begin{tabular}{r|cccccc}
- & M & C & L & E & S & All\tabularnewline
-\hline
-DPDS      &    55 &  47  & 23  & 14  &  0 & 139\tabularnewline 
-WA-DPDS   &    97 &  89  & 50  & 34  &  1 & 271\tabularnewline 
-SPSS      &    38 &  37  & 190 & 74  & 50 & 389\tabularnewline
-\end{tabular}
-\medskip
-
-(b) Total amount of computing time completed $T_C$:
-\medskip
-\begin{tabular}{r|cccccc}
- & M & C & L & E & S & All\tabularnewline
-\hline
-DPDS      &   55  & 47  &  26  & 14  &  0 & 142\tabularnewline
-WA-DPDS   &   96  & 98  &  59  & 26  & 10 & 289\tabularnewline
-SPSS      &   32  & 12  & 165  & 75  & 41 & 325\tabularnewline
-\end{tabular}
-\medskip
-
-(c) Exponential score $s$:
-\medskip
-\begin{tabular}{r|cccccc}
- & M & C & L & E & S & All\tabularnewline
-\hline
-DPDS      &   55  & 47  &  26  & 14  &  0 & 142\tabularnewline
-WA-DPDS   &   96  & 98  &  59  & 26  & 10 & 289\tabularnewline
-SPSS      &   32  & 12  & 165  & 75  & 41 & 325\tabularnewline
-\end{tabular}
-\medskip
-
-\caption{Summary table for Pareto-distributed workflows showing number of times
-each algorithm achieved the highest score over all budgets and deadlines. Data from a total of 525 runs on 
-Pareto-distributed ensembles for Montage (M), CyberShake (C), LIGO (L), 
-Epigenomics (E), SIPHT (S).
-\label{tab:num-dags-pareto}}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}[tb]
-\centering
-
-(a) Number of workflows completed $N_c$:
-\medskip
-\begin{tabular}{r|cccccc}
- & M & C & L & E & S & All\tabularnewline
-\hline
-DPDS      &   79  & 92  & 170 &  24 &  29 & 394\tabularnewline
-WA-DPDS   &   82  & 100 & 180 &  58 &  27 & 447\tabularnewline
-SPSS      &   72  & 77  & 195 &  75 &  50 & 469 \tabularnewline
-\end{tabular}
-\medskip
-
-(b) Total amount of computing time completed $T_C$:
-\medskip
-\begin{tabular}{r|cccccc}
- & M & C & L & E & S & All\tabularnewline
-\hline
-DPDS      &   79  & 92  & 170 &  24 &  29 & 394\tabularnewline
-WA-DPDS   &   82  & 100 & 180 &  46 &  27 & 435\tabularnewline
-SPSS      &   72  & 75  & 195 &  75 &  50 & 467 \tabularnewline
-\end{tabular}
-\medskip
-
-(c) Exponential score $s$:
-\medskip
-\begin{tabular}{r|cccccc}
- & M & C & L & E & S & All\tabularnewline
-\hline
-DPDS      &   79  & 92  & 170 &  24 &  29 & 394\tabularnewline
-WA-DPDS   &   82  & 100 & 180 &  46 &  27 & 435\tabularnewline
-SPSS      &   72  & 75  & 195 &  75 &  50 & 467 \tabularnewline
-\end{tabular}
-\medskip
-
-\caption{Summary table for equal size workflows showing number of times each
-algorithm achieved the highest score over all budgets and deadlines. Data from a total of 525 runs on 
-ensembles of equal size workflows for Montage (M), CyberShake (C), LIGO (L), 
-Epigenomics (E), SIPHT (S).
-\label{tab:num-dags-equal}}
-\end{table}
-
-Figure~\ref{fig:total-time} presents the same results using the total computing
-time metric, $T_C$. These plots confirm our previous observations about the
-algorithms' performance, but they also allow us observe a more general property:
-for a given budget, when the deadline is tightened the amount of work done
-steeply decreases. This is partially due to the fact that fewer workflows can be
-completed given a short deadline because their critical paths are simply too long
-and they are squeezed out of early start times by high priority workflows.
-In addition, assigning a shorter deadline means that more VMs need to be
-provisioned to complete the work. However, because of the dependencies in the 
-workflows, many of these VMs will either be idle, or will be backfilled with 
-tasks from low-priority workflows that never finish, resulting, effectively, 
-in reduced resource utilization and fewer completed workflows.
+\TODO{Rewrite this next paragraph}
 
@@
 
-% \begin{figure*}[htb] \centering
-% \includegraphics[width=0.19\textwidth]{score2-MONTAGE-n-1000-8-dagh1-20m0}
-% \includegraphics[width=0.19\textwidth]{score2-CYBERSHAKE-n-1000-8-dagh1-20m0}
-% \includegraphics[width=0.19\textwidth]{score2-LIGO-n-1000-8-dagh1-40m0}
-% \includegraphics[width=0.19\textwidth]{score2-GENOME-n-1000-8-dagh100-1500m0}
-% \includegraphics[width=0.19\textwidth]{score2-SIPHT-n-1000-8-dagh5-50m0}
-% \caption{Comparison of DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
-% Pareto-distributed workflows. Score is computed as $\sum(2^(-p))$ where
-% priorities of subsequent workflows are $p=0,1,2,\ldots$ (0 means highest
-% priority).} \label{fig:} \end{figure*}
-
-\begin{figure}[htb] \centering
-\includegraphics[width=0.7\columnwidth]{pareto-budget-SIPHT-n-1000-8-dagh5-50h4m0}\\
-(a) ensemble of Pareto-distributed workflows
-\includegraphics[width=0.7\columnwidth]{constant-budget-SIPHT-n-1000-0-dagh5-50h4m0}\\
-(b) ensemble of workflows of equal size \caption{Comparison of number of
-workflows completed for a given deadline, plotted versus different budget
-amounts.}
-\label{fig:budgets}
-\end{figure}
-
-In order to compare the efficiency of our algorithms, we counted the number of times
-each algorithm achieved the best result for all simulated applications, deadlines
-and budgets. Table~\ref{tab:num-dags-pareto} shows these summary values for 
-Pareto-distri\-buted workflows. Note that all algorithms that achieve the best 
-result are credited, so the sum for a given application can be greater than the
-total number of scenarios. These results depend on the metric: number of
-workflows completed does not count the priorities or sizes of workflows, which
-means that it happens more often that all the algorithms achieve the same best
-result. However, it does not mean that all algorithms complete the same
-workflows: that is why the values in (a), (b) and (c) are different. The 
-exponential score and total amount of computing time completed give the same
-rankings, since in our test ensembles larger workflows were given higher
-priorities. It can be observed based on the sampled problem space that the
-workflow-aware algorithms achieve the highest score in $\sim$2--3 times more 
-cases than the simple DPDS algorithm, which does not take into account workflow 
-structure or size. SPSS performs the best in $\sim$20\% more cases than WA-DPDS,
-including the Montage and CyberShake ensembles, for which SPSS does not 
-perform as well as WA-DPDS.
-
-Similar results were achieved for ensembles consisting of workflows of equal
-size set to 1000. A summary of these results is shown in
-Table~\ref{tab:num-dags-equal}. Again, we can see that the static SPSS algorithm performed better for coarse-grained
-workflows (LIGO, Epigenomics, SIPHT). The differences between algorithm scores
-are smaller since the workflows have equal size, therefore rejecting a particular
-workflow does not make it any easier for the algorithms to complete any other 
-workflow.
-
-Figure~\ref{fig:budgets} shows the same data for the SIPHT ensemble from another
-perspective. It helps answer the question: Given a deadline, how many workflows
-can be finished when we increase the budget? In Figure~\ref{fig:budgets}(a) the
-growth is super-linear, which results from the Pareto-like distribution of 
-workflow sizes in the ensemble (see Figure~\ref{fig:ensemble-pareto}). 
-For ensembles of constant size workflows (see Figure~\ref{fig:budgets}(b)) 
-the growth is linear and the costs are higher, since all the workflows in the
-test consisted of 1000 tasks.
-
 \begin{figure}[htb]
+    \TODO{Does this figure need removed?}
     \centering

### 2012-04-23 paper/paper.tex
Rewritten old discussion of results section

@@
 \subsection{Experimental Parameters}
-
+\label{sec:exp-parameters}
 \TODO{Edit this first paragraph, which was moved from the old Scenarios section}
@@
 
-\begin{figure*}[htb] 
-\TODO{Cut this figure down}\\
-\centering
-\includegraphics[width=0.19\textwidth]{pareto-cost-MONTAGE-n-1000-8-dagh1-20m0}
-\includegraphics[width=0.19\textwidth]{pareto-cost-CYBERSHAKE-n-1000-8-dagh1-20m0}
-\includegraphics[width=0.19\textwidth]{pareto-cost-LIGO-n-1000-8-dagh1-40m0}
-\includegraphics[width=0.19\textwidth]{pareto-cost-GENOME-n-1000-8-dagh100-1500m0}
-\includegraphics[width=0.19\textwidth]{pareto-cost-SIPHT-n-1000-8-dagh5-50m0}
-\caption{Effective computing cost in \$ per hour calculated by dividing the sum of
-task runtimes by the total budget. Higher costs result from lower resource utilization. Data based on DPDS, AW-DPDS and SPSS algorithms for ensembles of 100
-Pareto-distributed workflows. The ranges on $y$ axis are chosen to 
-emphasize the differences  between  the performance of the algorithms.}
-\label{fig:cost}
-\end{figure*}
-
-\TODO{Rewrite this next paragraph}
-
-These observations are further confirmed when examining the effective computing
-cost, as shown in Figure~\ref{fig:cost}. It can be seen that lower resource
-utilization and lower parallel efficiency for short deadlines lead to
-substantial increases in the effective computing cost. This increase can be 
-significant. For example, a SIPHT ensemble with short deadlines and a budget
-of \$1,000 has an effective computing cost $\sim$8--20\% higher than the 
-minimum effective cost of \$1/hour, depending on the algorithm used. This means 
-that out of the \$1,000 budget, \$80--\$200 is spent on idle resources.
-
-Another general conclusion can be drawn from Figure~\ref{fig:cost}: despite some
-random fluctuations resulting mostly from the finite result space, all the
-curves for all the applications and budgets have a similar shape. This shape
-represents a trade-off curve between two conflicting objectives: time and cost.
-We observe that the selection of deadline and cost can be formulated as a
-multi-criteria optimization problem, and that the curves presented in 
-Figure~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
-solutions to this problem. In our case the solutions were achieved by maximizing
-the amount of work done, which is the inverse of minimizing the effective 
-computing cost metric for a given deadline. Therefore, the obtained results 
-can be used to aid in resource allocation planning, when both cost and time 
-criteria are given not as constraints but as objectives.
+% New rewritten text for section ``discussion of Results'' starts here
 
 \begin{figure}[htb]
-    \TODO{Does this figure need removed?}
     \centering
@@
 \end{figure}
+
+To analyze the behaviour of algorithms we ran simulations for constraint values
+ranging between the estimated limits indicated in Section~\ref{sec:exp-parameters},
+resulting in $10 * 10 = 100$ (budget, deadline) data points for each of 5 applications, 3 algorithms and
+using 5 ensemble size distributions and 10 random seeds. From this result space
+of 75000 we selected representative plots.
+
+Fig.~\ref{fig:size-cost}a shows example of a performance curve using the total computing
+time metric, $T_C$ for LIGO application with pareto Sorted distribution of
+workflows. It can be seen that SPSS gives the best performance and WA-DPDS
+performs better than the workflow-unaware version, DPDS. 
+This means that for the workflows we tested the simple admission procedure based 
+on estimation of workflow cost does not degrade the solution, but rather in 
+many cases it enables larger workflows that would lead to budget overrun to
+be rejected. Thus it can save the space for smaller workflows to complete.
+
+This plot allows us observe a more general property: for a given budget,
+when the deadline is tightened the amount of work done steeply decreases. 
+This is partially due to the fact that fewer workflows can be
+completed given a short deadline because their critical paths are simply too long
+and they are squeezed out of early start times by high priority workflows.
+In addition, assigning a shorter deadline means that more VMs need to be
+provisioned to complete the work. However, because of the dependencies in the 
+workflows, many of these VMs will either be idle, or will be backfilled with 
+tasks from low-priority workflows that never finish, resulting, effectively, 
+in reduced resource utilization and fewer completed workflows.
+
+These observations are further confirmed when examining the effective computing
+cost, as shown in Figure~\ref{fig:size-cost}b. It can be seen that lower
+resource utilization and lower parallel efficiency for short deadlines lead to
+substantial increases in the effective computing cost. This increase can be 
+significant. For example, for short deadlines an effective computing cost
+is $\sim$10--40\% higher than the minimum effective cost of \$1/hour, depending
+on the algorithm used. This means that out of the \$1,000 budget, \$100--\$400
+is spent on idle resources.
+
+Another general conclusion can be drawn from Figure~\ref{fig:size-cost}b:
+all the curves for all the applications and budgets have a similar shape. This
+shape represents a trade-off curve between two conflicting objectives: time and cost.
+We observe that the selection of deadline and cost can be formulated as a
+multi-criteria optimization problem, and that the curves presented in 
+Figure~\ref{fig:size-cost}b are approximations of the Pareto front (or Pareto
+set) of solutions to this problem. In our case the solutions were achieved by maximizing
+the amount of work done, which is the inverse of minimizing the effective 
+computing cost metric for a given deadline. Therefore, the obtained results 
+can be used to aid in multi-criteria resource allocation planning, when both
+cost and time criteria are given not as constraints but as objectives.
 

### 2012-04-23 paper/paper.tex
better Fig 5
Merge SPDS with DPDS, change algorithms' font sizes

--- 2012-04-23 paper/paper.tex
@@
 
-This section describes several algorithms were developed to schedule and provision resources for ensembles of workflows on the cloud under budget and deadline constraints.
-
-\subsection{Static Provisioning Dynamic Scheduling (SPDS) }
-\label{sec:spds}
-
-The simplest strategy for executing ensembles of workflows in the cloud is to provision resources statically and schedule workflow tasks dynamically.
+This section describes three algorithms were developed to schedule and provision resources for ensembles of workflows on the cloud under budget and deadline constraints.
+
+\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
+\label{sec:dpds}
+
+DPDS is an online algorithm that provisions resources and schedules tasks at runtime. It consists of two main parts: a provisioning procedure, and a scheduling procedure.
+
+DPDS' provisioning procedure is based on resource utilization.
+DPDS starts with a fixed number of resources calculated based on the available time and budget, and adjusts the number of resources according to how well they are utilized by the application.
+
 Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a VM in dollars $p$, it is easy to calculate the number of VMs, $N_{VM}$, to provision so that the entire budget is consumed before the deadline:
@@
 %
-The SPDS algorithm statically provisions $N_{VM}$ VMs at the start of the ensemble execution and keeps them running until the deadline is reached, or the budget runs out.
-This provisioning plan has the advantage that it minimizes the number of provisioning and de-provisioning requests.
-
-Once the VMs are provisioned, the tasks are mapped onto idle VMs using the dynamic priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
-
-\begin{algorithm}[tb]
-\caption{Priority-based scheduling algorithm for SPDS}
-\label{alg:ds}
-\begin{algorithmic}[1]
-\Procedure{Schedule}{}
-    \State $P\gets$ empty priority queue
-	\State $IdleVMs\gets$ set of idle VMs
-	\For{root task $t$ in all workflows}
-    	\State \Call{Insert}{$t,P$} 
-    \EndFor
-    \While{deadline not reached}
-    	\While{$IdleVMs \neq \emptyset\ and\ P \neq \emptyset$}
-    		\State $v\gets$ \Call{SelectRandom}{$IdleVMs$}
-    		\State $t\gets$ \Call{Pop}{$P$}
-    		\State \Call{Submit}{$t,v$}
-    	\EndWhile
-    	\State Wait for task $t$ to finish on VM $v$
-    	\State Update $P$ with ready children of $t$
-		\State \Call{Insert}{$v,IdleVMs$}
-    \EndWhile
-\EndProcedure
-\end{algorithmic} 
-\end{algorithm}
-
-Initially, the ready tasks from all workflows in the ensemble are added to a priority queue based on the priority of the workflow to which they belong.
-If there are idle VMs available, and the priority queue is not empty, the next task from the priority queue is submitted to an arbitrarily chosen idle VM.
-The process is repeated until there are no idle VMs or the priority queue is empty.
-The scheduler then waits for a task to finish, adds its ready children to the priority queue, marks the VM as idle, and the entire process repeats until the deadline is reached.
-
-This algorithm guarantees that tasks from lower priority workflows are always deferred when higher-priority tasks are available, but lower-priority tasks can still occupy idle VMs when higher-priority tasks are not available.
-However, because there is no preemption, long-running low-priority tasks may delay the execution of higher-priority tasks.
-In addition, tasks from low priority workflows may be executed even though there is no chance that those workflows will be completed within the current budget and deadline.
-Figure~\ref{fig:spds-example} shows an example schedule generated using the SPDS algorithm.
-The figure illustrates how tasks from lower priority workflows backfill idle VMs when tasks from higher priority workflows are not available.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=0.5\textwidth]{figures/spds-gantt}
-  \caption[Example schedule generated by the SPDS algorithm.]{Example schedule generated by the SPDS algorithm. Each row is a different VM. Tasks are represented as boxes and are colored by workflow.}
-  \label{fig:spds-example}
-\end{figure}
-
-\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
-
-The static provisioning approach of SPDS may perform poorly when resource demand changes during ensemble execution.
-In those cases, SPDS may either over-provision, wasting the budget on idle VMs, or under-provision, using up the deadline while starving the application of usable VMs.
-One of the advantages of elasticity in cloud computing is that the number of resources can be changed to meet the demands of an application.
-In policy-based provisioning (or auto-scaling) systems, metrics such as resource utilization or task queue length are monitored to estimate application demand.
-This is the approach used by the Dynamic Provisioning Dynamic Scheduling (DPDS) algorithm to determine when to provision and de-provision VMs.
-The algorithm periodically computes resource utilization using the percentage of idle VMs over time, and adjusts the number of VMs if the utilization is above or below given thresholds.
+The DPDS algorithm provisions $N_{VM}$ VMs at the start of the ensemble execution.
+
+As the application runs, the DPDS algorithm periodically computes resource utilization using the percentage of idle VMs over time, and adjusts the number of VMs if the utilization is above or below given thresholds.
 Because it is assumed that VMs are billed by the hour, DPDS only considers VMs that are approaching their hourly billing cycle when deciding which VMs to terminate.
-The DPDS algorithm is shown in Algorithm~\ref{alg:prov}.
+This dynamic provisioning algorithm is shown in Algorithm~\ref{alg:prov}.
 
@@
 \label{alg:prov}
+{\small
 \begin{algorithmic}[1]
@@
 \end{algorithmic}
+}
 \end{algorithm}
@@
 
+In order to schedule individual workflow tasks onto available VMs, DPDS uses the dynamic, priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
+
+\begin{algorithm}[tb]
+\caption{Priority-based scheduling algorithm for DPDS}
+\label{alg:ds}
+{\small
+\begin{algorithmic}[1]
+\Procedure{Schedule}{}
+    \State $P\gets$ empty priority queue
+	\State $IdleVMs\gets$ set of idle VMs
+	\For{root task $t$ in all workflows}
+    	\State \Call{Insert}{$t,P$} 
+    \EndFor
+    \While{deadline not reached}
+    	\While{$IdleVMs \neq \emptyset\ and\ P \neq \emptyset$}
+    		\State $v\gets$ \Call{SelectRandom}{$IdleVMs$}
+    		\State $t\gets$ \Call{Pop}{$P$}
+    		\State \Call{Submit}{$t,v$}
+    	\EndWhile
+    	\State Wait for task $t$ to finish on VM $v$
+    	\State Update $P$ with ready children of $t$
+		\State \Call{Insert}{$v,IdleVMs$}
+    \EndWhile
+\EndProcedure
+\end{algorithmic}
+}
+\end{algorithm}
+
+Initially, the ready tasks from all workflows in the ensemble are added to a priority queue based on the priority of the workflow to which they belong.
+If there are idle VMs available, and the priority queue is not empty, the next task from the priority queue is submitted to an arbitrarily chosen idle VM.
+The process is repeated until there are no idle VMs or the priority queue is empty.
+The scheduler then waits for a task to finish, adds its ready children to the priority queue, marks the VM as idle, and the entire process repeats until the deadline is reached.
+
+The DPDS algorithm guarantees that tasks from lower priority workflows are always deferred when higher-priority tasks are available, but lower-priority tasks can still occupy idle VMs when higher-priority tasks are not available.
+However, because there is no preemption, long-running low-priority tasks may delay the execution of higher-priority tasks.
+In addition, tasks from low priority workflows may be executed even though there is no chance that those workflows will be completed within the current budget and deadline.
+Figure~\ref{fig:spds-example} shows an example schedule generated using the SPDS algorithm.
+The figure illustrates how tasks from lower priority workflows backfill idle VMs when tasks from higher priority workflows are not available.
+
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=0.5\textwidth]{figures/spds-gantt}
+  \caption[Example schedule generated by the DPDS algorithm.]{Example schedule generated by the DPDS algorithm. Each row is a different VM. Tasks are represented as boxes and are colored by workflow.}
+  \label{fig:spds-example}
+\end{figure}
+
 \subsection{Workflow-Aware DPDS (WA-DPDS)}
@@
 \label{alg:wa-dpds}
+{\small
 \begin{algorithmic}[1]
@@
 \label{alg:admit}
+{\small
 \begin{algorithmic}[1]
@@
 \label{alg:planworkflow}
+{\small
 \begin{algorithmic}[1]

### 2012-04-23 paper/paper.tex

--- 2012-04-23 paper/paper.tex
@@
     \centering
-    \includegraphics[width=0.75\textwidth]{run-finish-variations-test-0-output-distributions}
+    \includegraphics[width=1.06\textwidth]{run-finish-variations-test-0-output-distributions}
     \caption[Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.]{Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.

### 2012-04-24 paper/paper.tex
add failures

@@
 
+\subsection{Failures}
+\label{sec:failures}
+
+Running workflows consisting of large numbers of tasks on distributed systems
+almost always faces different types of failures. The goal of this experiment was
+to asses the behaviour of the algorithms in the presence of task execution
+failures by introducing a failure model into the simulator. The model is
+characterized by a failure rate $f$ that defines the probability that a task
+will fail. The failure time of the task is determined by randomly sampling a
+time between task start and finish time. If the task fails, it is reported to
+the workflow engine, which then resubmits the task for execution. Dynamic algorithms
+add the task again to the priority queue (queue $P$ in Algorithm~\ref{alg:ds})
+so that it can be resubmitted by the scheduler. Static algorithm resubmits the
+failed task on the same VM that was selected in the plan.
+
+\begin{figure*}[tb]
+    \subfloat[Ratio of Actual Cost to Budget]{
+        \includegraphics[width=0.45\textwidth]{run-finish-failures-test-0-output-cost-all}
+        }
+    \subfloat[Ratio of Actual Makespan to Deadline]{
+        \includegraphics[width=0.45\textwidth]{run-finish-failures-test-0-output-makespan-all}
+        }
+    \caption{Boxplots for budget and deadline ratios when failure rate varies
+    from 0 seconds to 50 \% for all three algorithms. A value of 1
+    indicates that the actual cost is equal to the budget, or the actual makespan is equal to the deadline, respectively. Values greater than 1 indicate that the budget/deadline has been exceeded.}
+    \label{fig:failures}
+\end{figure*}
+
+Fig.~\ref{fig:failures} shows the ratios of actual values to constraints when
+the failures are introduced. The high failure rates have a high impact on the
+performance of the static algorithm and can degrade the solution considerably.
+Comparing Fig.~\ref{fig:failures} to Figs.~\ref{fig:variances}
+and~\ref{fig:delays} one may conclude that failures are worse than
+provisioning delays and runtime estimate errors, since their impact is larger.
+However, small failure rates (below 10\%) do not cause constraint overruns
+larger than 10\%. We consider higher failure rates as rare events that suggest
+some significant system malfunction or improper resource selection
+process. It is thus more important that all the algorithms can handle task
+failures by resubmission (as is the case of Pegasus and Condor DAGMan).
+
 \subsection{SPSS Planning Time}

### 2012-04-24 paper/paper.tex
Merge metrics sections

@@
 
-In order to assess the relative performance of our algorithms it is necessary to define a metric that can be used to score the performance of the different algorithms on a given problem (ensemble, budget, and deadline).
-The simplest approach is to simply count the number of workflows in the ensemble that each algorithm is able to complete within the budget before the deadline.
-The problem with this approach is that it does not account for the priority-based utility function specified by the user.
-Using the counting approach, a less efficient algorithm may be able to complete a large number of low-priority workflows by executing the smallest workflows first.
-
-In order to account for the user's priority, we used an exponential scoring approach where the score for an ensemble $e$ is:
-%
-$$
-Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
+In order to assess the relative performance of our algorithms, we have defined the following metrics:
+
+\begin{itemize}
+  \item $N_c$: number of workflows completed,
+  \item $T_C$: total amount of computing time completed,
+  \item $C_{eff}$: effective cost of computing per hour, which includes overheads due to the cost of idle resources.
+  \item $s$: user preference score (utility function),
+\end{itemize}
+
+These metrics measure the relative performance of the different algorithms on a given problem (ensemble, budget, and deadline).
+
+The simplest approach is to simply count the number of workflows in the ensemble that each algorithm is able to complete within the budget, before the deadline, $N_c$.
+
+The total amount of computing time completed, $T_C$, is the sum of the runtimes of all tasks in all completed workflows.
+Because we assumed that the resource cost per hour is equal to \$1, the total computing time can be easily compared to the allocated budget $b$.
+Their ratio: $u = (T_C * \$1)/b$, is the resource utilization, and its inverse: $C_{eff} = b/T_C$, is the effective resource cost in dollars per hour.
+To calculate $C_{eff}$ and $u$ we use the value of allocated budget $b$, not the actual total cost, since we are interested in the scenarios when the budget is limited and does not allow completing all the workflows from the ensemble.
+
+The $N_c$, $T_C$ and $C_{eff}$ metrics do not account for the priority-based utility function specified by the user.
+Using $N_c$, for example, a less efficient algorithm may be able to complete a large number of low-priority workflows by executing the smallest workflows first.
+In order to account for the user's priority, we used an exponential scoring approach where the score $s$ for an ensemble $e$ is:
+%
+$$
+s = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
 $$
@@
 %
-This scoring is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.
-
-\subsection{Metrics}
-
-\TODO{Merge this with Section \ref{sec:perf_metric}}
-
-\begin{itemize}
-  \item $N_c$: number of workflows completed,
-  \item $s$: user preference score (utility function),
-  \item $T_C$: total amount of computing time completed,
-  \item $C_{eff}$: effective cost of computing per hour, which includes
-  overheads due to the cost of idle resources.
-\end{itemize}
-
-The total amount of computing time completed, $T_C$, is calculated as the sum 
-of the runtimes of all tasks in all completed workflows. Because we assumed 
-that the resource cost per hour is equal to \$1, the total computing time 
-can be easily compared to the allocated budget $b$. Their ratio: 
-$u = (T_C * \$1)/b$, is the resource utilization, and its inverse: 
-$C_{eff} = b/T_C$, is the effective resource cost in dollars per hour. To
-calculate $C_{eff}$ and $u$ we use the value of allocated budget $b$, not the
-actual total cost, since we are interested in the scenarios when the budget is
-limited and does not allow completing all the workflows from the ensemble.
+This metric is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.
 

### 2012-04-24 paper/paper.tex
Clean up/merge old scenarios text with experimental parameters section

@@
 
+\subsection{Experimental Parameters}
+\label{sec:exp-parameters}
+
+In order to observe the interesting characteristics of the proposed algorithms, we evaluated them using a broad range of constraint values.
+For each ensemble, we selected ranges for deadline and budget that cover a broad parameter space: from tight constraints, where only a small number of workflows can be completed, to more liberal constraints where all, or almost all, of the workflows can be completed.
+Because the interesting parameter space varies among the applications, ensemble types and sizes, we computed constraint ranges based on the characteristics of each ensemble.
+The budget constraints are calculated by identifying the smallest budget required to execute one of the workflows in the ensemble (MinBudget), and the smallest budget required to execute all workflows in the ensemble (MaxBudget):
+%
+$$
+MinBudget = \min_{w\ \in\ e}{Cost(w)}
+$$
+$$
+MaxBudget = \sum_{w\ \in\ e}{Cost(w)}
+$$
+%
+This range---$[MinBudget, MaxBudget]$---is then divided into equal intervals to determine the budgets to use in each experiment.
+Similarly, the deadline constraints for each ensemble are calculated by identifying the smallest amount of time required to execute a single workflow in the ensemble (MinDeadline), which is the length of the critical path for the workflow with the shortest critical path, and by identifying the smallest amount of time required to execute all workflows (MaxDeadline), which is the sum of the critical paths of all the workflows:
+%
+$$
+MinDeadline = \min_{w\ \in\ e}{CriticalPath(w)}
+$$
+$$
+MaxDeadline = \sum_{w\ \in\ e}{CriticalPath(w)}
+$$
+%
+This range---$[MinDeadline, MaxDeadline]$---is then divided into equal intervals to determine the deadlines to use in each experiment.
+By computing the budget and deadline constraits in this way we ensure that the experiments for each ensemble cover the most interesting area of the parameter space for the ensemble.
+
+In all the experiments we assumed that the VMs have a price of \$1 per VM-hour.
+This price was chosen to simplify interpretation of results and should not affect the relative performance of the different algorithms.
+In this study we do not take into account the heterogeneity of the infrastructure since we assume that it is always possible to select a VM type that has the best price to performance ratio for a given application.
+
+All the experiments were run with maximum autoscaling factor ($v_{max}$) set to 1.0 for DPDS and WA-DPDS.
+After experimenting with DPDS and WA-DPDS we found that, due to the high parallelism of workflows used, the resource utilization remains high enough without adjusting the autoscaling rate.
+Based on experiments with the target applications, we set the SPSS $\alpha$ parameter for deadline distribution to be 0.7, which allocates slightly more time to levels with many tasks.
+
 \subsection{Performance Metrics}
@@
 This metric is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.
-
-\subsection{Experimental Parameters}
-\label{sec:exp-parameters}
-\TODO{Edit this first paragraph, which was moved from the old Scenarios section}
-
-We have prepared test scenarios to evaluate the proposed algorithms under a
-range of parameters that enables us to observe the interesting characteristics
-of the different algorithms for smaller and larger constraints. For a given
-ensemble we can expect, in general, that the amount of work done (in terms of
-workflows completed) depends mainly on the budget constraint, which determines
-the number of machine-hours that can be provisioned. The deadline, although
-important in terms of scheduling, is assumed to be greater than the longest critical
-path of all workflows in the ensemble. We can then have a trivial but unrealistic solution
-where the deadline  is long enough to allow all the workflow tasks to be
-executed sequentially: this eliminates all the overheads of parallel execution
-and leads to the highest resource utilization. More interesting and more
-realistic, however, are the cases where the deadline is shorter than the serial
-schedule, since it forces the system to provision multiple VMs to execute
-workflow tasks in parallel.
-
-For each application, we selected ranges of constraints (deadline and budget) to cover a broad parameter space: from tight constraints, where only a small number of workflows can be completed, to more liberal constraints where all, or almost all, of the workflows can be completed.
-These parameters vary among the applications and ensemble types and sizes.
-The budget constraints for each ensemble are calculated by identifying the smallest budget required to execute one of the workflows in the ensemble (MinBudget), and the smallest budget required to execute all workflows in the ensemble (MaxBudget):
-%
-$$
-MinBudget = \min_{w\ \in\ e}{Cost(w)}
-$$
-$$
-MaxBudget = \sum_{w\ \in\ e}{Cost(w)}
-$$
-%
-This range---$[MinBudget, MaxBudget]$---is then divided into equal intervals to determine the budgets to use in each experiment.
-Similarly, the deadline constraints for each ensemble are calculated by identifying the smallest amount of time required to execute a single workflow in the ensemble (MinDeadline), which is the length of the critical path for the workflow with the shortest critical path, and by identifying the smallest amount of time required to execute all workflows (MaxDeadline), which is the sum of the critical paths of all the workflows:
-%
-$$
-MinDeadline = \min_{w\ \in\ e}{CriticalPath(w)}
-$$
-$$
-MaxDeadline = \sum_{w\ \in\ e}{CriticalPath(w)}
-$$
-%
-This range---$[MinDeadline, MaxDeadline]$---is then divided into equal intervals to determine the deadlines to use in each experiment.
-By computing the budget and deadline constraits in this way we ensure that the experiments for each ensemble cover the most interesting area of the parameter space for the ensemble.
-
-In all the experiments we assumed that the VMs have a price of \$1 per VM-hour.
-This price was chosen to simplify interpretation of results and should not affect the relative performance of the different algorithms.
-In this study we do not take into account the heterogeneity of the infrastructure since we assume that it is always possible to select a VM type that has the best price to performance ratio for a given application.
-
-All the experiments were run with maximum autoscaling factor ($v_{max}$) set to 1.0 for DPDS and WA-DPDS.
-After experimenting with DPDS and WA-DPDS we found that, due to the high parallelism of workflows used, the resource utilization remains high enough without adjusting the autoscaling rate.
-Based on experiments with the target applications, we set the SPSS $\alpha$ parameter for deadline distribution to be 0.7, which allocates slightly more time to levels with many tasks.
 

### 2012-04-24 paper/paper.tex
update section V.E

@@
 
-\subsection{Discussion of Results}
+\subsection{Example of Results}
 
@@
 
-\begin{figure}[htb]
+\begin{figure}[tb]
     \centering
     \subfloat[size]{
-        \includegraphics[width=0.75\columnwidth]{size-selected-LIGO}
+        \includegraphics[width=0.6\columnwidth]{size-selected-LIGO}
         }\\
     \subfloat[cost]{
-        \includegraphics[width=0.75\columnwidth]{cost-selectedLIGO}
+        \includegraphics[width=0.6\columnwidth]{cost-selectedLIGO}
         }
@@
     sum of dag sizes and effective cost, data for LIGO application, Pareto
-    Sorted ensemble of 100 workflows.}
+    Sorted ensemble of 100 workflows, budget = \$600}
     \label{fig:size-cost}
@@
 
-To analyze the behaviour of algorithms we ran simulations for constraint values
-ranging between the estimated limits indicated in Section~\ref{sec:exp-parameters},
-resulting in $10 * 10 = 100$ (budget, deadline) data points for each of 5 applications, 3 algorithms and
-using 5 ensemble size distributions and 10 random seeds. From this result space
-of 75000 we selected representative plots.
-
-Fig.~\ref{fig:size-cost}a shows example of a performance curve using the total computing
-time metric, $T_C$ for LIGO application with pareto Sorted distribution of
-workflows. It can be seen that SPSS gives the best performance and WA-DPDS
-performs better than the workflow-unaware version, DPDS.
-This means that for the workflows we tested the simple admission procedure based
-on estimation of workflow cost does not degrade the solution, but rather in
-many cases it enables larger workflows that would lead to budget overrun to
-be rejected. Thus it can save the space for smaller workflows to complete.
+%To analyze the behaviour of algorithms we ran simulations for constraint values
+%ranging between the estimated limits indicated in
+%Section~\ref{sec:exp-parameters}, resulting in $10 * 10 = 100$ (budget,
+%deadline) data points for each of 5 applications, 3 algorithms and using 5
+%ensemble size distributions and 10 random seeds. From this result space of
+%75000 we selected representative plots.
+
+To illustrate the behavior of the algorithms we show here examples of
+performance curves obtained for selected sample parameters.
+Fig.~\ref{fig:size-cost}a uses the total computing time metric, $T_C$ for LIGO
+application with pareto Sorted distribution of workflows. It can be seen that
+SPSS gives the best performance and WA-DPDS performs better than the
+workflow-unaware version, DPDS. This means that for the workflows we tested the
+simple admission procedure based on estimation of workflow cost does not degrade
+the solution, but rather in many cases it enables larger workflows that would
+lead to budget overrun to be rejected. Thus it can save the space for smaller
+workflows to complete.
 
@@
 is $\sim$10--40\% higher than the minimum effective cost of \$1/hour, depending
-on the algorithm used. This means that out of the \$1,000 budget, \$100--\$400
-is spent on idle resources.
+on the algorithm used. This means that out of the \$600 budget,
+$\sim$\$60--\$240 is spent on idle resources.
 

### 2012-04-24 paper/paper.tex
Update task granularity plot/text

--- 2012-04-24 paper/paper.tex
@@
 Figure~\ref{fig:stretching} shows the relative performance of the algorithms as the scaling factor is increased from 1 to 16.
-Each data point represents the result of 500 simulations (5 random seeds for 10 budgets and 10 deadlines).
+This figure only shows results for the CyberShake application with pareto unsorted ensembles, but similar results were found for other ensemble types for both Montage and CyberShake.
+Each data point in the figure represents the result of 500 simulations (5 random seeds for 10 budgets and 10 deadlines).
 The best scores percent was calculated the same way as it was for Figure~\ref{fig:distributions}.
 
-The figure clearly shows that, as the scaling factor increases, the relative performance of SPSS increases as well.
-In all cases, the performance of SPSS is superior to the dynamic algorithms for scaling factors greater than about 8.
-This suggests that, in general, for fine-grained workflows, the dynamic algorithms will produce better scores, and for coarse-grained workflows SPSS will produce better scores.
-
-\begin{figure*}[p]
+The figure shows that, as the scaling factor increases, the relative performance of SPSS increases as well.
+In this case SPSS performance surpasses DPDS and WA-DPDS for scaling factors larger than 2.
+In all other cases (other ensemble types for both applications), the performance of SPSS is superior to the dynamic algorithms for scaling factors greater than about 8.
+This suggests that, in general, for fine-grained workflows the dynamic algorithms will produce better scores, and for coarse-grained workflows SPSS will produce better scores.
+
+\begin{figure}[htb]
     \centering
-    \subfloat[CyberShake]{
-        \includegraphics[width=0.75\textwidth]{stretching_cybershake}
-    }\\
-    \subfloat[Montage]{
-        \includegraphics[width=0.75\textwidth]{stretching_montage}
-    }
-    \caption[Percentage of high scores achieved by each algorithm for Montage and Cybershake ensembles when task runtime is stretched]{Percentage of high scores achieved by each algorithm for Montage and Cybershake ensembles when task runtime is stretched.
+    \includegraphics[width=0.4\textwidth]{stretching_cybershake}
+    \caption[Percentage of high scores achieved by each algorithm for Cybershake pareto unsorted ensembles when task runtime is stretched]{Percentage of high scores achieved by each algorithm for Cybershake pareto unsorted ensembles when task runtime is stretched.
 A stretching factor of $x$ means that the runtime of tasks in each workflow was multiplied by $x$.}
     \label{fig:stretching}
-\end{figure*}
+\end{figure}
 

### 2012-04-24 paper/paper.tex
Update SPSS planning time figure/text

@@
 
-\subsection{Failures}
+\subsection{Task Failures}
 \label{sec:failures}
@@
 
-Figure~\ref{fig:spss_planning_time} shows the SPSS planning time for three different ensemble sizes (50, 100, 200), and five different workflow sizes (50, 200, 400, 600, 800).
+Figure~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles of 100 workflows with five different workflow sizes: 50, 200, 400, 600, and 800 tasks.
 The ensembles were generated using a constant distribution equal to the workflow size desired.
@@
 Each box summarizes the results of 2 applications, 10 random seeds, 10 budgets and 10 deadlines, or 2000 simulations per box.
-For ensemble size 200, the results for workflow size 800 are omitted because the simulations took too long to get a representative sample.
-
-Figure~\ref{fig:spss_planning_time} shows that, for small ensembles and small workflows, the SPSS planning time is reasonable, taking on the order of tens of seconds to a few minutes.
-For larger ensembles of large workflows, however, the SPSS planning time can easily reach 10-15 minutes.
-Considering that the largest workflows used in this experiment are rather small (maximum of 800 tasks), and that real workflows are often much larger (workflows with tens of thousands of tasks are common, and workflows with hundreds of thousands or millions of tasks are rare but do exist), it is unlikely that SPSS will be practical for ensembles of very large workflows.
-
-\begin{figure*}[t]
-    \includegraphics[width=\textwidth]{spss_planning_time}
-    \caption{Planning time of SPSS algorithm for different ensemble and workflow sizes.}
+
+Figure~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS planning time is reasonable, taking on the order of tens of seconds to a few minutes.
+For larger ensembles of large workflows, however, the SPSS planning time can easily reach 10 minutes.
+Considering that the largest workflows used in this experiment are still relatively small (maximum of 800 tasks), and that real workflows are often much larger (workflows with tens of thousands of tasks are common, and workflows with hundreds of thousands or millions of tasks are not impossible), it is unlikely that SPSS will be practical for ensembles of very large workflows.
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics[width=0.4\textwidth]{spss_planning_time}
+    \caption{Planning time of SPSS algorithm for ensembles of 100 workflows and different workflow sizes.}
     \label{fig:spss_planning_time}
-\end{figure*}
+\end{figure}
 

### 2012-04-24 paper/paper.tex
split evaluation section
new boxplots

@@
 
-\section{Evaluation}
+\section{Evaluation Metods}
 \label{sec:performance}
@@
 
-\subsection{Example of Results}
+\section{Discussion of Results}
+\subsection{Results Example}
 
@@
     \subfloat[Ratio of Actual Cost to Budget]{
-        \includegraphics[width=0.45\textwidth]{run-finish-variations-test-0-output-cost-all}
+        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-cost-all}
         }
     \subfloat[Ratio of Actual Makespan to Deadline]{
-        \includegraphics[width=0.45\textwidth]{run-finish-variations-test-0-output-makespan-all}
-        }
+    \hspace{2cm}
+        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all} }
     \caption[Boxplots for budget and deadline ratios when runtime estimates are inaccurate]{Boxplots for budget and deadline ratios when runtime estimate error varies from $\pm$0\% to $\pm$50\% for all three algorithms.
@@
     \subfloat[Ratio of Actual Cost to Budget]{
-        \includegraphics[width=0.45\textwidth]{run-finish-delays-test-0-output-cost-montage-uniform_unsorted}
+        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-cost-montage-uniform_unsorted}
         }
     \subfloat[Ratio of Actual Makespan to Deadline]{
-        \includegraphics[width=0.45\textwidth]{run-finish-delays-test-0-output-dagfinish-montage-uniform_unsorted}
+        \hspace{2cm}
+        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-dagfinish-montage-uniform_unsorted}
         }
@@
     \subfloat[Ratio of Actual Cost to Budget]{
-        \includegraphics[width=0.45\textwidth]{run-finish-failures-test-0-output-cost-all}
+        \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-cost-all}
         }
     \subfloat[Ratio of Actual Makespan to Deadline]{
-        \includegraphics[width=0.45\textwidth]{run-finish-failures-test-0-output-makespan-all}
+        \hspace{2cm}
+        \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-makespan-all}
         }

### 2012-04-25 paper/paper.tex
Minor edits

@@
 
-\section{Evaluation Metods}
+\section{Evaluation Methods}
 \label{sec:performance}
 
-This section describes a simulation study of the relative performance of three algorithms: DPDS, WA-DPDS, and SPSS.
-SPDS is not evaluated because, being a simplified version of DPDS and WA-DPDS, it will always produce results worse than the other algorithms.
-
 \subsection{Simulator}
 
-To evaluate and compare the algorithms, we developed a cloud workflow simulator based on CloudSim~\cite{Calheiros2011}.
+To evaluate and compare the three proposed algorithms, we developed a cloud workflow simulator based on CloudSim~\cite{Calheiros2011}.
 Our simulation model consists of Cloud, VM and WorkflowEngine entities.
@@
 \section{Discussion of Results}
+
 \subsection{Results Example}

### 2012-04-25 paper/paper.tex
Update conclusions

@@
 
-In this paper we addressed the interesting and important new problem of
-scheduling and resource provisioning for scientific workflow ensembles
-on IaaS clouds. The problem is different from previous work on grid and
-utility grid scheduling in that cloud infrastructures provide more control
-over the resources, so the resource provisioning plan can be adjusted
-according to the requirements of the application. Therefore the problem
-space becomes larger: it requires not only an efficient mapping of tasks
-to available resources, but also the selection of the best resource
-provisioning plan. Formulating the problem as a maximization of the
-number of prioritized workflows completed from the ensemble is also novel and  
-requires workflows to be admitted or rejected based on their estimated
-resource demands. We believe that this bi-constrained problem is highly
-relevant because such constraints are commonly imposed on many real-world
-projects. The approach is also directly applicable to grid environments
-that provide resource reservations and charge service units for resource use.
-
-We have considered both static and dynamic scheduling approaches and
-developed the DPDS, WA-DPDS and SPSS algorithms based on these
-strategies. The algorithms were evaluated on ensembles of synthetic
-workflows, which were generated based on real scientific applications. For
-the purposes of evaluation we have developed a simulator that models: the
-cloud infrastructure, and a workflow engine with tightly-coupled scheduler
-and provisioner modules. The results of our simulation studies indicate
-that the two algorithms that take into account information about the
-structure of the workflow and task runtime estimates (WA-DPDS and SPSS)
-yield better results than the simple on-line priority-based scheduling
-strategy (DPDS).
-
-During our work, several interesting lessons were learned from simulating the execution of
-workflow ensembles on cloud infrastructures. One result is that the simple
-strategy used by DPDS performs relatively well in terms of the number of
-workflows completed and the high resource utilization it can achieve. However,
-some of these workflows may be of lower priority, and the high utilization
-may be achieved by running tasks from workflows that cannot be completed
-given the constraints. On the other hand, DPDS may be the only choice when
-estimates of task runtimes are not available in advance. We also observed
-that the WA-DPDS and SPSS algorithms are able to find better quality
-solutions in terms of the target metrics: this is primarily a result of the
-admission algorithms, which reject workflows that are too large to be
-completed given the constraints, and accept more smaller workflows that
-can be completed.
-
-Another lesson learned from our studies is the effect of task granularity on
-workflow ensembles. The Montage and CyberShake workflows have fine-grained tasks
-with execution times on the order of seconds to minutes, while LIGO, Epigenomics
-and SIPHT have coarser-grained tasks with runtimes ranging from several minutes
-to hours. We observed that the dynamic algorithms achieved better results on the
-fine grained workflows since the simple random scheduling strategy leads to good
-load-balancing and high resource utilization for these ensembles. On the other
-hand, when tasks are longer, the dynamic algorithms tend to use up the budget by
-running tasks from low-priority workflows while the more sophisticated static
-scheduling strategy of SPSS is more circumspect. In the future we plan to
-investigate further the effects of task granularity on ensemble scheduling in a more
-systematic fashion.
-
-Our current study suggests many areas for future work. One interesting
-issue that we plan to investigate is how the static and dynamic algorithms
-perform with uncertain task runtimes and failures, which are known to have
-an important influence on overall system performance~\cite{Sakellariou2010,Dongarra2007}.
-We also plan to extend our application and infrastructure model to explicitly
-include the various data storage options available on clouds (right now we only
-model data access as part of the task time and do not include data costs). A previous
-experimental study on that subject~\cite{Juve2010} suggests that the data
-demands of scientific workflows have a large impact on not only the
-execution time, but also on the cost of workflows in commercial clouds.
-Finally, we plan to investigate heterogeneous environments that include
-multiple VM types and cloud providers, including private and community clouds,
-which will make the problem even more complex and challenging~\cite{Marshall2010, Vockler2011, Juve2010}.
-
-We believe that the results of our study can be applied in practice to develop
-tools that assist researchers in planning their large-scale computational
-experiments. The estimates of costs, runtimes and number of workflows completed
-that can be obtained from both the static algorithms and from the simulation
-runs, constitute valuable hints for the planning decisions and assessing the
-associated trade-offs. For this purpose our future work includes also further
-extension of our experimental studies~\cite{Juve2009, Vockler2011} of workflows on
-available cloud infrastructures to develop more accurate performance and cost
-models of the applications. This should lead to better understanding of the
-problems and costs of doing science in the cloud.
+In this paper we addressed the interesting and important new problem of scheduling and resource provisioning for scientific workflow ensembles on IaaS clouds.
+The problem is different from previous work on grid and utility grid scheduling in that cloud infrastructures provide more control over the resources, so the resource provisioning plan can be adjusted according to the requirements of the application.
+Therefore the problem space becomes larger; it requires not only an efficient mapping of tasks to available resources, but also the selection of the best resource provisioning plan.
+Formulating the problem as a maximization of the number of prioritized workflows completed from the ensemble is also novel and requires workflows to be admitted or rejected based on their estimated resource demands.
+We believe that this bi-constrained problem is highly relevant because such constraints are commonly imposed on many real-world projects.
+The approach is also directly applicable to grid environments that provide resource reservations and charge service units for resource use.
+
+The DPDS, WA-DPDS and SPSS algorithms were developed based on both static and dynamic scheduling approaches.
+The algorithms were evaluated on ensembles of synthetic workflows, which were generated based on real scientific applications.
+For the purposes of evaluation we have developed a simulator that models the cloud infrastructure, and a workflow engine with tightly-coupled scheduler and provisioner modules.
+
+The results of our simulation studies indicate that the two algorithms that take into account information about the structure of the workflow and task runtime estimates (WA-DPDS and SPSS) yield better results than the simple priority-based scheduling strategy (DPDS), which makes provisioning decisions purely based on resource utilization.
+On the other hand, DPDS may be the only choice when estimates of task runtimes are not available in advance.
+
+In cases where there are no provisioning delays, task runtime estimates are good, and failures are rare, we found that the static algorithm performs significantly better than the dynamic algorithms.
+However, when runtime estimates are poor, delays are present, or failures are common, the plans produced by the static algorithm are obstructed and it frequently exceeds the budget and deadline constraints.
+In comparison, the dynamic algorithms are able to quickly adapt to even verly large delays, poor estimates, and failures to prevent the ensemble from exceeding the constraints.
+
+In addition, we found that SPSS tends to perform better on coarse-grained workflows than the dynamic algorithms.
+For wide, fine-grained workflows, such as CyberShake and Montage, however, the dynamic algorithms frequently produce performance that is as good or better than SPSS,  because they are better able to pack the smaller tasks onto idle VMs close to the deadline than SPSS, which distributes deadlines to tasks in a way that prevents them from starting late.
+
+For very large workflows and ensembles, we found that the planning time of the SPSS algorithm is prohibitive due to its $O(n^2)$ algorithmic complexity.
+Our simulation results showed that ensembles of 100 workflows with 800 tasks often took 10 minutes or more to plan.
+This suggests that ensembles of workflows with tens of thousands of tasks, which are commonly encountered in real workflow applications, would take many hours to plan.
+In comparison, the $O(n)$ complexity of the dynamic algorithms makes them much more attractive for large scale problems.
+
+This study suggests many areas for future work.
+It would be interesting to extend the application and infrastructure model to explicitly include the various data storage options available on clouds (right now data access is modeled as part of the task execution time and data costs are not considered).
+A previous experimental study \cite{Juve2010} suggests that the data demands of scientific workflows have a large impact on not only the execution time, but also on the cost of workflows in commercial clouds.
+Finally, it would be interesting to investigate heterogeneous environments that include multiple VM types and cloud providers, including private and community clouds, which will make the problem even more complex and challenging~\cite{Marshall2010, Vockler2011, Juve2010}.
+
+The results of this study can also be applied to develop tools that assist researchers in planning their large-scale computational experiments.
+The estimates of cost, runtime, and number of workflows completed that can be obtained from both the static algorithms and from the simulation runs, constitute valuable hints for planning ensembles and evaluating the associated trade-offs.
 
@@
 %\bibliographystyle{abbrv}
+
 % IEEE style
@@
 
+% This balances the columns by inserting a column break at the given reference
+\IEEEtriggeratref{13}
+
 \bibliography{paper}

### 2012-04-25 paper/paper.tex
authors

@@
 % IEEE Format
-\author{
-  \IEEEauthorblockN{Maciej Malawski}
-  \IEEEauthorblockA{AGH University of Science and Technology\\
-Dept. of Computer Science\\
-Krakow, Poland; and\\
-University of Notre Dame\\
-Center for Research Computing\\
-111 ITC, Notre Dame, IN, USA\\
-Email: mmalawsk@nd.edu}
-  \and
-  \IEEEauthorblockN{Gideon Juve and Ewa Deelman}
-  \IEEEauthorblockA{USC Information Sciences Institute\\
-4676 Admiralty Way\\
-Marina del Rey, CA, USA\\
-Email: \{gideon,deelman\}@isi.edu}
-  \and
-  \IEEEauthorblockN{Jarek Nabrzyski}
-  \IEEEauthorblockA{University of Notre Dame\\
-Center for Research Computing\\
-111 ITC, Notre Dame, IN, USA\\
-Email: naber@nd.edu}}
+\author{\IEEEauthorblockN{Maciej
+Malawski\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Gideon
+Juve\IEEEauthorrefmark{2}, Ewa Deelman\IEEEauthorrefmark{2} and Jarek
+Nabrzyski\IEEEauthorrefmark{3}} 
+\IEEEauthorblockA{\IEEEauthorrefmark{1}AGH University of Science and
+Technology, Dept. of Computer Science, Krakow, Poland, Email:
+malawski@agh.edu.pl} 
+\IEEEauthorblockA{\IEEEauthorrefmark{2}USC Information Sciences Institute, 4676
+Admiralty Way, Marina del Rey, CA, USA, Email: \{gideon,deelman\}@isi.edu} 
+\IEEEauthorblockA{\IEEEauthorrefmark{3}University of Notre Dame, Center for
+Research Computing, 111 ITC, Notre Dame, IN, USA, Email: naber@nd.edu}
+}
 

### 2012-04-25 paper/paper.tex
Reduce conclusion more

--- 2012-04-25 paper/paper.tex
@@
 
-% New rewritten text for section ``discussion of Results'' starts here
-
 \begin{figure}[tb]
@@
     }
-    \caption{Example performance of algorithms expressed in
-    sum of dag sizes and effective cost, data for LIGO application, Pareto
-    Sorted ensemble of 100 workflows, budget = \$600}
+    \caption{Example performance of algorithms expressed in sum of workflow sizes and effective cost, data for LIGO application, Pareto Sorted ensemble of 100 workflows, budget = \$600}
     \label{fig:size-cost}
@@
 
-%To analyze the behaviour of algorithms we ran simulations for constraint values
-%ranging between the estimated limits indicated in
-%Section~\ref{sec:exp-parameters}, resulting in $10 * 10 = 100$ (budget,
-%deadline) data points for each of 5 applications, 3 algorithms and using 5
-%ensemble size distributions and 10 random seeds. From this result space of
-%75000 we selected representative plots.
-
-To illustrate the behavior of the algorithms we show here examples of
-performance curves obtained for selected sample parameters.
-Fig.~\ref{fig:size-cost}a uses the total computing time metric, $T_C$ for LIGO
-application with pareto Sorted distribution of workflows. It can be seen that
-SPSS gives the best performance and WA-DPDS performs better than the
-workflow-unaware version, DPDS. This means that for the workflows we tested the
-simple admission procedure based on estimation of workflow cost does not degrade
-the solution, but rather in many cases it enables larger workflows that would
-lead to budget overrun to be rejected. Thus it can save the space for smaller
-workflows to complete.
-
-This plot allows us observe a more general property: for a given budget,
-when the deadline is tightened the amount of work done steeply decreases.
-This is partially due to the fact that fewer workflows can be
-completed given a short deadline because their critical paths are simply too long
-and they are squeezed out of early start times by high priority workflows.
-In addition, assigning a shorter deadline means that more VMs need to be
-provisioned to complete the work. However, because of the dependencies in the
-workflows, many of these VMs will either be idle, or will be backfilled with
-tasks from low-priority workflows that never finish, resulting, effectively,
-in reduced resource utilization and fewer completed workflows.
-
-These observations are further confirmed when examining the effective computing
-cost, as shown in Figure~\ref{fig:size-cost}b. It can be seen that lower
-resource utilization and lower parallel efficiency for short deadlines lead to
-substantial increases in the effective computing cost. This increase can be
-significant. For example, for short deadlines an effective computing cost
-is $\sim$10--40\% higher than the minimum effective cost of \$1/hour, depending
-on the algorithm used. This means that out of the \$600 budget,
-$\sim$\$60--\$240 is spent on idle resources.
-
-Another general conclusion can be drawn from Figure~\ref{fig:size-cost}b:
-all the curves for all the applications and budgets have a similar shape. This
-shape represents a trade-off curve between two conflicting objectives: time and cost.
-We observe that the selection of deadline and cost can be formulated as a
-multi-criteria optimization problem, and that the curves presented in
-Figure~\ref{fig:size-cost}b are approximations of the Pareto front (or Pareto
-set) of solutions to this problem. In our case the solutions were achieved by maximizing
-the amount of work done, which is the inverse of minimizing the effective
-computing cost metric for a given deadline. Therefore, the obtained results
-can be used to aid in multi-criteria resource allocation planning, when both
-cost and time criteria are given not as constraints but as objectives.
+%To analyze the behavior of the algorithms we ran simulations for constraint values ranging between the estimated limits indicated in Section~\ref{sec:exp-parameters}, resulting in $10 * 10 = 100$ (budget, deadline) data points for each of 5 applications, 3 algorithms and using 5 ensemble size distributions and 10 random seeds.
+From this result space of 75,000 we selected representative plots.
+
+To illustrate the behavior of the algorithms we show here examples of performance curves obtained for selected sample parameters.
+Fig.~\ref{fig:size-cost}a uses the total computing time metric, $T_C$ for LIGO application with Pareto Sorted distribution of workflows.
+It can be seen that SPSS gives the best performance and WA-DPDS performs better than the workflow-unaware version, DPDS.
+This means that for the workflows we tested the simple admission procedure based on estimation of workflow cost does not degrade the solution, but rather in many cases it enables larger workflows that would lead to budget overrun to be rejected.
+Thus it can save the space for smaller workflows to complete.
+
+This plot allows us observe a more general property: for a given budget, when the deadline is tightened the amount of work done steeply decreases.
+This is partially due to the fact that fewer workflows can be completed given a short deadline because their critical paths are simply too long and they are squeezed out of early start times by high priority workflows.
+In addition, assigning a shorter deadline means that more VMs need to be provisioned to complete the work.
+However, because of the dependencies in the workflows, many of these VMs will either be idle, or will be backfilled with tasks from low-priority workflows that never finish, resulting, effectively, in reduced resource utilization and fewer completed workflows.
+
+These observations are further confirmed when examining the effective computing cost, as shown in Figure~\ref{fig:size-cost}b.
+It can be seen that lower resource utilization and lower parallel efficiency for short deadlines lead to substantial increases in the effective computing cost.
+This increase can be significant.
+For example, for short deadlines an effective computing cost is $\sim$10--40\% higher than the minimum effective cost of \$1/hour, depending on the algorithm used.
+This means that out of the \$600 budget, $\sim$\$60--\$240 is spent on idle resources.
+
+Another general conclusion can be drawn from Figure~\ref{fig:size-cost}b: all the curves for all the applications and budgets have a similar shape.
+This shape represents a trade-off curve between two conflicting objectives: time and cost.
+We observe that the selection of deadline and cost can be formulated as a multi-criteria optimization problem, and that the curves presented in Figure~\ref{fig:size-cost}b are approximations of the Pareto front (or Pareto set) of solutions to this problem.
+In our case the solutions were achieved by maximizing the amount of work done, which is the inverse of minimizing the effective computing cost metric for a given deadline.
+Therefore, the obtained results can be used to aid in multi-criteria resource allocation planning, when both cost and time criteria are given not as constraints but as objectives.
 
@@
 
-The goal of the first experiment is to characterize the relative performance of the DPDS, WA-DPDS, and SPSS algorithms.
-This was done by simulating the algorithms on many different ensembles and comparing the scores computed using the technique described in Section~\ref{sec:perf_metric}.
+The goal of the the next experiment is to characterize the relative performance of the proposed algorithms.
+This was done by simulating the algorithms on many different ensembles and comparing the scores computed using the exponential scoring technique described in Section~\ref{sec:perf_metric}.
 The algorithm that achieves the largest number of high scores is judged to have the best performance.
@@
 In this paper we addressed the interesting and important new problem of scheduling and resource provisioning for scientific workflow ensembles on IaaS clouds.
-The problem is different from previous work on grid and utility grid scheduling in that cloud infrastructures provide more control over the resources, so the resource provisioning plan can be adjusted according to the requirements of the application.
+The goal of this work is to maximize the number of prioritized workflows that can be completed given budget and deadline constraints.
+
+%This problem differs from previous work on grid and utility grid scheduling in that cloud infrastructures provide more control over provisioning, so that the number of resources can be adjusted according to the requirements of the application.
 Therefore the problem space becomes larger; it requires not only an efficient mapping of tasks to available resources, but also the selection of the best resource provisioning plan.
-Formulating the problem as a maximization of the number of prioritized workflows completed from the ensemble is also novel and requires workflows to be admitted or rejected based on their estimated resource demands.
+
+%Formulating the problem as a maximization of the number of prioritized workflows completed from the ensemble is also novel and requires workflows to be admitted or rejected based on their estimated resource demands.
 We believe that this bi-constrained problem is highly relevant because such constraints are commonly imposed on many real-world projects.
@@
 
-The DPDS, WA-DPDS and SPSS algorithms were developed based on both static and dynamic scheduling approaches.
-The algorithms were evaluated on ensembles of synthetic workflows, which were generated based on real scientific applications.
-For the purposes of evaluation we have developed a simulator that models the cloud infrastructure, and a workflow engine with tightly-coupled scheduler and provisioner modules.
-
-The results of our simulation studies indicate that the two algorithms that take into account information about the structure of the workflow and task runtime estimates (WA-DPDS and SPSS) yield better results than the simple priority-based scheduling strategy (DPDS), which makes provisioning decisions purely based on resource utilization.
-On the other hand, DPDS may be the only choice when estimates of task runtimes are not available in advance.
-
-In cases where there are no provisioning delays, task runtime estimates are good, and failures are rare, we found that the static algorithm performs significantly better than the dynamic algorithms.
-However, when runtime estimates are poor, delays are present, or failures are common, the plans produced by the static algorithm are obstructed and it frequently exceeds the budget and deadline constraints.
-In comparison, the dynamic algorithms are able to quickly adapt to even verly large delays, poor estimates, and failures to prevent the ensemble from exceeding the constraints.
-
-In addition, we found that SPSS tends to perform better on coarse-grained workflows than the dynamic algorithms.
+We developed three algorithms to solve this problem: two dynamic algorithms, DPDS and WA-DPDS, and one static algorithm, SPSS.
+The algorithms were evaluated via simulation on ensembles of synthetic workflows, which were generated based on statistics from real scientific applications.
+
+The results of our simulation studies indicate that the two algorithms that take into account the structure of the workflow and task runtime estimates (WA-DPDS and SPSS) yield better results than the simple priority-based scheduling strategy (DPDS), which makes provisioning decisions based purely on resource utilization.
+
+In cases where there are no provisioning delays, task runtime estimates are good, and failures are rare, we found that the static algorithm, SPSS, performs significantly better than both dynamic algorithms.
+However, when conditions are less than perfect, the static plans produced by SPSS are disrupted and it frequently exceeds the budget and deadline constraints.
+In comparison, the dynamic algorithms are able to adapt to a wide variety of conditions, and rarely exceed the constraints even with long delays, poor estimates, and high failure rates.
+
+%In addition, we found that SPSS tends to perform better on coarse-grained workflows than the dynamic algorithms.
 For wide, fine-grained workflows, such as CyberShake and Montage, however, the dynamic algorithms frequently produce performance that is as good or better than SPSS, because they are better able to pack the smaller tasks onto idle VMs close to the deadline than SPSS, which distributes deadlines to tasks in a way that prevents them from starting late.
 
-For very large workflows and ensembles, we found that the planning time of the SPSS algorithm is prohibitive due to its $O(n^2)$ algorithmic complexity.
-Our simulation results showed that ensembles of 100 workflows with 800 tasks often took 10 minutes or more to plan.
-This suggests that ensembles of workflows with tens of thousands of tasks, which are commonly encountered in real workflow applications, would take many hours to plan.
-In comparison, the $O(n)$ complexity of the dynamic algorithms makes them much more attractive for large scale problems.
-
-This study suggests many areas for future work.
-It would be interesting to extend the application and infrastructure model to explicitly include the various data storage options available on clouds (right now data access is modeled as part of the task execution time and data costs are not considered).
+%For very large workflows and ensembles, we found that the planning time of the SPSS algorithm is prohibitive.
+SPSS often took 10 minutes or more to plan ensembles of 100 workflows with 800 tasks each.
+This suggests that ensembles of workflows with tens of thousands of tasks, which are commonly encountered in real workflow applications, would take many hours to plan using SPSS.
+This makes the dynamic algorithms much more attractive for large scale problems.
+
+This study suggests several areas for future work.
+Our current approach models data access as part of the task execution time and does not consider data storage and transfer costs.
+In the future we plan to extend the application and infrastructure model to  include the various data storage options available on clouds.
 A previous experimental study \cite{Juve2010} suggests that the data demands of scientific workflows have a large impact on not only the execution time, but also on the cost of workflows in commercial clouds.
-Finally, it would be interesting to investigate heterogeneous environments that include multiple VM types and cloud providers, including private and community clouds, which will make the problem even more complex and challenging~\cite{Marshall2010, Vockler2011, Juve2010}.
-
-The results of this study can also be applied to develop tools that assist researchers in planning their large-scale computational experiments.
+We also plan to investigate heterogeneous environments that include multiple VM types and cloud providers, including private and community clouds, which will make the problem even more complex and challenging.
+
+%The results of this study can be applied to develop tools that assist researchers in planning their large-scale computational experiments.
 The estimates of cost, runtime, and number of workflows completed that can be obtained from both the static algorithms and from the simulation runs, constitute valuable hints for planning ensembles and evaluating the associated trade-offs.

### 2012-04-25 paper/paper.tex
Remove Results Example section

@@
 \usepackage{subfig} % for \subfloat
+\usepackage{verbatim}
 
@@
 
-In order to assess the relative performance of our algorithms, we have defined the following metrics:
-
-\begin{itemize}
-  \item $N_c$: number of workflows completed,
-  \item $T_C$: total amount of computing time completed,
-  \item $C_{eff}$: effective cost of computing per hour, which includes overheads due to the cost of idle resources.
-  \item $s$: user preference score (utility function),
-\end{itemize}
-
-These metrics measure the relative performance of the different algorithms on a given problem (ensemble, budget, and deadline).
-
-The simplest approach is to simply count the number of workflows in the ensemble that each algorithm is able to complete within the budget, before the deadline, $N_c$.
-
-The total amount of computing time completed, $T_C$, is the sum of the runtimes of all tasks in all completed workflows.
-Because we assumed that the resource cost per hour is equal to \$1, the total computing time can be easily compared to the allocated budget $b$.
-Their ratio: $u = (T_C * \$1)/b$, is the resource utilization, and its inverse: $C_{eff} = b/T_C$, is the effective resource cost in dollars per hour.
-To calculate $C_{eff}$ and $u$ we use the value of allocated budget $b$, not the actual total cost, since we are interested in the scenarios when the budget is limited and does not allow completing all the workflows from the ensemble.
-
-The $N_c$, $T_C$ and $C_{eff}$ metrics do not account for the priority-based utility function specified by the user.
-Using $N_c$, for example, a less efficient algorithm may be able to complete a large number of low-priority workflows by executing the smallest workflows first.
-In order to account for the user's priority, we used an exponential scoring approach where the score $s$ for an ensemble $e$ is:
-%
-$$
-s = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
+In order to assess the relative performance of our algorithms it is necessary to define a metric that can be used to score the performance of the different algorithms on a given problem (ensemble, budget, and deadline).
+The simplest approach is to simply count the number of workflows in the ensemble that each algorithm is able to complete within the budget before the deadline.
+The problem with this approach is that it does not account for the priority-based utility function specified by the user.
+Using the counting approach, a less efficient algorithm may be able to complete a large number of low-priority workflows by executing the smallest workflows first.
+
+In order to account for the user's priority, we used an exponential scoring approach where the score for an ensemble $e$ is:
+%
+$$
+Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
 $$
@@
 %
-This metric is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.
+This scoring is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.
 
 \section{Discussion of Results}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{comment} % Remove the Results Example section
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{comment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Relative Performance of Algorithms}
 
-The goal of the the next experiment is to characterize the relative performance of the proposed algorithms.
+The goal of the first experiment is to characterize the relative performance of the proposed algorithms.
 This was done by simulating the algorithms on many different ensembles and comparing the scores computed using the exponential scoring technique described in Section~\ref{sec:perf_metric}.
-The algorithm that achieves the largest number of high scores is judged to have the best performance.
+In this experiment, the algorithm that achieves the largest number of high scores has the best performance.
 

### 2012-04-25 paper/paper.tex
Really remove the Results Example section

@@
 \usepackage{subfig} % for \subfloat
-\usepackage{verbatim}
 
@@
 
-\subsection{Performance Metrics}
+\subsection{Performance Metric}
 \label{sec:perf_metric}
@@
 \section{Discussion of Results}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{comment} % Remove the Results Example section
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Results Example}
-
-\begin{figure}[tb]
-    \centering
-    \subfloat[size]{
-        \includegraphics[width=0.6\columnwidth]{size-selected-LIGO}
-    }\\
-    \subfloat[cost]{
-        \includegraphics[width=0.6\columnwidth]{cost-selectedLIGO}
-    }
-    \caption{Example performance of algorithms expressed in sum of workflow sizes and effective cost, data for LIGO application, Pareto Sorted ensemble of 100 workflows, budget = \$600}
-    \label{fig:size-cost}
-\end{figure}
-
-%To analyze the behavior of the algorithms we ran simulations for constraint values ranging between the estimated limits indicated in Section~\ref{sec:exp-parameters}, resulting in $10 * 10 = 100$ (budget, deadline) data points for each of 5 applications, 3 algorithms and using 5 ensemble size distributions and 10 random seeds.
-From this result space of 75,000 we selected representative plots.
-
-To illustrate the behavior of the algorithms we show here examples of performance curves obtained for selected sample parameters.
-Fig.~\ref{fig:size-cost}a uses the total computing time metric, $T_C$ for LIGO application with Pareto Sorted distribution of workflows.
-It can be seen that SPSS gives the best performance and WA-DPDS performs better than the workflow-unaware version, DPDS.
-This means that for the workflows we tested the simple admission procedure based on estimation of workflow cost does not degrade the solution, but rather in many cases it enables larger workflows that would lead to budget overrun to be rejected.
-Thus it can save the space for smaller workflows to complete.
-
-This plot allows us observe a more general property: for a given budget, when the deadline is tightened the amount of work done steeply decreases.
-This is partially due to the fact that fewer workflows can be completed given a short deadline because their critical paths are simply too long and they are squeezed out of early start times by high priority workflows.
-In addition, assigning a shorter deadline means that more VMs need to be provisioned to complete the work.
-However, because of the dependencies in the workflows, many of these VMs will either be idle, or will be backfilled with tasks from low-priority workflows that never finish, resulting, effectively, in reduced resource utilization and fewer completed workflows.
-
-These observations are further confirmed when examining the effective computing cost, as shown in Figure~\ref{fig:size-cost}b.
-It can be seen that lower resource utilization and lower parallel efficiency for short deadlines lead to substantial increases in the effective computing cost.
-This increase can be significant.
-For example, for short deadlines an effective computing cost is $\sim$10--40\% higher than the minimum effective cost of \$1/hour, depending on the algorithm used.
-This means that out of the \$600 budget, $\sim$\$60--\$240 is spent on idle resources.
-
-Another general conclusion can be drawn from Figure~\ref{fig:size-cost}b: all the curves for all the applications and budgets have a similar shape.
-This shape represents a trade-off curve between two conflicting objectives: time and cost.
-We observe that the selection of deadline and cost can be formulated as a multi-criteria optimization problem, and that the curves presented in Figure~\ref{fig:size-cost}b are approximations of the Pareto front (or Pareto set) of solutions to this problem.
-In our case the solutions were achieved by maximizing the amount of work done, which is the inverse of minimizing the effective computing cost metric for a given deadline.
-Therefore, the obtained results can be used to aid in multi-criteria resource allocation planning, when both cost and time criteria are given not as constraints but as objectives.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{comment}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2012-04-25 paper/paper.tex
intro and related work

@@
 Scientific workflows, usually represented as directed acyclic graphs (DAGs), are
-an important class of distributed applications that have been studied in the context of
-resource management and task scheduling on grid and utility computing systems.
-Workflow applications for large computational problems are often composed of 
-several inter-related workflows grouped into {\em ensembles}. All the
-workflows in an ensemble typically have a similar structure, but they differ in their
-input data, number of tasks, and individual task sizes. The CyberShake 
-application~\cite{Callaghan2011} is a good example of an application that requires 
-scientific workflow ensembles. CyberShake computes seismic hazard curves for a given
-geographical location. In order to produce a hazard map for a region such as 
-Southern California, many hazard curves need to be computed for a large number of 
-geographical sites, where each site requires running a single, complex scientific 
-workflow. Workflows for individual sites may differ not only in their parameters, 
-but also in their priority. For example, some sites may be in heavily populated areas 
-or in strategic locations such as power plants, while others may be less important. 
-Other examples of workflow ensembles come from astronomical applications that create
-science-grade image mosaics~\cite{Deelman2008} or search for extrasolar 
-planets~\cite{Vockler2011}, where multiple, different workflows are used to
-process data for different areas of the sky.
+an important class of distributed applications that lead to challenging problems
+in resource management on grid and utility computing systems. Workflow
+applications for large computational problems are often composed of several
+inter-related workflows grouped into {\em ensembles}. All the workflows in an
+ensemble typically have a similar structure, but they differ in their input
+data, number of tasks, and individual task sizes.
+
+The CyberShake~\cite{Callaghan2011} is a good example of an application that
+requires scientific workflow ensembles. Generating a hazard map for a region
+such as Southern California requires an ensemble of many related workflows to
+compute individual hazard curves. The hazard map generated by the CyberShake
+group in 2009, for example, required an ensemble of 239 workflows. In astronomy,
+for example, a user might wish to use multiple Montage~\cite{Deelman2008}
+workflows with different parameters to generate a set of image mosaics that
+cover a given area of the sky in different wavelengths, or use an ensemble to generate
+the parts of a very large mosaic. The Pegasus team is now working on a Galactic
+Plane project that is focused on generating a mosaic of the entire sky. This
+ensemble consists of 17 workflows, each of which contains 900 sub-workflows and
+computes one ``tile'' of the full mosaic. Another ensemble example is the
+Periodograms application~\cite{Vockler2011} that searches for extrasolar planet
+by detecting the periodic dips in light intensity. Due to the large scale of the
+input data, this application is often split up into multiple batches processed
+by different workflow. Additional workflows are created to run the analysis
+using different parameters. A recent analysis of data from the Kepler satellite
+required three ensembles of 15 workflows.
+
+Ensemble workflows may differ not only in their parameters, but also in their
+priority. For example, in CyberShake some sites may be in heavily populated
+areas or in strategic locations such as power plants, while others may be less
+important. Scientists typically prioritize the workflows in an ensemble such
+that important workflows are finished first. This enables them to see critical
+results early, and helps them to choose the most important workflows when the
+time and financial resources available for running the computations are limited.
 
@@
 
-In this paper, we aim to gain insight into resource management challenges when executing 
-scientific workflow ensembles on clouds. In particular, we address a new and important 
-problem of maximizing the number of completed workflows from an ensemble under both budget and
-deadline constraints. The motivation for this work is to answer the fundamental question of 
-concern to a researcher: How much computation can be completed given the limited budget and 
-timeframe of a research project? The goals of this paper are to discuss and assess possible
-static and dynamic strategies for both task scheduling and resource
-provisioning. Based on the knowledge of workflow scheduling algorithms we
-analyze strategies for on-line scheduling of individual tasks as well as static
-algorithms that rely on the information about the workflow structure (critical
-paths and workflow levels) and estimates of task runtimes. In addition, we
-analyze a hybrid workflow-aware dynamic scheduling algorithm, which uses the
-workflow structure information to estimate which workflows should be rejected
-from the ensemble based on the constraints. We evaluate these algorithms using a
-simulator we have developed using CloudSim~\cite{Calheiros2011}, which models the 
-infrastructure and the application. The simulator includes also tightly-coupled 
-scheduler and provisioner modules. The algorithms were evaluated on a set 
-of scientific workflow ensembles from several different applications using a 
-broad range of budget and deadline parameters.
+In this paper, we aim to gain insight into resource management challenges when
+executing scientific workflow ensembles on clouds. In particular, we address a
+new and important problem of maximizing the number of completed workflows from
+an ensemble under both budget and deadline constraints. The motivation for this
+work is to answer the fundamental question of concern to a researcher: How much
+computation can be completed given the limited budget and timeframe of a
+research project?
 
@@
 \begin{itemize}
-  \item a definition of the problem of scheduling prioritized workflows grouped into
-  ensembles, under both budget and deadline constraints,
-  \item an analysis of static and dynamic strategies for task scheduling and
-  resource provisioning on clouds, together with admission procedures to
-  predict which workflows from the ensemble can be completed,
-  \item the development of several static and dynamic algorithms and of a simulator based
-  on the defined application and infrastructure model,
-  \item a performance evaluation of the algorithms on a set of synthetic benchmark
-  workflow ensembles based on important, real scientific applications,
-  \item an analysis of performance results that assesses the advantages and
-  disadvantages of the proposed algorithms on different application scenarios and
+  \item we define the problem of scheduling prioritized workflows grouped
+  into ensembles, under both budget and deadline constraints,
+  \item we analyze and develop dynamic (on-line) strategies for task scheduling
+  and resource provisioning as well as static algorithms that
+  rely on the information about the workflow structure (critical paths and
+  workflow levels) and estimates of task runtimes. 
+  \item we analyze and develop a hybrid workflow-aware dynamic scheduling
+  algorithm, together with admission procedures to predict which workflows from
+  the ensemble can be completed,
+  \item we evaluate these algorithms using a simulator we have
+  developed using CloudSim~\cite{Calheiros2011}, which models the infrastructure
+  and the application, taking into account uncertainties in task runtime
+  estimations, resource provisioning delays and failures, 
+  \item we discuss the performance of the algorithms on a set of
+  synthetic benchmark workflow ensembles based on important, real scientific
+  applications, in the presence of such of such factors as 
+  uncertainties in task runtime estimations, resource provisioning delays and
+  failures, on a broad range of different application scenarios and
   varying constraint values.
@@
 Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
-infrastructure and application model we are targeting in this paper.
-In Section~\ref{sec:algorithms} we describe the dynamic and
-static algorithms we developed. Section~\ref{sec:performance} presents test
-scenarios, performance metrics, and the results of simulation studies of these algorithms.
-Finally, general conclusions, lessons learned, and future work are outlined in
-Section~\ref{sec:conclusions}.
+infrastructure and application model we are targeting in this paper. In
+Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
+developed. Section~\ref{sec:performance} presents test scenarios and performance
+metrics, while Section~\ref{sec:results} discusses the results of simulation
+studies of these algorithms. Finally, general conclusions, lessons learned, and
+future work are outlined in Section~\ref{sec:conclusions}.
 
@@
 General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
-Auto Scaling\footnote{\url{http://aws.amazon.com/autoscaling}} and
-RightScale\footnote{\url{http://www.rightscale.com}}) allow adjusting the size
+Auto Scaling~\cite{Autoscaling} and
+RightScale~\cite{RightScale}) allow adjusting the size
 of resource pool based on metrics related to infrastructure and application.
@@
 \section{Discussion of Results}
+\label{sec:results}
 

### 2012-04-25 paper/paper.tex
Sections I-V shortened.

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Cost- and Deadline-Constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds}
+\title{Cost- and Deadline-Constrained Provisioning for Scientific Workflow
+Ensembles in IaaS Clouds}
 
@@
 \section{Introduction}
+\label{sec:intro}
 
 Scientific workflows, usually represented as directed acyclic graphs (DAGs), are
-an important class of distributed applications that lead to challenging problems
-in resource management on grid and utility computing systems. Workflow
-applications for large computational problems are often composed of several
+an important class of applications that lead to challenging problems
+in resource management on grid and utility computing systems. Workflows
+for large computational problems are often composed of several
 inter-related workflows grouped into {\em ensembles}. All the workflows in an
@@
 such as Southern California requires an ensemble of many related workflows to
-compute individual hazard curves. The hazard map generated by the CyberShake
-group in 2009, for example, required an ensemble of 239 workflows. In astronomy,
+compute individual hazard curves. E.g. the hazard map generated by the
+CyberShake group in 2009 required an ensemble of 239 workflows. In astronomy,
 for example, a user might wish to use multiple Montage~\cite{Deelman2008}
@@
 computes one ``tile'' of the full mosaic. Another ensemble example is the
-Periodograms application~\cite{Vockler2011} that searches for extrasolar planet
+Periodograms application~\cite{Vockler2011} that searches for extrasolar planets
 by detecting the periodic dips in light intensity. Due to the large scale of the
 input data, this application is often split up into multiple batches processed
-by different workflow. Additional workflows are created to run the analysis
+by different workflows. Additional workflows are created to run the analysis
 using different parameters. A recent analysis of data from the Kepler satellite
@@
 are regarded by the scientific community as a potentially attractive source of 
-low-cost, on-demand computing resources~\cite{Ostermann2010, Keahey2009}. In contrast 
+low-cost computing resources~\cite{Ostermann2010, Keahey2009}. In contrast 
 to clusters and grids, which typically offer best-effort quality of service, clouds 
-give users more flexibility in creating a controlled and managed computing environment.
-Clouds provide the ability to adjust resource capacity according to the dynamically changing
-computing demands of the application, often called auto-scaling. However, giving users 
-more control over the resources also requires the development of new methods for task 
+give more flexibility in creating a controlled and managed computing environment.
+Clouds provide the ability to adjust resource capacity according to the changing
+demands of the application, often called auto-scaling. However, giving users 
+more control also requires the development of new methods for task 
 scheduling and resource provisioning. The resource management decisions required 
@@
 metrics such as workflow makespan or resource utilization, but must also now consider
-budget constraints, since the resources from public (commercial) cloud providers 
+budget constraints, since the resources from public (commercial) clouds
 usually have monetary costs associated with them~\cite{Durkee2010}.
@@
   the ensemble can be completed,
-  \item we evaluate these algorithms using a simulator we have
-  developed using CloudSim~\cite{Calheiros2011}, which models the infrastructure
-  and the application, taking into account uncertainties in task runtime
-  estimations, resource provisioning delays and failures, 
+  \item we evaluate these algorithms using our simulator based on
+  CloudSim~\cite{Calheiros2011}, which models the infrastructure and the application, 
+  taking into account uncertainties in task runtime
+  estimations, provisioning delays and failures, 
   \item we discuss the performance of the algorithms on a set of
-  synthetic benchmark workflow ensembles based on important, real scientific
-  applications, in the presence of such of such factors as 
-  uncertainties in task runtime estimations, resource provisioning delays and
-  failures, on a broad range of different application scenarios and
+  synthetic workflow ensembles based on important, real scientific
+  applications, on a broad range of different application scenarios and
   varying constraint values.
@@
 
-We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2).
-In this model, virtual machine (VM) instances are requested on-demand.
-The VMs are billed by the hour, with partial hours being rounded up.
-Although there are multiple VM types with different amounts of CPU, memory, disk space, and I/O, for this thesis we focus on a single VM type because we assume that for most applications there will typically be only one or two VM types with the best price/performance ratio for the application~\cite{Juve2009}.
-We also assume that a submitted task has exclusive access to a VM instance and that there is no preemption.
-
-We assume that there is a delay between the time that a new VM instance is requested and when it becomes available to execute tasks.
-In practice this delay is typically on the order of tens of seconds to a few minutes, and is highly dependent upon the cloud platform and the VM image size~\cite{Nurmi2008b}.
+We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2),
+where virtual machine (VM) instances are requested on-demand and are billed by
+the hour, with partial hours being rounded up. Although there may be
+heterogeneous VM types with different amounts of CPU, memory, disk space, and
+I/O, for this paper we focus on a single VM type because we assume that for most
+applications there will typically be only one or two VM types with the best
+price/performance ratio for the application~\cite{Juve2009}. We also assume that
+a submitted task has exclusive access to a VM instance and that there is no
+preemption. We assume that there is a delay between the time that a new VM
+instance is requested and when it becomes available to execute tasks. In
+practice this delay is typically on the order of tens of seconds to a few
+minutes, and is highly dependent upon the cloud platform and the VM image
+size~\cite{Nurmi2008b}.
 
@@
 
-The target applications are ensembles of scientific workflows that can be modeled as Directed Acyclic Graphs (DAGs), where the nodes in the graph represent computational tasks, and the edges represent data- or control-flow dependencies between the tasks.
-
-We assume that there are runtime estimates for each task in the workflow based on either a performance model of the application, or historical data that can be mined.
-Additionally, we assume that the runtime estimates for individual workflow tasks are not perfect, but may vary based on a uniform distribution of $\pm p\%$.
-
-This study uses synthetic workflows that were generated using historical data from real applications~\cite{Bharathi2008}.
-The applications come from a wide variety of domains including: bioinformatics (Epigenomics, SIPHT: sRNA identification protocol using high-throughput technology), astronomy (Montage), earthquake science (CyberShake), and physics (LIGO).
-The runtimes used to generate the synthetic workflows are based on distributions gathered from running real workflows.
-The synthetic workflows were generated using code developed by Bharathi, et al~\cite{WorkflowGenerator}.
-
-Although scientific workflows are often data-intensive, the algorithms described in the next section do not currently consider the size of input and output data when scheduling tasks.
-Instead it is assumed that all workflow data is stored in a shared cloud storage system, such as Amazon S3, and that intermediate data transfer times are included in task runtime estimates.
-It is also assumed that data transfer times between the shared storage and the VMs are equal for different VMs so that task placement decisions do not have an impact on the runtime of the tasks.
-
-This work focuses on scheduling and provisioning for workflow ensembles.
-A workflow ensemble consists of several related workflows that a user wishes to run.
-Each workflow in an ensemble is given a numeric priority that indicates how important the workflow is to the user.
-As such, the priorities indicate the utility function of the user.
-These priorities are absolute in the sense that completing a workflow with a given priority is more valuable (gives the user more utility) than completing all other workflows in the ensemble with lower priorities combined.
-
-The goal of the workflow ensemble scheduling and cloud provisioning problem is to complete as many high-priority workflows as possible given a fixed budget and deadline.
-Only workflows for which all tasks are finished by the deadline are considered to be complete---partial results are not usable in this model.
+The target applications are ensembles of scientific workflows that can be
+modeled as Directed Acyclic Graphs (DAGs), where the nodes in the graph
+represent computational tasks, and the edges represent data- or control-flow
+dependencies between the tasks. We assume that there are runtime estimates for
+each task in the workflow based on either a performance model of the
+application, or historical data that can be mined. Additionally, we assume that
+the runtime estimates for individual workflow tasks are not perfect, but may
+vary based on a uniform distribution of $\pm p\%$.
+
+This study uses synthetic workflows that were generated using historical data
+from real applications~\cite{Bharathi2008}. The applications come from a wide
+variety of domains including: bioinformatics (Epigenomics, SIPHT: sRNA
+identification protocol using high-throughput technology), astronomy (Montage),
+earthquake science (CyberShake), and physics (LIGO). The runtimes used to
+generate the synthetic workflows are based on distributions gathered from
+running real workflows. The synthetic workflows were generated using code
+developed by Bharathi, et al~\cite{WorkflowGenerator}.
+
+Although scientific workflows are often data-intensive, the algorithms described
+in the next section do not currently consider the size of input and output data
+when scheduling tasks. Instead it is assumed that all workflow data is stored in
+a shared cloud storage system, such as Amazon S3, and that intermediate data
+transfer times are included in task runtime estimates. It is also assumed that
+data transfer times between the shared storage and the VMs are equal for
+different VMs so that task placement decisions do not have an impact on the
+runtime of the tasks.
+
+A workflow ensemble consists of several related workflows that a user wishes to
+run. Each workflow in an ensemble is given a numeric priority that indicates how
+important the workflow is to the user. As such, the priorities indicate the
+utility function of the user. These priorities are absolute in the sense that
+completing a workflow with a given priority is more valuable (gives the user
+more utility) than completing all other workflows in the ensemble with lower
+priorities combined. The goal of the workflow ensemble scheduling and cloud
+provisioning problem is to complete as many high-priority workflows as possible
+given a fixed budget and deadline. Only workflows for which all tasks are
+finished by the deadline are considered to be complete---partial results are not
+usable in this model.
 
@@
 
-DPDS is an online algorithm that provisions resources and schedules tasks at runtime. It consists of two main parts: a provisioning procedure, and a scheduling procedure.
-
-DPDS' provisioning procedure is based on resource utilization.
-DPDS starts with a fixed number of resources calculated based on the available time and budget, and adjusts the number of resources according to how well they are utilized by the application.
-
-Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a VM in dollars $p$, it is easy to calculate the number of VMs, $N_{VM}$, to provision so that the entire budget is consumed before the deadline:
-%
+DPDS is an online algorithm that provisions resources and schedules tasks at
+runtime. It consists of two main parts: a provisioning procedure, and a
+scheduling procedure.
+
+DPDS' provisioning procedure is based on resource utilization. DPDS starts with
+a fixed number of resources calculated based on the available time and budget,
+and adjusts the number of resources according to how well they are utilized by
+the application.
+Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a
+VM in dollars $p$, it is easy to calculate the number of VMs, $N_{VM}$, to
+provision so that the entire budget is consumed before the deadline:
 \begin{equation}
@@
 
-As the application runs, the DPDS algorithm periodically computes resource utilization using the percentage of idle VMs over time, and adjusts the number of VMs if the utilization is above or below given thresholds.
-Because it is assumed that VMs are billed by the hour, DPDS only considers VMs that are approaching their hourly billing cycle when deciding which VMs to terminate.
+The DPDS algorithm periodically computes resource
+utilization using the percentage of idle VMs over time, and adjusts the number
+of VMs if the utilization is above or below given thresholds. Because it is
+assumed that VMs are billed by the hour, DPDS only considers VMs that are
+approaching their hourly billing cycle when deciding which VMs to terminate.
 This dynamic provisioning algorithm is shown in Algorithm~\ref{alg:prov}.
@@
 \label{alg:prov}
-{\small
+{\footnotesize
 \begin{algorithmic}[1]
@@
 
-The set of VMs completing their billing cycle is determined by considering both the provisioner interval, and the termination delay of the provider.
-This guarantees that VMs can be terminated before they start the next billing cycle and prevents the budget from being overrun.
-The VMs terminated in line \ref{l:terminate1} of Algorithm~\ref{alg:prov} are the ones that would overrun the budget if not terminated in the current provisioning cycle.
-The VMs terminated in line \ref{l:terminate2} are chosen to increase the resource utilization to the desired threshold.
-In order to prevent instances from being terminated too quickly, potentially wasting resources that have already been paid for but could be used later, no more than half of the idle resources are terminated during each provisioning cycle.
-To avoid an uncontrolled increase in the number of instances, which may happen in the case of highly parallel workflows, the provisioner will not start a new VM if the number of running VMs is greater than the product of $N_{VM}$ (from Equation~\ref{eq:static-plan}) and an autoscaling parameter, $v_{max}$.
-Unless otherwise specified, $v_{max}$ is assumed to be 1.
-
-In order to schedule individual workflow tasks onto available VMs, DPDS uses the dynamic, priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
+The set of VMs completing their billing cycle is determined by  both
+the provisioner interval, and the termination delay of the provider. This
+guarantees that VMs can be terminated before they start the next billing cycle
+and prevents the budget from being overrun. The VMs terminated in line
+\ref{l:terminate1} of Algorithm~\ref{alg:prov} are the ones that would overrun
+the budget if not terminated in the current provisioning cycle. The VMs
+terminated in line \ref{l:terminate2} are chosen to increase the resource
+utilization to the desired threshold. In order to prevent instances that have
+already been paid for from being terminated too quickly, no more than half of
+the idle resources are terminated during each provisioning cycle. To avoid an
+uncontrolled increase in the number of instances, which may happen in the case
+of highly parallel workflows, the provisioner will not start a new VM if the
+number of running VMs is greater than the product of $N_{VM}$ (from
+Equation~\ref{eq:static-plan}) and an autoscaling parameter, $v_{max}$. Unless
+otherwise specified, $v_{max}$ is assumed to be 1.
+
+In order to schedule individual workflow tasks onto available VMs, DPDS uses the
+dynamic, priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
+Initially, the ready tasks from all workflows in the ensemble are added to a
+priority queue based on the priority of the workflow to which they belong. If
+there are idle VMs available, and the priority queue is not empty, the next task
+from the priority queue is submitted to an arbitrarily chosen idle VM. The
+process is repeated until there are no idle VMs or the priority queue is empty.
+The scheduler then waits for a task to finish, adds its ready children to the
+priority queue, marks the VM as idle, and the entire process repeats until the
+deadline is reached.
 
@@
 \label{alg:ds}
-{\small
+{\footnotesize
 \begin{algorithmic}[1]
@@
 
-Initially, the ready tasks from all workflows in the ensemble are added to a priority queue based on the priority of the workflow to which they belong.
-If there are idle VMs available, and the priority queue is not empty, the next task from the priority queue is submitted to an arbitrarily chosen idle VM.
-The process is repeated until there are no idle VMs or the priority queue is empty.
-The scheduler then waits for a task to finish, adds its ready children to the priority queue, marks the VM as idle, and the entire process repeats until the deadline is reached.
-
-The DPDS algorithm guarantees that tasks from lower priority workflows are always deferred when higher-priority tasks are available, but lower-priority tasks can still occupy idle VMs when higher-priority tasks are not available.
-However, because there is no preemption, long-running low-priority tasks may delay the execution of higher-priority tasks.
-In addition, tasks from low priority workflows may be executed even though there is no chance that those workflows will be completed within the current budget and deadline.
-Figure~\ref{fig:spds-example} shows an example schedule generated using the SPDS algorithm.
-The figure illustrates how tasks from lower priority workflows backfill idle VMs when tasks from higher priority workflows are not available.
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=0.5\textwidth]{figures/spds-gantt}
-  \caption[Example schedule generated by the DPDS algorithm.]{Example schedule generated by the DPDS algorithm. Each row is a different VM. Tasks are represented as boxes and are colored by workflow.}
-  \label{fig:spds-example}
-\end{figure}
+The DPDS algorithm guarantees that tasks from lower priority workflows are
+always deferred when higher-priority tasks are available, but lower-priority
+tasks can still occupy idle VMs when higher-priority tasks are not available.
+However, because there is no preemption, long-running low-priority tasks may
+delay the execution of higher-priority tasks. In addition, tasks from low
+priority workflows may be executed even though there is no chance that those
+workflows will be completed within the current budget and deadline.
+Figure~\ref{fig:algorithm-example}a shows an example schedule generated using
+the DPDS algorithm. The figure illustrates how tasks from lower priority workflows
+backfill idle VMs when tasks from higher priority workflows are not available.
 
@@
 
-The DPDS algorithm does not use any information about the structure of the workflows in the ensemble when scheduling tasks: it looks only at the priorities of the ready tasks when deciding which task to schedule next.
-It does not consider whether a lower priority task belongs to a workflow that will never be able to complete given the current budget and deadline.
-As a result, DPDS may start lower priority tasks just to keep VMs busy that will end up delaying higher priority tasks later on, making it less likely that higher priority workflows will be able to finish.
-
-In order to address this issue, the Workflow-Aware DPDS (WA-DPDS) algorithm extends DPDS by introducing a workflow admission procedure.
-The admission procedure is invoked whenever WA-DPDS sees the first task of a new workflow at the head of the priority queue (i.e.
-when no other tasks from the workflow have been scheduled yet).
-The admission procedure---shown in Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough budget remaining to admit the new workflow; if there is not, then the workflow is rejected and its tasks are removed from the queue.
-This algorithm compares the current cost (consumed budget) and remaining budget, taking into account the cost of currently running VMs, and the cost of workflows that have already been admitted.
-In addition, it adds a small safety margin of \$0.10 to avoid going over the budget.
+The DPDS algorithm does not use any information about the structure of the
+workflows in the ensemble when scheduling tasks: it looks only at the priorities
+of the ready tasks when deciding which task to schedule next. It does not
+consider whether a lower priority task belongs to a workflow that will never be
+able to complete given the current budget and deadline. As a result, DPDS may
+start lower priority tasks just to keep VMs busy that will end up delaying
+higher priority tasks later on, making it less likely that higher priority
+workflows will be able to finish.
+
+In order to address this issue, the Workflow-Aware DPDS (WA-DPDS) algorithm
+extends DPDS by introducing a workflow admission procedure. The admission
+procedure is invoked whenever WA-DPDS sees the first task of a new workflow at
+the head of the priority queue (i.e. when no other tasks from the workflow have
+been scheduled yet). The admission procedure---shown in
+Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough budget remaining
+to admit the new workflow; if there is not, then the workflow is rejected and
+its tasks are removed from the queue. This algorithm compares the current cost
+(consumed budget) and remaining budget, taking into account the cost of
+currently running VMs, and the cost of workflows that have already been
+admitted. In addition, it adds a small safety margin of \$0.10 to avoid going
+over the budget.
 
@@
 \label{alg:wa-dpds}
-{\small
+{\footnotesize
 \begin{algorithmic}[1]
@@
   \State $c_w\gets$ \Call{EstimateCost}{w}
-    \If{$c_w<r_b$}
-      \State \textbf{return} $TRUE$
-    \Else
-      \State \textbf{return} $FALSE$
+    \If{$c_w<r_b$}  \textbf{return} $TRUE$
+    \Else ~  \textbf{return} $FALSE$
   \EndIf       
@@
 
-This admission procedure relies only on the total estimated resource consumption and compares it to the remaining budget.
-We found that this estimation is useful not only to prevent low-priority workflows from delaying high-priority ones, but also to reject large and costly workflows that would overrun the budget and admit smaller workflows that can efficiently utilize idle resources in ensembles containing workflows of non-uniform sizes.
-It would also be possible to extend this admission procedure to check other constraints, such as whether the estimated critical path of the new workflow exceeds the time remaining until the deadline.
+This admission procedure relies only on the total estimated resource consumption
+and compares it to the remaining budget. We found that this estimation is useful
+not only to prevent low-priority workflows from delaying high-priority ones, but
+also to reject large and costly workflows that would overrun the budget and
+admit smaller workflows that can efficiently utilize idle resources.
+% in
+%ensembles containing workflows of non-uniform sizes. It would also be possible
+%to extend this admission procedure to check other constraints, such as whether
+%the estimated critical path of the new workflow exceeds the time remaining
+%until the deadline.
 
@@
 
-The previous algorithms are all online algorithms that make provisioning and scheduling decisions at runtime.
-In comparison, the SPSS algorithm creates a provisioning and scheduling plan before running any workflow tasks.
-This enables SPSS to start only those workflows that it knows can be completed given the deadline and budget constraints, and eliminates any waste that may be allowed by the dynamic algorithms.
-
-The approach used by SPSS is to plan each workflow in the ensemble in priority order, rejecting any workflows that cannot be completed by the deadline or that cause the cost of the plan to exceed the budget.
-Once the plan is complete, the VMs are provisioned and tasks are executed according to the schedules given by the plan.
-
-The disadvantage of the static planning approach used by SPSS is that it is sensitive to dynamic changes in the environment and the application that may disrupt its carefully constructed plan.
-For example, if there are provisioning delays, or if the runtime estimates for the tasks are inaccurate, then workflow execution may diverge from the plan.
-This issue will be discussed further in Sections~\ref{sec:delays} and \ref{sec:variances}.
-
-Algorithm~\ref{alg:admit} shows how ensembles are planned in SPSS.
-The algorithm starts with an empty plan containing no VMs and no scheduled tasks.
-Workflows from the ensemble are considered in priority order.
-For each workflow, SPSS attempts to build on top of the current plan by provisioning VMs to schedule the tasks of the workflow so that it finishes before the deadline with the least possible cost.
-If the cost of the new plan is less than the budget, then the new plan is accepted and the workflow is admitted.
-If not, then the new plan is rejected and the process continues with the next workflow in the ensemble.
-The idea behind this algorithm is that, if each workflow can be completed by the deadline with the lowest possible cost, then the number of workflows that can be completed within the given budget will be maximized.
+The previous dynamic (online) algorithms make provisioning and scheduling
+decisions at runtime. In comparison, the SPSS algorithm creates a provisioning
+and scheduling plan before running any workflow tasks. This enables SPSS to
+start only those workflows that it knows can be completed given the deadline and
+budget constraints, and eliminates any waste that may be allowed by the dynamic
+algorithms. However, the disadvantage of SPSS is that it is sensitive to dynamic
+changes in the environment (see Sections~\ref{sec:delays} and
+\ref{sec:variances}).
+
+The approach used by SPSS is to plan each workflow in the ensemble in priority
+order, rejecting any workflows that would exceed the dedline or budget. Once the
+plan is complete, the VMs are provisioned and tasks are executed according to
+the schedules given by the plan.
+
+%The disadvantage of the static planning approach used by SPSS is that it is
+%sensitive to dynamic changes in the environment and the application that may
+%disrupt its carefully constructed plan. For example, if there are provisioning
+%delays, or if the runtime estimates for the tasks are inaccurate, then workflow
+%execution may diverge from the plan. This issue will be discussed further in
+%Sections~\ref{sec:delays} and \ref{sec:variances}.
+
+Algorithm~\ref{alg:admit} shows how ensembles are planned in SPSS. Workflows
+from the ensemble are considered in priority order. For each workflow, SPSS
+attempts to build on top of the current plan by provisioning VMs to schedule the
+tasks of the workflow so that it finishes before the deadline with the least
+possible cost. If the cost of the new plan is less than the budget, then the new
+plan is accepted and the workflow is admitted. If not, then the new plan is
+rejected and the process continues with the next workflow in the ensemble. The
+idea is that, if each workflow can be completed by the
+deadline with the lowest possible cost, then the number of workflows that can be
+completed within the given budget will be maximized.
+
+To plan a workflow, the SPSS algorithm assigns sub-deadlines to each individual
+task in the workflow, and then schedules each task so as to minimize the cost of
+the task while still meeting its assigned sub-deadline. If each task can be
+completed by its deadline in the least expensive way, then the cost of the
+entire workflow can be minimized without exceeding the deadline. SPSS assigns
+sub-deadlines to each task based on the slack time of the workflow, which is
+defined as the amount of extra time that a workflow can extend its critical path
+and still be completed by the ensemble deadline. For a workflow $w$, the slack
+time of $w$ is:
+$$ ST(w) = d - CP(w) $$
+where $d$ is the deadline and $CP(w)$ is the critical path of $w$. We
+assume that $CP(w) \leq d$, otherwise the workflow cannot be completed by the
+deadline and must be rejected. 
+%For large ensembles we expect the critical path
+%of any individual workflow to be much less than the deadline.
 
@@
 \label{alg:admit}
-{\small
+{\footnotesize
 \begin{algorithmic}[1]
@@
 
-In order to plan a workflow, the SPSS algorithm assigns sub-deadlines to each individual task in the workflow, and then schedules each task so as to minimize the cost of the task while still meeting its assigned sub-deadline.
-The idea is that if each task can be completed by its deadline in the least expensive manner possible, then the cost of the entire workflow can be minimized without exceeding the ensemble deadline.
+A task's level is the length of the longest path between the task and an entry task of the workflow:
+%
+$$
+Level(t) = \begin{cases}
+0,&$if~$Pred(t) = \emptyset\cr
+\max_{p \in Pred(t)}Level(p)+1,&$otherwise.$
+\end{cases}
+$$
+%
+SPSS distributes the slack time of the workflow by level, so that each level of the workflow gets a portion of the workflow's slack time proportional to the number of tasks in the level and the total runtime of tasks in the level.
+The idea is that levels containing many tasks and large runtimes should be given a larger portion of the slack time so that tasks in those levels may be serialized.
+Otherwise, many resources need to be allocated to run all of the tasks in parallel, which may be more costly.
+
+The slack time of a level $l$ in workflow $w$ is given by:
+%
+$$
+ST(l) = ST(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
+$$
+%
+where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between 0 and 1 that causes more slack time to be given to levels with more tasks (large $\alpha$) or more runtime (small $\alpha$).
+
+The deadline of a task $t$ is then:
+%
+$$
+DL(t) = LST(t) + RT(t) + ST(Level(t))
+$$
+%
+where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and $LST(t)$ is the latest start time of $t$ determined by:
+%
+$$
+LST(t) = \begin{cases}
+0,&$if~$Pred(t) = \emptyset\cr
+\max_{p \in Pred(t)} DL(p),&$otherwise.$
+\end{cases}
+$$
 
@@
 \label{alg:planworkflow}
-{\small
+{\footnotesize
 \begin{algorithmic}[1]
@@
 
-SPSS assigns sub-deadlines to each task based on the slack time of the workflow, which is defined as the amount of extra time that a workflow can extend its critical path and still be completed by the ensemble deadline.
-For a workflow $w$, the slack time of $w$ is:
-%
-$$
-ST(w) = d - CP(w)
-$$
-%
-where $d$ is the ensemble deadline and $CP(w)$ is the critical path of $w$.
-We assume that $CP(w) \leq d$, otherwise the workflow cannot be completed by the deadline and must be rejected.
-For large ensembles we expect the critical path of any individual workflow to be much less than the deadline.
-
-A task's level is the length of the longest path between the task and an entry task of the workflow:
-%
-$$
-Level(t) = \begin{cases}
-0,&$if~$Pred(t) = \emptyset\cr
-\max_{p \in Pred(t)}Level(p)+1,&$otherwise.$
-\end{cases}
-$$
-%
-SPSS distributes the slack time of the workflow by level, so that each level of the workflow gets a portion of the workflow's slack time proportional to the number of tasks in the level and the total runtime of tasks in the level.
-The idea is that levels containing many tasks and large runtimes should be given a larger portion of the slack time so that tasks in those levels may be serialized.
-Otherwise, many resources need to be allocated to run all of the tasks in parallel, which may be more costly.
-
-The slack time of a level $l$ in workflow $w$ is given by:
-%
-$$
-ST(l) = ST(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
-$$
-%
-where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between 0 and 1 that causes more slack time to be given to levels with more tasks (large $\alpha$) or more runtime (small $\alpha$).
-
-The deadline of a task $t$ is then:
-%
-$$
-DL(t) = LST(t) + RT(t) + ST(Level(t))
-$$
-%
-where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and $LST(t)$ is the latest start time of $t$ determined by:
-%
-$$
-LST(t) = \begin{cases}
-0,&$if~$Pred(t) = \emptyset\cr
-\max_{p \in Pred(t)} DL(p),&$otherwise.$
-\end{cases}
-$$
-
-Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for each workflow.
-The \textsc{PlanWorkflow} procedure first calls \textsc{DeadlineDistribution} to assign sub-deadlines to tasks according to the procedure described above.
-Then, the \textsc{PlanWorkflow} procedure schedules tasks onto VMs, allocating new VMs when necessary.
-For each task in the workflow, the least expensive slot is chosen to schedule the task so that it can be completed by  its deadline.
-VMs are allocated in blocks of one billing cycle (one hour) regardless of the size of the task.
-When computing the cost of scheduling a task on a given VM, the algorithm considers idle slots in blocks that were allocated for previous tasks to be free, while slots in new blocks cost the full price of a billing cycle.
-For example, if a task has a runtime of 10 minutes, and the price of a block is \$1, then the algorithm will either schedule the task on an existing VM that has an idle slot larger than 10 minutes for a cost of \$0, or it will allocate a new block on an existing VM, or provision a new VM, for a cost of \$1.
-If the cost of slots on two different VMs is equal, then the slot with the earliest start time is chosen.
-To prevent too many VMs from being provisioned, the algorithm always prefers to extend the runtime of existing VMs before allocating new VMs.
-The result of this is that the algorithm will only allocate a new block if there are no idle slots on existing blocks large enough or early enough to complete the task by its deadline, and it will only allocate a new VM if it cannot add a block to an existing VM to complete the task by its deadline.
-
-An example schedule generated by SPSS is shown in Figure.~\ref{fig:spss-example}.
-This example shows how SPSS tends to start many workflows in parallel, running each workflow over a longer period of time on only a few VMs to minimize cost.
-In comparison, the dynamic algorithms tend to run one workflow at a time across many VMs in parallel.
+Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for each
+workflow. The \textsc{PlanWorkflow} procedure first calls
+\textsc{DeadlineDistribution} to assign sub-deadlines to tasks.
+% according to the procedure described above. 
+Then, \textsc{PlanWorkflow} schedules
+tasks onto VMs, allocating new VMs when necessary. For each task in the
+workflow, the least expensive slot is chosen to schedule the task so that it can
+be completed by  its deadline. VMs are allocated in blocks of one billing cycle
+(one hour) regardless of the size of the task. When computing the cost of
+scheduling a task on a given VM, the algorithm considers idle slots in blocks
+that were allocated for previous tasks to be free, while slots in new blocks
+cost the full price of a billing cycle. 
+%For example, if a task has a runtime of
+%10 minutes, and the price of a block is \$1, then the algorithm will either
+%schedule the task on an existing VM that has an idle slot larger than 10
+%minutes for a cost of \$0, or it will allocate a new block on an existing VM,
+%or provision a new VM, for a cost of \$1. 
+If the cost of slots on two different
+VMs is equal, then the slot with the earliest start time is chosen. To prevent too
+many VMs from being provisioned, the algorithm always prefers to extend the
+runtime of existing VMs before allocating new VMs. 
+%The result of this is that
+%the algorithm will only allocate a new block if there are no idle slots on
+%existing blocks large enough or early enough to complete the task by its
+%deadline, and it will only allocate a new VM if it cannot add a block to an
+%existing VM to complete the task by its deadline.
 
@@
   \centering
-  \includegraphics[width=0.5\textwidth]{figures/spss-gantt}
-  \caption[Example schedule generated by the SPSS algorithm]{Example schedule generated by the SPSS algorithm. Each row is a different VM. Tasks are represented as boxes and are colored by workflow.}
-  \label{fig:spss-example}
+  \subfloat[DPDS]{
+  \includegraphics[width=0.45\textwidth]{figures/spds-gantt}
+  }\\
+  \subfloat[SPSS]{
+  \includegraphics[width=0.45\textwidth]{figures/spss-gantt}
+  }
+  \caption[Example schedules generated by the algorithms]{Example
+  schedules generated by the algorithms. Each row is a different
+  VM. Tasks are represented as boxes and are colored by workflow.}
+  \label{fig:algorithm-example}
 \end{figure}
+
+An example schedule generated by SPSS (Figure~\ref{fig:algorithm-example}b)
+shows how SPSS tends to start many workflows in parallel, running each workflow
+over a longer period of time on only a few VMs to minimize cost. In comparison,
+the dynamic algorithms tend to run one workflow at a time across many VMs in
+parallel.
 
@@
 
-To evaluate and compare the three proposed algorithms, we developed a cloud workflow simulator based on CloudSim~\cite{Calheiros2011}.
-Our simulation model consists of Cloud, VM and WorkflowEngine entities.
-The Cloud entity is responsible for staring and terminating VM entities using an API similar to that used by Amazon EC2.
-VM entities simulate the execution of individual tasks, including randomized variations in runtime.
-And the WorkflowEngine entity manages the scheduling of tasks and the provisioning of VMs according to the rules established by the different algorithms.
-We assume that the VMs have a single core and execute tasks sequentially.
-Although CloudSim provides a more advanced infrastructure model which includes time-sharing and space-sharing policies, we do not use these features since we are interested mainly in the execution of tasks on VMs and high-level workflow scheduling and provisioning.
-The simulator reads workflow description files in a modified version of the DAX format used by the Pegasus Workflow Management System~\cite{Deelman2005}.
-The modified file format was created for the synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.
+To evaluate and compare the three proposed algorithms, we developed a cloud
+workflow simulator based on CloudSim~\cite{Calheiros2011}. Our simulation model
+consists of Cloud, VM and WorkflowEngine entities. The Cloud entity starts and
+terminates VM entities using an Amazon EC2-like API. VM entities simulate the
+execution of individual tasks, including randomized variations in runtime. The
+WorkflowEngine entity manages the scheduling of tasks and the provisioning of 
+VMs based on the chosen algorithm. We assume that the VMs have a single core
+and execute tasks sequentially. 
+%Although CloudSim provides a more advanced infrastructure model which includes
+%time-sharing and space-sharing policies, we do not use these features since we
+%are interested mainly in the execution of tasks on VMs and high-level workflow
+%scheduling and provisioning. 
+The simulator reads workflow description files in
+a modified version of the DAX format used by the Pegasus Workflow Management
+System~\cite{Deelman2005}. 
+%The modified file format was created for the
+%synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.
 
@@
 
-In order to evaluate the algorithms on a standard set of workflows, we created randomized ensembles using workflows available from the workflow generator gallery published online by Bharathi, et al~\cite{WorkflowGenerator}.
-The gallery contains synthetic workflows modeled using structures and parameters that were taken from real applications.
-Ensembles were created using synthetic workflows from five real applications: SIPHT, a bioinformatics application, LIGO, a physics application, Epigenomics, another bioinformatics application, Montage, an astronomy application, and CyberShake, an earthquake science application.
-For each of these applications, workflows with 50, 100, 200, 300, 400, 500, 600, 700, 800, 900 and 1000 tasks were generated.
-For each workflow size, 20 different workflow instances were generated using parameters and task runtime distributions from traces of real workflows.
-The total collection of synthetic workflows contains 5 applications, 11 different workflow sizes, and 20 workflow instances, for a total of 1100 synthetic workflows.
+In order to evaluate the algorithms on a standard set of workflows, we created
+randomized ensembles using workflows available from the workflow generator
+gallery published online by Bharathi, et al~\cite{WorkflowGenerator}. The
+gallery contains synthetic workflows modeled using structures and parameters
+that were taken from real applications. Ensembles were created using synthetic
+workflows from five real applications: SIPHT, LIGO, Epigenomics, Montage and
+CyberShake. For each of these applications, workflows with 50, 100, 200, 300,
+400, 500, 600, 700, 800, 900 and 1000 tasks were generated. For each workflow
+size, 20 different workflow instances were generated using parameters and task
+runtime distributions from traces of real workflows. The total collection of
+synthetic workflows contains 5 applications, 11 different workflow sizes, and 20
+workflow instances, for a total of 1100 synthetic workflows.
 
 Using this collection of workflows, we constructed five different ensemble types:
-
-\begin{itemize}
-  \item constant
-  \item uniform, sorted
-  \item uniform, unsorted
-  \item Pareto, sorted
-  \item Pareto, unsorted
-\end{itemize}
-
-In the unsorted ensembles, workflows of different sizes are mixed together, and the priorities are assigned randomly so that there is no relationship between priority and workflow size.
-For many applications, however, large workflows are more important to users than small workflows because the represent more significant computations.
-To model this, the sorted ensembles are sorted by size, so that the largest workflows are given the highest priority.
+%
+%\begin{itemize}
+constant, uniform sorted, uniform unsorted, Pareto sorted and Pareto unsorted.
+%\end{itemize}
+In the unsorted ensembles, workflows of different sizes are mixed together, and
+the priorities are assigned randomly.
+% so that there is no relationship between
+%priority and workflow size. 
+For many applications, however, large workflows are
+more important to users than small workflows because they represent more
+significant computations. To model this, the sorted ensembles are sorted by
+size, so that the largest workflows have the highest priority.
 
 Constant ensembles contain workflows that all have the same number of tasks. 
-For each constant ensemble, the number of tasks is chosen randomly from the set of possible workflow sizes.
-Once the size is determined, then N workflows of that size are chosen randomly for the ensemble from the set of synthetic workflows for the given application.
-
-Uniform ensembles contain workflows with sizes that are uniformly distributed among the set of possible sizes.
-Each workflow in a uniform ensemble is selected by first randomly choosing the size the workflow according to a uniform distribution, and then randomly choosing a workflow of that size from the set of synthetic workflows for the given application.
-
-Pareto ensembles contain a small number of larger workflows and a large number of smaller workflows.
-The sizes of the workflows in a Pareto ensemble are chosen according to a Pareto distribution.
-The distribution was modified so that the number of large workflows (of size $\geq$ 900) is increased by a small amount to produce a ``heavy-tail''.
-This causes Pareto ensembles to have a slightly larger number of large workflows, which reflects behavior commonly observed in many computational workloads.
-An example of the distribution of workflow sizes that occurs in a Pareto ensemble is shown in Figure~\ref{fig:ensemble-pareto}.
-
-\begin{figure}[ht] 
-    \centering
-    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
-    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of workflow sizes in Pareto ensembles. Workflow size is measured in number of tasks.}
-    \label{fig:ensemble-pareto}
-\end{figure}
-
-The number of workflows in an ensemble depends on the particular application, but we assume that ensemble sizes are on the order of between 10 and 100 workflows.
-This range is motivated by several reasons: First, such sizes are typical of the real applications we have examined.
-For example, the number of geographical sites of interest to the users of the CyberShake application in the past has been on the order of 100.
-Second, smaller ensembles consisting of just a few workflows can be aggregated into a single workflow, so there is no need to treat them as an ensemble.
-Similarly, when the number of workflows grows, and each workflow has a large number of tasks, either the deadline and budget constraints are low enough to prevent many of the workflows from running, or the problem of efficiently allocating them to the resources becomes similar to a bag-of-tasks problem, which is easier to solve efficiently.
+%For each constant ensemble, 
+The number of tasks is chosen randomly from the set of
+possible workflow sizes. Once the size is determined, then N workflows of that
+size are chosen randomly for the ensemble from the set of synthetic workflows.
+%for the given application.
+
+Uniform ensembles contain workflows with sizes that are uniformly distributed
+among the set of possible sizes. Each workflow 
+%in a uniform ensemble 
+is selected by first randomly choosing the size the workflow
+% according to a uniform distribution, 
+and then randomly choosing a workflow of that size from the set of
+synthetic workflows.
+% for the given application.
+
+Pareto ensembles contain a small number of larger workflows and a large number
+of smaller workflows. Their sizes  
+%in a Pareto ensemble 
+are chosen according to a Pareto distribution. The distribution was modified so
+that the number of large workflows (of size $\geq$ 900) is increased by a small amount to
+produce a ``heavy-tail''. This causes Pareto ensembles to have a slightly larger
+number of large workflows, which reflects behavior commonly observed in many
+computational workloads. 
+%An example of the distribution of workflow sizes that
+%occurs in a Pareto ensemble is shown in Figure~\ref{fig:ensemble-pareto}.
+
+%\begin{figure}[ht] 
+%    \centering
+%    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
+%    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of
+%    workflow sizes in Pareto ensembles. Workflow size is measured in number of
+%    tasks.}
+%    \label{fig:ensemble-pareto}
+%\end{figure}
+
+The number of workflows in an ensemble depends on the particular application,
+but we assume that ensemble sizes are on the order of between 10 and 100
+workflows. This range is motivated by several reasons: First, such sizes are
+typical of the real applications we have examined (see Section~\ref{sec:intro}). 
+%For example, the number of
+%geographical sites of interest to the users of the CyberShake application in
+%the past has been on the order of 100. 
+Second, smaller ensembles consisting of
+just a few workflows can be aggregated into a single workflow, so there is no need to
+treat them as an ensemble. Similarly, when the number of workflows grows, and
+each workflow has a large number of tasks, either the deadline and budget
+constraints are low enough to prevent many of the workflows from running, or the
+problem of efficiently allocating them to the resources becomes similar to a
+bag-of-tasks problem, which is easier to solve efficiently.
 
@@
 
-In order to observe the interesting characteristics of the proposed algorithms, we evaluated them using a broad range of constraint values.
-For each ensemble, we selected ranges for deadline and budget that cover a broad parameter space: from tight constraints, where only a small number of workflows can be completed, to more liberal constraints where all, or almost all, of the workflows can be completed.
-Because the interesting parameter space varies among the applications, ensemble types and sizes, we computed constraint ranges based on the characteristics of each ensemble.
-The budget constraints are calculated by identifying the smallest budget required to execute one of the workflows in the ensemble (MinBudget), and the smallest budget required to execute all workflows in the ensemble (MaxBudget):
+In order to observe the interesting characteristics of the proposed algorithms,
+for each ensemble, we selected ranges for deadline and budget that cover a broad
+parameter space: from tight constraints, where only a small number of workflows can be completed,
+to more liberal constraints where all, or almost all, of the workflows can be
+completed. We computed constraint ranges based on
+the characteristics of each ensemble. The budget constraints are calculated by
+identifying the smallest budget required to execute one of the workflows in the
+ensemble (MinBudget), and the smallest budget required to execute all workflows
+in the ensemble (MaxBudget):
+%
 %
@@
 %
-This range---$[MinBudget, MaxBudget]$---is then divided into equal intervals to determine the budgets to use in each experiment.
-Similarly, the deadline constraints for each ensemble are calculated by identifying the smallest amount of time required to execute a single workflow in the ensemble (MinDeadline), which is the length of the critical path for the workflow with the shortest critical path, and by identifying the smallest amount of time required to execute all workflows (MaxDeadline), which is the sum of the critical paths of all the workflows:
+This range---$[MinBudget, MaxBudget]$---is then divided into equal intervals to
+determine the budgets to use in each experiment. Similarly, the deadline
+constraints are calculated by identifying the smallest amount
+of time required to execute a single workflow in the ensemble (MinDeadline),
+which is the length of the critical path for the workflow with the shortest
+critical path, and by identifying the smallest amount of time required to
+execute all workflows (MaxDeadline), which is the sum of the critical paths of
+all the workflows:
+%
 %
@@
 %
-This range---$[MinDeadline, MaxDeadline]$---is then divided into equal intervals to determine the deadlines to use in each experiment.
-By computing the budget and deadline constraits in this way we ensure that the experiments for each ensemble cover the most interesting area of the parameter space for the ensemble.
+This range---$[MinDeadline, MaxDeadline]$---is then divided into equal intervals. 
+By computing the budget and deadline constraits in this way we ensure that the
+experiments for each ensemble cover the most interesting area of the parameter space for the
+ensemble.
 
 In all the experiments we assumed that the VMs have a price of \$1 per VM-hour.
-This price was chosen to simplify interpretation of results and should not affect the relative performance of the different algorithms.
-In this study we do not take into account the heterogeneity of the infrastructure since we assume that it is always possible to select a VM type that has the best price to performance ratio for a given application.
-
-All the experiments were run with maximum autoscaling factor ($v_{max}$) set to 1.0 for DPDS and WA-DPDS.
-After experimenting with DPDS and WA-DPDS we found that, due to the high parallelism of workflows used, the resource utilization remains high enough without adjusting the autoscaling rate.
-Based on experiments with the target applications, we set the SPSS $\alpha$ parameter for deadline distribution to be 0.7, which allocates slightly more time to levels with many tasks.
+This price was chosen to simplify interpretation of results and should not
+affect the relative performance of the different algorithms. In this study the
+heterogeneity of the infrastructure is not relevant since we assume that it is
+always possible to select a VM type that has the best price to performance ratio
+for a given application~\cite{Juve2009}.
+
+All the experiments were run with maximum autoscaling factor ($v_{max}$) set to
+1.0 for DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found
+that, due to the high parallelism of workflows used, the resource utilization
+remains high enough without adjusting the autoscaling rate. Based on experiments
+with the target applications, we set the SPSS $\alpha$ parameter for deadline
+distribution to be 0.7, which allocates slightly more time to levels with many
+tasks.
 
@@
 
-In order to assess the relative performance of our algorithms it is necessary to define a metric that can be used to score the performance of the different algorithms on a given problem (ensemble, budget, and deadline).
-The simplest approach is to simply count the number of workflows in the ensemble that each algorithm is able to complete within the budget before the deadline.
-The problem with this approach is that it does not account for the priority-based utility function specified by the user.
-Using the counting approach, a less efficient algorithm may be able to complete a large number of low-priority workflows by executing the smallest workflows first.
-
-In order to account for the user's priority, we used an exponential scoring approach where the score for an ensemble $e$ is:
+In order to quantitively compare our algorithms it is necessary to define a
+metric that can be used to score the performance of the different algorithms on
+a given problem (ensemble, budget, and deadline). The simplest approach is to
+count the number of workflows in the ensemble that each algorithm is able to
+complete within the budget before the deadline, but this metric does not account
+for the priority-based utility function specified by the user. Using the
+counting approach, a less efficient algorithm may be able to complete a large
+number of low-priority workflows by executing the smallest workflows first.
+
+In order to account for the user's priority, we used an exponential scoring where the score for an ensemble $e$ is:
 %

### 2012-04-25 paper/paper.tex
fix boxplot figures

@@
     \subfloat[Ratio of Actual Cost to Budget]{
-        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-cost-montage-uniform_unsorted}
+        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-cost-montage}
         }
@@
         \hspace{2cm}
-        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-dagfinish-montage-uniform_unsorted}
+        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
         }

### 2012-04-25 scripts/R/data_cube_csv.R
more fancy plotting scripts

@@ -136,9 +136,9 @@ sumCutScores = apply(cutScores,c(1,5,6),sum)
 
 
 
-pdf(file=paste(csv_filename, "-distributions.pdf", sep=""), height=4, width=6, bg="white")
-par(mfrow=c(2,3))
-
+pdf(file=paste(csv_filename, "-distributions.pdf", sep=""), height=2, width=12, bg="white")
+par(mfrow=c(1,6))
+par(mar=c(4,4,2,0))
 for (i in 1:n_applications) {
 	print(applications[i])
 	bar_data = sumScores[,,i]
@@ -148,12 +148,14 @@ for (i in 1:n_applications) {
 	# reorder algorithms and change to %
 	bar_data = bar_data[c(1,3,2),]/10.0
 	print(bar_data)	
-	barplot(as.table(bar_data), beside = TRUE, xlab="Distribution", ylab = "Best Scores (%)", main = applications[i], col=gray( c(0.1,0.4,0.8) ),
-	ylim = c(0,100), cex.names=0.8)
+	barplot(as.table(bar_data), beside = TRUE, xlab="Distribution",  main = applications[i], 
+			col=gray( c(0.1,0.4,0.8) ),
+	ylim = c(0,100), cex.names=1.0)
+	if (i==1) {title(ylab="Best Scores (%)")}
 }
-plot.new() 
+plot.new()
 plot.window(c(0,1), c(0,1))
-legend("center", algorithms[c(1,3,2)], fill=gray( c(0.1,0.4,0.8) ))
+legend("left", algorithms[c(1,3,2)], fill=gray( c(0.1,0.4,0.8) ))
 dev.off()
 
 

### 2012-04-25 scripts/R/stats_delays_csv.R

@@ -1,0 +1,225 @@
+
+
+
+
+############################################################################################################################
+# Start
+############################################################################################################################
+
+
+
+csv_filename = "run-finish-delays-test-0-output"
+
+data = read.csv(paste(csv_filename,".dat", sep = ""))
+
+# reorder algorithms
+data$algorithm = factor(data$algorithm,c("DPDS","WADPDS","SPSS"))
+
+delays = unique(data$delay)
+print(delays)
+
+applications = levels(unique(data$application))
+print(applications)
+distributions = levels(unique(data$distribution))
+print(distributions)
+#algorithms = levels(unique(data$algorithm))
+algorithms = c("DPDS","WADPDS","SPSS")
+print(algorithms)
+
+pdf(file=paste(csv_filename, "-jobfinish.pdf", sep=""), height=4, width=12, bg="white")
+par(mfrow=c(1,7))
+for (delay in delays) {
+	print(delays)
+	boxplot(jobfinish/deadline~algorithm, las = 2, data = data[data$delay==delay,],main = sprintf("%2.0f",delay),cex.names=0.8,ylim = c(0,2),ylab = "jobfinish / deadline")
+	abline(h = 1, col = "red")
+}
+dev.off()
+
+pdf(file=paste(csv_filename, "-dagfinish.pdf", sep=""), height=4, width=12, bg="white")
+par(mfrow=c(1,7))
+for (delay in delays) {
+	print(delays)
+	boxplot(dagfinish/deadline~algorithm, las = 2, data = data[data$delay==delay,],main = sprintf("%2.0f",delay),cex.names=0.8,ylim = c(0,2),ylab = "dagfinish / deadline")
+	abline(h = 1, col = "red")
+}
+dev.off()
+
+pdf(file=paste(csv_filename, "-vmfinish.pdf", sep=""), height=4, width=12, bg="white")
+par(mfrow=c(1,7))
+for (delay in delays) {
+	print(delays)
[…]

### 2012-04-25 scripts/R/stats_sqlite.R

@@ -71,6 +71,56 @@ plot_series <- function(prefix, plot_title, application, budgets, algorithms, li
 	dev.off()
 }
 
+# Plot selected column from the series
+plot_selected <- function(prefix, plot_title, application, budget_id, algorithms, list_avg, hours, legend_position) {
+	pdf(file=paste(prefix, application, ".pdf", sep=""), height=4, width=5, bg="white")
+#	par(mfrow=c(1,5))	
+	par(mar=c(4,4,0.5,0.1))
+	#pdf(file=paste(prefix, dag, "h", deadline, "m", max_scaling, ".pdf", sep=""), height=3, width=10, bg="white")
+	#par(mfrow=c(2,1))
+	x_max = hours[length(hours)]
+	# create vectors for max and min values
+	list_max = 1:n_algorithms
+	list_min = 1:n_algorithms
+	
+	i = budget_id	
+	
+	for(alg in 1:length(algorithms)) {
+		list_max[alg]=max(list_avg[[alg]][i,])
+		list_min[alg]=min(list_avg[[alg]][i,])
+	}
+	y_max = max(list_max)
+	y_min = min(list_min)
+	# print(y_max)
+	# print(y_min)
+	
+	plot(hours, list_avg[[1]][1,]*1.2, type="n", col="blue", axes=FALSE, ann=FALSE, xlim=c(0,x_max*1.2), ylim=c(y_min,y_max))
+	for(alg in 1:length(algorithms)){	
+		lines(hours, list_avg[[alg]][i,], type="o", pch=alg, lty=5, col=alg)
+	}
+	
+	#title(main=paste("budget = $",budgets[i]), col.main="black", font.main=4)
+	legend(legend_position, c("SPSS","WADPDS","DPDS"), cex=1, col=c(2,1,3), pch=c(2,1,3) )
+	
+	
+	#legend(x_range*1.0, avg_m[i,length(hours)]+5, paste("$",budgets[i]), cex=1.2, bty="n");
+	
+	title(ylab=plot_title, cex.lab=1.2)
+	title(xlab="deadline in hours", cex.lab=1.2)
+	axis(2,cex.axis=1.2)
+	axis(1,cex.axis=1.2)
+	grid(col = "gray", lty = "dashed")
+	box()
+	
+	
+	
+	#legend("topright", c(algorithms$algorithm[1:6]), cex=1, col=1:6, pch=1:6 )
+#	mtext(application, side=3, outer=TRUE, line=-1.5) 
+	dev.off()
+}
+
+
+
 
 
[…]

### 2012-04-25 scripts/R/stats_variations_csv.R

@@ -12,52 +12,58 @@
 #csv_filename = "run-failures-test-0-output.dat"
 #csv_filename = "run-scaling-failures-test-0-output.dat"
 #csv_filename = "run-scaling-1-failures-test-0-output.dat"
-csv_filename = "run-finish-failures-test-0-output"
-#csv_filename = "run-finish-variations-test-0-output.dat"
+#csv_filename = "run-finish-failures-test-0-output.dat"
+csv_filename = "run-finish-variations-test-0-output"
 
 data = read.csv(paste(csv_filename,".dat", sep = ""))
-
 
 # reorder algorithms
 data$algorithm = factor(data$algorithm,c("DPDS","WADPDS","SPSS"))
 algorithms = c("DPDS","WADPDS","SPSS")
 print(algorithms)
 
-failureRates = unique(data$failureRate)
-print(failureRates)
+
+runtimeVariances = unique(data$runtimeVariance)
+print(runtimeVariances)
+
+applications = levels(unique(data$application))
+print(applications)
+distributions = levels(unique(data$distribution))
+print(distributions)
+
 
 pdf(file=paste(csv_filename, "-jobfinish.pdf", sep=""), height=4, width=12, bg="white")
 par(mfrow=c(1,7))
-for (failureRate in failureRates) {
-	print(failureRates)
-	boxplot(jobfinish/deadline~algorithm, las = 2, data = data[data$failureRate==failureRate,],main = sprintf("%2.0f %%",failureRate*100),cex.names=0.8,ylim = c(0,2),ylab = "jobfinish / deadline")
+for (runtimeVariance in runtimeVariances) {
+	print(runtimeVariances)
+	boxplot(jobfinish/deadline~algorithm, las = 2, data = data[data$runtimeVariance==runtimeVariance,],main = sprintf("%2.0f %%",runtimeVariance*100),cex.names=0.8,ylim = c(0,2),ylab = "jobfinish / deadline")
 	abline(h = 1, col = "red")
 }
 dev.off()
 
 pdf(file=paste(csv_filename, "-dagfinish.pdf", sep=""), height=4, width=12, bg="white")
 par(mfrow=c(1,7))
-for (failureRate in failureRates) {
-	print(failureRates)
[…]

### 2012-04-26 paper/paper.tex
Resize figures
abstract

@@
 simulation using a set of scientific workflow ensembles with a broad range of
-budget and deadline parameters, identify which algorithm performs best for 
-different application scenarios and constraint ranges. We find that the key 
+budget and deadline parameters, taking into account uncertainties in task runtime
+estimations, provisioning delays and failures. We find that the key 
 factor determining the performance of an algorithm is its ability to decide which

### 2012-04-26 paper/paper.tex
Fix broken references

--- 2012-04-25 paper/paper.tex
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{stretching_cybershake}
+    \includegraphics[width=0.3\textwidth]{stretching_cybershake}
     \caption[Percentage of high scores achieved by each algorithm for Cybershake pareto unsorted ensembles when task runtime is stretched]{Percentage of high scores achieved by each algorithm for Cybershake pareto unsorted ensembles when task runtime is stretched.
@@
 This is a consequence of the deadline distribution function in SPSS, which prevents workflows from starting late.
-This is shown by the gantt chart in Figure~\ref{fig:spss-example}, which illustrates how SPSS tends to pile up workflows at the beginning of the timeline.
-In comparison, the dynamic algorithms tend to spread out workflow starts over the entire duration as shown in Figure~\ref{fig:spds-example}.
+This is shown by the gantt chart in Figure~\ref{fig:algorithm-example}b, which illustrates how SPSS tends to pile up workflows at the beginning of the timeline.
+In comparison, the dynamic algorithms tend to spread out workflow starts over the entire duration as shown in Figure~\ref{fig:algorithm-example}a.
 The consequence of this behavior is that SPSS plans tend to use up the budget but leave plenty of time before the deadline.
@@
     }
+    \hspace{2cm}
     \subfloat[Ratio of Actual Makespan to Deadline]{
-    \hspace{2cm}
-        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all} }
+        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all}
+    }
     \caption[Boxplots for budget and deadline ratios when runtime estimates are inaccurate]{Boxplots for budget and deadline ratios when runtime estimate error varies from $\pm$0\% to $\pm$50\% for all three algorithms.
@@
     }
+    \hspace{2cm}
     \subfloat[Ratio of Actual Makespan to Deadline]{    
-        \hspace{2cm}
         \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
@@
     }
+    \hspace{2cm}
     \subfloat[Ratio of Actual Makespan to Deadline]{
-        \hspace{2cm}
         \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-makespan-all}
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{spss_planning_time}
+    \includegraphics[width=0.3\textwidth]{spss_planning_time}
     \caption{Planning time of SPSS algorithm for ensembles of 100 workflows and different workflow sizes.}
@@
 % This balances the columns by inserting a column break at the given reference
-\IEEEtriggeratref{13}
+\IEEEtriggeratref{23}
 

### 2012-04-26 paper/paper.tex
Update relative performance section

@@
 
-The goal of the first experiment is to characterize the relative performance of the proposed algorithms.
-This was done by simulating the algorithms on many different ensembles and comparing the scores computed using the exponential scoring technique described in Section~\ref{sec:perf_metric}.
-In this experiment, the algorithm that achieves the largest number of high scores has the best performance.
-
-Figure~\ref{fig:distributions} shows the percentage of simulations for which each algorithm achieved the highest score for a given ensemble type.
-This experiment was conducted using all five applications, with all five types of ensembles.
-For each application, ensembles of 50 workflows were created using 10 different random seeds.
-Each ensemble was simulated with all three algorithms using 10 budgets and 10 deadlines.
-The total number of simulations to produce each bar in the figure is: $10 * 10 * 10 = 1000$.
-The best scores percentage is computed by counting the number of times that a given algorithm achieved the highest score and dividing by 1000.
-Note that it is possible for multiple algorithms to get the same high score (to tie), so the numbers do not necessarily add up to 100\%.
-The sum is much higher than 100\% in cases where the dynamic algorithms perform relatively well because DPDS and WA-DPDS, which are very similar algorithms, often get the same high score.
+The goal of the first experiment is to characterize the relative performance
+of the proposed algorithms. This was done by simulating the algorithms on many
+different ensembles and comparing the scores computed using the exponential
+scoring technique described in Section~\ref{sec:perf_metric}.
+
+Figure~\ref{fig:distributions} shows the percentage of simulations for which
+each algorithm achieved the highest score for a given ensemble type. This
+experiment was conducted using all five applications, with all five types of
+ensembles. For each application and ensemble type, 10 random ensembles of 50
+workflows each were created. Each ensemble was simulated with all three
+algorithms using 10 budgets and 10 deadlines (1000 simulations per
+application, ensemble type, and algorithm). The best scores percentage is
+computed by counting the number of times that a given algorithm achieved the
+highest score and dividing by 1000. Note that it is possible for multiple
+algorithms to get the same high score (to tie), so the numbers do not
+necessarily add up to 100\%. The sum is much higher than 100\% in cases where
+the dynamic algorithms perform relatively well because DPDS and WA-DPDS, which
+are very similar algorithms, often get the same high score.
 
@@
     \centering
-    \includegraphics[width=1.06\textwidth]{run-finish-variations-test-0-output-distributions}
-    \caption[Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.]{Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.
+    \includegraphics[width=\textwidth]{run-finish-variations-test-0-output-distributions}
+    \caption{Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.
 C = Constant ensembles, PS = Pareto Sorted ensembles, PU = Pareto Unsorted ensembles, US = Uniform Sorted ensembles, and UU = Uniform Unsorted ensembles.}
@@
 
-There are several interesting things to notice about Figure~\ref{fig:distributions}.
-The first is that, in most cases, SPSS significantly outperforms the dynamic algorithms (DPDS and WA-DPDS).
-This is attributed to the fact that SPSS is able to make more intelligent scheduling and provisioning decisions because it has the opportunity to compare different decisions and choose the one that results in the best outcome.
-In comparison, the dynamic algorithms are online algorithms and are not able to project into the future to weigh the outcomes of their choices.
-
-The second thing to notice about the figure is that, for constant ensembles, the dynamic algorithms perform significantly better relative to SPSS than for other ensemble types.
-This is a result of the fact that, since all of the workflows are of approximately the same size and shape, the choice of which workflow to execute next is arbitrary.
-
-Another thing to notice about these results is that the workflow-aware algorithms (WA-DPDS and SPSS) both perform better in most cases than the simple online algorithm that uses resource utilization alone to make provisioning decisions.
-This suggests that there is a significant value in having information about the structure and estimated runtime of a workflow when making scheduling and provisioning decisions.
-
-Finally, it is interesting to see that, for Montage and CyberShake, the relative performance of SPSS compared to the dynamic algorithms is significantly less than it is for other applications.
-We attribute this to the structure of Montage and CyberShake.
-The workflows for both applications are very wide relative to their height, and both have very short-running tasks.
-As a result of these two characteristics, the Montage and CyberShake applications have very short critical paths and look more like bag-of-tasks applications, which are easier to execute than more structured applications.
-DPDS and WA-DPDS are able to pack more of the tasks into the available budget and deadline because there are a) more choices for where to place the tasks, and b) the different choices have a smaller impact on the algorithms' ability to execute the workflow within the constraints.
-In addition, the short critical paths put SPSS at a disadvantage.
-Because of the way SPSS assigns deadlines to individual tasks, it is prevented from starting workflows late, which prevents it from packing tasks into the idle VM slots at the end of the schedule.
+There are several interesting things to notice about
+Figure~\ref{fig:distributions}. The first is that, in most cases, SPSS
+significantly outperforms both dynamic algorithms (DPDS and WA-DPDS). This is
+attributed to the fact that SPSS is able to make more intelligent scheduling
+and provisioning decisions because it has the opportunity to compare different
+options and choose the one that results in the best outcome. In comparison,
+the dynamic algorithms are online algorithms and are not able to project into
+the future to weigh the outcomes of their choices.
+
+The second thing to notice about the figure is that, for constant ensembles,
+the dynamic algorithms perform significantly better relative to SPSS than for
+other ensemble types. This is a result of the fact that, since all of the
+workflows are of approximately the same size and shape, the choice of which
+workflow to execute next has little impact on the final result.
+
+Another thing to notice about this experiment is that the workflow-aware
+algorithms (WA-DPDS and SPSS) both perform better in most cases than the
+simple online algorithm that uses resource utilization alone to make
+provisioning decisions (DPDS). This suggests that there is a significant value
+in having information about the structure and estimated runtime of a workflow
+when making scheduling and provisioning decisions.
+
+Finally, it is interesting to see that, for Montage and CyberShake, the
+relative performance of SPSS is significantly less than it is for other
+applications. We attribute this to the structure of Montage and CyberShake.
+The workflows for both applications are very wide relative to their height,
+and both have very short-running tasks. As a result of these two
+characteristics, the Montage and CyberShake applications have very short
+critical paths and look more like bag-of-tasks applications, which are easier
+to execute than more structured applications. DPDS and WA-DPDS are able to
+pack more of the tasks into the available budget and deadline because there
+are a) more choices for where to place the tasks, and b) the different choices
+have a smaller impact on the algorithms' ability to execute the workflow
+within the constraints. In addition, the short critical paths put SPSS at a
+disadvantage. Because of the way SPSS assigns deadlines to individual tasks,
+it is prevented from starting workflows late, which prevents it from packing
+tasks into the idle VM slots at the end of the schedule.
 

### 2012-04-26 paper/paper.tex
Update task granularity section

@@
 \subsection{Task Granularity}
-
-In the last experiment we noted that, for Montage and CyberShake, the short runtimes of their tasks made it possible for the dynamic algorithms to perform better relative to SPSS when compared with other applications.
-In order to test this theory we adjusted the granularity of the tasks in several Montage and CyberShake ensembles to see how this would affect the relative performance of the algorithms.
-The granularity adjustment was achieved by multiplying the runtime of each task by a fixed scaling factor.
-
-Figure~\ref{fig:stretching} shows the relative performance of the algorithms as the scaling factor is increased from 1 to 16.
-This figure only shows results for the CyberShake application with pareto unsorted ensembles, but similar results were found for other ensemble types for both Montage and CyberShake.
-Each data point in the figure represents the result of 500 simulations (5 random seeds for 10 budgets and 10 deadlines).
-The best scores percent was calculated the same way as it was for Figure~\ref{fig:distributions}.
-
-The figure shows that, as the scaling factor increases, the relative performance of SPSS increases as well.
-In this case SPSS performance surpasses DPDS and WA-DPDS for scaling factors larger than 2.
-In all other cases (other ensemble types for both applications), the performance of SPSS is superior to the dynamic algorithms for scaling factors greater than about 8.
-This suggests that, in general, for fine-grained workflows the dynamic algorithms will produce better scores, and for coarse-grained workflows SPSS will produce better scores.
+\TODO{Should we remove this section?}
+
+In the last experiment we noted that, for Montage and CyberShake, the short
+runtimes of their tasks made the dynamic algorithms perform better relative to
+SPSS. In order to test this theory we adjusted the granularity of the tasks in
+several Montage and CyberShake ensembles to see how this would affect the
+relative performance of the algorithms. The granularity adjustment was
+achieved by multiplying the runtime of each task by a fixed scaling factor.
+
+Figure~\ref{fig:stretching} shows the relative performance of the algorithms
+as the scaling factor is increased from 1 to 16. The figure only shows results
+for the CyberShake application with pareto unsorted ensembles, but similar
+results were found for other ensemble types for both Montage and CyberShake.
+Each data point in the figure represents 500 simulations (5 ensembles with 10
+budgets and 10 deadlines). The best scores percent was calculated the same way
+as it was for Figure~\ref{fig:distributions}.
+
+The figure shows that, as the scaling factor increases beyond 2, the relative
+performance of SPSS surpasses that of the dynamic algorithms. Similar results
+were found for other ensemble types for both applications. This result
+suggests that, in general, for fine-grained workflows the dynamic algorithms
+will produce better scores, and for coarse-grained workflows SPSS will produce
+better scores.
 
@@
     \includegraphics[width=0.3\textwidth]{stretching_cybershake}
-    \caption[Percentage of high scores achieved by each algorithm for Cybershake pareto unsorted ensembles when task runtime is stretched]{Percentage of high scores achieved by each algorithm for Cybershake pareto unsorted ensembles when task runtime is stretched.
-A stretching factor of $x$ means that the runtime of tasks in each workflow was multiplied by $x$.}
+    \caption{Percentage of high scores achieved by each algorithm for pareto 
+    unsorted CyberShake ensembles when task runtime is scaled by a constant 
+    factor. A scaling factor of $x$ means that the runtime of tasks in each 
+    workflow was multiplied by $x$.}
     \label{fig:stretching}

### 2012-04-26 paper/paper.tex
Edit inaccurate runtime estimates section

@@
 
-Both of the workflow-aware algorithms rely on estimates of task runtimes to make better scheduling and provisioning decisions.
-However, these estimates can be inaccurate.
-It is typically not possible to predict exactly how long a task will take to execute on a given machine at a given time.
-The actual runtime depends on many factors that are beyond the control of the application, such as network or disk contention, background processing, etc.
-
-Given that task runtime estimates are bound to be inaccurate, the question is: How does the error in the task runtime estimate impact the performance of our scheduling and provisioning algorithms?
-To test this we created an experiment to measure how introducing uniform errors in the runtime estimates affects the ability of the algorithms to achieve the budget and deadline constraints.
-This experiment used a single application, Montage, and a single ensemble type, uniform unsorted.
-This application and ensemble type was chosen arbitrarily for simplicity.
-It is expected that the application and ensemble type have little impact on how well the algorithms are able to stay within constraints.
-
-The estimation errors are generated by randomly sampling a uniform distribution between $-p\%$ and $+p\%$ of the runtime of a task.
-So, for example, if the value of $p$ is $10\%$, then the actual runtime of a task ($r_a$) is set in the simulation to be: $r_a = r_e * (1.0 + q)$, where $r_e$ is the estimated runtime of the task and $q$ is a uniform random number beween $-0.1$ and $+0.1$.
-Since the sampling is done uniformly, we expect to get just as many overestimates as underestimates in any given simulation.
-
-Note that the algorithms have \textbf{not} been changed to account for inaccurate runtime estimates.
-It is likely that better performance could be achieved if the algorithms were given some hint as to the accuracy of the task runtimes, but investigating that optimization is left for future work.
-The goal here is to determine how well the current algorithms are able to stay within the constraints given inaccurate task runtime estimates.
-
-Figure~\ref{fig:variances} shows the results of simulations for errors ranging from $0\%$ to $50\%$.
-These simulations were done for ensembles of 50 workflows, using 10 random seeds, 10 budgets, and 10 deadlines for each error value and algorithm (3000 simulations per error value).
-Each plot in the figure shows box plots of the ratio of actual cost to budget, or actual makespan to deadline.
-The actual cost and makespan are the real cost and makespan of the workflow calculated after the simulation completed.
-The ratio of actual values to constraints indicates whether the actual value exeeded the constraint.
-Values equal to 1 mean that the actual value was equal to the constraint, values less than 1 indicate that the actual value was within the constraint, and values greater than 1 mean that the actual value exceeded the constraint.
-
-Figure~\ref{fig:variances}.a shows the ratio of actual cost to budget.
-These results show two important characteristics of algorithm behavior.
-First, both dynamic algorithms very rarely exceeded the budget, even with very large errors.
-This is indicative of the fact that the dynamic algorithms are easily able to adapt to changes at runtime to ensure good results without exceeding the constraints, even when the quality of information available to them is low.
-Second, unlike the dynamic algorithms, the static algorithm frequently exceeded the budget constraint when the quality of runtime estimates was low.
-For small errors the budget was exceeded infrequently by small amounts, and for large errors (more than 10\%) the budget exceeded frequently by large amounts.
-This is a result of the fact that the static algorithm makes all of its decisions before execution and is not able to adapt to changing circumstances.
-The SPSS plan describes what tasks to execute on which VMs, but not when to execute them.
-At runtime, the worklow engine is forced to spend extra money to extend the runtime of some VMs so that all of the tasks assigned to that VM can be completed.
-At the same time, other VMs can be terminated early because the tasks they were assigned finished earlier than expected.
-However, because of dependencies in the workflow, the latter case is less likely to happen, which causes gaps in the schedule.
-So the net result is that the overall cost is increased.
-
-Figure~\ref{fig:variances}.b shows the ratio of actual makespan to budget.
-The interesting thing to notice about this figure is that, for all three algorithms, the deadline constraint is rarely exceeded, even in cases with very poor quality estimates.
-For the dynamic algorithms this is a result of the fact that they can adapt at runtime to changing circumstances to avoid exceeding the deadline.
-For the static algorithm this is a result of the way that SPSS schedules workflows and not of a particularly clever optimization.
-The SPSS algorithm tends to schedule workflows early, using up the budget long before the deadline is reached.
-This is a consequence of the deadline distribution function in SPSS, which prevents workflows from starting late.
-This is shown by the gantt chart in Figure~\ref{fig:algorithm-example}b, which illustrates how SPSS tends to pile up workflows at the beginning of the timeline.
-In comparison, the dynamic algorithms tend to spread out workflow starts over the entire duration as shown in Figure~\ref{fig:algorithm-example}a.
-The consequence of this behavior is that SPSS plans tend to use up the budget but leave plenty of time before the deadline.
-So when the runtime of the plan is increased by introducing errors, the SPSS plan has some room to expand without exceeding the deadline.
-
-% Figure~\ref{fig:variances_scores} shows relative performance as a percentage of the number of high scores achieved by each algorithm as the error increases.
-The figure shows that, for this experiment, the scores remain the same as the error increases.
-This is a result of the way the error is applied.
-Since the error is selected uniformly from $-p\%$ to $+p\%$ it is equally likely that the error will be positive as negative.
-In other words, it is equally likely that a task will take longer as it is that the task will be shorter.
-Because the application used in this experiment, Montage, has a low task runtime variance, applying the error leaves the total runtime of the tasks in the workflow unchanged---the total number of CPU hours in each workflow remains the same regardless of the error.
-The end result is that the dynamic algorithms are able to achieve the same score without exceeding the constraints by using a different schedule that finishes the same total amount of work using the same budget and deadline.
-If we had applied a skewed error, for example, an error that was more likely to be positive than negative, then the scores for the dynamic algorithms would decrease relative to the static algorithm as the error increased.
-
-The behavior of these algorithms indicate that, when the quality of runtime estimates is low, the dynamic algorithms do a much better job of adapting to the changing circumstances to meet their obligations.
-In comparison, the static algorithm performs better with good information, but fails to meet the constraints when the quality of information is poor.
-It is possible that the static algorithm could be modified to add more breathing room in the schedule to account for situations where the runtime estimates may be inaccurate.
-Such a modification would result in slightly worse scores when the estimates are good, but would improve scores when the estimates are bad.
-Alternatively, the SPSS algorithm could be modified to generate a plan that specifies when to provision and deprovision each VM.
-In that case the workflow engine could be prevented from exceeding the budget and deadline constraints, but would prevent some workflows from finishing, which may significantly decrease the score.
-These two modifications are subjects for future work.
+Both of the workflow-aware algorithms rely on estimates of task runtimes to
+make better scheduling and provisioning decisions. In practice, however, these
+estimates are often inaccurate. Given inaccuate estimates, the question is:
+How do errors in task runtime estimates impact the performance of our
+scheduling and provisioning algorithms? We created an experiment to measure
+how introducing uniform errors in these estimates affects the ability of the
+algorithms to achieve the desired budget and deadline constraints.
+
+\TODO{Update the description of the experiment in this paragraph if necessary}
+
+This experiment used a single application, Montage, and a single ensemble
+type, uniform unsorted. This application and ensemble type was chosen
+arbitrarily for simplicity. It is expected that the application and ensemble
+type have little impact on how well the algorithms are able to stay within
+constraints.
+
+The actual runtime of each task is adjusted in the simulation by adding a
+random error to the estimated runtime of $\pm p\%$. Since the sampling is done
+uniformly, we expect to get just as many overestimates as underestimates in
+any given simulation.
+
+Figure~\ref{fig:variances} shows the results of simulations for estimate
+errors ranging from $0\%$ to $50\%$. These simulations were done for 10
+ensembles of 50 workflows using 10 budgets and 10 deadlines for each error
+value and algorithm (3000 simulations per error value). The figure shows box
+plots of the ratio of actual cost to budget and actual makespan to deadline.
+The actual cost and makespan are the simulated cost and makespan of the
+ensemble. The ratio of actual values to constraints indicates whether the
+actual value exeeded the constraint: Values less than 1 indicate that the
+constraint was not exceeded, and values greater than 1 indicate that the
+constraint was exceeded.
+
+Figure~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
+figure illustrates two important algorithm characteristics. First, the dynamic
+algorithms very rarely exceeded the budget, even with very large errors. This
+indicates that the dynamic algorithms are able to adapt to uncertainties at
+runtime to ensure that the constraints are not exceeded, even when the quality
+of information available to them is low. Second, unlike the dynamic
+algorithms, the static algorithm frequently exceeded the budget constraint; by
+large amounts in some cases. This is a result of the fact that the static
+algorithm makes all of its decisions before execution and is not able to adapt
+to changing circumstances.
+
+%The SPSS plan describes what tasks to execute on which VMs, but not when to
+%execute them. At runtime, the worklow engine is forced to spend extra money
+%to extend the runtime of some VMs so that all of the tasks assigned to that
+%VM can be completed. At the same time, other VMs can be terminated early
+%because the tasks they were assigned finished earlier than expected. However,
+%because of dependencies in the workflow, the latter case is less likely to
+%happen, which causes gaps in the schedule. So the net result is that the
+%overall cost is increased.
+
+Figure~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
+interesting thing to notice about this figure is that, for all three
+algorithms, the deadline constraint is rarely exceeded, even in cases with
+very poor quality estimates. For the dynamic algorithms this is a result of
+the fact that they can adapt to poor estimates. For the static algorithm this
+is a result of the way that SPSS schedules workflows, and not of a
+particularly clever optimization. The SPSS algorithm tends to schedule
+workflows early, using up the budget long before the deadline is reached. This
+is a consequence of the deadline distribution function in SPSS, which prevents
+workflows from starting late. This is shown by the gantt chart in
+Figure~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
+up workflows at the beginning of the timeline. In comparison, the dynamic
+algorithms tend to spread out workflow starts over the entire duration as
+shown in Figure~\ref{fig:algorithm-example}.a. The consequence of this
+behavior is that SPSS plans tend to use up the budget but leave plenty of time
+before the deadline. So when the runtime of the plan is increased by
+introducing errors, the SPSS plan has some room to expand without exceeding
+the deadline.
+
+Note that the algorithms have \textit{not} been changed to account for
+inaccurate runtime estimates in this experiment. It is likely that better
+performance could be achieved if the algorithms were given a hint as to the
+accuracy of the task runtimes. Investigating that optimization is left for
+future work. The goal here is to determine how well the current algorithms are
+able to stay within the constraints given inaccurate task runtime estimates.
+
+%It is possible that the static algorithm could be modified to add more
+%breathing room in the schedule to account for situations where the runtime
+%estimates may be inaccurate. Such a modification would result in slightly
+%worse scores when the estimates are good, but would improve scores when the
+%estimates are bad. Alternatively, the SPSS algorithm could be modified to
+%generate a plan that specifies when to provision and deprovision each VM. In
+%that case the workflow engine could be prevented from exceeding the budget
+%and deadline constraints, but would prevent some workflows from finishing,
+%which may significantly decrease the score. These two modifications are
+%subjects for future work.
 
@@
     }
-    \caption[Boxplots for budget and deadline ratios when runtime estimates are inaccurate]{Boxplots for budget and deadline ratios when runtime estimate error varies from $\pm$0\% to $\pm$50\% for all three algorithms.
-A value of 1 indicates that the actual cost is equal to the budget, or the actual makespan is equal to the deadline, respectively.
-Values greater than 1 indicate that the budget/deadline has been exceeded.}
+    \caption{Boxplots for budget and deadline ratios when runtime estimate 
+    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. A 
+    value of 1 indicates that the actual cost is equal to the budget, or 
+    the actual makespan is equal to the deadline, respectively. Values 
+    greater than 1 indicate that the constraint has been exceeded.}
     \label{fig:variances}

### 2012-04-26 paper/paper.tex
Edit provisioning delays section

@@
 
-\begin{figure}[htb]
+\begin{figure}[tb]
     \centering
@@
 \begin{figure*}[tb]
+    \centering
     \subfloat[Ratio of Actual Cost to Budget]{
@@
     \caption{Boxplots for budget and deadline ratios when runtime estimate 
-    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. A 
-    value of 1 indicates that the actual cost is equal to the budget, or 
-    the actual makespan is equal to the deadline, respectively. Values 
-    greater than 1 indicate that the constraint has been exceeded.}
+    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. Values 
+    greater than 1 indicate that the budget/deadline constraint was exceeded.}
     \label{fig:variances}
@@
 
-One important issue to consider when provisioning resources in the cloud is the amount of time between when a resource is requested, and when it becomes available to the application.
-Typically these provisioning delays are on the order of a few minutes, and are highly dependent upon the architecture of the cloud system and/or the size of the user's VM image~\cite{Nurmi2008b}.
-
-Recall that it is assumed that resources are billed from the minute that they are requested until they are terminated, and not from the moment that they become available.
-As a result, provisioning delays have an impact on both the cost (budget constraint) and makespan (deadline constraint) of the ensemble.
-
-Figure~\ref{fig:delays} shows the ratios of actual values to constraints when the provisioning delay is increased from 0 seconds up to 15 minutes.
-The simulations were done using uniform unsorted ensembles of 50 workflows from the Montage application.
-Each plot in the figure contains three boxes: one for each algorithm.
-Each box summarizes the results of 10 budgets, 10 deadlines, and 10 random seeds for a total of 1000 simulations per algorithm and provisioning delay (3000 simulations per plot).
-As in Figure~\ref{fig:variances}, the y-axis in each plot represents the ratio of the actual simulated value to the constraint value for the budget constraint (actual cost) or the deadline constraint (actual makespan).
-These ratios indicate how frequently and by how much the actual simulated values exceeded (or did not exceed) the constraints.
-
-%\begin{figure}[ht]
-%  \centering
-%  \includegraphics[width=0.5\textwidth]{variances_scores}
-%  \caption{Percentage of high scores achieved by each algorithm on Montage with
-%  uniform unsorted ensembles when runtime estimate error varies from $\pm$0\%
-%  to $\pm$50\%.}
-%  \label{fig:variances_scores}
-%\end{figure}
+One important issue to consider when provisioning resources in the cloud is
+the amount of time between when a resource is requested, and when it actually
+becomes available to the application. Typically these provisioning delays are
+on the order of a few minutes, and are highly dependent upon the architecture
+of the cloud system and/or the size of the user's VM image~\cite{Nurmi2008b}.
+
+We assume that resources are billed from the minute that they are requested
+until they are terminated, and not from the moment that they become available.
+As a result, provisioning delays have an impact on both the cost (budget
+constraint) and makespan (deadline constraint) of the ensemble.
+
+\TODO{Update the following paragraph depending on what results Maciek included}
+
+Figure~\ref{fig:delays} shows the ratios of actual values to constraints when
+the provisioning delay is increased from 0 seconds up to 15 minutes. These
+simulations were done using 10 uniform unsorted ensembles of 50 workflows from
+the Montage application, using 10 deadlines (1000 simulations per algorithm
+and provisioning delay). As in Figure~\ref{fig:variances}, the y-axis in each
+plot represents the ratio of the actual simulated value to the constraint
+value for the budget constraint (actual cost) or the deadline constraint
+(actual makespan). These ratios indicate how frequently and by how much the
+actual simulated values exceeded (or did not exceed) the constraints.
 
 \begin{figure*}[tb]
+    \centering
     \subfloat[Ratio of Actual Cost to Budget]{
@@
     }
-    \caption[Boxplots for budget and deadline ratios when provisioning delays occur]{Boxplots for budget and deadline ratios when provisioning delay varies from 0 seconds to 15 minutes for all three algorithms.
-A value of 1 indicates that the actual cost is equal to the budget, or the actual makespan is equal to the deadline, respectively.
-Values greater than 1 indicate that the budget/deadline has been exceeded.}
+    \caption{Boxplots for budget and deadline ratios when provisioning delay 
+    varies from 0 seconds to 15 minutes for all three algorithms. Values 
+    greater than 1 indicate that the budget/deadline constraint was exceeded.}
     \label{fig:delays}
 \end{figure*}
+
+\TODO{Update this next paragraph depending on what Maciek included in the figure. The results seem to have changed.}
 
@@
 The difference is that, a very small error in runtime estimates has little impact on the budget, while a very small provisioning delay has a large impact on the budget.
-Figure~\ref{fig:delays}.a shows that even for an impossibly short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget.
+Figure~\ref{fig:delays}.a shows that even for a very short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget.
 In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget.
@@
 
-% Figure~\ref{fig:delays_scores} shows the relative performance as a percentage of the number of high scores achieved by each algorithm as the provisioning delay increases.
-This figure shows that, as the delay increases, the relative performance of the static algorithm increases as well.
-This behavior is a result of the fact that the static algorithm is, in essence, cheating by using more time and money than the dynamic algorithms.
-In the previous section we showed how the relative performance of the algorithms remains the same when a uniform error is applied to the runtime constraints.
-In that case, the dynamic algorithms adapted to the error by rearranging the schedule to accomplish the same amount of work given the same deadline and budget.
-In this case, the dynamic algorithms adapt by performing less work (executing fewer workflows) to remain within the constraints while accounting for the delays.
-In both cases, the static algorithm completed the same amount of work, but did so by exceeding the constraints.
+% Figure~\ref{fig:delays_scores} shows the relative performance as a
+% percentage of the number of high scores achieved by each algorithm as the
+% provisioning delay increases. This figure shows that, as the delay
+% increases, the relative performance of the static algorithm increases as
+% well. This behavior is a result of the fact that the static algorithm is, in
+% essence, cheating by using more time and money than the dynamic algorithms.
+% In the previous section we showed how the relative performance of the
+% algorithms remains the same when a uniform error is applied to the runtime
+% constraints. In that case, the dynamic algorithms adapted to the error by
+% rearranging the schedule to accomplish the same amount of work given the
+% same deadline and budget. In this case, the dynamic algorithms adapt by
+% performing less work (executing fewer workflows) to remain within the
+% constraints while accounting for the delays. In both cases, the static
+% algorithm completed the same amount of work, but did so by exceeding the
+% constraints.
 
@@
 %   \includegraphics[width=0.5\textwidth]{delays_scores}
-%   \caption{Percentage of high scores achieved by each algorithm on Montage with uniform unsorted ensembles when provisioning delay varies from 0 seconds to 15 minutes.}
+%   \caption{Percentage of high scores achieved by each algorithm on Montage 
+%   with uniform unsorted ensembles when provisioning delay varies from 0 
+%   seconds to 15 minutes.}
 %   \label{fig:delays_scores}
 % \end{figure}
+
+\TODO{Update this paragraph when Maciek explains what was included in the plot.}
 
@@
     \caption{Boxplots for budget and deadline ratios when failure rate varies
-    from 0 seconds to 50 \% for all three algorithms. A value of 1
-    indicates that the actual cost is equal to the budget, or the actual makespan is equal to the deadline, respectively. Values greater than 1 indicate that the budget/deadline has been exceeded.}
+    from 0\% to 50\% for all three algorithms. Values greater than 1 indicate 
+    that the budget/deadline constraint was exceeded.}
     \label{fig:failures}
@@
 % This balances the columns by inserting a column break at the given reference
-\IEEEtriggeratref{23}
+\IEEEtriggeratref{19}
 

### 2012-04-27 paper/paper.tex
Edit task failures section

@@
 Running workflows consisting of large numbers of tasks on distributed systems
-almost always faces different types of failures. The goal of this experiment was
-to asses the behaviour of the algorithms in the presence of task execution
+often results in failures. The goal of the next experiment was
+to assess the behavior of the algorithms in the presence of task execution
 failures by introducing a failure model into the simulator. The model is
@@
 will fail. The failure time of the task is determined by randomly sampling a
-time between task start and finish time. If the task fails, it is reported to
-the workflow engine, which then resubmits the task for execution. Dynamic algorithms
-add the task again to the priority queue (queue $P$ in Algorithm~\ref{alg:ds})
-so that it can be resubmitted by the scheduler. Static algorithm resubmits the
-failed task on the same VM that was selected in the plan.
+value between the task start time and finish time. If the task fails, it is reported to
+the workflow engine, which retries the task until it succeeds. The dynamic algorithms
+re-add the task to the priority queue (queue $P$ in Algorithm~\ref{alg:ds})
+so that it can be resubmitted by the scheduler. The SPSS algorithm immediately 
+resubmits the failed task to the same VM that was selected in the plan.
 
@@
 \end{figure*}
+
+\TODO{Include a description of the ensembles that were simulated.}
 
@@
 larger than 10\%. We consider higher failure rates as rare events that suggest
-some significant system malfunction or improper resource selection
-process. It is thus more important that all the algorithms can handle task
-failures by resubmission (as is the case of Pegasus and Condor DAGMan).
+a significant system malfunction or invalid selection of resources.
 

### 2012-04-27 paper/paper.tex
Change Figure to Fig.

@@
 workflows will be completed within the current budget and deadline.
-Figure~\ref{fig:algorithm-example}a shows an example schedule generated using
+Fig.~\ref{fig:algorithm-example}a shows an example schedule generated using
 the DPDS algorithm. The figure illustrates how tasks from lower priority workflows
@@
 
-An example schedule generated by SPSS (Figure~\ref{fig:algorithm-example}b)
+An example schedule generated by SPSS (Fig.~\ref{fig:algorithm-example}b)
 shows how SPSS tends to start many workflows in parallel, running each workflow
@@
 %An example of the distribution of workflow sizes that
-%occurs in a Pareto ensemble is shown in Figure~\ref{fig:ensemble-pareto}.
+%occurs in a Pareto ensemble is shown in Fig.~\ref{fig:ensemble-pareto}.
 
@@
 
-Figure~\ref{fig:distributions} shows the percentage of simulations for which
+Fig.~\ref{fig:distributions} shows the percentage of simulations for which
 each algorithm achieved the highest score for a given ensemble type. This
@@
 There are several interesting things to notice about
-Figure~\ref{fig:distributions}. The first is that, in most cases, SPSS
+Fig.~\ref{fig:distributions}. The first is that, in most cases, SPSS
 significantly outperforms both dynamic algorithms (DPDS and WA-DPDS). This is
@@
 
-Figure~\ref{fig:stretching} shows the relative performance of the algorithms
+Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
 as the scaling factor is increased from 1 to 16. The figure only shows results
@@
 budgets and 10 deadlines). The best scores percent was calculated the same way
-as it was for Figure~\ref{fig:distributions}.
+as it was for Fig.~\ref{fig:distributions}.
 
@@
 
-Figure~\ref{fig:variances} shows the results of simulations for estimate
+Fig.~\ref{fig:variances} shows the results of simulations for estimate
 errors ranging from $0\%$ to $50\%$. These simulations were done for 10
@@
 
-Figure~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
+Fig.~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
 figure illustrates two important algorithm characteristics. First, the dynamic
@@
 
-Figure~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
+Fig.~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
 interesting thing to notice about this figure is that, for all three
@@
 workflows from starting late. This is shown by the gantt chart in
-Figure~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
+Fig.~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
 up workflows at the beginning of the timeline. In comparison, the dynamic
 algorithms tend to spread out workflow starts over the entire duration as
-shown in Figure~\ref{fig:algorithm-example}.a. The consequence of this
+shown in Fig.~\ref{fig:algorithm-example}.a. The consequence of this
 behavior is that SPSS plans tend to use up the budget but leave plenty of time
@@
 
-Figure~\ref{fig:delays} shows the ratios of actual values to constraints when
+Fig.~\ref{fig:delays} shows the ratios of actual values to constraints when
 the provisioning delay is increased from 0 seconds up to 15 minutes. These
@@
 the Montage application, using 10 deadlines (1000 simulations per algorithm
-and provisioning delay). As in Figure~\ref{fig:variances}, the y-axis in each
+and provisioning delay). As in Fig.~\ref{fig:variances}, the y-axis in each
 plot represents the ratio of the actual simulated value to the constraint
@@
 The difference is that, a very small error in runtime estimates has little impact on the budget, while a very small provisioning delay has a large impact on the budget.
-Figure~\ref{fig:delays}.a shows that even for a very short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget.
+Fig.~\ref{fig:delays}.a shows that even for a very short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget.
 In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget.
@@
 
-% Figure~\ref{fig:delays_scores} shows the relative performance as a
+% Fig.~\ref{fig:delays_scores} shows the relative performance as a
 % percentage of the number of high scores achieved by each algorithm as the
@@
 
-Figure~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles of 100 workflows with five different workflow sizes: 50, 200, 400, 600, and 800 tasks.
+Fig.~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles of 100 workflows with five different workflow sizes: 50, 200, 400, 600, and 800 tasks.
 The ensembles were generated using a constant distribution equal to the workflow size desired.
@@
 
-Figure~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS planning time is reasonable, taking on the order of tens of seconds to a few minutes.
+Fig.~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS planning time is reasonable, taking on the order of tens of seconds to a few minutes.
 For larger ensembles of large workflows, however, the SPSS planning time can easily reach 10 minutes.

### 2012-04-27 paper/paper.tex
Edit abstract to 150-word SC limit

@@
 
-Large-scale computational applications represented as scientific workflows, are
-often grouped into ensembles composed of several inter-related workflows. In
-this paper, we address a new and important problem concerning the efficient 
-management of such ensembles under both budget and deadline constraints on 
-Infrastructure-as-a-Service (IaaS) clouds. IaaS clouds are characterized by
-on-demand resource provisioning capabilities and a pay-per-use model.  We discuss,
-develop, and assess algorithms based on static and dynamic strategies for both
-task scheduling and resource provisioning. We assume that the workflows within
-an ensemble are prioritized by the user. We evaluate the algorithms via
-simulation using a set of scientific workflow ensembles with a broad range of
-budget and deadline parameters, taking into account uncertainties in task runtime
-estimations, provisioning delays and failures. We find that the key 
-factor determining the performance of an algorithm is its ability to decide which
-workflows in an ensemble to admit or reject for execution. The results
-show that the admission procedure based on workflow structure and estimates of
-task runtimes can significantly improve the quality of solutions.
+Large-scale computational applications expressed as scientific workflows are
+often grouped into ensembles of inter-related workflows. In this paper, we
+address a new and important problem concerning the efficient  management of
+such ensembles under budget and deadline constraints on  Infrastructure-
+as-a-Service (IaaS) clouds. We discuss, develop, and assess algorithms based
+on static and dynamic strategies for both task scheduling and resource
+provisioning. We evaluate the algorithms via simulation using a set of
+scientific workflow ensembles with a broad range of budget and deadline
+parameters, taking into account uncertainties in task runtime estimations,
+provisioning delays and failures. We find that the key factor determining the
+performance of an algorithm is its ability to decide which workflows in an
+ensemble to admit or reject for execution. Our results show that an admission
+procedure based on workflow structure and estimates of task runtimes can
+significantly improve the quality of solutions.
 

### 2012-04-28 paper/paper.tex
Edit results

@@
 
-In order to account for the user's priority, we used an exponential scoring where the score for an ensemble $e$ is:
+In order to account for the user's priority, we used an exponential scoring
+where the score for an ensemble $e$ is:
 %
@@
 %
-where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that the highest-priority workflow has $Priority(w)=0$, the next highest workflow has $Priority(w)=1$, and so on.
-This exponential scoring function gives the highest priority workflow a score that is higher than all the lower-priority workflows combined:
+where $Completed(e)$ is the set of workflows in ensemble $e$ that was
+completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$
+such that the highest-priority workflow has $Priority(w)=0$, the next highest
+workflow has $Priority(w)=1$, and so on. This exponential scoring function
+gives the highest priority workflow a score that is higher than all the lower-
+priority workflows combined:
 %
@@
 %
-This scoring is consistent with our definition of the problem, which is to complete as many workflows as possible, according to their priorities, given a set budget and deadline.
+This scoring is consistent with our definition of the problem, which is to
+complete as many workflows as possible, according to their priorities, given a
+set budget and deadline.
 
@@
 
+\begin{figure*}[tb]
+    \centering
+    \subfloat[Ratio of Actual Cost to Budget]{
+        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-cost-all}
+    }
+    \hspace{2cm}
+    \subfloat[Ratio of Actual Makespan to Deadline]{
+        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all}
+    }
+    \caption{Boxplots for budget and deadline ratios when runtime estimate 
+    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. Values 
+    greater than 1 indicate that the budget/deadline constraint was exceeded.}
+    \label{fig:variances}
+\end{figure*}
+
 Both of the workflow-aware algorithms rely on estimates of task runtimes to
@@
 
-\TODO{Update the description of the experiment in this paragraph if necessary}
-
-This experiment used a single application, Montage, and a single ensemble
-type, uniform unsorted. This application and ensemble type was chosen
-arbitrarily for simplicity. It is expected that the application and ensemble
-type have little impact on how well the algorithms are able to stay within
-constraints.
-
 The actual runtime of each task is adjusted in the simulation by adding a
@@
 
-Fig.~\ref{fig:variances} shows the results of simulations for estimate
-errors ranging from $0\%$ to $50\%$. These simulations were done for 10
-ensembles of 50 workflows using 10 budgets and 10 deadlines for each error
-value and algorithm (3000 simulations per error value). The figure shows box
-plots of the ratio of actual cost to budget and actual makespan to deadline.
-The actual cost and makespan are the simulated cost and makespan of the
-ensemble. The ratio of actual values to constraints indicates whether the
-actual value exeeded the constraint: Values less than 1 indicate that the
-constraint was not exceeded, and values greater than 1 indicate that the
-constraint was exceeded.
+Fig.~\ref{fig:variances} shows the results of simulations for estimate errors
+ranging from $0\%$ to $50\%$. The figure summarizes the results of an
+extensive suite of 525,000 simulations (10 ensembles of 50 workflows x 5
+applications x 5 distributions x 10 budgets x 10 deadlines x 7 error values x
+3 algorithms). The figure shows box plots of the ratio of actual cost to
+budget and actual makespan to deadline. The actual cost and makespan are the
+simulated cost and makespan of the ensemble. The ratio of actual values to
+constraints indicates whether the actual value exeeded the constraint: Values
+greater than 1 indicate that the constraint was exceeded.
 
 Fig.~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
-figure illustrates two important algorithm characteristics. First, the dynamic
+plot illustrates two important algorithm characteristics. First, the dynamic
 algorithms very rarely exceeded the budget, even with very large errors. This
@@
 Fig.~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
-interesting thing to notice about this figure is that, for all three
+interesting thing to notice about this plot is that, for all three
 algorithms, the deadline constraint is rarely exceeded, even in cases with
@@
 
-\begin{figure*}[tb]
-    \centering
-    \subfloat[Ratio of Actual Cost to Budget]{
-        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-cost-all}
-    }
-    \hspace{2cm}
-    \subfloat[Ratio of Actual Makespan to Deadline]{
-        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all}
-    }
-    \caption{Boxplots for budget and deadline ratios when runtime estimate 
-    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. Values 
-    greater than 1 indicate that the budget/deadline constraint was exceeded.}
-    \label{fig:variances}
-\end{figure*}
-
 \subsection{Provisioning Delays}
@@
 
-\TODO{Update the following paragraph depending on what results Maciek included}
-
 Fig.~\ref{fig:delays} shows the ratios of actual values to constraints when
-the provisioning delay is increased from 0 seconds up to 15 minutes. These
-simulations were done using 10 uniform unsorted ensembles of 50 workflows from
-the Montage application, using 10 deadlines (1000 simulations per algorithm
-and provisioning delay). As in Fig.~\ref{fig:variances}, the y-axis in each
-plot represents the ratio of the actual simulated value to the constraint
-value for the budget constraint (actual cost) or the deadline constraint
-(actual makespan). These ratios indicate how frequently and by how much the
-actual simulated values exceeded (or did not exceed) the constraints.
+the provisioning delay is increased from 0 seconds up to 15 minutes. The
+figure summarizes a suite of 105,000 simulations (10 ensembles of 50 workflows
+x 5 distributions x 10 budgets x 10 deadlines x 7 error values x 3
+algorithms). Only one application, Montage, was used in this experiment. As in
+Fig.~\ref{fig:variances}, the y-axis in each plot represents the ratio of the
+actual simulated value to the constraint value for the budget constraint
+(actual cost) or the deadline constraint (actual makespan). These ratios
+indicate how frequently and by how much the actual simulated values exceeded
+(or did not exceed) the constraints.
 
@@
 
-\TODO{Update this next paragraph depending on what Maciek included in the figure. The results seem to have changed.}
-
-The effect of provisioning delays on workflow performance is similar to that of inaccurate runtime estimates: when the delays are small, all algorithms are able to produce results within the constraints, but when the delays are larger, the dynamic algorithms are able to adapt to avoid exceeding the constraints while the static algorithm is not.
-The explanation for this behavior is also similar.
-The difference is that, a very small error in runtime estimates has little impact on the budget, while a very small provisioning delay has a large impact on the budget.
-Fig.~\ref{fig:delays}.a shows that even for a very short provisioning delay of 30 seconds (usually it takes more than than 30 seconds just to boot a VM), the SPSS algorithm exceeds the budget by as much as a factor of 2, and has a distribution with an upper quartile of 1.5 times the budget.
-In the case of delays more than 5 minutes, which is more typical of what has been observed on academic infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, fully three quarters of the simulations exceeded the budget.
-With the exception of a few outliers for 10 and 15 minute delays, the dynamic algorithms consistently met the budget and deadline constraints.
+The effect of provisioning delays on workflow performance is similar to that
+of inaccurate runtime estimates: when the delays are small, all algorithms are
+able to produce results within the constraints, but when the delays are
+larger, the dynamic algorithms are able to adapt to avoid exceeding the
+constraints while the static algorithm is not. In the case of delays of more
+than 5 minutes, which is typical of what has been observed on academic
+infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and
+FutureGrid~\cite{FutureGrid}, approximately half of the SPSS simulations
+exceeded the budget while none of the simulations using a dynamic algorithm
+exceeded the budget.
 
@@
 
-\TODO{Update this paragraph when Maciek explains what was included in the plot.}
-
-This behavior suggests that SPSS is far too sensitive to provisioning delays in its current form to be of practical use in real systems.
-It is likely, as was the case for inaccurate runtime estimates, that modifying the SPSS algorithm would improve its performance for provisioning delays.
-In fact, since provisioning operations are infrequent (because all the algorithms tend to provision resources for a long time), it is likely that the performance of SPSS could be improved significantly by simply adding an estimate of the provisioning delay to its scheduling function.
-Such an estimate may not have to be particularly accurate to get good results, and developing an estimate from historical data should be relatively simple.
-Testing this idea is left for future work.
+This experiment suggests that SPSS is far too sensitive to provisioning delays
+in its current form to be of practical use in real systems. It is possible
+that modifying the SPSS algorithm to account for provisioning delays would
+improve its performance on this experiment. In fact, since provisioning
+operations are infrequent (because all the algorithms tend to provision
+resources for a long time), it is likely that the performance of SPSS could be
+improved significantly by simply adding an estimate of the provisioning delay
+to its scheduling function. Such an estimate may not have to be particularly
+accurate to get good results, and developing an estimate from historical data
+should be relatively simple. Testing this idea is left for future work.
 
@@
 Running workflows consisting of large numbers of tasks on distributed systems
-often results in failures. The goal of the next experiment was
-to assess the behavior of the algorithms in the presence of task execution
-failures by introducing a failure model into the simulator. The model is
-characterized by a failure rate $f$ that defines the probability that a task
-will fail. The failure time of the task is determined by randomly sampling a
-value between the task start time and finish time. If the task fails, it is reported to
-the workflow engine, which retries the task until it succeeds. The dynamic algorithms
-re-add the task to the priority queue (queue $P$ in Algorithm~\ref{alg:ds})
-so that it can be resubmitted by the scheduler. The SPSS algorithm immediately 
-resubmits the failed task to the same VM that was selected in the plan.
+often results in failures. The goal of the next experiment was to assess the
+behavior of the algorithms in the presence of task execution failures by
+introducing a failure model into the simulator. The model is characterized by
+a failure rate $f$ that defines the probability that a task will fail. The
+failure time of the task is determined by randomly sampling a value between
+the task start time and finish time. If the task fails, it is reported to the
+workflow engine, which retries the task until it succeeds. The dynamic
+algorithms re-add the task to the priority queue (queue $P$ in
+Algorithm~\ref{alg:ds}) so that it can be resubmitted by the scheduler. The
+SPSS algorithm immediately  resubmits the failed task to the same VM that was
+selected in the plan.
 
@@
 
-Unlike the dynamic algorithms, which execute their scheduling and provisioning logic at runtime, the SPSS algorithm plans out its provisioning and scheduling decisions ahead of workflow execution.
-Because SPSS involves more complicated logic than the dynamic algorithms, it is important to understand what impact this has on execution time.
-
-Fig.~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles of 100 workflows with five different workflow sizes: 50, 200, 400, 600, and 800 tasks.
-The ensembles were generated using a constant distribution equal to the workflow size desired.
-Two different applications were used: SIPHT and CyberShake.
-Each box summarizes the results of 2 applications, 10 random seeds, 10 budgets and 10 deadlines, or 2000 simulations per box.
-
-Fig.~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS planning time is reasonable, taking on the order of tens of seconds to a few minutes.
-For larger ensembles of large workflows, however, the SPSS planning time can easily reach 10 minutes.
-Considering that the largest workflows used in this experiment are still relatively small (maximum of 800 tasks), and that real workflows are often much larger (workflows with tens of thousands of tasks are common, and workflows with hundreds of thousands or millions of tasks are not impossible), it is unlikely that SPSS will be practical for ensembles of very large workflows.
+Unlike the dynamic algorithms, which execute their scheduling and provisioning
+logic at runtime, the SPSS algorithm plans out its provisioning and scheduling
+decisions ahead of workflow execution. Because SPSS involves more complicated
+logic than the dynamic algorithms, it is important to understand what impact
+this has on execution time.
+
+Fig.~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles
+of 100 workflows with five different workflow sizes: 50, 200, 400, 600, and
+800 tasks. The ensembles were generated using a constant distribution equal to
+the workflow size desired. Two different applications were used: SIPHT and
+CyberShake. Each box summarizes the results of 2 applications, 10 random
+seeds, 10 budgets and 10 deadlines, or 2000 simulations per box.
+
+Fig.~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS
+planning time is reasonable, taking on the order of tens of seconds to a few
+minutes. For larger ensembles of large workflows, however, the SPSS planning
+time can easily reach 10 minutes. Considering that the largest workflows used
+in this experiment are still relatively small (maximum of 800 tasks), and that
+real workflows are often much larger (workflows with tens of thousands of
+tasks are common, and workflows with hundreds of thousands or millions of
+tasks are not impossible), it is unlikely that SPSS will be practical for
+ensembles of very large workflows.
 
@@
 
-The algorithmic complexity of SPSS is $O(n^2)$, where $n$ is the number of tasks in the ensemble.
-This derives from the fact that, for each task, SPSS considers scheduling the task on the cheapest available slot, which involves scanning all of the available slots on all the VMs.
-Since the number of available slots, in the worst case, is proportional to the number of tasks scheduled (because scheduling a task splits an existing slot into at most two slots), the complexity of SPSS is $O(n^2)$.
-In comparison, the dynamic algorithms all have a more scalable complexity of $O(n)$.
-DPDS only examines the tasks in the workflow once when they are scheduled, and WA-DPDS examines them twice: once in the admissions algorithm, and once when they are scheduled.
-This makes the dynamic algorithms a better fit for larger workflows and ensembles even though they may not produce results that are as good as SPSS.
-
-It may be possible to optimize SPSS to reduce its runtime by, for example, clustering the workflow to increase task granularity, which would decrease the ratio of planning time to ensemble makespan.
-It may also be possible to reduce the complexity of SPSS by employing more sophisticated data structures to store the available slots.
-Investigating these topics is left for future work.
+The algorithmic complexity of SPSS is $O(n^2)$, where $n$ is the number of
+tasks in the ensemble. This derives from the fact that, for each task, SPSS
+considers scheduling the task on the cheapest available slot, which involves
+scanning all of the available slots on all the VMs. Since the number of
+available slots, in the worst case, is proportional to the number of tasks
+scheduled (because scheduling a task splits an existing slot into at most two
+slots), the complexity of SPSS is $O(n^2)$. In comparison, the dynamic
+algorithms all have a more scalable complexity of $O(n)$. DPDS only examines
+the tasks in the workflow once when they are scheduled, and WA-DPDS examines
+them twice: once in the admissions algorithm, and once when they are
+scheduled. This makes the dynamic algorithms a better fit for larger workflows
+and ensembles even though they may not produce results that are as good as
+SPSS.
+
+It may be possible to optimize SPSS to reduce its runtime by, for example,
+clustering the workflow to increase task granularity, which would decrease the
+ratio of planning time to ensemble makespan. It may also be possible to reduce
+the complexity of SPSS by employing more sophisticated data structures to
+store the available slots. Investigating these topics is left for future work.
 
@@
 
-In this paper we addressed the interesting and important new problem of scheduling and resource provisioning for scientific workflow ensembles on IaaS clouds.
-The goal of this work is to maximize the number of prioritized workflows that can be completed given budget and deadline constraints.
-
-%This problem differs from previous work on grid and utility grid scheduling in that cloud infrastructures provide more control over provisioning, so that the number of resources can be adjusted according to the requirements of the application.
-Therefore the problem space becomes larger; it requires not only an efficient mapping of tasks to available resources, but also the selection of the best resource provisioning plan.
-
-%Formulating the problem as a maximization of the number of prioritized workflows completed from the ensemble is also novel and requires workflows to be admitted or rejected based on their estimated resource demands.
-We believe that this bi-constrained problem is highly relevant because such constraints are commonly imposed on many real-world projects.
-The approach is also directly applicable to grid environments that provide resource reservations and charge service units for resource use.
-
-We developed three algorithms to solve this problem: two dynamic algorithms, DPDS and WA-DPDS, and one static algorithm, SPSS.
-The algorithms were evaluated via simulation on ensembles of synthetic workflows, which were generated based on statistics from real scientific applications.
-
-The results of our simulation studies indicate that the two algorithms that take into account the structure of the workflow and task runtime estimates (WA-DPDS and SPSS) yield better results than the simple priority-based scheduling strategy (DPDS), which makes provisioning decisions based purely on resource utilization.
-
-In cases where there are no provisioning delays, task runtime estimates are good, and failures are rare, we found that the static algorithm, SPSS, performs significantly better than both dynamic algorithms.
-However, when conditions are less than perfect, the static plans produced by SPSS are disrupted and it frequently exceeds the budget and deadline constraints.
-In comparison, the dynamic algorithms are able to adapt to a wide variety of conditions, and rarely exceed the constraints even with long delays, poor estimates, and high failure rates.
-
-%In addition, we found that SPSS tends to perform better on coarse-grained workflows than the dynamic algorithms.
-For wide, fine-grained workflows, such as CyberShake and Montage, however, the dynamic algorithms frequently produce performance that is as good or better than SPSS, because they are better able to pack the smaller tasks onto idle VMs close to the deadline than SPSS, which distributes deadlines to tasks in a way that prevents them from starting late.
-
-%For very large workflows and ensembles, we found that the planning time of the SPSS algorithm is prohibitive.
-SPSS often took 10 minutes or more to plan ensembles of 100 workflows with 800 tasks each.
-This suggests that ensembles of workflows with tens of thousands of tasks, which are commonly encountered in real workflow applications, would take many hours to plan using SPSS.
-This makes the dynamic algorithms much more attractive for large scale problems.
-
-This study suggests several areas for future work.
-Our current approach models data access as part of the task execution time and does not consider data storage and transfer costs.
-In the future we plan to extend the application and infrastructure model to  include the various data storage options available on clouds.
-A previous experimental study \cite{Juve2010} suggests that the data demands of scientific workflows have a large impact on not only the execution time, but also on the cost of workflows in commercial clouds.
-We also plan to investigate heterogeneous environments that include multiple VM types and cloud providers, including private and community clouds, which will make the problem even more complex and challenging.
-
-%The results of this study can be applied to develop tools that assist researchers in planning their large-scale computational experiments.
-The estimates of cost, runtime, and number of workflows completed that can be obtained from both the static algorithms and from the simulation runs, constitute valuable hints for planning ensembles and evaluating the associated trade-offs.
+In this paper we addressed the interesting and important new problem of
+scheduling and resource provisioning for scientific workflow ensembles on IaaS
+clouds. The goal of this work is to maximize the number of prioritized
+workflows that can be completed given budget and deadline constraints.
+
+%This problem differs from previous work on grid and utility grid scheduling
+%in that cloud infrastructures provide more control over provisioning, so that
+%the number of resources can be adjusted according to the requirements of the
+%application. Therefore the problem space becomes larger; it requires not only
+%an efficient mapping of tasks to available resources, but also the selection
+%of the best resource provisioning plan.
+
+%Formulating the problem as a maximization of the number of prioritized
+%workflows completed from the ensemble is also novel and requires workflows to
+%be admitted or rejected based on their estimated resource demands. We believe
+%that this bi-constrained problem is highly relevant because such constraints
+%are commonly imposed on many real-world projects. The approach is also
+%directly applicable to grid environments that provide resource reservations
+%and charge service units for resource use.
+
+We developed three algorithms to solve this problem: two dynamic algorithms,
+DPDS and WA-DPDS, and one static algorithm, SPSS. The algorithms were
+evaluated via simulation on ensembles of synthetic workflows, which were
+generated based on statistics from real scientific applications.
+
+The results of our simulation studies indicate that the two algorithms that
+take into account the structure of the workflow and task runtime estimates
+(WA-DPDS and SPSS) yield better results than the simple priority-based
+scheduling strategy (DPDS), which makes provisioning decisions based purely on
+resource utilization.
+
+In cases where there are no provisioning delays, task runtime estimates are
+good, and failures are rare, we found that the static algorithm, SPSS,
+performs significantly better than both dynamic algorithms. However, when
+conditions are less than perfect, the static plans produced by SPSS are
+disrupted and it frequently exceeds the budget and deadline constraints. In
+comparison, the dynamic algorithms are able to adapt to a wide variety of
+conditions, and rarely exceed the constraints even with long delays, poor
+estimates, and high failure rates.
+
+%In addition, we found that SPSS tends to perform better on coarse-grained
+%workflows than the dynamic algorithms. For wide, fine-grained workflows, such
+%as CyberShake and Montage, however, the dynamic algorithms frequently produce
+%performance that is as good or better than SPSS, because they are better able
+%to pack the smaller tasks onto idle VMs close to the deadline than SPSS,
+%which distributes deadlines to tasks in a way that prevents them from
+%starting late.
+
+%For very large workflows and ensembles, we found that the planning time of
+%the SPSS algorithm is prohibitive. SPSS often took 10 minutes or more to plan
+%ensembles of 100 workflows with 800 tasks each. This suggests that ensembles
+%of workflows with tens of thousands of tasks, which are commonly encountered
+%in real workflow applications, would take many hours to plan using SPSS. This
+%makes the dynamic algorithms much more attractive for large scale problems.
+
+This study suggests several areas for future work. Our current approach models
+data access as part of the task execution time and does not consider data
+storage and transfer costs. In the future we plan to extend the application
+and infrastructure model to  include the various data storage options
+available on clouds. A previous experimental study \cite{Juve2010} suggests
+that the data demands of scientific workflows have a large impact on not only
+the execution time, but also on the cost of workflows in commercial clouds. We
+also plan to investigate heterogeneous environments that include multiple VM
+types and cloud providers, including private and community clouds, which will
+make the problem even more complex and challenging.
+
+%The results of this study can be applied to develop tools that assist
+%researchers in planning their large-scale computational experiments. The
+%estimates of cost, runtime, and number of workflows completed that can be
+%obtained from both the static algorithms and from the simulation runs,
+%constitute valuable hints for planning ensembles and evaluating the
+%associated trade-offs.
 
@@
 % This balances the columns by inserting a column break at the given reference
-\IEEEtriggeratref{19}
+\IEEEtriggeratref{16}
 

### 2012-05-02 paper/paper.tex
Add Ewa's changes

@@
 
-Large-scale computational applications expressed as scientific workflows are
+Large-scale applications expressed as scientific workflows are
 often grouped into ensembles of inter-related workflows. In this paper, we
@@
 on static and dynamic strategies for both task scheduling and resource
-provisioning. We evaluate the algorithms via simulation using a set of
+provisioning. We perform the evaluation via simulation using a set of
 scientific workflow ensembles with a broad range of budget and deadline
 parameters, taking into account uncertainties in task runtime estimations,
-provisioning delays and failures. We find that the key factor determining the
+provisioning delays, and failures. We find that the key factor determining the
 performance of an algorithm is its ability to decide which workflows in an
@@
 
-Scientific workflows, usually represented as directed acyclic graphs (DAGs), are
+Scientific workflows, usually represented as Directed Acyclic Graphs (DAGs), are
 an important class of applications that lead to challenging problems
@@
 for large computational problems are often composed of several
-inter-related workflows grouped into {\em ensembles}. All the workflows in an
+inter-related workflows grouped into {\em ensembles}.  Workflows in an
 ensemble typically have a similar structure, but they differ in their input
@@
 
-The CyberShake~\cite{Callaghan2011} is a good example of an application that
+CyberShake~\cite{Callaghan2011} is a good example of an application that
 requires scientific workflow ensembles. Generating a hazard map for a region
 such as Southern California requires an ensemble of many related workflows to
-compute individual hazard curves. E.g. the hazard map generated by the
-CyberShake group in 2009 required an ensemble of 239 workflows. In astronomy,
-for example, a user might wish to use multiple Montage~\cite{Deelman2008}
+compute individual hazard curves. E.g. the hazard map generated in 2009 required an ensemble of 239 workflows. In astronomy,
+a user might wish to use multiple Montage~\cite{Deelman2008}
 workflows with different parameters to generate a set of image mosaics that
 cover a given area of the sky in different wavelengths, or use an ensemble to generate
-the parts of a very large mosaic. The Pegasus team is now working on a Galactic
-Plane project that is focused on generating a mosaic of the entire sky. This
-ensemble consists of 17 workflows, each of which contains 900 sub-workflows and
+the parts of a very large mosaic. The
+ensemble for generating a mosaic of the entire sky consists of 17 workflows, each of which contains 900 sub-workflows and
 computes one ``tile'' of the full mosaic. Another ensemble example is the
@@
 by different workflows. Additional workflows are created to run the analysis
-using different parameters. A recent analysis of data from the Kepler satellite
+using different parameters. A recent analysis of Kepler satellite data
 required three ensembles of 15 workflows.
@@
 Infrastructure-as-a-Service (IaaS) clouds offer the ability to provision 
-computational resources on-demand according to a pay-per-use model. These systems 
+resources on-demand according to a pay-per-use model. These systems 
 are regarded by the scientific community as a potentially attractive source of 
@@
 more control also requires the development of new methods for task 
-scheduling and resource provisioning. The resource management decisions required 
-in cloud scenarios not only have to take into account traditional performance-related 
-metrics such as workflow makespan or resource utilization, but must also now consider
-budget constraints, since the resources from public (commercial) clouds
+scheduling and resource provisioning. Resource management decisions required 
+in cloud scenarios not only have to take into account performance-related 
+metrics such as workflow makespan or resource utilization, but must also consider
+budget constraints, since the resources from commercia clouds
 usually have monetary costs associated with them~\cite{Durkee2010}.
@@
 In this paper, we aim to gain insight into resource management challenges when
-executing scientific workflow ensembles on clouds. In particular, we address a
+executing scientific workflow ensembles on clouds. We address a
 new and important problem of maximizing the number of completed workflows from
@@
   taking into account uncertainties in task runtime
-  estimations, provisioning delays and failures, 
+  estimations, provisioning delays, and failures, 
   \item we discuss the performance of the algorithms on a set of
   synthetic workflow ensembles based on important, real scientific
-  applications, on a broad range of different application scenarios and
+  applications, on a broad range of different application scenarios, and
   varying constraint values.
@@
 
-The paper is organized as follows: after an analysis of related work in
-Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
-infrastructure and application model we are targeting in this paper. In
-Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
-developed. Section~\ref{sec:performance} presents test scenarios and performance
-metrics, while Section~\ref{sec:results} discusses the results of simulation
-studies of these algorithms. Finally, general conclusions, lessons learned, and
-future work are outlined in Section~\ref{sec:conclusions}.
+% Ewa: the following paragraph can be deleted if necessary
+%The paper is organized as follows: after an analysis of related work in
+%Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
+%infrastructure and application model we are targeting in this paper. In
+%Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
+%developed. Section~\ref{sec:performance} presents test scenarios and performance
+%metrics, while Section~\ref{sec:results} discusses the results of simulation
+%studies of these algorithms. Finally, general conclusions, lessons learned, and
+%future work are outlined in Section~\ref{sec:conclusions}.
 
@@
 of these autoscaling systems, but no support for complex applications is provided.
-
 Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
@@
 
-\subsection{Resource Model}
-
-We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2),
+\underline{Resource Model}  We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2),
 where virtual machine (VM) instances are requested on-demand and are billed by
@@
 
-\subsection{Application Model}
-
-The target applications are ensembles of scientific workflows that can be
-modeled as Directed Acyclic Graphs (DAGs), where the nodes in the graph
+\underline{Application Model} The target applications are ensembles of scientific workflows that can be
+modeled as DAGs, where the nodes in the graph
 represent computational tasks, and the edges represent data- or control-flow
-dependencies between the tasks. We assume that there are runtime estimates for
-each task in the workflow based on either a performance model of the
-application, or historical data that can be mined. Additionally, we assume that
-the runtime estimates for individual workflow tasks are not perfect, but may
+dependencies between the tasks. We assume that runtime estimates for the workflow tasks
+ are know and  that they are not perfect, but may
 vary based on a uniform distribution of $\pm p\%$.
@@
 identification protocol using high-throughput technology), astronomy (Montage),
-earthquake science (CyberShake), and physics (LIGO). The runtimes used to
-generate the synthetic workflows are based on distributions gathered from
-running real workflows. The synthetic workflows were generated using code
-developed by Bharathi, et al~\cite{WorkflowGenerator}.
-
-Although scientific workflows are often data-intensive, the algorithms described
-in the next section do not currently consider the size of input and output data
-when scheduling tasks. Instead it is assumed that all workflow data is stored in
+earthquake science (CyberShake), and physics (LIGO).  The synthetic workflows were generated using code
+developed in~\cite{WorkflowGenerator},  with tasks runtimes based on distributions gathered from
+running real workflows.
+
+Although workflows are often data-intensive, the algorithms described
+here do not currently consider the size of input and output data
+when scheduling tasks. Instead, it is assumed that all workflow data is stored in
 a shared cloud storage system, such as Amazon S3, and that intermediate data
@@
 data transfer times between the shared storage and the VMs are equal for
-different VMs so that task placement decisions do not have an impact on the
+different VMs so that task placement decisions do not impact the
 runtime of the tasks.
 
-A workflow ensemble consists of several related workflows that a user wishes to
-run. Each workflow in an ensemble is given a numeric priority that indicates how
+We assume that each workflow in an ensemble is given a numeric priority that indicates how
 important the workflow is to the user. As such, the priorities indicate the
 utility function of the user. These priorities are absolute in the sense that
-completing a workflow with a given priority is more valuable (gives the user
-more utility) than completing all other workflows in the ensemble with lower
+completing a workflow with a given priority is more valuable than completing all other workflows in the ensemble with lower
 priorities combined. The goal of the workflow ensemble scheduling and cloud
@@
 
-This section describes three algorithms were developed to schedule and provision resources for ensembles of workflows on the cloud under budget and deadline constraints.
+This section describes three algorithms developed to schedule and provision resources for ensembles of workflows on the cloud under budget and deadline constraints.
 
@@
 %
-The DPDS algorithm provisions $N_{VM}$ VMs at the start of the ensemble execution.
-
-The DPDS algorithm periodically computes resource
+DPDS provisions $N_{VM}$ VMs at the start of the ensemble execution, then
+it periodically computes resource
 utilization using the percentage of idle VMs over time, and adjusts the number
@@
 
-The DPDS algorithm guarantees that tasks from lower priority workflows are
+DPDS guarantees that tasks from lower priority workflows are
 always deferred when higher-priority tasks are available, but lower-priority
 tasks can still occupy idle VMs when higher-priority tasks are not available.
-However, because there is no preemption, long-running low-priority tasks may
+Since there is no preemption, long-running low-priority tasks may
 delay the execution of higher-priority tasks. In addition, tasks from low
@@
 
-The DPDS algorithm does not use any information about the structure of the
-workflows in the ensemble when scheduling tasks: it looks only at the priorities
-of the ready tasks when deciding which task to schedule next. It does not
+DPDS does not use any information about the structure of the
+workflows in the ensemble when scheduling tasks.  It does not
 consider whether a lower priority task belongs to a workflow that will never be
@@
 
-In order to address this issue, the Workflow-Aware DPDS (WA-DPDS) algorithm
-extends DPDS by introducing a workflow admission procedure. The admission
-procedure is invoked whenever WA-DPDS sees the first task of a new workflow at
+The Workflow-Aware DPDS (WA-DPDS) algorithm
+extends DPDS by introducing a workflow admission procedure, which is invoked whenever WA-DPDS sees the first task of a new workflow at
 the head of the priority queue (i.e. when no other tasks from the workflow have
@@
 to admit the new workflow; if there is not, then the workflow is rejected and
-its tasks are removed from the queue. This algorithm compares the current cost
+its tasks are removed from the queue. WA-DPDS compares the current cost
 (consumed budget) and remaining budget, taking into account the cost of
@@
 admitted. In addition, it adds a small safety margin of \$0.10 to avoid going
-over the budget.
+over budget. We found the admission procedure useful
+not only to prevent low-priority workflows from delaying high-priority ones, but
+also to reject large and costly workflows that would overrun the budget and
+admit smaller workflows that can efficiently utilize idle resources.
+% in
+%ensembles containing workflows of non-uniform sizes. It would also be possible
+%to extend this admission procedure to check other constraints, such as whether
+%the estimated critical path of the new workflow exceeds the time remaining
+%until the deadline.
 
@@
 
-This admission procedure relies only on the total estimated resource consumption
-and compares it to the remaining budget. We found that this estimation is useful
-not only to prevent low-priority workflows from delaying high-priority ones, but
-also to reject large and costly workflows that would overrun the budget and
-admit smaller workflows that can efficiently utilize idle resources.
-% in
-%ensembles containing workflows of non-uniform sizes. It would also be possible
-%to extend this admission procedure to check other constraints, such as whether
-%the estimated critical path of the new workflow exceeds the time remaining
-%until the deadline.
-
 \subsection{Static Provisioning Static Scheduling (SPSS)}
@@
 The previous dynamic (online) algorithms make provisioning and scheduling
-decisions at runtime. In comparison, the SPSS algorithm creates a provisioning
+decisions at runtime. By contrast, the SPSS algorithm creates a provisioning
 and scheduling plan before running any workflow tasks. This enables SPSS to
@@
 
-The approach used by SPSS is to plan each workflow in the ensemble in priority
-order, rejecting any workflows that would exceed the dedline or budget. Once the
-plan is complete, the VMs are provisioned and tasks are executed according to
-the schedules given by the plan.
+%The approach used by SPSS is to plan each workflow in the ensemble in priority
+%order, rejecting any workflows that would exceed the dedline or budget. Once the
+%plan is complete, the VMs are provisioned and tasks are executed according to
+%the schedules given by the plan.
 
@@
 and still be completed by the ensemble deadline. For a workflow $w$, the slack
-time of $w$ is:
-$$ ST(w) = d - CP(w) $$
+time of $w$ is: $ ST(w) = d - CP(w) $
 where $d$ is the deadline and $CP(w)$ is the critical path of $w$. We
@@
   \subfloat[DPDS]{
-  \includegraphics[width=0.45\textwidth]{figures/spds-gantt}
+  \includegraphics[width=0.40\textwidth]{figures/spds-gantt}
   }\\
   \subfloat[SPSS]{
-  \includegraphics[width=0.45\textwidth]{figures/spss-gantt}
+  \includegraphics[width=0.40\textwidth]{figures/spss-gantt}
   }
@@
 randomized ensembles using workflows available from the workflow generator
-gallery published online by Bharathi, et al~\cite{WorkflowGenerator}. The
+gallery~\cite{WorkflowGenerator}. 
+%% Ewa: maybe put url instead of citation
+The
 gallery contains synthetic workflows modeled using structures and parameters
@@
 workflows from five real applications: SIPHT, LIGO, Epigenomics, Montage and
-CyberShake. For each of these applications, workflows with 50, 100, 200, 300,
-400, 500, 600, 700, 800, 900 and 1000 tasks were generated. For each workflow
+CyberShake. For each application, workflows with 50, 100, 200, 300,
+400, 500, 600, 700, 800, 900 and 1000 tasks were created. For each workflow
 size, 20 different workflow instances were generated using parameters and task
-runtime distributions from traces of real workflows. The total collection of
+runtime distributions from real workflow traces. The total collection of
 synthetic workflows contains 5 applications, 11 different workflow sizes, and 20
@@
 but we assume that ensemble sizes are on the order of between 10 and 100
-workflows. This range is motivated by several reasons: First, such sizes are
-typical of the real applications we have examined (see Section~\ref{sec:intro}). 
+workflows, typical of the real applications we have examined (see Section~\ref{sec:intro}). 
 %For example, the number of
@@
 %the past has been on the order of 100. 
-Second, smaller ensembles consisting of
-just a few workflows can be aggregated into a single workflow, so there is no need to
-treat them as an ensemble. Similarly, when the number of workflows grows, and
-each workflow has a large number of tasks, either the deadline and budget
-constraints are low enough to prevent many of the workflows from running, or the
-problem of efficiently allocating them to the resources becomes similar to a
-bag-of-tasks problem, which is easier to solve efficiently.
+%Second, smaller ensembles consisting of
+%just a few workflows can be aggregated into a single workflow, so there is no need to
+%treat them as an ensemble. Similarly, when the number of workflows grows, and
+%each workflow has a large number of tasks, either the deadline and budget
+%constraints are low enough to prevent many of the workflows from running, or the
+%problem of efficiently allocating them to the resources becomes similar to a
+%bag-of-tasks problem, which is easier to solve efficiently.
 
@@
 number of low-priority workflows by executing the smallest workflows first.
-
-In order to account for the user's priority, we used an exponential scoring
-where the score for an ensemble $e$ is:
+In order to account for the user's priority, we used an exponential scoring defined as:
 %
@@
 %
-This scoring is consistent with our definition of the problem, which is to
-complete as many workflows as possible, according to their priorities, given a
-set budget and deadline.
+%This scoring is consistent with our definition of the problem, which is to
+%complete as many workflows as possible, according to their priorities, given a
+%set budget and deadline.
 
@@
 The workflows for both applications are very wide relative to their height,
-and both have very short-running tasks. As a result of these two
-characteristics, the Montage and CyberShake applications have very short
-critical paths and look more like bag-of-tasks applications, which are easier
+and both have very short-running tasks, resulting  in relatively short
+critical paths that makes them look more like bag-of-tasks applications, which are easier
 to execute than more structured applications. DPDS and WA-DPDS are able to
@@
 
-\subsection{Task Granularity}
-\TODO{Should we remove this section?}
-
-In the last experiment we noted that, for Montage and CyberShake, the short
-runtimes of their tasks made the dynamic algorithms perform better relative to
-SPSS. In order to test this theory we adjusted the granularity of the tasks in
-several Montage and CyberShake ensembles to see how this would affect the
-relative performance of the algorithms. The granularity adjustment was
-achieved by multiplying the runtime of each task by a fixed scaling factor.
-
-Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
-as the scaling factor is increased from 1 to 16. The figure only shows results
-for the CyberShake application with pareto unsorted ensembles, but similar
-results were found for other ensemble types for both Montage and CyberShake.
-Each data point in the figure represents 500 simulations (5 ensembles with 10
-budgets and 10 deadlines). The best scores percent was calculated the same way
-as it was for Fig.~\ref{fig:distributions}.
-
-The figure shows that, as the scaling factor increases beyond 2, the relative
-performance of SPSS surpasses that of the dynamic algorithms. Similar results
-were found for other ensemble types for both applications. This result
-suggests that, in general, for fine-grained workflows the dynamic algorithms
-will produce better scores, and for coarse-grained workflows SPSS will produce
-better scores.
-
-\begin{figure}[tb]
-    \centering
-    \includegraphics[width=0.3\textwidth]{stretching_cybershake}
-    \caption{Percentage of high scores achieved by each algorithm for pareto 
-    unsorted CyberShake ensembles when task runtime is scaled by a constant 
-    factor. A scaling factor of $x$ means that the runtime of tasks in each 
-    workflow was multiplied by $x$.}
-    \label{fig:stretching}
-\end{figure}
+%\subsection{Task Granularity}
+%\TODO{Should we remove this section?}
+%
+%In the last experiment we noted that, for Montage and CyberShake, the short
+%runtimes of their tasks made the dynamic algorithms perform better relative to
+%SPSS. In order to test this theory we adjusted the granularity of the tasks in
+%several Montage and CyberShake ensembles to see how this would affect the
+%relative performance of the algorithms. The granularity adjustment was
+%achieved by multiplying the runtime of each task by a fixed scaling factor.
+
+%Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
+%as the scaling factor is increased from 1 to 16. The figure only shows results
+%for the CyberShake application with pareto unsorted ensembles, but similar
+%results were found for other ensemble types for both Montage and CyberShake.
+%Each data point in the figure represents 500 simulations (5 ensembles with 10
+%budgets and 10 deadlines). The best scores percent was calculated the same way
+%as it was for Fig.~\ref{fig:distributions}.
+
+%The figure shows that, as the scaling factor increases beyond 2, the relative
+%performance of SPSS surpasses that of the dynamic algorithms. Similar results
+%were found for other ensemble types for both applications. This result
+%suggests that, in general, for fine-grained workflows the dynamic algorithms
+%will produce better scores, and for coarse-grained workflows SPSS will produce
+%better scores.
+
+%\begin{figure}[tb]
+ %   \centering
+ %   \includegraphics[width=0.3\textwidth]{stretching_cybershake}
+ %   \caption{Percentage of high scores achieved by each algorithm for pareto 
+ %   unsorted CyberShake ensembles when task runtime is scaled by a constant 
+ %   factor. A scaling factor of $x$ means that the runtime of tasks in each 
+ %   workflow was multiplied by $x$.}
+%    \label{fig:stretching}
+%\end{figure}
 
@@
     \centering
-    \subfloat[Ratio of Actual Cost to Budget]{
+    \subfloat[Ratio of Simulated Cost to Budget]{
         \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-cost-all}
@@
     \hspace{2cm}
-    \subfloat[Ratio of Actual Makespan to Deadline]{
+    \subfloat[Ratio of Simulated Makespan to Deadline]{
         \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all}
@@
 How do errors in task runtime estimates impact the performance of our
-scheduling and provisioning algorithms? We created an experiment to measure
-how introducing uniform errors in these estimates affects the ability of the
-algorithms to achieve the desired budget and deadline constraints.
+scheduling and provisioning algorithms?  Thus, we introduced uniform errors in the task runtime
+and observed the behavior of the algorithms in terms of meeting  the desired budget and deadline constraints.
 
@@
 applications x 5 distributions x 10 budgets x 10 deadlines x 7 error values x
-3 algorithms). The figure shows box plots of the ratio of actual cost to
-budget and actual makespan to deadline. The actual cost and makespan are the
-simulated cost and makespan of the ensemble. The ratio of actual values to
-constraints indicates whether the actual value exeeded the constraint: Values
+3 algorithms). The figure shows box plots of the ratio of the ensemble cost to
+budget and ensemble makespan to deadline. The ratio indicates whether the value (for example the simulated cost) exeeded the constraint (the budget): Values
 greater than 1 indicate that the constraint was exceeded.
 
-Fig.~\ref{fig:variances}.a shows the ratio of actual cost to budget. This
+Fig.~\ref{fig:variances}.a shows the ratio of simulated cost to budget. This
 plot illustrates two important algorithm characteristics. First, the dynamic
@@
 large amounts in some cases. This is a result of the fact that the static
-algorithm makes all of its decisions before execution and is not able to adapt
+algorithm makes all of its decisions ahead of the execution and is not able to adapt
 to changing circumstances.
@@
 
-Fig.~\ref{fig:variances}.b shows the ratio of actual makespan to budget. The
-interesting thing to notice about this plot is that, for all three
-algorithms, the deadline constraint is rarely exceeded, even in cases with
+Fig.~\ref{fig:variances}.b shows the ratio of simulated makespan to budget. 
+Interestingly, the deadline constraint is rarely exceeded, even in cases with
 very poor quality estimates. For the dynamic algorithms this is a result of
@@
 Fig.~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
-up workflows at the beginning of the timeline. In comparison, the dynamic
-algorithms tend to spread out workflow starts over the entire duration as
-shown in Fig.~\ref{fig:algorithm-example}.a. The consequence of this
-behavior is that SPSS plans tend to use up the budget but leave plenty of time
-before the deadline. So when the runtime of the plan is increased by
+up workflows at the beginning of the timeline. 
+%In comparison, the dynamic
+%algorithms tend to spread out workflow starts over the entire duration as
+%shown in Fig.~\ref{fig:algorithm-example}.a. The consequence of this
+5behavior is that SPSS plans tend to use up the budget but leave plenty of time
+%before the deadline. 
+As a result, when the runtime of the plan is increased by
 introducing errors, the SPSS plan has some room to expand without exceeding
@@
 accuracy of the task runtimes. Investigating that optimization is left for
-future work. The goal here is to determine how well the current algorithms are
-able to stay within the constraints given inaccurate task runtime estimates.
+future work. 
+%The goal here is to determine how well the current algorithms are
+%able to stay within the constraints given inaccurate task runtime estimates.
 
@@
 becomes available to the application. Typically these provisioning delays are
-on the order of a few minutes, and are highly dependent upon the architecture
-of the cloud system and/or the size of the user's VM image~\cite{Nurmi2008b}.
+on the order of a few minutes, and are highly dependent upon the cloud architecture
+and/or the size of the VM image~\cite{Nurmi2008b}.
 
 We assume that resources are billed from the minute that they are requested
-until they are terminated, and not from the moment that they become available.
-As a result, provisioning delays have an impact on both the cost (budget
-constraint) and makespan (deadline constraint) of the ensemble.
-
-Fig.~\ref{fig:delays} shows the ratios of actual values to constraints when
+until they are terminated. 
+As a result, provisioning delays have an impact on both the cost  and makespan of the ensemble.
+
+Fig.~\ref{fig:delays} shows the ratios of simulated values to constraints when
 the provisioning delay is increased from 0 seconds up to 15 minutes. The
@@
 algorithms). Only one application, Montage, was used in this experiment. As in
-Fig.~\ref{fig:variances}, the y-axis in each plot represents the ratio of the
-actual simulated value to the constraint value for the budget constraint
-(actual cost) or the deadline constraint (actual makespan). These ratios
-indicate how frequently and by how much the actual simulated values exceeded
-(or did not exceed) the constraints.
+Fig.~\ref{fig:variances}, the y-axis in each plot represents the ratio of the simulated value.
+ to the constraint value for the budget constraint or the deadline constraint.
 
@@
     \centering
-    \subfloat[Ratio of Actual Cost to Budget]{
+    \subfloat[Ratio of Simulated Ensemble Cost to Budget]{
         \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-cost-montage}
@@
     \hspace{2cm}
-    \subfloat[Ratio of Actual Makespan to Deadline]{    
+    \subfloat[Ratio of Simulated Ensemble Makespan to Deadline]{    
         \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
@@
 SPSS algorithm immediately  resubmits the failed task to the same VM that was
-selected in the plan.
+selected in the plan (to minimize the disruption to the overall plan).
 
 \begin{figure*}[tb]
-    \subfloat[Ratio of Actual Cost to Budget]{
+    \subfloat[Ratio of Simulated Ensemble Cost to Budget]{
         \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-cost-all}
@@
     \hspace{2cm}
-    \subfloat[Ratio of Actual Makespan to Deadline]{
+    \subfloat[Ratio of Simulated Ensemble Makespan to Deadline]{
         \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-makespan-all}
@@
 
-Fig.~\ref{fig:failures} shows the ratios of actual values to constraints when
+Fig.~\ref{fig:failures} shows the ratios of simulated values to constraints when
 the failures are introduced. The high failure rates have a high impact on the
@@
 
-Unlike the dynamic algorithms, which execute their scheduling and provisioning
-logic at runtime, the SPSS algorithm plans out its provisioning and scheduling
-decisions ahead of workflow execution. Because SPSS involves more complicated
-logic than the dynamic algorithms, it is important to understand what impact
+ Because SPSS involves more complicated
+logic than the dynamic algorithms and makes its decisions ahead of the execution, it is important to understand what impact
 this has on execution time.
@@
 
-The algorithmic complexity of SPSS is $O(n^2)$, where $n$ is the number of
-tasks in the ensemble. This derives from the fact that, for each task, SPSS
-considers scheduling the task on the cheapest available slot, which involves
+%The algorithmic complexity of SPSS is $O(n^2)$, where $n$ is the number of
+%tasks in the ensemble. This derives from the fact that, for each task,
+
+SPSS
+considers scheduling each task on the cheapest available slot, which involves
 scanning all of the available slots on all the VMs. Since the number of
 available slots, in the worst case, is proportional to the number of tasks
-scheduled (because scheduling a task splits an existing slot into at most two
-slots), the complexity of SPSS is $O(n^2)$. In comparison, the dynamic
+scheduled
+%Ewa:  I don't get that
+ (because scheduling a task splits an existing slot into at most two
+slots, one before the task, and one after), the complexity of SPSS is $O(n^2)$, where $n$ is the number of
+tasks in the ensemble. In comparison, the dynamic
 algorithms all have a more scalable complexity of $O(n)$. DPDS only examines
@@
 scheduled. This makes the dynamic algorithms a better fit for larger workflows
-and ensembles even though they may not produce results that are as good as
+and ensembles even though in some cases they may not produce results that are as good as
 SPSS.
@@
 scheduling and resource provisioning for scientific workflow ensembles on IaaS
-clouds. The goal of this work is to maximize the number of prioritized
+clouds. The goal of this work is to maximize the number of user-prioritized
 workflows that can be completed given budget and deadline constraints.
@@
 The results of our simulation studies indicate that the two algorithms that
-take into account the structure of the workflow and task runtime estimates
+take into account the structure of the workflow 
+%Ewa:  so it is really taking into account the workflow structure or the ensemble structure/members?
+and the members of the ensemble (through admission procedures) and task runtime estimates
 (WA-DPDS and SPSS) yield better results than the simple priority-based
 scheduling strategy (DPDS), which makes provisioning decisions based purely on
-resource utilization.
+resource utilization.  This underscores the importance of viewing workflow ensembles as a whole rather than just individual tasks or individual workflows.
 
 In cases where there are no provisioning delays, task runtime estimates are
-good, and failures are rare, we found that the static algorithm, SPSS,
+good, and failures are rare, we found that SPSS
 performs significantly better than both dynamic algorithms. However, when

### 2012-05-02 paper/paper.tex
Gideon's and Jarek's edits

@@
 
-Scientific workflows, usually represented as Directed Acyclic Graphs (DAGs), are
-an important class of applications that lead to challenging problems
-in resource management on grid and utility computing systems. Workflows
-for large computational problems are often composed of several
-inter-related workflows grouped into {\em ensembles}.  Workflows in an
-ensemble typically have a similar structure, but they differ in their input
-data, number of tasks, and individual task sizes.
-
-CyberShake~\cite{Callaghan2011} is a good example of an application that
-requires scientific workflow ensembles. Generating a hazard map for a region
-such as Southern California requires an ensemble of many related workflows to
-compute individual hazard curves. E.g. the hazard map generated in 2009 required an ensemble of 239 workflows. In astronomy,
-a user might wish to use multiple Montage~\cite{Deelman2008}
-workflows with different parameters to generate a set of image mosaics that
-cover a given area of the sky in different wavelengths, or use an ensemble to generate
-the parts of a very large mosaic. The
-ensemble for generating a mosaic of the entire sky consists of 17 workflows, each of which contains 900 sub-workflows and
-computes one ``tile'' of the full mosaic. Another ensemble example is the
-Periodograms application~\cite{Vockler2011} that searches for extrasolar planets
-by detecting the periodic dips in light intensity. Due to the large scale of the
-input data, this application is often split up into multiple batches processed
-by different workflows. Additional workflows are created to run the analysis
-using different parameters. A recent analysis of Kepler satellite data
-required three ensembles of 15 workflows.
-
-Ensemble workflows may differ not only in their parameters, but also in their
-priority. For example, in CyberShake some sites may be in heavily populated
-areas or in strategic locations such as power plants, while others may be less
-important. Scientists typically prioritize the workflows in an ensemble such
-that important workflows are finished first. This enables them to see critical
-results early, and helps them to choose the most important workflows when the
-time and financial resources available for running the computations are limited.
+Scientific workflows, usually represented as Directed Acyclic Graphs (DAGs),
+are an important class of applications that lead to challenging problems in
+resource management on grid and utility computing systems. Workflows for large
+computational problems are often composed of several inter-related workflows
+grouped into {\em ensembles}.  Workflows in an ensemble typically have a
+similar structure, but they differ in their input data, number of tasks, and
+individual task sizes.
+
+There are many applications that require scientific workflow ensembles.
+CyberShake~\cite{Callaghan2011}, for example, uses ensembles to generate
+seismic hazard maps.  Each workflow in a CyberShake ensemble generates a
+hazard curve for a particular geographic location, and several hazard curves
+are combined to create a hazard map.  In 2009 CyberShake was used to generate
+a map that required an ensemble of 239 workflows.  Similarly, users of
+Montage~\cite{Deelman2008} often need several workflows  with different
+parameters to generate a set of image mosaics that can be combined  into a
+single, large mosaic. The Galactic Plane ensemble, which generates several
+mosaics of the  entire sky in different wavelengths, consists of 17 workflows,
+each of which contains 900 sub-workflows.  Another ensemble example is the
+Periodograms application~\cite{Vockler2011}, which searches for extrasolar
+planets by detecting periodic dips in the light intensity of their host start.
+Due to the large scale of the input data, this application is often split up
+into multiple batches processed by different workflows. Additional workflows
+are created to run the analysis using different parameters. A recent analysis
+of Kepler satellite data required three ensembles of 15 workflows.
+
+Workflows in an ensemble may differ not only in their parameters, but also in
+their priority. For example, in CyberShake some sites may be in heavily
+populated areas or in strategic locations such as power plants, while others
+may be less important. Scientists typically prioritize the workflows in such
+an ensemble so that important workflows are finished first. This enables them
+to see critical results early, and helps them to choose the most important
+workflows when the time and financial resources available for running the
+computations are limited.
 
@@
 metrics such as workflow makespan or resource utilization, but must also consider
-budget constraints, since the resources from commercia clouds
+budget constraints, since the resources from commercial clouds
 usually have monetary costs associated with them~\cite{Durkee2010}.
@@
 \begin{itemize}
-  \item we define the problem of scheduling prioritized workflows grouped
-  into ensembles, under both budget and deadline constraints,
-  \item we analyze and develop dynamic (on-line) strategies for task scheduling
-  and resource provisioning as well as static algorithms that
-  rely on the information about the workflow structure (critical paths and
-  workflow levels) and estimates of task runtimes. 
-  \item we analyze and develop a hybrid workflow-aware dynamic scheduling
-  algorithm, together with admission procedures to predict which workflows from
-  the ensemble can be completed,
-  \item we evaluate these algorithms using our simulator based on
-  CloudSim~\cite{Calheiros2011}, which models the infrastructure and the application, 
-  taking into account uncertainties in task runtime
-  estimations, provisioning delays, and failures, 
+  \item we define the problem of scheduling prioritized workflow ensembles
+  under budget and deadline constraints,
+  \item we analyze and develop several dynamic (online) and static (offline) 
+  algorithms for task scheduling and resource provisioning that rely on workflow 
+  structure information (critical paths and workflow levels) and estimates of 
+  task runtimes,
+  \item we evaluate these algorithms using a simulator based on 
+  CloudSim~\cite{Calheiros2011}, which models the infrastructure and the 
+  application,taking into account uncertainties in task runtime estimates, 
+  provisioning delays, and failures, 
   \item we discuss the performance of the algorithms on a set of
   synthetic workflow ensembles based on important, real scientific
-  applications, on a broad range of different application scenarios, and
+  applications, using a broad range of different application scenarios and
   varying constraint values.
@@
 Auto Scaling~\cite{Autoscaling} and
-RightScale~\cite{RightScale}) allow adjusting the size
-of resource pool based on metrics related to infrastructure and application.
+RightScale~\cite{RightScale}) allow the size
+of a resource pool to be adjusted based on metrics related to infrastructure and application.
 A typical infrastruc\-ture-specific metric is system load, whereas
@@
 cost-minimization workflow scheduling, developed for utility grid systems. However, our problem is
-different from \cite{Yu2005, Abrishami2010} in that we consider ensembles of workflows in
+different from \cite{Yu2005} and \cite{Abrishami2010} in that we consider ensembles of workflows in
 IaaS clouds, which allow one to provision resources on a per-hour billing model,
@@
 more work to be done than the available budget, so some work must be rejected.
-Therefore cost is not something we optimize (i.e. an objective), but rather a constraint.
+Therefore, cost is not something we optimize (i.e. an objective), but rather a constraint.
 
@@
 metaheuristics that usually run for a long time before producing good results,
-which makes them less useful in the scenarios we consider in this paper. Our work can be also
-regarded as an extension of the budget-constrained workflow scheduling
+which makes them less useful in the scenarios we consider in this paper. Our work can also
+be regarded as an extension of the budget-constrained workflow scheduling
 \cite{Sakellariou2007} in the sense that we are dealing with workflow ensembles
@@
 \underline{Resource Model}  We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2),
-where virtual machine (VM) instances are requested on-demand and are billed by
+where virtual machine (VM) instances may be provisioned on-demand and are billed by
 the hour, with partial hours being rounded up. Although there may be
@@
 applications there will typically be only one or two VM types with the best
-price/performance ratio for the application~\cite{Juve2009}. We also assume that
+price/performance ratio for the application~\cite{Juve2009}. We assume that
 a submitted task has exclusive access to a VM instance and that there is no
-preemption. We assume that there is a delay between the time that a new VM
-instance is requested and when it becomes available to execute tasks. In
-practice this delay is typically on the order of tens of seconds to a few
-minutes, and is highly dependent upon the cloud platform and the VM image
-size~\cite{Nurmi2008b}.
-
-\underline{Application Model} The target applications are ensembles of scientific workflows that can be
-modeled as DAGs, where the nodes in the graph
+preemption. We also assume that there is a delay between the time that a new VM
+instance is requested and when it becomes available to execute tasks.
+
+\underline{Application Model} The target applications are ensembles of
+scientific workflows that can be modeled as DAGs, where the nodes in the graph
 represent computational tasks, and the edges represent data- or control-flow
-dependencies between the tasks. We assume that runtime estimates for the workflow tasks
- are know and  that they are not perfect, but may
-vary based on a uniform distribution of $\pm p\%$.
+dependencies between the tasks. We assume that runtime estimates for the
+workflow tasks are known, but that they are not perfect and may vary based on
+a uniform distribution of $\pm p\%$.
 
@@
 variety of domains including: bioinformatics (Epigenomics, SIPHT: sRNA
-identification protocol using high-throughput technology), astronomy (Montage),
-earthquake science (CyberShake), and physics (LIGO).  The synthetic workflows were generated using code
-developed in~\cite{WorkflowGenerator},  with tasks runtimes based on distributions gathered from
-running real workflows.
+identification protocol using high-throughput technology), astronomy
+(Montage), earthquake science (CyberShake), and physics (LIGO). The synthetic
+workflows were generated using code developed in~\cite{WorkflowGenerator},
+with task runtimes based on distributions gathered from running real
+workflows.
 
@@
 a shared cloud storage system, such as Amazon S3, and that intermediate data
-transfer times are included in task runtime estimates. It is also assumed that
+transfer times are included in task runtimes. It is also assumed that
 data transfer times between the shared storage and the VMs are equal for
@@
 
-We assume that each workflow in an ensemble is given a numeric priority that indicates how
-important the workflow is to the user. As such, the priorities indicate the
-utility function of the user. These priorities are absolute in the sense that
-completing a workflow with a given priority is more valuable than completing all other workflows in the ensemble with lower
-priorities combined. The goal of the workflow ensemble scheduling and cloud
-provisioning problem is to complete as many high-priority workflows as possible
-given a fixed budget and deadline. Only workflows for which all tasks are
-finished by the deadline are considered to be complete---partial results are not
-usable in this model.
+We assume that each workflow in an ensemble is given a numeric priority that
+indicates how important the workflow is to the user. As such, the priorities
+indicate the utility function of the user. These priorities are absolute in
+the sense that completing a workflow with a given priority is more valuable
+than completing all other workflows in the ensemble with lower priorities
+combined. The goal of the workflow ensemble scheduling and cloud provisioning
+problem is to complete as many high-priority workflows as possible given a
+fixed budget and deadline. Only workflows for which all tasks are finished by
+the deadline are considered to be complete---partial results are not usable in
+this model.
 
@@
 
-This section describes three algorithms developed to schedule and provision resources for ensembles of workflows on the cloud under budget and deadline constraints.
+This section describes three algorithms that were developed to schedule and
+provision resources for ensembles of workflows on the cloud under budget and
+deadline constraints.
 
@@
 Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a
-VM in dollars $p$, it is easy to calculate the number of VMs, $N_{VM}$, to
+VM in dollars $p$, it is possible to calculate the number of VMs, $N_{VM}$, to
 provision so that the entire budget is consumed before the deadline:
+%
 \begin{equation}
@@
 %
-DPDS provisions $N_{VM}$ VMs at the start of the ensemble execution, then
-it periodically computes resource
-utilization using the percentage of idle VMs over time, and adjusts the number
-of VMs if the utilization is above or below given thresholds. Because it is
-assumed that VMs are billed by the hour, DPDS only considers VMs that are
-approaching their hourly billing cycle when deciding which VMs to terminate.
-This dynamic provisioning algorithm is shown in Algorithm~\ref{alg:prov}.
+DPDS provisions $N_{VM}$ VMs at the start of the ensemble execution, then it
+periodically computes resource utilization using the percentage of idle VMs
+over time and adjusts the number of VMs if the utilization is above or below
+given thresholds. Because it is assumed that VMs are billed by the hour, DPDS
+only considers VMs that are approaching their hourly billing cycle when
+deciding which VMs to terminate. This dynamic provisioning algorithm is shown
+in Algorithm~\ref{alg:prov}.
 
@@
 \EndProcedure
-\end{algorithmic}
-}
+\end{algorithmic}}
 \end{algorithm}
@@
 
-DPDS guarantees that tasks from lower priority workflows are
-always deferred when higher-priority tasks are available, but lower-priority
-tasks can still occupy idle VMs when higher-priority tasks are not available.
-Since there is no preemption, long-running low-priority tasks may
-delay the execution of higher-priority tasks. In addition, tasks from low
-priority workflows may be executed even though there is no chance that those
-workflows will be completed within the current budget and deadline.
-Fig.~\ref{fig:algorithm-example}a shows an example schedule generated using
-the DPDS algorithm. The figure illustrates how tasks from lower priority workflows
-backfill idle VMs when tasks from higher priority workflows are not available.
+DPDS guarantees that tasks from lower priority workflows are always deferred
+when higher-priority tasks are available, but lower-priority tasks can still
+occupy idle VMs when higher-priority tasks are not available. Since there is
+no preemption, long-running low-priority tasks may delay the execution of
+higher-priority tasks. In addition, tasks from low priority workflows may be
+executed even though there is no chance that those workflows will be completed
+within the current budget and deadline. Fig.~\ref{fig:algorithm-example}a
+shows an example schedule generated using the DPDS algorithm. The figure
+illustrates how tasks from lower priority workflows backfill idle VMs when
+tasks from higher priority workflows are not available.
 
@@
 
-The Workflow-Aware DPDS (WA-DPDS) algorithm
-extends DPDS by introducing a workflow admission procedure, which is invoked whenever WA-DPDS sees the first task of a new workflow at
-the head of the priority queue (i.e. when no other tasks from the workflow have
-been scheduled yet). The admission procedure---shown in
-Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough budget remaining
-to admit the new workflow; if there is not, then the workflow is rejected and
-its tasks are removed from the queue. WA-DPDS compares the current cost
-(consumed budget) and remaining budget, taking into account the cost of
-currently running VMs, and the cost of workflows that have already been
-admitted. In addition, it adds a small safety margin of \$0.10 to avoid going
-over budget. We found the admission procedure useful
-not only to prevent low-priority workflows from delaying high-priority ones, but
-also to reject large and costly workflows that would overrun the budget and
-admit smaller workflows that can efficiently utilize idle resources.
+The Workflow-Aware DPDS (WA-DPDS) algorithm extends DPDS by introducing a
+workflow admission procedure, which is invoked whenever WA-DPDS sees the first
+task of a new workflow at the head of the priority queue (i.e. when no other
+tasks from the workflow have been scheduled yet). The admission procedure---
+shown in Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough
+budget remaining to admit the new workflow; if there is not, then the workflow
+is rejected and its tasks are removed from the queue. WA-DPDS compares the
+current cost (consumed budget) and remaining budget, taking into account the
+cost of currently running VMs, and the cost of workflows that have already
+been admitted. In addition, it adds a small safety margin of \$0.10 to avoid
+going over budget. We found the admission procedure useful not only to prevent
+low-priority workflows from delaying high-priority ones, but also to reject
+large and costly workflows that would overrun the budget and admit smaller
+workflows that can efficiently utilize idle resources.
+
 % in 
@@
 
-A task's level is the length of the longest path between the task and an entry task of the workflow:
+A task's level is the length of the longest path between the task and an entry
+task of the workflow:
 %
@@
 %
-SPSS distributes the slack time of the workflow by level, so that each level of the workflow gets a portion of the workflow's slack time proportional to the number of tasks in the level and the total runtime of tasks in the level.
-The idea is that levels containing many tasks and large runtimes should be given a larger portion of the slack time so that tasks in those levels may be serialized.
-Otherwise, many resources need to be allocated to run all of the tasks in parallel, which may be more costly.
+SPSS distributes the slack time of the workflow by level, so that each level
+of the workflow gets a portion of the workflow's slack time proportional to
+the number of tasks in the level and the total runtime of tasks in the level.
+The idea is that levels containing many tasks and large runtimes should be
+given a larger portion of the slack time so that tasks in those levels may be
+serialized. Otherwise, many resources need to be allocated to run all of the
+tasks in parallel, which may be more costly.
 
@@
 %
-where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of tasks in level $l$, $R(w)$ is the total runtime of all tasks in the workflow, $R(w)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a parameter between 0 and 1 that causes more slack time to be given to levels with more tasks (large $\alpha$) or more runtime (small $\alpha$).
+where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of
+tasks in level $l$, $R(w)$ is the total runtime of all tasks in the workflow,
+$R(w)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a
+parameter between 0 and 1 that causes more slack time to be given to levels
+with more tasks (large $\alpha$) or more runtime (small $\alpha$).
 
@@
 %
-where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and $LST(t)$ is the latest start time of $t$ determined by:
+where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and 
+$LST(t)$ is the latest start time of $t$ determined by:
 %
@@
 \textsc{DeadlineDistribution} to assign sub-deadlines to tasks.
-% according to the procedure described above. 
 Then, \textsc{PlanWorkflow} schedules
@@
 randomized ensembles using workflows available from the workflow generator
-gallery~\cite{WorkflowGenerator}. 
-%% Ewa: maybe put url instead of citation
-The
+gallery~\cite{WorkflowGenerator}. The
 gallery contains synthetic workflows modeled using structures and parameters
@@
 Using this collection of workflows, we constructed five different ensemble types:
-%
-%\begin{itemize}
 constant, uniform sorted, uniform unsorted, Pareto sorted and Pareto unsorted.
-%\end{itemize}
-In the unsorted ensembles, workflows of different sizes are mixed together, and
+In the unsorted ensembles, workflows of different sizes are mixed together and
 the priorities are assigned randomly.
@@
 %in a uniform ensemble 
-is selected by first randomly choosing the size the workflow
+is selected by first randomly choosing the size of the workflow
 % according to a uniform distribution, 
@@
 %just a few workflows can be aggregated into a single workflow, so there is no need to
-%treat them as an ensemble. Similarly, when the number of workflows grows, and
+%treat them as an ensemble. Similarly, when the number of workflows grows and
 %each workflow has a large number of tasks, either the deadline and budget
@@
 %
-%
 $$
@@
 This range---$[MinDeadline, MaxDeadline]$---is then divided into equal intervals. 
-By computing the budget and deadline constraits in this way we ensure that the
+By computing the budget and deadline constraints in this way we ensure that the
 experiments for each ensemble cover the most interesting area of the parameter space for the
@@
 
-The second thing to notice about the figure is that, for constant ensembles,
-the dynamic algorithms perform significantly better relative to SPSS than for
-other ensemble types. This is a result of the fact that, since all of the
-workflows are of approximately the same size and shape, the choice of which
-workflow to execute next has little impact on the final result.
+The second thing to notice is that, for constant ensembles, the dynamic
+algorithms perform significantly better relative to SPSS compared to  other
+ensemble types. This is a result of the fact that, since all of the workflows
+are of approximately the same size and shape, the choice of which workflow to
+execute next has a smaller impact on the final result.
 
@@
 %as the scaling factor is increased from 1 to 16. The figure only shows results
-%for the CyberShake application with pareto unsorted ensembles, but similar
+%for the CyberShake application with Pareto unsorted ensembles, but similar
 %results were found for other ensemble types for both Montage and CyberShake.
@@
  %   \includegraphics[width=0.3\textwidth]{stretching_cybershake}
- %   \caption{Percentage of high scores achieved by each algorithm for pareto 
+ %   \caption{Percentage of high scores achieved by each algorithm for Pareto 
  %   unsorted CyberShake ensembles when task runtime is scaled by a constant 
@@
 How do errors in task runtime estimates impact the performance of our
-scheduling and provisioning algorithms?  Thus, we introduced uniform errors in the task runtime
-and observed the behavior of the algorithms in terms of meeting  the desired budget and deadline constraints.
-
-The actual runtime of each task is adjusted in the simulation by adding a
-random error to the estimated runtime of $\pm p\%$. Since the sampling is done
-uniformly, we expect to get just as many overestimates as underestimates in
-any given simulation.
-
-Fig.~\ref{fig:variances} shows the results of simulations for estimate errors
-ranging from $0\%$ to $50\%$. The figure summarizes the results of an
+scheduling and provisioning algorithms?  To examine this we introduced uniform
+errors in the task runtime and observed the behavior of the algorithms in
+terms of meeting the desired budget and deadline constraints.
+
+In this experiment the actual runtime of each task is adjusted in the
+simulation by adding a random error to the estimated runtime of $\pm p\%$.
+Since the sampling is done uniformly, we expect to get just as many
+overestimates as underestimates in any given simulation.
+
+Fig.~\ref{fig:variances} shows the results for estimate errors
+ranging from $0\%$ to $50\%$. The figure summarizes the outcome of an
 extensive suite of 525,000 simulations (10 ensembles of 50 workflows x 5
@@
 3 algorithms). The figure shows box plots of the ratio of the ensemble cost to
-budget and ensemble makespan to deadline. The ratio indicates whether the value (for example the simulated cost) exeeded the constraint (the budget): Values
+budget, and of ensemble makespan to deadline. The ratio indicates whether the value 
+(for example, the simulated cost) exeeded the constraint (the budget). Values
 greater than 1 indicate that the constraint was exceeded.
@@
 is a consequence of the deadline distribution function in SPSS, which prevents
-workflows from starting late. This is shown by the gantt chart in
-Fig.~\ref{fig:algorithm-example}.b, which illustrates how SPSS tends to pile
+workflows from starting late. This is illustrated by the gantt chart in
+Fig.~\ref{fig:algorithm-example}.b, which shows how SPSS tends to pile
 up workflows at the beginning of the timeline. 
@@
 %shown in Fig.~\ref{fig:algorithm-example}.a. The consequence of this
-5behavior is that SPSS plans tend to use up the budget but leave plenty of time
+%behavior is that SPSS plans tend to use up the budget but leave plenty of time
 %before the deadline. 
@@
 We assume that resources are billed from the minute that they are requested
-until they are terminated. 
-As a result, provisioning delays have an impact on both the cost  and makespan of the ensemble.
+until they are terminated.  As a result, provisioning delays have an impact on
+both the cost and makespan of an ensemble.
 
@@
 x 5 distributions x 10 budgets x 10 deadlines x 7 error values x 3
-algorithms). Only one application, Montage, was used in this experiment. As in
-Fig.~\ref{fig:variances}, the y-axis in each plot represents the ratio of the simulated value.
- to the constraint value for the budget constraint or the deadline constraint.
+algorithms). To reduce the simulation time, only one application, Montage, was
+used in this experiment.
+
+%As in Fig.~\ref{fig:variances}, the y-axis in each
+%plot represents the ratio of the simulated value to the constraint value for
+%the budget constraint or the deadline constraint.
 
@@
 constraints while the static algorithm is not. In the case of delays of more
-than 5 minutes, which is typical of what has been observed on academic
+than one minute, which is typical of what has been observed on academic
 infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and
 FutureGrid~\cite{FutureGrid}, approximately half of the SPSS simulations
-exceeded the budget while none of the simulations using a dynamic algorithm
-exceeded the budget.
+exceeded the budget; in some cases by up to a factor of 2. In comparison,
+none of the simulations that used a dynamic algorithm exceeded the budget
+or the deadline constraint.
 
@@
 
-This experiment suggests that SPSS is far too sensitive to provisioning delays
+This experiment suggests that SPSS is too sensitive to provisioning delays
 in its current form to be of practical use in real systems. It is possible
@@
 
-\TODO{Include a description of the ensembles that were simulated.}
-
-Fig.~\ref{fig:failures} shows the ratios of simulated values to constraints when
-the failures are introduced. The high failure rates have a high impact on the
-performance of the static algorithm and can degrade the solution considerably.
-Comparing Fig.~\ref{fig:failures} to Figs.~\ref{fig:variances}
-and~\ref{fig:delays} one may conclude that failures are worse than
-provisioning delays and runtime estimate errors, since their impact is larger.
-However, small failure rates (below 10\%) do not cause constraint overruns
-larger than 10\%. We consider higher failure rates as rare events that suggest
-a significant system malfunction or invalid selection of resources.
+Fig.~\ref{fig:failures} shows the ratios of simulated values to constraints
+when the failures are introduced. The figure summarizes a suite of 525,000
+simulations (10 ensembles of 50 workflows x 5 applications x 5 distributions x
+10 budgets x 10 deadlines x 7 failure rates x 3 algorithms). As in the
+previous  experiments, these results show that high failure rates have a large
+impact on the performance of the static algorithm and can degrade the solution
+considerably, while the dynamic algorithms are able to adapt. Comparing
+Fig.~\ref{fig:failures} to Figs.~\ref{fig:variances} and~\ref{fig:delays} one
+may conclude that failures are worse than provisioning delays and runtime
+estimate errors, since their impact is larger. However, we consider
+higher failure rates as rare events that suggest a significant system
+malfunction or invalid selection of resources.
 
@@
 
- Because SPSS involves more complicated
-logic than the dynamic algorithms and makes its decisions ahead of the execution, it is important to understand what impact
-this has on execution time.
+Because SPSS involves more complicated logic than the dynamic algorithms and
+makes its decisions before execution, it is important to understand what
+impact planning time has on the overall execution time.
 
@@
 the workflow size desired. Two different applications were used: SIPHT and
-CyberShake. Each box summarizes the results of 2 applications, 10 random
-seeds, 10 budgets and 10 deadlines, or 2000 simulations per box.
+CyberShake. Each box summarizes the results of 2000 simulations (2
+applications x 10 ensembles x 10 budgets x 10 deadlines).
 
@@
 tasks are common, and workflows with hundreds of thousands or millions of
-tasks are not impossible), it is unlikely that SPSS will be practical for
+tasks are possible), it is unlikely that SPSS will be practical for
 ensembles of very large workflows.
@@
 
-%The algorithmic complexity of SPSS is $O(n^2)$, where $n$ is the number of
-%tasks in the ensemble. This derives from the fact that, for each task,
-
-SPSS
-considers scheduling each task on the cheapest available slot, which involves
-scanning all of the available slots on all the VMs. Since the number of
-available slots, in the worst case, is proportional to the number of tasks
-scheduled
-%Ewa:  I don't get that
- (because scheduling a task splits an existing slot into at most two
-slots, one before the task, and one after), the complexity of SPSS is $O(n^2)$, where $n$ is the number of
-tasks in the ensemble. In comparison, the dynamic
-algorithms all have a more scalable complexity of $O(n)$. DPDS only examines
-the tasks in the workflow once when they are scheduled, and WA-DPDS examines
-them twice: once in the admissions algorithm, and once when they are
+SPSS considers scheduling each task on the cheapest available slot, which
+involves scanning all of the available slots on all of the VMs. Since the number
+of available slots, in the worst case, is proportional to the number of tasks
+scheduled (because scheduling a task splits an existing slot into at most two
+slots: one before the task, and one after), the complexity of SPSS is
+$O(n^2)$, where $n$ is the number of tasks in the ensemble. In comparison, the
+dynamic algorithms all have a more scalable complexity of $O(n)$. DPDS only
+examines the tasks in the workflow once when they are scheduled, and WA-DPDS
+examines them twice: once in the admissions algorithm, and once when they are
 scheduled. This makes the dynamic algorithms a better fit for larger workflows
-and ensembles even though in some cases they may not produce results that are as good as
-SPSS.
+and ensembles even though in some cases they may not produce results that are
+as good as SPSS.
 
@@
 The results of our simulation studies indicate that the two algorithms that
-take into account the structure of the workflow 
-%Ewa:  so it is really taking into account the workflow structure or the ensemble structure/members?
-and the members of the ensemble (through admission procedures) and task runtime estimates
+take into account the structure of the workflow and task runtime estimates
 (WA-DPDS and SPSS) yield better results than the simple priority-based
 scheduling strategy (DPDS), which makes provisioning decisions based purely on
-resource utilization.  This underscores the importance of viewing workflow ensembles as a whole rather than just individual tasks or individual workflows.
+resource utilization. This underscores the importance of viewing workflow
+ensembles as a whole rather than as individual tasks or individual workflows.
 
@@
 % This balances the columns by inserting a column break at the given reference
-\IEEEtriggeratref{16}
+\IEEEtriggeratref{17}
 

### 2012-05-03 paper/paper.tex
makespan clarification

@@
 Periodograms application~\cite{Vockler2011}, which searches for extrasolar
-planets by detecting periodic dips in the light intensity of their host start.
+planets by detecting periodic dips in the light intensity of their host star.
 Due to the large scale of the input data, this application is often split up
@@
 3 algorithms). The figure shows box plots of the ratio of the ensemble cost to
-budget, and of ensemble makespan to deadline. The ratio indicates whether the value 
+budget, and of ensemble makespan to deadline. Whiskers on the plots
+indicate maximum and minimum values. The ratio indicates whether the value 
 (for example, the simulated cost) exeeded the constraint (the budget). Values
@@
 Interestingly, the deadline constraint is rarely exceeded, even in cases with
-very poor quality estimates. For the dynamic algorithms this is a result of
-the fact that they can adapt to poor estimates. For the static algorithm this
-is a result of the way that SPSS schedules workflows, and not of a
-particularly clever optimization. The SPSS algorithm tends to schedule
-workflows early, using up the budget long before the deadline is reached. This
-is a consequence of the deadline distribution function in SPSS, which prevents
-workflows from starting late. This is illustrated by the gantt chart in
-Fig.~\ref{fig:algorithm-example}.b, which shows how SPSS tends to pile
-up workflows at the beginning of the timeline. 
-%In comparison, the dynamic
-%algorithms tend to spread out workflow starts over the entire duration as
-%shown in Fig.~\ref{fig:algorithm-example}.a. The consequence of this
-%behavior is that SPSS plans tend to use up the budget but leave plenty of time
-%before the deadline. 
-As a result, when the runtime of the plan is increased by
-introducing errors, the SPSS plan has some room to expand without exceeding
-the deadline.
+very poor quality estimates. For the dynamic algorithms this is a result of the
+fact that they can adapt to poor estimates and stop submitting new tasks when
+the constraints are reached. Makespan is then computed as the finish time of the
+last fully completed workflow from the ensemble. For the static algorithm this
+is a result of the way that SPSS schedules workflows, and not of a particularly
+clever optimization. The SPSS algorithm tends to schedule workflows early, using
+up the budget long before the deadline is reached. This is a consequence of the
+deadline distribution function in SPSS, which prevents workflows from starting
+late. This is illustrated by the gantt chart in
+Fig.~\ref{fig:algorithm-example}.b, which shows how SPSS tends to pile up
+workflows at the beginning of the timeline.
+% In comparison, the dynamic algorithms tend to spread out workflow starts over
+% the entire duration as shown in Fig.~\ref{fig:algorithm-example}.a. The
+% consequence of this behavior is that SPSS plans tend to use up the budget but
+% leave plenty of time before the deadline.
+As a result, when the runtime of the plan is increased by introducing errors,
+the SPSS plan has some room to expand without exceeding the deadline.
 

### 2012-07-19 paper/paper.tex
Balance columns
applied reviewers comments

@@
 \documentclass[conference]{IEEEtran}
-
+\usepackage{fancyhdr}
 \usepackage[pdftex]{graphicx}
@@
 
+\thispagestyle{fancy}
+\lhead{}
+\rhead{}
+\chead{}
+\lfoot{\footnotesize{
+SC12, November 10-16, 2012, Salt Lake City, Utah, USA
+\newline
+978-1-4673-0806-9/12/\$31.00 \copyright 2012 IEEE}}
+\rfoot{}
+\cfoot{}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrulewidth}{0pt}
+
 \begin{abstract}
@@
 to see critical results early, and helps them to choose the most important
-workflows when the time and financial resources available for running the
-computations are limited.
+workflows when the time and financial resources available for computing are
+limited.
 
@@
 RightScale~\cite{RightScale}) allow the size
-of a resource pool to be adjusted based on metrics related to infrastructure and application.
-A typical infrastruc\-ture-specific metric is system load, whereas
-application-speci\-fic metrics include response time and length of a task or
+of a resource pool to be adjusted based on infrastructure and
+application metrics. A typical infrastruc\-ture-specific metric is system load,
+whereas application-speci\-fic metrics include response time and length of a task or
 of a request queue. It is possible to set thresholds and limits to tune the behavior
@@
 in that we consider workflows, while policy based approaches typically consider
-bags of independent tasks or unpredictable batch workloads. Our approach enables us to
-take advantage of workflow-aware scheduling heuristics that cannot be applied to
+bags of independent tasks or unpredictable batch workloads. This enables us to
+take advantage of workflow-aware heuristics that cannot be applied to
 independent tasks.
@@
 
+\underline{Performance Metric}
+\label{sec:perf_metric}
+In order to precisely define the objective of the algrithms it is necessary to
+introduce a metric that can be used to score the performance of the different
+algorithms on a given problem (ensemble, budget, and deadline). The simplest
+approach is to count the number of workflows in the ensemble that each algorithm
+is able to complete within the budget before the deadline, but this metric does
+not account for the priority-based utility function specified by the user. Using
+the counting approach, a less efficient algorithm may be able to complete a
+large number of low-priority workflows by executing the smallest workflows
+first. In order to account for the priority, we use an exponential
+scoring defined as:
+$$ Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}} $$
+where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed
+by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that
+the highest-priority workflow has $Priority(w)=0$, the next highest workflow has
+$Priority(w)=1$, and so on. This exponential scoring function gives the highest
+priority workflow a score that is higher than all the lower- priority workflows
+combined:
+$$ 2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i} $$
+%  This scoring is consistent with our definition of the problem, which is to
+% complete as many workflows as possible, according to their priorities, given a
+% set budget and deadline.
+
 \section{Algorithms}
@@
 deadline constraints.
-
-\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
-\label{sec:dpds}
-
-DPDS is an online algorithm that provisions resources and schedules tasks at
-runtime. It consists of two main parts: a provisioning procedure, and a
-scheduling procedure.
-
-DPDS' provisioning procedure is based on resource utilization. DPDS starts with
-a fixed number of resources calculated based on the available time and budget,
-and adjusts the number of resources according to how well they are utilized by
-the application.
-Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a
-VM in dollars $p$, it is possible to calculate the number of VMs, $N_{VM}$, to
-provision so that the entire budget is consumed before the deadline:
-%
-\begin{equation}
-\label{eq:static-plan}
-N_{VM} = \lceil b / (d * p) \rceil
-\end{equation}
-%
-DPDS provisions $N_{VM}$ VMs at the start of the ensemble execution, then it
-periodically computes resource utilization using the percentage of idle VMs
-over time and adjusts the number of VMs if the utilization is above or below
-given thresholds. Because it is assumed that VMs are billed by the hour, DPDS
-only considers VMs that are approaching their hourly billing cycle when
-deciding which VMs to terminate. This dynamic provisioning algorithm is shown
-in Algorithm~\ref{alg:prov}.
 
@@
 
-The set of VMs completing their billing cycle is determined by both
-the provisioner interval, and the termination delay of the provider. This
-guarantees that VMs can be terminated before they start the next billing cycle
-and prevents the budget from being overrun. The VMs terminated in line
-\ref{l:terminate1} of Algorithm~\ref{alg:prov} are the ones that would overrun
-the budget if not terminated in the current provisioning cycle. The VMs
-terminated in line \ref{l:terminate2} are chosen to increase the resource
-utilization to the desired threshold. In order to prevent instances that have
-already been paid for from being terminated too quickly, no more than half of
-the idle resources are terminated during each provisioning cycle. To avoid an
-uncontrolled increase in the number of instances, which may happen in the case
-of highly parallel workflows, the provisioner will not start a new VM if the
-number of running VMs is greater than the product of $N_{VM}$ (from
-Equation~\ref{eq:static-plan}) and an autoscaling parameter, $v_{max}$. Unless
-otherwise specified, $v_{max}$ is assumed to be 1.
-
-In order to schedule individual workflow tasks onto available VMs, DPDS uses the
-dynamic, priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
-Initially, the ready tasks from all workflows in the ensemble are added to a
-priority queue based on the priority of the workflow to which they belong. If
-there are idle VMs available, and the priority queue is not empty, the next task
-from the priority queue is submitted to an arbitrarily chosen idle VM. The
-process is repeated until there are no idle VMs or the priority queue is empty.
-The scheduler then waits for a task to finish, adds its ready children to the
-priority queue, marks the VM as idle, and the entire process repeats until the
-deadline is reached.
+\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
+\label{sec:dpds}
+
+DPDS is an online algorithm that provisions resources and schedules tasks at
+runtime. It consists of two main parts: a provisioning procedure, and a
+scheduling procedure.
+
+DPDS' provisioning procedure is based on resource utilization. DPDS starts with
+a fixed number of resources calculated based on the available time and budget,
+and adjusts the number of resources according to how well they are utilized by
+the application.
+Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a
+VM in dollars $p$, it is possible to calculate the number of VMs, $N_{VM}$, to
+provision so that the entire budget is consumed before the deadline:
+%
+\begin{equation}
+\label{eq:static-plan}
+N_{VM} = \lceil b / (d * p) \rceil
+\end{equation}
+%
+DPDS provisions $N_{VM}$ VMs at the start of the ensemble execution, then it
+periodically computes resource utilization using the percentage of idle VMs
+over time and adjusts the number of VMs if the utilization is above or below
+given thresholds. Because it is assumed that VMs are billed by the hour, DPDS
+only considers VMs that are approaching their hourly billing cycle when
+deciding which VMs to terminate. This dynamic provisioning algorithm is shown
+in Algorithm~\ref{alg:prov}.
 
@@
 
+The set of VMs completing their billing cycle is determined by both
+the provisioner interval, and the termination delay of the provider. This
+guarantees that VMs can be terminated before they start the next billing cycle
+and prevents the budget from being overrun. The VMs terminated in line
+\ref{l:terminate1} of Algorithm~\ref{alg:prov} are the ones that would overrun
+the budget if not terminated in the current provisioning cycle. The VMs
+terminated in line \ref{l:terminate2} are chosen to increase the resource
+utilization to the desired threshold. In order to prevent instances that have
+already been paid for from being terminated too quickly, no more than half of
+the idle resources are terminated during each provisioning cycle. To avoid an
+uncontrolled increase in the number of instances, which may happen in the case
+of highly parallel workflows, the provisioner will not start a new VM if the
+number of running VMs is greater than the product of $N_{VM}$ (from
+Equation~\ref{eq:static-plan}) and an autoscaling parameter, $v_{max}$. Unless
+otherwise specified, $v_{max}$ is assumed to be 1.
+
+In order to schedule individual workflow tasks onto available VMs, DPDS uses the
+dynamic, priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
+Initially, the ready tasks from all workflows in the ensemble are added to a
+priority queue based on the priority of the workflow to which they belong. If
+there are idle VMs available, and the priority queue is not empty, the next task
+from the priority queue is submitted to an arbitrarily chosen idle VM. The
+process is repeated until there are no idle VMs or the priority queue is empty.
+The scheduler then waits for a task to finish, adds its ready children to the
+priority queue, marks the VM as idle, and the entire process repeats until the
+deadline is reached.
+
 DPDS guarantees that tasks from lower priority workflows are always deferred
@@
 cost of currently running VMs, and the cost of workflows that have already
-been admitted. In addition, it adds a small safety margin of \$0.10 to avoid
-going over budget. We found the admission procedure useful not only to prevent
+been admitted. In addition, it adds a small safety margin of \$0.10 (10\% of the
+compute hour cost) to avoid going over budget. We found the admission procedure useful not only to prevent
 low-priority workflows from delaying high-priority ones, but also to reject
@@
 cost the full price of a billing cycle. 
-%For example, if a task has a runtime of
-%10 minutes, and the price of a block is \$1, then the algorithm will either
-%schedule the task on an existing VM that has an idle slot larger than 10
-%minutes for a cost of \$0, or it will allocate a new block on an existing VM,
-%or provision a new VM, for a cost of \$1. 
+For example, if a task has a runtime of
+10 minutes, and the price of a block is \$1, then the algorithm will either
+schedule the task on an existing VM that has an idle slot larger than 10
+minutes for a cost of \$0, or it will allocate a new block on an existing VM,
+or provision a new VM, for a cost of \$1. 
 If the cost of slots on two different
@@
   schedules generated by the algorithms. Each row is a different
-  VM. Tasks are represented as boxes and are colored by workflow.}
+  VM. Boxes are tasks colored by workflow.}
   \label{fig:algorithm-example}
@@
 
-\subsection{Performance Metric}
-\label{sec:perf_metric}
-
-In order to quantitively compare our algorithms it is necessary to define a
-metric that can be used to score the performance of the different algorithms on
-a given problem (ensemble, budget, and deadline). The simplest approach is to
-count the number of workflows in the ensemble that each algorithm is able to
-complete within the budget before the deadline, but this metric does not account
-for the priority-based utility function specified by the user. Using the
-counting approach, a less efficient algorithm may be able to complete a large
-number of low-priority workflows by executing the smallest workflows first.
-In order to account for the user's priority, we used an exponential scoring defined as:
-%
-$$
-Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
-$$
-%
-where $Completed(e)$ is the set of workflows in ensemble $e$ that was
-completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$
-such that the highest-priority workflow has $Priority(w)=0$, the next highest
-workflow has $Priority(w)=1$, and so on. This exponential scoring function
-gives the highest priority workflow a score that is higher than all the lower-
-priority workflows combined:
-%
-$$
-2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i}
-$$
-%
-%This scoring is consistent with our definition of the problem, which is to
-%complete as many workflows as possible, according to their priorities, given a
-%set budget and deadline.
-
 \section{Discussion of Results}
@@
 of the proposed algorithms. This was done by simulating the algorithms on many
-different ensembles and comparing the scores computed using the exponential
-scoring technique described in Section~\ref{sec:perf_metric}.
+different ensembles and comparing the scores computed using the
+performance metric based on exponential scoring defined in
+Section~\ref{sec:perf_metric}.
 
@@
 
-Another thing to notice about this experiment is that the workflow-aware
+We can also see that the workflow-aware
 algorithms (WA-DPDS and SPSS) both perform better in most cases than the
@@
 
-Finally, it is interesting to see that, for Montage and CyberShake, the
+Finally, it is interesting that, for Montage and CyberShake, the
 relative performance of SPSS is significantly less than it is for other
@@
 
-Both of the workflow-aware algorithms rely on estimates of task runtimes to
-make better scheduling and provisioning decisions. In practice, however, these
-estimates are often inaccurate. Given inaccuate estimates, the question is:
-How do errors in task runtime estimates impact the performance of our
-scheduling and provisioning algorithms?  To examine this we introduced uniform
-errors in the task runtime and observed the behavior of the algorithms in
-terms of meeting the desired budget and deadline constraints.
+Both of the workflow-aware algorithms rely on estimates of task runtimes to make
+better scheduling and provisioning decisions. Our experience suggests that such
+assumption is often reasonable, since we can obtain workflow performance
+characteristics from preliminary runs~\cite{Bharathi2008,Deelman2005,Juve2010}.
+Some applications, like Periodograms~\cite{Vockler2011} include even a
+performance model that estimates task runtimes and automatically annotates
+workflow description with these data. In practice, however, these
+estimates are often inaccurate. Given inaccuate estimates, the question is: How
+do errors in task runtime estimates impact the performance of our scheduling and
+provisioning algorithms?  To examine this we introduced uniform errors in the
+task runtime and observed the behavior of the algorithms in terms of meeting the
+desired budget and deadline constraints.
 
@@
 applications x 5 distributions x 10 budgets x 10 deadlines x 7 error values x
-3 algorithms). The figure shows box plots of the ratio of the ensemble cost to
+3 algorithms). Box plots show the ratio of the ensemble cost to
 budget, and of ensemble makespan to deadline. Whiskers on the plots
@@
 
-The effect of provisioning delays on workflow performance is similar to that
-of inaccurate runtime estimates: when the delays are small, all algorithms are
-able to produce results within the constraints, but when the delays are
-larger, the dynamic algorithms are able to adapt to avoid exceeding the
-constraints while the static algorithm is not. In the case of delays of more
-than one minute, which is typical of what has been observed on academic
-infrastructure clouds~\cite{Juve2011} such as Magellan~\cite{Magellan} and
-FutureGrid~\cite{FutureGrid}, approximately half of the SPSS simulations
-exceeded the budget; in some cases by up to a factor of 2. In comparison,
-none of the simulations that used a dynamic algorithm exceeded the budget
-or the deadline constraint.
+The effect of provisioning delays on workflow performance is similar to that of
+inaccurate runtime estimates: when the delays are small, all algorithms are able
+to produce results within the constraints, but for larger delays, the dynamic
+algorithms are able to adapt to avoid exceeding the constraints while the static
+algorithm is not. In the case of delays of more than one minute, which is
+typical of what has been observed on academic clouds~\cite{Juve2011} such as
+Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, approximately half of
+the SPSS simulations exceeded the budget; in some cases by up to a factor of 2.
+In comparison, none of the simulations that used a dynamic algorithm exceeded
+the budget or the deadline constraint.
 
@@
 
-Fig.~\ref{fig:failures} shows the ratios of simulated values to constraints
-when the failures are introduced. The figure summarizes a suite of 525,000
+Fig.~\ref{fig:failures} shows the ratios of simulated values to constraints when
+the failures are introduced. The figure summarizes a suite of 525,000
 simulations (10 ensembles of 50 workflows x 5 applications x 5 distributions x
-10 budgets x 10 deadlines x 7 failure rates x 3 algorithms). As in the
-previous  experiments, these results show that high failure rates have a large
-impact on the performance of the static algorithm and can degrade the solution
-considerably, while the dynamic algorithms are able to adapt. Comparing
-Fig.~\ref{fig:failures} to Figs.~\ref{fig:variances} and~\ref{fig:delays} one
-may conclude that failures are worse than provisioning delays and runtime
-estimate errors, since their impact is larger. However, we consider
-higher failure rates as rare events that suggest a significant system
-malfunction or invalid selection of resources.
+10 budgets x 10 deadlines x 7 failure rates x 3 algorithms). As in the previous 
+experiments, these results show that high failure rates can degrade the
+performance of the static algorithm considerably, while the dynamic algorithms
+are able to adapt. Comparing Fig.~\ref{fig:failures} to
+Figs.~\ref{fig:variances} and~\ref{fig:delays} one may conclude that failures
+are worse than provisioning delays and runtime estimate errors, since their
+impact is larger. However, we consider higher failure rates as rare events that
+suggest a significant system malfunction or invalid selection of resources.
 
@@
 real workflows are often much larger (workflows with tens of thousands of
-tasks are common, and workflows with hundreds of thousands or millions of
+tasks are common, and even workflows with millions of
 tasks are possible), it is unlikely that SPSS will be practical for
 ensembles of very large workflows.
-
-\begin{figure}[tb]
-    \centering
-    \includegraphics[width=0.3\textwidth]{spss_planning_time}
-    \caption{Planning time of SPSS algorithm for ensembles of 100 workflows and different workflow sizes.}
-    \label{fig:spss_planning_time}
-\end{figure}
 
@@
 examines the tasks in the workflow once when they are scheduled, and WA-DPDS
-examines them twice: once in the admissions algorithm, and once when they are
+does it twice: once in the admission algorithm, and once when they are
 scheduled. This makes the dynamic algorithms a better fit for larger workflows
-and ensembles even though in some cases they may not produce results that are
-as good as SPSS.
+and ensembles even though in some cases they may not produce as good results
+as SPSS.
 
@@
 store the available slots. Investigating these topics is left for future work.
+
+\begin{figure}[tb]
+    \centering
+    \vspace{-6mm}
+    \includegraphics[width=0.30\textwidth]{spss_planning_time}
+    \vspace{-3mm}
+    \caption{Planning time of SPSS algorithm for ensembles of 100 workflows and different workflow sizes.}
+    \label{fig:spss_planning_time}
+\end{figure}
 
@@
 that the data demands of scientific workflows have a large impact on not only
-the execution time, but also on the cost of workflows in commercial clouds. We
-also plan to investigate heterogeneous environments that include multiple VM
+the execution time, but also on the cost of workflows in commercial clouds. 
+In dynamic algorithms, it will be interesting to extend our
+utilization-based autoscaling with more advanced workflow-aware strategies
+based on feedback control.
+We also plan to investigate heterogeneous environments that include multiple VM
 types and cloud providers, including private and community clouds, which will
@@
 % This balances the columns by inserting a column break at the given reference
-\IEEEtriggeratref{17}
+\IEEEtriggeratref{15}
 

### 2012-07-25 paper/paper.tex
Add STCI acknowlegement
improved Fig. 1

@@
 executed even though there is no chance that those workflows will be completed
-within the current budget and deadline. Fig.~\ref{fig:algorithm-example}a
+within the current budget and deadline. Fig.~\ref{fig:algorithm-example-dpds}
 shows an example schedule generated using the DPDS algorithm. The figure
@@
   \subfloat[DPDS]{
-  \includegraphics[width=0.40\textwidth]{figures/spds-gantt}
+  \includegraphics[width=0.40\textwidth]{figures/output-test-DPDS}
+  \label{fig:algorithm-example-dpds}
   }\\
   \subfloat[SPSS]{
-  \includegraphics[width=0.40\textwidth]{figures/spss-gantt}
+  \includegraphics[width=0.40\textwidth]{figures/output-test-SPSS}
+  \label{fig:algorithm-example-spss}
   }
@@
   VM. Boxes are tasks colored by workflow.}
-  \label{fig:algorithm-example}
+
 \end{figure}
 
-An example schedule generated by SPSS (Fig.~\ref{fig:algorithm-example}b)
+An example schedule generated by SPSS (Fig.~\ref{fig:algorithm-example-spss})
 shows how SPSS tends to start many workflows in parallel, running each workflow
@@
     \subfloat[Ratio of Simulated Ensemble Cost to Budget]{
-        \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-cost-all}
+        \includegraphics[width=0.38\textwidth]{run-finish-failures-test-0-output-cost-all}
         }
@@
     \subfloat[Ratio of Simulated Ensemble Makespan to Deadline]{
-        \includegraphics[width=0.4\textwidth]{run-finish-failures-test-0-output-makespan-all}
+        \includegraphics[width=0.38\textwidth]{run-finish-failures-test-0-output-makespan-all}
         }
@@
     \vspace{-6mm}
-    \includegraphics[width=0.30\textwidth]{spss_planning_time}
+    \includegraphics[width=0.27\textwidth]{spss_planning_time}
     \vspace{-3mm}
@@
 
-%\section*{Acknowledgments}
-%Acknowledgements go here.
+\section*{Acknowledgments}
+This work was sponsored by the National Science Foundation under award 
+OCI-0943725 (STCI).
 

### 2012-07-25 scripts/ruby/plots.rb

@@ -71,8 +71,10 @@ def plot_schedule (filename)
   tasks = Tasks.new
   tasks.read_tasks(filename)
   
-  vms = read_array("cat #{filename}-vms.txt | grep -e '[0-9]'  ")
-
+  # change this to plot VMs
+  # vms = read_array("cat #{filename}-vms.txt | grep -e '[0-9]'  ")
+  vms = {}
+  
   ymax = vms.length + 0.8
   
   inputs = read_array("cat #{filename}-inputs-transfer.txt | grep -e '[0-9]'  ")
@@ -84,21 +86,35 @@ def plot_schedule (filename)
   Gnuplot.open do |gp|
     Gnuplot::Plot.new( gp ) do |plot|
 
-      plot.title  "Schedule " + File.basename(filename)
-      plot.xlabel "time"
-      plot.ylabel "VM ID"
-      plot.xtics 3600
+      #plot.title  "Schedule " + File.basename(filename)
+      plot.xlabel "Time"
+      plot.ylabel "VM"
+      #plot.xtics 3600
       #plot.ytics 50
       #plot.yrange "[-0.8:#{ymax}]"
-      #plot.terminal 'pdf size 8.5,11 font "arial,6" linewidth 1'
+      plot.terminal 'pdfcairo size 5,1.5 font "arial,8" linewidth 1'
       #plot.terminal 'pdf size 11,8.5 font "arial,6" linewidth 1'
-      #plot.output filename + ".pdf"
+      plot.output filename + ".pdf"
       #puts "Saving plot to file: " + filename + ".pdf"
-      plot.set "key right outside"
-      plot.set "grid"
-      plot.terminal "png size 1024,768"
+      #plot.set "key right outside"
+      plot.set "key off"
+      plot.noytics
+      plot.noxtics
+      #plot.set "grid"
+      plot.set "style line 1 lc rgb 'grey10' lt 1 lw 1"
+      plot.set "style line 2 lc rgb 'brown' lt 1 lw 1"
+      plot.set "style line 3 lc rgb 'orange' lt 1 lw 1"
+      plot.set "style line 4 lc rgb 'blue' lt 1 lw 1"
+      plot.set "style line 5 lc rgb 'green' lt 1 lw 1"
+      plot.set "style line 6 lc rgb 'grey10' lt 1 lw 1"
+      plot.set "style line 7 lc rgb 'brown' lt 1 lw 1"
+      plot.set "style line 8 lc rgb 'orange' lt 1 lw 1"
[…]

### 2012-07-25 src/cws/core/provisioner/DynamicProvisionerDynamicSchedulerTest.java

@@ -6,6 +6,7 @@ import java.io.File;
 import java.util.ArrayList;
 import java.util.HashSet;
 import java.util.List;
+import java.util.Random;
 
 import org.cloudbus.cloudsim.core.CloudSim;
 import org.junit.Test;
@@ -18,8 +19,11 @@ import cws.core.SimpleJobFactory;
 import cws.core.VM;
 import cws.core.WorkflowEngine;
 import cws.core.WorkflowEvent;
+import cws.core.algorithms.Algorithm;
 import cws.core.dag.DAG;
 import cws.core.dag.DAGParser;
+import cws.core.experiment.AlgorithmFactory;
+import cws.core.experiment.DAGListGenerator;
 import cws.core.experiment.Experiment;
 import cws.core.experiment.ExperimentDescription;
 import cws.core.experiment.ExperimentResult;
@@ -75,8 +79,46 @@ public class DynamicProvisionerDynamicSchedulerTest implements WorkflowEvent {
 
 		assertEquals(0, engine.getQueuedJobs().size());
 		
-		wfLog.printJobs("testEnsembleProvisionerDynamicSchedulerDag_CyberShake_100x10");
-		wfLog.printVmList("testEnsembleProvisionerDynamicSchedulerDag_CyberShake_100x10");
+		wfLog.printJobs("output/testEnsembleProvisionerDynamicSchedulerDag_CyberShake_100x10");
+		wfLog.printVmList("output/testEnsembleProvisionerDynamicSchedulerDag_CyberShake_100x10");
+	}
+	
+	
+	@Test
+	public void testGantt() {
+		String dagName = "CyberShake_100.dag";
+		double deadline = 3600.0; //seconds
+		double budget = 10.0;
+		double price = 1.0;
+		int numDAGs = 10;
+		double max_scaling = 1.0;
+		
+		
+		String dags[] = new String[numDAGs];
+		for (int i=0; i< numDAGs; i++) dags[i] = dagName;
+		
+        ExperimentDescription param = new ExperimentDescription("test", "DPDS", "output", dagPath, dags,
+                deadline, budget, price, max_scaling, 0.7, 1, 0.0, 0.0, "constant", 0);     
+		Algorithm algorithm = AlgorithmFactory.createAlgorithm(param);
+		algorithm.setGenerateLog(true);
+		String fileName = param.getRunDirectory() + File.separator + "output-test-DPDS";
+		algorithm.simulate(fileName);
+		
[…]

### 2012-08-02 paper/paper.tex
added ack

@@
     \subfloat[Ratio of Simulated Cost to Budget]{
-        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-cost-all}
+        \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-cost-all}
         }
@@
     \subfloat[Ratio of Simulated Makespan to Deadline]{
-        \includegraphics[width=0.4\textwidth]{run-finish-variations-test-0-output-makespan-all}
+        \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-makespan-all}
         }
@@
     \subfloat[Ratio of Simulated Ensemble Cost to Budget]{
-        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-cost-montage}
+        \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-cost-montage}
         }
@@
     \subfloat[Ratio of Simulated Ensemble Makespan to Deadline]{    
-        \includegraphics[width=0.4\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
+        \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
         }
@@
 \section*{Acknowledgments}
-This work was sponsored by the National Science Foundation under award 
-OCI-0943725 (STCI).
+This work was supported by the NSF under award OCI-0943725 (STCI) and
+by the EU FP7 projects UrbanFlood (248767) and VPH-Share (269978).
 

### 2012-08-08 paper/paper.tex
add ref to simulator
Add Ewa's changes

@@
 Workflows in an ensemble may differ not only in their parameters, but also in
-their priority. For example, in CyberShake some sites may be in heavily
+their priority. For example, in CyberShake some geographic locations may be in heavily
 populated areas or in strategic locations such as power plants, while others
@@
 tasks in level $l$, $R(w)$ is the total runtime of all tasks in the workflow,
-$R(w)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a
+$R(l)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a
 parameter between 0 and 1 that causes more slack time to be given to levels
@@
 To evaluate and compare the three proposed algorithms, we developed a cloud
-workflow simulator based on CloudSim~\cite{Calheiros2011}. Our simulation model
-consists of Cloud, VM and WorkflowEngine entities. The Cloud entity starts and
+workflow simulator~\cite{CloudWorkflowSimulator} based on CloudSim~\cite{Calheiros2011}. Our
+simulation model consists of Cloud, VM and WorkflowEngine entities. The Cloud entity starts and
 terminates VM entities using an Amazon EC2-like API. VM entities simulate the
@@
 workflows each were created. Each ensemble was simulated with all three
-algorithms using 10 budgets and 10 deadlines (1000 simulations per
-application, ensemble type, and algorithm). The best scores percentage is
+algorithms using 10 budgets and 10 deadlines (1,000 simulations per
+application, ensemble type, and algorithm, or 75,000 in total). The best scores percentage is
 computed by counting the number of times that a given algorithm achieved the
@@
         \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-cost-all}
+        \label{fig:variances-budget}
     }
@@
         \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-makespan-all}
+        \label{fig:variances-deadline}
     }
@@
 characteristics from preliminary runs~\cite{Bharathi2008,Deelman2005,Juve2010}.
-Some applications, like Periodograms~\cite{Vockler2011} include even a
+Some applications, like Periodograms~\cite{Vockler2011} even include a
 performance model that estimates task runtimes and automatically annotates
@@
 
-Fig.~\ref{fig:variances}.a shows the ratio of simulated cost to budget. This
+Fig.~\ref{fig:variances-budget} shows the ratio of simulated cost to budget. This
 plot illustrates two important algorithm characteristics. First, the dynamic
@@
 
-Fig.~\ref{fig:variances}.b shows the ratio of simulated makespan to budget.
+Fig.~\ref{fig:variances-deadline} shows the ratio of simulated makespan to budget.
 Interestingly, the deadline constraint is rarely exceeded, even in cases with
@@
 late. This is illustrated by the gantt chart in
-Fig.~\ref{fig:algorithm-example}.b, which shows how SPSS tends to pile up
+Fig.~\ref{fig:algorithm-example-spss}, which shows how SPSS tends to pile up
 workflows at the beginning of the timeline.
 % In comparison, the dynamic algorithms tend to spread out workflow starts over
-% the entire duration as shown in Fig.~\ref{fig:algorithm-example}.a. The
+% the entire duration as shown in Fig.~\ref{fig:algorithm-example-dpds}. The
 % consequence of this behavior is that SPSS plans tend to use up the budget but
@@
         \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-cost-montage}
+        \label{fig:delays-budget}
     }
@@
         \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
+        \label{fig:delays-deadline}
     }

### 2013-04-07 src/cws/core/experiment/Experiment.java
Finally got rid of all compiler's warnings, cleaned up code here and there and added some TODOs

@@ -94,22 +94,6 @@ public class Experiment {
     }
 
     /**
-     * TODO(bryk): Why this method is not used?
-     * Helper method to read DAX or DAG file format.
-     * XML-based DAX seems to be 10x slower.
-     * @param file
-     * @return DAG object
-     */
-    private static DAG parse(File file) {
-        if (file.getName().endsWith("dag"))
-            return DAGParser.parseDAG(file);
-        else if (file.getName().endsWith("dax"))
-            return DAGParser.parseDAX(file);
-        else
-            throw new RuntimeException("Unrecognized file: " + file.getName());
-    }
-
-    /**
      * Generates a series of experiments for varying deadlines
      * @param runDirectory the directory with input and output files for this series
      * @param group prefix to prepend to generated output files

### 2013-04-09 paper/paper.tex
Re-add granularity section

@@
 
-%\subsection{Task Granularity}
-%\TODO{Should we remove this section?}
-%
-%In the last experiment we noted that, for Montage and CyberShake, the short
-%runtimes of their tasks made the dynamic algorithms perform better relative to
-%SPSS. In order to test this theory we adjusted the granularity of the tasks in
-%several Montage and CyberShake ensembles to see how this would affect the
-%relative performance of the algorithms. The granularity adjustment was
-%achieved by multiplying the runtime of each task by a fixed scaling factor.
-
-%Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
-%as the scaling factor is increased from 1 to 16. The figure only shows results
-%for the CyberShake application with Pareto unsorted ensembles, but similar
-%results were found for other ensemble types for both Montage and CyberShake.
-%Each data point in the figure represents 500 simulations (5 ensembles with 10
-%budgets and 10 deadlines). The best scores percent was calculated the same way
-%as it was for Fig.~\ref{fig:distributions}.
-
-%The figure shows that, as the scaling factor increases beyond 2, the relative
-%performance of SPSS surpasses that of the dynamic algorithms. Similar results
-%were found for other ensemble types for both applications. This result
-%suggests that, in general, for fine-grained workflows the dynamic algorithms
-%will produce better scores, and for coarse-grained workflows SPSS will produce
-%better scores.
-
-%\begin{figure}[tb]
- %   \centering
- %   \includegraphics[width=0.3\textwidth]{stretching_cybershake}
- %   \caption{Percentage of high scores achieved by each algorithm for Pareto 
- %   unsorted CyberShake ensembles when task runtime is scaled by a constant 
- %   factor. A scaling factor of $x$ means that the runtime of tasks in each 
- %   workflow was multiplied by $x$.}
-%    \label{fig:stretching}
-%\end{figure}
+\subsection{Task Granularity}
+
+In the last experiment we noted that, for Montage and CyberShake, the short
+runtimes of their tasks made the dynamic algorithms perform better relative to
+SPSS. In order to test this theory we adjusted the granularity of the tasks in
+several Montage and CyberShake ensembles to see how this would affect the
+relative performance of the algorithms. The granularity adjustment was
+achieved by multiplying the runtime of each task by a fixed scaling factor.
+
+Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
+as the scaling factor is increased from 1 to 16. The figure only shows results
+for the CyberShake application with Pareto unsorted ensembles, but similar
+results were found for other ensemble types for both Montage and CyberShake.
+Each data point in the figure represents 500 simulations (5 ensembles with 10
+budgets and 10 deadlines). The best scores percent was calculated the same way
+as it was for Fig.~\ref{fig:distributions}.
+
+The figure shows that, as the scaling factor increases beyond 2, the relative
+performance of SPSS surpasses that of the dynamic algorithms. Similar results
+were found for other ensemble types for both applications. This result
+suggests that, in general, for fine-grained workflows the dynamic algorithms
+will produce better scores, and for coarse-grained workflows SPSS will produce
+better scores.
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics[width=0.3\textwidth]{stretching_cybershake}
+    \caption{Percentage of high scores achieved by each algorithm for Pareto 
+    unsorted CyberShake ensembles when task runtime is scaled by a constant 
+    factor. A scaling factor of $x$ means that the runtime of tasks in each 
+    workflow was multiplied by $x$.}
+    \label{fig:stretching}
+\end{figure}
 

### 2013-04-09 paper/paper.tex
Re-add sections that were commented-out

@@
 
-% Ewa: the following paragraph can be deleted if necessary
-%The paper is organized as follows: after an analysis of related work in
-%Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
-%infrastructure and application model we are targeting in this paper. In
-%Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
-%developed. Section~\ref{sec:performance} presents test scenarios and performance
-%metrics, while Section~\ref{sec:results} discusses the results of simulation
-%studies of these algorithms. Finally, general conclusions, lessons learned, and
-%future work are outlined in Section~\ref{sec:conclusions}.
+The paper is organized as follows: after an analysis of related work in
+Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
+infrastructure and application model we are targeting in this paper. In
+Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
+developed. Section~\ref{sec:performance} presents test scenarios and performance
+metrics, while Section~\ref{sec:results} discusses the results of simulation
+studies of these algorithms. Finally, general conclusions, lessons learned, and
+future work are outlined in Section~\ref{sec:conclusions}.
 
@@
 $$ 2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i} $$
-%  This scoring is consistent with our definition of the problem, which is to
-% complete as many workflows as possible, according to their priorities, given a
-% set budget and deadline.
+This scoring is consistent with our definition of the problem, which is to
+complete as many workflows as possible, according to their priorities, given a
+set budget and deadline.
 
@@
 large and costly workflows that would overrun the budget and admit smaller
-workflows that can efficiently utilize idle resources.
-
-% in 
-%ensembles containing workflows of non-uniform sizes. It would also be possible 
-%to extend this admission procedure to check other constraints, such as whether
-%the estimated critical path of the new workflow exceeds the time remaining
-%until the deadline.
+workflows that can efficiently utilize idle resources in 
+ensembles containing workflows of non-uniform sizes. It would also be possible 
+to extend this admission procedure to check other constraints, such as whether
+the estimated critical path of the new workflow exceeds the time remaining
+until the deadline.
 
@@
 
-%The approach used by SPSS is to plan each workflow in the ensemble in priority
-%order, rejecting any workflows that would exceed the dedline or budget. Once the
-%plan is complete, the VMs are provisioned and tasks are executed according to
-%the schedules given by the plan.
-
-%The disadvantage of the static planning approach used by SPSS is that it is
-%sensitive to dynamic changes in the environment and the application that may
-%disrupt its carefully constructed plan. For example, if there are provisioning
-%delays, or if the runtime estimates for the tasks are inaccurate, then workflow
-%execution may diverge from the plan. This issue will be discussed further in
-%Sections~\ref{sec:delays} and \ref{sec:variances}.
+The approach used by SPSS is to plan each workflow in the ensemble in priority
+order, rejecting any workflows that would exceed the dedline or budget. Once the
+plan is complete, the VMs are provisioned and tasks are executed according to
+the schedules given by the plan.
+
+The disadvantage of the static planning approach used by SPSS is that it is
+sensitive to dynamic changes in the environment and the application that may
+disrupt its carefully constructed plan. For example, if there are provisioning
+delays, or if the runtime estimates for the tasks are inaccurate, then workflow
+execution may diverge from the plan. This issue will be discussed further in
+Sections~\ref{sec:delays} and \ref{sec:variances}.
 
@@
 deadline and must be rejected. 
-%For large ensembles we expect the critical path
-%of any individual workflow to be much less than the deadline.
+For large ensembles we expect the critical path
+of any individual workflow to be much less than the deadline.
 
@@
 many VMs from being provisioned, the algorithm always prefers to extend the
-runtime of existing VMs before allocating new VMs. 
-%The result of this is that
-%the algorithm will only allocate a new block if there are no idle slots on
-%existing blocks large enough or early enough to complete the task by its
-%deadline, and it will only allocate a new VM if it cannot add a block to an
-%existing VM to complete the task by its deadline.
+runtime of existing VMs before allocating new VMs. The result of this is that
+the algorithm will only allocate a new block if there are no idle slots on
+existing blocks large enough or early enough to complete the task by its
+deadline, and it will only allocate a new VM if it cannot add a block to an
+existing VM to complete the task by its deadline.
 
@@
 and execute tasks sequentially. 
-%Although CloudSim provides a more advanced infrastructure model which includes
-%time-sharing and space-sharing policies, we do not use these features since we
-%are interested mainly in the execution of tasks on VMs and high-level workflow
-%scheduling and provisioning. 
+Although CloudSim provides a more advanced infrastructure model which includes
+time-sharing and space-sharing policies, we do not use these features since we
+are interested mainly in the execution of tasks on VMs and high-level workflow
+scheduling and provisioning. 
 The simulator reads workflow description files in
@@
 computational workloads. 
-%An example of the distribution of workflow sizes that
-%occurs in a Pareto ensemble is shown in Fig.~\ref{fig:ensemble-pareto}.
-
-%\begin{figure}[ht] 
-%    \centering
-%    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
-%    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of
-%    workflow sizes in Pareto ensembles. Workflow size is measured in number of
-%    tasks.}
-%    \label{fig:ensemble-pareto}
-%\end{figure}
+An example of the distribution of workflow sizes that
+occurs in a Pareto ensemble is shown in Fig.~\ref{fig:ensemble-pareto}.
+
+\begin{figure}[ht] 
+    \centering
+    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
+    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of
+    workflow sizes in Pareto ensembles. Workflow size is measured in number of
+    tasks.}
+    \label{fig:ensemble-pareto}
+\end{figure}
 
@@
 workflows, typical of the real applications we have examined (see Section~\ref{sec:intro}). 
-%For example, the number of
-%geographical sites of interest to the users of the CyberShake application in
-%the past has been on the order of 100. 
-%Second, smaller ensembles consisting of
-%just a few workflows can be aggregated into a single workflow, so there is no need to
-%treat them as an ensemble. Similarly, when the number of workflows grows and
-%each workflow has a large number of tasks, either the deadline and budget
-%constraints are low enough to prevent many of the workflows from running, or the
-%problem of efficiently allocating them to the resources becomes similar to a
-%bag-of-tasks problem, which is easier to solve efficiently.
+For example, the number of
+geographical sites of interest to the users of the CyberShake application in
+the past has been on the order of 100. 
+Second, smaller ensembles consisting of
+just a few workflows can be aggregated into a single workflow, so there is no need to
+treat them as an ensemble. Similarly, when the number of workflows grows and
+each workflow has a large number of tasks, either the deadline and budget
+constraints are low enough to prevent many of the workflows from running, or the
+problem of efficiently allocating them to the resources becomes similar to a
+bag-of-tasks problem, which is easier to solve efficiently.
 
@@
 
-%The SPSS plan describes what tasks to execute on which VMs, but not when to
-%execute them. At runtime, the worklow engine is forced to spend extra money
-%to extend the runtime of some VMs so that all of the tasks assigned to that
-%VM can be completed. At the same time, other VMs can be terminated early
-%because the tasks they were assigned finished earlier than expected. However,
-%because of dependencies in the workflow, the latter case is less likely to
-%happen, which causes gaps in the schedule. So the net result is that the
-%overall cost is increased.
+The SPSS plan describes what tasks to execute on which VMs, but not when to
+execute them. At runtime, the worklow engine is forced to spend extra money
+to extend the runtime of some VMs so that all of the tasks assigned to that
+VM can be completed. At the same time, other VMs can be terminated early
+because the tasks they were assigned finished earlier than expected. However,
+because of dependencies in the workflow, the latter case is less likely to
+happen, which causes gaps in the schedule. So the net result is that the
+overall cost is increased.
 
@@
 workflows at the beginning of the timeline.
-% In comparison, the dynamic algorithms tend to spread out workflow starts over
-% the entire duration as shown in Fig.~\ref{fig:algorithm-example-dpds}. The
-% consequence of this behavior is that SPSS plans tend to use up the budget but
-% leave plenty of time before the deadline.
+In comparison, the dynamic algorithms tend to spread out workflow starts over
+the entire duration as shown in Fig.~\ref{fig:algorithm-example-dpds}. The
+consequence of this behavior is that SPSS plans tend to use up the budget but
+leave plenty of time before the deadline.
 As a result, when the runtime of the plan is increased by introducing errors,
@@
 future work. 
-%The goal here is to determine how well the current algorithms are
-%able to stay within the constraints given inaccurate task runtime estimates.
-
-%It is possible that the static algorithm could be modified to add more
-%breathing room in the schedule to account for situations where the runtime
-%estimates may be inaccurate. Such a modification would result in slightly
-%worse scores when the estimates are good, but would improve scores when the
-%estimates are bad. Alternatively, the SPSS algorithm could be modified to
-%generate a plan that specifies when to provision and deprovision each VM. In
-%that case the workflow engine could be prevented from exceeding the budget
-%and deadline constraints, but would prevent some workflows from finishing,
-%which may significantly decrease the score. These two modifications are
-%subjects for future work.
+The goal here is to determine how well the current algorithms are
+able to stay within the constraints given inaccurate task runtime estimates.
+
+It is possible that the static algorithm could be modified to add more
+breathing room in the schedule to account for situations where the runtime
+estimates may be inaccurate. Such a modification would result in slightly
+worse scores when the estimates are good, but would improve scores when the
+estimates are bad. Alternatively, the SPSS algorithm could be modified to
+generate a plan that specifies when to provision and deprovision each VM. In
+that case the workflow engine could be prevented from exceeding the budget
+and deadline constraints, but would prevent some workflows from finishing,
+which may significantly decrease the score. These two modifications are
+subjects for future work.
 
@@
 
-%As in Fig.~\ref{fig:variances}, the y-axis in each
-%plot represents the ratio of the simulated value to the constraint value for
-%the budget constraint or the deadline constraint.
+As in Fig.~\ref{fig:variances}, the y-axis in each
+plot represents the ratio of the simulated value to the constraint value for
+the budget constraint or the deadline constraint.
 
@@
 
-% Fig.~\ref{fig:delays_scores} shows the relative performance as a
-% percentage of the number of high scores achieved by each algorithm as the
-% provisioning delay increases. This figure shows that, as the delay
-% increases, the relative performance of the static algorithm increases as
-% well. This behavior is a result of the fact that the static algorithm is, in
-% essence, cheating by using more time and money than the dynamic algorithms.
-% In the previous section we showed how the relative performance of the
-% algorithms remains the same when a uniform error is applied to the runtime
-% constraints. In that case, the dynamic algorithms adapted to the error by
-% rearranging the schedule to accomplish the same amount of work given the
-% same deadline and budget. In this case, the dynamic algorithms adapt by
-% performing less work (executing fewer workflows) to remain within the
-% constraints while accounting for the delays. In both cases, the static
-% algorithm completed the same amount of work, but did so by exceeding the
-% constraints.
-
-% \begin{figure}[ht]
-%   \centering
-%   \includegraphics[width=0.5\textwidth]{delays_scores}
-%   \caption{Percentage of high scores achieved by each algorithm on Montage 
-%   with uniform unsorted ensembles when provisioning delay varies from 0 
-%   seconds to 15 minutes.}
-%   \label{fig:delays_scores}
-% \end{figure}
+Fig.~\ref{fig:delays_scores} shows the relative performance as a
+percentage of the number of high scores achieved by each algorithm as the
+provisioning delay increases. This figure shows that, as the delay
+increases, the relative performance of the static algorithm increases as
+well. This behavior is a result of the fact that the static algorithm is, in
+essence, cheating by using more time and money than the dynamic algorithms.
+In the previous section we showed how the relative performance of the
+algorithms remains the same when a uniform error is applied to the runtime
+constraints. In that case, the dynamic algorithms adapted to the error by
+rearranging the schedule to accomplish the same amount of work given the
+same deadline and budget. In this case, the dynamic algorithms adapt by
+performing less work (executing fewer workflows) to remain within the
+constraints while accounting for the delays. In both cases, the static
+algorithm completed the same amount of work, but did so by exceeding the
+constraints.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.5\textwidth]{delays_scores}
+  \caption{Percentage of high scores achieved by each algorithm on Montage 
+  with uniform unsorted ensembles when provisioning delay varies from 0 
+  seconds to 15 minutes.}
+  \label{fig:delays_scores}
+\end{figure}
 
@@
 
-%This problem differs from previous work on grid and utility grid scheduling
-%in that cloud infrastructures provide more control over provisioning, so that
-%the number of resources can be adjusted according to the requirements of the
-%application. Therefore the problem space becomes larger; it requires not only
-%an efficient mapping of tasks to available resources, but also the selection
-%of the best resource provisioning plan.
-
-%Formulating the problem as a maximization of the number of prioritized
-%workflows completed from the ensemble is also novel and requires workflows to
-%be admitted or rejected based on their estimated resource demands. We believe
-%that this bi-constrained problem is highly relevant because such constraints
-%are commonly imposed on many real-world projects. The approach is also
-%directly applicable to grid environments that provide resource reservations
-%and charge service units for resource use.
+This problem differs from previous work on grid and utility grid scheduling
+in that cloud infrastructures provide more control over provisioning, so that
+the number of resources can be adjusted according to the requirements of the
+application. Therefore the problem space becomes larger; it requires not only
+an efficient mapping of tasks to available resources, but also the selection
+of the best resource provisioning plan.
+
+Formulating the problem as a maximization of the number of prioritized
+workflows completed from the ensemble is also novel and requires workflows to
+be admitted or rejected based on their estimated resource demands. We believe
+that this bi-constrained problem is highly relevant because such constraints
+are commonly imposed on many real-world projects. The approach is also
+directly applicable to grid environments that provide resource reservations
+and charge service units for resource use.
 
@@
 
-%In addition, we found that SPSS tends to perform better on coarse-grained
-%workflows than the dynamic algorithms. For wide, fine-grained workflows, such
-%as CyberShake and Montage, however, the dynamic algorithms frequently produce
-%performance that is as good or better than SPSS, because they are better able
-%to pack the smaller tasks onto idle VMs close to the deadline than SPSS,
-%which distributes deadlines to tasks in a way that prevents them from
-%starting late.
-
-%For very large workflows and ensembles, we found that the planning time of
-%the SPSS algorithm is prohibitive. SPSS often took 10 minutes or more to plan
-%ensembles of 100 workflows with 800 tasks each. This suggests that ensembles
-%of workflows with tens of thousands of tasks, which are commonly encountered
-%in real workflow applications, would take many hours to plan using SPSS. This
-%makes the dynamic algorithms much more attractive for large scale problems.
+In addition, we found that SPSS tends to perform better on coarse-grained
+workflows than the dynamic algorithms. For wide, fine-grained workflows, such
+as CyberShake and Montage, however, the dynamic algorithms frequently produce
+performance that is as good or better than SPSS, because they are better able
+to pack the smaller tasks onto idle VMs close to the deadline than SPSS,
+which distributes deadlines to tasks in a way that prevents them from
+starting late.
+
+For very large workflows and ensembles, we found that the planning time of
+the SPSS algorithm is prohibitive. SPSS often took 10 minutes or more to plan
+ensembles of 100 workflows with 800 tasks each. This suggests that ensembles
+of workflows with tens of thousands of tasks, which are commonly encountered
+in real workflow applications, would take many hours to plan using SPSS. This
+makes the dynamic algorithms much more attractive for large scale problems.
 
@@
 
-%The results of this study can be applied to develop tools that assist
-%researchers in planning their large-scale computational experiments. The
-%estimates of cost, runtime, and number of workflows completed that can be
-%obtained from both the static algorithms and from the simulation runs,
-%constitute valuable hints for planning ensembles and evaluating the
-%associated trade-offs.
+The results of this study can be applied to develop tools that assist
+researchers in planning their large-scale computational experiments. The
+estimates of cost, runtime, and number of workflows completed that can be
+obtained from both the static algorithms and from the simulation runs,
+constitute valuable hints for planning ensembles and evaluating the
+associated trade-offs.
 
@@
 % This balances the columns by inserting a column break at the given reference
-\IEEEtriggeratref{15}
+%\IEEEtriggeratref{15}
 

### 2013-04-09 paper/paper.tex
Convert paper to journal format

@@
-\documentclass[conference]{IEEEtran}
+\documentclass[preprint,12pt]{elsarticle}
+%\documentclass[final,1p,times,twocolumn]{elsarticle}
 \usepackage{fancyhdr}
-\usepackage[pdftex]{graphicx}
+\usepackage{graphicx}
 \graphicspath{{./figures/}}
@@
 
+\journal{Future Generation Computer Systems}
+
 \begin{document}
+
+\begin{frontmatter}
 
@@
 
-% IEEE Format
-\author{\IEEEauthorblockN{Maciej
-Malawski\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Gideon
-Juve\IEEEauthorrefmark{2}, Ewa Deelman\IEEEauthorrefmark{2} and Jarek
-Nabrzyski\IEEEauthorrefmark{3}} 
-\IEEEauthorblockA{\IEEEauthorrefmark{1}AGH University of Science and
-Technology, Dept. of Computer Science, Krakow, Poland, Email:
-malawski@agh.edu.pl} 
-\IEEEauthorblockA{\IEEEauthorrefmark{2}USC Information Sciences Institute, 4676
-Admiralty Way, Marina del Rey, CA, USA, Email: \{gideon,deelman\}@isi.edu} 
-\IEEEauthorblockA{\IEEEauthorrefmark{3}University of Notre Dame, Center for
-Research Computing, 111 ITC, Notre Dame, IN, USA, Email: naber@nd.edu}
-}
-
-% ACM Format
-%\numberofauthors{3}
-%\author{
-%    \alignauthor Maciej Malawski\\
-%       \affaddr{AGH University of Science and Technology}\\
-%       \affaddr{Dept. of Computer Science}\\
-%       \affaddr{Krakow, Poland; and}\\
-%       \affaddr{University of Notre Dame}\\
-%       \affaddr{Center for Research Computing}\\
-%       \affaddr{111 ITC, Notre Dame, IN, USA}\\
-%       \email{mmalawsk@nd.edu}
-%    \alignauthor Gideon Juve and Ewa Deelman\\
-%       \affaddr{USC Information Sciences Institute}\\
-%       \affaddr{4676 Admiralty Way}\\
-%       \affaddr{Marina del Rey, CA, USA}\\
-%       \email{\{gideon,deelman\}@isi.edu}
-%    \alignauthor Jarek Nabrzyski\\
-%       \affaddr{University of Notre Dame}\\
-%       \affaddr{Center for Research Computing}\\
-%       \affaddr{111 ITC, Notre Dame, IN, USA}\\
-%       \email{naber@nd.edu}
-%}
-
-\maketitle
-
-\thispagestyle{fancy}
-\lhead{}
-\rhead{}
-\chead{}
-\lfoot{\footnotesize{
-SC12, November 10-16, 2012, Salt Lake City, Utah, USA
-\newline
-978-1-4673-0806-9/12/\$31.00 \copyright 2012 IEEE}}
-\rfoot{}
-\cfoot{}
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrulewidth}{0pt}
+\author[agh]{Maciej Malawski}
+\ead{malawski@agh.edu.pl}
+\author[isi]{Gideon Juve}
+\ead{gideon@isi.edu}
+\author[isi]{Ewa Deelman}
+\ead{deelman@isi.edu}
+\author[nd]{Jarek Nabrzyski}
+\ead{naber@nd.edu}
+
+\address[agh]{AGH University of Science and Technology, Dept. of Computer Science, Krakow, Poland} 
+\address[isi]{USC Information Sciences Institute, 4676 Admiralty Way, Marina del Rey, CA, USA} 
+\address[nd]{University of Notre Dame, Center for Research Computing, 111 ITC, Notre Dame, IN, USA}
 
@@
 
-% IEEE does not have these
-%\category{H.3.4}{Systems and Software}{Distributed systems}
-%\category{D.4.8}{Performance}{Simulation}
-%\terms{DAG, Scheduling, Simulation}
-%\keywords{Scientific workflows, Workflow ensembles, Cloud computing, Infrastructure-as-a-Service}
+\begin{keyword}
+scientific workflows, workflow ensembles, cloud computing, resource provisioning
+\end{keyword}
+
+\end{frontmatter}
 
@@
 % IEEE style
-\bibliographystyle{IEEEtran}
+%\bibliographystyle{IEEEtran}
+
+% FGCS style
+\bibliographystyle{elsarticle-num}
 

### 2013-07-12 paper/paper.tex
New plots for FGCS paper

@@
-\documentclass[preprint,12pt]{elsarticle}
+\documentclass[preprint,5p]{elsarticle}
 %\documentclass[final,1p,times,twocolumn]{elsarticle}
@@
 
+This paper is an extension of our earlier conference
+publication~\cite{Malawski-SC12}.
+Here we present a more detailed performance evaluation, including the discussion
+of results based on deadline and budget constraints. We also report on new
+results and discussion of task granularity in the context of the resource
+billing cycle, which is important aspect of scheduling and resource provisioning
+on cloud infrastructures.
+
 The paper is organized as follows: after an analysis of related work in
@@
 cost-minimization workflow scheduling, developed for utility grid systems. However, our problem is
-different from \cite{Yu2005} and \cite{Abrishami2010} in that we consider ensembles of workflows in
+different from \cite{Yu2005} and \cite{Abrishami2013158} in that we consider ensembles of workflows in
 IaaS clouds, which allow one to provision resources on a per-hour billing model,
@@
   \subfloat[DPDS]{
-  \includegraphics[width=0.40\textwidth]{figures/output-test-DPDS}
+  \includegraphics[width=0.80\columnwidth]{figures/output-test-DPDS}
   \label{fig:algorithm-example-dpds}
@@
   \subfloat[SPSS]{
-  \includegraphics[width=0.40\textwidth]{figures/output-test-SPSS}
+  \includegraphics[width=0.80\columnwidth]{figures/output-test-SPSS}
   \label{fig:algorithm-example-spss}
@@
 \subsection{Relative Performance of Algorithms}
+\label{sec:relative-performance}
 
@@
 
+\subsection{Results vs. deadline and budget}
+
+In order to give more details about the performance of the algorithms, in this
+section we show how the results depend on the deadline and budget constraints.
+As the performance metric we use the number of completed workflows of the
+ensemble, since the exponential score is not suitable for showing trends.
+We use all deadline and budget parameters as defined in
+Section~\ref{sec:exp-parameters} and normalize them to the range from 0 to 1 to
+facilitate plotting.
+For each normalized deadline, we compute the mean number of completed workflows,
+averaged over all budgets and random seeds, so each point on the plot is based
+on 100 different simulation runs. In Fig.~\ref{fig:completed-deadlines}
+and~\ref{fig:completed-budgets} we present the results of CyberShake and LIGO
+applications, as representatives of fine-grained and coarse-grained workflows.
+Similar results were obtained for other applications.
+
+Fig.~\ref{fig:completed-deadlines} shows the results depending on normalized
+deadline. It can be observed, that with increasing deadline, the number of
+completed worklfows grows steeply and then stabilizes. This means that for
+longer deadlines it is usually not possible to complete more workflows due to
+the budget constraint. For short deadlines, the algorithms need to allocate more
+VM instances in parallel, making it more difficult to utilize the resources
+efficiently, which results in lower numbers of completed workflows.
+
+We can also observe that the algorithms are able to complete more workflows in
+the ensembles that have priorities assigned randomly (unsorted), in comparison
+to the ensembes where the priorities are assigned based on workflow size
+(sorted). This results from the fact that in the unsoted distributions the large
+workflows often have low priorities, so they can be rejected by the algoritms,
+thus making more space in the schedule for smaller high-priority workflows.
+
+One interesting observation from Fig.~\ref{fig:completed-deadlines} is that for
+CyberShake application the result curves have a sawtooth shape. This results
+from the combination of two effects. The first comes from the fact that dynamic
+algorithms perform better when the deadline is close to the full hour, due to
+the provisioning mechanism that tries to shut down unused VM instances before a
+full hour. The second one is that when the deadline increases, it may become
+possible to admit a larger workflow with higher priority instead of several
+smaller ones with lower priority. That is why the number of completed workflows
+may decrese, but the exponential score (utility function) based on priorities
+will be higher. These effects are visible for CyberShake application, for which
+the maximum deadline is 3 hours, but they become less important for LIGO
+application, for which the maximum deadline is 20 hours. See also the discussion
+on task granularity in Setion~\ref{sec:granularity}.
+
+Fig~\ref{fig:completed-budgets} shows the results based on normalized budget,
+each point being the average over all deadline and random seed. What becomes
+apparent is that for constant and unsorted distributions the number of completed
+workflows is almost linear. This confirms that the budget constraint has the
+linear influence on the results, when the influence of the deadline is
+eliminated by averaging. On the other hand, for sorted distributions the
+functions are convex, increasing slowly for small budgets and steeper for larger
+budgets. This reflects the fact that the in sorted ensembles the largest
+workflows have the highest priority, so it requires a large increment of a small
+budget to fir the next workflow into the schedule.
+
+The chacacteristics discussed in this section can be useful not only for
+analysis of our algorithms, but also for planning scientific experiments and
+understanding tradeoffs between cost, deadline and the number of work
+complete~\cite{Malawski-FGCS-13}.
+
+\begin{figure*}[t!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-deadlines-app-dist-selected}    
+    \caption{Number of completed workflows from the ensemble depending on normalized deadline.}
+    \label{fig:completed-deadlines}
+\end{figure*}
+
+\begin{figure*}[t!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-budgets-app-dist-selected}    
+    \caption{Number of completed workflows from the ensemble depending on normalized budget.}
+    \label{fig:completed-budgets}
+\end{figure*}
+
 \subsection{Task Granularity}
-
-In the last experiment we noted that, for Montage and CyberShake, the short
-runtimes of their tasks made the dynamic algorithms perform better relative to
-SPSS. In order to test this theory we adjusted the granularity of the tasks in
-several Montage and CyberShake ensembles to see how this would affect the
-relative performance of the algorithms. The granularity adjustment was
-achieved by multiplying the runtime of each task by a fixed scaling factor.
-
-Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
-as the scaling factor is increased from 1 to 16. The figure only shows results
-for the CyberShake application with Pareto unsorted ensembles, but similar
-results were found for other ensemble types for both Montage and CyberShake.
-Each data point in the figure represents 500 simulations (5 ensembles with 10
-budgets and 10 deadlines). The best scores percent was calculated the same way
-as it was for Fig.~\ref{fig:distributions}.
+\label{sec:granularity}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\columnwidth]{run-stretching-test-output-scaling-CYBERSHAKE}
+    \includegraphics[width=0.9\columnwidth]{run-stretching-test-output-scaling-MONTAGE}
+
+    \caption{Percentage of high scores achieved by each algorithm for CyberShake
+    and Montage ensembles when task runtime is scaled by a constant factor. A
+    scaling factor of $x$ means that the runtime of tasks in each workflow was
+    multiplied by $x$.}
+    \label{fig:stretching}
+\end{figure}
+
+In the experiments described in previous sections we noted that, for Montage and
+CyberShake, the short runtimes of their tasks made the dynamic algorithms
+perform better relative to SPSS. In order to test this theory we adjusted the
+granularity of the tasks in several Montage and CyberShake ensembles to see how
+this would affect the relative performance of the algorithms. The granularity
+adjustment was achieved by multiplying the runtime of each task by a fixed
+scaling factor.
+
+Fig.~\ref{fig:stretching} shows the relative performance of the algorithms as
+the scaling factor is increased from 1 to 16 for CyberShake  and Montage
+applications. Each data point in the figure represents 500 simulations (5
+ensembles with 10 budgets and 10 deadlines). The best scores percent was
+calculated the same way as it was for Fig.~\ref{fig:distributions}.
 
 The figure shows that, as the scaling factor increases beyond 2, the relative
-performance of SPSS surpasses that of the dynamic algorithms. Similar results
-were found for other ensemble types for both applications. This result
+performance of SPSS surpasses that of the dynamic algorithms.  This result
 suggests that, in general, for fine-grained workflows the dynamic algorithms
@@
 
-\begin{figure}[tb]
-    \centering
-    \includegraphics[width=0.3\textwidth]{stretching_cybershake}
-    \caption{Percentage of high scores achieved by each algorithm for Pareto 
-    unsorted CyberShake ensembles when task runtime is scaled by a constant 
-    factor. A scaling factor of $x$ means that the runtime of tasks in each 
-    workflow was multiplied by $x$.}
-    \label{fig:stretching}
-\end{figure}
+The granularity of the workflow tasks should be discussed in comparison with the
+length of the cloud billing cycle. In this paper we assumed that the resources
+are charged for hourly time slots, as is the case of Amazon EC2. However, there
+are cloud providers that have higher billing frequency, e.g. every 5 minutes in
+the case of CloudSigma\footnote{\url{http://www.cloudsigma.com}} or every 1
+minute in the case of Google Compute
+Engine\footnote{\url{https://cloud.google.com/pricing/compute-engine}} (after
+first 10 minutes). Our results indicate that when the task granularity is close
+to the length of the billing period, the SPSS produces better results, while for
+tasks that are shorter than the billing period, the dynamic algorithms perform
+better. This suggests that the these effects have to be taken into account when
+selecting a cloud provider, and also when using task clustering algorithms that
+group smaller tasks into larger groups~\cite{ChenD12}.
 

### 2013-07-12 scripts/R/data_cube_csv.R

@@ -1,6 +1,6 @@
-library(DBI)
-library(RSQLite)
-library(cwhmisc)
+#library(DBI)
+#library(RSQLite)
+#library(cwhmisc)
 
 
 

### 2013-07-12 scripts/R/data_cube_csv_budgets_deadlines.R

@@ -1,0 +1,122 @@
+
+csv_filename = "run-finish-variations-test-0-output.dat"
+
+
+alldata = read.csv(csv_filename, colClasses=c("scorebits"="character"))
+alldata = alldata[alldata$runtimeVariance==0,]
+
+attach(alldata)
+failureRates = unique(failureRate)
+#alldata$deadline = alldata$deadline/3600
+
+runIds = unique(seed)
+n_runIds=length(runIds)
+
+
+print(n_runIds)
+
+
+applications = levels(unique(application))
+
+naapplications = length(applications)
+
+distributions = levels(unique(distribution))
+print(distributions)
+ndistributions = length(distributions)
+
+algorithms = levels(unique(algorithm))
+print(algorithms)
+
+nalgorithms = length(algorithms)
+
+deadlines = unique(deadline)
+
+detach(alldata)
+
+mydata = alldata[c("application", "distribution", "seed", "budget", "deadline", "algorithm", "completed", "minDeadline", "maxDeadline")]
+
+# normalize deadlines to the values 1, 2, 3, ..., 10 for averaging
+mydata$deadline = round((mydata$deadline-mydata$minDeadline)/(mydata$maxDeadline-mydata$minDeadline)*90/10+1)
+
+deadlines = unique(mydata$deadline)
+print(deadline)
+
+
+attach(mydata)
+avgdata = aggregate( completed ~ application:distribution:deadline:algorithm, data=mydata, FUN=mean)
+detach(mydata)
+
+# renormalize deadlines to the values 0 to 1 for plotting
+avgdata$deadline = (avgdata$deadline-1)/90*10
+
+
+
+pdf(file=paste(csv_filename, "-deadlines-app-dist.pdf", sep=""), height=18, width=16, bg="white")
+par(mfrow=c(5,5))
+for (app in 1:naapplications) {
+for (j in 1:ndistributions) {
+	
+	plotdata = subset(avgdata, application==applications[app] & distribution==distributions[j])
+	
+	xrange = range(plotdata$deadline)
+	xrange[1]=0
+	yrange = range(plotdata$completed)
+	yrange[1]=0
+	
+	plot(xrange, yrange, type="n", xlab="Normalized deadline",
+			ylab="Completed Workflows", main = paste(c(applications[app],distributions[j])))
[…]

### 2013-07-12 scripts/ruby/generate_stretching.rb

@@ -1,0 +1,50 @@
+
+
+
+# Generates inputs for experiments with diferent failure rates
+# 
+# Author: Maciej Malawski
+
+prefix = "stretching-test"
+applications = ["CYBERSHAKE", "MONTAGE"]
+input_dirs = {"CYBERSHAKE" => "CyberShake", "GENOME" => "Genome", "LIGO" => "LIGO", "MONTAGE" => "Montage", "SIPHT" => "SIPHT"}
+algorithms = ["DPDS", "WADPDS", "SPSS"]
+distributions = ["uniform_unsorted", "uniform_sorted", "pareto_unsorted", "pareto_sorted", "constant"]
+delay = 0
+scaling_factors = [1, 2, 4, 8, 16]
+failure_rate = 0
+ensemble_size = 50
+variation = 0.0
+id = 0
+
+for seed in 0..9 do
+  task_id = 0
+  run_dir = "run-%s-%02d" % [prefix, seed]
+  begin 
+    Dir.mkdir run_dir
+  rescue
+  end
+  file_name = run_dir + "/" + run_dir+".txt"
+  f = File.open(file_name, "w")
+  for scaling_factor in scaling_factors do
+    for application in applications do
+      for algorithm in algorithms do
+        for distribution in distributions do
+          id = id + 1
+          task_id = task_id + 1
+          id_string = "%08d" % id
+          args = "-application #{application} -inputdir ../projects/pegasus/#{input_dirs[application]}/ -outputfile #{run_dir}/#{prefix}-output-#{id_string}.dat -distribution #{distribution} -ensembleSize #{ensemble_size} -algorithm #{algorithm} -seed #{seed} -failureRate #{failure_rate} -runtimeVariance #{variation} -delay #{delay} -scalingFactor #{scaling_factor}\n"
+          f.write args
+        end
+      end  
+    end
+  end
+  f.close
+  #puts "./scripts/run_test_local.sh " + file_name + " &"
+  puts "qsub -pe smp 2 -q short -t 1-#{task_id}  /afs/crc.nd.edu/user/m/mmalawsk/cloudworkflowsimulator/run_test_sge.sh " + file_name
+end
+
+puts `cd ~/cloudworkflowsimulator && ant`  
+puts `cp ~/cloudworkflowsimulator/dist/cloudworkflowsimulator.jar /afs/crc.nd.edu/user/m/mmalawsk/cloudworkflowsimulator/dist/`  
+
[…]

### 2013-08-30 paper/paper.tex
Update to paper after Ewa's comments

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Cost- and Deadline-Constrained Provisioning for Scientific Workflow
+\title{Algorithms for Cost- and Deadline-Constrained Provisioning for Scientific Workflow
 Ensembles in IaaS Clouds}
@@
 
-\address[agh]{AGH University of Science and Technology, Dept. of Computer Science, Krakow, Poland} 
+\address[agh]{AGH University of Science and Technology, Dept. of Computer Science, Al. Mickiewicza 30, Krakow, Poland} 
 \address[isi]{USC Information Sciences Institute, 4676 Admiralty Way, Marina del Rey, CA, USA} 
@@
 
-Large-scale applications expressed as scientific workflows are
-often grouped into ensembles of inter-related workflows. In this paper, we
-address a new and important problem concerning the efficient management of
-such ensembles under budget and deadline constraints on Infrastructure-
-as-a-Service (IaaS) clouds. We discuss, develop, and assess algorithms based
-on static and dynamic strategies for both task scheduling and resource
-provisioning. We perform the evaluation via simulation using a set of
-scientific workflow ensembles with a broad range of budget and deadline
-parameters, taking into account uncertainties in task runtime estimations,
+Large-scale applications expressed as scientific workflows are often grouped
+into ensembles of inter-related workflows. In this paper, we address a new and
+important problem concerning the efficient management of such ensembles under
+budget and deadline constraints on Infrastructure- as-a-Service (IaaS) clouds.
+IaaS clouds are characterized by on-demand resource provisioning capabilities
+and a pay-per-use model. We discuss, develop, and assess algorithms based on
+static and dynamic strategies for both task scheduling and resource
+provisioning. We perform the evaluation via simulation using a set of scientific
+workflow ensembles with a broad range of budget and deadline parameters, taking
+into account task granularity, uncertainties in task runtime estimations,
 provisioning delays, and failures. We find that the key factor determining the
@@
 \begin{keyword}
-scientific workflows, workflow ensembles, cloud computing, resource provisioning
+scientific workflows; workflow ensembles; cloud computing; resource provisioning
 \end{keyword}
@@
 
-Scientific workflows, usually represented as Directed Acyclic Graphs (DAGs),
-are an important class of applications that lead to challenging problems in
-resource management on grid and utility computing systems. Workflows for large
-computational problems are often composed of several inter-related workflows
-grouped into {\em ensembles}.  Workflows in an ensemble typically have a
-similar structure, but they differ in their input data, number of tasks, and
-individual task sizes.
+Large-scale scientific workflows, usually represented as Directed Acyclic Graphs
+(DAGs), are an important class of applications that lead to challenging problems
+in resource management on grid and cloud infrastructures. What becomes more
+important, workflows for large computational problems are often composed of
+several inter-related workflows grouped into {\em ensembles}.  Workflows in an
+ensemble typically have a similar structure, but they differ in their input
+data, number of tasks, and individual task sizes.
 
 There are many applications that require scientific workflow ensembles.
-CyberShake~\cite{Callaghan2011}, for example, uses ensembles to generate
-seismic hazard maps.  Each workflow in a CyberShake ensemble generates a
-hazard curve for a particular geographic location, and several hazard curves
-are combined to create a hazard map.  In 2009 CyberShake was used to generate
-a map that required an ensemble of 239 workflows.  Similarly, users of
+CyberShake~\cite{Callaghan2011}, for example, uses ensembles to generate seismic
+hazard maps.  Each workflow in a CyberShake ensemble generates a hazard curve
+for a particular geographic location, and several hazard curves are combined to
+create a hazard map.  In the recent study of 2013 CyberShake generates a map that
+requires an ensemble of 286 workflows.  Similarly, users of
 Montage~\cite{Deelman2008} often need several workflows  with different
@@
 Due to the large scale of the input data, this application is often split up
-into multiple batches processed by different workflows. Additional workflows
-are created to run the analysis using different parameters. A recent analysis
-of Kepler satellite data required three ensembles of 15 workflows.
+into multiple batches processed by different workflows. Additional workflows are
+created to run the analysis using different parameters. A recent analysis of
+Kepler satellite data required three ensembles of 15 workflows.
 
@@
 an ensemble under both budget and deadline constraints. The motivation for this
-work is to answer the fundamental question of concern to a researcher: How much
+work is to answer the fundamental question of concern to a researcher: {\em How much
 computation can be completed given the limited budget and timeframe of a
-research project?
+research project?}
 
@@
   CloudSim~\cite{Calheiros2011}, which models the infrastructure and the 
-  application,taking into account uncertainties in task runtime estimates, 
+  application, taking into account uncertainties in task runtime estimates, 
   provisioning delays, and failures, 
@@
 Here we present a more detailed performance evaluation, including the discussion
-of results based on deadline and budget constraints. We also report on new
+of results based on varying deadline and budget constraints. We also report on new
 results and discussion of task granularity in the context of the resource
-billing cycle, which is important aspect of scheduling and resource provisioning
+billing cycle, which is an important aspect of scheduling and resource provisioning
 on cloud infrastructures.
@@
 \label{sec:related}
-General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
+General policy-based and rule-based approaches to dynamic provisioning (such as Amazon
 Auto Scaling~\cite{Autoscaling} and
@@
 Our work is related to the strategies for deadline-con\-strained
-cost-minimization workflow scheduling, developed for utility grid systems. However, our problem is
-different from \cite{Yu2005} and \cite{Abrishami2013158} in that we consider ensembles of workflows in
-IaaS clouds, which allow one to provision resources on a per-hour billing model,
-rather than utility grids, which allow one to choose from a pool of existing
-resources with a per-job billing model. Our work is also different from
-cloud-targeted autoscaling solution~\cite{Mao2011} in that we consider ensembles
-of workflows rather than unpredictable workloads containing workflows. We also consider budget constraints
-rather than cost minimization as a goal. In other words, we assume that there is
-more work to be done than the available budget, so some work must be rejected.
-Therefore, cost is not something we optimize (i.e. an objective), but rather a constraint.
+cost-minimization workflow scheduling, developed for utility grid and cloud
+systems. However, our problem is different from
+\cite{Yu2005},\cite{Abrishami2013158} and~\cite{Young2013} in that we consider
+ensembles of workflows instead of individual workflows.
+Our work is also different from cloud-targeted autoscaling
+solution~\cite{Mao2011} in that we consider ensembles of workflows rather than
+unpredictable workloads containing workflows. We also consider budget
+constraints rather than cost minimization as a goal. In other words, we assume
+that there is more work to be done than the available budget, so some work must
+be rejected.
+Therefore, cost is not something we optimize (i.e. an objective), but rather a
+constraint.
 
 This work is related to bi-criteria scheduling and multi-criteria scheduling of
-workflows~\cite{Wieczorek2009,Prodan2010,Dongarra2007}. These approaches are
+workflows~\cite{Wieczorek2009,Prodan2010,Dongarra2007,Durillo2013}. These approaches are
 similar to ours in that we have two scheduling criteria: cost and makespan. The
@@
 of work completed) is subject to optimization, whereas time and cost are
-treated as constraints. Other approaches~\cite{Talukder2009,Pandey2010} use
+treated as constraints.
+
+Other approaches~\cite{Talukder2009,Pandey2010} use
 metaheuristics that usually run for a long time before producing good results,
@@
 
-\underline{Resource Model}  We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2),
+\subsection{Resource Model}  
+We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2),
 where virtual machine (VM) instances may be provisioned on-demand and are billed by
@@
 
-\underline{Application Model} The target applications are ensembles of
+\subsection{Application Model} 
+The target applications are ensembles of
 scientific workflows that can be modeled as DAGs, where the nodes in the graph
@@
 
-\underline{Performance Metric}
+\subsection{Performance Metric}
 \label{sec:perf_metric}
@@
 executed even though there is no chance that those workflows will be completed
-within the current budget and deadline. Fig.~\ref{fig:algorithm-example-dpds}
+within the current budget and deadline. Figure~\ref{fig:algorithm-example-dpds}
 shows an example schedule generated using the DPDS algorithm. The figure
 illustrates how tasks from lower priority workflows backfill idle VMs when
-tasks from higher priority workflows are not available.
+tasks from higher priority workflows are not ready for execution.
 
@@
 algorithms. However, the disadvantage of SPSS is that it is sensitive to dynamic
-changes in the environment (see Sections~\ref{sec:delays} and
-\ref{sec:variances}).
+changes in the environment (see Sections~\ref{sec:variances} and
+\ref{sec:delays}).
 
 The approach used by SPSS is to plan each workflow in the ensemble in priority
-order, rejecting any workflows that would exceed the dedline or budget. Once the
+order, rejecting any workflows that would exceed the deadline or budget. Once the
 plan is complete, the VMs are provisioned and tasks are executed according to
@@
 execution may diverge from the plan. This issue will be discussed further in
-Sections~\ref{sec:delays} and \ref{sec:variances}.
+Sections~\ref{sec:variances} and \ref{sec:delays}.
 
@@
 
-An example schedule generated by SPSS (Fig.~\ref{fig:algorithm-example-spss})
+An example schedule generated by SPSS (Figure~\ref{fig:algorithm-example-spss})
 shows how SPSS tends to start many workflows in parallel, running each workflow
@@
 and execute tasks sequentially. 
-Although CloudSim provides a more advanced infrastructure model which includes
+Although CloudSim provides a more advanced infrastructure model, which includes
 time-sharing and space-sharing policies, we do not use these features since we
@@
 An example of the distribution of workflow sizes that
-occurs in a Pareto ensemble is shown in Fig.~\ref{fig:ensemble-pareto}.
-
-\begin{figure}[ht] 
+occurs in a Pareto ensemble is shown in Figure~\ref{fig:ensemble-pareto}.
+
+\begin{figure}[t] 
     \centering
@@
 
+\begin{figure*}[tb]
+    \centering
+    \includegraphics[width=\textwidth]{run-finish-variations-test-0-output-distributions}
+    \caption{Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.
+C = Constant ensembles, PS = Pareto Sorted ensembles, PU = Pareto Unsorted ensembles, US = Uniform Sorted ensembles, and UU = Uniform Unsorted ensembles.}
+    \label{fig:distributions}
+\end{figure*}
+
 The goal of the first experiment is to characterize the relative performance
@@
 
-Fig.~\ref{fig:distributions} shows the percentage of simulations for which
+Figure~\ref{fig:distributions} shows the percentage of simulations for which
 each algorithm achieved the highest score for a given ensemble type. This
@@
 
-\begin{figure*}[ht]
-    \centering
-    \includegraphics[width=\textwidth]{run-finish-variations-test-0-output-distributions}
-    \caption{Percentage of high scores achieved by each algorithm on different ensemble types for all five applications.
-C = Constant ensembles, PS = Pareto Sorted ensembles, PU = Pareto Unsorted ensembles, US = Uniform Sorted ensembles, and UU = Uniform Unsorted ensembles.}
-    \label{fig:distributions}
-\end{figure*}
-
 There are several interesting things to notice about
-Fig.~\ref{fig:distributions}. The first is that, in most cases, SPSS
+Figure~\ref{fig:distributions}. The first is that, in most cases, SPSS
 significantly outperforms both dynamic algorithms (DPDS and WA-DPDS). This is
@@
 
-\subsection{Results vs. deadline and budget}
+\subsection{Results vs. varying deadline and budget}
+
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-deadlines-app-dist-selected}    
+    \caption{Number of completed workflows from the ensemble depending on normalized deadline.}
+    \label{fig:completed-deadlines}
+\end{figure*}
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-budgets-app-dist-selected}    
+    \caption{Number of completed workflows from the ensemble depending on normalized budget.}
+    \label{fig:completed-budgets}
+\end{figure*}
 
@@
 averaged over all budgets and random seeds, so each point on the plot is based
-on 100 different simulation runs. In Fig.~\ref{fig:completed-deadlines}
+on 100 different simulation runs. In Figure~\ref{fig:completed-deadlines}
 and~\ref{fig:completed-budgets} we present the results of CyberShake and LIGO
@@
 
-Fig.~\ref{fig:completed-deadlines} shows the results depending on normalized
+Figure~\ref{fig:completed-deadlines} shows the results depending on normalized
 deadline. It can be observed, that with increasing deadline, the number of
@@
 to the ensembes where the priorities are assigned based on workflow size
-(sorted). This results from the fact that in the unsoted distributions the large
+(sorted). This results from the fact that in the unsorted distributions the large
 workflows often have low priorities, so they can be rejected by the algoritms,
@@
 
-One interesting observation from Fig.~\ref{fig:completed-deadlines} is that for
+One interesting observation from Figure~\ref{fig:completed-deadlines} is that for
 CyberShake application the result curves have a sawtooth shape. This results
@@
 
-\begin{figure*}[t!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-deadlines-app-dist-selected}    
-    \caption{Number of completed workflows from the ensemble depending on normalized deadline.}
-    \label{fig:completed-deadlines}
-\end{figure*}
-
-\begin{figure*}[t!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-budgets-app-dist-selected}    
-    \caption{Number of completed workflows from the ensemble depending on normalized budget.}
-    \label{fig:completed-budgets}
-\end{figure*}
-
 \subsection{Task Granularity}
@@
 
-\begin{figure}[h!]
+\begin{figure*}[htb]
     \centering
-    \includegraphics[width=0.9\columnwidth]{run-stretching-test-output-scaling-CYBERSHAKE}
-    \includegraphics[width=0.9\columnwidth]{run-stretching-test-output-scaling-MONTAGE}
-
-    \caption{Percentage of high scores achieved by each algorithm for CyberShake
-    and Montage ensembles when task runtime is scaled by a constant factor. A
+%    \subfloat[CyberShake]{
+    \includegraphics[width=0.6\textwidth]{run-stretching-test-output-scaling-CYBERSHAKE}
+%    }\\
+%    \subfloat[Montage]{
+    \includegraphics[width=0.6\textwidth]{run-stretching-test-output-scaling-MONTAGE}
+%    }
+    \caption{Percentage of high scores achieved by each algorithm for CyberShake (top)
+    and Montage (bottom) ensembles when task runtime is scaled by a constant factor. A
     scaling factor of $x$ means that the runtime of tasks in each workflow was
@@
     \label{fig:stretching}
-\end{figure}
+\end{figure*}
 
@@
 
-Fig.~\ref{fig:stretching} shows the relative performance of the algorithms as
+Figure~\ref{fig:stretching} shows the relative performance of the algorithms as
 the scaling factor is increased from 1 to 16 for CyberShake  and Montage
@@
 ensembles with 10 budgets and 10 deadlines). The best scores percent was
-calculated the same way as it was for Fig.~\ref{fig:distributions}.
+calculated the same way as it was for Figure~\ref{fig:distributions}.
 
@@
 
-The granularity of the workflow tasks should be discussed in comparison with the
+The granularity of the workflow tasks should be discussed in the context of the
 length of the cloud billing cycle. In this paper we assumed that the resources
-are charged for hourly time slots, as is the case of Amazon EC2. However, there
-are cloud providers that have higher billing frequency, e.g. every 5 minutes in
-the case of CloudSigma\footnote{\url{http://www.cloudsigma.com}} or every 1
-minute in the case of Google Compute
+are charged for hourly time slots, as is the case of VM instances on Amazon EC2.
+However, there are cloud providers that have higher billing frequency, e.g.
+every 5 minutes in the case of
+CloudSigma\footnote{\url{http://www.cloudsigma.com}} or every 1 minute in the
+case of Google Compute
 Engine\footnote{\url{https://cloud.google.com/pricing/compute-engine}} (after
@@
 tasks that are shorter than the billing period, the dynamic algorithms perform
-better. This suggests that the these effects have to be taken into account when
+better. This suggests that these granularity effects have to be taken into account when
 selecting a cloud provider, and also when using task clustering algorithms that
@@
 
-\begin{figure*}[tb]
+\begin{figure*}[htb]
     \centering
@@
 \end{figure*}
+\begin{figure*}[htb]
+    \centering
+        \includegraphics[width=0.8\textwidth]{run-finish-variations-test-0-output-distributions-variances}
+    \caption{\label{fig:variances-scores}Percentage of high scores achieved by each algorithm on 
+    CyberShake and LIGO when runtime estimate error varies from 0\% to 50\%}
+\end{figure*}
 
@@
 
-Fig.~\ref{fig:variances} shows the results for estimate errors
+Figure~\ref{fig:variances} shows the results for estimate errors
 ranging from $0\%$ to $50\%$. The figure summarizes the outcome of an
@@
 
-Fig.~\ref{fig:variances-budget} shows the ratio of simulated cost to budget. This
+Figure~\ref{fig:variances-budget} shows the ratio of simulated cost to budget. This
 plot illustrates two important algorithm characteristics. First, the dynamic
@@
 
-Fig.~\ref{fig:variances-deadline} shows the ratio of simulated makespan to budget.
+Figure~\ref{fig:variances-deadline} shows the ratio of simulated makespan to budget.
 Interestingly, the deadline constraint is rarely exceeded, even in cases with
@@
 deadline distribution function in SPSS, which prevents workflows from starting
-late. This is illustrated by the gantt chart in
-Fig.~\ref{fig:algorithm-example-spss}, which shows how SPSS tends to pile up
+late. This is illustrated by the Gantt chart in
+Figure~\ref{fig:algorithm-example-spss}, which shows how SPSS tends to pile up
 workflows at the beginning of the timeline.
 In comparison, the dynamic algorithms tend to spread out workflow starts over
-the entire duration as shown in Fig.~\ref{fig:algorithm-example-dpds}. The
+the entire duration as shown in Figure~\ref{fig:algorithm-example-dpds}. The
 consequence of this behavior is that SPSS plans tend to use up the budget but
@@
 the SPSS plan has some room to expand without exceeding the deadline.
+
+Figure~\ref{fig:variances-scores} shows relative performance as a percentage of
+the number of high scores achieved by each algorithm as the error increases. As
+in Figure~\ref{fig:completed-budgets} and ~\ref{fig:completed-deadlines} we show
+results for CyberShake and LIGO as representatives of fine grained and coarse
+grained workflows. The figure shows that, for this experiment, the scores remain
+the same as the error increases. This is a result of the way the error is
+applied. Since the error is selected uniformly from $-p$ to $+p$ it is equally
+likely that the error will be positive as negative. Therefore, applying the
+error leaves the total runtime of the tasks in the workflow unchanged -- the
+total number of CPU hours in each workflow remains nearly the same regardless of
+the error. The end result is that the dynamic algorithms are able to achieve the
+same score without exceeding the constraints by using a different schedule that
+finishes the same total amount of work using the same budget and deadline.
 
@@
 
-We assume that resources are billed from the minute that they are requested
-until they are terminated.  As a result, provisioning delays have an impact on
-both the cost and makespan of an ensemble.
-
-Fig.~\ref{fig:delays} shows the ratios of simulated values to constraints when
-the provisioning delay is increased from 0 seconds up to 15 minutes. The
-figure summarizes a suite of 105,000 simulations (10 ensembles of 50 workflows
-x 5 distributions x 10 budgets x 10 deadlines x 7 error values x 3
-algorithms). To reduce the simulation time, only one application, Montage, was
-used in this experiment.
-
-As in Fig.~\ref{fig:variances}, the y-axis in each
-plot represents the ratio of the simulated value to the constraint value for
-the budget constraint or the deadline constraint.
-
-\begin{figure*}[tb]
+\begin{figure*}[htb]
     \centering
@@
 \end{figure*}
+\begin{figure*}[htb]
+  \centering
+  \includegraphics[width=0.8\textwidth]{run-finish-delays-test-0-output-distributions-delays}
+  \caption{Percentage of high scores achieved by each algorithm on CyberShake and LIGO
+  ensembles when provisioning delay varies from 0 seconds to 15 minutes.}
+  \label{fig:delays-scores}
+\end{figure*}
+
+We assume that resources are billed from the minute that they are requested
+until they are terminated.  As a result, provisioning delays have an impact on
+both the cost and makespan of an ensemble.
+
+Figure~\ref{fig:delays} shows the ratios of simulated values to constraints when
+the provisioning delay is increased from 0 seconds up to 15 minutes. The
+figure summarizes a suite of 105,000 simulations (10 ensembles of 50 workflows
+x 5 distributions x 10 budgets x 10 deadlines x 7 error values x 3
+algorithms). To reduce the simulation time, only one application, Montage, was
+used in this experiment.
+
+As in Figure~\ref{fig:variances}, the y-axis in each
+plot represents the ratio of the simulated value to the constraint value for
+the budget constraint or the deadline constraint.
 
@@
 
-Fig.~\ref{fig:delays_scores} shows the relative performance as a
+Figure~\ref{fig:delays-scores} shows the relative performance as a
 percentage of the number of high scores achieved by each algorithm as the
@@
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=0.5\textwidth]{delays_scores}
-  \caption{Percentage of high scores achieved by each algorithm on Montage 
-  with uniform unsorted ensembles when provisioning delay varies from 0 
-  seconds to 15 minutes.}
-  \label{fig:delays_scores}
-\end{figure}
-
 This experiment suggests that SPSS is too sensitive to provisioning delays
@@
 \end{figure*}
-
-Fig.~\ref{fig:failures} shows the ratios of simulated values to constraints when
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[width=0.8\textwidth]{run-finish-failures-test-0-output-distributions-failures}
+  \caption{Percentage of high scores achieved by each algorithm on CyberShake and LIGO
+  ensembles when failure rate varies 0 to 50\%.}
+  \label{fig:failures-scores}
+\end{figure*}
+
+Figure~\ref{fig:failures} shows the ratios of simulated values to constraints when
 the failures are introduced. The figure summarizes a suite of 525,000
@@
 performance of the static algorithm considerably, while the dynamic algorithms
-are able to adapt. Comparing Fig.~\ref{fig:failures} to
+are able to adapt.
+
+Figure~\ref{fig:failures-scores} shows the relative performance as a percentage of
+the number of high scores achieved by each algorithm as the failure rate
+increases. The results are similar to the ones of
+Figure~\ref{fig:delays-scores}. Comparing Figure~\ref{fig:failures} to
 Figs.~\ref{fig:variances} and~\ref{fig:delays} one may conclude that failures
@@
 
-Fig.~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles
+Figure~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles
 of 100 workflows with five different workflow sizes: 50, 200, 400, 600, and
@@
 
-Fig.~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS
+Figure~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS
 planning time is reasonable, taking on the order of tens of seconds to a few
@@
 In cases where there are no provisioning delays, task runtime estimates are
-good, and failures are rare, we found that SPSS
-performs significantly better than both dynamic algorithms. However, when
-conditions are less than perfect, the static plans produced by SPSS are
-disrupted and it frequently exceeds the budget and deadline constraints. In
-comparison, the dynamic algorithms are able to adapt to a wide variety of
-conditions, and rarely exceed the constraints even with long delays, poor
-estimates, and high failure rates.
+good, and failures are rare, we found that SPSS performs significantly better
+than both dynamic algorithms. However, when conditions are less than perfect,
+the static plans produced by SPSS are disrupted and it frequently exceeds the
+budget and deadline constraints. In comparison, the dynamic algorithms are able
+to adapt to a wide variety of conditions, and rarely exceed the constraints even
+with long delays, poor estimates, and high failure rates. This, however, comes
+at the cost of being able to complete less workflows than SPSS. Alternatively, a
+user may decide to accept that the budget and deadline constraints are exceeded,
+as long as more work is completed.
 
@@
 This work was supported by the NSF under award OCI-0943725 (STCI) and
-by the EU FP7 projects UrbanFlood (248767) and VPH-Share (269978).
+by the EU FP7 project VPH-Share (269978).
 

### 2013-08-30 scripts/R/data_cube_delays.R

@@ -1,0 +1,86 @@
+
+
+############################################################################################################################
+# Start
+############################################################################################################################
+
+#csv_filename = "run-failures-test-0-output.dat"
+#csv_filename = "run-scaling-failures-test-0-output.dat"
+#csv_filename = "run-scaling-1-failures-test-0-output.dat"
+#csv_filename = "run-finish-failures-test-0-output.dat"
+csv_filename = "run-finish-delays-test-0-output.dat"
+
+
+data = read.csv(csv_filename, colClasses=c("scorebits"="character"))
+#data = data[data$failureRate==0,]
+#data = data[data$runtimeVariance==0,]
+
+delays = unique(data$delay)
+
+n_delays = length(delays)
+
+runIds = unique(data$seed)
+n_runIds=length(runIds)
+
+
+print(n_runIds)
+
+
+applications = levels(unique(data$application))
+
+n_applications = length(applications)
+
+distributions = levels(unique(data$distribution))
+print(distributions)
+n_distributions = length(distributions)
+
+algorithms = levels(unique(data$algorithm))
+print(algorithms)
+
+n_algorithms = length(algorithms)
+
+n_budgets = 10
+n_deadlines = 10
+
+
+
+# create 7-dimensional matrix
+cube_dimensions = c(n_algorithms, n_runIds, n_deadlines, n_budgets, n_distributions, n_applications, n_delays)
+scoreBitString = data[order(data$delay, data$application, data$distribution, data$budget, data$deadline, data$seed, data$algorithm),]$scorebits
+scoreBitStrings = array(scoreBitString,  dim=cube_dimensions)
+
+
+maxScores = apply(scoreBitStrings, 2:7, max)
+topScores = array(0,  dim=cube_dimensions)
+for (i in 1:n_algorithms) {
+	topScores[i,,,,,,] = scoreBitStrings[i,,,,,,]>=maxScores
+}
+# sum scores, grouping by algorithm, application and delay
+sumScores = apply(topScores,c(1,6,7),sum)
+
+
+
+pdf(file=paste(csv_filename, "-distributions-delays.pdf", sep=""), height=2, width=7, bg="white")
+par(mfrow=c(1,3))
+par(mar=c(4,4,2,0))
+#select only cybershake and ligo
+for (i in c(1,3)) {
[…]

### 2013-08-30 scripts/R/data_cube_failures.R

@@ -1,0 +1,85 @@
+
+
+############################################################################################################################
+# Start
+############################################################################################################################
+
+#csv_filename = "run-failures-test-0-output.dat"
+#csv_filename = "run-scaling-failures-test-0-output.dat"
+#csv_filename = "run-scaling-1-failures-test-0-output.dat"
+csv_filename = "run-finish-failures-test-0-output.dat"
+
+
+data = read.csv(csv_filename, colClasses=c("scorebits"="character"))
+#data = data[data$failureRate==0,]
+#data = data[data$runtimeVariance==0,]
+
+failureRates = unique(data$failureRate)
+
+n_failureRates = length(failureRates)
+
+runIds = unique(data$seed)
+n_runIds=length(runIds)
+
+
+print(n_runIds)
+
+
+applications = levels(unique(data$application))
+
+n_applications = length(applications)
+
+distributions = levels(unique(data$distribution))
+print(distributions)
+n_distributions = length(distributions)
+
+algorithms = levels(unique(data$algorithm))
+print(algorithms)
+
+n_algorithms = length(algorithms)
+
+n_budgets = 10
+n_deadlines = 10
+
+
+
+# create 7-dimensional matrix
+cube_dimensions = c(n_algorithms, n_runIds, n_deadlines, n_budgets, n_distributions, n_applications, n_failureRates)
+scoreBitString = data[order(data$failureRate, data$application, data$distribution, data$budget, data$deadline, data$seed, data$algorithm),]$scorebits
+scoreBitStrings = array(scoreBitString,  dim=cube_dimensions)
+
+
+maxScores = apply(scoreBitStrings, 2:7, max)
+topScores = array(0,  dim=cube_dimensions)
+for (i in 1:n_algorithms) {
+	topScores[i,,,,,,] = scoreBitStrings[i,,,,,,]>=maxScores
+}
+# sum scores, grouping by algorithm, application and delay
+sumScores = apply(topScores,c(1,6,7),sum)
+
+
+
+pdf(file=paste(csv_filename, "-distributions-failures.pdf", sep=""), height=2, width=7, bg="white")
+par(mfrow=c(1,3))
+par(mar=c(4,4,2,0))
+#select only cybershake and ligo
+for (i in c(1,3)) {
+	print(applications[i])
[…]

### 2013-08-30 scripts/R/data_cube_variations.R

@@ -1,0 +1,86 @@
+
+
+############################################################################################################################
+# Start
+############################################################################################################################
+
+#csv_filename = "run-failures-test-0-output.dat"
+#csv_filename = "run-scaling-failures-test-0-output.dat"
+#csv_filename = "run-scaling-1-failures-test-0-output.dat"
+#csv_filename = "run-finish-failures-test-0-output.dat"
+csv_filename = "run-finish-variations-test-0-output.dat"
+
+
+data = read.csv(csv_filename, colClasses=c("scorebits"="character"))
+#data = data[data$failureRate==0,]
+#data = data[data$runtimeVariance==0,]
+
+runtimeVariances = unique(data$runtimeVariance)
+
+n_runtimeVariances = length(runtimeVariances)
+
+runIds = unique(data$seed)
+n_runIds=length(runIds)
+
+
+print(n_runIds)
+
+
+applications = levels(unique(data$application))
+
+n_applications = length(applications)
+
+distributions = levels(unique(data$distribution))
+print(distributions)
+n_distributions = length(distributions)
+
+algorithms = levels(unique(data$algorithm))
+print(algorithms)
+
+n_algorithms = length(algorithms)
+
+n_budgets = 10
+n_deadlines = 10
+
+
+
+# create 7-dimensional matrix
+cube_dimensions = c(n_algorithms, n_runIds, n_deadlines, n_budgets, n_distributions, n_applications, n_runtimeVariances)
+scoreBitString = data[order(data$runtimeVariance, data$application, data$distribution, data$budget, data$deadline, data$seed, data$algorithm),]$scorebits
+scoreBitStrings = array(scoreBitString,  dim=cube_dimensions)
+
+
+maxScores = apply(scoreBitStrings, 2:7, max)
+topScores = array(0,  dim=cube_dimensions)
+for (i in 1:n_algorithms) {
+	topScores[i,,,,,,] = scoreBitStrings[i,,,,,,]>=maxScores
+}
+# sum scores, grouping by algorithm, application and variance
+sumScores = apply(topScores,c(1,6,7),sum)
+
+
+
+pdf(file=paste(csv_filename, "-distributions-variances.pdf", sep=""), height=2, width=7, bg="white")
+par(mfrow=c(1,3))
+par(mar=c(4,4,2,0))
[…]

### 2013-08-30 scripts/R/gantt-test.R

@@ -1,0 +1,38 @@
+require(plotrix) # Required package
+
+# see simple example: http://www.inside-r.org/packages/cran/plotrix/docs/gantt.chart
+
+# if both vgidpos and vgridlab are specified,
+# starts and ends don't have to be dates
+info2<-list(labels=c("VM 1","VM 2","VM 1","VM 3","VM 3","VM 4","VM 2","VM 5","VM 4"),
+		starts=c(8.1,8.7,13.0,9.1,11.6,9.0,13.6,9.3,14.2),
+		ends=c(12.5,12.7,16.5,10.3,15.6,11.7,18.1,18.2,19.0))
+
+gantt.chart(info2,vgridlab=8:19,vgridpos=8:19,
+		main="A color for each interval - with borders",
+		taskcolors=c(2,3,7,4,8,5,3,6,"purple"),border.col="black")
+
+
+csv_filename = "/home/malawski/git/cloudworkflowsimulator-storage/task-info3.csv"
+alldata = read.csv(csv_filename, sep = " ")
+
+alldata <- alldata[order(alldata$vm),] 
+
+
+# if needed to start from 1, but probably better not rename them to keep consistent with logs
+#minvm = min(alldata$vm)
+#alldata$vm <- alldata$vm - minvm + 1
+
+info2<-list(labels=alldata$vm,		
+		starts=alldata$startTaskTime, 
+		ends=alldata$finishTaskTime)
+
+maxFinish = max(alldata$finishTaskTime)
+
+#png(file=paste(csv_filename, ".png", sep=""), width=1600, height=900, bg="white")
+pdf(file=paste(csv_filename, ".pdf", sep=""), width=16, height=9, bg="white")
+
+gantt.chart(info2,vgridlab=seq(0,maxFinish/3600, 1),vgridpos=seq(0,maxFinish, 3600),
+		taskcolors =alldata$dag, main="Tasks on VMs",border.col="black")
+
+dev.off()

### 2013-10-19 paper/paper.tex

@@
 (DAGs), are an important class of applications that lead to challenging problems
-in resource management on grid and cloud infrastructures. What becomes more
-important, workflows for large computational problems are often composed of
-several inter-related workflows grouped into {\em ensembles}.  Workflows in an
-ensemble typically have a similar structure, but they differ in their input
-data, number of tasks, and individual task sizes.
+in resource management on grid and cloud infrastructures. In addition, workflows
+for large computational problems are often composed of several inter-related
+workflows grouped into {\em ensembles}. Workflows in an ensemble typically have
+a similar structure, but they differ in their input data, number of tasks, and
+individual task sizes.
 
@@
 for a particular geographic location, and several hazard curves are combined to
-create a hazard map.  In the recent study of 2013 CyberShake generates a map that
-requires an ensemble of 286 workflows.  Similarly, users of
-Montage~\cite{Deelman2008} often need several workflows  with different
-parameters to generate a set of image mosaics that can be combined  into a
-single, large mosaic. The Galactic Plane ensemble, which generates several
-mosaics of the  entire sky in different wavelengths, consists of 17 workflows,
+create a hazard map.  In a 2013 study, CyberShake was used to generate a set
+of hazard maps over 286 sites that required an ensemble of 2288
+workflows~\cite{CyberShake13.4}. Similarly, users of Montage~\cite{Deelman2008}
+often need several workflows  with different parameters to generate a set of
+image mosaics that can be combined  into a single, large mosaic. The Galactic
+Plane ensemble~\cite{GalacticPlane}, which generates several mosaics of the
+entire sky in different wavelengths, consists of 17 workflows,
 each of which contains 900 sub-workflows.  Another ensemble example is the
@@
 their priority. For example, in CyberShake some geographic locations may be in heavily
-populated areas or in strategic locations such as power plants, while others
+populated areas or have sensitive facilities such as power plants, while others
 may be less important. Scientists typically prioritize the workflows in such
@@
 
-The granularity of the workflow tasks should be discussed in the context of the
+The granularity of the workflow tasks should be discussed in relation to the
 length of the cloud billing cycle. In this paper we assumed that the resources
-are charged for hourly time slots, as is the case of VM instances on Amazon EC2.
+are billed by the hour, as is the case of VM instances on Amazon EC2.
 However, there are cloud providers that have higher billing frequency, e.g.
@@
 in Figure~\ref{fig:completed-budgets} and ~\ref{fig:completed-deadlines} we show
-results for CyberShake and LIGO as representatives of fine grained and coarse
-grained workflows. The figure shows that, for this experiment, the scores remain
+results for CyberShake and LIGO as representatives of fine-grained and
+coarse-grained workflows. The figure shows that, for this experiment, the scores remain
 the same as the error increases. This is a result of the way the error is
@@
 to adapt to a wide variety of conditions, and rarely exceed the constraints even
-with long delays, poor estimates, and high failure rates. This, however, comes
-at the cost of being able to complete less workflows than SPSS. Alternatively, a
-user may decide to accept that the budget and deadline constraints are exceeded,
-as long as more work is completed.
+with long delays, poor estimates, and high failure rates. However, this comes
+at the cost of not being able to complete as many workflows as SPSS in ideal
+conditions.
 

### 2013-10-19 scripts/filter_columns.sh
Isuue #50: added scripts for managing CSV files.

@@ -1,0 +1,14 @@
+#!/bin/bash
+# Filters the given CSV file leaving only selected columns.
+# TODO(bryk): Add more comments.
+
+if [ "$#" -ne 2 ]; then
+  echo -e "Usage:\n\t$0 [column list] [simulation output CSV file]";
+  exit 1;
+fi
+
+COLS=$1;
+
+echo $COLS;
+csvtool namedcol $COLS $2;
+

### 2013-10-19 scripts/view_transfer_ouptuts.sh

@@ -1,0 +1,50 @@
+#!/bin/bash
+# This script can be used for lightweight analysis of transfer outputs of a simulation.
+# It runs graphical console tool for displaying CSV files, thus is far more faster than Excel or OpenOffice.
+# Furthermore, it can be run in a console-only environment.
+# TODO(bryk): Add more comments.
+
+COLSFILE='transter_outputs_columns.lst';
+
+if [ "$#" -ne 1 ]; then
+  echo -e "Usage:\n\t$0 [simulation output CSV file]";
+  exit 1;
+fi
+
+hash csvtool 2> /dev/null || {
+  echo >&2 "csvtool program is required to run this script. You can install it from your program repository.";
+  exit 1;
+}
+
+hash tabview 2> /dev/null || {
+  echo >&2 "tabview program is required to run this script. You can install it from github repository: https://github.com/firecat53/tabview";
+  exit 1;
+}
+
+TMPFILE=mktemp;
+
+cat $(COLSFILE) > $(TMPFILE) && csvtool namedcol `cat $(COLSFILE)` $1 | awk -F ',' '\
+{ \
+  hum[1024**5] = "PB"; \
+  hum[1024**4] = "TB"; \
+  hum[1024**3] = "GB"; \
+  hum[1024**2] = "MB"; \
+  hum[1024] = "KB"; \
+  for (i = 1; i <= 5; i++) { \
+    sum = $i; \
+    for (x = 1024**5; x >= 1024; x /= 1024) { \
+      if (sum >= x) { \
+        printf "%.2f%s,", sum/x, hum[x]; \
+        break; \
+      } \
+    } \
+  } \
+  for (i = 6; i < NF; i++) { \
+    printf "%s,", $i; \
+  } \
+  printf "%s\n", $NF; \
+}' >> $(TMPFILE) && tabview $(TMPFILE);
+
+rm $(TMPFILE);
+
+

### 2014-09-22 paper/paper.tex
Remove redundant sentence
added example to the paper

@@
 scoring defined as:
-$$ Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}} $$
+\begin{equation}
+\label{eq:score}
+Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
+\end{equation}
 where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed
@@
 executed even though there is no chance that those workflows will be completed
-within the current budget and deadline. Figure~\ref{fig:algorithm-example-dpds}
-shows an example schedule generated using the DPDS algorithm. The figure
-illustrates how tasks from lower priority workflows backfill idle VMs when
-tasks from higher priority workflows are not ready for execution.
+within the current budget and deadline.
 
@@
 budget constraints, and eliminates any waste that may be allowed by the dynamic
-algorithms. However, the disadvantage of SPSS is that it is sensitive to dynamic
-changes in the environment (see Sections~\ref{sec:variances} and
-\ref{sec:delays}).
+algorithms.
 
@@
 
+\subsection{Illustrative example}
+
+To illustrate how the algorithms perform, we parepared an artificial ensemble 
+of three workflows, as seen in Fig.~\ref{fig:example-ensemble}. 
+The deadline is set to 6 hours and the budget to \$18. In this example we assume
+that there are no VM provisioning delays, the runtime estimates are accurate, 
+there are no failures and the cost of 1 VM-hour is \$1.
+For dynamic algorithms we assume that the low utilization threshold ($u_l$) for deprovisioning is 50\%
+and that the autoscaling parameter $v_{max} = 1$.
+
+\begin{figure}[htb] 
+  \centering
+  \includegraphics[width=0.80\columnwidth]{figures/example-ensemble}
+  \caption{Artificial ensemble consisting of three workflows \{a, b, c\}, with priorities \{1, 2, 3\}.
+  Task labels indicate estimated execution time in minutes, 
+  so e.g. task $a.70$ has estimated execution time of 70 minutes.}
+  \label{fig:example-ensemble}
+\end{figure}
+
+The resulting schedule produced by DPDS algorithm is shown in Fig.~\ref{fig:algorithms-example-dpds}.
+The algorithm at the beginning estimates that it needs 3 VM instances by dividing the budget by the deadline ($18 / 6 = 3$). 
+Next, it schedules all the ready tasks in the priority order, so $a.70$ is started first, 
+and immediately $b.70$ and $c.60$ are started, since no other tasks of workflow a are ready. 
+When task $a.60$ completes, there are again other no higher priority tasks ready, 
+so the next task from workflow $c$ can be started ($c.45$). Since there is no preemption,
+this results in workflow $a$ being delayed by the lower priority workflow $c$.
+Finally, the outcome of this dynamic scheduling policy is that none of the workflows completes within the deadline.
+Actually, the tasks $b.100$, $c.65$ and $a.160$ are not finished since all the VMs 
+are terminated due to running out of budget. 
+The exponential score achieved by the algorithm computed according to equation~\ref{eq:score} is 0.
+On the other hand, we can observe that the utilization of the VMs is kept on the level of 100\% throughout the 
+whole execution time, which may be considered an advantage in some cases.
+
+\begin{figure}[htb] 
+  \centering
+  \includegraphics[width=0.80\columnwidth]{figures/algorithms-example-dpds}
+  \caption{Schedule produced by DPDS algorithm for the example ensemble from Figure~\ref{fig:example-ensemble}.
+  White boxes indicate VM instances provisioned for their hourly time slots, 
+  and solid boxes indicate running tasks.
+  None of the workflows completes within the deadline.}
+  \label{fig:algorithms-example-dpds}
+\end{figure}
+
+The resulting schedule produced by WA-DPDS algorithm is shown in Fig.~\ref{fig:algorithms-example-wadpds}.
+The schedule is generated in a similar way as in DPDS algorithm, with the one notable difference. 
+After scheduling tasks $a.70$ and $b.70$, the algorithm estimates whether it has enough budget to 
+admit workflow $c$. The answer is that based on the sum of task runtimes there is no room for workflow $c$,
+so the whole workflow is rejected. As a consequence, task $a.100$ can start at time $t=70$ minutes
+thus workflow $a$ is not delayed as in the case of DPDS. One can also observe that $VM.2$ and $VM.3$ remain idle for
+one hour, which is the consequence of the conservative provisioning settings we assume in this example.  
+This means that VM utilization at the level of 2/3 does not trigger deprovisioning in this case, and the maximum autoscaling
+factor set to 1 prevents from provisioning of new VMs over the initial number even if utilization is 100\%.
+In the resulting schedule only workflow $a$ completes, so the score is $2^{-1} = 0.5$, which is better than DPDS.
+
+\begin{figure}[htb] 
+  \centering
+  \includegraphics[width=0.80\columnwidth]{figures/algorithms-example-wadpds}
+  \caption{Schedule produced by WA-DPDS algorithm for the example ensemble from Figure~\ref{fig:example-ensemble}.
+  Workflow a completes within the deadline.}
+  \label{fig:algorithms-example-wadpds}
+\end{figure}
+
+The resulting schedule produced by SPSS algorithm is shown in Fig.~\ref{fig:algorithms-example-spss}.
+
+\begin{figure}[htb] 
+  \centering
+  \includegraphics[width=0.80\columnwidth]{figures/algorithms-example-spss}
+  \caption{Schedule produced by DPDS algorithm for the example ensemble from Figure~\ref{fig:example-ensemble}.
+  Workflows a and c complete within the deadline.}
+  \label{fig:algorithms-example-spss}
+\end{figure}
+
+The example given above illustrates the main concepts of the algorithms using the artificial ensemble.
+Two additional example schedules generated using our simulator are shown in 
+Figure~\ref{fig:examples-simulated}. 
+The schedule generated using DPDS (Figure~\ref{fig:algorithm-example-dpds})
+illustrates how tasks from lower priority workflows backfill idle VMs when
+tasks from higher priority workflows are not ready for execution.
+On the other hand, the schedule generated by SPSS (Figure~\ref{fig:algorithm-example-spss})
+shows how SPSS tends to start many workflows in parallel, running each workflow
+over a longer period of time on only a few VMs to minimize cost. In comparison,
+the dynamic algorithms tend to run one workflow at a time across many VMs in
+parallel.
+
 \begin{figure}[tb] 
@@
   VM. Boxes are tasks colored by workflow.}
+  \label{fig:examples-simulated}
 
 \end{figure}
-
-An example schedule generated by SPSS (Figure~\ref{fig:algorithm-example-spss})
-shows how SPSS tends to start many workflows in parallel, running each workflow
-over a longer period of time on only a few VMs to minimize cost. In comparison,
-the dynamic algorithms tend to run one workflow at a time across many VMs in
-parallel.
 

### 2014-09-25 paper/paper.tex
Add description for SPSS example schedule

@@
 
-The resulting schedule produced by SPSS algorithm is shown in Fig.~\ref{fig:algorithms-example-spss}.
+The resulting schedule produced by SPSS algorithm is shown in
+Fig.~\ref{fig:algorithms-example-spss}. SPSS generates the schedule by
+considering workflows in priority order. It first considers workflow $a$, and
+develops a schedule that completes $a$ by its deadline using the minimum
+number of node-hours. Once $a$ has been scheduled, SPSS will develop a
+low-cost schedule that includes workflow $b$. However, after adding $b$,
+the algorithm will determine that the total cost of the combined schedule
+for $a$ and $b$ is greater than the budget, and will reject workflow $b$ and
+remove it from the schedule. It will then consider workflow $c$, and develop
+a low-cost schedule that adds $c$ to the existing schedule for $a$. In order
+to complete each job in $c$ by its deadline, SPSS provisions 6 additional
+VM-hours. The resulting schedule has a total cost of 16 VM-hours, or
+\$16, which is less than the budget of \$18 and completes both $a$ and $c$ by
+their deadlines.
 
@@
   \includegraphics[width=0.80\columnwidth]{figures/algorithms-example-spss}
-  \caption{Schedule produced by DPDS algorithm for the example ensemble from Figure~\ref{fig:example-ensemble}.
-  Workflows a and c complete within the deadline.}
+  \caption{Schedule produced by SPSS algorithm for the example ensemble from Figure~\ref{fig:example-ensemble}.
+  Workflows $a$ and $c$ complete within the given budget and deadline.}
   \label{fig:algorithms-example-spss}

### 2014-09-29 paper/paper.tex
Update text for corrected SPSS example

@@
 %
-$$
+\begin{equation}
+\label{eq:spss-deadline}
 DL(t) = LST(t) + RT(t) + ST(Level(t))
-$$
+\end{equation}
 %
@@
 
-To illustrate how the algorithms perform, we parepared an artificial ensemble 
+To illustrate how the algorithms perform, we prepared an artificial ensemble 
 of three workflows, as seen in Fig.~\ref{fig:example-ensemble}. 
@@
 
-The resulting schedule produced by SPSS algorithm is shown in
-Fig.~\ref{fig:algorithms-example-spss}. SPSS generates the schedule by
+The final schedule produced by the SPSS algorithm is shown in
+Fig.~\ref{fig:algorithms-example-spss}. SPSS generates this schedule by
 considering workflows in priority order. It first considers workflow $a$, and
-develops a schedule that completes $a$ by its deadline using the minimum
-number of node-hours. Once $a$ has been scheduled, SPSS will develop a
-low-cost schedule that includes workflow $b$. However, after adding $b$,
+develops a low-cost schedule that completes $a$ by its deadline using a minimum
+number of VM-hours. This process involves assigning a deadline to each task
+according to Equation~\ref{eq:spss-deadline}, and then scheduling tasks in
+deadline-order on the VM that can complete each task by its deadline with the
+the minimum increase in cost. Once $a$ has been scheduled, SPSS will develop an
+updated low-cost schedule that includes workflow $b$. However, after adding $b$,
 the algorithm will determine that the total cost of the combined schedule
@@
 to complete each job in $c$ by its deadline, SPSS provisions 6 additional
-VM-hours. The resulting schedule has a total cost of 16 VM-hours, or
-\$16, which is less than the budget of \$18 and completes both $a$ and $c$ by
-their deadlines.
+VM-hours, which includes adding an hour to the start time of $VM.2$ for task
+$c.60$, adding an hour to the end of $VM.2$ for task $c.45$, adding an hour
+to the end of $VM.3$ for $c.50$, adding a new VM, $VM.4$, for 2 hours to run
+$c.65$, and finally adding another hour to the end of $VM.2$ for $c.55$. The
+resulting schedule has a total cost of 16 VM-hours, or \$16, which is less
+than the budget of \$18 and completes both $a$ and $c$ by the deadline.
 

### 2014-10-03 paper/paper.tex
updated spss example fig
updated related work

@@
 of these autoscaling systems, but no support for complex applications is provided.
+
 Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
-Kim2011}) also allow to scale the cloud resource pool or to extend the
+Kim2011,Chen2011}) also allow to scale the cloud resource pool or to extend the
 capabilities of clusters using cloud-burst techniques. Our approach is different
@@
 of work completed) is subject to optimization, whereas time and cost are
-treated as constraints.
-
-Other approaches~\cite{Talukder2009,Pandey2010} use
+treated as constraints, which is not addressed by other works.
+
+Our work can also be regarded as an extension of the budget-constrained 
+workflow scheduling~\cite{Sakellariou2007}, the cost optimization for workflows 
+under deadline constraints~\cite{Yu2005}, 
+or methods of stretching and compacting the workflow to optimize time and cost~\cite{Lee2013}, 
+in the sense that we are dealing with workflow ensembles
+and the deadline constraint is added. 
+Interesting discussion of uncertainties and the way to address them using Monte Carlo approach 
+is also presented in~\cite{Zheng2013b}, where the objective is makespan optimization.
+
+Multiple workflows has been also considered in previous research, 
+either in the context of heterogeneous systems~\cite{Zhao2006} 
+where makespan and fairness optimization is applied, 
+or in the context of grids~\cite{Hirales2010} 
+where multiple heuristics are evaluated with respect to the makespan of schedules they produce. 
+Scheduling multiple workflows in~\cite{Duan2012,Duan2014} applies multiobjective optimization 
+of time and cost under storage bandwidth and capacity constraints with the use of heuristics based on game theory. 
+The approach proposed in~\cite{TolosanaCalasanz20121300} focuses on enforcement of QoS 
+measured as throughput of processing by individual workflows. 
+The problem of workflow admission addressed in~\cite{Zheng2013a} is similar to the one solved by our algorithms, 
+but the difference is that we address the problem from the end-user perspective, 
+while in their work they assume that the resource provider decides 
+whether to admit or reject the workflow based on the resource estimates and the deadline constraints.
+
+Other approaches for scheduling workflows on grids and clouds such as ~\cite{Talukder2009,Pandey2010} use
 metaheuristics that usually run for a long time before producing good results,
-which makes them less useful in the scenarios we consider in this paper. Our work can also
-be regarded as an extension of the budget-constrained workflow scheduling
-\cite{Sakellariou2007} in the sense that we are dealing with workflow ensembles
-and the deadline constraint is added.
+which makes them less useful in the scenarios we consider in this paper.
+
+Noteworthy are also methods based on mathematical programming, 
+using linear programming or mixed-integer programming. 
+Such method is applied to scheduling workflows on hybrid clouds using time discretization~\cite{Genez13b}, 
+but the model presented there is limited to small scale workflows. 
+Large-scale applications on hybrid clouds are addressed in~\cite{VandenBossche2010,VandenBossche2013973}, 
+but their model considers bag-of-task applications instead of workflows. 
+The cloud bursting scenario where private cloud is combined with a public one, described in~\cite{BittencourtM11}, 
+addresses workflows, and the goal is to minimize the cost while keeping the deadline. 
+In our related paper~\cite{optimization-ppam13}, we also use the mixed-integer programming approach 
+to schedule multi-level deadline-constrained workflows, but workflow ensembles are not considered.
+
+\begin{table*}[tb]
+    \centering
+    \caption{Summary table of related work on workflow scheduling and provisioning.} 
+    \label{tab:related}
+    \includegraphics[width=\textwidth]{Cloud-workflows-table}
+\end{table*}
+
+The analysis of related work is summarized in Table~\ref{tab:related}, where we classify the methods based on 7 categories. 
+Infrastructure type can be grid or cloud, where grid includes also utility grids 
+and heterogeneous computing systems where cost can be also relevant (such as in~\cite{Sakellariou2007}). 
+Application types of interest are bags of tasks, single workflows or multiple workflows. 
+The workload can be static, where the scheduling process can plan the tasks in advance, 
+or dynamic, where the algorithm has to deal with unpredictable stream of jobs arriving. 
+The provisioning methods are divided into dynamic and static approaches, 
+where static means that the provisioning plan is prepared prior to execution. 
+For simplification, we classify all the algorithms for grid and heterogeneous systems into the static category, 
+since these systems do not allow for creation of new resources on demand, 
+even if some mechanisms for dynamic provisioning such as pilot jobs are available. 
+The related work addresses multiple optimization objectives, 
+including time, monetary cost, reliability, energy or fairness, 
+and there are examples of multicriteria optimization among them. 
+Most of the scheduling approaches consider also constraints for optimization, 
+and most commonly addressed ones are budget and deadline, 
+whereas such additional constraints as bandwidth, storage capacity or job throughput are also present. 
+Finally, we divide the algorithmic methods into heuristics, 
+including various list scheduling techniques based on graph analysis, 
+metaheuristics that use genetic or evolutionary algorithms, 
+and the approaches using mathematical programming such as linear 
+or mixed-integer programming for which ready to use solvers exist.
+
+The analysis of the table suggests that there are many approaches to workflow scheduling 
+and resource provisioning on clouds, but there is still room for improvement. 
+New infrastructure types, including hybrid and federated clouds should be investigated, 
+more emphasis should be put on problems with multiple workflows, 
+and new methods for combining static and dynamic workloads and algorithms offer potentially interesting area of research. 
+Finally, other objectives and constraints in addition to budget and execution time can be explored.
+
+In this context, we consider our attempt to address workflow ensembles using both static and dynamic algorithms 
+presented in this paper an interesting and original contribution that has not been explored before.
 
@@
 than the budget of \$18 and completes both $a$ and $c$ by the deadline.
+The exponential score is then $2^{-1}+2^{-3} = 0.625$, which is better than WA-DPDS.
 

### 2014-10-14 paper/paper.tex
replies to review comments

@@
 billing cycle, which is an important aspect of scheduling and resource provisioning
-on cloud infrastructures.
+on cloud infrastructures. The section on the related work and the examples of algorithms 
+have been also extended to provide a better positioning and increase the readability of the paper.
 
@@
 %synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.
+
+The decision to use a simulator instead of actual execution of workflows on clouds 
+was motivated by our intention to perform an extensive analysis of the algorithms
+in various aspects. This results in a large parameter space, since e.g. for analysis 
+presented in Section~\ref{sec:variances}, we needed to execute 525,000 simulations 
+of workflow ensembles, each ensemble consisting of 50 workflows of up to 1000 tasks.
+Not only would this be impractical to run on a real infrastructure for obvious reasons, 
+but also it would not allow to study such interesting aspects as the influence 
+of task graularity presented in Section~\ref{sec:granularity}. 
+The usage of simulator, although also quite compute intensive, allowed us
+to study all these aspects in detail.
 
@@
 overestimates as underestimates in any given simulation. 
+Our approach for generating errors is similar to the one used in~\cite{Zheng2013b},
+where the random error represents the Quality of Estimation metric.
 

### 2014-10-14 paper/paper.tex
Minor edits to paper

--- 2014-10-03 paper/paper.tex
@@
 billing cycle, which is an important aspect of scheduling and resource provisioning
-on cloud infrastructures.
+on cloud infrastructures. The related work section and the algorithm examples
+have been also extended to provide a better positioning and increase the readability of the paper.
 
@@
 Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
-Kim2011,Chen2011}) also allow to scale the cloud resource pool or to extend the
+Kim2011,Chen2011}) also enable scaling the cloud resource pool or to extending the
 capabilities of clusters using cloud-burst techniques. Our approach is different
@@
 
-Our work can also be regarded as an extension of the budget-constrained 
-workflow scheduling~\cite{Sakellariou2007}, the cost optimization for workflows 
+Our work can also be regarded as an extension of budget-constrained 
+workflow scheduling~\cite{Sakellariou2007}, cost optimization for workflows 
 under deadline constraints~\cite{Yu2005}, 
@@
 and the deadline constraint is added. 
-Interesting discussion of uncertainties and the way to address them using Monte Carlo approach 
+An interesting discussion of uncertainties and the way to address them using a Monte Carlo approach 
 is also presented in~\cite{Zheng2013b}, where the objective is makespan optimization.
 
-Multiple workflows has been also considered in previous research, 
+Scheduling optimization for multiple workflows has been also considered in previous research, 
 either in the context of heterogeneous systems~\cite{Zhao2006} 
@@
 
-Noteworthy are also methods based on mathematical programming, 
+Also noteworthy are methods based on mathematical programming, 
 using linear programming or mixed-integer programming. 
-Such method is applied to scheduling workflows on hybrid clouds using time discretization~\cite{Genez13b}, 
+Such a method is applied to scheduling workflows on hybrid clouds using time discretization in~\cite{Genez13b}, 
 but the model presented there is limited to small scale workflows. 
@@
 but their model considers bag-of-task applications instead of workflows. 
-The cloud bursting scenario where private cloud is combined with a public one, described in~\cite{BittencourtM11}, 
-addresses workflows, and the goal is to minimize the cost while keeping the deadline. 
+The cloud bursting scenario described in~\cite{BittencourtM11}, where a private cloud is combined with a public one, 
+addresses workflows, and the goal is to minimize the cost while satisfying the deadline. 
 In our related paper~\cite{optimization-ppam13}, we also use the mixed-integer programming approach 
@@
 The analysis of related work is summarized in Table~\ref{tab:related}, where we classify the methods based on 7 categories. 
-Infrastructure type can be grid or cloud, where grid includes also utility grids 
-and heterogeneous computing systems where cost can be also relevant (such as in~\cite{Sakellariou2007}). 
+Infrastructure type can be grid or cloud, where grid includes utility grids 
+and heterogeneous computing systems where cost can be relevant (such as in~\cite{Sakellariou2007}). 
 Application types of interest are bags of tasks, single workflows or multiple workflows. 
@@
 where static means that the provisioning plan is prepared prior to execution. 
-For simplification, we classify all the algorithms for grid and heterogeneous systems into the static category, 
+To simplify, we classify all the algorithms for grid and heterogeneous systems into the static category, 
 since these systems do not allow for creation of new resources on demand, 
@@
 including time, monetary cost, reliability, energy or fairness, 
-and there are examples of multicriteria optimization among them. 
-Most of the scheduling approaches consider also constraints for optimization, 
-and most commonly addressed ones are budget and deadline, 
-whereas such additional constraints as bandwidth, storage capacity or job throughput are also present. 
+and there are examples of multi-criteria optimization among them. 
+Most of the scheduling approaches also consider constraints for optimization, 
+and the most commonly addressed ones are budget and deadline. 
+Other constraints, such as bandwidth, storage capacity or job throughput, are also present. 
 Finally, we divide the algorithmic methods into heuristics, 
@@
 
-The analysis of the table suggests that there are many approaches to workflow scheduling 
+An analysis of the table suggests that there are many approaches to workflow scheduling 
 and resource provisioning on clouds, but there is still room for improvement. 
@@
 %synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.
+
+The decision to use a simulator instead of actually executing workflows on clouds 
+was motivated by our intention to perform an extensive analysis of the algorithms
+in various aspects. This results in a very large parameter space. For example, the analysis 
+presented in Section~\ref{sec:variances} required 525,000 simulations 
+of workflow ensembles, each ensemble consisting of 50 workflows with up to 1000 tasks per workflow.
+Not only would this be impractical to run on a real infrastructure for obvious reasons, 
+but it would also not enable us to study interesting aspects of the problem, such as the influence 
+of task granularity presented in Section~\ref{sec:granularity}. 
+Our use of a simulator, although quite compute-intensive itself, enabled us
+to study all these aspects in detail.
 

### 2014-10-15 paper/paper.tex
minor edits, misspelling corrections
minor edits

@@
 important problem concerning the efficient management of such ensembles under
-budget and deadline constraints on Infrastructure- as-a-Service (IaaS) clouds.
+budget and deadline constraints on Infrastructure-as-a-Service (IaaS) clouds.
 IaaS clouds are characterized by on-demand resource provisioning capabilities
@@
 \label{sec:perf_metric}
-In order to precisely define the objective of the algrithms it is necessary to
+In order to precisely define the objective of the algorithms it is necessary to
 introduce a metric that can be used to score the performance of the different
@@
 deadline. It can be observed, that with increasing deadline, the number of
-completed worklfows grows steeply and then stabilizes. This means that for
+completed workflows grows steeply and then stabilizes. This means that for
 longer deadlines it is usually not possible to complete more workflows due to
@@
 the ensembles that have priorities assigned randomly (unsorted), in comparison
-to the ensembes where the priorities are assigned based on workflow size
+to the ensembles where the priorities are assigned based on workflow size
 (sorted). This results from the fact that in the unsorted distributions the large
-workflows often have low priorities, so they can be rejected by the algoritms,
+workflows often have low priorities, so they can be rejected by the algorithms,
 thus making more space in the schedule for smaller high-priority workflows.
@@
 smaller ones with lower priority. That is why the number of completed workflows
-may decrese, but the exponential score (utility function) based on priorities
+may decrease, but the exponential score (utility function) based on priorities
 will be higher. These effects are visible for CyberShake application, for which
@@
 
-The chacacteristics discussed in this section can be useful not only for
+The characteristics discussed in this section can be useful not only for
 analysis of our algorithms, but also for planning scientific experiments and
@@
 workflow description with these data. In practice, however, these
-estimates are often inaccurate. Given inaccuate estimates, the question is: How
+estimates are often inaccurate. Given inaccurate estimates, the question is: How
 do errors in task runtime estimates impact the performance of our scheduling and
@@
 indicate maximum and minimum values. The ratio indicates whether the value 
-(for example, the simulated cost) exeeded the constraint (the budget). Values
+(for example, the simulated cost) exceeded the constraint (the budget). Values
 greater than 1 indicate that the constraint was exceeded.
@@
 The SPSS plan describes what tasks to execute on which VMs, but not when to
-execute them. At runtime, the worklow engine is forced to spend extra money
+execute them. At runtime, the workflow engine is forced to spend extra money
 to extend the runtime of some VMs so that all of the tasks assigned to that

### 2014-10-17 paper/paper.tex
updated table and references

@@
 systems. However, our problem is different from
-\cite{Yu2005},\cite{Abrishami2013158} and~\cite{Young2013} in that we consider
+\cite{Yu2005},\cite{Abrishami2013158} and~\cite{Lee2013} in that we consider
 ensembles of workflows instead of individual workflows.
@@
 constraint.
+In~\cite{Arabnejad2014} authors introduced a novel single DAG static scheduling PEFT algorithm, 
+which outperforms well-known list-based heuristics for makespan optimization, but the budget is not taken into account.
 
 This work is related to bi-criteria scheduling and multi-criteria scheduling of
-workflows~\cite{Wieczorek2009,Prodan2010,Dongarra2007,Durillo2013}. These approaches are
+workflows~\cite{Wieczorek2009,Prodan2010,Dongarra2007,Durillo2013,Durillo2012}. These approaches are
 similar to ours in that we have two scheduling criteria: cost and makespan. The
@@
 
-Other approaches for scheduling workflows on grids and clouds such as ~\cite{Talukder2009,Pandey2010} use
+Other approaches for scheduling workflows on grids and clouds such as ~\cite{Talukder2009,Pandey2010,Durillo2014} use
 metaheuristics that usually run for a long time before producing good results,

### 2014-10-18 src/cws/core/config/GlobalStorageParamsLoader.java
Silenced compilation warnings for configuration loader

@@ -172,7 +172,8 @@ public class GlobalStorageParamsLoader {
         }
     }
 
-    private Map<String, Object> loadConfigFromFile(CommandLine args) throws FileNotFoundException {
+    @SuppressWarnings("unchecked")
+	private Map<String, Object> loadConfigFromFile(CommandLine args) throws FileNotFoundException {
         String gsConfigFilename = args.getOptionValue(GS_TYPE_OPTION_NAME, DEFAULT_GS_TYPE_FILENAME);
         String gsConfigDirectory = args.getOptionValue(GS_CONFIGS_DIRECTORY_OPTION_NAME, DEFAULT_GS_CONFIGS_DIRECTORY);
 

### 2014-10-19 paper/paper.tex
applied Ewa comments

@@
 important problem concerning the efficient management of such ensembles under
-budget and deadline constraints on Infrastructure-as-a-Service (IaaS) clouds.
+budget and deadline constraints on Infrastructure as a Service (IaaS) clouds.
 IaaS clouds are characterized by on-demand resource provisioning capabilities
-and a pay-per-use model. We discuss, develop, and assess algorithms based on
+and a pay-per-use model. We discuss, develop, and assess novel algorithms based on
 static and dynamic strategies for both task scheduling and resource
@@
 image mosaics that can be combined  into a single, large mosaic. The Galactic
-Plane ensemble~\cite{GalacticPlane}, which generates several mosaics of the
-entire sky in different wavelengths, consists of 17 workflows,
+Plane ensemble~\cite{GalacticPlane}, which generates several mosaics 
+in different wavelengths, consists of 17 workflows,
 each of which contains 900 sub-workflows.  Another ensemble example is the
@@
 metrics such as workflow makespan or resource utilization, but must also consider
-budget constraints, since the resources from commercial clouds
+budget constraints, since commercial cloud resources
 usually have monetary costs associated with them~\cite{Durkee2010}.
@@
 Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
-Kim2011,Chen2011}) also enable scaling the cloud resource pool or to extending the
+Kim2011,Chen2011}) also enable scaling the cloud resource pool or the extension of the
 capabilities of clusters using cloud-burst techniques. Our approach is different
-in that we consider workflows, while policy based approaches typically consider
+in that we consider workflows, while policy-based approaches typically consider
 bags of independent tasks or unpredictable batch workloads. This enables us to
@@
 Scheduling optimization for multiple workflows has been also considered in previous research, 
-either in the context of heterogeneous systems~\cite{Zhao2006} 
+either in the context of heterogeneous systems~\cite{Zhao2006}, 
 where makespan and fairness optimization is applied, 
-or in the context of grids~\cite{Hirales2010} 
+or in the context of grids~\cite{Hirales2010}, 
 where multiple heuristics are evaluated with respect to the makespan of schedules they produce. 
@@
 The analysis of related work is summarized in Table~\ref{tab:related}, where we classify the methods based on 7 categories. 
-Infrastructure type can be grid or cloud, where grid includes utility grids 
-and heterogeneous computing systems where cost can be relevant (such as in~\cite{Sakellariou2007}). 
-Application types of interest are bags of tasks, single workflows or multiple workflows. 
+The infrastructure type can be a grid or a cloud, where a grid includes utility grids 
+and heterogeneous computing systems, where cost can be relevant (such as in~\cite{Sakellariou2007}). 
+Application types of interest are bags of tasks, single workflows, or multiple workflows. 
 The workload can be static, where the scheduling process can plan the tasks in advance, 
@@
 since these systems do not allow for creation of new resources on demand, 
-even if some mechanisms for dynamic provisioning such as pilot jobs are available. 
+even if some mechanisms for dynamic provisioning such as pilot jobs~\cite{Sfiligoi08} are available. 
 The related work addresses multiple optimization objectives, 
@@
 New infrastructure types, including hybrid and federated clouds should be investigated, 
-more emphasis should be put on problems with multiple workflows, 
-and new methods for combining static and dynamic workloads and algorithms offer potentially interesting area of research. 
+more emphasis should be put on problems with multiple interrelated workflows, 
+and new methods for combining static and dynamic workloads and algorithms offer potentially interesting areas of research. 
 Finally, other objectives and constraints in addition to budget and execution time can be explored.
@@
 To illustrate how the algorithms perform, we prepared an artificial ensemble 
-of three workflows, as seen in Fig.~\ref{fig:example-ensemble}. 
+of three workflows, as seen in Figure~\ref{fig:example-ensemble}. 
 The deadline is set to 6 hours and the budget to \$18. In this example we assume
@@
   \includegraphics[width=0.80\columnwidth]{figures/example-ensemble}
-  \caption{Artificial ensemble consisting of three workflows \{a, b, c\}, with priorities \{1, 2, 3\}.
+  \caption{Artificial ensemble consisting of three workflows \{$a$, $b$, $c$\}, with priorities \{1, 2, 3\}.
   Task labels indicate estimated execution time in minutes, 
@@
 
-The resulting schedule produced by DPDS algorithm is shown in Fig.~\ref{fig:algorithms-example-dpds}.
+The resulting schedule produced by DPDS algorithm is shown in Figure~\ref{fig:algorithms-example-dpds}.
 The algorithm at the beginning estimates that it needs 3 VM instances by dividing the budget by the deadline ($18 / 6 = 3$). 
 Next, it schedules all the ready tasks in the priority order, so $a.70$ is started first, 
-and immediately $b.70$ and $c.60$ are started, since no other tasks of workflow a are ready. 
+and immediately $b.70$ and $c.60$ are started, since no other tasks of workflow $a$ are ready. 
 When task $a.60$ completes, there are again other no higher priority tasks ready, 
@@
 Finally, the outcome of this dynamic scheduling policy is that none of the workflows completes within the deadline.
-Actually, the tasks $b.100$, $c.65$ and $a.160$ are not finished since all the VMs 
+Actually, tasks $b.100$, $c.65$ and $a.160$ are not finished since all the VMs 
 are terminated due to running out of budget. 
@@
 
-The resulting schedule produced by WA-DPDS algorithm is shown in Fig.~\ref{fig:algorithms-example-wadpds}.
-The schedule is generated in a similar way as in DPDS algorithm, with the one notable difference. 
+The resulting schedule produced by the WA-DPDS algorithm is shown in Figure~\ref{fig:algorithms-example-wadpds}.
+The schedule is generated in a similar way as in DPDS algorithm, with one notable difference. 
 After scheduling tasks $a.70$ and $b.70$, the algorithm estimates whether it has enough budget to 
-admit workflow $c$. The answer is that based on the sum of task runtimes there is no room for workflow $c$,
-so the whole workflow is rejected. As a consequence, task $a.100$ can start at time $t=70$ minutes
-thus workflow $a$ is not delayed as in the case of DPDS. One can also observe that $VM.2$ and $VM.3$ remain idle for
+admit workflow $c$. Since based on the sum of task runtimes there is no room for workflow $c$,
+the whole workflow is rejected. As a consequence, task $a.100$ can start at time $t=70$ minutes.
+Therefore workflow $a$ is not delayed as in the case of DPDS. One can also observe that $VM.2$ and $VM.3$ remain idle for
 one hour, which is the consequence of the conservative provisioning settings we assume in this example.  
@@
 The final schedule produced by the SPSS algorithm is shown in
-Fig.~\ref{fig:algorithms-example-spss}. SPSS generates this schedule by
+Figure~\ref{fig:algorithms-example-spss}. SPSS generates this schedule by
 considering workflows in priority order. It first considers workflow $a$, and
@@
   \subfloat[DPDS]{
-  \includegraphics[width=0.80\columnwidth]{figures/output-test-DPDS}
+  \includegraphics[width=\columnwidth]{figures/output-test-DPDS}
   \label{fig:algorithm-example-dpds}
@@
   \subfloat[SPSS]{
-  \includegraphics[width=0.80\columnwidth]{figures/output-test-SPSS}
+  \includegraphics[width=\columnwidth]{figures/output-test-SPSS}
   \label{fig:algorithm-example-spss}
@@
 randomized ensembles using workflows available from the workflow generator
-gallery~\cite{WorkflowGenerator}. The
+gallery~\cite{WorkflowGenerator,Silva2014}. The
 gallery contains synthetic workflows modeled using structures and parameters
@@
     \centering
-    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
+    \includegraphics[width=0.8\columnwidth]{figures/ensemble-pareto}
     \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-deadlines-app-dist-selected}    
+    \includegraphics[width=\textwidth]{run-finish-variations-test-0-output-deadlines-app-dist-selected}    
     \caption{Number of completed workflows from the ensemble depending on normalized deadline.}
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-budgets-app-dist-selected}    
+    \includegraphics[width=\textwidth]{run-finish-variations-test-0-output-budgets-app-dist-selected}    
     \caption{Number of completed workflows from the ensemble depending on normalized budget.}
@@
 
-Fig~\ref{fig:completed-budgets} shows the results based on normalized budget,
+Figure\ref{fig:completed-budgets} shows the results based on normalized budget,
 each point being the average over all deadline and random seed. What becomes
@@
 workflows have the highest priority, so it requires a large increment of a small
-budget to fir the next workflow into the schedule.
+budget to fit the next workflow into the schedule.
 
@@
 analysis of our algorithms, but also for planning scientific experiments and
-understanding tradeoffs between cost, deadline and the number of work
+understanding trade-offs between cost, deadline and the number of work
 complete~\cite{Malawski-FGCS-13}.
@@
 %    \subfloat[CyberShake]{
-    \includegraphics[width=0.6\textwidth]{run-stretching-test-output-scaling-CYBERSHAKE}
+    \includegraphics[width=0.64\textwidth]{run-stretching-test-output-scaling-CYBERSHAKE}
 %    }\\
 %    \subfloat[Montage]{
-    \includegraphics[width=0.6\textwidth]{run-stretching-test-output-scaling-MONTAGE}
+    \includegraphics[width=0.64\textwidth]{run-stretching-test-output-scaling-MONTAGE}
 %    }
@@
 first 10 minutes). Our results indicate that when the task granularity is close
-to the length of the billing period, the SPSS produces better results, while for
-tasks that are shorter than the billing period, the dynamic algorithms perform
+to the length of the billing period,  SPSS produces better results, while for
+tasks that are shorter than the billing period,  dynamic algorithms perform
 better. This suggests that these granularity effects have to be taken into account when
@@
 runtime to ensure that the constraints are not exceeded, even when the quality
-of information available to them is low. Second, unlike the dynamic
+of information available to them is low. Second, unlike dynamic
 algorithms, the static algorithm frequently exceeded the budget constraint; by
@@
 Figure~\ref{fig:delays-scores}. Comparing Figure~\ref{fig:failures} to
-Figs.~\ref{fig:variances} and~\ref{fig:delays} one may conclude that failures
+Figures~\ref{fig:variances} and~\ref{fig:delays} one may conclude that failures
 are worse than provisioning delays and runtime estimate errors, since their
@@
 This study suggests several areas for future work. Our current approach models
-data access as part of the task execution time and does not consider data
+data access as part of the task execution time and does not explicitly consider data
 storage and transfer costs. In the future we plan to extend the application
@@
 \section*{Acknowledgments}
-This work was supported by the NSF under award OCI-0943725 (STCI) and
+This work was supported by the NSF under award ACI-1148515 and
 by the EU FP7 project VPH-Share (269978).

### 2015-01-20 paper/paper.tex
added number of tasks in periodograms
Update acknowledgements

@@
 into multiple batches processed by different workflows. Additional workflows are
-created to run the analysis using different parameters. A recent analysis of
-Kepler satellite data required three ensembles of 15 workflows.
+created to run the analysis using different parameters. 
+An example analysis of Kepler satellite data required three ensembles of 15 workflows, 
+each workflow with about 14,000 tasks.
 
@@
 \section*{Acknowledgments}
-This work was supported by the NSF under award ACI-1148515 and
-by the EU FP7 project VPH-Share (269978).
+This work was supported by the U.S. Department of Energy Office of Science
+under award number ER26110, by the National Science Foundation ACI SI2-SSI
+program award number 1148515, and by the EU FP7 project VPH-Share (269978).
 

### 2015-04-25 paper/paper.tex

@@
 under award number ER26110, by the National Science Foundation ACI SI2-SSI
-program award number 1148515, and by the EU FP7 project VPH-Share (269978).
+program award number 1148515, by the AGH grant 11.11.230.124 
+and by the EU FP7 project VPH-Share (269978).
 

### 2015-04-25 integration_test/cws/core/scheduler/DAGDynamicSchedulerStorageAwareTest.java
Fix unused imports warnings.

@@ -10,11 +10,9 @@ import org.junit.Test;
 import cws.core.Cloud;
 import cws.core.EnsembleManager;
 import cws.core.Provisioner;
-import cws.core.provisioner.NullProvisioner;
 import cws.core.VM;
 import cws.core.VMFactory;
 import cws.core.WorkflowEngine;
-import cws.core.WorkflowEvent;
 import cws.core.cloudsim.CloudSimWrapper;
 import cws.core.core.VMType;
 import cws.core.core.VMTypeBuilder;
@@ -22,6 +20,7 @@ import cws.core.dag.DAG;
 import cws.core.dag.DAGParser;
 import cws.core.engine.Environment;
 import cws.core.log.WorkflowLog;
+import cws.core.provisioner.NullProvisioner;
 import cws.core.storage.StorageManager;
 import cws.core.storage.cache.FIFOCacheManager;
 import cws.core.storage.cache.VMCacheManager;

### 2015-04-25 integration_test/cws/core/scheduler/DAGDynamicSchedulerTest.java

@@ -13,11 +13,9 @@ import org.junit.Test;
 import cws.core.Cloud;
 import cws.core.EnsembleManager;
 import cws.core.Provisioner;
-import cws.core.provisioner.NullProvisioner;
 import cws.core.VM;
 import cws.core.VMFactory;
 import cws.core.WorkflowEngine;
-import cws.core.WorkflowEvent;
 import cws.core.cloudsim.CloudSimWrapper;
 import cws.core.core.VMType;
 import cws.core.core.VMTypeBuilder;
@@ -26,7 +24,7 @@ import cws.core.dag.DAGParser;
 import cws.core.dag.Task;
 import cws.core.engine.Environment;
 import cws.core.log.WorkflowLog;
-import cws.core.provisioner.ConstantDistribution;
+import cws.core.provisioner.NullProvisioner;
 import cws.core.storage.StorageManager;
 import cws.core.storage.VoidStorageManager;
 

### 2015-04-25 integration_test/cws/core/scheduler/EnsembleDynamicSchedulerTest.java

@@ -13,11 +13,9 @@ import org.junit.Test;
 import cws.core.Cloud;
 import cws.core.EnsembleManager;
 import cws.core.Provisioner;
-import cws.core.provisioner.NullProvisioner;
 import cws.core.VM;
 import cws.core.VMFactory;
 import cws.core.WorkflowEngine;
-import cws.core.WorkflowEvent;
 import cws.core.cloudsim.CloudSimWrapper;
 import cws.core.core.VMType;
 import cws.core.core.VMTypeBuilder;
@@ -27,6 +25,7 @@ import cws.core.dag.Task;
 import cws.core.engine.Environment;
 import cws.core.log.WorkflowLog;
 import cws.core.provisioner.ConstantDistribution;
+import cws.core.provisioner.NullProvisioner;
 import cws.core.storage.StorageManager;
 import cws.core.storage.VoidStorageManager;
 

### 2015-04-25 src/cws/core/Cloud.java

@@ -1,10 +1,9 @@
 package cws.core;
 
 import java.util.HashSet;
-import java.util.Set;
-import com.google.common.collect.ImmutableSet;
-
 import java.util.List;
+import java.util.Set;
+
 import com.google.common.collect.ImmutableList;
 import com.google.common.collect.ImmutableList.Builder;
 

### 2015-04-25 src/cws/core/algorithms/DynamicAlgorithm.java

@@ -8,7 +8,6 @@ import cws.core.Scheduler;
 import cws.core.VM;
 import cws.core.VMFactory;
 import cws.core.WorkflowEngine;
-import cws.core.WorkflowEvent;
 import cws.core.cloudsim.CloudSimWrapper;
 import cws.core.dag.DAG;
 import cws.core.engine.Environment;

### 2015-04-25 src/cws/core/provisioner/SimpleUtilizationBasedProvisioner.java

@@ -4,7 +4,6 @@ import java.util.HashSet;
 import java.util.Iterator;
 import java.util.Set;
 
-import cws.core.Provisioner;
 import cws.core.VM;
 import cws.core.VMFactory;
 import cws.core.WorkflowEngine;
